### 2023-01-06 manuscript/jmlr_manuscript.tex
Initial work on the manuscript.

\documentclass[twoside,11pt]{article}

\usepackage{blindtext}

% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
%
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf

% Available options for package jmlr2e are:
%
%   - abbrvbib : use abbrvnat for the bibliography style
%   - nohyperref : do not load the hyperref package
%   - preprint : remove JMLR specific information from the template,
%         useful for example for posting to preprint servers.
%
% Example of using the package with custom options:
%
% \usepackage[abbrvbib, preprint]{jmlr2e}

\usepackage{jmlr2e}

% Definitions of handy macros can go here

\newcommand{\dataset}{{\cal D}}
\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}

% Heading arguments are {volume}{year}{pages}{date submitted}{date published}{paper id}{author-full-names}

\usepackage{lastpage}
\jmlrheading{23}{2022}{1-\pageref{LastPage}}{1/21; Revised 5/22}{9/22}{21-0000}{Author One and Author Two}

% Short headings should be running head and authors last names

\ShortHeadings{Shift-sensitive Power Transformation}{Zwanenburg and Löck}
\firstpageno{1}

\begin{document}

\title{Shift-sensitive Power Transformations for Transforming Data to Normality}

\author{\name Alex Zwanenburg \email alexander.zwanenburg@nct-dresden.de\\
       \addr National Center for Tumour Diseases (NCT/UCC) Dresden\\
       University of Washington\\
       Seattle, WA 98195-4322, USA
       \AND
       \name Author Two \email two@cs.berkeley.edu \\
       \addr Division of Computer Science\\
       University of California\\
       Berkeley, CA 94720-1776, USA}

\editor{My editor}

\maketitle

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\blindtext
\end{abstract}

\begin{keywords}
  keyword one, keyword two, keyword three
\end{keywords}

\section{Introduction}

\blindmathpaper

\section{Methods}

\section{Simulation}

\section{Experimental Results}

\section{Discussion}

\section{Conclusion}

Here is a citation \cite{chow:68}.

% Acknowledgements and Disclosure of Funding should go at the end, before appendices and references

\acks{All acknowledgements go at the end of the paper before appendices and references.
Moreover, you are required to declare funding (financial activities supporting the
submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found on the JMLR website.}

% Manual newpage inserted to improve layout of sample file - not
% needed in general before appendices/bibliography.

\newpage

\appendix
\section*{Appendix A.}
\label{app:theorem}

% Note: in this sample, the section number is hard-coded in. Following
% proper LaTeX conventions, it should properly be coded as a reference:

%In this appendix we prove the following theorem from
%Section~\ref{sec:textree-generalization}:

In this appendix we prove the following theorem from
Section~6.2:

\noindent
{\bf Theorem} {\it Let $u,v,w$ be discrete variables such that $v, w$ do
not co-occur with $u$ (i.e., $u\neq0\;\Rightarrow \;v=w=0$ in a given
dataset $\dataset$). Let $N_{v0},N_{w0}$ be the number of data points for
which $v=0, w=0$ respectively, and let $I_{uv},I_{uw}$ be the
respective empirical mutual information values based on the sample
$\dataset$. Then
\[
	N_{v0} \;>\; N_{w0}\;\;\Rightarrow\;\;I_{uv} \;\leq\;I_{uw}
\]
with equality only if $u$ is identically 0.} \hfill\BlackBox

\noindent
{\bf Proof}. We use the notation:
\[
P_v(i) \;=\;\frac{N_v^i}{N},\;\;\;i \neq 0;\;\;\;
P_{v0}\;\equiv\;P_v(0)\; = \;1 - \sum_{i\neq 0}P_v(i).
\]
These values represent the (empirical) probabilities of $v$
taking value $i\neq 0$ and 0 respectively.  Entropies will be denoted
by $H$. We aim to show that $\fracpartial{I_{uv}}{P_{v0}} < 0$....\\

{\noindent \em Remainder omitted in this sample. See http://www.jmlr.org/papers/ for full paper.}

\vskip 0.2in
\bibliography{sample}

\end{document}

### 2023-01-06 manuscript/manuscript.Rmd
Initial work on the manuscript.

---
title: "Shift-sensitive Power Transformations for Transforming Data to Normality"
author: "Alex Zwanenburg"
date: "2023-01-06"
output: 
  pdf_document:
    latex_engine: lualatex
bibliography: "refs.bib"
header-includes:
   - \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Introduction

Numerical features in datasets may strongly deviate from normal distributions, e.g. by being skewed. This may complicate further analysis. Power transformations [@Tukey1957-rt] can help improve normality of such features. The two most commonly used transformations are that of @Box1964-mz and @Yeo2000-vw. The Box-Cox transformation of a feature $x$ under the transformation parameter $\lambda$ is defined as:

```{=latex}
\begin{equation}
\phi_{\text{BC}}^\lambda (x) = 
\begin{cases}
\left(x^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x) & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
One limitation of th Box-Cox transformation is that it is only defined for $x > 0$. In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x \in \mathbb{R}$:

```{=latex}
\begin{equation}
\phi_{\text{YJ}}^\lambda (x) = 
\begin{cases}
\left( \left( 1 + x \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x \geq 0\\
\log(1 + x) & \text{if } \lambda = 0 \text{ and } x \geq 0\\
-\left( \left( 1 - x\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x < 0\\
-\log(1 - x) & \text{if } \lambda = 2 \text{ and } x < 0
\end{cases}
\end{equation}
```
The $\lambda$-parameter is typically optimised using maximum likelihood optimisation under the assumption that the transformed feature is normally distributed. As noted by Raymaekers and Rousseeuw, this approach is very sensitive to the presence of outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].

Power transformation does not always

Figure showing Ames dataset feature "Year_Built". Panels: Histogram of feature; Q-Q plot of original feature, normal Box-Cox, robust Box-Cox, normal Yeo-Johnson, robust Yeo-Johnson.

# Methods

# Simulation

# Experimental Results

# Discussion

# Conclusion

# Code availability

Shift-sensitive power transformations are implemented in the `power.transform` package.

# Appendix A.

# References

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-01-06 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-Power transformation does not always
+Power transformation does not always yield transformed variables that
 
 Figure showing Ames dataset feature "Year_Built". Panels: Histogram of feature; Q-Q plot of original feature, normal Box-Cox, robust Box-Cox, normal Yeo-Johnson, robust Yeo-Johnson.
+
+We make the following contributions:
+
+-   We devise shift-sensitive versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
+
+-   We define an empirical measure of the goodness of fit for detecting cases when power transformations fail to yield a normally distributed transformed feature.
+
+-   We assess the effect of power transformations on the performance of machine learning models.
 

### 2023-01-10 manuscript/manuscript.Rmd
WIP on Figure 1.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(
+  echo = TRUE,
+  fig.align= "center")
+
+library(power.transform)
+
+# Base font size.
+base_size <- 10
+
+get_annotation_settings <- function(ggtheme=NULL){
+  # Import formatting settings from the provided ggtheme.
+
+  # Find the text size for the table. This is based on text sizes in the
+  # ggtheme.
+  fontsize <- ggtheme$text$size
+  fontsize_rel <- 1.0
+
+  # Attempt to base the text size on the general axis.text attribute.
+  if(!is.null(ggtheme$axis.text$size)){
+    if(inherits(ggtheme$axis.text$size, "rel")){
+      # Find the relative text size of axis text.
+      fontsize_rel <- as.numeric(ggtheme$axis.text$size)
+
+    } else {
+      # Set absolute text size.
+      fontsize <- ggtheme$axis.text$size
+      fontsize_rel <- 1.0
+    }
+  }
+
+  # Attempt to refine the text size using the axis.text.y attribute in
+  # particular.
+  if(!is.null(ggtheme$axis.text.y$size)){
+    if(inherits(ggtheme$axis.text.y$size, "rel")){
+      # Set relative text size of axis text
+      fontsize_rel <- as.numeric(ggtheme$axis.text.y$size)
+
+    } else {
+      # Set absolute text size.
+      fontsize <- as.numeric(ggtheme$axis.text.y$size)
+      fontsize_rel <- 1.0
+    }
+  }
+
+  # Update the text size using the magical ggplot2 point size (ggplot2:::.pt).
+  geom_text_size <- fontsize * fontsize_rel / 2.845276
+
+  # Obtain lineheight
+  lineheight <- ggtheme$text$lineheight
+  if(!is.null(ggtheme$axis.text$lineheight)) lineheight <- ggtheme$axis.text$lineheight
+  if(!is.null(ggtheme$axis.text.y$lineheight)) lineheight <- ggtheme$axis.text.y$lineheight
+
+  # Obtain family
+  fontfamily <- ggtheme$text$family
+  if(!is.null(ggtheme$axis.text$family)) fontfamily <- ggtheme$axis.text$family
+  if(!is.null(ggtheme$axis.text.y$family)) fontfamily <- ggtheme$axis.text.y$family
+  if(!is.null(ggtheme$axis.text.x$family)) fontfamily <- ggtheme$axis.text.x$family
+
+  # Obtain face
+  fontface <- ggtheme$text$face
+  if(!is.null(ggtheme$axis.text$face)) fontface <- ggtheme$axis.text$face
+  if(!is.null(ggtheme$axis.text.y$face)) fontface <- ggtheme$axis.text.y$face
+  if(!is.null(ggtheme$axis.text.x$face)) fontface <- ggtheme$axis.text.x$face
+
+  # Obtain colour
+  colour <- ggtheme$text$colour
+  if(!is.null(ggtheme$axis.text$colour)) colour <- ggtheme$axis.text$colour
+  if(!is.null(ggtheme$axis.text.y$colour)) colour <- ggtheme$axis.text.y$colour
+  if(!is.null(ggtheme$axis.text.x$colour)) colour <- ggtheme$axis.text.x$colour
+
+  return(list(
+    "geom_text_size"=geom_text_size,
+    "fontsize"=fontsize,
+    "fontsize_rel"=fontsize_rel,
+    "colour"=colour,
+    "family"=fontfamily,
+    "face"=fontface,
+    "lineheight"=lineheight))
+}
+
 ```
@@
 
-Numerical features in datasets may strongly deviate from normal distributions, e.g.
+Numerical variables in datasets may strongly deviate from normal distributions, e.g.
 by being skewed.
@@
 ```
-The $\lambda$-parameter is typically optimised using maximum likelihood optimisation under the assumption that the transformed feature is normally distributed.
+The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
 As noted by Raymaekers and Rousseeuw, this approach is very sensitive to the presence of outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].
 
-Power transformation does not always yield transformed variables that
+Power transformation does not guarantee that transformed variables are normally distributed.
+In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure \@ref(fig:reduced-normality).
+
+``` {r reduced-normality, echo="FALSE", include="FALSE", fig.cap="Power transformations may reduce normality depending on location.
+Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$.
+For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed.
+Panel A and B show the $\\lambda$ value found through maximum likelihood estimation for Box-Cox and Yeo-Johnson transformations, respectively."}
+
+compute_lambda <- function(mu, x, method){
+  # Create transformer object.
+  transformer <- suppressWarnings(
+    power.transform::find_transformation_parameters(
+      x = x + mu,
+      method = method,
+      robust = FALSE,
+      shift = FALSE,
+      lambda = NULL))
+
+  # Return lambda value.
+  return(transformer@lambda)
+}
+
+x <- stats::rnorm(10000L)
+shift_range <- 10^seq(from=0, to=6, by=0.1)
+
+box_cox_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "box_cox")
+
+yeo_johnson_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "yeo_johnson")
+
+data <- data.table::data.table(
+  "mu" = log10(c(shift_range, shift_range)),
+  "lambda" = c(box_cox_values, yeo_johnson_values),
+  "method" = factor(x=rep(c("Box-Cox", "Yeo-Johnson"), each = length(shift_range), levels=c("Box-Cox", "Yeo-Johnson"))))
+
+annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
+
+p <- ggplot2::ggplot(
+  data = data,
+  mapping = ggplot2::aes(
+    x = mu,
+    y = lambda,
+    colour = method))
+p <- p + ggplot2::theme_light(base_size=base_size)
+p <- p + ggplot2::geom_point()
+p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
+p <- p + ggplot2::annotate(
+  geom = "text",
+  x = 6,
+  y = 1,
+  label = "expected",
+  colour=annotation_settings$colour,
+  family=annotation_settings$family,
+  fontface=annotation_settings$face,
+  size=annotation_settings$geom_text_size,
+  vjust=1.5,
+  hjust=1.0)
+p <- p + ggplot2::scale_x_continuous(labels = scales::math_format())
+
+p
+```
 

### 2023-01-10 manuscript/manuscript.Rmd
Set seed for reproducibility.
Prevent printing code.
Adapted figure labels and inserted figure caption into chunk.

@@
 
-library(power.transform)
+# Allow for defining figure captions within a chunk.
+knitr::opts_knit$set(
+  eval.after = "fig.cap")
+
+require(power.transform)
+require(ggplot2)
+require(data.table)
+require(latex2exp)
 
@@
 base_size <- 10
+
+# Set seed.
+set.seed(19L)
 
@@
 Power transformation does not guarantee that transformed variables are normally distributed.
-In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure \@ref(fig:reduced-normality).
+In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure.
 
-``` {r reduced-normality, echo="FALSE", include="FALSE", fig.cap="Power transformations may reduce normality depending on location.
-Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$.
-For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed.
-Panel A and B show the $\\lambda$ value found through maximum likelihood estimation for Box-Cox and Yeo-Johnson transformations, respectively."}
+``` {r reduced-normality, echo=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Power transformations may reduce normality depending on location. ",
+  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
+  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
+  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
+  "and even breaks down due to numerical issues for $\\mu > 10^3$.")
 
@@
   hjust=1.0)
-p <- p + ggplot2::scale_x_continuous(labels = scales::math_format())
+p <- p + ggplot2::scale_x_continuous(
+  name = latex2exp::TeX("$\\mu$"),
+  labels = scales::math_format())
+
+p <- p + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"))
 
@@
 # References
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-01-13 manuscript/manuscript.Rmd
WIP on manuscript.

@@
    - \usepackage{amsmath}
+   - \DeclareMathOperator*{\argmax}{argmax}
+   - \DeclareMathOperator{\sgn}{sgn}
 ---
@@
 Power transformation does not guarantee that transformed variables are normally distributed.
-In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure.
+In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure **INSERT REFERENCE**.
+This is problematic for automated power transformations, for three reasons.
+First, normality is often desired in machine learning applications.
+Second, a large negative or positve $\lambda$-parameter may lead to numeric issues.
+Third, we currently do not have a test to automatically reject poor transformations.
+Statistical tests for normality exist, such as the Shapiro-Wilk test [@Shapiro1965-zd].
+Given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for practical applications of automated power transformations.
 
-``` {r reduced-normality, echo=FALSE, fig.cap=cap}
+```{r reduced-normality, echo=FALSE, fig.cap=cap}
 cap <- paste0(
@@
 }
+
+# Non-standard evaluation
+mu <- lambda <- method <- NULL
 
@@
 
-Figure showing Ames dataset feature "Year_Built". Panels: Histogram of feature; Q-Q plot of original feature, normal Box-Cox, robust Box-Cox, normal Yeo-Johnson, robust Yeo-Johnson.
-
-We make the following contributions:
+Here we make the following contributions:
 
@@
 
+## Shift-sensitive power transformation
+
+Box-Cox and Yeo-Johnson transformations are modified by adding a shift parameter $x_0$.
+The modified Box-Cox transformation of a feature $x$ under transformation parameter $\lambda$ and shift parameter $x_0$ is then:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{BC}}^{\lambda, x_0} (x) = 
+\begin{cases}
+\left( \left(x - x_0 \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log(x - x_0) & \text{if } \lambda = 0
+\end{cases}
+\end{equation}
+```
+Where $x - x_0 > 0$. Likewise, the modified Yeo-Johnson transformation of a feature $x$ under transformation parameter $\lambda$ and shift parameter $x_0$ is:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{YJ}}^{\lambda, x_0} (x) = 
+\begin{cases}
+\left( \left( 1 + x - x_0\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x - x_0 \geq 0\\
+\log(1 + x - x_0) & \text{if } \lambda = 0 \text{ and } x - x_0 \geq 0\\
+-\left( \left( 1 - x - x_0\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x - x_0< 0\\
+-\log(1 - x - x_0) & \text{if } \lambda = 2 \text{ and } x - x_0 < 0
+\end{cases}
+\end{equation}
+```
+For both modified transformations, $\lambda$ and $x_0$ are obtained by maximising the normal log-likelihood function.
+For the modified Box-Cox transformation, and ignoring constant terms, this can be written as:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{BC}}, \hat{x}_0^{\text{BC}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \log(x_i - x_0)
+\end{equation}
+```
+where $\sigma^2$ is the variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (x)$. Similarly for the modified Yeo-Johnson transformation:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{YJ}}, \hat{x}_0^{\text{YJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|)
+\end{equation}
+```
+where $\sigma^2$ is the variance of Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (x)$.
+
+## Robust shift-sensitive power transformations
+
+Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
+Their presence may lead to poor transformations to normality, as shown in Figure **INSERT REFERENCE**.
+As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e.
+achieve *central normality*.
+To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
+Because $\lambda$ is updated during this procedure, simultaneous estimation of $\lambda$ and $x_0$ is not possible.
+
 # Simulation
@@
 # Discussion
+
+In their work, Box and Cox already mention a transformation with a shift parameter, but preferred the now well-known version in Eq.
+**INSERT REFERENCE** for the theoretical analysis in their paper [@Box1964-mz].
 

### 2023-01-15 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
-Because $\lambda$ is updated during this procedure, simultaneous estimation of $\lambda$ and $x_0$ is not possible.
+Because $\lambda$ is updated during this procedure, simultaneous estimation of $\lambda$ and $x_0$ for modified power transformations is not possible.
+
+In essence, achieving central normality requires identifying outliers in the data, and weighting such instances during the optimisation process.
+@Raymaekers2021-kq do this (during the final steps of the iterative procedure) by introducing a weighted maximum likelihood estimation.
+This weighted maximum log-likelihood estimator is based on equations **INSERT REFERENCE**.
+The weighted maximum log-likelihood estimator for the modified Box-Cox power transformation is:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
+\end{equation}
+```
+where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (x)$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+Analogously, the weighted maximum log-likelihood estimator for the modified Yeo-Johnson power transformation is:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
+\end{equation}
+```
+where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (x)$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter.
 

### 2023-01-20 manuscript/manuscript.Rmd
WIP on manuscript.

@@
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and even breaks down due to numerical issues for $\\mu > 10^3$.")
+  "and even breaks down due to numerical issues for $\\mu > 10^3$. For values $\\mu \\leq 10^3$, $\\lambda$ values are similar between methods.")
 
@@
   "lambda" = c(box_cox_values, yeo_johnson_values),
-  "method" = factor(x=rep(c("Box-Cox", "Yeo-Johnson"), each = length(shift_range), levels=c("Box-Cox", "Yeo-Johnson"))))
+  "method" = factor(
+    x=rep(
+      c("Box-Cox", "Yeo-Johnson"),
+      each = length(shift_range)),
+    levels=c("Box-Cox", "Yeo-Johnson")))
 
@@
 
-In essence, achieving central normality requires identifying outliers in the data, and weighting such instances during the optimisation process.
+In essence, obtaining transformation parameters that provide central normality requires identifying outliers in the data, and weighting such instances during the optimisation process.
 @Raymaekers2021-kq do this (during the final steps of the iterative procedure) by introducing a weighted maximum likelihood estimation.
@@
 ```
-Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter.
+Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter:
+
+```{=latex}
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} \leq Q_{\mathcal{N}} (0.995)\\
+0 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} > Q_{\mathcal{N}} (0.995)
+\end{cases}
+\end{equation}
+```
+Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
+In short, weights for all standardised transformed feature values that exceed the value corresponding to the $99.5^\text{th}$ percentile of the normal distribution ($Q_{\mathcal{N}}(0.995)$) are set to 0, and remaining weights are left at 1.
+This works well if outliers are relatively rare in the transformed feature values, which we cannot *a priori* assume in our use case.
+Using lower quantile values comes at the cost of essentially trimming more and more instances from the dataset.
+
+Another option is to weight instances based on their absolute residuals, $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - Q_{\mathcal{N}}(q_i) \right|$, where $q_i$ is equispaced quantile probability for instance $i$.
+This weights instances by the distance from the expected normal distribution.
+Though we refrain from specifying the exact function here, one would down-weight instances with large absolute residuals more strongly than small absolute residuals.
+While this approach is able to deal with outliers, there is no guarantee that large absolute residuals are strictly outliers.
+They may also be found in the central region, and down-weighting there might lead to suboptimal solutions.
+
+We propose a practical approach.
+Let $\kappa$ be the fraction of central instances of feature $x$, i.e.
+those located between $q_i=(1-\kappa)/2$ and $q_i=(1 + \kappa)/2$.
+For example, if $\kappa = 0.80$, those instances with values between the $10^{\text{th}}$ and $90^{\text{th}}$ percentiles would be considered to be central.
+We then use Tukey's tapered cosine window to assign weights [@Tukey1967-eb]:
+
+```{=latex}
+\begin{equation}
+w_i = 
+\begin{cases}
+\frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 - \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right] & \text{, } 0 \leq i - 0.5 < (1 - \kappa) /2\\
+1 & \text{, } (1 - \kappa) /2 \leq i - 0.5 \leq (1 + \kappa) /2\\
+\frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 + \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right]& \text{, } (1 + \kappa) /2 < i - 0.5 \leq 1
+\end{cases}
+\end{equation}
+```
+Here, $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
+
+## Goodness of fit
 
 # Simulation
+
+For modified transformations show:
+- Comparison with figure 1.
+- A shifted left-skewed distribution, with the same shifts.
+- A shifted right-skewed distribution, with the same shifts.
+
+```{r shifted-distributions, echo=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location. ",
+  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
+  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
+  "Unlike the orginal power transformations, $\\lambda = 1$ is consistently found using shift-sensitive power transformations.")
+
+compute_lambda <- function(mu, x, method, shift=FALSE, robust=FALSE){
+  # Create transformer object.
+  transformer <- suppressWarnings(
+    power.transform::find_transformation_parameters(
+      x = x + mu,
+      method = method,
+      robust = robust,
+      shift = shift,
+      lambda = NULL))
+
+  # Return lambda value.
+  return(transformer@lambda)
+}
+
+# Non-standard evaluation
+mu <- lambda <- method <- NULL
+
+x <- stats::rnorm(10000L)
+shift_range <- 10^seq(from=0, to=6, by=0.1)
+
+box_cox_shift_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "box_cox",
+  shift = TRUE,
+  robust = FALSE)
+
+box_cox_shift_robust_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "box_cox",
+  shift = TRUE,
+  robust = TRUE)
+
+yeo_johnson_shift_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "yeo_johnson",
+  shift = TRUE,
+  robust = FALSE)
+
+yeo_johnson_shift_robust_values <- sapply(
+  shift_range,
+  compute_lambda,
+  x = x,
+  method = "yeo_johnson",
+  shift = TRUE,
+  robust = TRUE)
+
+data <- data.table::data.table(
+  "mu" = log10(c(shift_range, shift_range)),
+  "lambda" = c(
+    box_cox_values,
+    box_cox_shift_values,
+    box_cox_shift_robust_values,
+    yeo_johnson_values,
+    yeo_johnson_shift_values,
+    yeo_johnson_shift_robust_values),
+  "method" = factor(
+    x=rep(
+      c("Box-Cox", "Yeo-Johnson"),
+      each = 3L * length(shift_range)),
+    levels=c("Box-Cox", "Yeo-Johnson")),
+  "version" = factor(
+    x=rep(
+      c("original", "shift-sensitive", "robust shift-sensitive"),
+      each = length(shift_range),
+      times = 2L),
+    levels = c("original", "shift-sensitive", "robust shift-sensitive")))
+
+annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
+
+p <- ggplot2::ggplot(
+  data = data,
+  mapping = ggplot2::aes(
+    x = mu,
+    y = lambda,
+    colour = method,
+    shape = version))
+p <- p + ggplot2::theme_light(base_size=base_size)
+p <- p + ggplot2::geom_point()
+p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
+p <- p + ggplot2::annotate(
+  geom = "text",
+  x = 6,
+  y = 1,
+  label = "expected",
+  colour=annotation_settings$colour,
+  family=annotation_settings$family,
+  fontface=annotation_settings$face,
+  size=annotation_settings$geom_text_size,
+  vjust=1.5,
+  hjust=1.0)
+p <- p + ggplot2::scale_x_continuous(
+  name = latex2exp::TeX("$\\mu$"),
+  labels = scales::math_format())
+
+p <- p + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"))
+
+p
+```
+
+For robustness simulate:
+- A shifted normal distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
+- A shifted left-skewed distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
+- A shifted right-skewed distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
+- To show: lambda for modified, and robust modified Yeo-Johnson distribution for: Raymaekers trimmed, Tukey Window
 

### 2023-01-20 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 base_size <- 10
+
+# Set general theme.
+plot_theme <- ggplot2::theme_light(base_size=base_size)
 
@@
 
-p <- ggplot2::ggplot(
-  data = data,
+p_bc <- ggplot2::ggplot(
+  data = data[method == "Box-Cox"],
   mapping = ggplot2::aes(
@@
     colour = method))
-p <- p + ggplot2::theme_light(base_size=base_size)
-p <- p + ggplot2::geom_point()
-p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
-p <- p + ggplot2::annotate(
+p_bc <- p_bc + plot_theme
+p_bc <- p_bc + ggplot2::geom_point()
+p_bc <- p_bc + ggplot2::geom_hline(
+  yintercept = 1.0, 
+  linetype="longdash", 
+  colour="grey40")
+p_bc <- p_bc + ggplot2::annotate(
   geom = "text",
@@
   size=annotation_settings$geom_text_size,
-  vjust=1.5,
+  vjust=-1.0,
   hjust=1.0)
-p <- p + ggplot2::scale_x_continuous(
+p_bc <- p_bc + ggplot2::scale_x_continuous(
   name = latex2exp::TeX("$\\mu$"),
   labels = scales::math_format())
+p_bc <- p_bc + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"),
+  limits = c(-3.6, 1.1))
+p_bc <- p_bc + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE,
+  guide="none")
 
-p <- p + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"))
+p_yj <- ggplot2::ggplot(
+  data = data[method == "Yeo-Johnson"],
+  mapping = ggplot2::aes(
+    x = mu,
+    y = lambda,
+    colour = method))
+p_yj <- p_yj + plot_theme
+p_yj <- p_yj + ggplot2::geom_point()
+p_yj <- p_yj + ggplot2::geom_hline(
+  yintercept = 1.0,
+  linetype="longdash",
+  colour="grey40")
+p_yj <- p_yj + ggplot2::annotate(
+  geom = "text",
+  x = 6,
+  y = 1,
+  label = "expected",
+  colour=annotation_settings$colour,
+  family=annotation_settings$family,
+  fontface=annotation_settings$face,
+  size=annotation_settings$geom_text_size,
+  vjust=-1.0,
+  hjust=1.0)
+p_yj <- p_yj + ggplot2::scale_x_continuous(
+  name = latex2exp::TeX("$\\mu$"),
+  labels = scales::math_format())
+p_yj <- p_yj + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"),
+  limits = c(-3.6, 1.1))
+p_yj <- p_yj + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE)
+p_yj <- p_yj + ggplot2::theme(
+  axis.title.y=ggplot2::element_blank(),
+  axis.text.y=ggplot2::element_blank())
 
+p <- egg::ggarrange(plots=list(p_bc, p_yj), ncol=2, draw=FALSE)
 p
@@
 
-# Non-standard evaluation
-mu <- lambda <- method <- NULL
+.create_lambda_shift_data <- function(x, shift_range){
 
-x <- stats::rnorm(10000L)
-shift_range <- 10^seq(from=0, to=6, by=0.1)
+  box_cox_values <- sapply(
+    shift_range,
+    compute_lambda,
+    x = x,
+    method = "box_cox")
+
+  yeo_johnson_values <- sapply(
+    shift_range,
+    compute_lambda,
+    x = x,
+    method = "yeo_johnson")
 
@@
 
-annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
+  return(data)
+}
+
+.create_density_plot <- function(x, plot_theme){
+  p <- ggplot2::ggplot(
+    data = data.table::data.table("x"=x),
+    mapping = ggplot2::aes(
+      x = x))
+  p <- p + plot_theme
+  p <- p + ggplot2::geom_density()
+  p <- p + ggplot2::theme(
+    axis.title = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    panel.grid = ggplot2::element_blank(),
+    panel.border = ggplot2::element_blank())
+
+  return(p)
+}
+
+.create_lambda_shift_plot <- function(data, plot_theme, limits, expected=FALSE){
+
+  annotation_settings <- get_annotation_settings(plot_theme)
 
@@
       shape = version))
-p <- p + ggplot2::theme_light(base_size=base_size)
+  p <- p + plot_theme
   p <- p + ggplot2::geom_point()
-p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
+  p <- p + ggplot2::geom_hline(
+    yintercept = 1.0, 
+    linetype="longdash", 
+    colour="grey40")
+  p <- p + ggplot2::scale_x_continuous(
+    name = latex2exp::TeX("$\\mu$"),
+    labels = scales::math_format())
+  p <- p + ggplot2::scale_y_continuous(
+    name = latex2exp::TeX("$\\lambda$"),
+    limits = limits)
+
+  if(expected){
       p <- p + ggplot2::annotate(
@@
     size=annotation_settings$geom_text_size,
-  vjust=1.5,
+    vjust=-1.0,
     hjust=1.0)
-p <- p + ggplot2::scale_x_continuous(
-  name = latex2exp::TeX("$\\mu$"),
-  labels = scales::math_format())
+  }
 
-p <- p + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"))
+  return(p)
+}
 
+# Non-standard evaluation
+mu <- lambda <- method <- NULL
+
+# Normal distribution.
+x <- stats::rnorm(10000L)
+shift_range <- 10^seq(from=0, to=6, by=0.1)
+
+# Right skewed data
+x_right_skewed <- (x - min(x) + 1)^0.1
+x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)
+
+# Left skewed data
+x_left_skewed <- (x - min(x))^3
+x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)
+
+#### Normal distribution -------------------------------------------------------
+
+data <- .create_lambda_shift_data(
+  x = x,
+  shift_range = shift_range)
+
+# Density
+p_dens_normal <- .create_density_plot(
+  x=x,
+  plot_theme=plot_theme)
+
+# Box-Cox
+p_bc_normal <- .create_lambda_shift_plot(
+  data=data[method == "Box-Cox"],
+  plot_theme=plot_theme,
+  limits = c(-3.6, 1.1))
+p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE,
+  guide="none")
+p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")
+
+# Yeo-Johnson
+p_yj_normal <- .create_lambda_shift_plot(
+  data=data[method == "Yeo-Johnson"],
+  plot_theme=plot_theme,
+  limits = c(-3.6, 1.1))
+p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE,
+  guide="none")
+p_yj_normal <- p_yj_normal + ggplot2::theme(
+  axis.title.y=ggplot2::element_blank(),
+  axis.text.y=ggplot2::element_blank())
+
+#### Right skewed distribution -------------------------------------------------
+
+data <- .create_lambda_shift_data(
+  x = x_right_skewed,
+  shift_range = shift_range)
+
+# Density
+p_dens_right <- .create_density_plot(
+  x = x_right_skewed,
+  plot_theme = plot_theme)
+
+# Box-Cox
+p_bc_right <- .create_lambda_shift_plot(
+  data=data[method == "Box-Cox"],
+  plot_theme=plot_theme,
+  limits = c(1.0, 65))
+p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE,
+  guide="none")
+p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")
+
+# Yeo-Johnson
+p_yj_right <- .create_lambda_shift_plot(
+  data=data[method == "Yeo-Johnson"],
+  plot_theme=plot_theme,
+  limits = c(1.0, 65))
+p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE)
+p_yj_right <- p_yj_right + ggplot2::theme(
+  axis.title.y=ggplot2::element_blank(),
+  axis.text.y=ggplot2::element_blank())
+
+#### Left skewed distribution --------------------------------------------------
+
+data <- .create_lambda_shift_data(
+  x = x_left_skewed,
+  shift_range = shift_range)
+
+# Density
+p_dens_left <- .create_density_plot(
+  x = x_left_skewed,
+  plot_theme = plot_theme)
+
+# Box-Cox
+p_bc_left <- .create_lambda_shift_plot(
+  data=data[method == "Box-Cox"],
+  plot_theme=plot_theme,
+  limits = c(-10.5, 0.5))
+p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE,
+  guide="none")
+p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")
+
+# Yeo-Johnson
+p_yj_left <- .create_lambda_shift_plot(
+  data=data[method == "Yeo-Johnson"],
+  plot_theme=plot_theme,
+  limits = c(-10.5, 0.5))
+p_yj_left <- p_yj_left + paletteer::scale_color_paletteer_d(
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE,
+  guide = "none")
+p_yj_left <- p_yj_left + ggplot2::theme(
+  axis.title.y=ggplot2::element_blank(),
+  axis.text.y=ggplot2::element_blank())
+
+p <- egg::ggarrange(
+  plots=list(
+    p_dens_normal, p_bc_normal, p_yj_normal,
+    p_dens_right, p_bc_right, p_yj_right,
+    p_dens_left, p_bc_left, p_yj_left),
+  ncol=3,
+  draw=FALSE)
 p

### 2023-01-20 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-For modified transformations show:
-- Comparison with figure 1.
-- A shifted left-skewed distribution, with the same shifts.
-- A shifted right-skewed distribution, with the same shifts.
+To assess whether the modified power transformations lead to $\lambda$ values that are invariant to shifts, we simulate three different distributions.
+The first distribution is a normal distribution, as previously shown in Figure **INSERT REFERENCE**.
+We first randomly draw $10000$ values from a normal distribution: $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$.
+The second distribution is a right-skewed distribution $x_{\text{right}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) + 1\right)^{\frac{1}{10}}$.
+The third distribution is a left-skewed distribution $x_{\text{left}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) \right)^3$.
+Subsequently, all three distributions are standardised to have mean $0$ and standard deviation $1$.
+We then compute transformation parameter $\lambda$ using the original definitions (Eq.
+**INSERT REFERENCE** and **INSERT REFERENCE**) and the modified definitions (Eq.
+**INSERT REFERENCE** and **INSERT REFERENCE**) for each distribution, after adding a positive value $d$ to each distribution.
+Here, $d \in [1, 10^6]$.
 
@@
 cap <- paste0(
-  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location. ",
-  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
-  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
-  "Unlike the orginal power transformations, $\\lambda = 1$ is consistently found using shift-sensitive power transformations.")
+  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location.")
 
@@
   p <- p + ggplot2::scale_x_continuous(
-    name = latex2exp::TeX("$\\mu$"),
+    name = latex2exp::TeX("$d$"),
     labels = scales::math_format())
@@
 
-data <- .create_lambda_shift_data(
+data_normal <- .create_lambda_shift_data(
   x = x,
@@
 p_bc_normal <- .create_lambda_shift_plot(
-  data=data[method == "Box-Cox"],
-  plot_theme=plot_theme,
+  data = data_normal[method == "Box-Cox"],
+  plot_theme = plot_theme,
   limits = c(-3.6, 1.1))
 p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE,
-  guide="none")
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE,
+  guide = "none")
 p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")
@@
 p_yj_normal <- .create_lambda_shift_plot(
-  data=data[method == "Yeo-Johnson"],
-  plot_theme=plot_theme,
+  data=data_normal[method == "Yeo-Johnson"],
+  plot_theme = plot_theme,
   limits = c(-3.6, 1.1))
 p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE,
-  guide="none")
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE,
+  guide = "none")
 p_yj_normal <- p_yj_normal + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank())
 
@@
 
-data <- .create_lambda_shift_data(
+data_right <- .create_lambda_shift_data(
   x = x_right_skewed,
@@
 p_bc_right <- .create_lambda_shift_plot(
-  data=data[method == "Box-Cox"],
-  plot_theme=plot_theme,
+  data = data_right[method == "Box-Cox"],
+  plot_theme = plot_theme,
   limits = c(1.0, 65))
 p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE,
-  guide="none")
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE,
+  guide = "none")
 p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")
@@
 p_yj_right <- .create_lambda_shift_plot(
-  data=data[method == "Yeo-Johnson"],
-  plot_theme=plot_theme,
+  data = data_right[method == "Yeo-Johnson"],
+  plot_theme = plot_theme,
   limits = c(1.0, 65))
 p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE)
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE)
 p_yj_right <- p_yj_right + ggplot2::theme(
@@
 
-data <- .create_lambda_shift_data(
+data_left <- .create_lambda_shift_data(
   x = x_left_skewed,
@@
 p_bc_left <- .create_lambda_shift_plot(
-  data=data[method == "Box-Cox"],
-  plot_theme=plot_theme,
+  data = data_left[method == "Box-Cox"],
+  plot_theme = plot_theme,
   limits = c(-10.5, 0.5))
 p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE,
-  guide="none")
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE,
+  guide = "none")
 p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")
@@
 p_yj_left <- .create_lambda_shift_plot(
-  data=data[method == "Yeo-Johnson"],
-  plot_theme=plot_theme,
+  data = data_left[method == "Yeo-Johnson"],
+  plot_theme = plot_theme,
   limits = c(-10.5, 0.5))
@@
   axis.text.y=ggplot2::element_blank())
+
+#### Aggregate -----------------------------------------------------------------
+
+data_normal[, "distribution":="normal"]
+data_left[, "distribution":="left-skewed"]
+data_right[, "distribution":="right-skewed"]
+
+aggregate_data <- data.table::rbindlist(list(data_normal, data_left, data_right))[, list("mean_lambda"=mean(lambda), "sd_lambda"=stats::sd(lambda)), by=c("method", "version", "distribution")]
 
@@
 ```
+
+The result is shown in Figure **INSERT REFERENCE**.
+For each distribution, transformation parameter $\lambda$ varies with $d$ when estimated using the original definitions.
+In contrast, estimations of $\lambda$ modified power transformations were invariant to $d$.
 

### 2023-01-24 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 cap <- paste0(
-  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location.")
+  "Power transformations modified to be shift-sensitive produce transformation parameters are invariant to location.")
 
@@
     panel.border = ggplot2::element_blank())
+  p <- p + ggplot2::xlim(c(-5, 5))
 
@@
   plot_theme=plot_theme)
+p_dens_normal <- p_dens_normal + ggplot2::ggtitle(label="normal distribution")
+p_dens_normal <- p_dens_normal + ggplot2::theme(plot.title=ggplot2::element_text(
+  hjust=0.5,
+  size=ggplot2::rel(1.0)))
 
@@
   guide = "none")
-p_yj_normal <- p_yj_normal + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
+p_yj_normal <- p_yj_normal + ggplot2::scale_shape_discrete(guide="none")
+# p_yj_normal <- p_yj_normal + ggplot2::theme(
+#   axis.title.y = ggplot2::element_blank(),
+#   axis.text.y = ggplot2::element_blank())
 
@@
   plot_theme = plot_theme)
+p_dens_right <- p_dens_right + ggplot2::ggtitle(label="right-skewed distribution")
+p_dens_right <- p_dens_right + ggplot2::theme(plot.title=ggplot2::element_text(
+  hjust=0.5,
+  size=ggplot2::rel(1.0)))
 
@@
   palette = "ggthemes::Tableau_10",
-  drop = FALSE)
-p_yj_right <- p_yj_right + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
+  drop = FALSE,
+  guide = "none")
+p_yj_right <- p_yj_right + ggplot2::scale_shape_discrete(guide="none")
+# p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
+#   palette = "ggthemes::Tableau_10",
+#   drop = FALSE)
+# p_yj_right <- p_yj_right + ggplot2::theme(
+#   axis.title.y=ggplot2::element_blank(),
+#   axis.text.y=ggplot2::element_blank())
 
@@
   plot_theme = plot_theme)
+p_dens_left <- p_dens_left + ggplot2::ggtitle(label="left-skewed distribution")
+p_dens_left <- p_dens_left + ggplot2::theme(plot.title=ggplot2::element_text(
+  hjust=0.5,
+  size=ggplot2::rel(1.0)))
 
@@
   palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")
+  drop = FALSE)
+# p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
+#   palette = "ggthemes::Tableau_10",
+#   drop = FALSE,
+#   guide = "none")
+# p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")
 
@@
   guide = "none")
-p_yj_left <- p_yj_left + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
+p_yj_left <- p_yj_left + ggplot2::scale_shape_discrete(guide="none")
+# p_yj_left <- p_yj_left + ggplot2::theme(
+#   axis.title.y=ggplot2::element_blank(),
+#   axis.text.y=ggplot2::element_blank())
 
@@
   ncol=3,
+  heights=c(0.5, 1.0, 1.0),
+  byrow=FALSE,
   draw=FALSE)

### 2023-01-27 manuscript/manuscript.Rmd
Update manuscript.Rmd

--- 2023-01-20 manuscript/manuscript.Rmd
@@
    - \usepackage{amsmath}
+   - \usepackage{booktabs}
    - \DeclareMathOperator*{\argmax}{argmax}
@@
 require(latex2exp)
+require(coro)
 
@@
 plot_theme <- ggplot2::theme_light(base_size=base_size)
-
-# Set seed.
-set.seed(19L)
 
@@
     "lineheight"=lineheight))
+}
+
+parse_mean_sd_latex <- function(mu, sigma, digits=2L){
+  return(paste0(
+    "$", round(mu, digits=digits), " \\pm ", round(sigma, digits=digits), "$"
+  ))
 }
@@
 mu <- lambda <- method <- NULL
+
+# Set seed.
+set.seed(19L)
 
@@
 
-```{r shifted-distributions, echo=FALSE, fig.cap=cap}
+```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
-  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location.")
+  "Power transformations modified to be shift-sensitive produce transformation parameters are invariant to location.")
 
-compute_lambda <- function(mu, x, method, shift=FALSE, robust=FALSE){
-  # Create transformer object.
-  transformer <- suppressWarnings(
-    power.transform::find_transformation_parameters(
-      x = x + mu,
-      method = method,
-      robust = robust,
-      shift = shift,
-      lambda = NULL))
+# Non-standard evaluation
+d <- distribution <- method <- mean_lambda <- sd_lambda <- NULL
 
-  # Return lambda value.
-  return(transformer@lambda)
-}
-
-.create_lambda_shift_data <- function(x, shift_range){
-
-  box_cox_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "box_cox")
-
-  yeo_johnson_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "yeo_johnson")
-
-  box_cox_shift_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "box_cox",
-    shift = TRUE,
-    robust = FALSE)
-
-  box_cox_shift_robust_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "box_cox",
-    shift = TRUE,
-    robust = TRUE)
-
-  yeo_johnson_shift_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "yeo_johnson",
-    shift = TRUE,
-    robust = FALSE)
-
-  yeo_johnson_shift_robust_values <- sapply(
-    shift_range,
-    compute_lambda,
-    x = x,
-    method = "yeo_johnson",
-    shift = TRUE,
-    robust = TRUE)
-
-  data <- data.table::data.table(
-    "mu" = log10(c(shift_range, shift_range)),
-    "lambda" = c(
-      box_cox_values,
-      box_cox_shift_values,
-      box_cox_shift_robust_values,
-      yeo_johnson_values,
-      yeo_johnson_shift_values,
-      yeo_johnson_shift_robust_values),
-    "method" = factor(
-      x=rep(
-        c("Box-Cox", "Yeo-Johnson"),
-        each = 3L * length(shift_range)),
-      levels=c("Box-Cox", "Yeo-Johnson")),
-    "version" = factor(
-      x=rep(
-        c("original", "shift-sensitive", "robust shift-sensitive"),
-        each = length(shift_range),
-        times = 2L),
-      levels = c("original", "shift-sensitive", "robust shift-sensitive")))
-
-  return(data)
-}
-
+# Density plot.
 .create_density_plot <- function(x, plot_theme){
@@
     panel.grid = ggplot2::element_blank(),
-    panel.border = ggplot2::element_blank())
+    panel.border = ggplot2::element_blank(),
+    plot.title=ggplot2::element_text(
+      hjust=0.5,
+      size=ggplot2::rel(1.0)))
+  p <- p + ggplot2::xlim(c(-5, 5))
 
@@
 
-.create_lambda_shift_plot <- function(data, plot_theme, limits, expected=FALSE){
-
-  annotation_settings <- get_annotation_settings(plot_theme)
+# Lambda plot,
+.create_lambda_shift_plot <- function(data, plot_theme, limits, guide=FALSE){
 
@@
     mapping = ggplot2::aes(
-      x = mu,
+      x = d,
       y = lambda,
@@
   p <- p + ggplot2::geom_point()
-  p <- p + ggplot2::geom_hline(
-    yintercept = 1.0, 
-    linetype="longdash", 
-    colour="grey40")
   p <- p + ggplot2::scale_x_continuous(
@@
 
-  if(expected){
-      p <- p + ggplot2::annotate(
-    geom = "text",
-    x = 6,
-    y = 1,
-    label = "expected",
-    colour=annotation_settings$colour,
-    family=annotation_settings$family,
-    fontface=annotation_settings$face,
-    size=annotation_settings$geom_text_size,
-    vjust=-1.0,
-    hjust=1.0)
+  if(guide){
+     p <- p + paletteer::scale_color_paletteer_d(
+      palette = "ggthemes::Tableau_10",
+      drop = FALSE)
+
+  } else {
+    p <- p + paletteer::scale_color_paletteer_d(
+      palette = "ggthemes::Tableau_10",
+      drop = FALSE,
+      guide = "none")
+    p <- p + ggplot2::scale_shape_discrete(guide="none")
   }
@@
 
-# Non-standard evaluation
-mu <- lambda <- method <- NULL
+# Generator for creating parameters and data for all experiments.
+generate_experiment_data <- coro::generator(
+  function(
+    x_normal,
+    x_right_skewed,
+    x_left_skewed){
+
+    shift_range <- 10^seq(from=0, to=6, by=0.1)
+
+    for(d in shift_range){
+      for(distribution in c("normal", "right-skewed", "left-skewed")){
+        if(distribution == "normal"){
+          x <- x_normal
+
+        } else if(distribution == "right-skewed"){
+          x <- x_right_skewed
+
+        } else if(distribution == "left-skewed"){
+          x <- x_left_skewed
+        }
+
+        for(method in c("box_cox", "yeo_johnson")){
+          for(shift in c(FALSE, TRUE)){
+            for(robust in c(FALSE, TRUE)){
+              # Skip if robust, but not shifted.
+              if(robust && !shift) next
+
+              if(!shift && !robust){
+                version <- "original"
+
+              } else if(!robust){
+                version <- "shift-sensitive"
+
+              } else {
+                version <- "robust shift-sensitive"
+              }
+
+              yield(list(
+                "x" = x + d,
+                "d" = d,
+                "distribution" = distribution,
+                "method" = method,
+                "shift" = shift,
+                "robust" = robust,
+                "version" = version
+              ))
+            }
+          }
+        }
+      }
+    }
+  })
+
+.compute_lambda <- function(parameter_set){
+  # Create transformer object.
+  transformer <- suppressWarnings(
+    power.transform::find_transformation_parameters(
+      x = parameter_set$x,
+      method = parameter_set$method,
+      robust = parameter_set$robust,
+      shift = parameter_set$shift,
+      lambda = NULL))
+
+  return(
+    data.table::data.table(
+      "distribution" = parameter_set$distribution,
+      "method" = parameter_set$method,
+      "version" = parameter_set$version,
+      "d" = log10(parameter_set$d),
+      "lambda" = transformer@lambda
+    ))
+}
+
+#### Computations --------------------------------------------------------------
+
+# Set seed.
+set.seed(19L)
 
 # Normal distribution.
-x <- stats::rnorm(10000L)
-shift_range <- 10^seq(from=0, to=6, by=0.1)
+x_normal <- stats::rnorm(10000L)
 
 # Right skewed data
-x_right_skewed <- (x - min(x) + 1)^0.1
+x_right_skewed <- (x_normal - min(x_normal) + 1)^0.1
 x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)
@@
 # Left skewed data
-x_left_skewed <- (x - min(x))^3
+x_left_skewed <- (x_normal - min(x_normal))^3
 x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)
 
+# Generate all experiments.
+experiments <- coro::collect(generate_experiment_data(
+  x_normal = x_normal,
+  x_right_skewed = x_right_skewed,
+  x_left_skewed = x_left_skewed
+))
+
+# Start cluster
+cl <- parallel::makeCluster(16L)
+
+# Compute all data in parallel.
+data <- parallel::parLapply(
+  cl=cl,
+  X=experiments,
+  fun=.compute_lambda
+)
+
+# Stop cluster.
+parallel::stopCluster(cl)
+
+# Combine data into a single table.
+data <- data.table::rbindlist(data)
+
+# Update distribution, method and version to factors.
+data$distribution <- factor(
+  x = data$distribution,
+  levels = c("normal", "right-skewed", "left-skewed"))
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+data$version <- factor(
+  x = data$version,
+  levels = c("original", "shift-sensitive", "robust shift-sensitive"))
+
 #### Normal distribution -------------------------------------------------------
-
-data_normal <- .create_lambda_shift_data(
-  x = x,
-  shift_range = shift_range)
-
 # Density
 p_dens_normal <- .create_density_plot(
-  x=x,
+  x=x_normal,
   plot_theme=plot_theme)
+p_dens_normal <- p_dens_normal + ggplot2::ggtitle(label="normal distribution")
 
@@
 p_bc_normal <- .create_lambda_shift_plot(
-  data = data_normal[method == "Box-Cox"],
+  data = data[distribution == "normal" & method == "Box-Cox"],
   plot_theme = plot_theme,
   limits = c(-3.6, 1.1))
-p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")
 
@@
 p_yj_normal <- .create_lambda_shift_plot(
-  data=data_normal[method == "Yeo-Johnson"],
+  data=data[distribution == "normal" & method == "Yeo-Johnson"],
   plot_theme = plot_theme,
   limits = c(-3.6, 1.1))
-p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_yj_normal <- p_yj_normal + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
 
 #### Right skewed distribution -------------------------------------------------
-
-data_right <- .create_lambda_shift_data(
-  x = x_right_skewed,
-  shift_range = shift_range)
 
@@
   plot_theme = plot_theme)
+p_dens_right <- p_dens_right + ggplot2::ggtitle(label="right-skewed distribution")
 
@@
 p_bc_right <- .create_lambda_shift_plot(
-  data = data_right[method == "Box-Cox"],
+  data =data[distribution == "right-skewed" & method == "Box-Cox"],
   plot_theme = plot_theme,
   limits = c(1.0, 65))
-p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")
 
@@
 p_yj_right <- .create_lambda_shift_plot(
-  data = data_right[method == "Yeo-Johnson"],
+  data = data[distribution == "right-skewed" & method == "Yeo-Johnson"],
   plot_theme = plot_theme,
   limits = c(1.0, 65))
-p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE)
-p_yj_right <- p_yj_right + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
 
 #### Left skewed distribution --------------------------------------------------
-
-data_left <- .create_lambda_shift_data(
-  x = x_left_skewed,
-  shift_range = shift_range)
 
@@
   plot_theme = plot_theme)
+p_dens_left <- p_dens_left + ggplot2::ggtitle(label="left-skewed distribution")
 
@@
 p_bc_left <- .create_lambda_shift_plot(
-  data = data_left[method == "Box-Cox"],
+  data = data[distribution == "left-skewed" & method == "Box-Cox"],
   plot_theme = plot_theme,
-  limits = c(-10.5, 0.5))
-p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")
+  limits = c(-10.5, 0.5),
+  guide = TRUE)
 
@@
 p_yj_left <- .create_lambda_shift_plot(
-  data = data_left[method == "Yeo-Johnson"],
+  data = data[distribution == "left-skewed" & method == "Yeo-Johnson"],
   plot_theme = plot_theme,
   limits = c(-10.5, 0.5))
-p_yj_left <- p_yj_left + paletteer::scale_color_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE,
-  guide = "none")
-p_yj_left <- p_yj_left + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
 
@@
 
-data_normal[, "distribution":="normal"]
-data_left[, "distribution":="left-skewed"]
-data_right[, "distribution":="right-skewed"]
-
-aggregate_data <- data.table::rbindlist(list(data_normal, data_left, data_right))[, list("mean_lambda"=mean(lambda), "sd_lambda"=stats::sd(lambda)), by=c("method", "version", "distribution")]
+aggregate_data <- data[, list("mean_lambda"=mean(lambda), "sd_lambda"=stats::sd(lambda)), by=c("method", "version", "distribution")]
+aggregate_data$distribution <- factor(
+  x=aggregate_data$distribution,
+  levels = c("normal", "left-skewed", "right-skewed"))
+aggregate_data[, "text":=parse_mean_sd_latex(mu=mean_lambda, sigma=sd_lambda)]
 
@@
 
+```{r, echo=FALSE, results="asis"}
+table_data <- data.table::dcast(
+  aggregate_data,
+  distribution ~ version + method,
+  value.var = "text",
+  sep = " ")
+
+knitr::kable(
+  table_data,
+  caption = "Estimates and deviation of transformation parameter $\\lambda$ under location shifts",
+  label = "shifted-distributions-table",
+  format = "latex",
+  escape = FALSE,
+  booktabs = TRUE)
+```
+
 For robustness simulate:
@@
 - To show: lambda for modified, and robust modified Yeo-Johnson distribution for: Raymaekers trimmed, Tukey Window
+
+```{r}
+# Generator for creating parameters and data for all experiments.
+generate_experiment_data <- coro::generator(
+  function(
+    x_normal,
+    x_right_skewed,
+    x_left_skewed){
+
+    n <- 20
+    k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 5)  
+    # k_range <- c(0.0, 0.001, 0.002, 0.005, 0.010, 0.020, 0.050, 0.100, 0.200)
+
+    for(distribution in c("normal", "right-skewed", "left-skewed")){
+      if(distribution == "normal"){
+        x <- x_normal
+
+      } else if(distribution == "right-skewed"){
+        x <- x_right_skewed
+
+      } else if(distribution == "left-skewed"){
+        x <- x_left_skewed
+      }
+      for(ii in seq_len(20)){
+        for(k in k_range){
+          # Compute interquartile range.
+          interquartile_range <- stats::IQR(x)
+
+          # Compute upper and lower quartiles.
+          q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+          q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+          # Set data where the outliers will be copied into.
+          x_outlier <- x
+
+          if(k != 0.0){
+            n_draw <- ceiling(k * 0.5 * length(x))
+
+            # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+            # than Q3 + 1.5 IQR.
+            outlier <- c(
+              q_lower - 1.5 * interquartile_range - 3.0 * stats::runif(n_draw),
+              q_upper + 1.5 * interquartile_range + 3.0 * stats::runif(n_draw))
+
+            # Randomly insert outlier values.
+            x_outlier[sample(seq_along(x), size=2*n_draw, replace=FALSE)] <- outlier
+          }
+
+          #for(method in c("box_cox", "yeo_johnson")){
+          for(method in c("yeo_johnson")){
+            for(weight_method in c("none", "tukey_window", "tapered_step_window")){
+            # for(weight_method in c("none", "tukey_window", "step_window", "tapered_step_window", "trim_transformation", "trim_residual", "tukey_biweight_transformation", "tukey_biweight_residual", "huber_weight_transformed", "huber_weight_residual")){
+
+              weight_method_name <- weight_method
+
+              for(tukey_window_k in c(0.80, 0.60, 0.40, 0.20, 0.00)){
+                if(weight_method == "none" && tukey_window_k != 0.00) next
+
+                yield(list(
+                  "ii" = ii,
+                  "x" = x_outlier,
+                  "k" = k,
+                  "distribution" = distribution,
+                  "method" = method,
+                  "robust" = weight_method != "none",
+                  "weight_method" = weight_method,
+                  "weight_method_name" = paste0(weight_method_name, "_k_", tukey_window_k),
+                  "tukey_window_k" = tukey_window_k
+                ))
+              }
+
+              # if(weight_method == "tukey_window"){
+              #   for(tukey_window_k in c(0.9, 0.8, 0.7, 0.6, 0.5)){
+              #     yield(list(
+              #       "x" = x_outlier,
+              #       "k" = k,
+              #       "distribution" = distribution,
+              #       "method" = method,
+              #       "robust" = weight_method != "none",
+              #       "weight_method" = weight_method,
+              #       "weight_method_name" = paste0(weight_method_name, "_k_", tukey_window_k),
+              #       "tukey_window_k" = tukey_window_k
+              #     ))
+              #   }
+              # } else {
+              #   yield(list(
+              #     "x" = x_outlier,
+              #     "k" = k,
+              #     "distribution" = distribution,
+              #     "method" = method,
+              #     "robust" = weight_method != "none",
+              #     "weight_method" = weight_method,
+              #     "weight_method_name" = weight_method_name
+              #   ))
+              # }
+            }
+          }
+        }
+      }
+    }
+  })
+
+.compute_lambda <- function(parameter_set){
+  # Create transformer object.
+  transformer <- suppressWarnings(
+    power.transform::find_transformation_parameters(
+      x = parameter_set$x,
+      method = parameter_set$method,
+      robust = parameter_set$robust,
+      shift = TRUE,
+      weight_method = parameter_set$weight_method,
+      k = parameter_set$tukey_window_k,
+      lambda = NULL))
+
+  return(
+    data.table::data.table(
+      "distribution" = parameter_set$distribution,
+      "method" = parameter_set$method,
+      "weight_method" = parameter_set$weight_method,
+      "weight_method_name" = parameter_set$weight_method_name,
+      "k" = parameter_set$k,
+      "lambda" = transformer@lambda,
+      "kappa" = parameter_set$tukey_window_k,
+      "ii" = parameter_set$ii
+    ))
+}
+
+# Set seed.
+set.seed(19L)
+
+# Normal distribution.
+x_normal <- stats::rnorm(10000L)
+
+# Right skewed data
+x_right_skewed <- (x_normal - min(x_normal) + 1)^0.1
+x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)
+
+# Left skewed data
+x_left_skewed <- (x_normal - min(x_normal))^3
+x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)
+
+# Generate all experiments.
+experiments <- coro::collect(generate_experiment_data(
+  x_normal = x_normal,
+  x_right_skewed = x_right_skewed,
+  x_left_skewed = x_left_skewed
+))
+
+# Start cluster
+cl <- parallel::makeCluster(16L)
+
+# Compute all data in parallel.
+data <- parallel::parLapply(
+  cl=cl,
+  X=experiments,
+  fun=.compute_lambda
+)
+
+# Stop cluster.
+parallel::stopCluster(cl)
+
+# Combine data into a single table.
+data <- data.table::rbindlist(data)
+
+# Update distribution, method and version to factors.
+data$distribution <- factor(
+  x = data$distribution,
+  levels = c("normal", "right-skewed", "left-skewed"))
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+data$kappa <- factor(data$kappa)
+
+# Find target lambda values in the absence of any outliers.
+target_data <- data[weight_method=="none" & k == 0.0, mget(c("distribution", "ii", "lambda", "method"))]
+data.table::setnames(target_data, old="lambda", new="target_lambda")
+
+# Merge target lambda values into the main data.
+data <- data[target_data, on=.NATURAL]
+
+# Compute difference with target lambda.
+data[, "diff_lambda":=lambda - target_lambda]
+
+summary_data[, list(deviation=sum(abs(diff_lambda))), by=c("distribution", "weight_method_name")][order(distribution, deviation)]
+
+# Compute average.
+new_data <- data[, list(lambda = mean(lambda)), by=c("k", "kappa", "weight_method", "distribution")]
+
+p <- ggplot2::ggplot(
+  data = new_data[distribution=="left-skewed"],
+  mapping = ggplot2::aes(
+    x = k,
+    y = lambda,
+    colour = kappa,
+    linetype = weight_method))
+p <- p + plot_theme
+# p <- p + ggplot2::geom_point()
+p <- p + ggplot2::geom_line()
+p <- p + ggplot2::scale_x_continuous(
+  name = latex2exp::TeX("$k$"),
+  trans = "sqrt")
+p <- p + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"))
+
+```
 

### 2023-01-27 manuscript/manuscript.Rmd
Updated manuscript to include asymmetric generalised normal distributions.

@@
 
+## Asymmetric generalised normal distributions
+
+Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
+Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk, @Nadarajah2005-xe] is a suitable option to simulate realistic feature distributions.
+This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:
+
+```{=latex}
+\begin{equation}
+f_{\beta}(x) = \frac{\beta}{2\Gamma \left(1 / \beta \right)} e^{-\left| x \right|^\beta}
+\end{equation}
+```
+Here, $\Gamma$ is the gamma function, and $\beta$ is a strictly positive shape parameter.
+For $\beta = 1$, the probability density function describes a Laplace distribution.
+A normal distribution is found for $\beta=2$, and for large $\beta$, the distribution approaches a uniform distribution.
+We will refrain from introducing scale and location parameters here directly.
+
+Realistic feature distributions may be skewed.
+Gijbels et al.
+describe a recipe for introducing skewness into the otherwise symmetric generalised norm distribution [@Gijbels2019-te], leading to the following probability density function:
+
+```{=latex}
+\begin{equation}
+f_{\alpha}(x; \mu, \sigma, \beta) = \frac{2 \alpha \left(1 - \alpha\right)}{\sigma}
+\begin{cases}
+f_{\beta}\left( \left(1 - \alpha \right) \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x \leq \mu \\
+f_{\beta}\left( \alpha \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x > \mu
+\end{cases}
+\end{equation}
+```
+Here $\alpha \in (0,1)$ is a skewness parameter.
+$\alpha > 0.5$ creates a distribution with a negative skew, i.e.
+a left-skewed distribution.
+A right-skewed distribution is created for $\alpha < 0.5$.
+$\mu$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively.
+$f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution, which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.
+
+We require a quantile function (or an approximation thereof) to draw random values from an asymmetric generalised normal distribution using inverse transform sampling.
+Gijbels et al.
+derived the following quantile function $F_{\alpha}^{-1}(p)$:
+
+```{=latex}
+\begin{equation}
+F_{\alpha}^{-1}(p; \mu, \sigma, \beta) =
+\begin{cases}
+\mu + \frac{\sigma}{1 - \alpha} F_{\beta}^{-1} \left( \frac{p}{2 \alpha}\right) & \text{, } p \leq \alpha \\
+\mu + \frac{\sigma}{\alpha} F_{\beta}^{-1} \left( \frac{1 + p - 2 \alpha}{2 \left(1 - \alpha \right)} \right) & \text{, } p > \alpha
+\end{cases}
+\end{equation}
+```
+The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}(p)$ thus incorporates the quantile function $F_{\beta}^{-1}(p^{*})$ of the symmetric generalised normal distribution.
+$F_{\beta}^{-1}(p^{*})$ was derived by Griffin to be [@Griffin2018-bf]:
+
+```{=latex}
+\begin{equation}
+F_{\beta}^{-1}(p^{*}) = \sgn\left(p^{*} - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p^{*} - 0.5 \right|; 1 / \beta \right)
+\end{equation}
+```
+Here, $F_{\Gamma}^{-1}$ is the quantile function of the gamma distribution with shape $1 / \beta$, which can be numerically approximated.
+
 ## Goodness of fit
@@
 The first distribution is a normal distribution, as previously shown in Figure **INSERT REFERENCE**.
-We first randomly draw $10000$ values from a normal distribution: $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$.
-The second distribution is a right-skewed distribution $x_{\text{right}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) + 1\right)^{\frac{1}{10}}$.
-The third distribution is a left-skewed distribution $x_{\text{left}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) \right)^3$.
-Subsequently, all three distributions are standardised to have mean $0$ and standard deviation $1$.
+We first randomly draw $10000$ values from a normal distribution: $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution is a right-skewed normal distribution $x_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution is a left-skewed normal distribution $x_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
 We then compute transformation parameter $\lambda$ using the original definitions (Eq.
@@
 
-```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, include=FALSE, eval = FALSE}
 cap <- paste0(
@@
 # Normal distribution.
-x_normal <- stats::rnorm(10000L)
+x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
 
 # Right skewed data
-x_right_skewed <- (x_normal - min(x_normal) + 1)^0.1
-x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)
+x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
 
 # Left skewed data
-x_left_skewed <- (x_normal - min(x_normal))^3
-x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)
+x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
 
@@
 
-```{r, echo=FALSE, results="asis"}
+```{r, echo=FALSE, results="asis", include=FALSE, eval = FALSE}
 table_data <- data.table::dcast(
@@
 
-```{r}
+```{r, include=FALSE, eval = FALSE}
 # Generator for creating parameters and data for all experiments.

### 2023-02-01 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and even breaks down due to numerical issues for $\\mu > 10^3$. For values $\\mu \\leq 10^3$, $\\lambda$ values are similar between methods.")
+  "and even breaks down due to numerical issues for $\\mu > 5 10^4$.")
 
@@
 
-x <- stats::rnorm(10000L)
+x <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
 shift_range <- 10^seq(from=0, to=6, by=0.1)
@@
   name = latex2exp::TeX("$\\lambda$"),
-  limits = c(-3.6, 1.1))
+  limits = c(0.0, 35.0))
 p_bc <- p_bc + paletteer::scale_color_paletteer_d(
@@
   name = latex2exp::TeX("$\\lambda$"),
-  limits = c(-3.6, 1.1))
+  limits = c(0.0, 35.0))
 p_yj <- p_yj + paletteer::scale_color_paletteer_d(
@@
 # Simulation
+
+## Modified power transformations are invariant to location
 
@@
 
-```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, include=FALSE, eval = FALSE}
+```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, eval = FALSE}
 cap <- paste0(
-  "Power transformations modified to be shift-sensitive produce transformation parameters are invariant to location.")
+  "Power transformations modified to be shift-sensitive produce transformation parameters are invariant to location. ")
 
@@
           for(shift in c(FALSE, TRUE)){
-            for(robust in c(FALSE, TRUE)){
+            for(robust in c(FALSE)){
+            # for(robust in c(FALSE, TRUE)){
+              # Only 
               # Skip if robust, but not shifted.
@@
   x = data$version,
-  levels = c("original", "shift-sensitive", "robust shift-sensitive"))
+  levels = c("original", "shift-sensitive"))
 
@@
   plot_theme = plot_theme,
-  limits = c(-3.6, 1.1))
+  limits = c(0.0, 35.0))
 
@@
   plot_theme = plot_theme,
-  limits = c(1.0, 65))
+  limits = c(-9.0, 1.0))
 
@@
   plot_theme = plot_theme,
-  limits = c(-10.5, 0.5),
+  limits = c(0.0, 60.0),
   guide = TRUE)
@@
   plot_theme = plot_theme,
-  limits = c(-10.5, 0.5))
+  limits = c(0.0, 60.0))
 
@@
 
-The result is shown in Figure **INSERT REFERENCE**.
+The result is shown in Figure **INSERT REFERENCE**, and summarised in Table **INSERT REFERENCE**.
 For each distribution, transformation parameter $\lambda$ varies with $d$ when estimated using the original definitions.
@@
 ```
+
+## Weighting based on feature values offers best robustness
+
+Outliers may be presented in the data, and affect estimation of power transformation parameters.
+Such outliers should be down-weighted during the optimisation process.
+Such weights can be assigned using three different data representations: 1.
+the original feature values; 2.
+the feature values after power transformation; 3.
+the residual errors between transformed feature values and their expected values according to the normal distribution.
 

### 2023-02-02 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 First, normality is often desired in machine learning applications.
-Second, a large negative or positve $\lambda$-parameter may lead to numeric issues.
+Second, a large negative or positive $\lambda$-parameter may lead to numeric issues.
 Third, we currently do not have a test to automatically reject poor transformations.
@@
 
-```{r reduced-normality, echo=FALSE, fig.cap=cap}
+```{r reduced-normality, echo=FALSE, fig.cap=cap, eval=FALSE}
 cap <- paste0(
@@
 ```
-Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter:
-
+
+There are a few options based on which weights may be set. Here, we investigate three:
+
+- Set weights based on quantiles of the original feature $x$. Quantiles are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
+
+- Set weights based on z-score of the transformed feature $\phi^{\lambda, x_0} (x)$.
+After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$.
+Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
+
+- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+
+Based on either of the above, weights are set using a weight function. We use the following functions:
+
+- A step function, with $k_1$ as threshold parameter:
 ```{=latex}
@@
 \begin{cases}
-1 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} \leq Q_{\mathcal{N}} (0.995)\\
-0 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} > Q_{\mathcal{N}} (0.995)
+1 & \text{if } \left| x_i \right| \leq k_1\\
+0 & \text{if } \left| x_i \right| > k_1
 \end{cases}
@@
 ```
-Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
+
+- A triangle function (or generalised Huber weight), with $k_1$ and $k_2 \geq k_1$ as threshold parameters:
+```{=latex}
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \left| x_i \right| \leq k_1\\
+ \\
+0 & \text{if } \left| x_i \right| > k_1
+\end{cases}
+\end{equation}
+```
+
+<!-- Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter: -->
+
+<!-- ```{=latex} -->
+<!-- \begin{equation} -->
+<!-- w_i = -->
+<!-- \begin{cases} -->
+<!-- 1 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} \leq Q_{\mathcal{N}} (0.995)\\ -->
+<!-- 0 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} > Q_{\mathcal{N}} (0.995) -->
+<!-- \end{cases} -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
 In short, weights for all standardised transformed feature values that exceed the value corresponding to the $99.5^\text{th}$ percentile of the normal distribution ($Q_{\mathcal{N}}(0.995)$) are set to 0, and remaining weights are left at 1.
@@
 Using lower quantile values comes at the cost of essentially trimming more and more instances from the dataset.
-
-Another option is to weight instances based on their absolute residuals, $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - Q_{\mathcal{N}}(q_i) \right|$, where $q_i$ is equispaced quantile probability for instance $i$.
+-->
+
+<!-- Another option is to weight instances based on their absolute residuals, $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - Q_{\mathcal{N}}(q_i) \right|$, where $q_i$ is equispaced quantile probability for instance $i$.
 This weights instances by the distance from the expected normal distribution.
-Though we refrain from specifying the exact function here, one would down-weight instances with large absolute residuals more strongly than small absolute residuals.
+-->
+
+<!-- Though we refrain from specifying the exact function here, one would down-weight instances with large absolute residuals more strongly than small absolute residuals.
 While this approach is able to deal with outliers, there is no guarantee that large absolute residuals are strictly outliers.
 They may also be found in the central region, and down-weighting there might lead to suboptimal solutions.
-
-We propose a practical approach.
+-->
+
+<!-- We propose a practical approach.
 Let $\kappa$ be the fraction of central instances of feature $x$, i.e.
@@
 For example, if $\kappa = 0.80$, those instances with values between the $10^{\text{th}}$ and $90^{\text{th}}$ percentiles would be considered to be central.
-We then use Tukey's tapered cosine window to assign weights [@Tukey1967-eb]:
-
-```{=latex}
-\begin{equation}
-w_i = 
-\begin{cases}
-\frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 - \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right] & \text{, } 0 \leq i - 0.5 < (1 - \kappa) /2\\
-1 & \text{, } (1 - \kappa) /2 \leq i - 0.5 \leq (1 + \kappa) /2\\
-\frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 + \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right]& \text{, } (1 + \kappa) /2 < i - 0.5 \leq 1
-\end{cases}
-\end{equation}
-```
-Here, $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
+We then use Tukey's tapered cosine window to assign weights [@Tukey1967-eb]: -->
+
+<!-- ```{=latex} -->
+<!-- \begin{equation} -->
+<!-- w_i =  -->
+<!-- \begin{cases} -->
+<!-- \frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 - \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right] & \text{, } 0 \leq i - 0.5 < (1 - \kappa) /2\\ -->
+<!-- 1 & \text{, } (1 - \kappa) /2 \leq i - 0.5 \leq (1 + \kappa) /2\\ -->
+<!-- \frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 + \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right]& \text{, } (1 + \kappa) /2 < i - 0.5 \leq 1 -->
+<!-- \end{cases} -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- Here, $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$. -->
 
@@
 Such outliers should be down-weighted during the optimisation process.
-Such weights can be assigned using three different data representations: 1.
+Weights can be assigned using three different data representations: 1.
 the original feature values; 2.

### 2023-02-03 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-- Set weights based on quantiles of the original feature $x$. Quantiles are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
+- Set weights based on probabilities of the empirical distribution of the original feature $x$.
+Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
 
@@
 
-- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
-
-Based on either of the above, weights are set using a weight function. We use the following functions:
-
-- A step function, with $k_1$ as threshold parameter:
+- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+
+Based on either of the above, weights are set using a weighting function.
+We will investigate the following functions, where $x_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:
+
+- A step function, with $k_1 \geq 0$ as threshold parameter:
 ```{=latex}
@@
 
-- A triangle function (or generalised Huber weight), with $k_1$ and $k_2 \geq k_1$ as threshold parameters:
+- A triangle function (or generalised Huber weight), with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
 ```{=latex}
@@
 \begin{cases}
-1 & \text{if } \left| x_i \right| \leq k_1\\
- \\
-0 & \text{if } \left| x_i \right| > k_1
+1 & \text{if } \left| x_i \right| < k_1\\
+1 - \frac{\left| x_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
+0 & \text{if } \left| x_i \right| > k_2
 \end{cases}
 \end{equation}
+```
+
+- A tapered cosine function [@Tukey1967-eb], with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
+```{=latex}
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \left| x_i \right| < k_1\\
+0.5 + 0.5 \cos\left(\pi \frac{\left| x_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
+0 & \text{if } \left| x_i \right| > k_2
+\end{cases}
+\end{equation}
+```
+
+All weighting functions share the characteristic that for $x_i < k_1$, instances are fully weighted, i.e.
+when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top.
+The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq x_i \leq k_2$, and assign no weight to instances $x_i > k_2$.
+
+```{r weight-functions , echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, eval = FALSE}
+cap <- paste0(
+  "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
+  "The step function was parameterised with $k_1 = 0.60$. ",
+  "The triangle and tapered cosine functions were both parameterised with $k_1 = 0.30$ and $k_2 = 0.90$.")
+
+# Non-standard evaluation
+d <- distribution <- method <- mean_lambda <- sd_lambda <- NULL
+
+data_step <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
+data_step[, "weight":=power.transform:::.step_window(x, k1=0.60)]
+
+p_step <- ggplot2::ggplot(
+  data = data_step,
+  mapping = ggplot2::aes(
+    x = x,
+    y = weight))
+p_step <- p_step + plot_theme
+p_step <- p_step + ggplot2::geom_line()
+p_step <- p_step + ggplot2::ggtitle(label="step")
+p_step <- p_step + ggplot2::theme(
+  plot.title=ggplot2::element_text(
+    hjust=0.5,
+    size=ggplot2::rel(1.0)))
+
+data_triangle <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
+data_triangle[, "weight":=power.transform:::.triangular_window(x, k1=0.30, k2=0.90)]
+
+p_triangle <- ggplot2::ggplot(
+  data = data_triangle,
+  mapping = ggplot2::aes(
+    x = x,
+    y = weight))
+p_triangle <- p_triangle + plot_theme
+p_triangle <- p_triangle + ggplot2::geom_line()
+p_triangle <- p_triangle + ggplot2::ggtitle(label="triangle")
+p_triangle <- p_triangle + ggplot2::theme(
+  plot.title=ggplot2::element_text(
+    hjust=0.5,
+    size=ggplot2::rel(1.0)),
+  axis.text.y = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank())
+
+data_cosine <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
+data_cosine[, "weight":=power.transform:::.tapered_cosine_window(x, k1=0.30, k2=0.90)]
+
+p_cosine <- ggplot2::ggplot(
+  data = data_cosine,
+  mapping = ggplot2::aes(
+    x = x,
+    y = weight))
+p_cosine <- p_cosine + plot_theme
+p_cosine <- p_cosine + ggplot2::geom_line()
+p_cosine <- p_cosine + ggplot2::ggtitle(label="tapered cosine")
+p_cosine <- p_cosine + ggplot2::theme(
+  plot.title = ggplot2::element_text(
+    hjust = 0.5,
+    size = ggplot2::rel(1.0)),
+  axis.text.y = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank())
+
+p <- egg::ggarrange(plots=list(p_step, p_triangle, p_cosine), ncol=3, draw=FALSE)
+p
 ```
@@
 
+```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = FALSE}
+
+set.seed(95)
+
+.compute_lambda <- function(x){
+  # Create transformer object.
+  transformer <- suppressWarnings(
+    power.transform::find_transformation_parameters(
+      x = x,
+      method = "yeo_johnson",
+      robust = FALSE,
+      shift = TRUE))
+
+  return(transformer@lambda)
+}
+
+# Generate table of asymmetric generalised normal distribution parameters.
+n_distributions <- 100
+
+# Generate alpha, beta and n.
+n <- stats::runif(n=n_distributions, min=2, max=4)
+n <- round(10^n)
+
+alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
+beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
+
+# Generate corresponding distributions.
+x <- mapply(
+  power.transform::ragn,
+  n = n,
+  alpha = alpha,
+  beta = beta)
+
+# Compute lambda values without weighting for Yeo-Johnson.
+target_lambda <- sapply(
+  x,
+  .compute_lambda)
+
+# Add outliers, between 0.00 and 0.10.
+n_outliers <- 10
+outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
+
+# Find optimised k1 and k2 values for all combinations of weight sources and
+# weighting functions.
+x <- mapply(
+  function(x, n, alpha, beta, target_lambda, outlier_fraction){
+
+    parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda"=target_lambda)
+
+    x <- lapply(
+      outlier_fraction,
+      function(k, x, parameters){
+        # Set parameters.
+        parameters$k <- k
+
+        # Compute interquartile range.
+        interquartile_range <- stats::IQR(x)
+
+        # Compute upper and lower quartiles.
+        q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+        q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+        # Set data where the outliers will be copied into.
+        x_outlier <- x
+
+        if(k != 0.0){
+          n_draw <- ceiling(k * 0.5 * length(x))
+
+          # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+          # than Q3 + 1.5 IQR.
+          outlier <- c(
+            q_lower - 1.5 * interquartile_range - 3.0 * stats::runif(n_draw),
+            q_upper + 1.5 * interquartile_range + 3.0 * stats::runif(n_draw))
+
+          # Randomly insert outlier values.
+          x_outlier[sample(seq_along(x), size=2*n_draw, replace=FALSE)] <- outlier
+        }
+
+        return(list(
+          "x" = x_outlier,
+          "parameters" = parameters))
+      },
+      x = x,
+      parameters)
+
+    return(x)
+  },
+  x = x,
+  n = n,
+  alpha = alpha,
+  beta = beta,
+  target_lambda = target_lambda,
+  MoreArgs=list("outlier_fraction"=outlier_fraction),
+  SIMPLIFY=FALSE,
+  USE.NAMES=FALSE)
+
+x <- unlist(x, recursive = FALSE)
+
+# Optimise k1, k2 so that sum of absolute differences with target-lambda is
+# minimised for each combination of source and weighting method.
+
+```
+
 For robustness simulate:
@@
 
-```{r, include=FALSE, eval = FALSE}
+```{r, eval = FALSE}
 # Generator for creating parameters and data for all experiments.

### 2023-02-06 manuscript/manuscript.Rmd
Added code for the robustness experiment.

@@
 # Generate table of asymmetric generalised normal distribution parameters.
-n_distributions <- 100
+n_distributions <- 10
 
@@
 # minimised for each combination of source and weighting method.
+experiment_args <- coro::generator(
+  function(){
+
+    # Non-robust transformation.
+    fun_args <- list(
+      "method" = "yeo_johnson",
+      "robust" = FALSE,
+      "shift" = TRUE)
+
+    opt_args <- list()
+
+    yield(list(
+      "name" = "non-robust",
+      "fun_args" = fun_args,
+      "opt_args" = opt_args
+    ))
+
+    for(robustness_source in c("original", "transformed", "residual")){
+
+      def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
+
+      if(robustness_source == "original"){
+        def_opt_limits$k1[2] <- 1.0
+        def_opt_limits$k2[2] <- 1.0
+
+        def_opt_init <- list("k1" = 0.8, "k2" = 0.95)
+
+      } else if(robustness_source == "transformed"){
+        def_opt_init <- list("k1" = 1.5, "k2" = 3.0)
+
+      } else if(robustness_source == "residual"){
+        def_opt_init <- list("k1" = 0.15, "k2" = 1.0)
+      }
+
+      for(robustness_weighting_function in c("step", "triangle", "cosine")){
+
+        fun_args <- list(
+          "method" = "yeo_johnson",
+          "robust" = TRUE,
+          "shift" = TRUE,
+          "weight_method" = paste0(robustness_source, "_", robustness_weighting_function),
+          "backup_use_default" = FALSE)
+
+        opt_limits <- def_opt_limits
+        opt_init <- def_opt_init
+
+        # Step-weighting only has k1 as a parameter.
+        if(robustness_weighting_function == "step"){
+          opt_limits$k2 <- NULL
+          opt_init$k2 <- NULL
+        }
+
+        yield(list(
+          "name" = paste0(robustness_source, "-", robustness_weighting_function),
+          "fun_args" = fun_args,
+          "opt_args" = list("initial" = opt_init, "limits" = opt_limits)
+        ))
+      }
+    }
+  })
+
+experiments <- coro::collect(experiment_args())
+
+# Function for computing transformer parameters under optimisation constraints.
+compute_robust_lambda <- function(
+    x,
+    fun_args,
+    opt_args){
+
+  # Custom parser.
+  ..outlier_parser <- function(
+    x,
+    fun_args,
+    opt_args
+  ){
+    # Create transformer.
+    transformer <- suppressWarnings(do.call(
+      power.transform::find_transformation_parameters,
+      args=c(
+        list("x" = x$x),
+        fun_args,
+        opt_args)))
+
+    parameter_data <- c(
+      x$parameters,
+      opt_args,
+      list(
+        "weight_method" = fun_args$weight_method,
+        "lambda" = transformer@lambda,
+        "shift" = transformer@shift)
+    )
+
+    return(parameter_data)
+  }
+
+  parameter_data <- lapply(
+    x,
+    ..outlier_parser,
+    fun_args = fun_args,
+    opt_args = opt_args)
+
+  return(parameter_data)
+}
+
+# Inner optimisation function that sets the loss.
+..optimisation_function <- function(
+  opt_param,
+  x,
+  fun_args){
+
+  # Parse options.
+  opt_args <- list()
+  if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
+  if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
+
+  parameter_data <- compute_robust_lambda(
+    x = x,
+    fun_args = fun_args,
+    opt_args = opt_args)
+
+  lambda_error <- sapply(
+    parameter_data,
+    function(x) (abs(x$lambda - x$target_lambda)))
+
+  return(sum(lambda_error))
+}
+
+# Outer optimisation function
+.optimisation_function <- function(
+    experiment,
+    x){
+
+  # Run optimiser.
+  if(length(experiment$opt_args) > 0){
+
+    x0 <- unlist(experiment$opt_args$initial)
+
+    lower <- experiment$opt_args$limits$k1[1]
+    upper <- experiment$opt_args$limits$k1[2]
+
+    if(length(experiment$opt_args$limits) > 1){
+      lower <- c(lower, experiment$opt_args$limits$k2[1])
+      upper <- c(upper, experiment$opt_args$limits$k2[2])
+    }
+
+    h <- nloptr::sbplx(
+      x0 = x0,
+      fn = ..optimisation_function,
+      lower = lower,
+      upper = upper,
+      control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
+      x = x,
+      fun_args = experiment$fun_args)
+
+    if(length(h$par) == 1){
+      return(list("name"=experiment$name, "k1" = h$par, "value" = h$value))
+
+    } else {
+      return(list("name"=experiment$name, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
+    }
+
+  } else {
+    value <- ..optimisation_function(
+      opt_param=list(),
+      x = x,
+      fun_args=experiment$fun_args)
+
+    return(list("name"=experiment$name, "value" = value))
+  }
+}
+
+lapply(
+  experiments,
+  .optimisation_function,
+  x=x
+)
 

### 2023-02-06 manuscript/manuscript.Rmd
Added computation for robustness comparison.

@@
 require(coro)
+
+manuscript_dir <- file.path(
+  getwd())
 
@@
 # Generate table of asymmetric generalised normal distribution parameters.
-n_distributions <- 10
+n_distributions <- 100
 
@@
 
-experiments <- coro::collect(experiment_args())
-
 # Function for computing transformer parameters under optimisation constraints.
@@
     fun_args,
-    opt_args){
+    opt_args,
+    cl=NULL){
 
@@
 
+  if(is.null(cl)){
     parameter_data <- lapply(
@@
 
+  } else {
+    parameter_data <- parallel::parLapply(
+      cl = cl,
+      X = x,
+      fun = ..outlier_parser,
+      fun_args = fun_args,
+      opt_args = opt_args)
+  }
+
   return(parameter_data)
@@
   x,
-  fun_args){
+  fun_args,
+  cl = NULL){
 
@@
     fun_args = fun_args,
-    opt_args = opt_args)
+    opt_args = opt_args,
+    cl = cl)
 
@@
     experiment,
-    x){
+    x,
+    cl = NULL){
 
@@
       x = x,
-      fun_args = experiment$fun_args)
+      fun_args = experiment$fun_args,
+      cl = cl)
 
@@
       x = x,
-      fun_args=experiment$fun_args)
+      fun_args=experiment$fun_args,
+      cl = cl)
 
@@
 
-lapply(
-  experiments,
+# Start cluster
+cl <- parallel::makeCluster(16L)
+
+experiment_results <- lapply(
+  coro::collect(experiment_args()),
   .optimisation_function,
-  x=x
-)
+  x = x,
+  cl = cl)
+
+# Stop cluster.
+parallel::stopCluster(cl)
 

### 2023-02-07 manuscript/manuscript.Rmd
WIP

@@
    - \DeclareMathOperator*{\argmax}{argmax}
+   - \DeclareMathOperator*{\argmin}{argmin}
    - \DeclareMathOperator{\sgn}{sgn}
@@
 The two most commonly used transformations are that of @Box1964-mz and @Yeo2000-vw.
-The Box-Cox transformation of a feature $x$ under the transformation parameter $\lambda$ is defined as:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{BC}}^\lambda (x) = 
+The Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}$ under the transformation parameter $\lambda$ is defined as:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{BC}}^\lambda (x_i) = 
 \begin{cases}
-\left(x^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
-\log(x) & \text{if } \lambda = 0
+\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log(x_i) & \text{if } \lambda = 0
 \end{cases}
@@
 ```
-One limitation of th Box-Cox transformation is that it is only defined for $x > 0$.
-In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x \in \mathbb{R}$:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{YJ}}^\lambda (x) = 
+One limitation of th Box-Cox transformation is that it is only defined for $x_i > 0$.
+In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{YJ}}^\lambda (x_i) = 
 \begin{cases}
-\left( \left( 1 + x \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x \geq 0\\
-\log(1 + x) & \text{if } \lambda = 0 \text{ and } x \geq 0\\
--\left( \left( 1 - x\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x < 0\\
--\log(1 - x) & \text{if } \lambda = 2 \text{ and } x < 0
+\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
+\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
+-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
+-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
 \end{cases}
@@
 
-```{r reduced-normality, echo=FALSE, fig.cap=cap, eval=FALSE}
+```{r reduced-normality, echo=FALSE, fig.cap=cap, eval=FALSE, warning=FALSE}
 cap <- paste0(
@@
 Box-Cox and Yeo-Johnson transformations are modified by adding a shift parameter $x_0$.
-The modified Box-Cox transformation of a feature $x$ under transformation parameter $\lambda$ and shift parameter $x_0$ is then:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{BC}}^{\lambda, x_0} (x) = 
+The shift-sensitive Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$ and shift parameter $x_0$ is then:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{BC}}^{\lambda, x_0} (x_i) = 
 \begin{cases}
-\left( \left(x - x_0 \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
-\log(x - x_0) & \text{if } \lambda = 0
+\left( \left(x_i - x_0 \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log(x_i - x_0) & \text{if } \lambda = 0
 \end{cases}
@@
 ```
-Where $x - x_0 > 0$. Likewise, the modified Yeo-Johnson transformation of a feature $x$ under transformation parameter $\lambda$ and shift parameter $x_0$ is:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{YJ}}^{\lambda, x_0} (x) = 
+Where $x_i - x_0 > 0$. Likewise, the shift-sensitive Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$ and shift parameter $x_0$ is:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{YJ}}^{\lambda, x_0} (x_i) = 
 \begin{cases}
-\left( \left( 1 + x - x_0\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x - x_0 \geq 0\\
-\log(1 + x - x_0) & \text{if } \lambda = 0 \text{ and } x - x_0 \geq 0\\
--\left( \left( 1 - x - x_0\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x - x_0< 0\\
--\log(1 - x - x_0) & \text{if } \lambda = 2 \text{ and } x - x_0 < 0
+\left( \left( 1 + x_i - x_0\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
+\log(1 + x_i - x_0) & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+-\left( \left( 1 - x_i - x_0\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0< 0\\
+-\log(1 - x_i - x_0) & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
 \end{cases}
@@
 ```
-For both modified transformations, $\lambda$ and $x_0$ are obtained by maximising the normal log-likelihood function.
-For the modified Box-Cox transformation, and ignoring constant terms, this can be written as:
+For both shift-sensitive transformations, $\lambda$ and $x_0$ are obtained by maximising the normal log-likelihood function.
+For the shift-sensitive Box-Cox transformation, and ignoring constant terms, this can be written as:
 
@@
 ```
-where $\sigma^2$ is the variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (x)$. Similarly for the modified Yeo-Johnson transformation:
+where $\sigma^2$ is the variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$. Similarly for the shift-sensitive Yeo-Johnson transformation:
 
@@
 ```
-where $\sigma^2$ is the variance of Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (x)$.
+where $\sigma^2$ is the variance of Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$.
 
@@
 \begin{equation}
-\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
+\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
 \end{equation}
@@
 ## Goodness of fit
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
+\end{equation}
+```
 
@@
 
-## Weighting based on feature values offers best robustness
+## Weighting based on empirical distribution is the most robust
 
 Outliers may be presented in the data, and affect estimation of power transformation parameters.
-Such outliers should be down-weighted during the optimisation process.
-Weights can be assigned using three different data representations: 1.
-the original feature values; 2.
-the feature values after power transformation; 3.
-the residual errors between transformed feature values and their expected values according to the normal distribution.
+The log-likelihood function is weighted to assign less weight to outlier instances.
+Weighting is based on either probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
+A weight is then assigned to each instance using one of three weighting functions: step, triangle and tapered cosine.
+The weighting functions have one (step) or two (triangle, tapered cosine) parameters.
+Here we investigate how these parameters should be set to yield weights that are robust to outliers.
+
+Let us randomly draw $m_d=100$ asymmetric generalised normal distributions.
+Each distribution is parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+Location and scale parameters are set as $\mu = 0$ and $\sigma = 1$, respectively.
+$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(2, 4\right)$.
+That is between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
+
+Outlier values are then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
+This is repeated $m_{out} = 10$ times, with outlier fractions $f_{out,j}$ regularly spaced in $[0.00, 0.10]$.
+Thus up to 10 percent of the values may be replaced by outliers.
+Outlier values are set according to @Tukey1977-xm.
+Let $x^{*} \sim U\left(-2, 2\right)$.
+Then the corresponding outlier value is:
+
+```{=latex}
+\begin{equation}
+x_{out} =
+\begin{cases}
+Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
+Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
+\end{cases}
+\end{equation}
+```
+
+Here, $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
+Outlier values randomly replace values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$.
+
+To find the optimal values for the weighting function parameters $k_1$ and $k_2$ (if applicable), we minimise the absolute difference between the $\hat{\lambda}^r$ parameter obtained for transformation in the presence of outliers, and the $\hat{\lambda}_0$ parameter obtained using the non-robust modified transformation in absence of outliers:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{k}_1, \hat{k}_2 \right\} = \argmin_{k_1, k_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \hat{\lambda}^r \left(\mathbf{X}_{i, j}; k_1, k_2 \right) - \hat{\lambda}_{0} \left(\mathbf{X}_i \right) \right|
+\end{equation}
+```
+
+Optimisation is performed using the subplex algorithm implemented in the *nloptr* package for R [@Rowan1990-sn, @Ypma2022-tc].
 
@@
 n <- stats::runif(n=n_distributions, min=2, max=4)
-n <- round(10^n)
+n <- ceiling(10^n)
 
@@
         if(k != 0.0){
-          n_draw <- ceiling(k * 0.5 * length(x))
+          n_draw <- ceiling(k * length(x))
 
@@
           # than Q3 + 1.5 IQR.
-          outlier <- c(
-            q_lower - 1.5 * interquartile_range - 3.0 * stats::runif(n_draw),
-            q_upper + 1.5 * interquartile_range + 3.0 * stats::runif(n_draw))
+          x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+          outlier <- numeric(n_draw)
+          if(any(x_random < 0)){
+            outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+          }
+
+          if(any(x_random >= 0)){
+            outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+          }
 
           # Randomly insert outlier values.
-          x_outlier[sample(seq_along(x), size=2*n_draw, replace=FALSE)] <- outlier
+          x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
         }
@@
 # Start cluster
-cl <- parallel::makeCluster(16L)
+cl <- parallel::makeCluster(8L)
 
@@
 
+saveRDS(experiment_results, file = file.path(manuscript_dir, "robustness_comparison.RDS"))
 ```

### 2023-02-08 manuscript/manuscript.Rmd
WIP on manuscript.

@@
 knitr::opts_chunk$set(
-  echo = TRUE,
+  echo = FALSE,
+  eval = TRUE,
   fig.align= "center")
@@
 
-manuscript_dir <- file.path(
-  getwd())
+manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")
 
@@
 Power transformation does not guarantee that transformed variables are normally distributed.
-In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure **INSERT REFERENCE**.
+In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure **INSERT**.
 This is problematic for automated power transformations, for three reasons.
@@
 Statistical tests for normality exist, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-Given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for practical applications of automated power transformations.
-
-```{r reduced-normality, echo=FALSE, fig.cap=cap, eval=FALSE, warning=FALSE}
+However, given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for practical applications of automated power transformations.
+
+```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
 cap <- paste0(
@@
 Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
-Their presence may lead to poor transformations to normality, as shown in Figure **INSERT REFERENCE**.
+Their presence may lead to poor transformations to normality, as shown in Figure **INSERT**.
 As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e.
@@
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
-Because $\lambda$ is updated during this procedure, simultaneous estimation of $\lambda$ and $x_0$ for modified power transformations is not possible.
-
-In essence, obtaining transformation parameters that provide central normality requires identifying outliers in the data, and weighting such instances during the optimisation process.
-@Raymaekers2021-kq do this (during the final steps of the iterative procedure) by introducing a weighted maximum likelihood estimation.
-This weighted maximum log-likelihood estimator is based on equations **INSERT REFERENCE**.
-The weighted maximum log-likelihood estimator for the modified Box-Cox power transformation is:
+Because $\lambda$ is updated during their procedure, simultaneous estimation of $\lambda$ and $x_0$ for shift-sensitive power transformations is not possible.
+
+In essence, obtaining robust transformation parameters that provide central normality requires identifying outliers in the data and weighting such instances during the optimisation process.
+@Raymaekers2021-kq do this (during the final steps of their iterative procedure) through a weighted maximum likelihood estimation.
+This weighted maximum log-likelihood estimator is based on equations **INSERT**.
+The weighted maximum log-likelihood estimator for the shift-sensitive Box-Cox power transformation is:
 
@@
 ```
-where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (x)$:
+where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
 
@@
 ```
-Analogously, the weighted maximum log-likelihood estimator for the modified Yeo-Johnson power transformation is:
+Analogously, the weighted maximum log-likelihood estimator for the shift-sensitive Yeo-Johnson power transformation is:
 
@@
 ```
-where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (x)$:
+where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$:
 
@@
 
-- Set weights based on probabilities of the empirical distribution of the original feature $x$.
-Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.
-
-- Set weights based on z-score of the transformed feature $\phi^{\lambda, x_0} (x)$.
+- Set weights based on probabilities of the empirical distribution of the original feature $\mathbf{X}$.
+Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $\mathbf{X}$.
+
+- Set weights based on z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$.
 After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$.
-Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
-
-- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+
+- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
 
 Based on either of the above, weights are set using a weighting function.
-We will investigate the following functions, where $x_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:
+We will investigate the following functions, where $\dot{x}_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:
 
@@
 \begin{cases}
-1 & \text{if } \left| x_i \right| \leq k_1\\
-0 & \text{if } \left| x_i \right| > k_1
+1 & \text{if } \left| \dot{x}_i \right| \leq k_1\\
+0 & \text{if } \left| \dot{x}_i \right| > k_1
 \end{cases}
@@
 \begin{cases}
-1 & \text{if } \left| x_i \right| < k_1\\
-1 - \frac{\left| x_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
-0 & \text{if } \left| x_i \right| > k_2
+1 & \text{if } \left| \dot{x}_i \right| < k_1\\
+1 - \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > k_2
 \end{cases}
@@
 \begin{cases}
-1 & \text{if } \left| x_i \right| < k_1\\
-0.5 + 0.5 \cos\left(\pi \frac{\left| x_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
-0 & \text{if } \left| x_i \right| > k_2
+1 & \text{if } \left| \dot{x}_i \right| < k_1\\
+0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > k_2
 \end{cases}
@@
 
-All weighting functions share the characteristic that for $x_i < k_1$, instances are fully weighted, i.e.
+All weighting functions share the characteristic that for $\dot{x}_i < k_1$, instances are fully weighted, i.e.
 when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top.
-The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq x_i \leq k_2$, and assign no weight to instances $x_i > k_2$.
-
-```{r weight-functions , echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, eval = FALSE}
+The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \dot{x}_i \leq k_2$, and assign no weight to instances $x_i > k_2$.
+
+```{r weight-functions, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
@@
 # Non-standard evaluation
-d <- distribution <- method <- mean_lambda <- sd_lambda <- NULL
+d <- distribution <- method <- mean_lambda <- sd_lambda <- weight <- NULL
 
@@
 
-<!-- Weights $w_i$ can be set based on the original feature $x$, as well as the transformed feature $\phi^{\lambda, x_0} (x)$. @Raymaekers2021-kq use the latter: -->
-
-<!-- ```{=latex} -->
-<!-- \begin{equation} -->
-<!-- w_i = -->
-<!-- \begin{cases} -->
-<!-- 1 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} \leq Q_{\mathcal{N}} (0.995)\\ -->
-<!-- 0 & \text{if } \frac{\left| \phi^{\lambda}(x) - \mu_M \right|}{\sigma_M} > Q_{\mathcal{N}} (0.995) -->
-<!-- \end{cases} -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (x)$ [@Huber1981-su].
-In short, weights for all standardised transformed feature values that exceed the value corresponding to the $99.5^\text{th}$ percentile of the normal distribution ($Q_{\mathcal{N}}(0.995)$) are set to 0, and remaining weights are left at 1.
-This works well if outliers are relatively rare in the transformed feature values, which we cannot *a priori* assume in our use case.
-Using lower quantile values comes at the cost of essentially trimming more and more instances from the dataset.
--->
-
-<!-- Another option is to weight instances based on their absolute residuals, $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - Q_{\mathcal{N}}(q_i) \right|$, where $q_i$ is equispaced quantile probability for instance $i$.
-This weights instances by the distance from the expected normal distribution.
--->
-
-<!-- Though we refrain from specifying the exact function here, one would down-weight instances with large absolute residuals more strongly than small absolute residuals.
-While this approach is able to deal with outliers, there is no guarantee that large absolute residuals are strictly outliers.
-They may also be found in the central region, and down-weighting there might lead to suboptimal solutions.
--->
-
-<!-- We propose a practical approach.
-Let $\kappa$ be the fraction of central instances of feature $x$, i.e.
-those located between $q_i=(1-\kappa)/2$ and $q_i=(1 + \kappa)/2$.
-For example, if $\kappa = 0.80$, those instances with values between the $10^{\text{th}}$ and $90^{\text{th}}$ percentiles would be considered to be central.
-We then use Tukey's tapered cosine window to assign weights [@Tukey1967-eb]: -->
-
-<!-- ```{=latex} -->
-<!-- \begin{equation} -->
-<!-- w_i =  -->
-<!-- \begin{cases} -->
-<!-- \frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 - \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right] & \text{, } 0 \leq i - 0.5 < (1 - \kappa) /2\\ -->
-<!-- 1 & \text{, } (1 - \kappa) /2 \leq i - 0.5 \leq (1 + \kappa) /2\\ -->
-<!-- \frac{1}{2} \left[1 + \cos\left(2 \pi \frac{i - \frac{1}{2}\left[ n \left(1 + \kappa\right) + 1\right]} {n \left(1 - \kappa \right)} \right) \right]& \text{, } (1 + \kappa) /2 < i - 0.5 \leq 1 -->
-<!-- \end{cases} -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- Here, $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$. -->
-
 ## Asymmetric generalised normal distributions
@@
 Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
-Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk, @Nadarajah2005-xe] is a suitable option to simulate realistic feature distributions.
+Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option to simulate realistic feature distributions.
 This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:
@@
 A right-skewed distribution is created for $\alpha < 0.5$.
-$\mu$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively.
+$\mu \in \mathcal{R}$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively.
 $f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution, which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.
@@
 
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
-\end{equation}
-```
+To be added.
 
@@
 
-## Modified power transformations are invariant to location
+## Shift-sensitive transformations are location-invariant
 
 To assess whether the modified power transformations lead to $\lambda$ values that are invariant to shifts, we simulate three different distributions.
-The first distribution is a normal distribution, as previously shown in Figure **INSERT REFERENCE**.
-We first randomly draw $10000$ values from a normal distribution: $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
-The second distribution is a right-skewed normal distribution $x_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
-The third distribution is a left-skewed normal distribution $x_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then compute transformation parameter $\lambda$ using the original definitions (Eq.
-**INSERT REFERENCE** and **INSERT REFERENCE**) and the modified definitions (Eq.
-**INSERT REFERENCE** and **INSERT REFERENCE**) for each distribution, after adding a positive value $d$ to each distribution.
+The first distribution is a normal distribution, as previously shown in Figure **INSERT**.
+We first randomly draw $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution is a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+We then compute transformation parameter $\hat{\lambda}$ using the original definitions (Eq.
+**INSERT** and **INSERT**) and the shift-sensitive definitions (Eq.
+**INSERT** and **INSERT**) for each distribution, after adding a positive value $d$ to each distribution.
 Here, $d \in [1, 10^6]$.
 
-```{r shifted-distributions, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, eval = FALSE}
+```{r shifted-distributions, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
-  "Power transformations modified to be shift-sensitive produce transformation parameters are invariant to location. ")
+  "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
+  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively,  which then underwent a shift $d$. ",
+  "Estimates of transformation parameter $\\lambda$ with the original power transformations (circle) show strong dependency on the overall location of the distribution, ",
+  "whereas estimates with the shift-sensitive power transformations (triangle) are constant.")
 
@@
 
+# Set seed.
+set.seed(19L)
+
+# Normal distribution.
+x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
+
+# Right skewed data
+x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
+
+# Left skewed data
+x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
+
+if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
+
   # Generator for creating parameters and data for all experiments.
@@
 
-# Set seed.
-set.seed(19L)
-
-# Normal distribution.
-x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
-
-# Right skewed data
-x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
-
-# Left skewed data
-x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
-
   # Generate all experiments.
@@
   data <- data.table::rbindlist(data)
+
+  # Save to file.
+  saveRDS(data, file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
+
+} else {
+  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
+}
 
@@
 p_dens_normal <- .create_density_plot(
-  x=x_normal,
-  plot_theme=plot_theme)
+  x = x_normal,
+  plot_theme = plot_theme)
 p_dens_normal <- p_dens_normal + ggplot2::ggtitle(label="normal distribution")
@@
 p_bc_right <- .create_lambda_shift_plot(
-  data =data[distribution == "right-skewed" & method == "Box-Cox"],
+  data = data[distribution == "right-skewed" & method == "Box-Cox"],
   plot_theme = plot_theme,
@@
 
-The result is shown in Figure **INSERT REFERENCE**, and summarised in Table **INSERT REFERENCE**.
-For each distribution, transformation parameter $\lambda$ varies with $d$ when estimated using the original definitions.
-In contrast, estimations of $\lambda$ modified power transformations were invariant to $d$.
-
-```{r, echo=FALSE, results="asis", include=FALSE, eval = FALSE}
+The result is shown in Figure **INSERT**, and summarised in Table **INSERT**.
+For each distribution, transformation parameter $\hat{\lambda}$ varies with $d$ when estimated using the original definitions.
+In contrast, estimation of $\lambda$ using shift-sensitive power transformations was invariant to $d$.
+
+```{r, results="asis"}
 table_data <- data.table::dcast(
@@
 
+data.table::setnames(
+  table_data,
+  old = c("original Box-Cox", "original Yeo-Johnson", "shift-sensitive Box-Cox", "shift-sensitive Yeo-Johnson"),
+  new = c("orig. Box-Cox", "orig. Yeo-Johnson", "shift-sens. Box-Cox", "shift-sens. Yeo-Johnson"))
+
 knitr::kable(
   table_data,
-  caption = "Estimates and deviation of transformation parameter $\\lambda$ under location shifts",
+  align = c("l", "c", "c", "c", "c"),
+  caption = paste0(
+    "Estimates and variability of transformation parameter $\\lambda$ under location shifts. ",
+    "Whereas the original (orig.) power transformations show considerable variability, ",
+    "whereas for shift-sensitive (shift-sens.) power transformations variability in the value of $\\lambda$ is negligible."),
   label = "shifted-distributions-table",
@@
 
-## Weighting based on empirical distribution is the most robust
+## Empirical distribution yields consistent robust transformations
 
@@
 The weighting functions have one (step) or two (triangle, tapered cosine) parameters.
-Here we investigate how these parameters should be set to yield weights that are robust to outliers.
+Here we investigate how these parameters should be set to find transformation parameter $\hat{\lambda}^r$ that are robust to outliers.
 
@@
 Outlier values are then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
-This is repeated $m_{out} = 10$ times, with outlier fractions $f_{out,j}$ regularly spaced in $[0.00, 0.10]$.
+This is repeated $m_{out} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
 Thus up to 10 percent of the values may be replaced by outliers.
@@
 ```
-
-Optimisation is performed using the subplex algorithm implemented in the *nloptr* package for R [@Rowan1990-sn, @Ypma2022-tc].
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = FALSE}
+The double sum in right-hand side of the above equation can be considered as the loss that should be minimised.
+Minimisation is performed using the subplex algorithm implemented in the *nloptr* package for R [@Rowan1990-sn; @Ypma2022-tc].
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
 
@@
 
-For robustness simulate:
-- A shifted normal distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
-- A shifted left-skewed distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
-- A shifted right-skewed distribution without, 0.1%, 0.2%, 0.5%, 1.0%, 2.0%, 5.0%, 10%, and 20% outliers mixed in.
-- To show: lambda for modified, and robust modified Yeo-Johnson distribution for: Raymaekers trimmed, Tukey Window
-
-```{r, eval = FALSE}
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Optimal weighting parameters and corresponding loss.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
+The {optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
+Lower loss indicates better robustness against outliers.}
+\begin{tabular}{l r r r r r r r}
+
+\toprule
+method & \multicolumn{3}{c}{$k_1$} & \multicolumn{3}{c}{$k_2$} & loss \\
+ & initial & limits & optimal & initial & limits & optimal & \\
+
+\midrule
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---   & 225 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.92 & ---  & ---       & ---   & 125 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.93  & 126 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.89 & 0.95 & $(0, 1]$  & 0.93  & 124 \\
+$z$ (step)               & 1.50 & $(0, 10]$ & 1.03 & ---  & ---       & ---   & 331 \\
+$z$ (triangle)           & 1.50 & $(0, 10]$ & 2.79 & 3.00 & $(0, 10]$ & 4.12  & 768 \\
+$z$ (tapered cosine)     & 1.50 & $(0, 10]$ & 2.69 & 3.00 & $(0, 10]$ & 4.21  & 772 \\
+$r$ (step)               & 0.15 & $(0, 10]$ & 0.05 & ---  & ---       & ---   & 1957 \\
+$r$ (triangle)           & 0.15 & $(0, 10]$ & 9.96 & 1.00 & $(0, 10]$ & 10.00 & 483 \\
+$r$ (tapered cosine)     & 0.15 & $(0, 10]$ & 9.97 & 1.00 & $(0, 10]$ & 10.00 & 484 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
+Table **INSERT** shows weighting parameters and corresponding loss, after optimisation.
+Loss without any mitigation of outliers was $225$.
+Only weighting based on empirical probabilities ($p^(*)$) achieved better loss under optimised weighting parameters.
+For these data, all three window functions achieved a similar loss.
+Weighting based on the z-score of the transformed feature, or on their residual error, performed worse than non-robust shift-sensitive power transformation.
+Thus weighting of outliers should not be based on either of these two options.
+The likely reason for their poor performance is that they allow for a direct interaction with estimation of the $\lambda$ parameter.
+Because they both use transformed feature values, weights actually vary during the optimisation process for maximising the weighted log-likelihood function.
+This complicates the parameter search space, and leads to different parameter estimates.
+Weighting based on empirical probabilities does not suffer from this issue, because all weights are fixed during optimising the weighted log-likelihood function.
+For weighting based on the z-score of the transformed feature, or on their residual error, we can see two extreme strategies emerging.
+Firstly, the strategy to down-weight most instances ($r$ (step), and to a lesser degree $z$ (step)) and thereby force normality on the most central parts of the distribution.
+Second is an attempt to assign full-weight to almost all instances, with large values for $k_1$ and $k_2$, which leads to behaviour similar to non-robust power transformation.
+
+```{r outlier-values, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "CHECK. ")
+
+# Prevent warnings due to non-standard evaluation.
+weight_method <- distribution <- k <- NULL
+
+.create_outlier_fraction_plot <- function(
+    data,
+    plot_theme,
+    limits,
+    lambda_target,
+    guide=FALSE){
+
+  # Prevent warnings due to non-standard evaluation.
+  weight_method <- k <- mean_lambda <- min_lambda <- max_lambda <- NULL
+
+  # Annotation settings.
+  annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
+
+  p <- ggplot2::ggplot(
+    data = data,
+    mapping = ggplot2::aes(
+      x = k,
+      y = mean_lambda,
+      ymin = min_lambda,
+      ymax = max_lambda))
+  p <- p + plot_theme
+
+  p <- p + ggplot2::geom_ribbon(mapping=ggplot2::aes(fill=weight_method), alpha=0.10)
+  p <- p + ggplot2::geom_line(mapping=ggplot2::aes(colour=weight_method))
+  p <- p + ggplot2::scale_x_continuous(
+    name = latex2exp::TeX("outlier fraction"),
+    trans = "sqrt")
+  p <- p + ggplot2::scale_y_continuous(
+    name = latex2exp::TeX("$\\lambda$"),
+    limits = limits)
+
+  p <- p + ggplot2::geom_hline(
+    yintercept = lambda_target, 
+    linetype="longdash", 
+    colour="grey40")
+
+  p <- p + ggplot2::annotate(
+    geom = "text",
+    x = 0.1,
+    y = lambda_target,
+    label = "target",
+    colour=annotation_settings$colour,
+    family=annotation_settings$family,
+    fontface=annotation_settings$face,
+    size=annotation_settings$geom_text_size,
+    vjust=-1.0,
+    hjust=1.0)
+
+  if(guide){
+    p <- p + paletteer::scale_color_paletteer_d(
+      palette="ggthemes::Tableau_10",
+      drop=FALSE,
+      guide=ggplot2::guide_legend("weighting method"))
+
+    p <- p + paletteer::scale_fill_paletteer_d(
+      palette="ggthemes::Tableau_10",
+      drop=FALSE,
+      guide=ggplot2::guide_legend("weighting method"))
+
+  } else {
+    p <- p + paletteer::scale_color_paletteer_d(
+      palette="ggthemes::Tableau_10",
+      drop=FALSE,
+      guide="none")
+
+    p <- p + paletteer::scale_fill_paletteer_d(
+      palette="ggthemes::Tableau_10",
+      drop=FALSE,
+      guide="none")
+  }
+
+  p <- p + ggplot2::scale_alpha(guide="none")
+
+  # Update title.
+  p <- p + ggplot2::theme(
+    plot.title=ggplot2::element_text(
+      hjust=0.5,
+      size=ggplot2::rel(1.0)))
+
+  return(p)
+}
+# Only process data if the file has not been created.
+if(!file.exists(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))){
+
   # Generator for creating parameters and data for all experiments.
@@
       n <- 20
-    k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 5)  
-    # k_range <- c(0.0, 0.001, 0.002, 0.005, 0.010, 0.020, 0.050, 0.100, 0.200)
+      k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 10)
 
@@
         }
-      for(ii in seq_len(20)){
+        for(ii in seq_len(100)){
           for(k in k_range){
@@
             if(k != 0.0){
-            n_draw <- ceiling(k * 0.5 * length(x))
+              n_draw <- ceiling(k * length(x))
 
@@
               # than Q3 + 1.5 IQR.
-            outlier <- c(
-              q_lower - 1.5 * interquartile_range - 3.0 * stats::runif(n_draw),
-              q_upper + 1.5 * interquartile_range + 3.0 * stats::runif(n_draw))
+              x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+              outlier <- numeric(n_draw)
+              if(any(x_random < 0)){
+                outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+              }
+
+              if(any(x_random >= 0)){
+                outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+              }
 
               # Randomly insert outlier values.
-            x_outlier[sample(seq_along(x), size=2*n_draw, replace=FALSE)] <- outlier
+              x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
             }
@@
             for(method in c("yeo_johnson")){
-            for(weight_method in c("none", "tukey_window", "tapered_step_window")){
-            # for(weight_method in c("none", "tukey_window", "step_window", "tapered_step_window", "trim_transformation", "trim_residual", "tukey_biweight_transformation", "tukey_biweight_residual", "huber_weight_transformed", "huber_weight_residual")){
-
-              weight_method_name <- weight_method
-
-              for(tukey_window_k in c(0.80, 0.60, 0.40, 0.20, 0.00)){
-                if(weight_method == "none" && tukey_window_k != 0.00) next
+              for(weight_method in c("none", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
+
+                k1 <- k2 <- NULL
+
+                # Optimal parameter values from the previous experiment.
+                if(weight_method == "original_cosine"){
+                  k1 <- 0.89
+                  k2 <- 0.93
+
+                } else if(weight_method == "transformed_step"){
+                  k1 <- 1.03
+
+                } else if(weight_method == "transformed_cosine"){
+                  k1 <- 2.69
+                  k2 <- 4.21
+
+                } else if(weight_method == "residual_step"){
+                  k1 <- 0.05
+
+                } else if(weight_method == "residual_cosine"){
+                  k1 <- 9.97
+                  k2 <- 10.0
+                }
 
@@
                   "weight_method" = weight_method,
-                  "weight_method_name" = paste0(weight_method_name, "_k_", tukey_window_k),
-                  "tukey_window_k" = tukey_window_k
-                ))
-              }
-
-              # if(weight_method == "tukey_window"){
-              #   for(tukey_window_k in c(0.9, 0.8, 0.7, 0.6, 0.5)){
-              #     yield(list(
-              #       "x" = x_outlier,
-              #       "k" = k,
-              #       "distribution" = distribution,
-              #       "method" = method,
-              #       "robust" = weight_method != "none",
-              #       "weight_method" = weight_method,
-              #       "weight_method_name" = paste0(weight_method_name, "_k_", tukey_window_k),
-              #       "tukey_window_k" = tukey_window_k
-              #     ))
-              #   }
-              # } else {
-              #   yield(list(
-              #     "x" = x_outlier,
-              #     "k" = k,
-              #     "distribution" = distribution,
-              #     "method" = method,
-              #     "robust" = weight_method != "none",
-              #     "weight_method" = weight_method,
-              #     "weight_method_name" = weight_method_name
-              #   ))
-              # }
+                  "k1" = k1,
+                  "k2" = k2))
               }
@@
         weight_method = parameter_set$weight_method,
-      k = parameter_set$tukey_window_k,
-      lambda = NULL))
+        k1 = parameter_set$k1,
+        k2 = parameter_set$k2))
 
@@
         "weight_method" = parameter_set$weight_method,
-      "weight_method_name" = parameter_set$weight_method_name,
         "k" = parameter_set$k,
         "lambda" = transformer@lambda,
-      "kappa" = parameter_set$tukey_window_k,
-      "ii" = parameter_set$ii
-    ))
-}
+        "ii" = parameter_set$ii))
+  }
+
+  #### Computations --------------------------------------------------------------
 
@@
   # Normal distribution.
-x_normal <- stats::rnorm(10000L)
+  x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
 
   # Right skewed data
-x_right_skewed <- (x_normal - min(x_normal) + 1)^0.1
-x_right_skewed <- (x_right_skewed - mean(x_right_skewed)) / sd(x_right_skewed)
+  x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
 
   # Left skewed data
-x_left_skewed <- (x_normal - min(x_normal))^3
-x_left_skewed <- (x_left_skewed - mean(x_left_skewed)) / sd(x_left_skewed)
+  x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
 
@@
   ))
-
   # Start cluster
-cl <- parallel::makeCluster(16L)
+  cl <- parallel::makeCluster(18L)
 
@@
 
+  saveRDS(data, file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
+
+} else {
+  data <- readRDS(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
+}
+
 # Update distribution, method and version to factors.
@@
   x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-data$kappa <- factor(data$kappa)
-
-# Find target lambda values in the absence of any outliers.
-target_data <- data[weight_method=="none" & k == 0.0, mget(c("distribution", "ii", "lambda", "method"))]
-data.table::setnames(target_data, old="lambda", new="target_lambda")
-
-# Merge target lambda values into the main data.
-data <- data[target_data, on=.NATURAL]
-
-# Compute difference with target lambda.
-data[, "diff_lambda":=lambda - target_lambda]
-
-summary_data[, list(deviation=sum(abs(diff_lambda))), by=c("distribution", "weight_method_name")][order(distribution, deviation)]
-
-# Compute average.
-new_data <- data[, list(lambda = mean(lambda)), by=c("k", "kappa", "weight_method", "distribution")]
-
-p <- ggplot2::ggplot(
-  data = new_data[distribution=="left-skewed"],
-  mapping = ggplot2::aes(
-    x = k,
-    y = lambda,
-    colour = kappa,
-    linetype = weight_method))
-p <- p + plot_theme
-# p <- p + ggplot2::geom_point()
-p <- p + ggplot2::geom_line()
-p <- p + ggplot2::scale_x_continuous(
-  name = latex2exp::TeX("$k$"),
-  trans = "sqrt")
-p <- p + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"))
-
+  levels = c("yeo_johnson"),
+  labels = c("Yeo-Johnson"))
+data$weight_method <- factor(
+  x = data$weight_method,
+  levels = c("none", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
+  labels = c("non-robust", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
+
+data <- data[, list(
+  "mean_lambda"=mean(lambda),
+  "min_lambda"=min(lambda),
+  "max_lambda"=max(lambda)),
+  by=c("distribution", "method", "weight_method", "k")]
+
+top_group <- c("non-robust", "emp. prob. (tap. cos.)", "z-score (step)")
+bottom_group <- c("z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
+
+#### Normal distribution -------------------------------------------------------
+p_normal_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "normal" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits= c(0.0, 2.0),
+  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_normal_t <- p_normal_t + ggplot2::ggtitle(label="normal distribution")
+p_normal_t <- p_normal_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank())
+
+p_normal_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "normal" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits= c(-2.5,  4.0),
+  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_normal_b <- p_normal_b + ggplot2::theme(
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+#### Right skewed distribution -------------------------------------------------
+p_right_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "right-skewed" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits= c(0.0, 2.0),
+  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_right_t <- p_right_t + ggplot2::ggtitle(label="right-skewed distribution")
+p_right_t <- p_right_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank())
+
+p_right_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "right-skewed" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits= c(-2.5,  4.0),
+  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_right_b <- p_right_b + ggplot2::theme(
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+#### Left skewed distribution --------------------------------------------------
+p_left_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "left-skewed" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits= c(0.0, 2.0),
+  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda,
+  guide = TRUE)
+p_left_t <- p_left_t + ggplot2::ggtitle(label="left-skewed distribution")
+p_left_t <- p_left_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank())
+
+p_left_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "left-skewed" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits= c(-2.5,  4.0),
+  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_left_b <- p_left_b + ggplot2::theme(
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+#### Aggregate -----------------------------------------------------------------
+
+p <- egg::ggarrange(
+  plots=list(
+    p_normal_t, p_right_t, p_left_t,
+    p_normal_b, p_right_b, p_left_b),
+  ncol=3,
+  heights=c(1.0, 1.0),
+  byrow=TRUE,
+  draw=FALSE)
+p
 ```

### 2023-02-08 manuscript/manuscript.Rmd
WIP

@@
             for(method in c("yeo_johnson")){
-              for(weight_method in c("none", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
+              for(weight_method in c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
 
@@
                 # Optimal parameter values from the previous experiment.
-                if(weight_method == "original_cosine"){
+                if(weight_method == "original_step"){
+                  k1 <- 0.92
+
+                } else if(weight_method == "original_cosine"){
                   k1 <- 0.89
@@
   x = data$weight_method,
-  levels = c("none", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
-  labels = c("non-robust", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
+  levels = c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
+  labels = c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
 
@@
 
-top_group <- c("non-robust", "emp. prob. (tap. cos.)", "z-score (step)")
-bottom_group <- c("z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
+top_group <- c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)")
+bottom_group <- c("z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
+
+top_limits <- c(0.2, 1.8)
+bottom_limits <- c(-2.5, 4.0)
 
@@
   plot_theme = plot_theme,
-  limits= c(0.0, 2.0),
+  limits = top_limits,
   lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
@@
   plot_theme = plot_theme,
-  limits= c(-2.5,  4.0),
+  limits = bottom_limits,
   lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
@@
   plot_theme = plot_theme,
-  limits= c(0.0, 2.0),
+  limits= top_limits,
   lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
@@
   plot_theme = plot_theme,
-  limits= c(-2.5,  4.0),
+  limits = bottom_limits,
   lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
@@
   plot_theme = plot_theme,
-  limits= c(0.0, 2.0),
+  limits = top_limits,
   lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda,
@@
   plot_theme = plot_theme,
-  limits= c(-2.5,  4.0),
+  limits = bottom_limits,
   lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)

### 2023-02-10 manuscript/manuscript.Rmd
Update to manuscript.

@@
     latex_engine: lualatex
-bibliography: "refs.bib"
+bibliography: refs.bib
 header-includes:
@@
 
-.compute_lambda <- function(x){
+.compute_lambda <- function(x, method){
   # Create transformer object.
@@
       x = x,
-      method = "yeo_johnson",
+      method = method,
       robust = FALSE,
@@
 
-# Compute lambda values without weighting for Yeo-Johnson.
-target_lambda <- sapply(
+# Compute lambda values without weighting for Box-Cox.
+target_lambda_bc <- sapply(
   x,
-  .compute_lambda)
+  .compute_lambda,
+  method = "box_cox")
+
+# Compute lambda values without weighting for Yeo-Johnson
+target_lambda_yj <- sapply(
+  x,
+  .compute_lambda,
+  method = "yeo_johnson")
 
@@
 x <- mapply(
-  function(x, n, alpha, beta, target_lambda, outlier_fraction){
-
-    parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda"=target_lambda)
+  function(x, n, alpha, beta, target_lambda_bc, target_lambda_yj, outlier_fraction){
+
+    parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda_bc"=target_lambda_bc, "target_lambda_yj"=target_lambda_yj)
 
@@
   beta = beta,
-  target_lambda = target_lambda,
+  target_lambda_bc = target_lambda_bc,
+  target_lambda_yj = target_lambda_yj,
   MoreArgs=list("outlier_fraction"=outlier_fraction),
@@
 
+    for(method in c("box_cox", "yeo_johnson")){
       # Non-robust transformation.
       fun_args <- list(
-      "method" = "yeo_johnson",
+        "method" = method,
         "robust" = FALSE,
@@
         "name" = "non-robust",
+        "method" = method,
         "fun_args" = fun_args,
@@
       ))
+    }
 
@@
 
-        def_opt_init <- list("k1" = 0.8, "k2" = 0.95)
+        def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
 
       } else if(robustness_source == "transformed"){
-        def_opt_init <- list("k1" = 1.5, "k2" = 3.0)
+        def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
 
       } else if(robustness_source == "residual"){
-        def_opt_init <- list("k1" = 0.15, "k2" = 1.0)
-      }
-
+        def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
+      }
+
+      for(method in c("box_cox", "yeo_johnson")){
         for(robustness_weighting_function in c("step", "triangle", "cosine")){
@@
           fun_args <- list(
-          "method" = "yeo_johnson",
+            "method" = method,
             "robust" = TRUE,
@@
             "name" = paste0(robustness_source, "-", robustness_weighting_function),
+            "method" = method,
             "fun_args" = fun_args,
@@
           ))
+        }
       }
@@
     opt_args,
-    cl=NULL){
+    cl_internal=NULL){
 
@@
       list(
+        "method" = fun_args$method,
         "weight_method" = fun_args$weight_method,
@@
 
-  if(is.null(cl)){
+  if(is.null(cl_internal)){
     parameter_data <- lapply(
@@
     parameter_data <- parallel::parLapply(
-      cl = cl,
+      cl = cl_internal,
       X = x,
@@
   fun_args,
-  cl = NULL){
+  cl_internal = NULL){
 
@@
     opt_args = opt_args,
-    cl = cl)
-
+    cl_internal = cl_internal)
+
+  if(fun_args$method == "box_cox"){
     lambda_error <- sapply(
       parameter_data,
-    function(x) (abs(x$lambda - x$target_lambda)))
+      function(x) (abs(x$lambda - x$target_lambda_bc)))
+
+  } else {
+    # Yeo-Johnson.
+    lambda_error <- sapply(
+      parameter_data,
+      function(x) (abs(x$lambda - x$target_lambda_yj)))
+  }
 
@@
     experiment,
-    x,
-    cl = NULL){
+    data,
+    cl_internal = NULL){
+
+  set.seed(9)
 
@@
       control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
-      x = x,
+      x = data,
       fun_args = experiment$fun_args,
-      cl = cl)
+      cl_internal = cl_internal)
 
     if(length(h$par) == 1){
-      return(list("name"=experiment$name, "k1" = h$par, "value" = h$value))
+      return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par, "value" = h$value))
 
     } else {
-      return(list("name"=experiment$name, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
+      return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
     }
@@
       opt_param=list(),
-      x = x,
+      x = data,
       fun_args=experiment$fun_args,
-      cl = cl)
-
-    return(list("name"=experiment$name, "value" = value))
+      cl = cl_internal)
+
+    return(list("name"=experiment$name, "method"=experiment$method, "value" = value))
   }
@@
 # Start cluster
-cl <- parallel::makeCluster(8L)
-
-experiment_results <- lapply(
-  coro::collect(experiment_args()),
-  .optimisation_function,
-  x = x,
-  cl = cl)
+cl <- parallel::makeCluster(18L)
+
+parallel::clusterExport(
+  cl=cl,
+  varlist = c("compute_robust_lambda", "..optimisation_function")
+)
+
+experiment_results <- parallel::parLapplyLB(
+  cl = cl,
+  X = coro::collect(experiment_args()),
+  fun = .optimisation_function,
+  data = x,
+  chunk.size = 1L)
 
@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss.
+\caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Box-Cox} power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
 \midrule
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 783 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.84 & ---  & ---       & ---  & 587 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.78 & 0.95 & $(0, 1]$  & 0.99 & 587 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.98 & 583 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 1135 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.26 & 1.96 & $(0, 10]$ & 5.47 & 1813 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.03 & 1.96 & $(0, 10]$ & 6.91 & 1844 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.56 & ---  & ---       & ---  & 1837 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.49 & 1.00 & $(0, 10]$ & 1.51 & 1877 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.32 & 1.00 & $(0, 10]$ & 1.65 & 1834 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Yeo-Johnson} power transformations.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
+The {optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
+Lower loss indicates better robustness against outliers.}
+\begin{tabular}{l r r r r r r r}
+
+\toprule
+method & \multicolumn{3}{c}{$k_1$} & \multicolumn{3}{c}{$k_2$} & loss \\
+ & initial & limits & optimal & initial & limits & optimal & \\
+
+\midrule
 non-robust               & ---  & ---       & ---   & ---  & ---       & ---   & 225 \\
@@
 $p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.89  & 0.95 & $(0, 1]$  & 0.93  & 124 \\
-$z$ (step)               & 1.50 & $(0, 10]$ & 1.03 & ---  & ---       & ---   & 331 \\
-$z$ (triangle)           & 1.50 & $(0, 10]$ & 2.79 & 3.00 & $(0, 10]$ & 4.12  & 768 \\
-$z$ (tapered cosine)     & 1.50 & $(0, 10]$ & 2.69 & 3.00 & $(0, 10]$ & 4.21  & 772 \\
-$r$ (step)               & 0.15 & $(0, 10]$ & 0.05 & ---  & ---       & ---   & 1957 \\
-$r$ (triangle)           & 0.15 & $(0, 10]$ & 9.96 & 1.00 & $(0, 10]$ & 10.00 & 483 \\
-$r$ (tapered cosine)     & 0.15 & $(0, 10]$ & 9.97 & 1.00 & $(0, 10]$ & 10.00 & 484 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.04  & ---  & ---       & ---   & 333 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.12  & 1.96 & $(0, 10]$ & 9.91  & 620 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.23  & 1.96 & $(0, 10]$ & 5.63  & 716 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 10.00 & ---  & ---       & ---   & 485 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 9.96  & 1.00 & $(0, 10]$ & 10.00 & 483 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 9.97  & 1.00 & $(0, 10]$ & 10.00 & 484 \\
 \bottomrule
@@
 
-            #for(method in c("box_cox", "yeo_johnson")){
-            for(method in c("yeo_johnson")){
+            for(method in c("box_cox", "yeo_johnson")){
               for(weight_method in c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
@@
                 if(weight_method == "original_step"){
-                  k1 <- 0.92
+                  k1 <- ifelse(method == "box_cox", 0.84, 0.92)
 
                 } else if(weight_method == "original_cosine"){
-                  k1 <- 0.89
-                  k2 <- 0.93
+                  k1 <- ifelse(method == "box_cox", 0.76, 0.89)
+                  k2 <- ifelse(method == "box_cox", 0.98, 0.93)
 
                 } else if(weight_method == "transformed_step"){
-                  k1 <- 1.03
+                  k1 <- ifelse(method == "box_cox", 1.09, 1.04)
 
                 } else if(weight_method == "transformed_cosine"){
-                  k1 <- 2.69
-                  k2 <- 4.21
+                  k1 <- ifelse(method == "box_cox", 0.03, 0.12)
+                  k2 <- ifelse(method == "box_cox", 6.91, 5.63)
 
                 } else if(weight_method == "residual_step"){
-                  k1 <- 0.05
+                  k1 <- ifelse(method == "box_cox", 1.56, 10.00)
 
                 } else if(weight_method == "residual_cosine"){
-                  k1 <- 9.97
-                  k2 <- 10.0
+                  k1 <- ifelse(method == "box_cox", 1.32, 9.97)
+                  k2 <- ifelse(method == "box_cox", 1.65, 10.0)
                 }
@@
 
-  #### Computations --------------------------------------------------------------
+  # Computations ---------------------------------------------------------------
 
@@
   x = data$method,
-  levels = c("yeo_johnson"),
-  labels = c("Yeo-Johnson"))
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
 data$weight_method <- factor(
@@
 
-#### Normal distribution -------------------------------------------------------
+# Normal distribution ----------------------------------------------------------
 p_normal_t <- .create_outlier_fraction_plot(
@@
 
-#### Right skewed distribution -------------------------------------------------
+# Right skewed distribution ----------------------------------------------------
 p_right_t <- .create_outlier_fraction_plot(
@@
 
-#### Left skewed distribution --------------------------------------------------
+# Left skewed distribution -----------------------------------------------------
 p_left_t <- .create_outlier_fraction_plot(
@@
 
-#### Aggregate -----------------------------------------------------------------
+# Aggregate --------------------------------------------------------------------
 
@@
 
+## Empirical goodness of fit test
+
+To answer:
+ 1.
+What is the residual threshold and the volume required to avoid rejecting transformations?
+Optimise: 1.
+window; 2.
+threshold; 3.
+threshold fraction.
+?
+Create a figure that summarises residual error thresholds (50%, 80%, 90%, 95%, 99%, 100%), below which x% of the data falls.
+ 2. How does this correspond to distance between bimodal normal distributions?
+
+```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "CHECK. ")
+
+residual <- residual_error <- threshold <- NULL
+
+# Generate table of asymmetric generalised normal distribution parameters.
+n_distributions <- 100
+
+# Generate alpha, beta and n.
+n <- stats::runif(n=n_distributions, min=2, max=4)
+n <- ceiling(10^n)
+
+alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
+beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
+
+# Generate corresponding distributions.
+x <- mapply(
+  power.transform::ragn,
+  n = n,
+  alpha = alpha,
+  beta = beta)
+
+# Add outliers, between 0.00 and 0.10.
+n_outliers <- 10
+outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
+
+x <- mapply(
+  function(x, n, alpha, beta, outlier_fraction){
+
+    parameters <- list("n" = n, "alpha"=alpha, "beta" = beta)
+
+    x <- lapply(
+      outlier_fraction,
+      function(k, x, parameters){
+        # Set parameters.
+        parameters$k <- k
+
+        # Compute interquartile range.
+        interquartile_range <- stats::IQR(x)
+
+        # Compute upper and lower quartiles.
+        q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+        q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+        # Set data where the outliers will be copied into.
+        x_outlier <- x
+
+        if(k != 0.0){
+          n_draw <- ceiling(k * length(x))
+
+          # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+          # than Q3 + 1.5 IQR.
+          x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+          outlier <- numeric(n_draw)
+          if(any(x_random < 0)){
+            outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+          }
+
+          if(any(x_random >= 0)){
+            outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+          }
+
+          # Randomly insert outlier values.
+          x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
+        }
+
+        return(list(
+          "x" = x_outlier,
+          "parameters" = parameters))
+      },
+      x = x,
+      parameters)
+
+    return(x)
+  },
+  x = x,
+  n = n,
+  alpha = alpha,
+  beta = beta,
+  MoreArgs=list("outlier_fraction"=outlier_fraction),
+  SIMPLIFY=FALSE,
+  USE.NAMES=FALSE)
+
+x <- unlist(x, recursive = FALSE)
+x <- lapply(seq_along(x), function(ii, x){
+  x$distribution_id <- ii
+
+  return(x)
+})
+
+compute_residuals <- function(
+    x,
+    n_sample = 200){
+
+  # Set interpolation points.
+  p_sample <- (seq_len(n_sample) - 1/3) / (n_sample + 1/3)
+
+  # Determine residuals for Box-Cox transformations.
+  transformer_bc <- suppressWarnings(power.transform::find_transformation_parameters(
+    method = "box_cox",
+    shift = TRUE,
+    robust = TRUE,
+    weight_method = "original_cosine",
+    k1 = 0.89,
+    k2 = 0.93))
+
+  residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_bc)
+
+  residual <- stats::approx(
+    x = residual_data$p,
+    y = abs(residual_data$residual),
+    xout = p_sample,
+    rule = 2,
+    ties = max)
+
+  residual_bc <- data.table::data.table(
+    "distribution_id" = x$distribution_id,
+    "p" = p_sample,
+    "residual" = residual,
+    "method" = "box_cox")
+
+  # Determine residuals for Yeo-Johnson transformations.
+  transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
+    method = "yeo_johnson",
+    shift = TRUE,
+    robust = TRUE,
+    weight_method = "original_cosine",
+    k1 = 0.89,
+    k2 = 0.93))
+
+  residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_yj)
+
+  residual <- stats::approx(
+    x = residual_data$p,
+    y = abs(residual_data$residual),
+    xout = p_sample,
+    rule = 2,
+    ties = max)
+
+  residual_yj <- data.table::data.table(
+    "distribution_id" = x$distribution_id,
+    "p" = p_sample,
+    "residual" = residual,
+    "method" = "yeo_johnson")
+
+  return(data.table::rbindlist(list(residual_bc, residual_yj)))
+}
+
+data <- lapply(
+  x,
+  compute_residuals)
+
+data <- data.table::rbindlist(data)
+
+saveRDS(
+  object=data,
+  file=file.path(manuscript_dir, "residual_plot.RDS"))
+
+data <- data[, list(
+  "residual_50"=stats::quantile(residual, probs=0.50),
+  "residual_90"=stats::quantile(residual, probs=0.90),
+  "residual_95"=stats::quantile(residual, probs=0.95),
+  "residual_99"=stats::quantile(residual, probs=0.99),
+  "residual_max"=max(residual)),
+  by=c("p", "method")]
+
+data <- data.table::melt(
+  data = data,
+  id.vars = c("p", "method"),
+  variable.name = "threshold",
+  value.name = "residual_error")
+
+data$threshold <- factor(
+  x = data$threshold,
+  levels = c("residual_50", "residual_90", "residual_95", "residual_99", "residual_max"),
+  labels = c("50 %", "90 %", "95 %", "99 %", "100 %"))
+
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+
+p_bc <- ggplot2::ggplot(
+  data = data[method == "Box-Cox"],
+  mapping = ggplot2::aes(
+    x = p,
+    y = residual_error,
+    colour = threshold))
+p_bc <- p_bc + plot_theme
+p_bc <- p_bc + ggplot2::geom_line()
+p_bc <- p_bc + ggplot2::scale_colour_discrete(
+  name = "Box-Cox threshold",
+  type=list("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E"))
+p_bc <- p_bc <- ggplot2::xlab("empirical probability")
+
+p_yj <- ggplot2::ggplot(
+  data = data[method == "Yeo-Johnson"],
+  mapping = ggplot2::aes(
+    x = p,
+    y = residual_error,
+    colour = threshold))
+p_yj <- p_yj + plot_theme
+p_yj <- p_yj + ggplot2::geom_line()
+p_yj <- p_yj + ggplot2::scale_colour_discrete(
+  name = "Yeo-Johnson threshold",
+  type=list("50 %" = "#FFBD7D", "90 %" = "#FFA954", "95 %" = "#F28E2B", "99 %" = "#CD6B0B", "100 %" = "#A25000"))
+p_yj <- p_yj <- ggplot2::xlab("empirical probability")
+
+```
+
 # Experimental Results

### 2023-02-11 manuscript/manuscript.Rmd
WIP

@@
 # Density plot.
-.create_density_plot <- function(x, plot_theme){
+.create_density_plot <- function(x, plot_theme, limits=c(-5, 5)){
   p <- ggplot2::ggplot(
@@
       size=ggplot2::rel(1.0)))
-  p <- p + ggplot2::xlim(c(-5, 5))
+  p <- p + ggplot2::xlim(limits)
 
@@
         "d" = log10(parameter_set$d),
-        "lambda" = transformer@lambda
+        "lambda" = transformer@lambda,
+        "x_0" = transformer@shift
       ))
@@
 The double sum in right-hand side of the above equation can be considered as the loss that should be minimised.
-Minimisation is performed using the subplex algorithm implemented in the *nloptr* package for R [@Rowan1990-sn; @Ypma2022-tc].
+Minimisation is performed using the subplex algorithm implemented in the `nloptr` package for R [@Rowan1990-sn; @Ypma2022-tc].
 
@@
 
+if(!file.exists(file.path(manuscript_dir, "robustness_comparison.RDS"))){
   set.seed(95)
@@
   saveRDS(experiment_results, file = file.path(manuscript_dir, "robustness_comparison.RDS"))
+}
 ```
@@
 
-```{r outlier-values, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r, warning=FALSE, message=FALSE}
+# Only process data if the file has not been created.
+if(!file.exists(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))){
+
+  # Generator for creating parameters and data for all experiments.
+  generate_experiment_data <- coro::generator(
+    function(
+    x_normal,
+    x_right_skewed,
+    x_left_skewed){
+
+      n <- 20
+      k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 10)
+
+      for(distribution in c("normal", "right-skewed", "left-skewed")){
+        if(distribution == "normal"){
+          x <- x_normal
+
+        } else if(distribution == "right-skewed"){
+          x <- x_right_skewed
+
+        } else if(distribution == "left-skewed"){
+          x <- x_left_skewed
+        }
+        for(ii in seq_len(100)){
+          for(k in k_range){
+            # Compute interquartile range.
+            interquartile_range <- stats::IQR(x)
+
+            # Compute upper and lower quartiles.
+            q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+            q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+            # Set data where the outliers will be copied into.
+            x_outlier <- x
+
+            if(k != 0.0){
+              n_draw <- ceiling(k * length(x))
+
+              # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+              # than Q3 + 1.5 IQR.
+              x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+              outlier <- numeric(n_draw)
+              if(any(x_random < 0)){
+                outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+              }
+
+              if(any(x_random >= 0)){
+                outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+              }
+
+              # Randomly insert outlier values.
+              x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
+            }
+
+            for(method in c("box_cox", "yeo_johnson")){
+              for(weight_method in c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
+
+                k1 <- k2 <- NULL
+
+                # Optimal parameter values from the previous experiment.
+                if(weight_method == "original_step"){
+                  k1 <- ifelse(method == "box_cox", 0.84, 0.92)
+
+                } else if(weight_method == "original_cosine"){
+                  k1 <- ifelse(method == "box_cox", 0.76, 0.89)
+                  k2 <- ifelse(method == "box_cox", 0.98, 0.93)
+
+                } else if(weight_method == "transformed_step"){
+                  k1 <- ifelse(method == "box_cox", 1.09, 1.04)
+
+                } else if(weight_method == "transformed_cosine"){
+                  k1 <- ifelse(method == "box_cox", 0.03, 0.12)
+                  k2 <- ifelse(method == "box_cox", 6.91, 5.63)
+
+                } else if(weight_method == "residual_step"){
+                  k1 <- ifelse(method == "box_cox", 1.56, 10.00)
+
+                } else if(weight_method == "residual_cosine"){
+                  k1 <- ifelse(method == "box_cox", 1.32, 9.97)
+                  k2 <- ifelse(method == "box_cox", 1.65, 10.0)
+                }
+
+                yield(list(
+                  "ii" = ii,
+                  "x" = x_outlier,
+                  "k" = k,
+                  "distribution" = distribution,
+                  "method" = method,
+                  "robust" = weight_method != "none",
+                  "weight_method" = weight_method,
+                  "k1" = k1,
+                  "k2" = k2))
+              }
+            }
+          }
+        }
+      }
+    })
+
+  .compute_lambda <- function(parameter_set){
+    # Create transformer object.
+    transformer <- suppressWarnings(
+      power.transform::find_transformation_parameters(
+        x = parameter_set$x,
+        method = parameter_set$method,
+        robust = parameter_set$robust,
+        shift = TRUE,
+        weight_method = parameter_set$weight_method,
+        k1 = parameter_set$k1,
+        k2 = parameter_set$k2))
+
+    return(
+      data.table::data.table(
+        "distribution" = parameter_set$distribution,
+        "method" = parameter_set$method,
+        "weight_method" = parameter_set$weight_method,
+        "k" = parameter_set$k,
+        "lambda" = transformer@lambda,
+        "ii" = parameter_set$ii))
+  }
+
+  # Computations ---------------------------------------------------------------
+
+  # Set seed.
+  set.seed(19L)
+
+  # Normal distribution.
+  x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
+
+  # Right skewed data
+  x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
+
+  # Left skewed data
+  x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
+
+  # Generate all experiments.
+  experiments <- coro::collect(generate_experiment_data(
+    x_normal = x_normal,
+    x_right_skewed = x_right_skewed,
+    x_left_skewed = x_left_skewed
+  ))
+
+  # Start cluster
+  cl <- parallel::makeCluster(18L)
+
+  # Compute all data in parallel.
+  data <- parallel::parLapply(
+    cl=cl,
+    X=experiments,
+    fun=.compute_lambda
+  )
+
+  # Stop cluster.
+  parallel::stopCluster(cl)
+
+  # Combine data into a single table.
+  data <- data.table::rbindlist(data)
+
+  saveRDS(data, file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
+
+}
+```
+
+```{r outlier-values_yeo-johnson, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
-  "CHECK. ")
+  "Estimated $\\lambda$ values for \\textbf{shift-sensitive Yeo-Johnson} power transformations as a function of the fraction of outlier values. ",
+  "Features were randomly drawn from centered, right-skewed and left-skewed normal distributions, and a fraction of values was randomly replaced by outlier values.",
+  "The top row shows weighting methods that perform better than the non-robust version, ",
+  "namely empirical probabilities weighted using step (emp. prob. (step)) and tapered cosine (emp. prob. (tap. cos.)) weighting functions. ",
+  "The bottom row shows weighting methods that perform worse than the non-robust version. ",
+  "Note that the range of $\\lambda$ values differs between top and bottom rows.")
 
@@
 }
-# Only process data if the file has not been created.
-if(!file.exists(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))){
-
-  # Generator for creating parameters and data for all experiments.
-  generate_experiment_data <- coro::generator(
-    function(
-    x_normal,
-    x_right_skewed,
-    x_left_skewed){
-
-      n <- 20
-      k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 10)
-
-      for(distribution in c("normal", "right-skewed", "left-skewed")){
-        if(distribution == "normal"){
-          x <- x_normal
-
-        } else if(distribution == "right-skewed"){
-          x <- x_right_skewed
-
-        } else if(distribution == "left-skewed"){
-          x <- x_left_skewed
-        }
-        for(ii in seq_len(100)){
-          for(k in k_range){
-            # Compute interquartile range.
-            interquartile_range <- stats::IQR(x)
-
-            # Compute upper and lower quartiles.
-            q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
-            q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
-
-            # Set data where the outliers will be copied into.
-            x_outlier <- x
-
-            if(k != 0.0){
-              n_draw <- ceiling(k * length(x))
-
-              # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-              # than Q3 + 1.5 IQR.
-              x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
-              outlier <- numeric(n_draw)
-              if(any(x_random < 0)){
-                outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
-              }
-
-              if(any(x_random >= 0)){
-                outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
-              }
-
-              # Randomly insert outlier values.
-              x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
-            }
-
-            for(method in c("box_cox", "yeo_johnson")){
-              for(weight_method in c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
-
-                k1 <- k2 <- NULL
-
-                # Optimal parameter values from the previous experiment.
-                if(weight_method == "original_step"){
-                  k1 <- ifelse(method == "box_cox", 0.84, 0.92)
-
-                } else if(weight_method == "original_cosine"){
-                  k1 <- ifelse(method == "box_cox", 0.76, 0.89)
-                  k2 <- ifelse(method == "box_cox", 0.98, 0.93)
-
-                } else if(weight_method == "transformed_step"){
-                  k1 <- ifelse(method == "box_cox", 1.09, 1.04)
-
-                } else if(weight_method == "transformed_cosine"){
-                  k1 <- ifelse(method == "box_cox", 0.03, 0.12)
-                  k2 <- ifelse(method == "box_cox", 6.91, 5.63)
-
-                } else if(weight_method == "residual_step"){
-                  k1 <- ifelse(method == "box_cox", 1.56, 10.00)
-
-                } else if(weight_method == "residual_cosine"){
-                  k1 <- ifelse(method == "box_cox", 1.32, 9.97)
-                  k2 <- ifelse(method == "box_cox", 1.65, 10.0)
-                }
-
-                yield(list(
-                  "ii" = ii,
-                  "x" = x_outlier,
-                  "k" = k,
-                  "distribution" = distribution,
-                  "method" = method,
-                  "robust" = weight_method != "none",
-                  "weight_method" = weight_method,
-                  "k1" = k1,
-                  "k2" = k2))
-              }
-            }
-          }
-        }
-      }
-    })
-
-  .compute_lambda <- function(parameter_set){
-    # Create transformer object.
-    transformer <- suppressWarnings(
-      power.transform::find_transformation_parameters(
-        x = parameter_set$x,
-        method = parameter_set$method,
-        robust = parameter_set$robust,
-        shift = TRUE,
-        weight_method = parameter_set$weight_method,
-        k1 = parameter_set$k1,
-        k2 = parameter_set$k2))
-
-    return(
-      data.table::data.table(
-        "distribution" = parameter_set$distribution,
-        "method" = parameter_set$method,
-        "weight_method" = parameter_set$weight_method,
-        "k" = parameter_set$k,
-        "lambda" = transformer@lambda,
-        "ii" = parameter_set$ii))
-  }
-
-  # Computations ---------------------------------------------------------------
-
-  # Set seed.
-  set.seed(19L)
-
-  # Normal distribution.
-  x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
-
-  # Right skewed data
-  x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
-
-  # Left skewed data
-  x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
-
-  # Generate all experiments.
-  experiments <- coro::collect(generate_experiment_data(
-    x_normal = x_normal,
-    x_right_skewed = x_right_skewed,
-    x_left_skewed = x_left_skewed
-  ))
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  # Compute all data in parallel.
-  data <- parallel::parLapply(
-    cl=cl,
-    X=experiments,
-    fun=.compute_lambda
-  )
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-
-  # Combine data into a single table.
-  data <- data.table::rbindlist(data)
-
-  saveRDS(data, file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
-
-} else {
+
 data <- readRDS(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
-}
 
@@
 
-data <- data[, list(
+data <- data[method == "Yeo-Johnson", list(
   "mean_lambda"=mean(lambda),
@@
 
+```{r outlier-values-box-cox, warning=FALSE, message=FALSE, fig.cap=cap}
+
+data <- readRDS(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
+
+cap <- paste0(
+  "Estimated $\\lambda$ values for \\textbf{shift-sensitive Box-Cox} power transformations as a function of the fraction of outlier values. ",
+  "Features were randomly drawn from centered, right-skewed and left-skewed normal distributions, and a fraction of values was randomly replaced by outlier values.",
+  "The top row shows weighting methods that perform better than the non-robust version, ",
+  "namely empirical probabilities weighted using step (emp. prob. (step)) and tapered cosine (emp. prob. (tap. cos.)) weighting functions. ",
+  "The bottom row shows weighting methods that perform worse than the non-robust version. ",
+  "Note that the range of $\\lambda$ values differs between top and bottom rows.")
+
+# Update distribution, method and version to factors.
+data$distribution <- factor(
+  x = data$distribution,
+  levels = c("normal", "right-skewed", "left-skewed"))
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+data$weight_method <- factor(
+  x = data$weight_method,
+  levels = c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
+  labels = c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
+
+data <- data[method == "Box-Cox", list(
+  "mean_lambda"=mean(lambda),
+  "min_lambda"=min(lambda),
+  "max_lambda"=max(lambda)),
+  by=c("distribution", "method", "weight_method", "k")]
+
+top_group <- c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)")
+bottom_group <- c("z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
+
+top_limits <- c(-3, 4.5)
+bottom_limits <- c(-4.01, 6.01)
+
+# Normal distribution ----------------------------------------------------------
+p_normal_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "normal" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits = top_limits,
+  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_normal_t <- p_normal_t + ggplot2::ggtitle(label="normal distribution")
+p_normal_t <- p_normal_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank())
+
+p_normal_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "normal" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits = bottom_limits,
+  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_normal_b <- p_normal_b + ggplot2::theme(
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+# Right skewed distribution ----------------------------------------------------
+p_right_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "right-skewed" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits= top_limits,
+  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_right_t <- p_right_t + ggplot2::ggtitle(label="right-skewed distribution")
+p_right_t <- p_right_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank())
+
+p_right_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "right-skewed" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits = bottom_limits,
+  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_right_b <- p_right_b + ggplot2::theme(
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+# Left skewed distribution -----------------------------------------------------
+p_left_t <- .create_outlier_fraction_plot(
+  data = data[distribution == "left-skewed" & weight_method %in% top_group],
+  plot_theme = plot_theme,
+  limits = top_limits,
+  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda,
+  guide = TRUE)
+p_left_t <- p_left_t + ggplot2::ggtitle(label="left-skewed distribution")
+p_left_t <- p_left_t + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank(),
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank())
+
+p_left_b <- .create_outlier_fraction_plot(
+  data = data[distribution == "left-skewed" & weight_method %in% bottom_group],
+  plot_theme = plot_theme,
+  limits = bottom_limits,
+  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
+p_left_b <- p_left_b + ggplot2::theme(
+  axis.title.y = ggplot2::element_blank(),
+  axis.text.y = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
+
+# Aggregate --------------------------------------------------------------------
+
+p <- egg::ggarrange(
+  plots=list(
+    p_normal_t, p_right_t, p_left_t,
+    p_normal_b, p_right_b, p_left_b),
+  ncol=3,
+  heights=c(1.0, 1.0),
+  byrow=TRUE,
+  draw=FALSE)
+p
+
+```
+
 ## Empirical goodness of fit test
@@
 
+if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))){
+
+  set.seed(95)
+
   # Generate table of asymmetric generalised normal distribution parameters.
-n_distributions <- 100
+  n_distributions <- 1000
 
@@
   x <- unlist(x, recursive = FALSE)
-x <- lapply(seq_along(x), function(ii, x){
+  x <- mapply(
+    function(ii, x){
       x$distribution_id <- ii
 
-  return(x)
-})
+      return(x)},
+    ii = seq_along(x),
+    x = x,
+    SIMPLIFY = FALSE)
 
@@
     transformer_bc <- suppressWarnings(power.transform::find_transformation_parameters(
+      x = x$x,
       method = "box_cox",
@@
       weight_method = "original_cosine",
-    k1 = 0.89,
-    k2 = 0.93))
+      k1 = 0.76,
+      k2 = 0.95))
 
@@
       rule = 2,
-    ties = max)
+      ties = max)$y
 
@@
       "residual" = residual,
-    "method" = "box_cox")
+      "method" = "box_cox",
+      "has_outliers" = x$parameters$k > 0)
 
@@
     transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
+      x = x$x,
       method = "yeo_johnson",
@@
       "residual" = residual,
-    "method" = "yeo_johnson")
+      "method" = "yeo_johnson",
+      "has_outliers" = x$parameters$k > 0)
 
@@
 
-data <- lapply(
-  x,
-  compute_residuals)
+  # Start cluster
+  cl <- parallel::makeCluster(18L)
+
+  # Compute all data in parallel.
+  data <- parallel::parLapply(
+    cl=cl,
+    X=x,
+    fun=compute_residuals)
+
+  # Stop cluster.
+  parallel::stopCluster(cl)
+
+  # data <- lapply(
+  #   x,
+  #   compute_residuals)
 
@@
     file=file.path(manuscript_dir, "residual_plot.RDS"))
+
+} else {
+  data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
+}
+
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+
+outlier_free_data <- data[has_outliers == FALSE, list("residual_error"=stats::quantile(residual, probs=0.99)), by=c("p", "method")]
 
@@
   labels = c("50 %", "90 %", "95 %", "99 %", "100 %"))
-
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
 
@@
 p_bc <- p_bc + ggplot2::geom_line()
+p_bc <- p_bc + ggplot2::geom_line(
+  data = outlier_free_data[method == "Box-Cox"],
+  mapping = ggplot2::aes(
+    x = p,
+    y = residual_error),
+  colour = "gray40",
+  linetype = "dashed")
 p_bc <- p_bc + ggplot2::scale_colour_discrete(
   name = "Box-Cox threshold",
-  type=list("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E"))
-p_bc <- p_bc <- ggplot2::xlab("empirical probability")
+  type=c("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E"))
+p_bc <- p_bc + ggplot2::xlab("empirical probability")
+p_bc <- p_bc + ggplot2::ylab("absolute residual error")
+p_bc <- p_bc + ggplot2::coord_cartesian(
+  xlim = c(0, 1),
+  ylim = c(0, 1))
+p_bc <- p_bc + ggplot2::theme(
+  axis.title.x = ggplot2::element_blank(),
+  axis.text.x = ggplot2::element_blank())
 
@@
 p_yj <- p_yj + ggplot2::geom_line()
+p_yj <- p_yj + ggplot2::geom_line(
+  data = outlier_free_data[method == "Yeo-Johnson"],
+  mapping = ggplot2::aes(
+    x = p,
+    y = residual_error),
+  colour = "gray40",
+  linetype = "dashed")
 p_yj <- p_yj + ggplot2::scale_colour_discrete(
   name = "Yeo-Johnson threshold",
-  type=list("50 %" = "#FFBD7D", "90 %" = "#FFA954", "95 %" = "#F28E2B", "99 %" = "#CD6B0B", "100 %" = "#A25000"))
-p_yj <- p_yj <- ggplot2::xlab("empirical probability")
-
+  type=c("50 %" = "#FFBD7D", "90 %" = "#FFA954", "95 %" = "#F28E2B", "99 %" = "#CD6B0B", "100 %" = "#A25000"))
+p_yj <- p_yj + ggplot2::xlab("empirical probability")
+p_yj <- p_yj + ggplot2::ylab("absolute residual error")
+p_yj <- p_yj + ggplot2::coord_cartesian(
+  xlim = c(0, 1),
+  ylim = c(0, 1))
+
+p <- egg::ggarrange(
+  plots=list(
+    p_bc,
+    p_yj),
+  ncol=1,
+  heights=c(1.0, 1.0),
+  byrow=TRUE,
+  draw=FALSE)
+p
+```
+
+```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
+
+if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))) stop("Residual data for goodness-of-fit test does not exist.")
+
+data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
+
+central_width <- c(0.60, 0.70, 0.80, 0.90, 0.95, 1.00)
+
+p_lower <- 0.50 - central_width / 2
+p_upper <- 0.50 + central_width / 2
+
+x <- data.table::copy(data)
+
+# Clip empirical probabilities to centre.
+x <- x[p >= p_lower & p <= p_upper]
+
+# Compute mean absolute residual error per feature.
+x <- x[, list("mare"=mean(residual)), by=c("distribution_id", "method")]
+x <- x[, list("n"=.N), by=c("mare", "method")][order(mare, method)]
+x[, "rejected":= 1.0 - cumsum(n) / sum(n)]
+x <- rbind(data.table::data.table("mare"=c(0.0, 0.0), "method"=c("box_cox", "yeo_johnson"), "n"=0L, "rejected"=0.0))
+
+p <- ggplot2::ggplot(
+  data = x,
+  mapping = ggplot2::aes(
+    x = mare,
+    y = rejected,
+    colour = method))
+p <- p + plot_theme
+p <- p + ggplot2::geom_step()
+p <- p + ggplot2::xlab("mean absolute residual error")
+p <- p + paletteer::scale_color_paletteer_d(
+  palette="ggthemes::Tableau_10",
+  drop=FALSE)
 ```
@@
 
+```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+
+if(!file.exists(file.path(manuscript_dir, "robustness_comparison_rightside_outlier.RDS"))){
+  set.seed(95)
+
+  .compute_lambda <- function(x, method){
+    # Create transformer object.
+    transformer <- suppressWarnings(
+      power.transform::find_transformation_parameters(
+        x = x,
+        method = method,
+        robust = FALSE,
+        shift = TRUE))
+
+    return(transformer@lambda)
+  }
+
+  # Generate table of asymmetric generalised normal distribution parameters.
+  n_distributions <- 100
+
+  # Generate alpha, beta and n.
+  n <- stats::runif(n=n_distributions, min=2, max=4)
+  n <- ceiling(10^n)
+
+  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
+  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
+
+  # Generate corresponding distributions.
+  x <- mapply(
+    power.transform::ragn,
+    n = n,
+    alpha = alpha,
+    beta = beta)
+
+  # Compute lambda values without weighting for Box-Cox.
+  target_lambda_bc <- sapply(
+    x,
+    .compute_lambda,
+    method = "box_cox")
+
+  # Compute lambda values without weighting for Yeo-Johnson
+  target_lambda_yj <- sapply(
+    x,
+    .compute_lambda,
+    method = "yeo_johnson")
+
+  # Add outliers, between 0.00 and 0.10.
+  n_outliers <- 10
+  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 20)
+
+  # Find optimised k1 and k2 values for all combinations of weight sources and
+  # weighting functions.
+  x <- mapply(
+    function(x, n, alpha, beta, target_lambda_bc, target_lambda_yj, outlier_fraction){
+
+      parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda_bc"=target_lambda_bc, "target_lambda_yj"=target_lambda_yj)
+
+      x <- lapply(
+        outlier_fraction,
+        function(k, x, parameters){
+          # Set parameters.
+          parameters$k <- k
+
+          # Compute interquartile range.
+          interquartile_range <- stats::IQR(x)
+
+          # Compute upper and lower quartiles.
+          q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+          q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+          # Set data where the outliers will be copied into.
+          x_outlier <- x
+
+          if(k != 0.0){
+            n_draw <- ceiling(k * length(x))
+
+            # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+            # than Q3 + 1.5 IQR.
+            x_random <- stats::runif(n_draw, min=0.0, max=2.0)
+            outlier <- q_upper + 1.5 * interquartile_range + x_random * interquartile_range
+
+            # Randomly insert outlier values.
+            x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
+          }
+
+          return(list(
+            "x" = x_outlier,
+            "parameters" = parameters))
+        },
+        x = x,
+        parameters)
+
+      return(x)
+    },
+    x = x,
+    n = n,
+    alpha = alpha,
+    beta = beta,
+    target_lambda_bc = target_lambda_bc,
+    target_lambda_yj = target_lambda_yj,
+    MoreArgs=list("outlier_fraction"=outlier_fraction),
+    SIMPLIFY=FALSE,
+    USE.NAMES=FALSE)
+
+  x <- unlist(x, recursive = FALSE)
+
+  # Optimise k1, k2 so that sum of absolute differences with target-lambda is
+  # minimised for each combination of source and weighting method.
+  experiment_args <- coro::generator(
+    function(){
+
+      for(method in c("box_cox", "yeo_johnson")){
+        # Non-robust transformation.
+        fun_args <- list(
+          "method" = method,
+          "robust" = FALSE,
+          "shift" = TRUE)
+
+        opt_args <- list()
+
+        yield(list(
+          "name" = "non-robust",
+          "method" = method,
+          "fun_args" = fun_args,
+          "opt_args" = opt_args
+        ))
+      }
+
+      for(robustness_source in c("original", "transformed", "residual")){
+
+        def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
+
+        if(robustness_source == "original"){
+          def_opt_limits$k1[2] <- 1.0
+          def_opt_limits$k2[2] <- 1.0
+
+          def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
+
+        } else if(robustness_source == "transformed"){
+          def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
+
+        } else if(robustness_source == "residual"){
+          def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
+        }
+
+        for(method in c("box_cox", "yeo_johnson")){
+          for(robustness_weighting_function in c("step", "triangle", "cosine")){
+
+            fun_args <- list(
+              "method" = method,
+              "robust" = TRUE,
+              "shift" = TRUE,
+              "weight_method" = paste0(robustness_source, "_", robustness_weighting_function),
+              "backup_use_default" = FALSE)
+
+            opt_limits <- def_opt_limits
+            opt_init <- def_opt_init
+
+            # Step-weighting only has k1 as a parameter.
+            if(robustness_weighting_function == "step"){
+              opt_limits$k2 <- NULL
+              opt_init$k2 <- NULL
+            }
+
+            yield(list(
+              "name" = paste0(robustness_source, "-", robustness_weighting_function),
+              "method" = method,
+              "fun_args" = fun_args,
+              "opt_args" = list("initial" = opt_init, "limits" = opt_limits)
+            ))
+          }
+        }
+      }
+    })
+
+  # Function for computing transformer parameters under optimisation constraints.
+  compute_robust_lambda <- function(
+    x,
+    fun_args,
+    opt_args,
+    cl_internal=NULL){
+
+    # Custom parser.
+    ..outlier_parser <- function(
+    x,
+    fun_args,
+    opt_args
+    ){
+      # Create transformer.
+      transformer <- suppressWarnings(do.call(
+        power.transform::find_transformation_parameters,
+        args=c(
+          list("x" = x$x),
+          fun_args,
+          opt_args)))
+
+      parameter_data <- c(
+        x$parameters,
+        opt_args,
+        list(
+          "method" = fun_args$method,
+          "weight_method" = fun_args$weight_method,
+          "lambda" = transformer@lambda,
+          "shift" = transformer@shift)
+      )
+
+      return(parameter_data)
+    }
+
+    if(is.null(cl_internal)){
+      parameter_data <- lapply(
+        x,
+        ..outlier_parser,
+        fun_args = fun_args,
+        opt_args = opt_args)
+
+    } else {
+      parameter_data <- parallel::parLapply(
+        cl = cl_internal,
+        X = x,
+        fun = ..outlier_parser,
+        fun_args = fun_args,
+        opt_args = opt_args)
+    }
+
+    return(parameter_data)
+  }
+
+  # Inner optimisation function that sets the loss.
+  ..optimisation_function <- function(
+    opt_param,
+    x,
+    fun_args,
+    cl_internal = NULL){
+
+    # Parse options.
+    opt_args <- list()
+    if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
+    if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
+
+    parameter_data <- compute_robust_lambda(
+      x = x,
+      fun_args = fun_args,
+      opt_args = opt_args,
+      cl_internal = cl_internal)
+
+    if(fun_args$method == "box_cox"){
+      lambda_error <- sapply(
+        parameter_data,
+        function(x) (abs(x$lambda - x$target_lambda_bc)))
+
+    } else {
+      # Yeo-Johnson.
+      lambda_error <- sapply(
+        parameter_data,
+        function(x) (abs(x$lambda - x$target_lambda_yj)))
+    }
+
+    return(sum(lambda_error))
+  }
+
+  # Outer optimisation function
+  .optimisation_function <- function(
+    experiment,
+    data,
+    cl_internal = NULL){
+
+    set.seed(9)
+
+    # Run optimiser.
+    if(length(experiment$opt_args) > 0){
+
+      x0 <- unlist(experiment$opt_args$initial)
+
+      lower <- experiment$opt_args$limits$k1[1]
+      upper <- experiment$opt_args$limits$k1[2]
+
+      if(length(experiment$opt_args$limits) > 1){
+        lower <- c(lower, experiment$opt_args$limits$k2[1])
+        upper <- c(upper, experiment$opt_args$limits$k2[2])
+      }
+
+      h <- nloptr::sbplx(
+        x0 = x0,
+        fn = ..optimisation_function,
+        lower = lower,
+        upper = upper,
+        control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
+        x = data,
+        fun_args = experiment$fun_args,
+        cl_internal = cl_internal)
+
+      if(length(h$par) == 1){
+        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par, "value" = h$value))
+
+      } else {
+        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
+      }
+
+    } else {
+      value <- ..optimisation_function(
+        opt_param=list(),
+        x = data,
+        fun_args=experiment$fun_args,
+        cl = cl_internal)
+
+      return(list("name"=experiment$name, "method"=experiment$method, "value" = value))
+    }
+  }
+
+  # Start cluster
+  cl <- parallel::makeCluster(18L)
+
+  parallel::clusterExport(
+    cl=cl,
+    varlist = c("compute_robust_lambda", "..optimisation_function")
+  )
+
+  experiment_results <- parallel::parLapplyLB(
+    cl = cl,
+    X = coro::collect(experiment_args()),
+    fun = .optimisation_function,
+    data = x,
+    chunk.size = 1L)
+
+  # Stop cluster.
+  parallel::stopCluster(cl)
+
+  saveRDS(experiment_results, file = file.path(manuscript_dir, "robustness_comparison_rightside_outlier.RDS"))
+}
+```
+
 # References

### 2023-02-14 manuscript/manuscript.Rmd
WIP on generalising estimators and weights.

@@
 Statistical tests for normality exist, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-However, given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for practical applications of automated power transformations.
+However, given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for applications of automated power transformations in practice.
 

### 2023-02-14 R/ParameterEstimatorMLE.R

@@ -1,0 +1,82 @@
+#' @include ParameterEstimators.R
+NULL
+
+
+
+# estimatorMaximumLikelihoodEstimation definition ------------------------------
+setClass(
+  "estimatorMaximumLikelihoodEstimation",
+  contains = "estimatorGeneric")
+
+
+
+# .compute_objective (MLE) -----------------------------------------------------
+setMethod(
+  ".compute_objective",
+  signature(object="estimatorMaximumLikelihoodEstimation"),
+  function(object, transformer, x, ...){
+
+    # Make sure that x is sorted.
+    if(is.unsorted(x)) x <- sort(x)
+
+    w <- .get_weights(
+      object = object,
+      transformer = transformer,
+      x = x)
+
+    if(!all(is.finite(w))) return(NA_real_)
+
+    # Transform x under the provided lambda.
+    y <- ..transform(
+      object = transformer,
+      x = x)
+
+    if(any(!is.finite(y))) return(NA_real_)
+
+    # Compute log-likelihood
+    llf <- .log_likelihood(
+      transformer = transformer,
+      y = y,
+      w = w)
+
+    # Log-likelihood should be maximised, hence return -llf because optimisers
+    # use minimisation.
+    return(-llf)
+  }
+)
+
+
+.log_likelihood <- function(
+    transformer,
+    y,
+    w){
+
+  # Compute the sum of the weights.
+  sum_w <- sum(w)
+  if(sum_w == 0.0) return(NA_real_)
+
+  # Compute the weighted estimates of the mean mu and variance sigma squared
+  # for y.
+  mu_hat <- sum(w * y) / sum_w
+  sigma_hat_squared <- sum(w * (y - mu_hat)^2) / sum_w
+
+  # Log-likelihood cannot be estimated if sigma is NaN, or equals 0.0.
+  if(!is.finite(sigma_hat_squared)) return(NA_real_)
+  if(sigma_hat_squared == 0) return(NA_real_)
+
+  # Compute the log likelihood under the assumption that the transformed
+  # variable y follows the normal distribution.
+  llf <- ..log_likelihood(
+    object = transformer,
+    x = x,
+    w = w,
+    sigma_hat_squared = sigma_hat_squared)
+
+  return(llf)
+}
+
+
+
+..estimators_mle <- function(){
+  return(c("mle"))
+}

### 2023-02-14 R/ParameterEstimatorRaymaekers.R

@@ -1,0 +1,236 @@
+#' @include ParameterEstimators.R
+NULL
+
+
+# estimatorRaymaekersRobust definition -----------------------------------------
+setClass(
+  "estimatorRaymaekersRobust",
+  contains = "estimatorGeneric")
+
+
+
+# ..optimise_transformation_parameters (Raymaekers robust) ---------------------
+setMethod(
+  ".optimise_transformation_parameters",
+  signature(object = "estimatorRaymaekersRobust"),
+  function(
+    object,
+    transformer,
+    optimiser,
+    optimisation_parameters,
+    x,
+    ...){
+
+    # Check that we do not inadvertently pass problems that do not require
+    # optimisation to the optimiser.
+    if(is.null(optimisation_parameters)) stop("DEV: optimisation_parameters cannot be empty.")
+    if(!is(transformer, "transformationPowerTransform")) stop("DEV: transformer should be a valid power transformation object.")
+    if(!is.numeric(x)) stop("DEV: x should be numeric.")
+
+    # Perform method-specific checks on proper usage.
+    if(!setequal(optimisation_parameters$parameter_type, "lambda")) stop("DEV: Raymaekers and Rousseeuws robust method can only be used for optimising lambda.")
+    if(!transformer@robust) stop("DEV: Raymaekers and Rousseuws robust method is intended for robust optimisation, but the robust attribute is FALSE.")
+
+    # Sort x if necessary.
+    if(is.unsorted(x)) x <- sort(x)
+
+    # Compute z-values according to the inverse cumulative density function.
+    z_expected <- compute_expected_z(x=x)
+
+    # Step 1: Compute initial estimate for lambda, based on residuals with
+    # rectified residual errors.
+    lambda <- suppressWarnings(
+      stats::optimise(
+        .raymaekers_robust_rectified_optimisation,
+        interval=c(optimisation_parameters$lower, optimisation_parameters$upper),
+        x = x,
+        z = z_expected,
+        transformer = transformer,
+        maximum = FALSE))
+
+    lambda <- ifelse(
+      is.finite(lambda$objective),
+      lambda$minimum,
+      1.0)
+
+    # Step 2: Compute lambda from reweighted maximum likelihood.
+    lambda <- .raymaekers_robust_reweighted_optimisation(
+      lambda = lambda,
+      x = x,
[…]

### 2023-02-14 R/ParameterEstimators.R

@@ -1,0 +1,249 @@
+..estimator_wrapper <- function(
+    par,
+    transformer,
+    estimator,
+    x,
+    parameter_type,
+    ...){
+
+  # Set lambda, shift, as provided by the optimisation algorithm.
+  if(setequal(parameter_type, c("lambda", "shift"))){
+    transformer@shift <- par[1]
+    transformer@lambda <- par[2]
+
+  } else if(parameter_type == "lambda" & length(par) == 1){
+    transformer@lambda <- par[1]
+
+  } else if(parameter_type == "shift" & length(par) == 1){
+    transformer@shift <- par[1]
+
+  } else {
+    stop("DEV: Optimisation function was not specified correctly.")
+  }
+
+  return(.compute_objective(
+    object = estimator,
+    transformer = transformer,
+    x = x,
+    ...))
+}
+
+
+
+# estimatorGeneric definition --------------------------------------------------
+setClass(
+  "estimatorGeneric",
+  slots = list(
+    "weighting_method" = "ANY"),
+  prototype = list(
+    "weighting_method" = NULL))
+
+setClass(
+  "estimatorEmpiricalDistributionFunction",
+  contains = "estimatorGeneric")
+
+setClass(
+  "estimatorAndersonDarling",
+  contains = "estimatorEmpiricalDistributionFunction")
+
+setClass(
+  "estimatorCramervonMises",
+  contains = "estimatorEmpiricalDistributionFunction")
+
+setClass(
+  "estimatorKolmogorovSmirnov",
+  contains = "estimatorEmpiricalDistributionFunction")
+
+setClass(
+  "estimatorSkewnessKurtosis",
+  contains = "estimatorGeneric")
+
+setClass(
+  "estimatorJarqueBera",
+  contains = "estimatorSkewnessKurtosis")
+
+setClass(
+  "estimatorDAgostino",
+  contains = "estimatorSkewnessKurtosis")
+
+
+
+# .compute_objective (generic) -------------------------------------------------
+setGeneric(
+  ".compute_objective",
+  function(object, ...) standardGeneric(".compute_estimator"))
+
+
+
+# .compute_objective (general) -------------------------------------------------
+setMethod(
+  ".compute_objective",
+  signature(object = "estimatorGeneric"),
+  function(object, ...){
+
+    # Estimators should be specified for each estimator - an error is thrown to
+    # ensure that this general method is never used.
[…]

### 2023-02-14 R/RobustTransformation.R

@@ -1,157 +1,0 @@
-.transformation_robust_optimisation <- function(x, type, lambda_range){
-  # This follows the algorithm from Raymaekers J, Rousseeuw PJ. Transforming
-  # variables to central normality. Mach Learn. 2021.
-  # doi:10.1007/s10994-021-05960-5
-
-  # Pick very wide lambda-range, if lambda_range is NULL.
-  if(is.null(lambda_range)) lambda_range <- c(-100.0, 100.0)
-
-  # Sort x if necessary.
-  if(is.unsorted(x)) x <- sort(x)
-
-  # Compute z-values according to the inverse cumulative density function.
-  z_expected <- compute_expected_z(x=x)
-
-  # Step 1: Compute initial estimate for lambda.
-  # Standard method based on optimising log-likelihood of the normal
-  # distribution.
-
-  optimal_lambda <- suppressWarnings(
-    stats::optimise(
-      ..transformation_rectified_optimisation,
-      interval=lambda_range,
-      x=x,
-      z=z_expected,
-      type=type,
-      maximum=FALSE))
-
-  optimal_lambda <- ifelse(
-    is.finite(optimal_lambda$objective),
-    optimal_lambda$minimum,
-    1.0)
-
-  # Step 2: Compute lambda from reweighted maximum likelihood.
-  optimal_lambda <- ..transformation_reweighted_optimisation(
-    lambda_0=optimal_lambda,
-    x=x,
-    type=type,
-    ii=1L,
-    lambda_range=lambda_range)
-
-  # Step 3: Compute lambda from reweighted maximum likelihood again.
-  optimal_lambda <- ..transformation_reweighted_optimisation(
-    lambda_0=optimal_lambda,
-    x=x,
-    type=type,
-    ii=2L,
-    lambda_range=lambda_range)
-
-  if(!is.finite(optimal_lambda)) return(1.0)
-
-  return(optimal_lambda)
-}
-
-
-
-..transformation_rectified_optimisation <- function(lambda, x, z, type){
-
-  rectifier_FUN <- switch(
-    type,
-    "box_cox"=..box_cox_transform_rectified,
-    "yeo_johnson"=..yeo_johnson_transform_rectified
-  )
-
-  # Compute values after power transformation with the appropriate lambda value.
-  y <- do.call(
-    rectifier_FUN,
-    args=list(
-      "lambda"=lambda,
-      "x"=x))
[…]

### 2023-02-14 R/TransformationObjects.R

@@ -1,3 +1,5 @@
+# transformationPowerTransform  definition -------------------------------------
+
 #' Generic transformation object
 #'
 #' This is the superclass for transformation objects.
@@ -16,7 +18,7 @@ setClass(
     "method" = "none",
     "complete" = FALSE))
 
-
+# transformationNone definition ------------------------------------------------
 
 #' No transformation object
 #'
@@ -33,101 +35,17 @@ setClass(
 
 
 
-#' Box-Cox transformation object
-#'
-#' This class is used for Box-Cox transformations.
-#'
-#' @slot method Main transformation method, i.e. `"box_cox"`.
-#' @slot robust Indicates whether a robust version of the Box-Cox transformation
-#'   is used to set transformation parameters. The value depends on the `robust`
-#'   argument of the `find_transformation_parameters` function.
-#' @slot lambda Numeric lambda parameter for the Box-Cox transformation.
-#' @slot shift Numeric shift parameter for the Box-Cox transformation. The value
-#'   depends on the data used for setting transformation parameters. If all data
-#'   are strictly positive, `shift` has a value of `0.0`. When negative or zero
-#'   values are present, data are shifted to be strictly positive. If
-#'   `shift=TRUE` in the `find_transformation_parameters` function, `lambda` and
-#'   `shift` parameters are optimised simultaneously.
-#' @slot complete Indicates whether transformation parameters were set.
-#'
-#' @seealso [find_transformation_parameters]
-#' @export
-#' @rdname transformation_box_cox
-
-setClass(
-  "transformationBoxCox",
-  contains="transformationPowerTransform",
-  slots=list(
-    "method" = "character",
-    "robust" = "logical",
-    "lambda" = "numeric",
-    "shift" = "numeric"),
-  prototype=list(
-    "method" = "box_cox",
-    "robust" = FALSE,
-    "lambda" = NA_real_,
-    "shift" = 0.0))
-
-#' @rdname transformation_box_cox
-#' @export
-setClass(
-  "transformationBoxCoxShift",
-  contains="transformationBoxCox")
-
-
-
-#' Yeo-Johnson transformation object
-#'
-#' This class is used for Yeo-Johnson transformations.
-#'
-#' @slot method Main transformation method, i.e. `"yeo_johnson"`.
-#' @slot robust Indicates whether a robust version of the Yeo-Johnson
[…]

### 2023-02-17 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
       shift = FALSE,
-      lambda = NULL))
+      lambda = NULL,
+      optimiser_control = list("xtol_rel"=1e-3)))
 
@@
         shift = parameter_set$shift,
-        lambda = NULL))
+        lambda = NULL,
+        optimiser_control = list("xtol_rel"=1e-3)))
 
@@
 
-      for(robustness_source in c("original", "transformed", "residual")){
+      for(robustness_source in c("empirical_probability", "transformed", "residual")){
 
@@
 
-        if(robustness_source == "original"){
+        if(robustness_source == "empirical_probability"){
           def_opt_limits$k1[2] <- 1.0
@@
               "shift" = TRUE,
-              "weight_method" = paste0(robustness_source, "_", robustness_weighting_function),
+              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
               "backup_use_default" = FALSE)
@@
       }
-    })
+    }
+  )
 
@@
           list("x" = x$x),
-          fun_args,
-          opt_args)))
+          list("weighting_function_parameters" = opt_args),
+          fun_args)))
 
@@
     }
-
-    return(sum(lambda_error))
+    cat(paste0("target: ", mean(lambda_error), "; k1: ", opt_param[1], "; k2: ", opt_param[2], "\n"))
+    return(mean(lambda_error))
   }
@@
         upper = upper,
-        control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
+        control=list("xtol_rel"=1e-2),
         x = data,

### 2023-02-21 manuscript/manuscript.Rmd
Added description of additional criteria to the manuscript.

@@
 This weighted maximum log-likelihood estimator is based on equations **INSERT**.
-The weighted maximum log-likelihood estimator for the shift-sensitive Box-Cox power transformation is:
-
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
+Here, the estimator is further normalised by the total weight and extended to include the shift parameter.
+The shift-sensitive Box-Cox power transformation is:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
 \end{equation}
@@
 \begin{equation}
-\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
+\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
 \end{equation}
@@
 data_step <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_step[, "weight":=power.transform:::.step_window(x, k1=0.60)]
+data_step[, "weight":=power.transform:::..weight_function_external(x, "step", k1=0.60)]
 
@@
 data_triangle <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_triangle[, "weight":=power.transform:::.triangular_window(x, k1=0.30, k2=0.90)]
+data_triangle[, "weight":=power.transform:::..weight_function_external(x, "triangle", k1=0.30, k2=0.90)]
 
@@
 data_cosine <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_cosine[, "weight":=power.transform:::.tapered_cosine_window(x, k1=0.30, k2=0.90)]
+data_cosine[, "weight":=power.transform:::..weight_function_external(x, "cosine", k1=0.30, k2=0.90)]
 
@@
 
-# Appendix A.
+# Appendix A: Non-MLE-based optimisation of transformation parameters
+
+Maximum likelihood estimation (MLE) is commonly used to optimise parameters for power transformation.
+Generally, optimisation requires minimisation or maximisation of a criterion.
+In MLE, the maximised criterion is the log-likelihood function of the normal distribution.
+Here, we investigate power transformation using optimisation criteria that are closely related to test statistics for normality tests.
+
+Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift-sensitive power transformations, respectively.
+Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
+
+Below we will focus on criteria based on empirical density function, and on skewness and kurtosis of the transformed featured.
+Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
+In the case of the Shapiro-Wilk test statistic this is because of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.
+
+## Empirical density function-based criteria
+
+The first class of criteria is based on the empirical distribution function (EDF).
+Transformation parameters are then fit through minimisation of the distance between the empirical distribution function $F_{\epsilon}$ and the cumulative density function (CDF) of the normal distribution $F_{\mathcal{N}}$.
+Let $F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}$ be the empirical probability of instance $i$.
+The normal distribution is parametrised by location parameter $\mu$ and scale parameter $\sigma$, both of which have to be estimated from the data.
+For non-robust power transformation, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
+For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+
+### Anderson-Darling criterion
+
+The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$. We define this criterion as follows:
+
+```{=latex}
+\begin{equation}
+U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right) }
+\end{equation}
+```
+
+Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+For non-robust power transformations, all $w_i = 1$.
+Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include any extraneous weights $w_i$.
+The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.
+
+For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Anderson-Darling criterion.
+
+### Cramér-von Mises criterion
+
+The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
+```{=latex}
+\begin{equation}
+U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2}
+\end{equation}
+```
+
+Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+For non-robust power transformations, all $w_i = 1$.
+The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of extraneous weights.
+This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
+Unlike the Anderson-Darling criterion, this criterion weights all instances equally.
+
+For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
+
+## Skewness-kurtosis-based criteria
+
+The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
+We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
+Here, $\mu$ is defined as:
+
+```{=latex}
+\begin{equation}
+\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0} \left(x_i \right)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+The location, or mean, is weighted using weights $w_i$, as described in section **Insert**. For non-robust transformations, $w_i = 1$. Then, $sigma^2$ is defined as:
+```{=latex}
+\begin{equation}
+\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Skewness is defined as:
+```{=latex}
+\begin{equation}
+s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Kurtosis is defined as:
+```{=latex}
+\begin{equation}
+k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
+\end{equation}
+```
+
+### D'Agostino criterion
+
+The D'Agostino criterion defined here follows the D'Agostino $K^2$ test statistic [@DAgostino1990-kp].
+This test statistic is composed of two separate test statistics, one of which is related to skewness, and the other to kurtosis.
+Both test statistics are computed in several steps.
+Let us first define $\nu=\sum_{i=1}^n w_i$.
+Thus for non-robust power transformations, $\nu = n$.
+
+For the skewness test statistic we first compute [@DAgostino1990-kp]:
+
+```{=latex}
+\begin{equation}
+\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
+\end{equation}
+
+\begin{equation}
+\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
+\end{equation}
+
+\begin{equation}
+\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 1} - 2}}
+\end{equation}
+```
+
+The skewness test statistic is then:
+
+```{=latex}
+\begin{equation}
+Z_s^2 = \frac{2 \left[\log\left(\beta_1 / \alpha + \sqrt{\beta_1^2 / \alpha^2 + 1} \right) \right]^2}{\log\left(\sqrt{2 \beta_2 - 1} - 1 \right)}
+\end{equation}
+```
+
+For the kurtosis test statistic we first compute [@DAgostino1990-kp, @Anscombe1983-nz]:
+
+```{=latex}
+\begin{equation}
+\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
+\end{equation}
+
+\begin{equation}
+\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
+\end{equation}
+
+\begin{equation}
+\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
+\end{equation}
+
+\begin{equation}
+\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
+\end{equation}
+
+\begin{equation}
+\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
+\end{equation}
+```
+
+The kurtosis test statistic is then:
+
+```{=latex}
+\begin{equation}
+Z_k^2 = \frac{9 \alpha_1}{2} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]^2
+\end{equation}
+```
+
+The D'Agostino $K^2$ test statistic and our criterion are the same, and are defined as:
+```{=latex}
+\begin{equation}
+U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
+\end{equation}
+```
+
+The main difference between the test statistic as originally formulated, and the criterion proposed here is the presence of weights for robust power transformation.
+
+### Jarque-Bera criterion
+
+The second criterion based on skewness and kurtosis is the Jarque-Bera criterion. It is relatively simple to compute, compared to the D'Agostino criterion:
+```{=latex}
+\begin{equation}
+U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
+\end{equation}
+```
+
+The main difference between the above criterion and the Jarque-Bera test statistic [@Jarque1980-hw] is that a scalar multiplication is absent.
 

### 2023-02-22 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 # Non-standard evaluation
-d <- distribution <- method <- mean_lambda <- sd_lambda <- NULL
+d <- distribution <- method <- mean_lambda <- sd_lambda <- estimation_method <- NULL
 
@@
               for(robust in c(FALSE)){
-                # for(robust in c(FALSE, TRUE)){
-                # Only 
+                for(estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())){
+
                   # Skip if robust, but not shifted.
@@
                     "robust" = robust,
-                  "version" = version
+                    "version" = version,
+                    "estimation_method" = estimation_method
                   ))
@@
         }
-    })
+      }
+    }
+  )
 
@@
     # Create transformer object.
+    if(parameter_set$version == "original"){
       transformer <- suppressWarnings(
@@
         shift = parameter_set$shift,
+        estimation_method = parameter_set$estimation_method,
         lambda = NULL,
-        optimiser_control = list("xtol_rel"=1e-3)))
+        optimiser_control = list("xtol_rel"=1e-5)))
+
+    } else {
+      transformer <- suppressWarnings(
+      power.transform::find_transformation_parameters(
+        x = parameter_set$x,
+        method = parameter_set$method,
+        robust = parameter_set$robust,
+        shift = parameter_set$shift,
+        estimation_method = parameter_set$estimation_method))
+    }
 
@@
         "method" = parameter_set$method,
+        "estimation_method" = parameter_set$estimation_method,
         "version" = parameter_set$version,
@@
 
+  # data <- lapply(
+  #   X = experiments,
+  #   FUN = .compute_lambda)
+
   # Start cluster
@@
 }
+
+# Only evaluate the MLE method here.
+data <- data[estimation_method == "mle"]
 
@@
 
-For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Anderson-Darling criterion.
-
 ### Cramér-von Mises criterion
@@
 The main difference between the above criterion and the Jarque-Bera test statistic [@Jarque1980-hw] is that a scalar multiplication is absent.
+
+## Optimisation using non-MLE criteria
+
+Each of the criteria above can be used for optimisation, i.e.:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}, \hat{x}_0 \right\} = \argmin_{\lambda, x_0} U\left(\mathbf{X}, \lambda, x_0 \right)
+\end{equation}
+```
+
+For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, or equivalently, $x_0$ is fixed:
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0 \right)
+\end{equation}
+```
 

### 2023-02-22 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
+```{r shifted-distributions-other-criteria, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
+  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively,  which then underwent a shift $d$. ",
+  "Estimates of transformation parameter $\\lambda$ with the original power transformations (circle) show strong dependency on the overall location of the distribution, ",
+  "whereas estimates with the shift-sensitive power transformations (triangle) are constant.")
+
+# Lambda plot,
+.create_lambda_shift_plot <- function(data, plot_theme, limits, guide=FALSE){
+
+  p <- ggplot2::ggplot(
+    data = data,
+    mapping = ggplot2::aes(
+      x = d,
+      y = lambda,
+      colour = estimation_method))
+  p <- p + plot_theme
+  p <- p + ggplot2::geom_point()
+  p <- p + ggplot2::scale_x_continuous(
+    name = latex2exp::TeX("$d$"),
+    labels = scales::math_format())
+  p <- p + ggplot2::scale_y_continuous(
+    name = latex2exp::TeX("$\\lambda$"),
+    limits = limits)
+  p <- p + ggplot2::theme(
+    plot.title=ggplot2::element_text(
+      hjust=0.5,
+      size=ggplot2::rel(1.0)))
+
+  if(guide){
+    p <- p + paletteer::scale_color_paletteer_d(
+      palette = "ggthemes::Tableau_10",
+      drop = FALSE)
+    p <- p + guides(colour = ggplot2::guide_legend(title = "Optimisation criterion"))
+
+  } else {
+    p <- p + paletteer::scale_color_paletteer_d(
+      palette = "ggthemes::Tableau_10",
+      drop = FALSE,
+      guide = "none")
+  }
+
+  return(p)
+}
+
+if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
+  stop("Data are missing for creating shifted-distribution plots.")
+
+} else {
+  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
+}
+
+# Update distribution, method and version to factors.
+data$distribution <- factor(
+  x = data$distribution,
+  levels = c("normal", "right-skewed", "left-skewed"))
+data$method <- factor(
+  x = data$method,
+  levels = c("box_cox", "yeo_johnson"),
+  labels = c("Box-Cox", "Yeo-Johnson"))
+data$version <- factor(
+  x = data$version,
+  levels = c("original", "shift-sensitive"))
+data$estimation_method <- factor(
+  x = data$estimation_method,
+  levels = c("mle", "anderson_darling", "cramer_von_mises", "jarque_bera", "dagostino"),
+  labels = c("MLE", "Anderson-Darling", "Cramér-von Mises", "Jarque-Bera", "D'Agostino")
+)
+
+#### Normal distribution -------------------------------------------------------
+
+# Box-Cox-original
+p_bc_normal <- .create_lambda_shift_plot(
+  data = data[distribution == "normal" & method == "Box-Cox" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-20.0, 35.0))
+p_bc_normal <- p_bc_normal + ggplot2::ggtitle(label="normal distribution")
+
+# Yeo-Johnson
+p_yj_normal <- .create_lambda_shift_plot(
+  data=data[distribution == "normal" & method == "Yeo-Johnson" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-20.0, 35.0))
+
+#### Right skewed distribution -------------------------------------------------
+
+# Box-Cox
+p_bc_right <- .create_lambda_shift_plot(
+  data = data[distribution == "right-skewed" & method == "Box-Cox" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-15.0, 1.0))
+p_bc_right <- p_bc_right + ggplot2::ggtitle(label="right-skewed distribution")
+
+# Yeo-Johnson
+p_yj_right <- .create_lambda_shift_plot(
+  data = data[distribution == "right-skewed" & method == "Yeo-Johnson" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-15.0, 1.0))
+
+#### Left skewed distribution --------------------------------------------------
+
+# Box-Cox
+p_bc_left <- .create_lambda_shift_plot(
+  data = data[distribution == "left-skewed" & method == "Box-Cox" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-50.0, 60.0),
+  guide = TRUE)
+p_bc_left <- p_bc_left + ggplot2::ggtitle(label="left-skewed distribution")
+
+# Yeo-Johnson
+p_yj_left <- .create_lambda_shift_plot(
+  data = data[distribution == "left-skewed" & method == "Yeo-Johnson" & version == "original"],
+  plot_theme = plot_theme,
+  limits = c(-50.0, 60.0))
+
+p <- egg::ggarrange(
+  plots=list(
+    p_bc_normal, p_yj_normal,
+    p_bc_right, p_yj_right,
+    p_bc_left, p_yj_left),
+  ncol=3,
+  heights=c(1.0, 1.0),
+  byrow=FALSE,
+  draw=FALSE)
+p
+
+p <- ggplot2::ggplot(
+  data = data,
+  mapping = ggplot2::aes(
+    x = d,
+    y = lambda,
+    colour = estimation_method))
+p <- p + plot_theme
+p <- p + ggplot2::geom_point()
+p <- p + ggplot2::scale_x_continuous(
+  name = latex2exp::TeX("$d$"),
+  labels = scales::math_format())
+p <- p + ggplot2::scale_y_continuous(
+  name = latex2exp::TeX("$\\lambda$"))
+p <- p + paletteer::scale_color_paletteer_d(
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE)
+p <- p + guides(colour = ggplot2::guide_legend(title = "optimisation criterion"))
+
+p <- p + ggplot2::facet_grid(
+  method + version ~ distribution,
+  scales = "free")
+
+```
+
 ```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

### 2023-02-23 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 The double sum in right-hand side of the above equation can be considered as the loss that should be minimised.
-Minimisation is performed using the subplex algorithm implemented in the `nloptr` package for R [@Rowan1990-sn; @Ypma2022-tc].
+Minimisation is performed using the `sbplx` algorithm implemented in the `nloptr` package for R [@Rowan1990-sn; @Ypma2022-tc].
 
@@
 
-if(!file.exists(file.path(manuscript_dir, "robustness_comparison.RDS"))){
-  set.seed(95)
-
-  .compute_lambda <- function(x, method){
+# Generate experiments.
+experiment_args <- coro::generator(
+  function(
+    manuscript_dir,
+    transformation_methods = NULL,
+    estimation_methods = NULL){
+
+    if(is.null(transformation_methods)){
+      transformation_methods <- c("box_cox", "yeo_johnson")
+    }
+
+    if(is.null(estimation_methods)){
+      estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
+    }
+
+    for(estimation_method in estimation_methods){
+
+      for(transformation_method in transformation_methods){
+        # Non-robust transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = estimation_method,
+          "robust" = FALSE,
+          "shift" = TRUE)
+
+        opt_args <- list()
+
+        file_name <- file.path(
+          manuscript_dir,
+          paste0(
+            "robustness_comparison_",
+            transformation_method, "_",
+            estimation_method, "_", "non_robust.RDS"))
+
+        if(!file.exists(file_name)){
+          # Only yield arguments if the file does not exist.
+          yield(list(
+            "name" = "non-robust",
+            "method" = transformation_method,
+            "estimation_method" = estimation_method,
+            "file_name" = file_name,
+            "fun_args" = fun_args,
+            "opt_args" = opt_args))
+        } else {
+          yield(file_name)
+        }
+      }
+
+      for(robustness_source in c("empirical_probability", "transformed", "residual")){
+
+        def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
+
+        if(robustness_source == "empirical_probability"){
+          def_opt_limits$k1[2] <- 1.0
+          def_opt_limits$k2[2] <- 1.0
+
+          def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
+
+        } else if(robustness_source == "transformed"){
+          def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
+
+        } else if(robustness_source == "residual"){
+          def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
+        }
+
+        for(method in c("box_cox", "yeo_johnson")){
+          for(robustness_weighting_function in c("step", "triangle", "cosine")){
+
+            fun_args <- list(
+              "method" = transformation_method,
+              "estimation_method" = estimation_method,
+              "robust" = TRUE,
+              "shift" = TRUE,
+              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
+              "backup_use_default" = FALSE)
+
+            opt_limits <- def_opt_limits
+            opt_init <- def_opt_init
+
+            # Step-weighting only has k1 as a parameter.
+            if(robustness_weighting_function == "step"){
+              opt_limits$k2 <- NULL
+              opt_init$k2 <- NULL
+            }
+
+            file_name <- file.path(
+              manuscript_dir,
+              paste0(
+                "robustness_comparison_",
+                transformation_method, "_",
+                estimation_method, "_",
+                robustness_source, "_",
+                robustness_weighting_function, ".RDS"))
+
+            if(!file.exists(file_name)){
+              # Only yield arguments if the file does not exist.
+              yield(list(
+                "name" = paste0(robustness_source, "-", robustness_weighting_function),
+                "method" = transformation_method,
+                "estimation_method" = estimation_method,
+                "file_name" = file_name,
+                "fun_args" = fun_args,
+                "opt_args" = list("initial" = opt_init, "limits" = opt_limits)))
+
+            } else {
+              yield(file_name)
+            }
+          }
+        }
+      }
+    }
+  }
+)
+
+# Helper function for computing the target lambda value.
+.compute_target_lambda <- function(
+    x,
+    transformation_methods = NULL,
+    estimation_methods = NULL){
+
+  if(is.null(transformation_methods)){
+    transformation_methods <- c("box_cox", "yeo_johnson")
+  }
+
+  if(is.null(estimation_methods)){
+    estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
+  }
+
+  lambda_list <- list()
+  ii <- 1L
+  for(transformation_method in transformation_methods){
+    for(estimation_method in estimation_methods){
       # Create transformer object.
@@
           x = x,
-        method = method,
+          method = transformation_method,
+          estimation_method = estimation_method,
           robust = FALSE,
@@
 
-    return(transformer@lambda)
-  }
-
-  # Generate table of asymmetric generalised normal distribution parameters.
-  n_distributions <- 100
-
-  # Generate alpha, beta and n.
-  n <- stats::runif(n=n_distributions, min=2, max=4)
-  n <- ceiling(10^n)
-
-  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
-  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
-
-  # Generate corresponding distributions.
-  x <- mapply(
-    power.transform::ragn,
-    n = n,
-    alpha = alpha,
-    beta = beta)
-
-  # Compute lambda values without weighting for Box-Cox.
-  target_lambda_bc <- sapply(
+      lambda_list[[ii]] <- data.table::data.table(
+        "method" = transformation_method,
+        "estimation_method" = estimation_method,
+        "lambda" = transformer@lambda)
+
+      ii <- ii + 1L
+    }
+  }
+
+  return(data.table::rbindlist(lambda_list))
+}
+
+# Helper function for population distributions with outliers.
+.populate_outliers <- function(
     x,
-    .compute_lambda,
-    method = "box_cox")
-
-  # Compute lambda values without weighting for Yeo-Johnson
-  target_lambda_yj <- sapply(
-    x,
-    .compute_lambda,
-    method = "yeo_johnson")
-
-  # Add outliers, between 0.00 and 0.10.
-  n_outliers <- 10
-  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
-
-  # Find optimised k1 and k2 values for all combinations of weight sources and
-  # weighting functions.
-  x <- mapply(
-    function(x, n, alpha, beta, target_lambda_bc, target_lambda_yj, outlier_fraction){
-
-      parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda_bc"=target_lambda_bc, "target_lambda_yj"=target_lambda_yj)
-
+    n,
+    alpha,
+    beta,
+    target_lambda,
+    outlier_fraction){
+
+  # Set parameters.
+  parameters <- list(
+    "n" = n,
+    "alpha" = alpha,
+    "beta" = beta,
+    "target_lambda" = target_lambda)
+
+  # Add outliers.
   x <- lapply(
@@
   return(x)
-    },
-    x = x,
-    n = n,
-    alpha = alpha,
-    beta = beta,
-    target_lambda_bc = target_lambda_bc,
-    target_lambda_yj = target_lambda_yj,
-    MoreArgs=list("outlier_fraction"=outlier_fraction),
-    SIMPLIFY=FALSE,
-    USE.NAMES=FALSE)
-
-  x <- unlist(x, recursive = FALSE)
-
-  # Optimise k1, k2 so that sum of absolute differences with target-lambda is
-  # minimised for each combination of source and weighting method.
-  experiment_args <- coro::generator(
-    function(){
-
-      for(method in c("box_cox", "yeo_johnson")){
-        # Non-robust transformation.
-        fun_args <- list(
-          "method" = method,
-          "robust" = FALSE,
-          "shift" = TRUE)
-
-        opt_args <- list()
-
-        yield(list(
-          "name" = "non-robust",
-          "method" = method,
-          "fun_args" = fun_args,
-          "opt_args" = opt_args
-        ))
-      }
-
-      for(robustness_source in c("empirical_probability", "transformed", "residual")){
-
-        def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
-
-        if(robustness_source == "empirical_probability"){
-          def_opt_limits$k1[2] <- 1.0
-          def_opt_limits$k2[2] <- 1.0
-
-          def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
-
-        } else if(robustness_source == "transformed"){
-          def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
-
-        } else if(robustness_source == "residual"){
-          def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
-        }
-
-        for(method in c("box_cox", "yeo_johnson")){
-          for(robustness_weighting_function in c("step", "triangle", "cosine")){
-
-            fun_args <- list(
-              "method" = method,
-              "robust" = TRUE,
-              "shift" = TRUE,
-              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
-              "backup_use_default" = FALSE)
-
-            opt_limits <- def_opt_limits
-            opt_init <- def_opt_init
-
-            # Step-weighting only has k1 as a parameter.
-            if(robustness_weighting_function == "step"){
-              opt_limits$k2 <- NULL
-              opt_init$k2 <- NULL
-            }
-
-            yield(list(
-              "name" = paste0(robustness_source, "-", robustness_weighting_function),
-              "method" = method,
-              "fun_args" = fun_args,
-              "opt_args" = list("initial" = opt_init, "limits" = opt_limits)
-            ))
-          }
-        }
-      }
-    }
-  )
-
-  # Function for computing transformer parameters under optimisation constraints.
-  compute_robust_lambda <- function(
+}
+
+# Helper function for computing transformer parameters under optimisation
+# constraints.
+.compute_robust_lambda <- function(
     x,
     fun_args,
-    opt_args,
-    cl_internal=NULL){
+    opt_args){
 
@@
     fun_args,
-    opt_args
-    ){
+    opt_args){
     # Create transformer.
@@
 
+    target_lambda <- x$parameters$target_lambda[method == fun_args$method & estimation_method == fun_args$estimation_method]$lambda
+
     parameter_data <- c(
-        x$parameters,
       opt_args,
@@
         "method" = fun_args$method,
+        "estimation_method" = fun_args$estimation_method,
         "weight_method" = fun_args$weight_method,
         "lambda" = transformer@lambda,
+        "target_lambda" = target_lambda,
         "shift" = transformer@shift)
@@
 
-    if(is.null(cl_internal)){
   parameter_data <- lapply(
@@
 
-    } else {
-      parameter_data <- parallel::parLapply(
-        cl = cl_internal,
-        X = x,
-        fun = ..outlier_parser,
-        fun_args = fun_args,
-        opt_args = opt_args)
-    }
-
   return(parameter_data)
@@
 
-  # Inner optimisation function that sets the loss.
-  ..optimisation_function <- function(
-    opt_param,
-    x,
-    fun_args,
-    cl_internal = NULL){
-
-    # Parse options.
-    opt_args <- list()
-    if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
-    if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
-
-    parameter_data <- compute_robust_lambda(
-      x = x,
-      fun_args = fun_args,
-      opt_args = opt_args,
-      cl_internal = cl_internal)
-
-    if(fun_args$method == "box_cox"){
-      lambda_error <- sapply(
-        parameter_data,
-        function(x) (abs(x$lambda - x$target_lambda_bc)))
-
-    } else {
-      # Yeo-Johnson.
-      lambda_error <- sapply(
-        parameter_data,
-        function(x) (abs(x$lambda - x$target_lambda_yj)))
-    }
-    cat(paste0("target: ", mean(lambda_error), "; k1: ", opt_param[1], "; k2: ", opt_param[2], "\n"))
-    return(mean(lambda_error))
-  }
-
-  # Outer optimisation function
-  .optimisation_function <- function(
+# Outer optimisation function used to optimise the weighting function parameters
+# over a dataset.
+.optimisation_outer <- function(
     experiment,
     data,
-    cl_internal = NULL){
+    verbose = FALSE){
 
   set.seed(9)
+
+  results_list <- list(
+    "name" = experiment$name,
+    "method" = experiment$method,
+    "estimation_method" = experiment$estimation_method)
 
@@
       x0 = x0,
-        fn = ..optimisation_function,
+      fn = .optimisation_inner,
       lower = lower,
@@
       fun_args = experiment$fun_args,
-        cl_internal = cl_internal)
-
-      if(length(h$par) == 1){
-        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par, "value" = h$value))
+      verbose = verbose)
+
+    results_list <- c(results_list, list("k1" = h$par[1]))
+    if(length(h$par) > 1) results_list <- c(results_list, list("k2" = h$par[2]))
+    results_list <- c(results_list, list("value" = h$value))
 
   } else {
-        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
-      }
-
-    } else {
-      value <- ..optimisation_function(
+    value <- .optimisation_inner(
       opt_param=list(),
       x = data,
-        fun_args=experiment$fun_args,
-        cl = cl_internal)
-
-      return(list("name"=experiment$name, "method"=experiment$method, "value" = value))
-    }
-  }
+      fun_args=experiment$fun_args)
+
+    results_list <- c(results_list, list("value" = value))
+  }
+
+  # Save to file.
+  saveRDS(results_list, file=experiment$file_name)
+
+  return(invisible(TRUE))
+}
+
+# Inner optimisation function that sets the loss.
+.optimisation_inner <- function(
+    opt_param,
+    x,
+    fun_args,
+    verbose = FALSE){
+
+  # Parse options.
+  opt_args <- list()
+  if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
+  if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
+
+  parameter_data <- .compute_robust_lambda(
+    x = x,
+    fun_args = fun_args,
+    opt_args = opt_args)
+
+  lambda_error <- sapply(
+    parameter_data,
+    function(x) (abs(x$lambda - x$target_lambda)))
+
+  if(verbose){
+    message <- paste0("target: ", mean(lambda_error))
+    if(length(opt_args) > 0) message <- c(message, paste0("k1: ", opt_param[1]))
+    if(length(opt_args) > 1) message <- c(message, paste0("k2: ", opt_param[2]))
+    cat(paste0(message, collapse="; "))
+    cat("\n")
+  }
+
+  return(mean(lambda_error))
+}
+
+# Select experiments to conduct by filtering out experiments that are file
+# names.
+experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
+experiments <- experiments[!sapply(experiments, is.character)]
+
+if(length(experiments) > 0){
+  set.seed(95)
+
+  # Generate table of asymmetric generalised normal distribution parameters.
+  n_distributions <- 3
+
+  # Generate alpha, beta and n.
+  n <- stats::runif(n=n_distributions, min=2, max=4)
+  n <- ceiling(10^n)
+
+  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
+  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
+
+  # Generate corresponding distributions.
+  x <- mapply(
+    power.transform::ragn,
+    n = n,
+    alpha = alpha,
+    beta = beta)
+
+  # Compute lambda values without weighting.
+  target_lambda <- lapply(x, .compute_target_lambda)
+
+  # Add outliers, between 0.00 and 0.10.
+  n_outliers <- 10
+  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
+
+  # Generate outliers in the data.
+  x <- mapply(
+    FUN = .populate_outliers,
+    x = x,
+    n = n,
+    alpha = alpha,
+    beta = beta,
+    target_lambda = target_lambda,
+    MoreArgs=list("outlier_fraction"=outlier_fraction),
+    SIMPLIFY=FALSE,
+    USE.NAMES=FALSE)
+
+  x <- unlist(x, recursive = FALSE)
+
+  lapply(experiments, .optimisation_outer, data=x, verbose=TRUE)
 
@@
     cl=cl,
-    varlist = c("compute_robust_lambda", "..optimisation_function")
+    varlist = c(".compute_robust_lambda", ".optimisation_inner")
   )
 
-  experiment_results <- parallel::parLapplyLB(
+  parallel::parLapplyLB(
     cl = cl,
-    X = coro::collect(experiment_args()),
-    fun = .optimisation_function,
+    X = experiments,
+    fun = .optimisation_outer,
     data = x,
@@
   parallel::stopCluster(cl)
-
-  saveRDS(experiment_results, file = file.path(manuscript_dir, "robustness_comparison.RDS"))
-}
+}
+
+# Collect all files
+experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
+optimisation_data <- lapply(experiments, readRDS)
+optimisation_data <- lapply(optimisation_data, data.table::as.data.table)
+optimisation_data <- data.table::rbindlist(optimisation_data, fill=TRUE)
 ```
@@
         x0 = x0,
-        fn = ..optimisation_function,
+        fn = .optimisation_inner,
         lower = lower,

### 2023-02-23 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-        for(method in c("box_cox", "yeo_johnson")){
+        for(transformation_method in transformation_methods){
           for(robustness_weighting_function in c("step", "triangle", "cosine")){
@@
 
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  parallel::clusterExport(
-    cl=cl,
-    varlist = c(".compute_robust_lambda", ".optimisation_inner")
-  )
-
-  parallel::parLapplyLB(
-    cl = cl,
-    X = experiments,
-    fun = .optimisation_outer,
-    data = x,
-    chunk.size = 1L)
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
+  # 
+  # 
+  # # Start cluster
+  # cl <- parallel::makeCluster(18L)
+  # 
+  # parallel::clusterExport(
+  #   cl=cl,
+  #   varlist = c(".compute_robust_lambda", ".optimisation_inner")
+  # )
+  # 
+  # parallel::parLapplyLB(
+  #   cl = cl,
+  #   X = experiments,
+  #   fun = .optimisation_outer,
+  #   data = x,
+  #   chunk.size = 1L)
+  # 
+  # # Stop cluster.
+  # parallel::stopCluster(cl)
 }

### 2023-02-27 manuscript/manuscript.Rmd
More work on manuscript.

--- 2023-02-23 manuscript/manuscript.Rmd
@@
           manuscript_dir,
+          "robustness_comparison",
           paste0(
-            "robustness_comparison_",
             transformation_method, "_",
-            estimation_method, "_", "non_robust.RDS"))
+            estimation_method, "_",
+            "non_robust.RDS"))
 
@@
               manuscript_dir,
+              "robustness_comparison",
               paste0(
-                "robustness_comparison_",
                 transformation_method, "_",
@@
       upper = upper,
-      control=list("xtol_rel"=1e-2),
+      control=list("xtol_rel"=1e-2, "maxeval"=50),
       x = data,
@@
   # Generate table of asymmetric generalised normal distribution parameters.
-  n_distributions <- 3
+  n_distributions <- 100L
 
@@
 
-  lapply(experiments, .optimisation_outer, data=x, verbose=TRUE)
+  # lapply(experiments, .optimisation_outer, data=x, verbose=TRUE)
 
@@
 optimisation_data <- data.table::rbindlist(optimisation_data, fill=TRUE)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+# Compute the (marginal) effect of each parameter.
+
+# Generate experiments.
+experiment_args <- coro::generator(
+  function(
+    manuscript_dir,
+    transformation_methods = NULL,
+    estimation_methods = NULL){
+
+    if(is.null(transformation_methods)){
+      transformation_methods <- c("box_cox", "yeo_johnson")
+    }
+
+    if(is.null(estimation_methods)){
+      estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
+    }
+
+    for(estimation_method in estimation_methods){
+
+      for(transformation_method in transformation_methods){
+        # Non-robust transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = estimation_method,
+          "robust" = FALSE,
+          "shift" = TRUE)
+
+        file_name <- file.path(
+          manuscript_dir,
+          "robustness_comparison",
+          paste0(
+            transformation_method, "_",
+            estimation_method, "_",
+            "non_robust.RDS"))
+
+        # Only yield arguments if the file does not exist.
+        yield(list(
+          "name" = "non-robust",
+          "method" = transformation_method,
+          "estimation_method" = estimation_method,
+          "file_name" = file_name,
+          "fun_args" = fun_args))
+      }
+
+      for(robustness_source in c("empirical_probability", "transformed", "residual")){
+        for(transformation_method in transformation_methods){
+          for(robustness_weighting_function in c("step", "triangle", "cosine")){
+
+            file_name <- file.path(
+              manuscript_dir,
+              "robustness_comparison",
+              paste0(
+                transformation_method, "_",
+                estimation_method, "_",
+                robustness_source, "_",
+                robustness_weighting_function, ".RDS"))
+
+            optimisation_settings <- readRDS(file_name)
+            weighting_function_parameters <- list("k1"=optimisation_settings$k1)
+            if(!is.null(optimisation_settings$k2)){
+              weighting_function_parameters <- c(
+                weighting_function_parameters,
+                list("k2"=optimisation_settings$k2))
+            }
+
+            fun_args <- list(
+              "method" = transformation_method,
+              "estimation_method" = estimation_method,
+              "robust" = TRUE,
+              "shift" = TRUE,
+              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
+              "weighting_function_parameters" = weighting_function_parameters)
+
+            # Only yield arguments if the file does not exist.
+            yield(list(
+              "name" = paste0(robustness_source, "-", robustness_weighting_function),
+              "method" = transformation_method,
+              "estimation_method" = estimation_method,
+              "file_name" = file_name,
+              "fun_args" = fun_args))
+          }
+        }
+      }
+    }
+  }
+)
+
+# Helper function for computing the target lambda value.
+.compute_target_lambda <- function(
+    x,
+    transformation_methods = NULL,
+    estimation_methods = NULL){
+
+  if(is.null(transformation_methods)){
+    transformation_methods <- c("box_cox", "yeo_johnson")
+  }
+
+  if(is.null(estimation_methods)){
+    estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
+  }
+
+  lambda_list <- list()
+  ii <- 1L
+  for(transformation_method in transformation_methods){
+    for(estimation_method in estimation_methods){
+      # Create transformer object.
+      transformer <- suppressWarnings(
+        power.transform::find_transformation_parameters(
+          x = x,
+          method = transformation_method,
+          estimation_method = estimation_method,
+          robust = FALSE,
+          shift = TRUE))
+
+      lambda_list[[ii]] <- data.table::data.table(
+        "method" = transformation_method,
+        "estimation_method" = estimation_method,
+        "lambda" = transformer@lambda)
+
+      ii <- ii + 1L
+    }
+  }
+
+  return(data.table::rbindlist(lambda_list))
+}
+
+# Helper function for population distributions with outliers.
+.populate_outliers <- function(
+    x,
+    n,
+    alpha,
+    beta,
+    target_lambda,
+    outlier_fraction){
+
+  # Set parameters.
+  parameters <- list(
+    "n" = n,
+    "alpha" = alpha,
+    "beta" = beta,
+    "target_lambda" = target_lambda)
+
+  # Add outliers.
+  x <- lapply(
+    outlier_fraction,
+    function(k, x, parameters){
+      # Set parameters.
+      parameters$k <- k
+
+      # Compute interquartile range.
+      interquartile_range <- stats::IQR(x)
+
+      # Compute upper and lower quartiles.
+      q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+      q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+      # Set data where the outliers will be copied into.
+      x_outlier <- x
+
+      if(k != 0.0){
+        n_draw <- ceiling(k * length(x))
+
+        # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+        # than Q3 + 1.5 IQR.
+        x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+        outlier <- numeric(n_draw)
+        if(any(x_random < 0)){
+          outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+        }
+
+        if(any(x_random >= 0)){
+          outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+        }
+
+        # Randomly insert outlier values.
+        x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
+      }
+
+      return(list(
+        "x" = x_outlier,
+        "parameters" = parameters))
+    },
+    x = x,
+    parameters)
+
+  return(x)
+}
+
+# Helper function for computing transformer parameters under optimisation
+# constraints.
+.compute_robust_lambda <- function(
+    experiment,
+    data){
+
+  # Custom parser.
+  ..outlier_parser <- function(
+    x,
+    fun_args){
+    # Create transformer.
+    transformer <- suppressWarnings(do.call(
+      power.transform::find_transformation_parameters,
+      args=c(
+        list("x" = x$x),
+        fun_args)))
+
+    target_lambda <- x$parameters$target_lambda[method == fun_args$method & estimation_method == fun_args$estimation_method]$lambda
+
+    parameter_data <- list(
+      "method" = fun_args$method,
+      "estimation_method" = fun_args$estimation_method,
+      "weight_method" = fun_args$weight_method,
+      "lambda" = transformer@lambda,
+      "target_lambda" = target_lambda,
+      "shift" = transformer@shift)
+
+    return(parameter_data)
+  }
+
+  parameter_data <- lapply(
+    data,
+    ..outlier_parser,
+    fun_args = experiment$fun_args)
+
+  return(parameter_data)
+}
+
+set.seed(95)
+
+# Generate table of asymmetric generalised normal distribution parameters.
+n_distributions <- 100L
+
+# Generate alpha, beta and n.
+n <- stats::runif(n=n_distributions, min=2, max=4)
+n <- ceiling(10^n)
+
+alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
+beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
+
+# Generate corresponding distributions.
+x <- mapply(
+  power.transform::ragn,
+  n = n,
+  alpha = alpha,
+  beta = beta)
+
+# Compute lambda values without weighting.
+target_lambda <- lapply(x, .compute_target_lambda)
+
+# Add outliers, between 0.00 and 0.10.
+n_outliers <- 10
+outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
+
+# Generate outliers in the data.
+x <- mapply(
+  FUN = .populate_outliers,
+  x = x,
+  n = n,
+  alpha = alpha,
+  beta = beta,
+  target_lambda = target_lambda,
+  MoreArgs=list("outlier_fraction"=outlier_fraction),
+  SIMPLIFY=FALSE,
+  USE.NAMES=FALSE)
+
+x <- unlist(x, recursive = FALSE)
+
+experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
+
+cl <- parallel::makeCluster(18L)
+
+data <- parallel::parLapplyLB(
+  cl = cl,
+  X = experiments,
+  fun = .compute_robust_lambda,
+  data = x,
+  chunk.size = 1L)
+
+# Stop cluster.
+parallel::stopCluster(cl)
+
 ```

### 2023-02-27 manuscript/manuscript.Rmd
Moved some simulation-related code to simulations.R

--- 2023-02-23 manuscript/manuscript.Rmd
@@
 manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")
+source(file.path(manuscript_dir, "simulations.R"))
 
@@
 
-if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
-
-  # Generator for creating parameters and data for all experiments.
-  generate_experiment_data <- coro::generator(
-    function(
-    x_normal,
-    x_right_skewed,
-    x_left_skewed){
-
-      shift_range <- 10^seq(from=0, to=6, by=0.1)
-
-      for(d in shift_range){
-        for(distribution in c("normal", "right-skewed", "left-skewed")){
-          if(distribution == "normal"){
-            x <- x_normal
-
-          } else if(distribution == "right-skewed"){
-            x <- x_right_skewed
-
-          } else if(distribution == "left-skewed"){
-            x <- x_left_skewed
-          }
-
-          for(method in c("box_cox", "yeo_johnson")){
-            for(shift in c(FALSE, TRUE)){
-              for(robust in c(FALSE)){
-                for(estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())){
-
-                  # Skip if robust, but not shifted.
-                  if(robust && !shift) next
-
-                  if(!shift && !robust){
-                    version <- "original"
-
-                  } else if(!robust){
-                    version <- "shift-sensitive"
-
-                  } else {
-                    version <- "robust shift-sensitive"
-                  }
-
-                  yield(list(
-                    "x" = x + d,
-                    "d" = d,
-                    "distribution" = distribution,
-                    "method" = method,
-                    "shift" = shift,
-                    "robust" = robust,
-                    "version" = version,
-                    "estimation_method" = estimation_method
-                  ))
-                }
-              }
-            }
-          }
-        }
-      }
-    }
-  )
-
-  .compute_lambda <- function(parameter_set){
-    # Create transformer object.
-    if(parameter_set$version == "original"){
-      transformer <- suppressWarnings(
-      power.transform::find_transformation_parameters(
-        x = parameter_set$x,
-        method = parameter_set$method,
-        robust = parameter_set$robust,
-        shift = parameter_set$shift,
-        estimation_method = parameter_set$estimation_method,
-        lambda = NULL,
-        optimiser_control = list("xtol_rel"=1e-5)))
-
-    } else {
-      transformer <- suppressWarnings(
-      power.transform::find_transformation_parameters(
-        x = parameter_set$x,
-        method = parameter_set$method,
-        robust = parameter_set$robust,
-        shift = parameter_set$shift,
-        estimation_method = parameter_set$estimation_method))
-    }
-
-    return(
-      data.table::data.table(
-        "distribution" = parameter_set$distribution,
-        "method" = parameter_set$method,
-        "estimation_method" = parameter_set$estimation_method,
-        "version" = parameter_set$version,
-        "d" = log10(parameter_set$d),
-        "lambda" = transformer@lambda,
-        "x_0" = transformer@shift
-      ))
-  }
-
-  #### Computations --------------------------------------------------------------
-
-  # Generate all experiments.
-  experiments <- coro::collect(generate_experiment_data(
-    x_normal = x_normal,
-    x_right_skewed = x_right_skewed,
-    x_left_skewed = x_left_skewed
-  ))
-
-  # data <- lapply(
-  #   X = experiments,
-  #   FUN = .compute_lambda)
-
-  # Start cluster
-  cl <- parallel::makeCluster(16L)
-
-  # Compute all data in parallel.
-  data <- parallel::parLapply(
-    cl=cl,
-    X=experiments,
-    fun=.compute_lambda
-  )
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-
-  # Combine data into a single table.
-  data <- data.table::rbindlist(data)
-
-  # Save to file.
-  saveRDS(data, file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
-
-} else {
-  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
-}
-
-# Only evaluate the MLE method here.
-data <- data[estimation_method == "mle"]
-
-# Update distribution, method and version to factors.
-data$distribution <- factor(
-  x = data$distribution,
-  levels = c("normal", "right-skewed", "left-skewed"))
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-data$version <- factor(
-  x = data$version,
-  levels = c("original", "shift-sensitive"))
+# Load data.
+data <- .get_shifted_distribution_data(manuscript_dir=manuscript_dir)
+data[estimation_method == "MLE"]
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-
-# Generate experiments.
-experiment_args <- coro::generator(
-  function(
-    manuscript_dir,
-    transformation_methods = NULL,
-    estimation_methods = NULL){
-
-    if(is.null(transformation_methods)){
-      transformation_methods <- c("box_cox", "yeo_johnson")
-    }
-
-    if(is.null(estimation_methods)){
-      estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
-    }
-
-    for(estimation_method in estimation_methods){
-
-      for(transformation_method in transformation_methods){
-        # Non-robust transformation.
-        fun_args <- list(
-          "method" = transformation_method,
-          "estimation_method" = estimation_method,
-          "robust" = FALSE,
-          "shift" = TRUE)
-
-        opt_args <- list()
-
-        file_name <- file.path(
-          manuscript_dir,
-          paste0(
-            "robustness_comparison_",
-            transformation_method, "_",
-            estimation_method, "_", "non_robust.RDS"))
-
-        if(!file.exists(file_name)){
-          # Only yield arguments if the file does not exist.
-          yield(list(
-            "name" = "non-robust",
-            "method" = transformation_method,
-            "estimation_method" = estimation_method,
-            "file_name" = file_name,
-            "fun_args" = fun_args,
-            "opt_args" = opt_args))
-        } else {
-          yield(file_name)
-        }
-      }
-
-      for(robustness_source in c("empirical_probability", "transformed", "residual")){
-
-        def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
-
-        if(robustness_source == "empirical_probability"){
-          def_opt_limits$k1[2] <- 1.0
-          def_opt_limits$k2[2] <- 1.0
-
-          def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
-
-        } else if(robustness_source == "transformed"){
-          def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
-
-        } else if(robustness_source == "residual"){
-          def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
-        }
-
-        for(method in c("box_cox", "yeo_johnson")){
-          for(robustness_weighting_function in c("step", "triangle", "cosine")){
-
-            fun_args <- list(
-              "method" = transformation_method,
-              "estimation_method" = estimation_method,
-              "robust" = TRUE,
-              "shift" = TRUE,
-              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
-              "backup_use_default" = FALSE)
-
-            opt_limits <- def_opt_limits
-            opt_init <- def_opt_init
-
-            # Step-weighting only has k1 as a parameter.
-            if(robustness_weighting_function == "step"){
-              opt_limits$k2 <- NULL
-              opt_init$k2 <- NULL
-            }
-
-            file_name <- file.path(
-              manuscript_dir,
-              paste0(
-                "robustness_comparison_",
-                transformation_method, "_",
-                estimation_method, "_",
-                robustness_source, "_",
-                robustness_weighting_function, ".RDS"))
-
-            if(!file.exists(file_name)){
-              # Only yield arguments if the file does not exist.
-              yield(list(
-                "name" = paste0(robustness_source, "-", robustness_weighting_function),
-                "method" = transformation_method,
-                "estimation_method" = estimation_method,
-                "file_name" = file_name,
-                "fun_args" = fun_args,
-                "opt_args" = list("initial" = opt_init, "limits" = opt_limits)))
-
-            } else {
-              yield(file_name)
-            }
-          }
-        }
-      }
-    }
-  }
-)
-
-# Helper function for computing the target lambda value.
-.compute_target_lambda <- function(
-    x,
-    transformation_methods = NULL,
-    estimation_methods = NULL){
-
-  if(is.null(transformation_methods)){
-    transformation_methods <- c("box_cox", "yeo_johnson")
-  }
-
-  if(is.null(estimation_methods)){
-    estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
-  }
-
-  lambda_list <- list()
-  ii <- 1L
-  for(transformation_method in transformation_methods){
-    for(estimation_method in estimation_methods){
-      # Create transformer object.
-      transformer <- suppressWarnings(
-        power.transform::find_transformation_parameters(
-          x = x,
-          method = transformation_method,
-          estimation_method = estimation_method,
-          robust = FALSE,
-          shift = TRUE))
-
-      lambda_list[[ii]] <- data.table::data.table(
-        "method" = transformation_method,
-        "estimation_method" = estimation_method,
-        "lambda" = transformer@lambda)
-
-      ii <- ii + 1L
-    }
-  }
-
-  return(data.table::rbindlist(lambda_list))
-}
-
-# Helper function for population distributions with outliers.
-.populate_outliers <- function(
-    x,
-    n,
-    alpha,
-    beta,
-    target_lambda,
-    outlier_fraction){
-
-  # Set parameters.
-  parameters <- list(
-    "n" = n,
-    "alpha" = alpha,
-    "beta" = beta,
-    "target_lambda" = target_lambda)
-
-  # Add outliers.
-  x <- lapply(
-    outlier_fraction,
-    function(k, x, parameters){
-      # Set parameters.
-      parameters$k <- k
-
-      # Compute interquartile range.
-      interquartile_range <- stats::IQR(x)
-
-      # Compute upper and lower quartiles.
-      q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
-      q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
-
-      # Set data where the outliers will be copied into.
-      x_outlier <- x
-
-      if(k != 0.0){
-        n_draw <- ceiling(k * length(x))
-
-        # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-        # than Q3 + 1.5 IQR.
-        x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
-        outlier <- numeric(n_draw)
-        if(any(x_random < 0)){
-          outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
-        }
-
-        if(any(x_random >= 0)){
-          outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
-        }
-
-        # Randomly insert outlier values.
-        x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
-      }
-
-      return(list(
-        "x" = x_outlier,
-        "parameters" = parameters))
-    },
-    x = x,
-    parameters)
-
-  return(x)
-}
-
-# Helper function for computing transformer parameters under optimisation
-# constraints.
-.compute_robust_lambda <- function(
-    x,
-    fun_args,
-    opt_args){
-
-  # Custom parser.
-  ..outlier_parser <- function(
-    x,
-    fun_args,
-    opt_args){
-    # Create transformer.
-    transformer <- suppressWarnings(do.call(
-      power.transform::find_transformation_parameters,
-      args=c(
-        list("x" = x$x),
-        list("weighting_function_parameters" = opt_args),
-        fun_args)))
-
-    target_lambda <- x$parameters$target_lambda[method == fun_args$method & estimation_method == fun_args$estimation_method]$lambda
-
-    parameter_data <- c(
-      opt_args,
-      list(
-        "method" = fun_args$method,
-        "estimation_method" = fun_args$estimation_method,
-        "weight_method" = fun_args$weight_method,
-        "lambda" = transformer@lambda,
-        "target_lambda" = target_lambda,
-        "shift" = transformer@shift)
-    )
-
-    return(parameter_data)
-  }
-
-  parameter_data <- lapply(
-    x,
-    ..outlier_parser,
-    fun_args = fun_args,
-    opt_args = opt_args)
-
-  return(parameter_data)
-}
-
-# Outer optimisation function used to optimise the weighting function parameters
-# over a dataset.
-.optimisation_outer <- function(
-    experiment,
-    data,
-    verbose = FALSE){
-
-  set.seed(9)
-
-  results_list <- list(
-    "name" = experiment$name,
-    "method" = experiment$method,
-    "estimation_method" = experiment$estimation_method)
-
-  # Run optimiser.
-  if(length(experiment$opt_args) > 0){
-
-    x0 <- unlist(experiment$opt_args$initial)
-
-    lower <- experiment$opt_args$limits$k1[1]
-    upper <- experiment$opt_args$limits$k1[2]
-
-    if(length(experiment$opt_args$limits) > 1){
-      lower <- c(lower, experiment$opt_args$limits$k2[1])
-      upper <- c(upper, experiment$opt_args$limits$k2[2])
-    }
-
-    h <- nloptr::sbplx(
-      x0 = x0,
-      fn = .optimisation_inner,
-      lower = lower,
-      upper = upper,
-      control=list("xtol_rel"=1e-2),
-      x = data,
-      fun_args = experiment$fun_args,
-      verbose = verbose)
-
-    results_list <- c(results_list, list("k1" = h$par[1]))
-    if(length(h$par) > 1) results_list <- c(results_list, list("k2" = h$par[2]))
-    results_list <- c(results_list, list("value" = h$value))
-
-  } else {
-    value <- .optimisation_inner(
-      opt_param=list(),
-      x = data,
-      fun_args=experiment$fun_args)
-
-    results_list <- c(results_list, list("value" = value))
-  }
-
-  # Save to file.
-  saveRDS(results_list, file=experiment$file_name)
-
-  return(invisible(TRUE))
-}
-
-# Inner optimisation function that sets the loss.
-.optimisation_inner <- function(
-    opt_param,
-    x,
-    fun_args,
-    verbose = FALSE){
-
-  # Parse options.
-  opt_args <- list()
-  if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
-  if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
-
-  parameter_data <- .compute_robust_lambda(
-    x = x,
-    fun_args = fun_args,
-    opt_args = opt_args)
-
-  lambda_error <- sapply(
-    parameter_data,
-    function(x) (abs(x$lambda - x$target_lambda)))
-
-  if(verbose){
-    message <- paste0("target: ", mean(lambda_error))
-    if(length(opt_args) > 0) message <- c(message, paste0("k1: ", opt_param[1]))
-    if(length(opt_args) > 1) message <- c(message, paste0("k2: ", opt_param[2]))
-    cat(paste0(message, collapse="; "))
-    cat("\n")
-  }
-
-  return(mean(lambda_error))
-}
-
-# Select experiments to conduct by filtering out experiments that are file
-# names.
-experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
-experiments <- experiments[!sapply(experiments, is.character)]
-
-if(length(experiments) > 0){
-  set.seed(95)
-
-  # Generate table of asymmetric generalised normal distribution parameters.
-  n_distributions <- 3
-
-  # Generate alpha, beta and n.
-  n <- stats::runif(n=n_distributions, min=2, max=4)
-  n <- ceiling(10^n)
-
-  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
-  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
-
-  # Generate corresponding distributions.
-  x <- mapply(
-    power.transform::ragn,
-    n = n,
-    alpha = alpha,
-    beta = beta)
-
-  # Compute lambda values without weighting.
-  target_lambda <- lapply(x, .compute_target_lambda)
-
-  # Add outliers, between 0.00 and 0.10.
-  n_outliers <- 10
-  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
-
-  # Generate outliers in the data.
-  x <- mapply(
-    FUN = .populate_outliers,
-    x = x,
-    n = n,
-    alpha = alpha,
-    beta = beta,
-    target_lambda = target_lambda,
-    MoreArgs=list("outlier_fraction"=outlier_fraction),
-    SIMPLIFY=FALSE,
-    USE.NAMES=FALSE)
-
-  x <- unlist(x, recursive = FALSE)
-
-  lapply(experiments, .optimisation_outer, data=x, verbose=TRUE)
-
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  parallel::clusterExport(
-    cl=cl,
-    varlist = c(".compute_robust_lambda", ".optimisation_inner")
-  )
-
-  parallel::parLapplyLB(
-    cl = cl,
-    X = experiments,
-    fun = .optimisation_outer,
-    data = x,
-    chunk.size = 1L)
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-}
-
-# Collect all files
-experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
-optimisation_data <- lapply(experiments, readRDS)
-optimisation_data <- lapply(optimisation_data, data.table::as.data.table)
-optimisation_data <- data.table::rbindlist(optimisation_data, fill=TRUE)
+two_sided_function_parameters <- .get_optimised_weighting_function_parameters_two_sided(manuscript_dir = manuscript_dir)
+
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+two_sided_transformation_parameters <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+
 ```
@@
 
-if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
-  stop("Data are missing for creating shifted-distribution plots.")
-
-} else {
-  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
-}
-
-# Update distribution, method and version to factors.
-data$distribution <- factor(
-  x = data$distribution,
-  levels = c("normal", "right-skewed", "left-skewed"))
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-data$version <- factor(
-  x = data$version,
-  levels = c("original", "shift-sensitive"))
-data$estimation_method <- factor(
-  x = data$estimation_method,
-  levels = c("mle", "anderson_darling", "cramer_von_mises", "jarque_bera", "dagostino"),
-  labels = c("MLE", "Anderson-Darling", "Cramér-von Mises", "Jarque-Bera", "D'Agostino")
-)
+data <- .get_shifted_distribution_data(manuscript_dir = manuscript_dir)
 

### 2023-02-27 manuscript/simulations.R

@@ -1,0 +1,913 @@
+.get_shifted_distribution_data <- function(manuscript_dir){
+
+  # Set seed.
+  set.seed(19L)
+
+  if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
+
+    # Normal distribution.
+    x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
+
+    # Right skewed data
+    x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
+
+    # Left skewed data
+    x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
+
+
+    # generator ----------------------------------------------------------------
+    generate_experiment_data <- coro::generator(
+      function(
+    x_normal,
+    x_right_skewed,
+    x_left_skewed){
+
+        shift_range <- 10^seq(from=0, to=6, by=0.1)
+
+        for(d in shift_range){
+          for(distribution in c("normal", "right-skewed", "left-skewed")){
+            if(distribution == "normal"){
+              x <- x_normal
+
+            } else if(distribution == "right-skewed"){
+              x <- x_right_skewed
+
+            } else if(distribution == "left-skewed"){
+              x <- x_left_skewed
+            }
+
+            for(method in c("box_cox", "yeo_johnson")){
+              for(shift in c(FALSE, TRUE)){
+                for(robust in c(FALSE)){
+                  for(estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())){
+
+                    # Skip if robust, but not shifted.
+                    if(robust && !shift) next
+
+                    if(!shift && !robust){
+                      version <- "original"
+
+                    } else if(!robust){
+                      version <- "shift-sensitive"
+
+                    } else {
+                      version <- "robust shift-sensitive"
+                    }
+
+                    yield(list(
+                      "x" = x + d,
+                      "d" = d,
+                      "distribution" = distribution,
+                      "method" = method,
+                      "shift" = shift,
+                      "robust" = robust,
+                      "version" = version,
+                      "estimation_method" = estimation_method
+                    ))
+                  }
+                }
+              }
+            }
+          }
+        }
+      }
+    )
+
[…]

### 2023-03-01 manuscript/manuscript.Rmd
Moved plotting routines to separate script to maintain flow of manuscript better.

@@
 require(ggplot2)
+require(patchwork)
 require(data.table)
@@
 source(file.path(manuscript_dir, "simulations.R"))
+source(file.path(manuscript_dir, "plots.R"))
 
@@
 plot_theme <- ggplot2::theme_light(base_size=base_size)
-
-get_annotation_settings <- function(ggtheme=NULL){
-  # Import formatting settings from the provided ggtheme.
-
-  # Find the text size for the table. This is based on text sizes in the
-  # ggtheme.
-  fontsize <- ggtheme$text$size
-  fontsize_rel <- 1.0
-
-  # Attempt to base the text size on the general axis.text attribute.
-  if(!is.null(ggtheme$axis.text$size)){
-    if(inherits(ggtheme$axis.text$size, "rel")){
-      # Find the relative text size of axis text.
-      fontsize_rel <- as.numeric(ggtheme$axis.text$size)
-
-    } else {
-      # Set absolute text size.
-      fontsize <- ggtheme$axis.text$size
-      fontsize_rel <- 1.0
-    }
-  }
-
-  # Attempt to refine the text size using the axis.text.y attribute in
-  # particular.
-  if(!is.null(ggtheme$axis.text.y$size)){
-    if(inherits(ggtheme$axis.text.y$size, "rel")){
-      # Set relative text size of axis text
-      fontsize_rel <- as.numeric(ggtheme$axis.text.y$size)
-
-    } else {
-      # Set absolute text size.
-      fontsize <- as.numeric(ggtheme$axis.text.y$size)
-      fontsize_rel <- 1.0
-    }
-  }
-
-  # Update the text size using the magical ggplot2 point size (ggplot2:::.pt).
-  geom_text_size <- fontsize * fontsize_rel / 2.845276
-
-  # Obtain lineheight
-  lineheight <- ggtheme$text$lineheight
-  if(!is.null(ggtheme$axis.text$lineheight)) lineheight <- ggtheme$axis.text$lineheight
-  if(!is.null(ggtheme$axis.text.y$lineheight)) lineheight <- ggtheme$axis.text.y$lineheight
-
-  # Obtain family
-  fontfamily <- ggtheme$text$family
-  if(!is.null(ggtheme$axis.text$family)) fontfamily <- ggtheme$axis.text$family
-  if(!is.null(ggtheme$axis.text.y$family)) fontfamily <- ggtheme$axis.text.y$family
-  if(!is.null(ggtheme$axis.text.x$family)) fontfamily <- ggtheme$axis.text.x$family
-
-  # Obtain face
-  fontface <- ggtheme$text$face
-  if(!is.null(ggtheme$axis.text$face)) fontface <- ggtheme$axis.text$face
-  if(!is.null(ggtheme$axis.text.y$face)) fontface <- ggtheme$axis.text.y$face
-  if(!is.null(ggtheme$axis.text.x$face)) fontface <- ggtheme$axis.text.x$face
-
-  # Obtain colour
-  colour <- ggtheme$text$colour
-  if(!is.null(ggtheme$axis.text$colour)) colour <- ggtheme$axis.text$colour
-  if(!is.null(ggtheme$axis.text.y$colour)) colour <- ggtheme$axis.text.y$colour
-  if(!is.null(ggtheme$axis.text.x$colour)) colour <- ggtheme$axis.text.x$colour
-
-  return(list(
-    "geom_text_size"=geom_text_size,
-    "fontsize"=fontsize,
-    "fontsize_rel"=fontsize_rel,
-    "colour"=colour,
-    "family"=fontfamily,
-    "face"=fontface,
-    "lineheight"=lineheight))
-}
+plot_theme$strip.background <- ggplot2::element_blank()
+plot_theme$strip.text$colour <- NULL
+plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
 
@@
 
-compute_lambda <- function(mu, x, method){
-  # Create transformer object.
-  transformer <- suppressWarnings(
-    power.transform::find_transformation_parameters(
-      x = x + mu,
-      method = method,
-      robust = FALSE,
-      shift = FALSE,
-      lambda = NULL,
-      optimiser_control = list("xtol_rel"=1e-3)))
-
-  # Return lambda value.
-  return(transformer@lambda)
-}
-
-# Non-standard evaluation
-mu <- lambda <- method <- NULL
-
-# Set seed.
-set.seed(19L)
-
-x <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
-shift_range <- 10^seq(from=0, to=6, by=0.1)
-
-box_cox_values <- sapply(
-  shift_range,
-  compute_lambda,
-  x = x,
-  method = "box_cox")
-
-yeo_johnson_values <- sapply(
-  shift_range,
-  compute_lambda,
-  x = x,
-  method = "yeo_johnson")
-
-data <- data.table::data.table(
-  "mu" = log10(c(shift_range, shift_range)),
-  "lambda" = c(box_cox_values, yeo_johnson_values),
-  "method" = factor(
-    x=rep(
-      c("Box-Cox", "Yeo-Johnson"),
-      each = length(shift_range)),
-    levels=c("Box-Cox", "Yeo-Johnson")))
-
-annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
-
-p_bc <- ggplot2::ggplot(
-  data = data[method == "Box-Cox"],
-  mapping = ggplot2::aes(
-    x = mu,
-    y = lambda,
-    colour = method))
-p_bc <- p_bc + plot_theme
-p_bc <- p_bc + ggplot2::geom_point()
-p_bc <- p_bc + ggplot2::geom_hline(
-  yintercept = 1.0, 
-  linetype="longdash", 
-  colour="grey40")
-p_bc <- p_bc + ggplot2::annotate(
-  geom = "text",
-  x = 6,
-  y = 1,
-  label = "expected",
-  colour=annotation_settings$colour,
-  family=annotation_settings$family,
-  fontface=annotation_settings$face,
-  size=annotation_settings$geom_text_size,
-  vjust=-1.0,
-  hjust=1.0)
-p_bc <- p_bc + ggplot2::scale_x_continuous(
-  name = latex2exp::TeX("$\\mu$"),
-  labels = scales::math_format())
-p_bc <- p_bc + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"),
-  limits = c(0.0, 35.0))
-p_bc <- p_bc + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE,
-  guide="none")
-
-p_yj <- ggplot2::ggplot(
-  data = data[method == "Yeo-Johnson"],
-  mapping = ggplot2::aes(
-    x = mu,
-    y = lambda,
-    colour = method))
-p_yj <- p_yj + plot_theme
-p_yj <- p_yj + ggplot2::geom_point()
-p_yj <- p_yj + ggplot2::geom_hline(
-  yintercept = 1.0,
-  linetype="longdash",
-  colour="grey40")
-p_yj <- p_yj + ggplot2::annotate(
-  geom = "text",
-  x = 6,
-  y = 1,
-  label = "expected",
-  colour=annotation_settings$colour,
-  family=annotation_settings$family,
-  fontface=annotation_settings$face,
-  size=annotation_settings$geom_text_size,
-  vjust=-1.0,
-  hjust=1.0)
-p_yj <- p_yj + ggplot2::scale_x_continuous(
-  name = latex2exp::TeX("$\\mu$"),
-  labels = scales::math_format())
-p_yj <- p_yj + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"),
-  limits = c(0.0, 35.0))
-p_yj <- p_yj + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE)
-p_yj <- p_yj + ggplot2::theme(
-  axis.title.y=ggplot2::element_blank(),
-  axis.text.y=ggplot2::element_blank())
-
-p <- egg::ggarrange(plots=list(p_bc, p_yj), ncol=2, draw=FALSE)
-p
+.plot_reduced_normality(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
+
 ```
@@
 
-# Non-standard evaluation
-d <- distribution <- method <- mean_lambda <- sd_lambda <- estimation_method <- NULL
-
-# Density plot.
-.create_density_plot <- function(x, plot_theme, limits=c(-5, 5)){
-  p <- ggplot2::ggplot(
-    data = data.table::data.table("x"=x),
-    mapping = ggplot2::aes(
-      x = x))
-  p <- p + plot_theme
-  p <- p + ggplot2::geom_density()
-  p <- p + ggplot2::theme(
-    axis.title = ggplot2::element_blank(),
-    axis.line = ggplot2::element_blank(),
-    axis.ticks = ggplot2::element_blank(),
-    axis.text = ggplot2::element_blank(),
-    panel.grid = ggplot2::element_blank(),
-    panel.border = ggplot2::element_blank(),
-    plot.title=ggplot2::element_text(
-      hjust=0.5,
-      size=ggplot2::rel(1.0)))
-  p <- p + ggplot2::xlim(limits)
-
-  return(p)
-}
-
-# Lambda plot,
-.create_lambda_shift_plot <- function(data, plot_theme, limits, guide=FALSE){
-
-  p <- ggplot2::ggplot(
-    data = data,
-    mapping = ggplot2::aes(
-      x = d,
-      y = lambda,
-      colour = method,
-      shape = version))
-  p <- p + plot_theme
-  p <- p + ggplot2::geom_point()
-  p <- p + ggplot2::scale_x_continuous(
-    name = latex2exp::TeX("$d$"),
-    labels = scales::math_format())
-  p <- p + ggplot2::scale_y_continuous(
-    name = latex2exp::TeX("$\\lambda$"),
-    limits = limits)
-
-  if(guide){
-    p <- p + paletteer::scale_color_paletteer_d(
-      palette = "ggthemes::Tableau_10",
-      drop = FALSE)
-
-  } else {
-    p <- p + paletteer::scale_color_paletteer_d(
-      palette = "ggthemes::Tableau_10",
-      drop = FALSE,
-      guide = "none")
-    p <- p + ggplot2::scale_shape_discrete(guide="none")
-  }
-
-  return(p)
-}
-
-# Set seed.
-set.seed(19L)
-
-# Normal distribution.
-x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
-
-# Right skewed data
-x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
-
-# Left skewed data
-x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
-
-# Load data.
-data <- .get_shifted_distribution_data(manuscript_dir=manuscript_dir)
-data[estimation_method == "MLE"]
-
-#### Normal distribution -------------------------------------------------------
-# Density
-p_dens_normal <- .create_density_plot(
-  x = x_normal,
-  plot_theme = plot_theme)
-p_dens_normal <- p_dens_normal + ggplot2::ggtitle(label="normal distribution")
-
-# Box-Cox
-p_bc_normal <- .create_lambda_shift_plot(
-  data = data[distribution == "normal" & method == "Box-Cox"],
-  plot_theme = plot_theme,
-  limits = c(0.0, 35.0))
-
-# Yeo-Johnson
-p_yj_normal <- .create_lambda_shift_plot(
-  data=data[distribution == "normal" & method == "Yeo-Johnson"],
-  plot_theme = plot_theme,
-  limits = c(0.0, 35.0))
-
-#### Right skewed distribution -------------------------------------------------
-
-# Density
-p_dens_right <- .create_density_plot(
-  x = x_right_skewed,
-  plot_theme = plot_theme)
-p_dens_right <- p_dens_right + ggplot2::ggtitle(label="right-skewed distribution")
-
-# Box-Cox
-p_bc_right <- .create_lambda_shift_plot(
-  data = data[distribution == "right-skewed" & method == "Box-Cox"],
-  plot_theme = plot_theme,
-  limits = c(-9.0, 1.0))
-
-# Yeo-Johnson
-p_yj_right <- .create_lambda_shift_plot(
-  data = data[distribution == "right-skewed" & method == "Yeo-Johnson"],
-  plot_theme = plot_theme,
-  limits = c(-9.0, 1.0))
-
-#### Left skewed distribution --------------------------------------------------
-
-# Density
-p_dens_left <- .create_density_plot(
-  x = x_left_skewed,
-  plot_theme = plot_theme)
-p_dens_left <- p_dens_left + ggplot2::ggtitle(label="left-skewed distribution")
-
-# Box-Cox
-p_bc_left <- .create_lambda_shift_plot(
-  data = data[distribution == "left-skewed" & method == "Box-Cox"],
-  plot_theme = plot_theme,
-  limits = c(0.0, 60.0),
-  guide = TRUE)
-
-# Yeo-Johnson
-p_yj_left <- .create_lambda_shift_plot(
-  data = data[distribution == "left-skewed" & method == "Yeo-Johnson"],
-  plot_theme = plot_theme,
-  limits = c(0.0, 60.0))
-
-#### Aggregate -----------------------------------------------------------------
-
-aggregate_data <- data[, list("mean_lambda"=mean(lambda), "sd_lambda"=stats::sd(lambda)), by=c("method", "version", "distribution")]
-aggregate_data$distribution <- factor(
-  x=aggregate_data$distribution,
-  levels = c("normal", "left-skewed", "right-skewed"))
-aggregate_data[, "text":=parse_mean_sd_latex(mu=mean_lambda, sigma=sd_lambda)]
-
-p <- egg::ggarrange(
-  plots=list(
-    p_dens_normal, p_bc_normal, p_yj_normal,
-    p_dens_right, p_bc_right, p_yj_right,
-    p_dens_left, p_bc_left, p_yj_left),
-  ncol=3,
-  heights=c(0.5, 1.0, 1.0),
-  byrow=FALSE,
-  draw=FALSE)
-p
+.plot_shifted_distributions(manuscript_dir = manuscript_dir, plot_theme = plot_theme)
 ```
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+```{r optimal_criteria, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Things."
+)
+
+# This find the optimal weighting parameter settings that are used within the
+# package.
 two_sided_function_parameters <- .get_optimised_weighting_function_parameters_two_sided(manuscript_dir = manuscript_dir)
 
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-two_sided_transformation_parameters <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+# This finds lambda values at the optimised weighting parameter settings.
+data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+
+data <- data[estimation_method == "MLE"]
+data <- data[, "lambda_error" := abs(lambda - target_lambda)]
+non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
+data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
+
+data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
+data[k == 0.0, "better_than_non_robust" := NA_integer_]
+data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
+
+pass_rate_data <- data[, list(
+  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
+  "n" = sum(!is.na(better_than_non_robust)),
+  "median_error" = stats::median(lambda_error)),
+  by=c("method", "estimation_method", "weight_method")]
+pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)][order(pass_rate)]
+
+p <- ggplot2::ggplot(
+  data = data,
+  mapping = ggplot2::aes(
+    x = weight_method,
+    y = lambda_error,
+    fill = method))
+p <- p + plot_theme
+p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
+p <- p + ggplot2::facet_wrap("method", nrow=2)
+p <- p + paletteer::scale_fill_paletteer_d(
+  palette = "ggthemes::Tableau_10",
+  drop = FALSE)
+p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
+p <- p + ggplot2::xlab("weighting method")
+p <- p + ggplot2::theme(
+  axis.text.x = ggplot2::element_text(
+    angle=60.0,
+    hjust=1.0))
 
@@
 
-Each of the criteria above can be used for optimisation, i.e.:
+Each of the criteria described in the previous section can be used for optimisation, i.e.:
 
@@
 ```
+
+# Appendix B: Simulations with other optimisation criteria
 
@@
 
-# Lambda plot,
-.create_lambda_shift_plot <- function(data, plot_theme, limits, guide=FALSE){
+  .plot_shifted_distributions_other_criteria(manuscript_dir = manuscript_dir, plot_theme = plot_theme)
+```
+
+```{r optimal_weighting_function_all_criteria_box_cox, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Things."
+)
+# This finds lambda values at the optimised weighting parameter settings.
+data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+
+data <- data[method == "Box-Cox"]
+data <- data[, "lambda_error" := abs(lambda - target_lambda)]
+non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
+data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
+
+data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
+data[k == 0.0, "better_than_non_robust" := NA_integer_]
+data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
+
+pass_rate_data <- data[, list(
+  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
+  "n" = sum(!is.na(better_than_non_robust)),
+  "median_error" = stats::median(lambda_error)),
+  by=c("method", "estimation_method", "weight_method")]
+pass_rate_data <- pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)]
 
@@
   mapping = ggplot2::aes(
-      x = d,
-      y = lambda,
-      colour = estimation_method))
+    x = weight_method,
+    y = lambda_error,
+    fill = estimation_method))
 p <- p + plot_theme
-  p <- p + ggplot2::geom_point()
-  p <- p + ggplot2::scale_x_continuous(
-    name = latex2exp::TeX("$d$"),
-    labels = scales::math_format())
-  p <- p + ggplot2::scale_y_continuous(
-    name = latex2exp::TeX("$\\lambda$"),
-    limits = limits)
-  p <- p + ggplot2::theme(
-    plot.title=ggplot2::element_text(
-      hjust=0.5,
-      size=ggplot2::rel(1.0)))
-
-  if(guide){
-    p <- p + paletteer::scale_color_paletteer_d(
+p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
+p <- p + ggplot2::facet_wrap("estimation_method", nrow=nlevels(data$estimation_method))
+p <- p + paletteer::scale_fill_paletteer_d(
   palette = "ggthemes::Tableau_10",
   drop = FALSE)
-    p <- p + guides(colour = ggplot2::guide_legend(title = "Optimisation criterion"))
-
-  } else {
-    p <- p + paletteer::scale_color_paletteer_d(
-      palette = "ggthemes::Tableau_10",
-      drop = FALSE,
-      guide = "none")
-  }
-
-  return(p)
-}
-
-data <- .get_shifted_distribution_data(manuscript_dir = manuscript_dir)
-
-#### Normal distribution -------------------------------------------------------
-
-# Box-Cox-original
-p_bc_normal <- .create_lambda_shift_plot(
-  data = data[distribution == "normal" & method == "Box-Cox" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-20.0, 35.0))
-p_bc_normal <- p_bc_normal + ggplot2::ggtitle(label="normal distribution")
-
-# Yeo-Johnson
-p_yj_normal <- .create_lambda_shift_plot(
-  data=data[distribution == "normal" & method == "Yeo-Johnson" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-20.0, 35.0))
-
-#### Right skewed distribution -------------------------------------------------
-
-# Box-Cox
-p_bc_right <- .create_lambda_shift_plot(
-  data = data[distribution == "right-skewed" & method == "Box-Cox" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-15.0, 1.0))
-p_bc_right <- p_bc_right + ggplot2::ggtitle(label="right-skewed distribution")
-
-# Yeo-Johnson
-p_yj_right <- .create_lambda_shift_plot(
-  data = data[distribution == "right-skewed" & method == "Yeo-Johnson" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-15.0, 1.0))
-
-#### Left skewed distribution --------------------------------------------------
-
-# Box-Cox
-p_bc_left <- .create_lambda_shift_plot(
-  data = data[distribution == "left-skewed" & method == "Box-Cox" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-50.0, 60.0),
-  guide = TRUE)
-p_bc_left <- p_bc_left + ggplot2::ggtitle(label="left-skewed distribution")
-
-# Yeo-Johnson
-p_yj_left <- .create_lambda_shift_plot(
-  data = data[distribution == "left-skewed" & method == "Yeo-Johnson" & version == "original"],
-  plot_theme = plot_theme,
-  limits = c(-50.0, 60.0))
-
-p <- egg::ggarrange(
-  plots=list(
-    p_bc_normal, p_yj_normal,
-    p_bc_right, p_yj_right,
-    p_bc_left, p_yj_left),
-  ncol=3,
-  heights=c(1.0, 1.0),
-  byrow=FALSE,
-  draw=FALSE)
+p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
+p <- p + ggplot2::xlab("weighting method")
+p <- p + ggplot2::ggtitle("shift-sensitive Box-Cox transformation")
+p <- p + ggplot2::theme(
+  axis.text.x = ggplot2::element_text(
+    angle=30.0,
+    hjust=1.0))
 p
+```
+
+```{r optimal_weighting_function_all_criteria_yeo_johnson, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Things."
+)
+# This finds lambda values at the optimised weighting parameter settings.
+data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+
+data <- data[method == "Yeo-Johnson"]
+data <- data[, "lambda_error" := abs(lambda - target_lambda)]
+non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
+data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
+
+data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
+data[k == 0.0, "better_than_non_robust" := NA_integer_]
+data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
+
+pass_rate_data <- data[, list(
+  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
+  "n" = sum(!is.na(better_than_non_robust)),
+  "median_error" = stats::median(lambda_error)),
+  by=c("method", "estimation_method", "weight_method")]
+pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)][order(pass_rate)]
 
@@
   mapping = ggplot2::aes(
-    x = d,
-    y = lambda,
-    colour = estimation_method))
+    x = weight_method,
+    y = lambda_error,
+    fill = estimation_method))
 p <- p + plot_theme
-p <- p + ggplot2::geom_point()
-p <- p + ggplot2::scale_x_continuous(
-  name = latex2exp::TeX("$d$"),
-  labels = scales::math_format())
-p <- p + ggplot2::scale_y_continuous(
-  name = latex2exp::TeX("$\\lambda$"))
-p <- p + paletteer::scale_color_paletteer_d(
+p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
+p <- p + ggplot2::facet_wrap("estimation_method", nrow=nlevels(data$estimation_method))
+p <- p + paletteer::scale_fill_paletteer_d(
   palette = "ggthemes::Tableau_10",
   drop = FALSE)
-p <- p + guides(colour = ggplot2::guide_legend(title = "optimisation criterion"))
-
-p <- p + ggplot2::facet_grid(
-  method + version ~ distribution,
-  scales = "free")
+p <- p + guides(fill = ggplot2::guide_legend(title = "optimisation criterion"))
+p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
+p <- p + ggplot2::xlab("weighting method")
+p <- p + ggplot2::ggtitle("shift-sensitive Yeo-Johnson transformation")
+p <- p + ggplot2::theme(
+  axis.text.x = ggplot2::element_text(
+    angle=30.0,
+    hjust=1.0))
+p
 

### 2023-03-01 manuscript/plots.R

@@ -1,0 +1,552 @@
+get_annotation_settings <- function(ggtheme = NULL) {
+  # Import formatting settings from the provided ggtheme.
+
+  # Find the text size for the table. This is based on text sizes in the
+  # ggtheme.
+  fontsize <- ggtheme$text$size
+  fontsize_rel <- 1.0
+
+  # Attempt to base the text size on the general axis.text attribute.
+  if (!is.null(ggtheme$axis.text$size)) {
+    if (inherits(ggtheme$axis.text$size, "rel")) {
+      # Find the relative text size of axis text.
+      fontsize_rel <- as.numeric(ggtheme$axis.text$size)
+
+    } else {
+      # Set absolute text size.
+      fontsize <- ggtheme$axis.text$size
+      fontsize_rel <- 1.0
+    }
+  }
+
+  # Attempt to refine the text size using the axis.text.y attribute in
+  # particular.
+  if (!is.null(ggtheme$axis.text.y$size)) {
+    if (inherits(ggtheme$axis.text.y$size, "rel")) {
+      # Set relative text size of axis text
+      fontsize_rel <- as.numeric(ggtheme$axis.text.y$size)
+
+    } else {
+      # Set absolute text size.
+      fontsize <- as.numeric(ggtheme$axis.text.y$size)
+      fontsize_rel <- 1.0
+    }
+  }
+
+  # Update the text size using the magical ggplot2 point size (ggplot2:::.pt).
+  geom_text_size <- fontsize * fontsize_rel / 2.845276
+
+  # Obtain lineheight
+  lineheight <- ggtheme$text$lineheight
+  if (!is.null(ggtheme$axis.text$lineheight)) lineheight <- ggtheme$axis.text$lineheight
+  if (!is.null(ggtheme$axis.text.y$lineheight)) lineheight <- ggtheme$axis.text.y$lineheight
+
+  # Obtain family
+  fontfamily <- ggtheme$text$family
+  if (!is.null(ggtheme$axis.text$family)) fontfamily <- ggtheme$axis.text$family
+  if (!is.null(ggtheme$axis.text.y$family)) fontfamily <- ggtheme$axis.text.y$family
+  if (!is.null(ggtheme$axis.text.x$family)) fontfamily <- ggtheme$axis.text.x$family
+
+  # Obtain face
+  fontface <- ggtheme$text$face
+  if (!is.null(ggtheme$axis.text$face)) fontface <- ggtheme$axis.text$face
[…]

### 2023-03-02 manuscript/manuscript.Rmd
Moved weighting function plot to plots.R

@@
 
-# Non-standard evaluation
-d <- distribution <- method <- mean_lambda <- sd_lambda <- weight <- NULL
-
-data_step <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_step[, "weight":=power.transform:::..weight_function_external(x, "step", k1=0.60)]
-
-p_step <- ggplot2::ggplot(
-  data = data_step,
-  mapping = ggplot2::aes(
-    x = x,
-    y = weight))
-p_step <- p_step + plot_theme
-p_step <- p_step + ggplot2::geom_line()
-p_step <- p_step + ggplot2::ggtitle(label="step")
-p_step <- p_step + ggplot2::theme(
-  plot.title=ggplot2::element_text(
-    hjust=0.5,
-    size=ggplot2::rel(1.0)))
-
-data_triangle <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_triangle[, "weight":=power.transform:::..weight_function_external(x, "triangle", k1=0.30, k2=0.90)]
-
-p_triangle <- ggplot2::ggplot(
-  data = data_triangle,
-  mapping = ggplot2::aes(
-    x = x,
-    y = weight))
-p_triangle <- p_triangle + plot_theme
-p_triangle <- p_triangle + ggplot2::geom_line()
-p_triangle <- p_triangle + ggplot2::ggtitle(label="triangle")
-p_triangle <- p_triangle + ggplot2::theme(
-  plot.title=ggplot2::element_text(
-    hjust=0.5,
-    size=ggplot2::rel(1.0)),
-  axis.text.y = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank())
-
-data_cosine <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
-data_cosine[, "weight":=power.transform:::..weight_function_external(x, "cosine", k1=0.30, k2=0.90)]
-
-p_cosine <- ggplot2::ggplot(
-  data = data_cosine,
-  mapping = ggplot2::aes(
-    x = x,
-    y = weight))
-p_cosine <- p_cosine + plot_theme
-p_cosine <- p_cosine + ggplot2::geom_line()
-p_cosine <- p_cosine + ggplot2::ggtitle(label="tapered cosine")
-p_cosine <- p_cosine + ggplot2::theme(
-  plot.title = ggplot2::element_text(
-    hjust = 0.5,
-    size = ggplot2::rel(1.0)),
-  axis.text.y = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank())
-
-p <- egg::ggarrange(plots=list(p_step, p_triangle, p_cosine), ncol=3, draw=FALSE)
-p
+.plot_weighting_functions(plot_theme = plot_theme)
 ```

### 2023-03-02 manuscript/plots.R

@@ -550,3 +550,60 @@ get_annotation_settings <- function(ggtheme = NULL) {
   return(p)
 }
 
+
+
+.plot_weighting_functions <- function(plot_theme){
+  # Non-standard evaluation
+  weight <- NULL
+
+  # Step function --------------------------------------------------------------
+  data_step <- data.table::data.table("x" = seq(from = -1.0, to = 1.0, by = 0.001))
+  data_step[, "weight" := power.transform:::..weight_function_external(x, "step", k1 = 0.60)]
+
+  p_step <- ggplot2::ggplot(
+    data = data_step,
+    mapping = ggplot2::aes(
+      x = x,
+      y = weight))
+  p_step <- p_step + plot_theme
+  p_step <- p_step + ggplot2::geom_line()
+  p_step <- p_step + ggplot2::facet_grid(cols = vars("step"))
+
+  # Triangle function ----------------------------------------------------------
+  data_triangle <- data.table::data.table("x" = seq(from = -1.0, to = 1.0, by = 0.001))
+  data_triangle[, "weight" := power.transform:::..weight_function_external(x, "triangle", k1 = 0.30, k2 = 0.90)]
+
+  p_triangle <- ggplot2::ggplot(
+    data = data_triangle,
+    mapping = ggplot2::aes(
+      x = x,
+      y = weight))
+  p_triangle <- p_triangle + plot_theme
+  p_triangle <- p_triangle + ggplot2::geom_line()
+  p_triangle <- p_triangle + ggplot2::facet_grid(cols = vars("triangle"))
+  p_triangle <- p_triangle + ggplot2::theme(
+    axis.text.y = ggplot2::element_blank(),
+    axis.ticks.y = ggplot2::element_blank(),
+    axis.title.y = ggplot2::element_blank())
+
+  # Cosine function ------------------------------------------------------------
+  data_cosine <- data.table::data.table(x = seq(from = -1.0, to = 1.0, by = 0.001))
+  data_cosine[, "weight" := power.transform:::..weight_function_external(x, "cosine", k1 = 0.30, k2 = 0.90)]
+
+  p_cosine <- ggplot2::ggplot(
+    data = data_cosine,
+    mapping = ggplot2::aes(
+      x = x,
+      y = weight))
+  p_cosine <- p_cosine + plot_theme
+  p_cosine <- p_cosine + ggplot2::geom_line()
[…]

### 2023-03-02 manuscript/manuscript.Rmd
Updated robust optimisation comparison.

@@
 ```
+Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **insert**.
 
@@
 
-The result is shown in Figure **INSERT**, and summarised in Table **INSERT**.
+The result is shown in Figure **INSERT**.
 For each distribution, transformation parameter $\hat{\lambda}$ varies with $d$ when estimated using the original definitions.
 In contrast, estimation of $\lambda$ using shift-sensitive power transformations was invariant to $d$.
-
-```{r, results="asis"}
-table_data <- data.table::dcast(
-  aggregate_data,
-  distribution ~ version + method,
-  value.var = "text",
-  sep = " ")
-
-data.table::setnames(
-  table_data,
-  old = c("original Box-Cox", "original Yeo-Johnson", "shift-sensitive Box-Cox", "shift-sensitive Yeo-Johnson"),
-  new = c("orig. Box-Cox", "orig. Yeo-Johnson", "shift-sens. Box-Cox", "shift-sens. Yeo-Johnson"))
-
-knitr::kable(
-  table_data,
-  align = c("l", "c", "c", "c", "c"),
-  caption = paste0(
-    "Estimates and variability of transformation parameter $\\lambda$ under location shifts. ",
-    "Whereas the original (orig.) power transformations show considerable variability, ",
-    "whereas for shift-sensitive (shift-sens.) power transformations variability in the value of $\\lambda$ is negligible."),
-  label = "shifted-distributions-table",
-  format = "latex",
-  escape = FALSE,
-  booktabs = TRUE)
-```
 
@@
 
-```{r optimal_criteria, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Things."
-)
-
-# This find the optimal weighting parameter settings that are used within the
-# package.
-two_sided_function_parameters <- .get_optimised_weighting_function_parameters_two_sided(manuscript_dir = manuscript_dir)
-
-# This finds lambda values at the optimised weighting parameter settings.
-data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
-
-data <- data[estimation_method == "MLE"]
-data <- data[, "lambda_error" := abs(lambda - target_lambda)]
-non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
-data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
-
-data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
-data[k == 0.0, "better_than_non_robust" := NA_integer_]
-data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
-
-pass_rate_data <- data[, list(
-  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
-  "n" = sum(!is.na(better_than_non_robust)),
-  "median_error" = stats::median(lambda_error)),
-  by=c("method", "estimation_method", "weight_method")]
-pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)][order(pass_rate)]
-
-p <- ggplot2::ggplot(
-  data = data,
-  mapping = ggplot2::aes(
-    x = weight_method,
-    y = lambda_error,
-    fill = method))
-p <- p + plot_theme
-p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
-p <- p + ggplot2::facet_wrap("method", nrow=2)
-p <- p + paletteer::scale_fill_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE)
-p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
-p <- p + ggplot2::xlab("weighting method")
-p <- p + ggplot2::theme(
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-```
+The optimal weighting function parameters for weighted MLE are shown in Tables **insert** and **insert** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 783 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.84 & ---  & ---       & ---  & 587 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.78 & 0.95 & $(0, 1]$  & 0.99 & 587 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.98 & 583 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 1135 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.26 & 1.96 & $(0, 10]$ & 5.47 & 1813 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.03 & 1.96 & $(0, 10]$ & 6.91 & 1844 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.56 & ---  & ---       & ---  & 1837 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.49 & 1.00 & $(0, 10]$ & 1.51 & 1877 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.32 & 1.00 & $(0, 10]$ & 1.65 & 1834 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 773 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.84 & ---  & ---       & ---  & 598 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.79 & 0.95 & $(0, 1]$  & 0.98 & 594 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.77 & 0.95 & $(0, 1]$  & 1.00 & 580 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.21 & ---  & ---       & ---  & 1214 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.05 & 1.96 & $(0, 10]$ & 4.47 & 1052 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.09 & 1.96 & $(0, 10]$ & 5.30 & 1086 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.36 & ---  & ---       & ---  & 1483 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.43 & 1.00 & $(0, 10]$ & 1.45 & 1498 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.20 & 1.00 & $(0, 10]$ & 1.48 & 1467 \\
 \bottomrule
@@
 \midrule
-non-robust               & ---  & ---       & ---   & ---  & ---       & ---   & 225 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.92  & ---  & ---       & ---   & 125 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88  & 0.95 & $(0, 1]$  & 0.93  & 126 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.89  & 0.95 & $(0, 1]$  & 0.93  & 124 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.04  & ---  & ---       & ---   & 333 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.12  & 1.96 & $(0, 10]$ & 9.91  & 620 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.23  & 1.96 & $(0, 10]$ & 5.63  & 716 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 10.00 & ---  & ---       & ---   & 485 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 9.96  & 1.00 & $(0, 10]$ & 10.00 & 483 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 9.97  & 1.00 & $(0, 10]$ & 10.00 & 484 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 305 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.87 & ---  & ---       & ---  & 136 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.88 & 133 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.81 & 0.95 & $(0, 1]$  & 0.88 & 133 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.07 & ---  & ---       & ---  & 333 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.00 & 1.96 & $(0, 10]$ & 4.57 & 227 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.55 & 1.96 & $(0, 10]$ & 5.16 & 236 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.57 & ---  & ---       & ---  & 631 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.28 & 1.00 & $(0, 10]$ & 1.81 & 562 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.32 & 1.00 & $(0, 10]$ & 1.75 & 567 \\
 \bottomrule
@@
 
-Table **INSERT** shows weighting parameters and corresponding loss, after optimisation.
-Loss without any mitigation of outliers was $225$.
-Only weighting based on empirical probabilities ($p^(*)$) achieved better loss under optimised weighting parameters.
-For these data, all three window functions achieved a similar loss.
-Weighting based on the z-score of the transformed feature, or on their residual error, performed worse than non-robust shift-sensitive power transformation.
-Thus weighting of outliers should not be based on either of these two options.
-The likely reason for their poor performance is that they allow for a direct interaction with estimation of the $\lambda$ parameter.
-Because they both use transformed feature values, weights actually vary during the optimisation process for maximising the weighted log-likelihood function.
-This complicates the parameter search space, and leads to different parameter estimates.
-Weighting based on empirical probabilities does not suffer from this issue, because all weights are fixed during optimising the weighted log-likelihood function.
-For weighting based on the z-score of the transformed feature, or on their residual error, we can see two extreme strategies emerging.
-Firstly, the strategy to down-weight most instances ($r$ (step), and to a lesser degree $z$ (step)) and thereby force normality on the most central parts of the distribution.
-Second is an attempt to assign full-weight to almost all instances, with large values for $k_1$ and $k_2$, which leads to behaviour similar to non-robust power transformation.
+Figure **insert** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
+In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation.
+For both transformations, weighting based on empirical probabilities yields the most consistent $\lambda$ parameter estimates in the presence of outliers.
+Weighting based on residual errors tended to perform least well, and were generally similar to (Yeo-Johnson) or worse than (Box-Cox) non-robust transformations.
+
+```{r optimised_weighting_function_parameters, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Robustness of power transformations after optimising weighting function parameters. ",
+  "The distribution of errors, i.e. the difference between robustly fitted $\\hat{\\lambda}^r$ ",
+  "and expected $\\hat{\\lambda}_{0}$ found in the absence of outliers, is shown for 1000 randomly ",
+  "parametrised asymmetric generalised normal distributions with up to 10% outliers on either side ",
+  "of the distribution. The top and bottom panel show errors for the shift-sensitive Box-Cox transformation, ",
+  "and shift-sensitive Yeo-Johnson transformation, respectively. ",
+  "In each panel, the error distribution for the non-robust transformation is shown to the left ",
+  "for comparison with different weighting functions. ",
+  "Note that a square root transformation was used along the $y$-axis. Median errors are shown above each distribution."
+)
+
+.plot_optimised_weighting_function_parameters(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir)
+```
 

### 2023-03-02 manuscript/plots.R

@@ -607,3 +607,78 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
   return(p)
 }
+
+
+
+.plot_optimised_weighting_function_parameters <- function(plot_theme, manuscript_dir) {
+
+  # Prevent warnings due to non-standard evaluation.
+  estimation_method <- weight_method <- method <- k <- NULL
+  lambda <- target_lambda <- lambda_error <- i.lambda_error <- non_robust_lambda_error <- NULL
+  better_than_non_robust <- pass_rate <- median_error_round <- NULL
+
+  # This find the optimal weighting parameter settings that are used within the
+  # package. The output is not used here, but its presence is required for
+  # processing.
+  two_sided_function_parameters <- .get_optimised_weighting_function_parameters_two_sided(
+    manuscript_dir = manuscript_dir)
+
+  # This finds lambda values at the optimised weighting parameter settings.
+  data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+
+  # Select only MLE here.
+  data <- data[estimation_method == "MLE"]
+
+  # Compute difference to target.
+  data <- data[, "lambda_error" := abs(lambda - target_lambda)]
+
+  # Join with error for non-robust values.
+  non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
+  data[
+    non_robust_data,
+    on = c("method", "estimation_method", "k", "ii"),
+    "non_robust_lambda_error" := i.lambda_error]
+
+  data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
+  data[k == 0.0, "better_than_non_robust" := NA_integer_]
+  data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
+
+  pass_rate_data <- data[, list(
+    "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
+    "n" = sum(!is.na(better_than_non_robust)),
+    "median_error" = stats::median(lambda_error)),
+    by=c("method", "estimation_method", "weight_method")]
+  pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)]
+  pass_rate_data[, "median_error_round" := format(pass_rate_data$median_error, digits = 1)]
+
[…]

### 2023-03-04 manuscript/manuscript.Rmd
Computed optimised parameters for data where outliers appear on the right-hand (positive) side of the distribution.

@@
   "In each panel, the error distribution for the non-robust transformation is shown to the left ",
-  "for comparison with different weighting functions. ",
+  "for comparison with different weighting methods. ",
   "Note that a square root transformation was used along the $y$-axis. Median errors are shown above each distribution."
@@
 ```
+# Appendix C: Right-sided outliers
 
 ```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+
+.get_optimised_weighting_function_parameters(manuscript_dir = manuscript_dir, side = "right")
 

### 2023-03-04 manuscript/simulations.R

@@ -165,14 +165,22 @@
 
 
 
-.get_optimised_weighting_function_parameters_two_sided <- function(manuscript_dir){
+.get_optimised_weighting_function_parameters <- function(manuscript_dir, side = "both") {
 
   # generator ------------------------------------------------------------------
   experiment_args <- coro::generator(
     function(
     manuscript_dir,
+    side,
     transformation_methods = NULL,
-    estimation_methods = NULL){
+    estimation_methods = NULL) {
+
+      if (side == "both") {
+        file_dir <- "robustness_comparison"
+
+      } else if (side == "right") {
+        file_dir <- "robustness_comparison_right"
+      }
 
       if(is.null(transformation_methods)){
         transformation_methods <- c("box_cox", "yeo_johnson")
@@ -196,7 +204,7 @@
 
           file_name <- file.path(
             manuscript_dir,
-            "robustness_comparison",
+            file_dir,
             paste0(
               transformation_method, "_",
               estimation_method, "_",
@@ -255,7 +263,7 @@
 
               file_name <- file.path(
                 manuscript_dir,
-                "robustness_comparison",
+                file_dir,
                 paste0(
                   transformation_method, "_",
                   estimation_method, "_",
@@ -330,7 +338,8 @@
     alpha,
     beta,
     target_lambda,
-    outlier_fraction){
+    outlier_fraction,
+    side){
 
     # Set parameters.
     parameters <- list(
@@ -361,7 +370,16 @@
 
           # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
           # than Q3 + 1.5 IQR.
-          x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+          if (side == "both") {
+            x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+          } else if (side == "right") {
+            x_random <- stats::runif(n_draw, min=0.0, max=2.0)
+          } else if (side == "left") {
+            x_random <- stats::runif(n_draw, min=-2.0, max=0.0)
+          } else {
+            stop(paste0("side was not recognised: ", side))
+          }
+
           outlier <- numeric(n_draw)
           if(any(x_random < 0)){
[…]

### 2023-03-05 manuscript/simulations.R
Compute right-sided outliers

@@ -351,7 +351,7 @@
     # Add outliers.
     x <- lapply(
       outlier_fraction,
-      function(k, x, parameters){
+      function(k, x, parameters, side){
         # Set parameters.
         parameters$k <- k
 
@@ -398,7 +398,8 @@
           "parameters" = parameters))
       },
       x = x,
-      parameters)
+      parameters = parameters,
+      side = side)
 
     return(x)
   }
@@ -627,18 +628,33 @@
 
 
 
-.get_transformation_parameters_two_sided <- function(manuscript_dir){
+.get_transformation_parameters <- function(manuscript_dir, side = "both"){
 
-  if(!file.exists(file.path(manuscript_dir, "robustness_comparison_marginal_parameter_data.RDS"))){
+  if (side == "both"){
+    file_name <- file.path(manuscript_dir, "robustness_comparison_marginal_parameter_data.RDS")
+
+  } else if (side == "right") {
+    file_name <- file.path(manuscript_dir, "robustness_comparison_marginal_parameter_data_right.RDS")
+  }
+
+  if(!file.exists(file_name)){
 
     # generator ----------------------------------------------------------------
     experiment_args <- coro::generator(
       function(
-    manuscript_dir,
-    transformation_methods = NULL,
-    estimation_methods = NULL){
+        manuscript_dir,
+        side,
+        transformation_methods = NULL,
+        estimation_methods = NULL) {
 
-        if(is.null(transformation_methods)){
+        if (side == "both") {
+          file_dir <- "robustness_comparison"
+
+        } else if (side == "right") {
+          file_dir <- "robustness_comparison_right"
+        }
+
+        if (is.null(transformation_methods)) {
           transformation_methods <- c("box_cox", "yeo_johnson")
         }
 
@@ -658,7 +674,7 @@
 
             file_name <- file.path(
               manuscript_dir,
-              "robustness_comparison",
+              file_dir,
               paste0(
                 transformation_method, "_",
                 estimation_method, "_",
@@ -679,7 +695,7 @@
 
                 file_name <- file.path(
                   manuscript_dir,
-                  "robustness_comparison",
+                  file_dir,
                   paste0(
                     transformation_method, "_",
                     estimation_method, "_",
@@ -720,9 +736,9 @@
 
     # Helper function for computing the target lambda value.
     .compute_target_lambda <- function(
-    x,
[…]

### 2023-03-30 manuscript/manuscript.Rmd
Compute right-sided outliers
Added section on goodness of fit.

@@
 
-To be added.
+Power transformations aim to transform features to a normal distribution.
+However, this may not always be successful or possible.
+Deviations from normality can be detected by normality tests, such as the Shapiro-Wilk test [@Shapiro1965-zd].
+In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers or a combination of the two.
+Here we develop an empirical test for central normality.
+
+Let central normality be defined as normality of central $\kappa$ portion of the data, i.e.
+$\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$, with $p_i$ probabilities of the empirical distribution, as previously.
+We then compute the residual errors between the z-scores of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ and the expected z-scores from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+The set of residual errors for the central portion of the data is then $\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+To be continued.
 
@@
 # This finds lambda values at the optimised weighting parameter settings.
-data <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)
+data <- .get_transformation_parameters(manuscript_dir = manuscript_dir)
 

### 2023-03-30 manuscript/manuscript.Rmd
Removed code used for showing robustness.

@@
   manuscript_dir = manuscript_dir)
-```
-
-```{r, warning=FALSE, message=FALSE}
-# Only process data if the file has not been created.
-if(!file.exists(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))){
-
-  # Generator for creating parameters and data for all experiments.
-  generate_experiment_data <- coro::generator(
-    function(
-    x_normal,
-    x_right_skewed,
-    x_left_skewed){
-
-      n <- 20
-      k_range <- (seq_len(n) - 1)^2 /((n-1)^2 * 10)
-
-      for(distribution in c("normal", "right-skewed", "left-skewed")){
-        if(distribution == "normal"){
-          x <- x_normal
-
-        } else if(distribution == "right-skewed"){
-          x <- x_right_skewed
-
-        } else if(distribution == "left-skewed"){
-          x <- x_left_skewed
-        }
-        for(ii in seq_len(100)){
-          for(k in k_range){
-            # Compute interquartile range.
-            interquartile_range <- stats::IQR(x)
-
-            # Compute upper and lower quartiles.
-            q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
-            q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
-
-            # Set data where the outliers will be copied into.
-            x_outlier <- x
-
-            if(k != 0.0){
-              n_draw <- ceiling(k * length(x))
-
-              # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-              # than Q3 + 1.5 IQR.
-              x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
-              outlier <- numeric(n_draw)
-              if(any(x_random < 0)){
-                outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
-              }
-
-              if(any(x_random >= 0)){
-                outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
-              }
-
-              # Randomly insert outlier values.
-              x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
-            }
-
-            for(method in c("box_cox", "yeo_johnson")){
-              for(weight_method in c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine")){
-
-                k1 <- k2 <- NULL
-
-                # Optimal parameter values from the previous experiment.
-                if(weight_method == "original_step"){
-                  k1 <- ifelse(method == "box_cox", 0.84, 0.92)
-
-                } else if(weight_method == "original_cosine"){
-                  k1 <- ifelse(method == "box_cox", 0.76, 0.89)
-                  k2 <- ifelse(method == "box_cox", 0.98, 0.93)
-
-                } else if(weight_method == "transformed_step"){
-                  k1 <- ifelse(method == "box_cox", 1.09, 1.04)
-
-                } else if(weight_method == "transformed_cosine"){
-                  k1 <- ifelse(method == "box_cox", 0.03, 0.12)
-                  k2 <- ifelse(method == "box_cox", 6.91, 5.63)
-
-                } else if(weight_method == "residual_step"){
-                  k1 <- ifelse(method == "box_cox", 1.56, 10.00)
-
-                } else if(weight_method == "residual_cosine"){
-                  k1 <- ifelse(method == "box_cox", 1.32, 9.97)
-                  k2 <- ifelse(method == "box_cox", 1.65, 10.0)
-                }
-
-                yield(list(
-                  "ii" = ii,
-                  "x" = x_outlier,
-                  "k" = k,
-                  "distribution" = distribution,
-                  "method" = method,
-                  "robust" = weight_method != "none",
-                  "weight_method" = weight_method,
-                  "k1" = k1,
-                  "k2" = k2))
-              }
-            }
-          }
-        }
-      }
-    })
-
-  .compute_lambda <- function(parameter_set){
-    # Create transformer object.
-    transformer <- suppressWarnings(
-      power.transform::find_transformation_parameters(
-        x = parameter_set$x,
-        method = parameter_set$method,
-        robust = parameter_set$robust,
-        shift = TRUE,
-        weight_method = parameter_set$weight_method,
-        k1 = parameter_set$k1,
-        k2 = parameter_set$k2))
-
-    return(
-      data.table::data.table(
-        "distribution" = parameter_set$distribution,
-        "method" = parameter_set$method,
-        "weight_method" = parameter_set$weight_method,
-        "k" = parameter_set$k,
-        "lambda" = transformer@lambda,
-        "ii" = parameter_set$ii))
-  }
-
-  # Computations ---------------------------------------------------------------
-
-  # Set seed.
-  set.seed(19L)
-
-  # Normal distribution.
-  x_normal <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.5, beta=2)
-
-  # Right skewed data
-  x_right_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.2, beta=2)
-
-  # Left skewed data
-  x_left_skewed <- power.transform::ragn(10000L, location=0, scale=1/sqrt(2), alpha=0.8, beta=2)
-
-  # Generate all experiments.
-  experiments <- coro::collect(generate_experiment_data(
-    x_normal = x_normal,
-    x_right_skewed = x_right_skewed,
-    x_left_skewed = x_left_skewed
-  ))
-
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  # Compute all data in parallel.
-  data <- parallel::parLapply(
-    cl=cl,
-    X=experiments,
-    fun=.compute_lambda
-  )
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-
-  # Combine data into a single table.
-  data <- data.table::rbindlist(data)
-
-  saveRDS(data, file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
-
-}
-```
-
-```{r outlier-values_yeo-johnson, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Estimated $\\lambda$ values for \\textbf{shift-sensitive Yeo-Johnson} power transformations as a function of the fraction of outlier values. ",
-  "Features were randomly drawn from centered, right-skewed and left-skewed normal distributions, and a fraction of values was randomly replaced by outlier values.",
-  "The top row shows weighting methods that perform better than the non-robust version, ",
-  "namely empirical probabilities weighted using step (emp. prob. (step)) and tapered cosine (emp. prob. (tap. cos.)) weighting functions. ",
-  "The bottom row shows weighting methods that perform worse than the non-robust version. ",
-  "Note that the range of $\\lambda$ values differs between top and bottom rows.")
-
-# Prevent warnings due to non-standard evaluation.
-weight_method <- distribution <- k <- NULL
-
-.create_outlier_fraction_plot <- function(
-    data,
-    plot_theme,
-    limits,
-    lambda_target,
-    guide=FALSE){
-
-  # Prevent warnings due to non-standard evaluation.
-  weight_method <- k <- mean_lambda <- min_lambda <- max_lambda <- NULL
-
-  # Annotation settings.
-  annotation_settings <- get_annotation_settings(ggplot2::theme_light(base_size = base_size))
-
-  p <- ggplot2::ggplot(
-    data = data,
-    mapping = ggplot2::aes(
-      x = k,
-      y = mean_lambda,
-      ymin = min_lambda,
-      ymax = max_lambda))
-  p <- p + plot_theme
-
-  p <- p + ggplot2::geom_ribbon(mapping=ggplot2::aes(fill=weight_method), alpha=0.10)
-  p <- p + ggplot2::geom_line(mapping=ggplot2::aes(colour=weight_method))
-  p <- p + ggplot2::scale_x_continuous(
-    name = latex2exp::TeX("outlier fraction"),
-    trans = "sqrt")
-  p <- p + ggplot2::scale_y_continuous(
-    name = latex2exp::TeX("$\\lambda$"),
-    limits = limits)
-
-  p <- p + ggplot2::geom_hline(
-    yintercept = lambda_target, 
-    linetype="longdash", 
-    colour="grey40")
-
-  p <- p + ggplot2::annotate(
-    geom = "text",
-    x = 0.1,
-    y = lambda_target,
-    label = "target",
-    colour=annotation_settings$colour,
-    family=annotation_settings$family,
-    fontface=annotation_settings$face,
-    size=annotation_settings$geom_text_size,
-    vjust=-1.0,
-    hjust=1.0)
-
-  if(guide){
-    p <- p + paletteer::scale_color_paletteer_d(
-      palette="ggthemes::Tableau_10",
-      drop=FALSE,
-      guide=ggplot2::guide_legend("weighting method"))
-
-    p <- p + paletteer::scale_fill_paletteer_d(
-      palette="ggthemes::Tableau_10",
-      drop=FALSE,
-      guide=ggplot2::guide_legend("weighting method"))
-
-  } else {
-    p <- p + paletteer::scale_color_paletteer_d(
-      palette="ggthemes::Tableau_10",
-      drop=FALSE,
-      guide="none")
-
-    p <- p + paletteer::scale_fill_paletteer_d(
-      palette="ggthemes::Tableau_10",
-      drop=FALSE,
-      guide="none")
-  }
-
-  p <- p + ggplot2::scale_alpha(guide="none")
-
-  # Update title.
-  p <- p + ggplot2::theme(
-    plot.title=ggplot2::element_text(
-      hjust=0.5,
-      size=ggplot2::rel(1.0)))
-
-  return(p)
-}
-
-data <- readRDS(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
-
-# Update distribution, method and version to factors.
-data$distribution <- factor(
-  x = data$distribution,
-  levels = c("normal", "right-skewed", "left-skewed"))
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-data$weight_method <- factor(
-  x = data$weight_method,
-  levels = c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
-  labels = c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
-
-data <- data[method == "Yeo-Johnson", list(
-  "mean_lambda"=mean(lambda),
-  "min_lambda"=min(lambda),
-  "max_lambda"=max(lambda)),
-  by=c("distribution", "method", "weight_method", "k")]
-
-top_group <- c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)")
-bottom_group <- c("z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
-
-top_limits <- c(0.2, 1.8)
-bottom_limits <- c(-2.5, 4.0)
-
-# Normal distribution ----------------------------------------------------------
-p_normal_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "normal" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits = top_limits,
-  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_normal_t <- p_normal_t + ggplot2::ggtitle(label="normal distribution")
-p_normal_t <- p_normal_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank())
-
-p_normal_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "normal" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_normal_b <- p_normal_b + ggplot2::theme(
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Right skewed distribution ----------------------------------------------------
-p_right_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "right-skewed" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits= top_limits,
-  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_right_t <- p_right_t + ggplot2::ggtitle(label="right-skewed distribution")
-p_right_t <- p_right_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
-
-p_right_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "right-skewed" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_right_b <- p_right_b + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Left skewed distribution -----------------------------------------------------
-p_left_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "left-skewed" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits = top_limits,
-  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda,
-  guide = TRUE)
-p_left_t <- p_left_t + ggplot2::ggtitle(label="left-skewed distribution")
-p_left_t <- p_left_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
-
-p_left_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "left-skewed" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_left_b <- p_left_b + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Aggregate --------------------------------------------------------------------
-
-p <- egg::ggarrange(
-  plots=list(
-    p_normal_t, p_right_t, p_left_t,
-    p_normal_b, p_right_b, p_left_b),
-  ncol=3,
-  heights=c(1.0, 1.0),
-  byrow=TRUE,
-  draw=FALSE)
-p
-```
-
-```{r outlier-values-box-cox, warning=FALSE, message=FALSE, fig.cap=cap}
-
-data <- readRDS(file.path(manuscript_dir, "robustness_comparison_plot.RDS"))
-
-cap <- paste0(
-  "Estimated $\\lambda$ values for \\textbf{shift-sensitive Box-Cox} power transformations as a function of the fraction of outlier values. ",
-  "Features were randomly drawn from centered, right-skewed and left-skewed normal distributions, and a fraction of values was randomly replaced by outlier values.",
-  "The top row shows weighting methods that perform better than the non-robust version, ",
-  "namely empirical probabilities weighted using step (emp. prob. (step)) and tapered cosine (emp. prob. (tap. cos.)) weighting functions. ",
-  "The bottom row shows weighting methods that perform worse than the non-robust version. ",
-  "Note that the range of $\\lambda$ values differs between top and bottom rows.")
-
-# Update distribution, method and version to factors.
-data$distribution <- factor(
-  x = data$distribution,
-  levels = c("normal", "right-skewed", "left-skewed"))
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-data$weight_method <- factor(
-  x = data$weight_method,
-  levels = c("none", "original_step", "original_cosine", "transformed_step", "transformed_cosine", "residual_step", "residual_cosine"),
-  labels = c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)", "z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)"))
-
-data <- data[method == "Box-Cox", list(
-  "mean_lambda"=mean(lambda),
-  "min_lambda"=min(lambda),
-  "max_lambda"=max(lambda)),
-  by=c("distribution", "method", "weight_method", "k")]
-
-top_group <- c("non-robust", "emp. prob. (step)", "emp. prob. (tap. cos.)")
-bottom_group <- c("z-score (step)", "z-score (tap. cos.)", "residual (step)", "residual (tap. cos.)")
-
-top_limits <- c(-3, 4.5)
-bottom_limits <- c(-4.01, 6.01)
-
-# Normal distribution ----------------------------------------------------------
-p_normal_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "normal" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits = top_limits,
-  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_normal_t <- p_normal_t + ggplot2::ggtitle(label="normal distribution")
-p_normal_t <- p_normal_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank())
-
-p_normal_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "normal" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "normal" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_normal_b <- p_normal_b + ggplot2::theme(
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Right skewed distribution ----------------------------------------------------
-p_right_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "right-skewed" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits= top_limits,
-  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_right_t <- p_right_t + ggplot2::ggtitle(label="right-skewed distribution")
-p_right_t <- p_right_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
-
-p_right_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "right-skewed" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "right-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_right_b <- p_right_b + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Left skewed distribution -----------------------------------------------------
-p_left_t <- .create_outlier_fraction_plot(
-  data = data[distribution == "left-skewed" & weight_method %in% top_group],
-  plot_theme = plot_theme,
-  limits = top_limits,
-  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda,
-  guide = TRUE)
-p_left_t <- p_left_t + ggplot2::ggtitle(label="left-skewed distribution")
-p_left_t <- p_left_t + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank(),
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank())
-
-p_left_b <- .create_outlier_fraction_plot(
-  data = data[distribution == "left-skewed" & weight_method %in% bottom_group],
-  plot_theme = plot_theme,
-  limits = bottom_limits,
-  lambda_target = data[distribution == "left-skewed" & weight_method == "non-robust" & k == 0.0]$mean_lambda)
-p_left_b <- p_left_b + ggplot2::theme(
-  axis.title.y = ggplot2::element_blank(),
-  axis.text.y = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_text(
-    angle=60.0,
-    hjust=1.0))
-
-# Aggregate --------------------------------------------------------------------
-
-p <- egg::ggarrange(
-  plots=list(
-    p_normal_t, p_right_t, p_left_t,
-    p_normal_b, p_right_b, p_left_b),
-  ncol=3,
-  heights=c(1.0, 1.0),
-  byrow=TRUE,
-  draw=FALSE)
-p
-
 ```

### 2023-04-02 manuscript/manuscript.Rmd
WIP on central normality tests.

@@
 cap <- paste0(
-  "CHECK. ")
-
-residual <- residual_error <- threshold <- NULL
-
-if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))){
-
-  set.seed(95)
-
-  # Generate table of asymmetric generalised normal distribution parameters.
-  n_distributions <- 1000
-
-  # Generate alpha, beta and n.
-  n <- stats::runif(n=n_distributions, min=2, max=4)
-  n <- ceiling(10^n)
-
-  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
-  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
-
-  # Generate corresponding distributions.
-  x <- mapply(
-    power.transform::ragn,
-    n = n,
-    alpha = alpha,
-    beta = beta)
-
-  # Add outliers, between 0.00 and 0.10.
-  n_outliers <- 10
-  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)
-
-  x <- mapply(
-    function(x, n, alpha, beta, outlier_fraction){
-
-      parameters <- list("n" = n, "alpha"=alpha, "beta" = beta)
-
-      x <- lapply(
-        outlier_fraction,
-        function(k, x, parameters){
-          # Set parameters.
-          parameters$k <- k
-
-          # Compute interquartile range.
-          interquartile_range <- stats::IQR(x)
-
-          # Compute upper and lower quartiles.
-          q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
-          q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
-
-          # Set data where the outliers will be copied into.
-          x_outlier <- x
-
-          if(k != 0.0){
-            n_draw <- ceiling(k * length(x))
-
-            # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-            # than Q3 + 1.5 IQR.
-            x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
-            outlier <- numeric(n_draw)
-            if(any(x_random < 0)){
-              outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
-            }
-
-            if(any(x_random >= 0)){
-              outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
-            }
-
-            # Randomly insert outlier values.
-            x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
-          }
-
-          return(list(
-            "x" = x_outlier,
-            "parameters" = parameters))
-        },
-        x = x,
-        parameters)
-
-      return(x)
-    },
-    x = x,
-    n = n,
-    alpha = alpha,
-    beta = beta,
-    MoreArgs=list("outlier_fraction"=outlier_fraction),
-    SIMPLIFY=FALSE,
-    USE.NAMES=FALSE)
-
-  x <- unlist(x, recursive = FALSE)
-  x <- mapply(
-    function(ii, x){
-      x$distribution_id <- ii
-
-      return(x)},
-    ii = seq_along(x),
-    x = x,
-    SIMPLIFY = FALSE)
-
-  compute_residuals <- function(
-    x,
-    n_sample = 200){
-
-    # Set interpolation points.
-    p_sample <- (seq_len(n_sample) - 1/3) / (n_sample + 1/3)
-
-    # Determine residuals for Box-Cox transformations.
-    transformer_bc <- suppressWarnings(power.transform::find_transformation_parameters(
-      x = x$x,
-      method = "box_cox",
-      shift = TRUE,
-      robust = TRUE,
-      weight_method = "original_cosine",
-      k1 = 0.76,
-      k2 = 0.95))
-
-    residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_bc)
-
-    residual <- stats::approx(
-      x = residual_data$p,
-      y = abs(residual_data$residual),
-      xout = p_sample,
-      rule = 2,
-      ties = max)$y
-
-    residual_bc <- data.table::data.table(
-      "distribution_id" = x$distribution_id,
-      "p" = p_sample,
-      "residual" = residual,
-      "method" = "box_cox",
-      "has_outliers" = x$parameters$k > 0)
-
-    # Determine residuals for Yeo-Johnson transformations.
-    transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
-      x = x$x,
-      method = "yeo_johnson",
-      shift = TRUE,
-      robust = TRUE,
-      weight_method = "original_cosine",
-      k1 = 0.89,
-      k2 = 0.93))
-
-    residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_yj)
-
-    residual <- stats::approx(
-      x = residual_data$p,
-      y = abs(residual_data$residual),
-      xout = p_sample,
-      rule = 2,
-      ties = max)$y
-
-    residual_yj <- data.table::data.table(
-      "distribution_id" = x$distribution_id,
-      "p" = p_sample,
-      "residual" = residual,
-      "method" = "yeo_johnson",
-      "has_outliers" = x$parameters$k > 0)
-
-    return(data.table::rbindlist(list(residual_bc, residual_yj)))
-  }
-
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  # Compute all data in parallel.
-  data <- parallel::parLapply(
-    cl=cl,
-    X=x,
-    fun=compute_residuals)
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-
-  # data <- lapply(
-  #   x,
-  #   compute_residuals)
-
-  data <- data.table::rbindlist(data)
-
-  saveRDS(
-    object=data,
-    file=file.path(manuscript_dir, "residual_plot.RDS"))
-
-} else {
-  data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
-}
-
-data$method <- factor(
-  x = data$method,
-  levels = c("box_cox", "yeo_johnson"),
-  labels = c("Box-Cox", "Yeo-Johnson"))
-
-outlier_free_data <- data[has_outliers == FALSE, list("residual_error"=stats::quantile(residual, probs=0.99)), by=c("p", "method")]
-
-data <- data[, list(
-  "residual_50"=stats::quantile(residual, probs=0.50),
-  "residual_90"=stats::quantile(residual, probs=0.90),
-  "residual_95"=stats::quantile(residual, probs=0.95),
-  "residual_99"=stats::quantile(residual, probs=0.99),
-  "residual_max"=max(residual)),
-  by=c("p", "method")]
-
-data <- data.table::melt(
-  data = data,
-  id.vars = c("p", "method"),
-  variable.name = "threshold",
-  value.name = "residual_error")
-
-data$threshold <- factor(
-  x = data$threshold,
-  levels = c("residual_50", "residual_90", "residual_95", "residual_99", "residual_max"),
-  labels = c("50 %", "90 %", "95 %", "99 %", "100 %"))
-
-p_bc <- ggplot2::ggplot(
-  data = data[method == "Box-Cox"],
-  mapping = ggplot2::aes(
-    x = p,
-    y = residual_error,
-    colour = threshold))
-p_bc <- p_bc + plot_theme
-p_bc <- p_bc + ggplot2::geom_line()
-p_bc <- p_bc + ggplot2::geom_line(
-  data = outlier_free_data[method == "Box-Cox"],
-  mapping = ggplot2::aes(
-    x = p,
-    y = residual_error),
-  colour = "gray40",
-  linetype = "dashed")
-p_bc <- p_bc + ggplot2::scale_colour_discrete(
-  name = "Box-Cox threshold",
-  type=c("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E"))
-p_bc <- p_bc + ggplot2::xlab("empirical probability")
-p_bc <- p_bc + ggplot2::ylab("absolute residual error")
-p_bc <- p_bc + ggplot2::coord_cartesian(
-  xlim = c(0, 1),
-  ylim = c(0, 1))
-p_bc <- p_bc + ggplot2::theme(
-  axis.title.x = ggplot2::element_blank(),
-  axis.text.x = ggplot2::element_blank())
-
-p_yj <- ggplot2::ggplot(
-  data = data[method == "Yeo-Johnson"],
-  mapping = ggplot2::aes(
-    x = p,
-    y = residual_error,
-    colour = threshold))
-p_yj <- p_yj + plot_theme
-p_yj <- p_yj + ggplot2::geom_line()
-p_yj <- p_yj + ggplot2::geom_line(
-  data = outlier_free_data[method == "Yeo-Johnson"],
-  mapping = ggplot2::aes(
-    x = p,
-    y = residual_error),
-  colour = "gray40",
-  linetype = "dashed")
-p_yj <- p_yj + ggplot2::scale_colour_discrete(
-  name = "Yeo-Johnson threshold",
-  type=c("50 %" = "#FFBD7D", "90 %" = "#FFA954", "95 %" = "#F28E2B", "99 %" = "#CD6B0B", "100 %" = "#A25000"))
-p_yj <- p_yj + ggplot2::xlab("empirical probability")
-p_yj <- p_yj + ggplot2::ylab("absolute residual error")
-p_yj <- p_yj + ggplot2::coord_cartesian(
-  xlim = c(0, 1),
-  ylim = c(0, 1))
-
-p <- egg::ggarrange(
-  plots=list(
-    p_bc,
-    p_yj),
-  ncol=1,
-  heights=c(1.0, 1.0),
-  byrow=TRUE,
-  draw=FALSE)
-p
+  "Residual error as a function of empirical probability. ",
+  "Percentiles of the error are shown for robust, shift-sensitive transformations of 1000 randomly drawn asymmetric ",
+  "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
+  "(up to 10% of samples) each. Larger errors occur for more extreme values at the edges of each distribution. ",
+  "For comparison, the 99th percentile of residual errors for transformations of outlier-free distributions are shown (grey dashed line).")
+
+.plot_residuals(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir)
+
 ```
@@
 ```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
-
-if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))) stop("Residual data for goodness-of-fit test does not exist.")
-
-data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
-
-central_width <- c(0.60, 0.70, 0.80, 0.90, 0.95, 1.00)
-
-p_lower <- 0.50 - central_width / 2
-p_upper <- 0.50 + central_width / 2
-
-x <- data.table::copy(data)
-
-# Clip empirical probabilities to centre.
-x <- x[p >= p_lower & p <= p_upper]
-
-# Compute mean absolute residual error per feature.
-x <- x[, list("mare"=mean(residual)), by=c("distribution_id", "method")]
-x <- x[, list("n"=.N), by=c("mare", "method")][order(mare, method)]
-x[, "rejected":= 1.0 - cumsum(n) / sum(n)]
-x <- rbind(data.table::data.table("mare"=c(0.0, 0.0), "method"=c("box_cox", "yeo_johnson"), "n"=0L, "rejected"=0.0))
-
-p <- ggplot2::ggplot(
-  data = x,
-  mapping = ggplot2::aes(
-    x = mare,
-    y = rejected,
-    colour = method))
-p <- p + plot_theme
-p <- p + ggplot2::geom_step()
-p <- p + ggplot2::xlab("mean absolute residual error")
-p <- p + paletteer::scale_color_paletteer_d(
-  palette="ggthemes::Tableau_10",
-  drop=FALSE)
+cap <- paste0(
+  "Check")
+.plot_
+
 ```

### 2023-04-02 manuscript/plots.R

@@ -613,7 +613,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 .plot_optimised_weighting_function_parameters <- function(plot_theme, manuscript_dir) {
 
   # Prevent warnings due to non-standard evaluation.
-  estimation_method <- weight_method <- method <- k <- NULL
+  estimation_method <- weight_method <- method <- k <- n <- NULL
   lambda <- target_lambda <- lambda_error <- i.lambda_error <- non_robust_lambda_error <- NULL
   better_than_non_robust <- pass_rate <- median_error_round <- NULL
 
@@ -682,4 +682,150 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p <- p + ggplot2::xlab("weighting method")
 
   return(p)
+}
+
+
+
+.plot_residuals <- function(plot_theme, manuscript_dir){
+
+  # Prevent warnings due to non-standard evaluation.
+  has_outliers <- residual <- threshold <- residual_error <- method <- p <- NULL
+
+  # Get data
+  data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
+
+  outlier_free_data <- data[
+    has_outliers == FALSE,
+    list("residual_error"=stats::quantile(residual, probs=0.99)),
+    by=c("p", "method")]
+
+  data <- data[, list(
+    "residual_50"=stats::quantile(residual, probs=0.50),
+    "residual_90"=stats::quantile(residual, probs=0.90),
+    "residual_95"=stats::quantile(residual, probs=0.95),
+    "residual_99"=stats::quantile(residual, probs=0.99),
+    "residual_max"=max(residual)),
+    by=c("p", "method")]
+
+  data <- data.table::melt(
+    data = data,
+    id.vars = c("p", "method"),
+    variable.name = "threshold",
+    value.name = "residual_error")
+
+  data$threshold <- factor(
+    x = data$threshold,
+    levels = c("residual_50", "residual_90", "residual_95", "residual_99", "residual_max"),
+    labels = c("50 %", "90 %", "95 %", "99 %", "100 %"))
+
+  p_bc <- ggplot2::ggplot(
+    data = data[method == "Box-Cox"],
+    mapping = ggplot2::aes(
[…]

### 2023-04-02 manuscript/simulations.R

@@ -994,3 +994,218 @@
 
   return(data)
 }
+
+
+
+.get_goodness_of_fit_data <- function(manuscript_dir) {
+
+  # Helper function for population distributions with outliers.
+  .populate_outliers <- function(
+    x,
+    n,
+    alpha,
+    beta,
+    ii,
+    outlier_fraction,
+    side = "both") {
+
+    # Set parameters.
+    parameters <- list(
+      "n" = n,
+      "alpha" = alpha,
+      "beta" = beta,
+      "ii" = ii)
+
+    # Add outliers.
+    x <- lapply(
+      outlier_fraction,
+      function(k, x, parameters, side){
+        # Set parameters.
+        parameters$k <- k
+
+        # Compute interquartile range.
+        interquartile_range <- stats::IQR(x)
+
+        # Compute upper and lower quartiles.
+        q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
+        q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
+
+        # Set data where the outliers will be copied into.
+        x_outlier <- x
+
+        if(k != 0.0){
+          n_draw <- ceiling(k * length(x))
+
+          # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
+          # than Q3 + 1.5 IQR.
+          if (side == "both") {
+            x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
+          } else if (side == "right") {
+            x_random <- stats::runif(n_draw, min=0.0, max=2.0)
+          } else if (side == "left") {
+            x_random <- stats::runif(n_draw, min=-2.0, max=0.0)
+          } else {
+            stop(paste0("side was not recognised: ", side))
+          }
+
+          outlier <- numeric(n_draw)
+          if(any(x_random < 0)){
+            outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
+          }
+
+          if(any(x_random >= 0)){
+            outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
+          }
+
+          # Randomly insert outlier values.
+          x_outlier[sample(seq_along(x), size = n_draw, replace = FALSE)] <- outlier
+        }
+
+        return(list(
+          "x" = x_outlier,
[…]

### 2023-04-03 manuscript/manuscript.Rmd
Further work on empirical test.

@@
 Here we develop an empirical test for central normality.
-
-Let central normality be defined as normality of central $\kappa$ portion of the data, i.e.
+The null hypothesis $H_0$ is that the distribution is centrally normal.
+The alternative hypothesis $H_1$ is that the distribution is not centrally normal.
+
+Let central normality be defined as normality of central portion $\kappa$ of the data, i.e.
 $\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$, with $p_i$ probabilities of the empirical distribution, as previously.
@@
 The set of residual errors for the central portion of the data is then $\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+The test statistic $\tau_{\text{gof}}$ is then defined as:
+```{=latex}
+\begin{equation}
+\tau_{\text{gof}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
+\end{equation}
+```
+Here $[\quad]$ denotes an Iverson bracket. The test statistic is equal to the mean of the residual errors of the central portion of the data.
 
@@
 
-To answer:
-1.
-What is the residual threshold and the volume required to avoid rejecting transformations?
-Optimise: 1.
-window; 2.
-threshold; 3.
-threshold fraction.
-?
-Create a figure that summarises residual error thresholds (50%, 80%, 90%, 95%, 99%, 100%), below which x% of the data falls.
-2. How does this correspond to distance between bimodal normal distributions?
-
-```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
+To develop an empirical goodness of fit test we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{gof}}$.
+We will first define the central portion $\kappa$.
+
+First we draw $m_d=100000$ random asymmetric generalised normal distributions, parametrised as above.
+As before, for each distribution, up to $10 \%$ of instances are replaced by outlier values, and this is repeated $10$ times.
+We then compute residuals after optimising robust shift-sensitive transformations with the empirical tapered cosine weighting method.
+
+Figure *Insert reference* shows the residual errors of the transformed distribution as function of the empirical probability.
+As expected, the largest deviations from normality appear on the extremities of this range.
+For Box-Cox, the effect of outliers on the lower end of the distribution leads to noticeable asymmetry.
+Between very low and very high empirical probabilities, residual errors are constrained and relatively flat.
+This is the candidate range for the central portion of the data.
+
+```{r goodness-of-fit-1, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
   "Residual error as a function of empirical probability. ",
-  "Percentiles of the error are shown for robust, shift-sensitive transformations of 1000 randomly drawn asymmetric ",
+  "Percentiles of the error are shown for robust, shift-sensitive transformations of 100000 randomly drawn asymmetric ",
   "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
@@
 ```
-
-```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
+We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic.
+Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure *Insert*.
+Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
+This leads to test statistic values listed in Table *insert*.
+
+```{r goodness-of-fit-2, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
-  "Check")
-.plot_
-
-```
+  "Rejected transformed asymmetric generalised normal distributions as function of ",
+  "the mean absolute residual error of the central portion $\\kappa$ of the distribution.")
+
+.plot_mean_absolute_residual_error(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir)
+
+```
+
+The second error,
 

### 2023-04-03 manuscript/plots.R

@@ -795,13 +795,13 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x <- x[p >= p_lower & p <= p_upper]
 
     # Compute mean absolute residual error per feature.
-    x <- x[, list("mare"=mean(residual)), by=c("distribution_id", "method")]
+    x <- x[, list("mare"=mean(residual)), by=c("distribution_id", "outlier_id", "method")]
     x <- x[, list("n"=.N), by=c("mare", "method")][order(mare, method)]
     x[, "rejected" := 1.0 - cumsum(n) / sum(n), by = "method"]
     x <- rbind(
       data.table::data.table(
         "mare"=c(0.0, 0.0),
-        "method"=c("box_cox", "yeo_johnson"),
+        "method"=c("Box-Cox", "Yeo-Johnson"),
         "n"=0L,
         "rejected"=1.0),
       x)
@@ -811,21 +811,58 @@ get_annotation_settings <- function(ggtheme = NULL) {
     mare_data[[ii]] <- x
   }
 
-  mare_data <- data.table::rbindlist(mare_data)
+  data <- data.table::rbindlist(mare_data)
 
-  p <- ggplot2::ggplot(
-    data = mare_data,
+  data$kappa <- factor(
+    x = data$kappa,
+    levels = central_width,
+    labels = c("60 %", "70 %", "80 %", "90 %", "95 %", "100 %"))
+
+  p_bc <- ggplot2::ggplot(
+    data = data[method == "Box-Cox"],
     mapping = ggplot2::aes(
       x = mare,
       y = rejected,
-      colour = method))
-  p <- p + plot_theme
-  p <- p + ggplot2::geom_step()
-  p <- p + ggplot2::xlab("mean absolute residual error")
-  p <- p + paletteer::scale_color_paletteer_d(
-    palette="ggthemes::Tableau_10",
-    drop=FALSE)
-  p <- p + ggplot2::coord_cartesian(xlim = c(0, 0.25))
-  p <- p + ggplot2::facet_wrap("kappa", ncol = 3)
+      colour = kappa))
+  p_bc <- p_bc + plot_theme
+  p_bc <- p_bc + ggplot2::geom_step()
+  p_bc <- p_bc + ggplot2::scale_colour_discrete(
+    name = "central portion κ\n(Box-Cox)",
+    type=c(
[…]

### 2023-04-03 manuscript/simulations.R

@@ -1150,7 +1150,7 @@
     set.seed(95)
 
     # Generate table of asymmetric generalised normal distribution parameters.
-    n_distributions <- 100L
+    n_distributions <- 10000L
 
     # Generate alpha, beta and n.
     n <- stats::runif(n=n_distributions, min=2, max=4)

### 2023-04-03 manuscript/manuscript.Rmd
Added table processor.

@@
 
-First we draw $m_d=100000$ random asymmetric generalised normal distributions, parametrised as above.
+First we draw $m_d=10000$ random asymmetric generalised normal distributions, parametrised as above.
 As before, for each distribution, up to $10 \%$ of instances are replaced by outlier values, and this is repeated $10$ times.
@@
   "Residual error as a function of empirical probability. ",
-  "Percentiles of the error are shown for robust, shift-sensitive transformations of 100000 randomly drawn asymmetric ",
+  "Percentiles of the error are shown for robust, shift-sensitive transformations of 10000 randomly drawn asymmetric ",
   "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
@@
 ```
-
-The second error,
+```{r include = FALSE, eval = FALSE}
+data <- .get_test_statistic_data(manuscript_dir = manuscript_dir)
+
+```
 

### 2023-04-03 manuscript/simulations.R

@@ -1105,11 +1105,9 @@
     residual_bc <- data.table::data.table(
       "distribution_id" = x$parameter$ii,
       "outlier_id" = x$parameters$jj,
-      "p" = p_sample,
-      "k" = x$parameter$k,
+      "p" = seq_along(p_sample),
       "residual" = residual,
-      "method" = "box_cox",
-      "has_outliers" = x$parameters$k > 0)
+      "method" = "box_cox")
 
     # Determine residuals for Yeo-Johnson transformations.
     transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
@@ -1134,13 +1132,18 @@
     residual_yj <- data.table::data.table(
       "distribution_id" = x$parameter$ii,
       "outlier_id" = x$parameters$jj,
-      "p" = p_sample,
-      "k" = x$parameter$k,
+      "p" = seq_along(p_sample),
       "residual" = residual,
-      "method" = "yeo_johnson",
-      "has_outliers" = x$parameters$k > 0)
+      "method" = "yeo_johnson")
 
-    return(data.table::rbindlist(list(residual_bc, residual_yj)))
+    data <- data.table::rbindlist(list(residual_bc, residual_yj))
+
+    data$method <- factor(
+      x = data$method,
+      levels = c("box_cox", "yeo_johnson"),
+      labels = c("Box-Cox", "Yeo-Johnson"))
+
+    return(data)
   }
 
 
@@ -1209,10 +1212,57 @@
     data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
   }
 
-  data$method <- factor(
-    x = data$method,
-    levels = c("box_cox", "yeo_johnson"),
-    labels = c("Box-Cox", "Yeo-Johnson"))
+  # Convert p to empirical probabilities.
+  data[, "p" := (p - 1/3) / (200 + 1/3)]
+  data[, "has_outliers" := outlier_id > 1L]
 
   return(data)
+}
+
+
+
+.get_test_statistics_data <- function(manuscript_dir){
+  # Compute test statistic values.
+  data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
+
+  central_width <- 0.80
+
+  significance_values <- c(0.50, 0.20, 0.10, 0.05, 0.02, 0.01, 0.001)
+
+  p_lower <- 0.50 - central_width / 2
+  p_upper <- 0.50 + central_width / 2
+
[…]

### 2023-04-04 manuscript/manuscript.Rmd
Added code for empirical gof statistics in the appendix

@@
 ```
+
 ```{r include = FALSE, eval = FALSE}
-data <- .get_test_statistic_data(manuscript_dir = manuscript_dir)
+data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
 
@@
 
+# Appendix D: Generalisability of empirical goodness of fit test
+
+Add details on how generalisable the test statistic is for fits generated by other methods:
+
+- using different estimation method (Cramér-von Mises)
+- using conventional power transformation
+- using non-robust power transformation (only outlier-free data)
+
+```{r}
+data <- .get_test_statistics_data_appendix(manuscript_dir = manuscript_dir)
+
+```
+
 # References

### 2023-04-04 manuscript/plots.R

@@ -778,6 +778,9 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 .plot_mean_absolute_residual_error <- function(plot_theme, manuscript_dir){
 
+  # Prevent warnings due to non-standard evaluation.
+  residual <- p <- n <- mare <- method <- NULL
+
   # Get data
   data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
 

### 2023-04-05 manuscript/manuscript.Rmd
Aggregated plotting data are now stored instead of raw data due to large file size.
Implemented empirical goodness of fit test into the package.

--- 2023-04-03 manuscript/manuscript.Rmd
@@
 Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
-This leads to test statistic values listed in Table *insert*.
+This leads to test statistic values listed in Table *insert*, which derive from the type I error rates of the Box-Cox transformation.
 
@@
 
-.plot_mean_absolute_residual_error(
+.plot_type_1_error_rate(
   plot_theme = plot_theme,
@@
 ```
+
 ```{r include = FALSE, eval = FALSE}
-data <- .get_test_statistic_data(manuscript_dir = manuscript_dir)
-
+table_data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
+```
+
+```{r include = FALSE, eval = FALSE}
+load(file = "./R/sysdata.rda")
+
+gof_lookup_table <- .get_test_statistics_data(manuscript_dir = manuscript_dir, reduce = TRUE, as_table = FALSE)
+gof_lookup_table[, "method" := NULL]
+
+usethis::use_data(
+  gof_lookup_table, 
+  two_sided_function_parameters, 
+  internal = TRUE, 
+  overwrite = TRUE)
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Probability of central normality, based on type I error rates for Box-Cox transformations.}
+\begin{tabular}{c c c c c c c}
+
+\toprule
+0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+
+\midrule
+0.029 & 0.039 & 0.047 & 0.054 & 0.062 & 0.069 & 0.090 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 ```
@@
 
+# Appendix D: Generalisability of empirical goodness of fit test
+
+The empirical goodness of fit test was characterised using the empirical distribution of type I errors related to the mean residual error in the central portion ($\kappa = 0.80$) of the transformed feature distribution, in the presence of outliers.
+The corresponding power transformations were both robust and shift sensitive, and used maximum likelihood estimation (MLE) for optimisation of transformation parameters.
+An important concern is how well the test performs for conventional power transformations (in absence of outliers), and for optimisation criteria that are objectively better than MLE, i.e.
+the Cramér-von Mises criterion.
+
+For Figure *insert* determined type 1 error rate as before, but for several variants.
+In addition to robust, shift sensitive transformations using MLE, conventional transformations and robust, shift-sensitive using the Crámer-von Mises criterion were assessed.
+Conventional transformations only used randomly drawn asymmetric generalised normal distributions without added outliers.
+This figure suggests two things: If conventional transformations were used instead of robust, shift sensitive variants, the test characteristics would be highly similar.
+However, the robust, shift sensitive variant shows that the test characteristics are even stable in the presence of outliers.
+Secondly, the optimisation criterion does directly matter, and compared to MLE, the Crámer-von Mises criterion results in transformed distributions that better approximate central normality.
+In this case, one could argue whether characteristics of the empirical goodness of fit test should be based on the Crámer-von Mises criterion instead of MLE.
+We would argue that MLE-based test characteristics form a good foundation, for two reasons.
+One, MLE is the most prevalent optimisation criterion for power transformations, and two, the resulting transformations still approach central normality.
+
+```{r gof_test_type_1_error_rate_appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Type I error rates for power transformation variants. ",
+  "The robust, shift sensitive power transformation using maximum likelihood estimation (MLE) is shown as comparison. ",
+  "Optimisation of transformation parameters using the Crámer-von Mises (C-vM) criterion leads to lower residual errors compared to MLE, ",
+  "particularly for Box-Cox transformations, i.e. they lead to transformations that are closer to central normality. ",
+  "Conventional transformations (in the absence of outliers) yield curves that are comparable to their robust, shift sensitive variants."
+)
+
+.plot_type_1_error_rate_appendix(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir)
+
+```
+
 # References

### 2023-04-05 R/GoodnessOfFit.R

@@ -11,17 +11,11 @@ NULL
 #'
 #' @param transformer A transformer object created using
 #'   `find_transformation_parameters`.
-#' @param threshold A numeric value corresponding to the upper limit of the
-#'   difference between the z-score of one instance of the transformed vector
-#'   `x` and its expected z-score according to the normal distribution.
-#' @param centre_width A numeric value between 0.0 and 1.0 that describes the
-#'   width of the centre of the data. Only instances within this centre are
-#'   compared against the threshold. Can be `NULL` to use all instances.
 #' @param ... Unused arguments.
 #'
 #' @inheritParams power_transform
 #'
-#' @return Fraction of instances whose residual exceeds the threshold.
+#' @return p-value for empirical goodness of fit test.
 #' @export
 #'
 #' @examples
@@ -36,33 +30,35 @@ NULL
 assess_transformation <- function(
     x,
     transformer,
-    threshold = 0.15,
-    centre_width = 0.80,
     ...) {
 
   # Prevent CRAN NOTE due to non-standard use of variables by data.table.
-  z_expected <- NULL
+  p_observed <- NULL
 
   # Compute fit data.
   residual_data <- get_residuals(
     x = x,
     transformer = transformer)
 
-  # Assign full weight to central elements. We do this to ensure that poor fits
-  # don't get down-weighted in the centre, and in that way produce
-  # log-likelihood values that are too optimistic.
-  if (!is.null(centre_width)) {
-    if (centre_width < 0.0 || centre_width > 1.0) {
-      stop(paste0("centre_width should be NULL or between 0.0 and 1.0. Found: ", centre_width))
-    }
+  # The test uses a central portion kappa = 0.80.
+  residual_data <- residual_data[p_observed >= 0.10 & p_observed <= 0.90]
 
-    residual_data <- residual_data[abs(z_expected) <= stats::qnorm(0.50 + centre_width / 2), ]
-    if (nrow(residual_data) == 0) {
-      stop(paste0("centre_width may be too small: no residuals were selected."))
-    }
+  test_statistic_value <- mean(abs(residual_data$residual))
+
+  if(test_statistic_value < min(gof_lookup_table$test_statistic)) return(1.0)
+  if(test_statistic_value > gof_lookup_table[alpha == 0.0001]$test_statistic){
+    message("p-value is smaller than 10^-4")
+    return(0.0)
   }
 
[…]

### 2023-04-05 R/PlotResidualPlot.R

@@ -3,6 +3,8 @@
 #' Create a figure that plots the residuals of the data. These residuals are the
 #' difference between expected normal quantiles and observed quantiles.
 #'
+#' @param centre_width A numeric value between 0.0 and 1.0 that describes the
+#'   width of the centre of the data. Can be NULL.
 #' @param show_original Show residuals for original, untransformed, data in
 #'   addition to transformed data.
 #' @param use_alpha Use transparency for points in case the data contains many

### 2023-04-05 manuscript/simulations.R

@@ -1241,10 +1241,43 @@
 
 
 
-.get_test_statistics_data <- function(manuscript_dir){
+.get_test_statistics_data <- function(manuscript_dir, as_table = TRUE, reduce = FALSE){
 
   # Non-standard evaluation in data.table.
-  residual <- p <- n <- mare <- method <- rejected <- NULL
+  residual <- p <- n <- test_statistic <- method <- alpha <- NULL
+
+  ..down_sample_data <- function(x, alpha, upper, lower, n){
+    # Find the x-values.
+    alpha_out <- mapply(
+      function(upper, lower, n){
+        # Linear placement of interpolation points. The starting point (upper)
+        # is not included.
+        x <- rev(seq(from = lower, to = upper, length.out = n + 1L))
+        return(tail(x, n = n))
+      },
+      upper = upper,
+      lower = lower,
+      n = n,
+      SIMPLIFY = FALSE,
+      USE.NAMES = FALSE
+    )
+
+    alpha_out <- unlist(alpha_out, use.names = FALSE)
+
+    # Find the corresponding
+    x_out <- stats::spline(
+      x = alpha,
+      y = x,
+      method = "hyman",
+      xout = alpha_out
+    )$y
+
+    # Add start and end points.
+    x_out <- c(min(x), x_out, max(x))
+    alpha_out <- c(1.0, alpha_out, 0.0)
+
+    return(list("alpha" = alpha_out, "test_statistic" = x_out))
+  }
 
   # Compute test statistic values.
   data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
@@ -1260,35 +1293,62 @@
   data <- data[p >= p_lower & p <= p_upper]
 
   # Compute mean absolute residual error per feature.
-  data <- data[, list("mare"=mean(residual)), by=c("distribution_id", "outlier_id", "method")]
-  data <- data[, list("n"=.N), by=c("mare", "method")][order(mare, method)]
-  data[, "rejected" := 1.0 - cumsum(n) / sum(n), by = "method"]
-  data <- rbind(
-    data.table::data.table(
-      "mare"=c(0.0, 0.0),
-      "method"=c("Box-Cox", "Yeo-Johnson"),
-      "n"=0L,
-      "rejected"=1.0),
-    data)
+  data <- data[, list("test_statistic" = mean(residual)), by=c("distribution_id", "outlier_id", "method")]
[…]

### 2023-04-05 tests/testthat/test-goodness-of-fit.R

@@ -1,12 +1,12 @@
 parameter_list <- list()
 ii <- 1
-for(method in c("box_cox", "yeo_johnson", "none")){
-  for(robust in c(FALSE, TRUE)){
-    for(shift in c(FALSE, TRUE)){
+for (method in c("box_cox", "yeo_johnson", "none")) {
+  for (robust in c(FALSE, TRUE)) {
+    for (shift in c(FALSE, TRUE)) {
       parameter_list[[ii]] <- list(
-        "method"=method,
-        "robust"=robust,
-        "shift"=shift
+        "method" = method,
+        "robust" = robust,
+        "shift" = shift
       )
 
       ii <- ii + 1
@@ -15,8 +15,7 @@ for(method in c("box_cox", "yeo_johnson", "none")){
 }
 
 # Iterate over different sample numbers.
-for(n in c(100, 1000, 10000)){
-
+for (n in c(100, 1000, 10000)) {
   # Set seed.
   set.seed(19L)
 
@@ -28,31 +27,33 @@ for(n in c(100, 1000, 10000)){
 
   # Draw bimodal with large separation
   x_bimodal_large <- c(
-    stats::rnorm(n / 2, mean=0.0),
-    stats::rnorm(n / 2, mean=6.0))
+    stats::rnorm(n / 2, mean = 0.0),
+    stats::rnorm(n / 2, mean = 6.0)
+  )
 
   # Draw bimodal with intermediate separation
   x_bimodal_intermediate <- c(
-    stats::rnorm(n / 2, mean=0.0),
-    stats::rnorm(n / 2, mean=3.0))
+    stats::rnorm(n / 2, mean = 0.0),
+    stats::rnorm(n / 2, mean = 3.0)
+  )
 
   # Draw bimodal with small separation
   x_bimodal_small <- c(
-    stats::rnorm(n / 2, mean=0.0),
-    stats::rnorm(n / 2, mean=1.0))
+    stats::rnorm(n / 2, mean = 0.0),
+    stats::rnorm(n / 2, mean = 1.0)
+  )
 
   # Draw uniform.
-  x_uniform <- stats::runif(n, min=1E-5)
+  x_uniform <- stats::runif(n, min = 1E-5)
 
   # Draw Gamma (left skewed)
-  x_gamma_left <- stats::rgamma(n, shape=2.0, scale=2.0)
+  x_gamma_left <- stats::rgamma(n, shape = 2.0, scale = 2.0)
 
   # Draw Gamma (monotonic decreasing)
[…]

### 2023-04-05 manuscript/manuscript.Rmd
WIP on multi-modal distribution plot.

@@
 
--   We define an empirical measure of the goodness of fit for detecting cases when power transformations fail to yield a normally distributed transformed feature.
+-   We define an empirical goodness of fit test for detecting cases where power transformations fail to yield a normally distributed transformed feature.
 
@@
 Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
-This leads to test statistic values listed in Table *insert*, which derive from the type I error rates of the Box-Cox transformation.
+This leads to test statistic values listed in Table *insert*, which derive from the type I error rates of the Yeo-Johnson transformation.
 
@@
 cap <- paste0(
-  "Rejected transformed asymmetric generalised normal distributions as function of ",
-  "the mean absolute residual error of the central portion $\\kappa$ of the distribution.")
+  "Type 1 error rate of transformed asymmetric generalised normal distributions as function of ",
+  "the test statistic for the central portion $\\kappa$ of the distribution.")
 
@@
 ```{r include = FALSE, eval = FALSE}
-table_data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
-```
-
-```{r include = FALSE, eval = FALSE}
+# This writes the lookup table for the goodness of fit test to the
+# power.transform package.
 load(file = "./R/sysdata.rda")
 
-gof_lookup_table <- .get_test_statistics_data(manuscript_dir = manuscript_dir, reduce = TRUE, as_table = FALSE)
+gof_lookup_table <- .get_test_statistics_data(
+  manuscript_dir = manuscript_dir,
+  reduce = TRUE,
+  as_table = FALSE)
 gof_lookup_table[, "method" := NULL]
@@
   overwrite = TRUE)
+```
+```{r include = FALSE, eval = FALSE}
+# This generates the data for the table below.
+table_data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
 ```
@@
 \end{table}
+```
+
+Figure *insert* shows bi-modal distributions, and their respective p-values using the empirical goodness of fit test.
+
+```{r goodness-of-fit-3, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0("Bi-modal distributions and empirical goodness of fit test results.")
+
+.plot_bimodal_distribution_test(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir)
+
 ```

### 2023-04-06 manuscript/manuscript.Rmd
Empirical goodness of fit test is now derived from distributions with fewer instances.

@@
 ```
+
 ```{r include = FALSE, eval = FALSE}
@@
 \midrule
-0.029 & 0.039 & 0.047 & 0.054 & 0.062 & 0.069 & 0.090 \\
+0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
 \bottomrule

### 2023-04-06 manuscript/manuscript.Rmd
Added configuration parameters.

@@
 
-First we draw $m_d=10000$ random asymmetric generalised normal distributions, parametrised as above.
+First we draw $m_d=10000$ random asymmetric generalised normal distributions.
+As before, each distribution is parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+Location and scale parameters are set as $\mu = 0$ and $\sigma = 1$, respectively.
+$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(1.47, 3.00\right)$, which leads to between $30$ and $1000$ values being drawn to create $\mathbf{X}_i$.
 As before, for each distribution, up to $10 \%$ of instances are replaced by outlier values, and this is repeated $10$ times.

### 2023-04-06 manuscript/manuscript.Rmd
Finished simulation section.

@@
 
-Figure *insert* shows bi-modal distributions, and their respective p-values using the empirical goodness of fit test.
+In figure *insert* we apply the empirical goodness of fit test to assess central normality of features that are composed of a mixture of samples drawn from two normal distributions.
+With increased separation of the underlying normal distributions, the probability of the feature being centrally normal decreases, as expected.
 
 ```{r goodness-of-fit-3, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0("Bi-modal distributions and empirical goodness of fit test results.")
+cap <- paste0(
+  "Bi-modal distributions and empirical goodness of fit test results. ",
+  "The feature (black) is a mixture of two identical sample sets (blue and orange) ",
+  "drawn from normal distributions that are offset by a distance $d$. ",
+  "We use the empirical goodness of fit test to compute p-values for the hypothesis that ",
+  "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
+  "the probability that the feature is centrally normal decreases. ",
+  "Quantile-quantile plots are drawn below each distribution."
+)
 

### 2023-04-06 manuscript/plots.R

@@ -1077,10 +1077,11 @@ get_annotation_settings <- function(ggtheme = NULL) {
 .plot_bimodal_distribution_test <- function(plot_theme, manuscript_dir) {
 
   # Lambda plot,
-  .create_density_plot <- function(
+  .create_density_and_qq_plot <- function(
     location_shift = 0.0,
     plot_theme,
-    limits = c(-6.0, 16.0)) {
+    limits = c(-7.0, 7.0),
+    drop_qq_y_axis = TRUE) {
     # Prevent warnings due to non-standard evaluation.
     distribution <- NULL
 
@@ -1089,7 +1090,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     # Normal distribution.
     x_1 <- power.transform::ragn(
       10000L,
-      location = 0,
+      location = -location_shift / 2,
       scale = 1 / sqrt(2),
       alpha = 0.5,
       beta = 2
@@ -1124,32 +1125,38 @@ get_annotation_settings <- function(ggtheme = NULL) {
     annotation_settings <- get_annotation_settings(plot_theme)
 
     p <- ggplot2::ggplot(
-      data = data,
+      data = data)
+    p <- p + plot_theme
+    p <- p + ggplot2::geom_density(
       mapping = ggplot2::aes(
         x = x,
+        y = ggplot2::after_stat(scaled),
         fill = distribution,
-        colour = distribution)
-    )
-    p <- p + plot_theme
-    p <- p + ggplot2::geom_density(alpha = 0.2)
+        colour = distribution),
+      alpha = 0.2)
+    p <- p + ggplot2::geom_density(
+      mapping = ggplot2::aes(
+        x = x,
+        y = ggplot2::after_stat(scaled)),
+      colour = "grey10")
 
     p <- p + ggplot2::geom_segment(
-      x = 0.0,
-      xend = location_shift,
+      x = -location_shift / 2.0,
+      xend = location_shift / 2.0,
       y = 0.10,
       yend = 0.10,
       colour = "grey40")
     p <- p + ggplot2::annotate(
       geom = "text",
-      x = location_shift / 2.0,
+      x = 0.0,
       y = 0.1,
       label = p_value,
       colour = "grey40",
       family = annotation_settings$family,
       fontface = annotation_settings$face,
[…]

### 2023-04-06 manuscript/manuscript.Rmd
Outline experimental section.

@@
 # Experimental Results
+
+## Datasets with outliers
+
+### Car data
+
+Top gear dataset, weight
+
+### Housing data
+
+Ames housing data set, year built
+
+### Other data ???
+
+## Integration into end-to-end machine learning
+
+- Assess whether shift-invariant robust transformation improves machine learning results.
+
+  - No transformation.
+  - Conventional Yeo-Johnson transformation.
+  - Shift sensitive Yeo-Johnson transformation.
+  - Shift sensitive, robust Yeo-Johnson transformation.
+  - Shift sensitive, robust Yeo-Johnson with central normality test.
+
+  TODO: 
+  - Add set_lambda, get_lambda, set_shift, get_shift methods.
+  - Add option to set p-value for central normality test. Default NULL (no filtering)
+  - implement into familiar
 

### 2023-04-10 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
   - Add set_lambda, get_lambda, set_shift, get_shift methods.
-  - Add option to set p-value for central normality test. Default NULL (no filtering)
   - implement into familiar
+  - update familiar.experiment to allow for setting external configurations (or maybe make that the default -- its definitely easier and more transparent.)
+
+  - Add tests for out-of-bound actions (power_transform) and GOF test rejection (find_parameters).
 
@@
 Shift-sensitive power transformations are implemented in the `power.transform` package.
+This manuscript was created using R Markdown, and is available from [https://github.com/alexzwanenburg/power.transform].
 

### 2023-04-11 manuscript/manuscript.Rmd
Updated united tests to identify classed warnings.

@@
   TODO: 
-  - Add set_lambda, get_lambda, set_shift, get_shift methods.
   - implement into familiar

### 2023-04-11 R/FindParameters.R

@@ -163,8 +163,8 @@ find_transformation_parameters <- function(
       rlang::warn(
         message = paste0(
           "The p-value of the transformed data (", gof_test_p, ") is below the required ",
-          "significance level (", empirical_gof_normality_p_value, ").
-          The transformation is rejected, and data are kept as is."),
+          "significance level (", empirical_gof_normality_p_value, "). ",
+          "The transformation is rejected, and data are kept as is."),
         class = "power_transform_no_transform")
 
       object <- methods::new("transformationNone")

### 2023-04-11 tests/testthat/test-estimators.R

@@ -186,7 +186,7 @@ for(ii in seq_along(parameter_list)){
           power.transform::find_transformation_parameters,
           args=c(list("x"=x_few_unique),
                  parameter_list[[ii]])),
-        "x contains ten or fewer unique values")
+        class = "power_transform_few_unique_values")
 
       # Check lambda values.
       if(parameter_list[[ii]]$method == "box_cox"){

### 2023-04-11 tests/testthat/test-transformation.R

@@ -537,7 +537,7 @@ for(ii in seq_along(parameter_list)){
             power.transform::find_transformation_parameters,
             args=c(list("x"=x_single_value),
                    parameter_list[[ii]])),
-          "x contains three or fewer unique values")
+          class = "power_transform_no_transform")
       }
 
       testthat::expect_s4_class(transformer, "transformationNone")
@@ -564,13 +564,14 @@ for(ii in seq_along(parameter_list)){
             power.transform::find_transformation_parameters,
             args=c(list("x"=x_three_unique),
                    parameter_list[[ii]])))
+
       } else {
         testthat::expect_warning(
           transformer <- do.call(
             power.transform::find_transformation_parameters,
             args=c(list("x"=x_three_unique),
                    parameter_list[[ii]])),
-          "x contains three or fewer unique values")
+          class = "power_transform_no_transform")
       }
 
       testthat::expect_s4_class(transformer, "transformationNone")
@@ -596,7 +597,7 @@ for(ii in seq_along(parameter_list)){
           power.transform::find_transformation_parameters,
           args=c(list("x"=x_few_unique),
                  parameter_list[[ii]])),
-        "x contains ten or fewer unique values")
+        class = "power_transform_few_unique_values")
 
       # Check lambda values.
       if(parameter_list[[ii]]$method == "box_cox"){

### 2023-11-01 manuscript/manuscript.Rmd
Fixes to manuscript.

@@
 ```
-One limitation of th Box-Cox transformation is that it is only defined for $x_i > 0$.
+One limitation of the Box-Cox transformation is that it is only defined for $x_i > 0$.
 In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$:
@@
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and even breaks down due to numerical issues for $\\mu > 5 10^4$.")
+  "and even breaks down due to numerical issues for $\\mu > 5 \cdot 10^4$.")
 
@@
 \begin{equation}
-\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
+\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
 \end{equation}
@@
 \begin{equation}
-\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
+\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
 \end{equation}
@@
 
-- Set weights based on z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$.
+- Set weights based on the z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$.
 After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$.
@@
 
-All weighting functions share the characteristic that for $\dot{x}_i < k_1$, instances are fully weighted, i.e.
+All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< k_1$, instances are fully weighted, i.e.
 when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top.
-The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \dot{x}_i \leq k_2$, and assign no weight to instances $x_i > k_2$.
+The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.
 
@@
 ```
-Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **insert**.
+Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **INSERT**.
 
@@
 \begin{equation}
-U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2}
+U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2
 \end{equation}

### 2023-11-01 manuscript/plots.R

@@ -792,6 +792,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   } else {
     data <- readRDS(file.path(manuscript_dir, "residual_error_plot_main_manuscript.RDS"))
+
+    outlier_free_data <- data[
+      has_outliers == FALSE,
+      list("residual_error" = stats::quantile(residual, probs = 0.99)),
+      by = c("p", "method")
+    ]
   }
 
   p_bc <- ggplot2::ggplot(

### 2023-11-01 manuscript/manuscript.Rmd
WIP

@@
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and even breaks down due to numerical issues for $\\mu > 5 \cdot 10^4$.")
+  "and even breaks down due to numerical issues for $\\mu > 5 \\cdot 10^4$.")
 
@@
 Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
-Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option to simulate realistic feature distributions.
+Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option for simulating realistic feature distributions.
 This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:
@@
 
-The optimal weighting function parameters for weighted MLE are shown in Tables **insert** and **insert** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
+The optimal weighting function parameters for weighted MLE are shown in Tables **INSERT** and **INSERT** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 
-Figure **insert** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
+Figure **INSERT** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
 In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation.
@@
 ### Car data
+
+The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*. Fuel consumption contains outliers.
+
+```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Insert caption here"
+)
+
+.plot_top_gear(
+  plot_theme = plot_theme
+)
+
+```
 

### 2023-11-01 manuscript/plots.R

@@ -1226,3 +1226,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
   return(p)
 }
+
+
+.plot_top_gear <- function(plot_theme) {
+
+  # Plot MPG. Show Q-Q plots for:
+  # - conventional transformation
+  # - Raymaekers
+  # - Shift-sensitive robust
+}

### 2023-11-03 manuscript/manuscript.Rmd
Work on manuscript.

@@
 
-## Goodness of fit
+## Empirical central normality test
 
@@
 
-## Empirical goodness of fit test
-
-To develop an empirical goodness of fit test we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{gof}}$.
+## Empirical central normality test
+
+To develop an empirical test for central normality we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{gof}}$.
 We will first define the central portion $\kappa$.
@@
 ```
+
+**MISSING TABLE DATA FOR YEO-JOHNSON**
 
@@
 
+In this section, we assess shift sensitive power transformations using real datasets.
+
 ## Datasets with outliers
 
-### Car data
-
-The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*. Fuel consumption contains outliers.
-
-```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap}
+### Top Gear dataset
+
+```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
 cap <- paste0(
-  "Insert caption here"
+  "\\label{fig:top-gear}",
+  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
+  "This data contains three strong outliers for highly efficient vehicles. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust shift-sensitive transformations. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
+  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
 )
@@
 
-Top gear dataset, weight
-
-### Housing data
+The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*.
+Fuel consumption contains outliers.
+Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$).
+Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
+
+### Housing dataset
+
+modeldata::Sacramento$price
 

### 2023-11-03 manuscript/plots.R

@@ -1238,5 +1238,168 @@ get_annotation_settings <- function(ggtheme = NULL) {
   #
   # Compute empirical test p-value.
 
+  plot_limits <- c(-3.0, 3.0)
 
+  ..parse_data <- function(
+    x,
+    transformer,
+    name
+  ) {
+    residual_data <- power.transform::get_residuals(
+      x = x,
+      transformer = transformer
+    )
+
+    residual_data[, "transformation" := name]
+
+    central_normality_data <- data.table::data.table(
+      p_value = power.transform::assess_transformation(
+        x = x,
+        transformer = transformer,
+        verbose = FALSE
+      ),
+      transformation = name
+    )
+
+    return(list(
+      "residual_data" = residual_data,
+      "central_normality_data" = central_normality_data
+    ))
+  }
+
+  # Load TopGear data into the current environment
+  utils::data("TopGear", package = "robustHD", envir = environment())
+  data <- data.table::as.data.table(TopGear)
+
+  # Extract MPG data.
+  x <- data$MPG
+  x <- x[!is.na(x)]
+
+  # No transformer
+  no_transformer <- power.transform::find_transformation_parameters(
+    x = x,
+    method = "none"
+  )
+
+  # Conventional transformer
+  yj_transformer <- power.transform::find_transformation_parameters(
+    x = x,
+    method = "yeo_johnson",
+    robust = FALSE,
+    shift = FALSE
+  )
+
+  # Raymaekers robust transformer
+  yj_rr_transformer <- power.transform::find_transformation_parameters(
+    x = x,
+    method = "yeo_johnson",
+    robust = TRUE,
+    shift = FALSE,
+    estimation_method = "raymaekers_robust"
+  )
+
+  # Shift-sensitive transformer
+  yj_shift_transformer <- power.transform::find_transformation_parameters(
+    x = x,
+    method = "yeo_johnson",
+    robust = FALSE,
+    shift = TRUE
+  )
+
+  # Robust shift-sensitive transformer
+  yj_shift_robust_transformer <- power.transform::find_transformation_parameters(
+    x = x,
+    method = "yeo_johnson",
+    robust = TRUE,
+    shift = TRUE
+  )
+
+  transformer_labels <- c(
+    "none", "original", "Raymaekers-Rousseeuw", "shift-sensitive", "robust shift-sensitive"
+  )
+
[…]

### 2023-11-03 manuscript/manuscript.Rmd
Added references.

@@
 Power transformation does not guarantee that transformed variables are normally distributed.
-In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure **INSERT**.
+In fact, depending on location and scale of the variable, power transformation may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
 This is problematic for automated power transformations, for three reasons.
@@
 
-```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
+```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
 cap <- paste0(
+  "\\label{fig:decreased-normality}",
   "Power transformations may reduce normality depending on location. ",
@@
 The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.
-
-```{r weight-functions, warning=FALSE, message=FALSE, fig.cap=cap}
+Examples of the three types of weighting function are shown in Figure \ref{fig:weighting-functions}.
+
+```{r weighting-functions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
 cap <- paste0(
+  "\\label{fig:weighting-functions}",
   "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
@@
 cap <- paste0(
+  "\\label{fig:shifted-distributions}",
   "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
@@
 
-The result is shown in Figure **INSERT**.
+The result is shown in Figure \ref{fig:shifted-distributions}.
 For each distribution, transformation parameter $\hat{\lambda}$ varies with $d$ when estimated using the original definitions.

### 2023-11-03 manuscript/manuscript.Rmd
Added internal figure references.

@@
 \begin{equation}
+\label{eqn:box-cox-original}
 \phi_{\text{BC}}^\lambda (x_i) = 
@@
 \begin{equation}
+\label{eqn:yeo-johnson-original}
 \phi_{\text{YJ}}^\lambda (x_i) = 
@@
 
-Box-Cox and Yeo-Johnson transformations are modified by adding a shift parameter $x_0$.
+Box-Cox and Yeo-Johnson transformations are modified by introducing a shift parameter $x_0$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
 The shift-sensitive Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$ and shift parameter $x_0$ is then:
@@
 \begin{equation}
+\label{eqn:box-cox-shift}
 \phi_{\text{BC}}^{\lambda, x_0} (x_i) = 
@@
 \begin{equation}
+\label{eqn:yeo-johnson-shift}
 \phi_{\text{YJ}}^{\lambda, x_0} (x_i) = 
@@
 \begin{equation}
+\label{eqn:box-cox-shift-log-likelihood}
 \left\{ \hat{\lambda}^{\text{BC}}, \hat{x}_0^{\text{BC}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \log(x_i - x_0)
@@
 \begin{equation}
+\label{eqn:yeo-johnson-shift-log-likelihood}
 \left\{ \hat{\lambda}^{\text{YJ}}, \hat{x}_0^{\text{YJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|)
@@
 @Raymaekers2021-kq do this (during the final steps of their iterative procedure) through a weighted maximum likelihood estimation.
-This weighted maximum log-likelihood estimator is based on equations **INSERT**.
+This weighted maximum log-likelihood estimator is based on equations \ref{eqn:box-cox-shift-log-likelihood} and \ref{eqn:yeo-johnson-shift-log-likelihood}.
 Here, the estimator is further normalised by the total weight and extended to include the shift parameter.
@@
 ```
+
+Other criteria aside from MLE can be used for finding robust transformation parameters, see the Appendix. For the remainder, we will use MLE as the optimisation criterion.
 
@@
 ```
-Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **INSERT**.
 
@@
 
-The test statistic $\tau_{\text{gof}}$ is then defined as:
-```{=latex}
-\begin{equation}
-\tau_{\text{gof}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
+The test statistic $\tau_{\text{ecn}}$ is then defined as:
+```{=latex}
+\begin{equation}
+\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
 \end{equation}
@@
 To assess whether the modified power transformations lead to $\lambda$ values that are invariant to shifts, we simulate three different distributions.
-The first distribution is a normal distribution, as previously shown in Figure **INSERT**.
+The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
 We first randomly draw $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
@@
 The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then compute transformation parameter $\hat{\lambda}$ using the original definitions (Eq.
-**INSERT** and **INSERT**) and the shift-sensitive definitions (Eq.
-**INSERT** and **INSERT**) for each distribution, after adding a positive value $d$ to each distribution.
+We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{box-cox-original} and \ref{yeo-johnson-original}) and the shift-sensitive definitions (equations \ref{box-cox-shift} and \ref{yeo-johnson-shift}) for each distribution, after adding a positive value $d$ to each distribution.
 Here, $d \in [1, 10^6]$.
@@
 
-The optimal weighting function parameters for weighted MLE are shown in Tables **INSERT** and **INSERT** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
+The optimal weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 \begin{center}
+\label{tab:optimal-weighting-parameters-box-cox}
 \caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Box-Cox} power transformations.
@@
 \begin{center}
+\label{tab:optimal-weighting-parameters-yeo-johnson}
 \caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Yeo-Johnson} power transformations.
@@
 
-Figure **INSERT** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
+Figure \ref{fig:optimised-weighting-function-parameters} shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
 In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation.
@@
 
-```{r optimised_weighting_function_parameters, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r optimised-weighting-function-parameters, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
+  "\\label{fig:optimised-weighting-function-parameters}",
   "Robustness of power transformations after optimising weighting function parameters. ",
@@
 
-Figure *Insert reference* shows the residual errors of the transformed distribution as function of the empirical probability.
+Figure \ref{tab:empirical-central-normality-residual-error} shows the residual errors of the transformed distribution as function of the empirical probability.
 As expected, the largest deviations from normality appear on the extremities of this range.
@@
 
-```{r goodness-of-fit-1, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
+  "\\label{tab:empirical-central-normality-residual-error}",
   "Residual error as a function of empirical probability. ",
@@
 We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic.
-Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure *Insert*.
+Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{tab:empirical-central-normality-type-1-error-rate}.
 Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
-This leads to test statistic values listed in Table *insert*, which derive from the type I error rates of the Yeo-Johnson transformation.
-
-```{r goodness-of-fit-2, warning=FALSE, message=FALSE, fig.cap=cap}
+This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}, which derive from the type I error rates of the Yeo-Johnson transformation.
+
+```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
+  "\\label{tab:empirical-central-normality-type-1-error-rate}",
   "Type 1 error rate of transformed asymmetric generalised normal distributions as function of ",
@@
 \begin{center}
+\label{tab:empirical-central-normality}
 \caption{Probability of central normality, based on type I error rates for Box-Cox transformations.}
@@
 
-In figure *insert* we apply the empirical goodness of fit test to assess central normality of features that are composed of a mixture of samples drawn from two normal distributions.
+In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical central normality test to assess central normality of features that are composed of a mixture of samples drawn from two normal distributions.
 With increased separation of the underlying normal distributions, the probability of the feature being centrally normal decreases, as expected.
 
-```{r goodness-of-fit-3, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
+  "\\label{fig:empirical-central-normality-examples}",
   "Bi-modal distributions and empirical goodness of fit test results. ",

### 2023-11-06 manuscript/manuscript.Rmd
Initial implementation of sensitivity-to-outlier plots.

@@
   "Power transformations may reduce normality depending on location. ",
-  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
+  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
   "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
@@
 
+```{r sensitivity-to-outliers, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
+cap <- paste0(
+  "\\label{fig:sensitivity-to-outliers}",
+  "Power transformations are affected by outliers. ",
+  "Variable $x$ consists of $1000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$, ",
+  "with a single outlier with value $d$. ",
+  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
+  "The figure shows that instead, $\\lambda$ is strongly affected by the presence of the outlier at $d$, ",
+  "and even breaks down due to numerical issues for $d > 5 \\cdot 10^4$.")
+
+.plot_sensitivity_to_outlier(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
+
+```
+
 Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
-Their presence may lead to poor transformations to normality, as shown in Figure **INSERT**.
+Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:sensitivity-to-outliers}.
 As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e.

### 2023-11-06 manuscript/simulations.R

@@ -165,6 +165,101 @@
 
 
 
+.get_shifted_outlier_plot_data <- function(manuscript_dir) {
+  # Set seed.
+  set.seed(19L)
+
+  if (!file.exists(file.path(manuscript_dir, "shifted_outlier_plot.RDS"))) {
+    # Normal distribution.
+    x <- power.transform::ragn(1000L, location = 0, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
+
+    # generator ----------------------------------------------------------------
+    generate_experiment_data <- coro::generator(
+      function(x) {
+        outlier_shift_range <- 10^seq(from = 0, to = 6, by = 0.1)
+
+        for (d in shift_range) {
+          for (method in c("box_cox", "yeo_johnson")) {
+            yield(list(
+              "x" = c(x, d),
+              "d" = d,
+              "method" = method,
+              "shift" = FALSE,
+              "robust" = FALSE
+            ))
+          }
+        }
+      }
+    )
+
+    .compute_lambda <- function(parameter_set) {
+      # Create transformer object.
+      transformer <- suppressWarnings(
+        power.transform::find_transformation_parameters(
+          x = parameter_set$x,
+          method = parameter_set$method,
+          robust = parameter_set$robust,
+          shift = parameter_set$shift,
+          lambda = NULL,
+          optimiser_control = list("xtol_rel" = 1e-5)
+        )
+      )
+
+      return(
+        data.table::data.table(
+          "method" = parameter_set$method,
+          "d" = log10(parameter_set$d),
+          "lambda" = transformer@lambda,
+          "x_0" = transformer@shift
+        )
+      )
+    }
+
+    # computations -------------------------------------------------------------
+
+    # Generate all experiments.
+    experiments <- coro::collect(generate_experiment_data(x=x))
+
+    data <- lapply(
+      X = experiments,
+      FUN = .compute_lambda
+    )
+
+    # # Start cluster
+    # cl <- parallel::makeCluster(16L)
+    #
+    # # Compute all data in parallel.
+    # data <- parallel::parLapply(
+    #   cl = cl,
+    #   X = experiments,
+    #   fun = .compute_lambda
+    # )
+    #
+    # # Stop cluster.
+    # parallel::stopCluster(cl)
+
+    # Combine data into a single table.
+    data <- data.table::rbindlist(data)
+
+    # Save to file.
[…]

### 2023-11-07 manuscript/plots.R
Added plot that shows effect of outliers.

@@ -267,7 +267,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   }
 
   # Prevent warnings due to non-standard evaluation.
-  estimation_method <- distribution <- method <- version <- NULL
+  method <- NULL
 
   # Process data.
   data <- .get_shifted_outlier_plot_data(manuscript_dir = manuscript_dir)
@@ -278,7 +278,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p_bc_normal <- .create_lambda_shift_plot(
     data = data[method == "Box-Cox"],
     plot_theme = plot_theme,
-    limits = c(-5.0, 35.0),
+    limits = c(-0.2, 1.2),
     strip_y_axis = FALSE
   )
 
@@ -286,7 +286,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p_yj_normal <- .create_lambda_shift_plot(
     data = data[method == "Yeo-Johnson"],
     plot_theme = plot_theme,
-    limits = c(-5.0, 35.0),
+    limits = c(-0.2, 1.2),
     guide = TRUE
   )
 

### 2023-11-07 manuscript/simulations.R

@@ -165,23 +165,23 @@
 
 
 
-.get_shifted_outlier_plot_data <- function(manuscript_dir) {
+.get_shifted_outlier_plot_data <- function(manuscript_dir, offset = 0.0) {
   # Set seed.
   set.seed(19L)
 
   if (!file.exists(file.path(manuscript_dir, "shifted_outlier_plot.RDS"))) {
     # Normal distribution.
-    x <- power.transform::ragn(1000L, location = 0, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
+    x <- power.transform::ragn(1000L, location = offset, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
 
     # generator ----------------------------------------------------------------
     generate_experiment_data <- coro::generator(
-      function(x) {
-        outlier_shift_range <- 10^seq(from = 0, to = 6, by = 0.1)
+      function(x, offset) {
+        outlier_shift_range <- 10^seq(from = -1, to = 6, by = 0.1)
 
-        for (d in shift_range) {
+        for (d in outlier_shift_range) {
           for (method in c("box_cox", "yeo_johnson")) {
             yield(list(
-              "x" = c(x, d),
+              "x" = c(x, offset + d),
               "d" = d,
               "method" = method,
               "shift" = FALSE,
@@ -200,7 +200,6 @@
           method = parameter_set$method,
           robust = parameter_set$robust,
           shift = parameter_set$shift,
-          lambda = NULL,
           optimiser_control = list("xtol_rel" = 1e-5)
         )
       )
@@ -218,7 +217,7 @@
     # computations -------------------------------------------------------------
 
     # Generate all experiments.
-    experiments <- coro::collect(generate_experiment_data(x=x))
+    experiments <- coro::collect(generate_experiment_data(x=x, offset = offset))
 
     data <- lapply(
       X = experiments,

### 2023-11-07 manuscript/manuscript.Rmd
Added plot that shows effect of outliers.
Fixed labels.

@@
 cap <- paste0(
-  "\\label{fig:decreased-normality}",
   "Power transformations may reduce normality depending on location. ",
@@
 cap <- paste0(
-  "\\label{fig:sensitivity-to-outliers}",
   "Power transformations are affected by outliers. ",
@@
   "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
-  "The figure shows that instead, $\\lambda$ is strongly affected by the presence of the outlier at $d$, ",
-  "and even breaks down due to numerical issues for $d > 5 \\cdot 10^4$.")
+  "The figure shows that instead, $\\lambda$ is strongly affected by the presence of the outlier at $d$.")
 
@@
 cap <- paste0(
-  "\\label{fig:weighting-functions}",
   "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
@@
 Gijbels et al.
-describe a recipe for introducing skewness into the otherwise symmetric generalised norm distribution [@Gijbels2019-te], leading to the following probability density function:
+describe a recipe for introducing skewness into the otherwise symmetric generalised normal distribution [@Gijbels2019-te], leading to the following probability density function:
 
@@
 The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{box-cox-original} and \ref{yeo-johnson-original}) and the shift-sensitive definitions (equations \ref{box-cox-shift} and \ref{yeo-johnson-shift}) for each distribution, after adding a positive value $d$ to each distribution.
+We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the shift-sensitive definitions (equations \ref{eqn:box-cox-shift} and \ref{eqn:yeo-johnson-shift}) for each distribution, after adding a positive value $d$ to each distribution.
 Here, $d \in [1, 10^6]$.
@@
 cap <- paste0(
-  "\\label{fig:shifted-distributions}",
   "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
@@
 \begin{center}
-\label{tab:optimal-weighting-parameters-box-cox}
 \caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Box-Cox} power transformations.
@@
 Lower loss indicates better robustness against outliers.}
+\label{tab:optimal-weighting-parameters-box-cox}
 \begin{tabular}{l r r r r r r r}
@@
 \begin{center}
-\label{tab:optimal-weighting-parameters-yeo-johnson}
 \caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Yeo-Johnson} power transformations.
@@
 Lower loss indicates better robustness against outliers.}
+\label{tab:optimal-weighting-parameters-yeo-johnson}
 \begin{tabular}{l r r r r r r r}
@@
 cap <- paste0(
-  "\\label{fig:optimised-weighting-function-parameters}",
   "Robustness of power transformations after optimising weighting function parameters. ",
@@
 
-Figure \ref{tab:empirical-central-normality-residual-error} shows the residual errors of the transformed distribution as function of the empirical probability.
+Figure \ref{fig:empirical-central-normality-residual-error} shows the residual errors of the transformed distribution as function of the empirical probability.
 As expected, the largest deviations from normality appear on the extremities of this range.
@@
 cap <- paste0(
-  "\\label{tab:empirical-central-normality-residual-error}",
   "Residual error as a function of empirical probability. ",
@@
 We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic.
-Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{tab:empirical-central-normality-type-1-error-rate}.
+Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}.
 Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
@@
 cap <- paste0(
-  "\\label{tab:empirical-central-normality-type-1-error-rate}",
   "Type 1 error rate of transformed asymmetric generalised normal distributions as function of ",
@@
 
-**MISSING TABLE DATA FOR YEO-JOHNSON**
-
 ```{=latex}
@@
 \begin{center}
+\caption{Test statistic for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
 \label{tab:empirical-central-normality}
-\caption{Probability of central normality, based on type I error rates for Box-Cox transformations.}
-\begin{tabular}{c c c c c c c}
+\begin{tabular}{l c c c c c c c}
 
 \toprule
-0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+transformation & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
 
 \midrule
-0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
+rob. shift-sensitive Box-Cox     & 0.046 & 0.072 & 0.090 & 0.106 & 0.126 & 0.140 & 0.189 \\
+rob. shift-sensitive Yeo-Johnson & 0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
 \bottomrule
@@
 cap <- paste0(
-  "\\label{fig:empirical-central-normality-examples}",
   "Bi-modal distributions and empirical goodness of fit test results. ",
@@
 cap <- paste0(
-  "\\label{fig:top-gear}",
   "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",

### 2023-11-07 manuscript/simulations.R

@@ -1471,8 +1471,8 @@
     )
   }
 
-  # Use Yeo-Johnson for test statistics.
-  data <- data[method == "Yeo-Johnson"]
+  # # Use Yeo-Johnson for test statistics.
+  # data <- data[method == "Yeo-Johnson"]
 
   if (as_table) {
     # Find thresholds.

### 2023-11-08 manuscript/manuscript.Rmd
Housing data

@@
 
-To be continued.
-
 # Simulation
@@
 Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}.
-Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80.
-This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}, which derive from the type I error rates of the Yeo-Johnson transformation.
+Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80.
+In the remaining we will use the empirical central normality test statistic values defined using robust shift-sensitive Yeo-Johnson transformations, as it is slightly more strict.
+This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}
 
@@
 \label{tab:empirical-central-normality}
-\begin{tabular}{l c c c c c c c}
+\begin{tabular}{c c c c c c c}
 
 \toprule
-transformation & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
 
 \midrule
-rob. shift-sensitive Box-Cox     & 0.046 & 0.072 & 0.090 & 0.106 & 0.126 & 0.140 & 0.189 \\
-rob. shift-sensitive Yeo-Johnson & 0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
+0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
 \bottomrule
@@
 
-In this section, we assess shift sensitive power transformations using real datasets.
+In the previous section, we simulated data to assess shift-sensitive power transformations and their robustness against outliers. In this section, we assess these aspects using real datasets.
 
@@
 ```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
+cap <- paste0(
+  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
+  "This data contains three strong outliers for highly efficient vehicles. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \cite{Raymaekers2021-kq}. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust shift-sensitive transformations. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
+  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
+)
+
+.plot_top_gear(
+  plot_theme = plot_theme
+)
+```
+
+The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*.
+Fuel consumption contains outliers.
+Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$).
+Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
+
+### Housing dataset
+
+```{r housing-data, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
 cap <- paste0(
@@
 
-.plot_top_gear(
+.plot_sacramento_house_prices(
   plot_theme = plot_theme
 )
-
-```
-
-The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*.
-Fuel consumption contains outliers.
-Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$).
-Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
-
-### Housing dataset
+```
 

### 2023-11-08 manuscript/plots.R

@@ -1344,14 +1344,54 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 .plot_top_gear <- function(plot_theme) {
+  require(robustHD)
 
-  # Plot MPG. Show Q-Q plots for:
+  # Load TopGear data into the current environment
+  utils::data("TopGear", package = "robustHD", envir = environment())
+  data <- data.table::as.data.table(TopGear)
+
+  # Extract MPG data.
+  x <- data$MPG
+  x <- x[!is.na(x)]
+
+  return(
+    ..plot_real_data(
+      plot_theme = plot_theme,
+      x = x
+    )
+  )
+}
+
+
+
+.plot_sacramento_house <- function(plot_theme) {
+  require(modeldata)
+
+  data <- data.table::as.data.table(modeldata::Sacramento)
+  x <- data$price
+
+  return(
+    ..plot_real_data(
+      plot_theme = plot_theme,
+      x = x
+    )
+  )
+}
+
+
+
+..plot_real_data <- function(plot_theme, x) {
+  # Show density and Q-Q plots for:
+  # - Original data
   # - conventional transformation
   # - Raymaekers
   # - Shift-sensitive
   # - Shift-sensitive robust
   #
   # Compute empirical test p-value.
+
+  # Prevent undue warnings.
+  transformation <- z_observed <- NULL
 
   plot_limits <- c(-3.0, 3.0)
 
@@ -1360,6 +1400,17 @@ get_annotation_settings <- function(ggtheme = NULL) {
     transformer,
     name
   ) {
+    transformed_data <- power.transform::power_transform(
+      x = x,
+      transformer = transformer
+    )
+
+    transformed_data <- data.table::data.table(
+      x = transformed_data,
+      x_standardised = (transformed_data - stats::median(transformed_data)) / stats::IQR(transformed_data),
+      transformation = name
+    )
+
     residual_data <- power.transform::get_residuals(
       x = x,
       transformer = transformer
@@ -1377,18 +1428,11 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
 
     return(list(
+      "transformed_data" = transformed_data,
       "residual_data" = residual_data,
       "central_normality_data" = central_normality_data
     ))
   }
-
-  # Load TopGear data into the current environment
[…]

### 2023-11-08 manuscript/simulations.R

@@ -1471,8 +1471,8 @@
     )
   }
 
-  # # Use Yeo-Johnson for test statistics.
-  # data <- data[method == "Yeo-Johnson"]
+  # Use Yeo-Johnson for test statistics.
+  data <- data[method == "Yeo-Johnson"]
 
   if (as_table) {
     # Find thresholds.

### 2023-11-10 manuscript/manuscript.Rmd
Added work on deriving maximum likelihood estimation.

@@
 ---
-title: "Shift-sensitive Power Transformations for Transforming Data to Normality"
+title: "Location and Scale-Invariant Power Transformations for Transforming Data to Normality"
 author: "Alex Zwanenburg"
-date: "2023-01-06"
+date: "`r Sys.Date()`"
 output:

### 2023-11-10 manuscript/manuscript_appendix_a.Rmd
Added work on deriving maximum likelihood estimation.

---
title: "Appendix A"
author: "Alex Zwanenburg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parameterised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:

```{=latex}
\begin{equation}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(\frac{x_i - x_0}{s}) & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
Where $x_i - x_0 > 0$.
The location and scale-invariant Yeo-Johnson transformation is:

```{=latex}
\begin{equation}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log(1 + \frac{x_i - x_0}{s}) & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log(1 - \frac{x_i - x_0}{s}) & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}
```

The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
However, the introduction of scaling parameter $s$ prevents their direct use.
Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.

Let $f(x_1, \ldots, x_n)$ be the probability density function of feature $\mathbf{X} = \{ x_1, \ldots, x_n\}$, 
and $f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))$ be the probability density function of 
the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$, that is assumed to follow a normal distribution.

The two probability density functions are related as follows:

```{=latex}
\begin{equation}
f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
\end{equation}
```
Where, $\left|\mathbf{J}\right|$ is the determinant of Jacobian $\mathbf{J}$.
The Jacobian takes the following form, with off-diagonal elements $0$:
```{=latex}
\begin{equation}
\mathbf{J} =
\begin{bmatrix}
    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
    \vdots & \vdots  & \ddots &  \vdots \\
    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
\end{bmatrix}
\end{equation}
```

Thus, $\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)$.

Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
 & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \left|\mathbf{J}\right| \\
 & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
 & = -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)
\end{split}
\end{equation}
```
With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
However, the final term differs between Box-Cox and Yeo-Johnson transformations.

For the Box-Cox transformation the partial derivative is:
```{=latex}
\begin{equation}
\begin{split}
\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
 & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
\end{split}
\end{equation}
```

Thus the final term is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = 
\end{split}
\end{equation}
```

For the Yeo-Johnson transformation, the partial derivative is:
```{=latex}
\begin{equation}
\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
\begin{cases}
\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
\end{cases}
\end{equation}
```

### 2023-11-10 manuscript/manuscript_appendix_a.Rmd
Update manuscript_appendix_a.Rmd

@@
 output: html_document
+header-includes:
+- \usepackage{amsmath}
+- \usepackage{booktabs}
+- \DeclareMathOperator*{\argmax}{argmax}
+- \DeclareMathOperator*{\argmin}{argmin}
+- \DeclareMathOperator{\sgn}{sgn}
 ---
@@
 \left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
-\log(\frac{x_i - x_0}{s}) & \text{if } \lambda = 0
+\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
 \end{cases}
@@
 \left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
-\log(1 + \frac{x_i - x_0}{s}) & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
 -\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
--\log(1 - \frac{x_i - x_0}{s}) & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
+-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
 \end{cases}
@@
 \mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
- & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \left|\mathbf{J}\right| \\
- & = \log f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
- & = -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
+ & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
 \end{split}
@@
 
-Thus the final term is:
+Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
 ```{=latex}
@@
 \begin{split}
-\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = 
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
+& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
+& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+
+This leads to the following log-likelihood for the Box-Cox transformation:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
 \end{split}
@@
 ```
+
+Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+
+This lead to the following log-likelihood for the Yeo-Johnson transformation:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```

### 2023-11-11 R/TransformationBoxCox.R
Updated log likelihood to correct formulation.

@@ -532,7 +532,7 @@ setMethod(
       x = x
     )
 
-    if (estimation_method == "mle") {
+    if (estimation_method %in% c("currently_no_method")) {
       if (length(lambda) == 1) {
         return(
           list(
@@ -598,12 +598,12 @@ setMethod(
     x <- (x - object@shift) / object@scale
 
     # Compute the log likelihood under the assumption that the transformed
-    # variable y follows the normal distribution. Note that the second term
-    # appears here and not in Raymaekers et al. because for scale-invariant
-    # transformations it is not a fixed value.
-    llf <- -sum(w) / 2.0 * log(sigma_hat_squared) +
-      -1.0 / (2.0 * sigma_hat_squared) * sum(w * (y - mu_hat)^2) +
-      (object@lambda - 1.0) * sum(w * log(x))
+    # variable y follows the normal distribution. See manuscript appendix for
+    # derivation.
+    llf <- -sum(w) / 2.0 * log(2 * pi * sigma_hat_squared) +
+           -1.0 / (2.0 * sigma_hat_squared) * sum(w * (y - mu_hat)^2) +
+           -sum(w) * log(object@scale) +
+           (object@lambda - 1.0) * sum(w * log(x))
 
     return(llf)
   }

### 2023-11-11 R/TransformationYeoJohnson.R

@@ -478,7 +478,7 @@ setMethod(
       x = x
     )
 
-    if (estimation_method == "mle") {
+    if (estimation_method %in% c("currently_no_method")) {
       if (length(lambda) == 1) {
         return(
           list(
@@ -544,11 +544,11 @@ setMethod(
     x <- (x - object@shift) / object@scale
 
     # Compute the log likelihood under the assumption that the transformed
-    # variable y follows the normal distribution. Note that the second term
-    # appears here and not in Raymaekers et al. because for scale-invariant
-    # transformations it is not a fixed value.
-    llf <- -sum(w) / 2.0 * log(sigma_hat_squared) +
+    # variable y follows the normal distribution. See manuscript appendix for
+    # derivation.
+    llf <- -sum(w) / 2.0 * log(2 * pi * sigma_hat_squared) +
            -1.0 / (2.0 * sigma_hat_squared) * sum(w * (y - mu_hat)^2) +
+           -sum(w) * log(object@scale) +
            (object@lambda - 1.0) * sum(w * sign(x) * log1p(abs(x)))
 
 

### 2023-11-11 R/ParameterEstimators.R
New mle-based optimisation passes checks and yields promising output.

@@ -418,7 +418,7 @@ setMethod(
       return("bobyqa")
 
     } else {
-      return("subplex")
+      return("bobyqa")
     }
   }
 )

### 2023-11-11 manuscript/plots.R

@@ -1445,7 +1445,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     method = "yeo_johnson",
     robust = FALSE,
-    shift = FALSE
+    invariant = FALSE
   )
 
   # Raymaekers robust transformer
@@ -1453,7 +1453,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     method = "yeo_johnson",
     robust = TRUE,
-    shift = FALSE,
+    invariant = FALSE,
     estimation_method = "raymaekers_robust"
   )
 
@@ -1462,7 +1462,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     method = "yeo_johnson",
     robust = FALSE,
-    shift = TRUE
+    invariant = TRUE
   )
 
   # Robust shift-sensitive transformer
@@ -1470,7 +1470,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     method = "yeo_johnson",
     robust = TRUE,
-    shift = TRUE
+    invariant = TRUE
   )
 
   transformer_labels <- c(

### 2023-11-13 manuscript/manuscript.Rmd
New mle-based optimisation passes checks and yields promising output.
Manuscript revision.

@@
 require(coro)
+require(paletteer)
+require(robustHD)
+require(modeldata)
 
@@
 
--   We devise shift-sensitive versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
-
--   We define an empirical goodness of fit test for detecting cases where power transformations fail to yield a normally distributed transformed feature.
+-   We devise shift and scale invariant versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
+
+-   We define an empirical central normality test for detecting cases where power transformations fail to yield a roughly normally distributed transformed feature.
 
@@
 
-## Shift-sensitive power transformation
-
-Box-Cox and Yeo-Johnson transformations are modified by introducing a shift parameter $x_0$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
-The shift-sensitive Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$ and shift parameter $x_0$ is then:
+## Shift and scale invariant power transformation
+
+Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
+The shift and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
 
@@
 \label{eqn:box-cox-shift}
-\phi_{\text{BC}}^{\lambda, x_0} (x_i) = 
+\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
 \begin{cases}
-\left( \left(x_i - x_0 \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
-\log(x_i - x_0) & \text{if } \lambda = 0
+\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
 \end{cases}
@@
 ```
-Where $x_i - x_0 > 0$. Likewise, the shift-sensitive Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$ and shift parameter $x_0$ is:
+
+Where $x_i - x_0 > 0$.
+Likewise, the shift and scale invariant Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is:
 
@@
 \label{eqn:yeo-johnson-shift}
-\phi_{\text{YJ}}^{\lambda, x_0} (x_i) = 
+\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
 \begin{cases}
-\left( \left( 1 + x_i - x_0\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
-\log(1 + x_i - x_0) & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
--\left( \left( 1 - x_i - x_0\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0< 0\\
--\log(1 - x_i - x_0) & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
+\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
+\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
+-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
 \end{cases}
@@
 ```
-For both shift-sensitive transformations, $\lambda$ and $x_0$ are obtained by maximising the normal log-likelihood function.
-For the shift-sensitive Box-Cox transformation, and ignoring constant terms, this can be written as:
-
-```{=latex}
-\begin{equation}
-\label{eqn:box-cox-shift-log-likelihood}
-\left\{ \hat{\lambda}^{\text{BC}}, \hat{x}_0^{\text{BC}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \log(x_i - x_0)
-\end{equation}
-```
-where $\sigma^2$ is the variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$. Similarly for the shift-sensitive Yeo-Johnson transformation:
-
+
+For both invariant transformations, $\lambda$, $x_0$ and $s$ are obtained by maximising the log-likelihood function.
+A full derivation is shown in Appendix A.
+The invariant Box-Cox log-likelihood function is:
+
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$, respectively.
+Similarly, for the invariant Yeo-Johnson log-likelihood function is:
+HEREHEHEHHRHEHHEHEHRER
 ```{=latex}

### 2023-11-13 manuscript/manuscript_appendix_a.Rmd
Update manuscript_appendix_a.Rmd
Updated log likelihood to correct formulation.
Manuscript revision.

@@
 
+## Box-Cox transformation
+
 For the Box-Cox transformation the partial derivative is:
@@
 \begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
@@
 ```
+
+We introduced per-sample weights $w_i$ to facilitate robust power transformations. The log likelihood of the robust Box-Cox transformation is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+## Yeo-Johnson transformation
 
@@
 \begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log{2 \pi \sigma^2} -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 ```
+
+The log likelihood of the robust Yeo-Johnson transformation is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```

### 2023-11-13 manuscript/manuscript.Rmd
Manuscript revision.

@@
 
--   We define an empirical central normality test for detecting cases where power transformations fail to yield a roughly normally distributed transformed feature.
+-   We define an empirical central normality test for detecting cases where power transformations fail to yield an approximately normally distributed transformed feature.
 
@@
 ```
-
-Where $x_i - x_0 > 0$.
+where $x_i - x_0 > 0$.
 Likewise, the shift and scale invariant Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is:
@@
 ```
-where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$, respectively.
+where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 Similarly, for the invariant Yeo-Johnson log-likelihood function is:
-HEREHEHEHHRHEHHEHEHRER
+
 ```{=latex}
@@
 \label{eqn:yeo-johnson-shift-log-likelihood}
-\left\{ \hat{\lambda}^{\text{YJ}}, \hat{x}_0^{\text{YJ}} \right\} = \argmax_{\lambda, x_0} \sum_{i=1}^n -\frac{1}{2}\log(\sigma^2) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|)
-\end{equation}
-```
-where $\sigma^2$ is the variance of Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$.
+\begin{split}
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 

### 2023-11-14 manuscript/manuscript.Rmd
Work on manuscript revision.

@@
 
-For both invariant transformations, $\lambda$, $x_0$ and $s$ are obtained by maximising the log-likelihood function.
+For both invariant transformations, $\lambda$, $x_0$ and $s$ parameters are obtained by maximising the log-likelihood function.
 A full derivation is shown in Appendix A.
@@
 \begin{equation}
+\label{eqn:box-cox-invariant-log-likelihood}
 \begin{split}
@@
 \begin{equation}
-\label{eqn:yeo-johnson-shift-log-likelihood}
+\label{eqn:yeo-johnson-invariant-log-likelihood}
 \begin{split}
@@
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
-Because $\lambda$ is updated during their procedure, simultaneous estimation of $\lambda$ and $x_0$ for shift-sensitive power transformations is not possible.
+Because $\lambda$ is updated during their procedure, it can not be used simultaneously estimate $\lambda$, $x_0$ and $s$ for invariant power transformations.
 
@@
 @Raymaekers2021-kq do this (during the final steps of their iterative procedure) through a weighted maximum likelihood estimation.
-This weighted maximum log-likelihood estimator is based on equations \ref{eqn:box-cox-shift-log-likelihood} and \ref{eqn:yeo-johnson-shift-log-likelihood}.
-Here, the estimator is further normalised by the total weight and extended to include the shift parameter.
-The shift-sensitive Box-Cox power transformation is:
-
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
-\end{equation}
-```
-where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
-Analogously, the weighted maximum log-likelihood estimator for the shift-sensitive Yeo-Johnson power transformation is:
-
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
-\end{equation}
-```
-where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
-
-Other criteria aside from MLE can be used for finding robust transformation parameters, see the Appendix. For the remainder, we will use MLE as the optimisation criterion.
+Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
+Compared to @Raymaekers2021-kq, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
+The weighted invariant Box-Cox log-likelihood function is:
+
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Analogously, the weighted invariant Yeo-Johnson log-likelihood function is:
+
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Other criteria aside from MLE can be used for finding robust transformation parameters, see Appendix B. For the remainder, we will use MLE as the optimisation criterion.
 
@@
 
-- Set weights based on the z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$.
-After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$.
-Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
-
-- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+- Set weights based on the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
+After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
+Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
+
+- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
 

### 2023-11-14 manuscript/manuscript_appendix_a.Rmd
Work on manuscript revision.

@@
 ```
-where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively:
 
@@
 ```
-where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:
 

### 2023-11-14 manuscript/simulations.R

@@ -1,8 +1,8 @@
-.get_shifted_distribution_data <- function(manuscript_dir) {
+.get_shifted_scaled_distribution_data <- function(manuscript_dir) {
   # Set seed.
   set.seed(19L)
 
-  if (!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))) {
+  if (!file.exists(file.path(manuscript_dir, "shifted_scaled_distributions_plot.RDS"))) {
     # Normal distribution.
     x_normal <- power.transform::ragn(10000L, location = 0, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
 
@@ -12,49 +12,68 @@
     # Left skewed data
     x_left_skewed <- power.transform::ragn(10000L, location = 0, scale = 1 / sqrt(2), alpha = 0.8, beta = 2)
 
-
     # generator ----------------------------------------------------------------
     generate_experiment_data <- coro::generator(
       function(
-          x_normal,
-          x_right_skewed,
-          x_left_skewed) {
+    x_normal,
+    x_right_skewed,
+    x_left_skewed) {
         shift_range <- 10^seq(from = 0, to = 6, by = 0.1)
+        scale_range <- 10^seq(from = 0, to = 6, by = 0.1)
 
-        for (d in shift_range) {
-          for (distribution in c("normal", "right-skewed", "left-skewed")) {
-            if (distribution == "normal") {
-              x <- x_normal
-            } else if (distribution == "right-skewed") {
-              x <- x_right_skewed
-            } else if (distribution == "left-skewed") {
-              x <- x_left_skewed
-            }
+        for (distribution in c("normal", "right-skewed", "left-skewed")) {
+          if (distribution == "normal") {
+            x <- x_normal
+          } else if (distribution == "right-skewed") {
+            x <- x_right_skewed
+          } else if (distribution == "left-skewed") {
+            x <- x_left_skewed
+          }
 
-            for (method in c("box_cox", "yeo_johnson")) {
-              for (invariant in c(FALSE, TRUE)) {
-                for (robust in c(FALSE)) {
-                  for (estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())) {
-                    # Skip if robust, but not invariant.
-                    if (robust && !invariant) next
+          for (method in c("box_cox", "yeo_johnson")) {
+            for (invariant in c(FALSE, TRUE)) {
+              for (robust in c(FALSE)) {
[…]

### 2023-11-14 manuscript/manuscript.Rmd
Further updartes.

@@
 .plot_reduced_normality(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
-
 ```
@@
   "Power transformations are affected by outliers. ",
-  "Variable $x$ consists of $1000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$, ",
+  "Variable $x$ consists of $1000$ values randomly drawn from a normal distribution $\\mathcal{N}(0, 1)$, ",
   "with a single outlier with value $d$. ",

### 2023-11-14 manuscript/simulations.R

@@ -225,8 +225,7 @@
           x = parameter_set$x,
           method = parameter_set$method,
           robust = parameter_set$robust,
-          shift = parameter_set$shift,
-          optimiser_control = list("xtol_rel" = 1e-5)
+          invariant = parameter_set$invariant
         )
       )
 

### 2023-11-14 manuscript/manuscript.Rmd
Further changes.

@@
 
-## Robust shift-sensitive power transformations
+## Robust location and scale invariant power transformations
 
@@
 
-Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
+Modifications intended to make power transformations invariant to location and scale of a feature and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
 Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option for simulating realistic feature distributions.
@@
 $\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$, with $p_i$ probabilities of the empirical distribution, as previously.
-We then compute the residual errors between the z-scores of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ and the expected z-scores from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+We then compute the residual errors between the z-scores of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the expected z-scores from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
 The set of residual errors for the central portion of the data is then $\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
@@
 
-## Shift-sensitive transformations are location-invariant
-
-To assess whether the modified power transformations lead to $\lambda$ values that are invariant to shifts, we simulate three different distributions.
+## Invariance to location and scale
+
+To assess whether the modified power transformations lead to $\lambda$ values that are invariant to location and scale of the distribution, we simulate three different distributions.
 The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
@@
 The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the shift-sensitive definitions (equations \ref{eqn:box-cox-shift} and \ref{eqn:yeo-johnson-shift}) for each distribution, after adding a positive value $d$ to each distribution.
-Here, $d \in [1, 10^6]$.
+We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
+To assess location invariance, a positive value $d_{\text{shift}}$ is added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
+Similarly, to assess scale invariance, each distribution is multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.
 
@@
 cap <- paste0(
-  "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
-  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively,  which then underwent a shift $d$. ",
-  "Estimates of transformation parameter $\\lambda$ with the original power transformations (circle) show strong dependency on the overall location of the distribution, ",
-  "whereas estimates with the shift-sensitive power transformations (triangle) are constant.")
+  "Invariant power transformation produces transformation parameters that are invariant to location and shape. ",
+  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively,  which then underwent a shift $d_{\\text{shift}}$ or multiplication by $d_{\\text{scale}}$. ",
+  "Estimates of transformation parameter $\\lambda$ with the conventional power transformations (circle) show strong dependency on the overall location and scale of the distribution, ",
+  "whereas estimates with the invariant power transformations (triangle) are constant.")
 
@@
 The result is shown in Figure \ref{fig:shifted-distributions}.
-For each distribution, transformation parameter $\hat{\lambda}$ varies with $d$ when estimated using the original definitions.
-In contrast, estimation of $\lambda$ using shift-sensitive power transformations was invariant to $d$.
+For each distribution, transformation parameter $\hat{\lambda}$ varies with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions.
+In contrast, estimation of $\lambda$ using invariant power transformations was invariant to $d_{\text{shift}}$ and $d_{\text{scale}}$.
 
@@
 
-# Appendix A: Non-MLE-based optimisation of transformation parameters
+# Appendix A: Log-likelihood functions for invariant transformations.
+
+Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parameterised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
+This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
+\end{cases}
+\end{equation}
+```
+Where $x_i - x_0 > 0$.
+The location and scale-invariant Yeo-Johnson transformation is:
+
+```{=latex}
+\begin{equation}
+\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
+\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
+-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
+\end{cases}
+\end{equation}
+```
+
+The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
+The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
+However, the introduction of scaling parameter $s$ prevents their direct use.
+Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.
+
+Let $f(x_1, \ldots, x_n)$ be the probability density function of feature $\mathbf{X} = \{ x_1, \ldots, x_n\}$, 
+and $f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))$ be the probability density function of 
+the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$, that is assumed to follow a normal distribution.
+
+The two probability density functions are related as follows:
+
+```{=latex}
+\begin{equation}
+f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
+\end{equation}
+```
+Where, $\left|\mathbf{J}\right|$ is the determinant of Jacobian $\mathbf{J}$.
+The Jacobian takes the following form, with off-diagonal elements $0$:
+```{=latex}
+\begin{equation}
+\mathbf{J} =
+\begin{bmatrix}
+    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
+    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
+    \vdots & \vdots  & \ddots &  \vdots \\
+    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
+\end{bmatrix}
+\end{equation}
+```
+
+Thus, $\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)$.
+
+Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
+The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
+ & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
+\end{split}
+\end{equation}
+```
+With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
+The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
+However, the final term differs between Box-Cox and Yeo-Johnson transformations.
+
+## Box-Cox transformation
+
+For the Box-Cox transformation the partial derivative is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
+ & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
+\end{split}
+\end{equation}
+```
+
+Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
+& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
+& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+
+This leads to the following log-likelihood for the Box-Cox transformation:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+
+We introduced per-sample weights $w_i$ to facilitate robust power transformations. The log likelihood of the robust Box-Cox transformation is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+```
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+## Yeo-Johnson transformation
+
+For the Yeo-Johnson transformation, the partial derivative is:
+```{=latex}
+\begin{equation}
+\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
+\begin{cases}
+\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
+\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
+\end{cases}
+\end{equation}
+```
+
+Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+
+This lead to the following log-likelihood for the Yeo-Johnson transformation:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+
+The log likelihood of the robust Yeo-Johnson transformation is:
+```{=latex}
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+```
+where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:
+
+```{=latex}
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+# Appendix B: Optimisation of transformation parameters
 
@@
 
-Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift-sensitive power transformations, respectively.
+Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
 Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
 
-Below we will focus on criteria based on empirical density function, and on skewness and kurtosis of the transformed featured.
+Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
 Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
-In the case of the Shapiro-Wilk test statistic this is because of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.
+In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.
 

### 2023-11-14 manuscript/manuscript.Rmd
Work in progress.

@@
 
-```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
+```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=4.0, warning=FALSE}
 cap <- paste0(
   "Power transformations may reduce normality depending on location. ",
-  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
+  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution. ",
+  "For the top row, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
+  "whereas for the bottom row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
   "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and even breaks down due to numerical issues for $\\mu > 5 \\cdot 10^4$.")
+  "and breaks down due to numerical issues for $\\mu > 10^3$ (both Box-Cox and Yeo-Johnson) ",
+  "and $\\sigma > 10^3 (Yeo-Johnson) only."
+)
 
@@
 \begin{equation}
-\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 1} - 2}}
+\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
+\end{equation}
+
+\begin{equation}
+\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
 \end{equation}
@@
 \begin{equation}
-Z_s^2 = \frac{2 \left[\log\left(\beta_1 / \alpha + \sqrt{\beta_1^2 / \alpha^2 + 1} \right) \right]^2}{\log\left(\sqrt{2 \beta_2 - 1} - 1 \right)}
+Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
 \end{equation}
@@
 \begin{equation}
-Z_k^2 = \frac{9 \alpha_1}{2} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]^2
+Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
 \end{equation}

### 2023-11-14 manuscript/plots.R

@@ -118,7 +118,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p <- p + ggplot2::scale_y_continuous(
     name = latex2exp::TeX("$\\lambda$"),
-    limits = c(-5.0, 35.0)
+    limits = c(0.0, 3.0)
   )
 
   # Branch into shift and scale-specific plots
@@ -199,7 +199,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p <- p + plot_theme
-  p <- p + ggplot2::geom_point()
   p <- p + ggplot2::geom_hline(
     yintercept = 1.0,
     linetype = "longdash",
@@ -296,7 +295,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     p <- ggplot2::ggplot(
       data = data,
       mapping = ggplot2::aes(
-        x = d,
+        x = shift,
         y = lambda,
         colour = method,
         shape = version
@@ -379,7 +378,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
 
   # Process / read data.
-  data <- .get_shifted_distribution_data(manuscript_dir = manuscript_dir)
+  data <- .get_shifted_scaled_distribution_data(manuscript_dir = manuscript_dir)
   data <- data[estimation_method == "MLE"]
 
   # Normal distribution --------------------------------------------------------

### 2023-11-15 manuscript/manuscript.Rmd
Work on manuscript.

@@
 plot_theme <- ggplot2::theme_light(base_size=base_size)
+plot_theme$plot.title$size <- ggplot2::rel(1.0)
 plot_theme$strip.background <- ggplot2::element_blank()
@@
   "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution. ",
-  "For the top row, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
-  "whereas for the bottom row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
+  "For the left-hand panels, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
+  "whereas for the right-hand row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
   "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and breaks down due to numerical issues for $\\mu > 10^3$ (both Box-Cox and Yeo-Johnson) ",
-  "and $\\sigma > 10^3 (Yeo-Johnson) only."
+  "and breaks down due to numerical issues for $\\mu > 5 \\cdot 10^3$ (both Box-Cox and Yeo-Johnson).",
+  "Scale did not affect Box-Cox and Yeo-Johnson transformations of variables drawn from normal distributions."
 )
@@
 
-## Shift and scale invariant power transformation
+## Location and scale invariant power transformation
 
 Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
-The shift and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
+The location and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
 
@@
 In contrast, estimation of $\lambda$ using invariant power transformations was invariant to $d_{\text{shift}}$ and $d_{\text{scale}}$.
+While the effect of location on the estimated value of $\lambda$ using conventional power transformations is noticeable, scale also affects $\lambda$ for features that are not normally distributed.
+The figure also shows that estimated $\lambda$ for Box-Cox power transformations in left-skewed distributions diverges from that found for invariant Box-Cox power transformations.
 

### 2023-11-15 R/TransformationYeoJohnson.R

@@ -376,10 +376,10 @@ setMethod(
     if (is.na(skewness)) skewness <- 0.0
 
     if (skewness < 0.0) {
-      x_initial <- stats::quantile(x, 0.35)
+      x_initial <- stats::quantile(x, 0.95)
 
     } else {
-      x_initial <- stats::quantile(x, 0.65)
+      x_initial <- stats::quantile(x, 0.05)
     }
 
     # Set shift range. Allow for moving the values entirely negative or
@@ -436,11 +436,11 @@ setMethod(
     if (!is.finite(scale)) scale <- 1.0
     if (scale == 0.0) scale <- 1.0
 
-    if (all(x > 0.0)) {
-      # Set shift so that the minimum value after scaling is 1, if all values
-      # are positive.
-      shift <- (min(x) / scale - 1.0) * scale
-    }
+    # if (all(x > 0.0)) {
+    #   # Set shift so that the minimum value after scaling is 1, if all values
+    #   # are positive.
+    #   shift <- (min(x) / scale - 1.0) * scale
+    # }
 
     # Shift and scale x.
     x <- (x - shift) / scale

### 2023-11-15 manuscript/plots.R

@@ -118,7 +118,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p <- p + ggplot2::scale_y_continuous(
     name = latex2exp::TeX("$\\lambda$"),
-    limits = c(0.0, 3.0)
+    limits = c(0.0, 6.0)
   )
 
   # Branch into shift and scale-specific plots
@@ -136,10 +136,23 @@ get_annotation_settings <- function(ggtheme = NULL) {
     data = data[method == "Box-Cox"],
     mapping = ggplot2::aes(x = .data$shift)
   )
+  p_bc_shift <- p_bc_shift + ggplot2::theme(
+    axis.text.x = ggplot2::element_blank(),
+    axis.title.x = ggplot2::element_blank(),
+    axis.ticks.x = ggplot2::element_blank()
+  )
 
   p_bc_scale <- p_scale + ggplot2::geom_point(
     data = data[method == "Box-Cox"],
     mapping = ggplot2::aes(x = .data$scale)
+  )
+  p_bc_scale <- p_bc_scale + ggplot2::theme(
+    axis.text.y = ggplot2::element_blank(),
+    axis.title.y = ggplot2::element_blank(),
+    axis.ticks.y = ggplot2::element_blank(),
+    axis.text.x = ggplot2::element_blank(),
+    axis.title.x = ggplot2::element_blank(),
+    axis.ticks.x = ggplot2::element_blank()
   )
 
   # Yeo-Johnson transformation -------------------------------------------------
@@ -148,17 +161,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
     mapping = ggplot2::aes(x = .data$shift)
   )
 
-  p_yj_shift <- p_yj_shift + ggplot2::theme(
-    axis.text.y = ggplot2::element_blank(),
-    axis.title.y = ggplot2::element_blank(),
-    axis.ticks.y = ggplot2::element_blank()
-  )
-
   p_yj_scale <- p_scale + ggplot2::geom_point(
     data = data[method == "Yeo-Johnson"],
-    mapping = ggplot2::aes(x = .data$shift)
+    mapping = ggplot2::aes(x = .data$scale)
   )
-
   p_yj_scale <- p_yj_scale + ggplot2::theme(
     axis.text.y = ggplot2::element_blank(),
     axis.title.y = ggplot2::element_blank(),
@@ -166,10 +172,11 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
 
   # Patch all the plots together.
[…]

### 2023-11-15 manuscript/simulations.R

@@ -33,7 +33,8 @@
           for (method in c("box_cox", "yeo_johnson")) {
             for (invariant in c(FALSE, TRUE)) {
               for (robust in c(FALSE)) {
-                for (estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())) {
+                for (estimation_method in "mle") {
+                # for (estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())) {
                   # Skip if robust, but not invariant.
                   if (robust && !invariant) next
 
@@ -95,10 +96,7 @@
             method = parameter_set$method,
             robust = parameter_set$robust,
             invariant = parameter_set$invariant,
-            estimation_method = parameter_set$estimation_method,
-            lambda = NULL,
-            optimiser = "subplex",
-            optimiser_control = list("xtol_rel"=1e-6)
+            estimation_method = parameter_set$estimation_method
           )
         )
       } else {
@@ -108,9 +106,7 @@
             method = parameter_set$method,
             robust = parameter_set$robust,
             invariant = parameter_set$invariant,
-            estimation_method = parameter_set$estimation_method,
-            optimiser = "subplex",
-            optimiser_control = list("xtol_rel"=1e-6)
+            estimation_method = parameter_set$estimation_method
           )
         )
       }

### 2023-11-16 manuscript/manuscript.Rmd
Further work.

@@
 
+In this section we simulate data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
+The $\lambda$ parameter for conventional power transformations (Eqn.
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn.
+\ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
+Appendix B describes other criteria, and Appendix C the results of simulations using these criteria.
+
 ## Invariance to location and scale
 
-To assess whether the modified power transformations lead to $\lambda$ values that are invariant to location and scale of the distribution, we simulate three different distributions.
+To assess whether the proposed power transformations lead to $\lambda$ values that are invariant to location and scale of the distribution, we simulate three different distributions.
 The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
@@
   "Estimates of transformation parameter $\\lambda$ with the conventional power transformations (circle) show strong dependency on the overall location and scale of the distribution, ",
-  "whereas estimates with the invariant power transformations (triangle) are constant.")
+  "whereas estimates with the invariant power transformations (triangle) are constant."
+)
 
@@
 This manuscript was created using R Markdown, and is available from [https://github.com/alexzwanenburg/power.transform].
-
-# Appendix A: Log-likelihood functions for invariant transformations.
-
-Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parameterised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
-This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
-\begin{cases}
-\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
-\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
-\end{cases}
-\end{equation}
-```
-Where $x_i - x_0 > 0$.
-The location and scale-invariant Yeo-Johnson transformation is:
-
-```{=latex}
-\begin{equation}
-\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
-\begin{cases}
-\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
-\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
--\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
--\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
-\end{cases}
-\end{equation}
-```
-
-The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
-The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
-However, the introduction of scaling parameter $s$ prevents their direct use.
-Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.
-
-Let $f(x_1, \ldots, x_n)$ be the probability density function of feature $\mathbf{X} = \{ x_1, \ldots, x_n\}$, 
-and $f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))$ be the probability density function of 
-the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$, that is assumed to follow a normal distribution.
-
-The two probability density functions are related as follows:
-
-```{=latex}
-\begin{equation}
-f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
-\end{equation}
-```
-Where, $\left|\mathbf{J}\right|$ is the determinant of Jacobian $\mathbf{J}$.
-The Jacobian takes the following form, with off-diagonal elements $0$:
-```{=latex}
-\begin{equation}
-\mathbf{J} =
-\begin{bmatrix}
-    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
-    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
-    \vdots & \vdots  & \ddots &  \vdots \\
-    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
-\end{bmatrix}
-\end{equation}
-```
-
-Thus, $\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)$.
-
-Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
-The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
-```{=latex}
-\begin{equation}
-\begin{split}
-\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
- & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
- & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
- & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
-\end{split}
-\end{equation}
-```
-With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
-The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
-However, the final term differs between Box-Cox and Yeo-Johnson transformations.
-
-## Box-Cox transformation
-
-For the Box-Cox transformation the partial derivative is:
-```{=latex}
-\begin{equation}
-\begin{split}
-\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
- & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
-\end{split}
-\end{equation}
-```
-
-Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
-```{=latex}
-\begin{equation}
-\begin{split}
-\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
-& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
-& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
-\end{split}
-\end{equation}
-```
-
-This leads to the following log-likelihood for the Box-Cox transformation:
-```{=latex}
-\begin{equation}
-\begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
-& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
-\end{split}
-\end{equation}
-```
-
-We introduced per-sample weights $w_i$ to facilitate robust power transformations. The log likelihood of the robust Box-Cox transformation is:
-```{=latex}
-\begin{equation}
-\begin{split}
-\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
-& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
-\end{split}
-\end{equation}
-```
-where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
-
-## Yeo-Johnson transformation
-
-For the Yeo-Johnson transformation, the partial derivative is:
-```{=latex}
-\begin{equation}
-\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
-\begin{cases}
-\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
-\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
-\end{cases}
-\end{equation}
-```
-
-Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
-```{=latex}
-\begin{equation}
-\begin{split}
-\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
-\end{split}
-\end{equation}
-```
-
-This lead to the following log-likelihood for the Yeo-Johnson transformation:
-```{=latex}
-\begin{equation}
-\begin{split}
-\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
-& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
-\end{split}
-\end{equation}
-```
-
-The log likelihood of the robust Yeo-Johnson transformation is:
-```{=latex}
-\begin{equation}
-\begin{split}
-\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
-& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
-\end{split}
-\end{equation}
-```
-where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
-
-# Appendix B: Optimisation of transformation parameters
-
-Maximum likelihood estimation (MLE) is commonly used to optimise parameters for power transformation.
-Generally, optimisation requires minimisation or maximisation of a criterion.
-In MLE, the maximised criterion is the log-likelihood function of the normal distribution.
-Here, we investigate power transformation using optimisation criteria that are closely related to test statistics for normality tests.
-
-Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
-Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
-
-Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
-Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
-In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.
-
-## Empirical density function-based criteria
-
-The first class of criteria is based on the empirical distribution function (EDF).
-Transformation parameters are then fit through minimisation of the distance between the empirical distribution function $F_{\epsilon}$ and the cumulative density function (CDF) of the normal distribution $F_{\mathcal{N}}$.
-Let $F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}$ be the empirical probability of instance $i$.
-The normal distribution is parametrised by location parameter $\mu$ and scale parameter $\sigma$, both of which have to be estimated from the data.
-For non-robust power transformation, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
-For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
-
-### Anderson-Darling criterion
-
-The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$. We define this criterion as follows:
-
-```{=latex}
-\begin{equation}
-U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right) }
-\end{equation}
-```
-
-Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
-For non-robust power transformations, all $w_i = 1$.
-Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include any extraneous weights $w_i$.
-The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.
-
-### Cramér-von Mises criterion
-
-The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
-```{=latex}
-\begin{equation}
-U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2
-\end{equation}
-```
-
-Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
-For non-robust power transformations, all $w_i = 1$.
-The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of extraneous weights.
-This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
-Unlike the Anderson-Darling criterion, this criterion weights all instances equally.
-
-For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
-
-## Skewness-kurtosis-based criteria
-
-The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
-We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
-Here, $\mu$ is defined as:
-
-```{=latex}
-\begin{equation}
-\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0} \left(x_i \right)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
-
-The location, or mean, is weighted using weights $w_i$, as described in section **Insert**. For non-robust transformations, $w_i = 1$. Then, $sigma^2$ is defined as:
-```{=latex}
-\begin{equation}
-\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
-\end{equation}
-```
-
-Skewness is defined as:
-```{=latex}
-\begin{equation}
-s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
-\end{equation}
-```
-
-Kurtosis is defined as:
-```{=latex}
-\begin{equation}
-k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
-\end{equation}
-```
-
-### D'Agostino criterion
-
-The D'Agostino criterion defined here follows the D'Agostino $K^2$ test statistic [@DAgostino1990-kp].
-This test statistic is composed of two separate test statistics, one of which is related to skewness, and the other to kurtosis.
-Both test statistics are computed in several steps.
-Let us first define $\nu=\sum_{i=1}^n w_i$.
-Thus for non-robust power transformations, $\nu = n$.
-
-For the skewness test statistic we first compute [@DAgostino1990-kp]:
-
-```{=latex}
-\begin{equation}
-\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
-\end{equation}
-
-\begin{equation}
-\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
-\end{equation}
-
-\begin{equation}
-\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
-\end{equation}
-
-\begin{equation}
-\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
-\end{equation}
-```
-
-The skewness test statistic is then:
-
-```{=latex}
-\begin{equation}
-Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
-\end{equation}
-```
-
-For the kurtosis test statistic we first compute [@DAgostino1990-kp, @Anscombe1983-nz]:
-
-```{=latex}
-\begin{equation}
-\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
-\end{equation}
-
-\begin{equation}
-\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
-\end{equation}
-
-\begin{equation}
-\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
-\end{equation}
-
-\begin{equation}
-\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
-\end{equation}
-
-\begin{equation}
-\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
-\end{equation}
-```
-
-The kurtosis test statistic is then:
-
-```{=latex}
-\begin{equation}
-Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
-\end{equation}
-```
-
-The D'Agostino $K^2$ test statistic and our criterion are the same, and are defined as:
-```{=latex}
-\begin{equation}
-U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
-\end{equation}
-```
-
-The main difference between the test statistic as originally formulated, and the criterion proposed here is the presence of weights for robust power transformation.
-
-### Jarque-Bera criterion
-
-The second criterion based on skewness and kurtosis is the Jarque-Bera criterion. It is relatively simple to compute, compared to the D'Agostino criterion:
-```{=latex}
-\begin{equation}
-U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
-\end{equation}
-```
-
-The main difference between the above criterion and the Jarque-Bera test statistic [@Jarque1980-hw] is that a scalar multiplication is absent.
-
-## Optimisation using non-MLE criteria
-
-Each of the criteria described in the previous section can be used for optimisation, i.e.:
-
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}, \hat{x}_0 \right\} = \argmin_{\lambda, x_0} U\left(\mathbf{X}, \lambda, x_0 \right)
-\end{equation}
-```
-
-For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, or equivalently, $x_0$ is fixed:
-```{=latex}
-\begin{equation}
-\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0 \right)
-\end{equation}
-```
-
-# Appendix B: Simulations with other optimisation criteria
-
-```{r shifted-distributions-other-criteria, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Shift-sensitive power transformation produces transformation parameters that are invariant to location. ",
-  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively,  which then underwent a shift $d$. ",
-  "Estimates of transformation parameter $\\lambda$ with the original power transformations (circle) show strong dependency on the overall location of the distribution, ",
-  "whereas estimates with the shift-sensitive power transformations (triangle) are constant.")
-
-  .plot_shifted_distributions_other_criteria(manuscript_dir = manuscript_dir, plot_theme = plot_theme)
-```
 

### 2023-11-16 manuscript/manuscript_appendix.Rmd
Further work.

--- 2023-11-14 manuscript/manuscript_appendix_a.Rmd
@@
 ---
-title: "Appendix A"
+title: "Appendix"
 author: "Alex Zwanenburg"
@@
 ```
+
+# Appendix A: Log-likelihood functions for location and scale invariant power transformation
 
@@
 ```
+
+# Appendix B: Optimisation of transformation parameters
+
+Maximum likelihood estimation (MLE) is commonly used to optimise parameters for power transformation.
+Generally, optimisation requires minimisation or maximisation of a criterion.
+In MLE, the maximised criterion is the log-likelihood function of the normal distribution.
+Here, we investigate power transformation using optimisation criteria that are closely related to test statistics for normality tests.
+
+Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
+Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
+
+Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
+Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
+In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.
+
+## Empirical density function-based criteria
+
+The first class of criteria is based on the empirical distribution function (EDF).
+Transformation parameters are then fit through minimisation of the distance between the empirical distribution function $F_{\epsilon}$ and the cumulative density function (CDF) of the normal distribution $F_{\mathcal{N}}$.
+Let $F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}$ be the empirical probability of instance $i$.
+The normal distribution is parametrised by location parameter $\mu$ and scale parameter $\sigma$, both of which have to be estimated from the data.
+For non-robust power transformation, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
+For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+
+### Anderson-Darling criterion
+
+The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$. We define this criterion as follows:
+
+```{=latex}
+\begin{equation}
+U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right) }
+\end{equation}
+```
+
+Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+For non-robust power transformations, all $w_i = 1$.
+Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include any extraneous weights $w_i$.
+The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.
+
+### Cramér-von Mises criterion
+
+The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
+```{=latex}
+\begin{equation}
+U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2
+\end{equation}
+```
+
+Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+For non-robust power transformations, all $w_i = 1$.
+The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of extraneous weights.
+This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
+Unlike the Anderson-Darling criterion, this criterion weights all instances equally.
+
+For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
+
+## Skewness-kurtosis-based criteria
+
+The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
+We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
+Here, $\mu$ is defined as:
+
+```{=latex}
+\begin{equation}
+\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0} \left(x_i \right)} {\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+The location, or mean, is weighted using weights $w_i$, as described in section **Insert**. For non-robust transformations, $w_i = 1$. Then, $sigma^2$ is defined as:
+```{=latex}
+\begin{equation}
+\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Skewness is defined as:
+```{=latex}
+\begin{equation}
+s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
+\end{equation}
+```
+
+Kurtosis is defined as:
+```{=latex}
+\begin{equation}
+k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
+\end{equation}
+```
+
+### D'Agostino criterion
+
+The D'Agostino criterion defined here follows the D'Agostino $K^2$ test statistic [@DAgostino1990-kp].
+This test statistic is composed of two separate test statistics, one of which is related to skewness, and the other to kurtosis.
+Both test statistics are computed in several steps.
+Let us first define $\nu=\sum_{i=1}^n w_i$.
+Thus for non-robust power transformations, $\nu = n$.
+
+For the skewness test statistic we first compute [@DAgostino1990-kp]:
+
+```{=latex}
+\begin{equation}
+\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
+\end{equation}
+
+\begin{equation}
+\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
+\end{equation}
+
+\begin{equation}
+\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
+\end{equation}
+
+\begin{equation}
+\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
+\end{equation}
+```
+
+The skewness test statistic is then:
+
+```{=latex}
+\begin{equation}
+Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
+\end{equation}
+```
+
+For the kurtosis test statistic we first compute [@DAgostino1990-kp, @Anscombe1983-nz]:
+
+```{=latex}
+\begin{equation}
+\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
+\end{equation}
+
+\begin{equation}
+\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
+\end{equation}
+
+\begin{equation}
+\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
+\end{equation}
+
+\begin{equation}
+\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
+\end{equation}
+
+\begin{equation}
+\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
+\end{equation}
+```
+
+The kurtosis test statistic is then:
+
+```{=latex}
+\begin{equation}
+Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
+\end{equation}
+```
+
+The D'Agostino $K^2$ test statistic and our criterion are the same, and are defined as:
+```{=latex}
+\begin{equation}
+U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
+\end{equation}
+```
+
+The main difference between the test statistic as originally formulated, and the criterion proposed here is the presence of weights for robust power transformation.
+
+### Jarque-Bera criterion
+
+The second criterion based on skewness and kurtosis is the Jarque-Bera criterion. It is relatively simple to compute, compared to the D'Agostino criterion:
+```{=latex}
+\begin{equation}
+U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
+\end{equation}
+```
+
+The main difference between the above criterion and the Jarque-Bera test statistic [@Jarque1980-hw] is that a scalar multiplication is absent.
+
+## Optimisation using non-MLE criteria
+
+Each of the above criteria can be used for optimisation, i.e.:
+
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}, \hat{x}_0, \hat{s}_0 \right\} = \argmin_{\lambda, x_0, s} U\left(\mathbf{X}, \lambda, x_0, s \right)
+\end{equation}
+```
+
+For conventional power transformations with fixed location and scale parameters, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, or equivalently, $x_0$ and $s$ may be fixed:
+```{=latex}
+\begin{equation}
+\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0, s \right)
+\end{equation}
+```
+
+# Appendix B: Simulations with other optimisation criteria
+
+```{r shifted-distributions-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Invariant power transformation produces transformation parameters that are invariant to location and shape. ",
+  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\\text{shift}}$ or multiplication by $d_{\\text{scale}}$. ",
+  "Estimates of transformation parameter $\\lambda$ with the conventional power transformations (circle) show strong dependency on the overall location and scale of the distribution, ",
+  "whereas estimates with the invariant power transformations (triangle) are constant."
+)
+
+.plot_shifted_distributions_appendix(manuscript_dir = manuscript_dir, plot_theme = plot_theme)
+```

### 2023-11-16 manuscript/plots.R

@@ -81,7 +81,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
   estimation_method <- distribution <- method <- version <- NULL
 
   # Process / read data.
-  data <- .get_shifted_scaled_distribution_data(manuscript_dir = manuscript_dir)
+  data <- .get_shifted_scaled_distribution_data(
+    manuscript_dir = manuscript_dir,
+    main_manuscript = TRUE
+  )
   data <- data[estimation_method == "MLE" & version == "conventional" & distribution == "normal"]
 
   # Get annotation settings.
@@ -260,7 +263,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 .plot_shifted_distributions <- function(plot_theme, manuscript_dir) {
-  # Creates the plot for lambda values under location shift for MLE
+  # Creates the plot for lambda values under location and scale shifts for MLE
 
   require(patchwork)
   require(ggplot2)
@@ -544,7 +547,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-.plot_shifted_distributions_other_criteria <- function(plot_theme, manuscript_dir) {
+.plot_shifted_distributions_appendix <- function(plot_theme, manuscript_dir) {
   # Creates the plot for lambda values under location shift for all optimisation
   # criteria.
 
@@ -616,7 +619,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
   distribution <- method <- version <- NULL
 
   # Process / read data.
-  data <- .get_shifted_distribution_data(manuscript_dir = manuscript_dir)
+  data <- .get_shifted_distribution_data(
+    manuscript_dir = manuscript_dir,
+    main_manuscript = FALSE
+  )
 
   # Normal distribution --------------------------------------------------------
 

### 2023-11-16 manuscript/manuscript.Rmd
Further update to document.

@@
 While the effect of location on the estimated value of $\lambda$ using conventional power transformations is noticeable, scale also affects $\lambda$ for features that are not normally distributed.
-The figure also shows that estimated $\lambda$ for Box-Cox power transformations in left-skewed distributions diverges from that found for invariant Box-Cox power transformations.
+Figure \ref{fig:shifted-distributions} also shows that estimated $\lambda$ for Box-Cox power transformations in left-skewed distributions diverges from that found for invariant Box-Cox power transformations.
 
@@
 Location and scale parameters are set as $\mu = 0$ and $\sigma = 1$, respectively.
-$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(2, 4\right)$.
-That is between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
+$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(2, 4\right)$, i.e., between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
 
@@
 The double sum in right-hand side of the above equation can be considered as the loss that should be minimised.
-Minimisation is performed using the `sbplx` algorithm implemented in the `nloptr` package for R [@Rowan1990-sn; @Ypma2022-tc].
+As mentioned earlier, transformation parameters are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
 
 The optimal weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
+
+```{r include = FALSE, eval = FALSE}
+# This writes the lookup table for two-sided weighting parameters.
+# power.transform package.
+load(file = "./R/sysdata.rda")
+
+two_sided_function_parameters <- .get_optimised_weighting_function_parameters(
+  manuscript_dir = manuscript_dir,
+  side = "both"
+)
+
+usethis::use_data(
+  gof_lookup_table, 
+  two_sided_function_parameters, 
+  internal = TRUE, 
+  overwrite = TRUE)
+```
 
@@
   "parametrised asymmetric generalised normal distributions with up to 10% outliers on either side ",
-  "of the distribution. The top and bottom panel show errors for the shift-sensitive Box-Cox transformation, ",
-  "and shift-sensitive Yeo-Johnson transformation, respectively. ",
+  "of the distribution. The top and bottom panel show errors for the invariant Box-Cox ",
+  "and invariant Yeo-Johnson transformations, respectively. ",
   "In each panel, the error distribution for the non-robust transformation is shown to the left ",
@@
   plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir)
+  manuscript_dir = manuscript_dir
+)
 ```

### 2023-11-16 manuscript/manuscript_appendix.Rmd
Further update to document.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(
+  echo = FALSE,
+  eval = TRUE,
+  fig.align= "center")
+
+# Allow for defining figure captions within a chunk.
+knitr::opts_knit$set(
+  eval.after = "fig.cap")
+
+require(power.transform)
+require(ggplot2)
+require(patchwork)
+require(data.table)
+require(latex2exp)
+require(coro)
+require(paletteer)
+require(robustHD)
+require(modeldata)
+
+manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")
+source(file.path(manuscript_dir, "simulations.R"))
+source(file.path(manuscript_dir, "plots.R"))
+
+# Base font size.
+base_size <- 10
+
+# Set general theme.
+plot_theme <- ggplot2::theme_light(base_size=base_size)
+plot_theme$plot.title$size <- ggplot2::rel(1.0)
+plot_theme$strip.background <- ggplot2::element_blank()
+plot_theme$strip.text$colour <- NULL
+plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
+
+parse_mean_sd_latex <- function(mu, sigma, digits=2L){
+  return(paste0(
+    "$", round(mu, digits=digits), " \\pm ", round(sigma, digits=digits), "$"
+  ))
+}
 ```

### 2023-11-16 manuscript/simulations.R

@@ -3,10 +3,10 @@
   set.seed(19L)
 
   if (main_manuscript) {
-    file_name <- file.path(manuscript_dir, "shifted_scaled_distributions_plot_main.RDS")
+    file_name <- file.path(manuscript_dir, "main_shifted_scaled_distributions_plot.RDS")
     estimation_methods <- "mle"
   } else {
-    file_name <- file.path(manuscript_dir, "shifted_scaled_distributions_plot_appendix.RDS")
+    file_name <- file.path(manuscript_dir, "appendix_shifted_scaled_distributions_plot.RDS")
     estimation_methods <- setdiff(
       power.transform:::..estimators_all(),
       power.transform:::..estimators_raymaekers_robust()
@@ -211,7 +211,7 @@
   # Set seed.
   set.seed(19L)
 
-  if (!file.exists(file.path(manuscript_dir, "shifted_outlier_plot.RDS"))) {
+  if (!file.exists(file.path(manuscript_dir, "main_shifted_outlier_plot.RDS"))) {
     # Normal distribution.
     x <- power.transform::ragn(1000L, location = offset, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
 
@@ -329,7 +329,7 @@
             "method" = transformation_method,
             "estimation_method" = estimation_method,
             "robust" = FALSE,
-            "shift" = TRUE
+            "invariant" = TRUE
           )
 
           opt_args <- list()
@@ -379,7 +379,7 @@
                 "method" = transformation_method,
                 "estimation_method" = estimation_method,
                 "robust" = TRUE,
-                "shift" = TRUE,
+                "invariant" = TRUE,
                 "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
                 "backup_use_default" = FALSE
               )
@@ -449,7 +449,7 @@
             method = transformation_method,
             estimation_method = estimation_method,
             robust = FALSE,
-            shift = TRUE
+            invariant = TRUE
           )
         )
 
@@ -576,7 +576,8 @@
           "weight_method" = fun_args$weighting_function,
           "lambda" = transformer@lambda,
           "target_lambda" = target_lambda,
-          "shift" = transformer@shift
+          "shift" = transformer@shift,
+          "scale" = transformer@scale
         )
       )
 

### 2023-11-21 manuscript/manuscript.Rmd
Updated weights.

@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 773 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.84 & ---  & ---       & ---  & 598 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.79 & 0.95 & $(0, 1]$  & 0.98 & 594 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.77 & 0.95 & $(0, 1]$  & 1.00 & 580 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.21 & ---  & ---       & ---  & 1214 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.05 & 1.96 & $(0, 10]$ & 4.47 & 1052 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.09 & 1.96 & $(0, 10]$ & 5.30 & 1086 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.36 & ---  & ---       & ---  & 1483 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.43 & 1.00 & $(0, 10]$ & 1.45 & 1498 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.20 & 1.00 & $(0, 10]$ & 1.48 & 1467 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
 \bottomrule
@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Yeo-Johnson} power transformations.
+\caption{Optimal weighting parameters and corresponding loss for \textbf{invariant Yeo-Johnson} power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 305 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.87 & ---  & ---       & ---  & 136 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.88 & 133 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.81 & 0.95 & $(0, 1]$  & 0.88 & 133 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.07 & ---  & ---       & ---  & 333 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.00 & 1.96 & $(0, 10]$ & 4.57 & 227 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.55 & 1.96 & $(0, 10]$ & 5.16 & 236 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.57 & ---  & ---       & ---  & 631 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.28 & 1.00 & $(0, 10]$ & 1.81 & 562 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.32 & 1.00 & $(0, 10]$ & 1.75 & 567 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
 \bottomrule

### 2023-11-21 manuscript/simulations.R
Updated .get_transformation_parameters

@@ -786,10 +786,11 @@
   if (!file.exists(file_name)) {
     # generator ----------------------------------------------------------------
     experiment_args <- coro::generator(
-      function(manuscript_dir,
-               side,
-               transformation_methods = NULL,
-               estimation_methods = NULL) {
+      function(
+    manuscript_dir,
+    side,
+    transformation_methods = NULL,
+    estimation_methods = NULL) {
         if (side == "both") {
           file_dir <- "robustness_comparison"
         } else if (side == "right") {
@@ -811,7 +812,7 @@
               "method" = transformation_method,
               "estimation_method" = estimation_method,
               "robust" = FALSE,
-              "shift" = TRUE
+              "invariant" = TRUE
             )
 
             file_name <- file.path(
@@ -824,7 +825,6 @@
               )
             )
 
-            # Only yield arguments if the file does not exist.
             yield(list(
               "name" = "non-robust",
               "method" = transformation_method,
@@ -861,7 +861,7 @@
                   "method" = transformation_method,
                   "estimation_method" = estimation_method,
                   "robust" = TRUE,
-                  "shift" = TRUE,
+                  "invariant" = TRUE,
                   "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
                   "weighting_function_parameters" = weighting_function_parameters
                 )
@@ -882,11 +882,11 @@
     )
 
 
-
     # Helper function for computing the target lambda value.
-    .compute_target_lambda <- function(x,
-                                       transformation_methods = NULL,
-                                       estimation_methods = NULL) {
+    .compute_target_lambda <- function(
+    x,
+    transformation_methods = NULL,
+    estimation_methods = NULL) {
       if (is.null(transformation_methods)) {
         transformation_methods <- c("box_cox", "yeo_johnson")
       }
@@ -906,7 +906,7 @@
               method = transformation_method,
               estimation_method = estimation_method,
               robust = FALSE,
-              shift = TRUE
+              invariant = TRUE
             )
           )
 
@@ -925,14 +925,15 @@
 
 
     # Helper function for population distributions with outliers.
-    .populate_outliers <- function(x,
-                                   n,
-                                   alpha,
-                                   beta,
-                                   target_lambda,
-                                   ii,
-                                   outlier_fraction,
-                                   side) {
+    .populate_outliers <- function(
+    x,
+    n,
+    alpha,
+    beta,
[…]

### 2023-11-21 manuscript/plots.R
Update residual data.

@@ -895,7 +895,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   # Prevent warnings due to non-standard evaluation.
   has_outliers <- residual <- threshold <- residual_error <- method <- p <- NULL
 
-  if (!file.exists(file.path(manuscript_dir, "residual_error_plot_main_manuscript.RDS"))) {
+  if (!file.exists(file.path(manuscript_dir, "manuscript_residual_error.RDS"))) {
     # Get data
     data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
 
@@ -930,10 +930,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
     saveRDS(
       data,
-      file.path(manuscript_dir, "residual_error_plot_main_manuscript.RDS")
+      file.path(manuscript_dir, "manuscript_residual_error.RDS")
     )
   } else {
-    data <- readRDS(file.path(manuscript_dir, "residual_error_plot_main_manuscript.RDS"))
+    data <- readRDS(file.path(manuscript_dir, "manuscript_residual_error.RDS"))
 
     outlier_free_data <- data[
       has_outliers == FALSE,

### 2023-11-21 manuscript/simulations.R

@@ -1266,7 +1266,7 @@
     transformer_bc <- suppressWarnings(power.transform::find_transformation_parameters(
       x = x$x,
       method = "box_cox",
-      shift = TRUE,
+      invariant = TRUE,
       robust = TRUE,
       estimation_method = "mle",
       weighting_function = "empirical_probability_cosine"
@@ -1297,7 +1297,7 @@
     transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
       x = x$x,
       method = "yeo_johnson",
-      shift = TRUE,
+      invariant = TRUE,
       robust = TRUE,
       estimation_method = "mle",
       weighting_function = "empirical_probability_cosine"

### 2023-11-21 manuscript/manuscript.Rmd
Updated .get_transformation_parameters
Update residual data.
Final analyses of the main manuscript.

@@
 For both transformations, weighting based on empirical probabilities yields the most consistent $\lambda$ parameter estimates in the presence of outliers.
-Weighting based on residual errors tended to perform least well, and were generally similar to (Yeo-Johnson) or worse than (Box-Cox) non-robust transformations.
+Other methods did not outperform the non-robust power transformation method.
 
@@
 \midrule
-0.042 & 0.063 & 0.077 & 0.090 & 0.106 & 0.118 & 0.159 \\
+0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
 \bottomrule
@@
 
-```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
+```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
@@
   "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \cite{Raymaekers2021-kq}. ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
   "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
@@
 
-```{r housing-data, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2}
+```{r housing-data, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
@@
 
-.plot_sacramento_house_prices(
+.plot_sacramento_house(
   plot_theme = plot_theme

### 2023-11-21 manuscript/plots.R

@@ -1501,7 +1501,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
 
   transformer_labels <- c(
-    "none", "original", "Raymaekers-Rousseeuw", "shift-sensitive", "robust shift-sensitive"
+    "none", "original", "Raymaekers-Rousseeuw", "invariant", "robust invariant"
   )
 
   data <- mapply(

### 2024-06-04 manuscript/manuscript.Rmd
Progress on manuscript and ML analysis.

@@
 ## Integration into end-to-end machine learning
+
+We used 285 datasets from the Penn Machine Learning Benchmarks collection [@].
+In this collection, 122 datasets correspond to regression tasks and 163 datasets to classification tasks.
+Using the familiar auto-machine learning library [@], each dataset was used to train a model for each of 16 process configurations.
+Each process configuration specifies the learner (generalised linear model or random forest), 
+transformation method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central normality test ($p = 0.01$)),
+and normalisation method (none, $z$-standardisation), yielding 16 distinct configurations.
+
+Before each experiment, each dataset was randomly split into a training (70%) and holdout test (30%) set five times. Thus, a total of 22800 models were created.
+Each model was then evaluated using the holdout test set.
+
+- Plot differences vs. none transformation method.
+- Plot differences vs. none normalisation method and conventional transformation.
 

### 2024-06-04 manuscript/simulations.R

@@ -1755,7 +1755,7 @@
 
 
 
-.get_ml_experiment_data <- function(manuscript_dir) {
+.get_ml_experiment_data <- function(manuscript_dir, plot_theme) {
   # Converts experiment results into something that can be interpreted.
   if (file.exists(file.path(manuscript_dir, "parsed_ml_results.RDS"))) {
     return(readRDS(file.path(manuscript_dir, "parsed_ml_results.RDS")))
@@ -1767,6 +1767,7 @@
 
   # non-standard evaluation
   experiment_parameters <- dataset_split <- value <- value_rank <- metric <- NULL
+  normalisation_method <- NULL
 
   data <- readRDS(file.path(manuscript_dir, "ml_experiment", "results", "performance.RDS"))
 
@@ -1821,7 +1822,9 @@
   model <- ranger::ranger(
     value_rank ~ learner + transformation_method + normalisation_method + task_difficulty,
     data = model_data,
-    num.trees = 2000
+    min.node.size = 2L,
+    num.trees = 2000L,
+    seed = 19L
   )
 
   # Generate all combinations.
@@ -1835,67 +1838,96 @@
       )
     )
   )
-
-  predict(model, data=abstract_data[normalisation_method == "none"])
-
-  # Convert dataset to category.
-  results$dataset <- factor(
-    x = results$dataset,
-    levels = sort(unique(results$dataset))
+  abstract_data$task_difficulty <- factor(
+    abstract_data$task_difficulty,
+    levels=levels(abstract_data$task_difficulty),
+    ordered=TRUE
   )
 
-  # Rename experiment_parameters and convert to category.
-  results$experiment_parameters <- factor(
-    x = results$experiment_parameters,
-    levels = c("config_no_transformation", "config_old_test", "config_new_test"),
-    labels = c("none", "conventional", "proposed")
+  ..to_data_table <- function(x) {
+    r <- methods::new("rstream.runif", seed=9L)
+    x <- as.vector(x)
+    n <- min(c(100L, length(x)))
+
+    y <- sapply(
+      seq_len(1000L),
+      function(ii, x, r, n) {
+        random_values <- rstream::rstream.sample(r, n)
+        x_indices <- ceiling(random_values * length(x))
+        x_indices[x_indices == 0.0] <- 1.0
+
+        return(mean(x[x_indices]))
+      }, x = x, r = r, n = n
+    )
+
+    return(data.table::data.table("value" = y))
+  }
+
+  # Effect of learner
[…]

### 2024-06-04 manuscript/manuscript.Rmd
Add friedman test.

@@
 
-We used 285 datasets from the Penn Machine Learning Benchmarks collection [@].
-In this collection, 122 datasets correspond to regression tasks and 163 datasets to classification tasks.
-Using the familiar auto-machine learning library [@], each dataset was used to train a model for each of 16 process configurations.
-Each process configuration specifies the learner (generalised linear model or random forest), 
-transformation method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central normality test ($p = 0.01$)),
-and normalisation method (none, $z$-standardisation), yielding 16 distinct configurations.
-
-Before each experiment, each dataset was randomly split into a training (70%) and holdout test (30%) set five times. Thus, a total of 22800 models were created.
-Each model was then evaluated using the holdout test set.
+We used 285 datasets from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
+In this collection, 122 datasets correspond to regression tasks and 163 datasets
+to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
+each dataset was used to train a model for each of 16 process configurations.
+Each process configuration specifies the learner (generalised linear model or
+random forest), transformation method (none, conventional Yeo-Johnson, robust
+invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
+normality test ($p = 0.01$)), and normalisation method (none,
+$z$-standardisation), yielding 16 distinct configurations.
+Before each experiment, each dataset was randomly split into a training (70%)
+and holdout test (30%) set five times. Thus, a total of 22800 models were
+created. Each model was then evaluated using the holdout test set using one of
+two metrics, i.e. the root relative squared error (RRSE) for regression tasks and the
+area under the receiver operating characteristic curve (AUC) for classification tasks.
+
+For the purpose of assessing the effect of the difficulty of the task, we computed the median performance score over all models for each dataset and assigned one the following categories:
+
+* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (87 datasets)
+* easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (46 datasets)
+* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (72 datasets)
+* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (60 datasets)
+* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (12 datasets)
+* unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (8 datasets)
+
+To remove the effect of the dataset, and allow for comparing metrics, we ranked all performance scores for each dataset so that a higher rank corresponds to better performance.
+Experiments yielding the same score received the same, average, rank.
+Subsequently ranks were normalised to the $[0.0, 1.0]$ range.
+
+```{r include = FALSE, eval = TRUE}
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Friedman test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ learner + transformation_method + normalisation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+
+friedman_test_result <- friedman.test(as.matrix(data[, "dataset" := NULL]))
+```
 

### 2024-06-04 manuscript/simulations.R

@@ -1755,19 +1755,9 @@
 
 
 
-.get_ml_experiment_data <- function(manuscript_dir, plot_theme) {
-  # Converts experiment results into something that can be interpreted.
-  if (file.exists(file.path(manuscript_dir, "parsed_ml_results.RDS"))) {
-    return(readRDS(file.path(manuscript_dir, "parsed_ml_results.RDS")))
-  }
-
-  if (!file.exists(file.path(manuscript_dir, "ml_exp_results_offset.RDS"))) {
-    stop("Execute ml_exp.R to create the results and copy it to the manuscript folder.")
-  }
-
+.get_ml_experiment_data <- function(manuscript_dir) {
   # non-standard evaluation
   experiment_parameters <- dataset_split <- value <- value_rank <- metric <- NULL
-  normalisation_method <- NULL
 
   data <- readRDS(file.path(manuscript_dir, "ml_experiment", "results", "performance.RDS"))
 
@@ -1808,16 +1798,20 @@
   # 1].
   data[
     metric == "root_relative_squared_error_winsor",
-    "value_rank" := (data.table::frank(-value, ties.method = "average") - 1.0) / (.N - 1.0),
+    "value_rank" := data.table::frank(-value, ties.method = "average"),
     by = c("dataset")
   ]
   data[
     metric == "auc_roc",
-    "value_rank" := (data.table::frank(value, ties.method = "average") - 1.0) / (.N - 1.0),
+    "value_rank" := data.table::frank(value, ties.method = "average"),
     by = c("dataset")
   ]
-  data[, "value_rank" := (value_rank - 0.5) * 2.0]
+  data[, "value_rank" := (value_rank - min(value_rank)) / (max(value_rank) - min(value_rank)), by = c("dataset") ]
 
+  return(data)
+}
+
+temp_plot_fun <-function() {
   model_data <- data[, mget(c("learner", "transformation_method", "normalisation_method", "task_difficulty", "value_rank"))]
   model <- ranger::ranger(
     value_rank ~ learner + transformation_method + normalisation_method + task_difficulty,

### 2024-06-05 manuscript/manuscript.Rmd
WIP

@@
 ```
+
+Significant differences exist between process configurations (Friedman test: $p < 10^{-5})$.
+
+```{r include = FALSE, eval = TRUE}
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Wilcoxon signed rank test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ learner + transformation_method + normalisation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+
+data[, "dataset" := NULL]
+configurations <- colnames(data)
+```
+
+Something about direct comparisons.
+
+To assess the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] (version 0.16.0); 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
 

### 2024-06-05 manuscript/manuscript.Rmd
Add marginal effect plot.

@@
 
-friedman_test_result <- friedman.test(as.matrix(data[, "dataset" := NULL]))
-```
-
-Significant differences exist between process configurations (Friedman test: $p < 10^{-5})$.
+h_overall <- friedman.test(as.matrix(data[, "dataset" := NULL]))
+```
+
+Significant differences exist between process configurations (Friedman test: $p < 10^{-8})$.
 
@@
   data,
-  dataset ~ learner + transformation_method + normalisation_method, value.var="value_rank",
+  dataset ~ learner, value.var="value_rank",
   fun.aggregate = mean
@@
 
-data[, "dataset" := NULL]
-configurations <- colnames(data)
-```
-
-Something about direct comparisons.
+h_learner <- stats::wilcox.test(
+    x = data[["random_forest_ranger"]],
+    y = data[["glm"]],
+    paired = TRUE,
+    alternative = "two.sided"
+)
+```
+
+```{r include = FALSE, eval = TRUE}
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Wilcoxon signed rank test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ normalisation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+
+h_normalisation <- stats::wilcox.test(
+    x = data[["none"]],
+    y = data[["robust_standardisation"]],
+    paired = TRUE,
+    alternative = "two.sided"
+)
+```
+
+```{r include = FALSE, eval = TRUE}
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Friedman test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ transformation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+
+h_transformation <- friedman.test(as.matrix(data[, "dataset" := NULL]))
+```
+
+Considering single process parameters, the choice of learner (Wilcoxon signed rank test: $p < 10^{-8})$) and normalisation method (Wilcoxon signed rank test: $p = 0.03$) had a significant impact (at significance level $p = 0.05$), but transformation method (Friedman test: $p=0.11$) did not.
 
@@
 
+```{r include = FALSE, eval = FALSE}
+# CANDIDATE FOR REMOVAL !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Wilcoxon signed rank test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ learner + transformation_method + normalisation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+data[, "dataset" := NULL]
+
+# Compute p-values through pair-wise comparison.
+config_pairs <- utils::combn(colnames(data), m = 2L)
+test_data <- apply(
+  config_pairs,
+  MARGIN = 2,
+  function(config_pair, data) {
+    h <- stats::wilcox.test(
+      x = data[[config_pair[1L]]],
+      y = data[[config_pair[2L]]],
+      paired = TRUE,
+      alternative = "two.sided"
+    )
+
+    return(data.table::data.table(
+      "config_1" = config_pair[1L],
+      "config_2" = config_pair[2L],
+      "raw_p_value" = h$p.value
+    ))
+  },
+  data = data
+)
+test_data <- data.table::rbindlist(test_data)
+
+# Perform multiple testing correction.
+test_data[, "p_value_corrected" := stats::p.adjust(raw_p_value)]
+```
+
 # References

### 2024-06-05 manuscript/plots.R

@@ -1642,3 +1642,159 @@ get_annotation_settings <- function(ggtheme = NULL) {
 }
 
 
+
+.plot_marginal_effects <- function(manuscript_dir, plot_theme) {
+
+  learner <- transformation_method <- normalisation_method <- task_difficulty <- NULL
+
+  # Get ML experiment data.
+  data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+
+  # Build a random forest to create latent representations of the data.
+  model_data <- data[, mget(c("learner", "transformation_method", "normalisation_method", "task_difficulty", "value_rank"))]
+  model <- ranger::ranger(
+    value_rank ~ learner + transformation_method + normalisation_method + task_difficulty,
+    data = model_data,
+    min.node.size = 2L,
+    num.trees = 2000L,
+    seed = 19L
+  )
+
+  # Generate all combinations of process parameters and task difficulty.
+  abstract_data <- data.table::data.table(
+    expand.grid(
+      list(
+        "learner" = levels(data$learner),
+        "transformation_method" = levels(data$transformation_method),
+        "normalisation_method" = levels(data$normalisation_method),
+        "task_difficulty" = levels(data$task_difficulty)
+      )
+    )
+  )
+  abstract_data$task_difficulty <- factor(
+    abstract_data$task_difficulty,
+    levels=levels(abstract_data$task_difficulty),
+    ordered=TRUE
+  )
+
+  # Helper for random
+  ..to_data_table <- function(x) {
+    r <- methods::new("rstream.runif", seed=9L)
+    x <- as.vector(x)
+    n <- min(c(100L, length(x)))
+
+    y <- sapply(
+      seq_len(1000L),
+      function(ii, x, r, n) {
+        random_values <- rstream::rstream.sample(r, n)
+        x_indices <- ceiling(random_values * length(x))
+        x_indices[x_indices == 0.0] <- 1.0
+
+        return(mean(x[x_indices]))
+      }, x = x, r = r, n = n
+    )
+
+    return(data.table::data.table("value" = y))
+  }
+
+  # Effect of learner
+  x <- predict(model, data = abstract_data[learner == "random_forest_ranger"], predict.all = TRUE)$predictions -
+    predict(model, data = abstract_data[learner == "glm"], predict.all = TRUE)$predictions
+  data_1 <- ..to_data_table(x)
+
+  # Effect of normalisation method
[…]

### 2024-06-05 manuscript/simulations.R

@@ -1811,117 +1811,3 @@
   return(data)
 }
 
-temp_plot_fun <-function() {
-  model_data <- data[, mget(c("learner", "transformation_method", "normalisation_method", "task_difficulty", "value_rank"))]
-  model <- ranger::ranger(
-    value_rank ~ learner + transformation_method + normalisation_method + task_difficulty,
-    data = model_data,
-    min.node.size = 2L,
-    num.trees = 2000L,
-    seed = 19L
-  )
-
-  # Generate all combinations.
-  abstract_data <- data.table::data.table(
-    expand.grid(
-      list(
-        "learner" = levels(data$learner),
-        "transformation_method" = levels(data$transformation_method),
-        "normalisation_method" = levels(data$normalisation_method),
-        "task_difficulty" = levels(data$task_difficulty)
-      )
-    )
-  )
-  abstract_data$task_difficulty <- factor(
-    abstract_data$task_difficulty,
-    levels=levels(abstract_data$task_difficulty),
-    ordered=TRUE
-  )
-
-  ..to_data_table <- function(x) {
-    r <- methods::new("rstream.runif", seed=9L)
-    x <- as.vector(x)
-    n <- min(c(100L, length(x)))
-
-    y <- sapply(
-      seq_len(1000L),
-      function(ii, x, r, n) {
-        random_values <- rstream::rstream.sample(r, n)
-        x_indices <- ceiling(random_values * length(x))
-        x_indices[x_indices == 0.0] <- 1.0
-
-        return(mean(x[x_indices]))
-      }, x = x, r = r, n = n
-    )
-
-    return(data.table::data.table("value" = y))
-  }
-
-  # Effect of learner
-  x <- predict(model, data = abstract_data[learner == "random_forest_ranger"], predict.all = TRUE)$predictions -
-    predict(model, data = abstract_data[learner == "glm"], predict.all = TRUE)$predictions
-  data_1 <- ..to_data_table(x)
-
-  # Effect of transformation method vs. none, by learner, task difficulty and
-  # normalisation method.
-  result_list <- list()
-
-  for (method in levels(abstract_data$transformation_method)) {
-    if (method == "none") next
-
-    for (lrnr in levels(abstract_data$learner)) {
-      for (norm_method in levels(abstract_data$normalisation_method)){
-
-        x <- predict(model, data = abstract_data[learner == lrnr & transformation_method == method & normalisation_method == norm_method], predict.all = TRUE)$predictions -
[…]

### 2024-06-07 manuscript/manuscript.Rmd
Marginal effect plot is done.

@@
 plot_theme <- ggplot2::theme_light(base_size=base_size)
-plot_theme$plot.title$size <- ggplot2::rel(1.0)
+plot_theme$plot.title$size <- ggplot2::rel(0.8)
+plot_theme$plot.title$hjust <- 0.5
 plot_theme$strip.background <- ggplot2::element_blank()
@@
 
-To assess the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] (version 0.16.0); 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
-
-- Plot differences vs. none transformation method.
-- Plot differences vs. none normalisation method and conventional transformation.
-
-- Assess whether shift-invariant robust transformation improves machine learning results.
-
-  - No transformation.
-  - Conventional Yeo-Johnson transformation.
-  - Shift sensitive Yeo-Johnson transformation.
-  - Shift sensitive, robust Yeo-Johnson transformation.
-  - Shift sensitive, robust Yeo-Johnson with central normality test.
-
-  TODO: 
-  - implement into familiar
-  - update familiar.experiment to allow for setting external configurations (or maybe make that the default -- its definitely easier and more transparent.)
-
-  - Add tests for out-of-bound actions (power_transform) and GOF test rejection (find_parameters).
+To estimate the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] (version 0.16.0); 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
+The estimated marginal effects are shown in Figure \ref{fig:marginal-effect-plot}.
+On the scale of normalised ranks ($[0.0, 1.0]$), the overall estimated marginal effect of using a random forest instead of generalised linear model was $0.226$.
+The overall marginal effect of using z-standardisation to normalise features was $0.003$.
+Transformation methods had the following marginal effects:
+$0.002$ for using conventional Yeo-Johnson transformation instead of no transformation;
+$-0.006$ for using robust invariant Yeo-Johnson transformation (with or without empirical central normality test) instead of no transformation; 
+and $-0.008$ for using robust invariant Yeo-Johnson transformation (with or without empirical central normality test) instead of conventional Yeo-Johnson transformation.
+
+```{r marginal-effect-plot, echo=FALSE, fig.cap=cap, fig.height=12.0, warning=FALSE}
+cap <- paste0(
+  "Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 22800 machine learning experiments on 285 datasets. ",
+  "The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). ",
+  "Random forests outperform GLM models for all task difficulties. ",
+  "The top-right panel shows the marginal effect of feature normalisation methods, i.e. no normalisation and z-standardisation. ",
+  "z-standardisation is generally beneficial, but the estimated effect is neglible. ",
+  "The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method. ",
+  "There is no consistent behaviour, and estimated effects are neglible. ",
+  "Note that the ranges of the $x$-axes of the three main panels differ. ",
+  "ECTN: empirical central normality test."
+)
+
+.plot_marginal_effects(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
+```
 
@@
 **INSERT REFERENCE** for the theoretical analysis in their paper [@Box1964-mz].
+
+We investigated the use of robust invariant power transformation for machine learning, and did not find a major positive or negative effect.
 

### 2024-06-07 manuscript/plots.R

@@ -1643,7 +1643,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-.plot_marginal_effects <- function(manuscript_dir, plot_theme) {
+.plot_marginal_effects <- function(plot_theme, manuscript_dir) {
 
   learner <- transformation_method <- normalisation_method <- task_difficulty <- NULL
 
@@ -1805,11 +1805,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
     labels = difficulty_labels
   )
 
+  data_3_x_range = c(-0.07, 0.07)
   data_3_labels <- data.table::data.table(
-    learner = factor("GLM", levels = levels(data_3$learner)),
-    normalisation_method = factor("no normalisation", levels = levels(data_3$normalisation_method)),
+    learner = factor("random forest", levels = levels(data_3$learner)),
+    normalisation_method = factor("z-standardisation", levels = levels(data_3$normalisation_method)),
     transformation_method = factor(transformation_method_labels),
-    task_difficulty  = factor("unsolvable", levels = difficulty_labels),
+    task_difficulty  = factor("overall", levels = difficulty_labels),
     label_1 = c("none better", "none better", "none better", "conv. better", "conv. better"),
     label_2 = c("conv. better", "invar. better", "invar. better", "invar. better", "invar. better")
   )
@@ -1829,7 +1830,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p3 <- p3 + ggplot2::geom_col(show.legend = FALSE)
   p3 <- p3 + ggplot2::facet_grid(learner + normalisation_method ~ transformation_method)
   p3 <- p3 + ggplot2::geom_text(
-    x = -0.10,
+    x = data_3_x_range[1L],
     ggplot2::aes(label=label_1),
     hjust = "inward",
     vjust = "outward",
@@ -1840,7 +1841,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     size = annotation_settings$geom_text_size * 0.8
   )
   p3 <- p3 + ggplot2::geom_text(
-    x = 0.10,
+    x = data_3_x_range[2L],
     ggplot2::aes(label=label_2),
     hjust = "inward",
     vjust = "outward",
[…]

### 2024-06-07 manuscript/manuscript.Rmd
Update to only consider datasets with numeric features.

@@
 
-Considering single process parameters, the choice of learner (Wilcoxon signed rank test: $p < 10^{-8})$) and normalisation method (Wilcoxon signed rank test: $p = 0.03$) had a significant impact (at significance level $p = 0.05$), but transformation method (Friedman test: $p=0.11$) did not.
+Here we focus on the subset of 232 datasets that contain numeric features. 
+Considering single process parameters, the choice of learner (Wilcoxon signed rank test: $p < 10^{-8})$) and normalisation method (Wilcoxon signed rank test: $p = 0.007$) had a significant impact (at significance level $p = 0.050$), but transformation method (Friedman test: $p = 0.054$) did not.
 
@@
 The estimated marginal effects are shown in Figure \ref{fig:marginal-effect-plot}.
-On the scale of normalised ranks ($[0.0, 1.0]$), the overall estimated marginal effect of using a random forest instead of generalised linear model was $0.226$.
-The overall marginal effect of using z-standardisation to normalise features was $0.003$.
+On the scale of normalised ranks ($[0.0, 1.0]$), the overall estimated marginal effect of using a random forest instead of generalised linear model was $0.272$.
+The overall marginal effect of using z-standardisation to normalise features was $0.008$.
 Transformation methods had the following marginal effects:
-$0.002$ for using conventional Yeo-Johnson transformation instead of no transformation;
-$-0.006$ for using robust invariant Yeo-Johnson transformation (with or without empirical central normality test) instead of no transformation; 
-and $-0.008$ for using robust invariant Yeo-Johnson transformation (with or without empirical central normality test) instead of conventional Yeo-Johnson transformation.
+$0.003$ for using conventional Yeo-Johnson transformation instead of no transformation;
+$-0.007$ for using robust invariant Yeo-Johnson transformation instead of no transformation; 
+$-0.009$ for using robust invariant Yeo-Johnson transformation with empirical central normality test instead of no transformation; 
+$-0.010$ for using robust invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson transformation;
+and $-0.012$ for using using robust invariant Yeo-Johnson transformation with empirical central normality test instead of conventional Yeo-Johnson transformation.
 
@@
 
-We investigated the use of robust invariant power transformation for machine learning, and did not find a major positive or negative effect.
+We investigated the use of robust invariant power transformation as part of the machine learning process, and did not find a major positive or negative effect.
+A possible explanation is that numeric features with large location shifts ($|\mu| > 1000.0$ were uncommon.
+Of the 4886 numeric features in the 232 datasets with numeric features, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
 

### 2024-06-07 manuscript/plots.R

@@ -1643,12 +1643,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-.plot_marginal_effects <- function(plot_theme, manuscript_dir) {
+.plot_marginal_effects <- function(plot_theme, manuscript_dir, subset = "numeric") {
 
   learner <- transformation_method <- normalisation_method <- task_difficulty <- NULL
 
   # Get ML experiment data.
-  data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+  data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir, subset = subset)
 
   # Build a random forest to create latent representations of the data.
   model_data <- data[, mget(c("learner", "transformation_method", "normalisation_method", "task_difficulty", "value_rank"))]
@@ -1805,7 +1805,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     labels = difficulty_labels
   )
 
-  data_3_x_range = c(-0.07, 0.07)
+  data_3_x_range = c(-0.08, 0.08)
   data_3_labels <- data.table::data.table(
     learner = factor("random forest", levels = levels(data_3$learner)),
     normalisation_method = factor("z-standardisation", levels = levels(data_3$normalisation_method)),
@@ -1868,7 +1868,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p3 <- p3 + ggplot2::ylab("task difficulty")
   p3 <- p3 + ggplot2::xlab("normalised rank difference")
 
-  data_2_x_range = c(-0.003, 0.0085)
+  data_2_x_range = c(-0.005, 0.03)
   data_2_labels <- data.table::data.table(
     x = data_2_x_range,
     task_difficulty  = factor(c("overall", "overall"), levels = difficulty_labels),
@@ -1914,7 +1914,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p2 <- p2 + ggplot2::xlab("normalised rank difference")
   p2 <- p2 + ggplot2::ggtitle("z-standardisation vs. none")
 
-  data_1_x_range = c(-0.05, 0.30)
+  data_1_x_range = c(-0.05, 0.35)
   data_1_labels <- data.table::data.table(
     x = data_1_x_range,
     task_difficulty  = factor(c("overall", "overall"), levels = difficulty_labels),

### 2024-06-07 manuscript/simulations.R

@@ -1755,7 +1755,7 @@
 
 
 
-.get_ml_experiment_data <- function(manuscript_dir) {
+.get_ml_experiment_data <- function(manuscript_dir, subset = "numeric") {
   # non-standard evaluation
   experiment_parameters <- dataset_split <- value <- value_rank <- metric <- NULL
 
@@ -1808,6 +1808,68 @@
   ]
   data[, "value_rank" := (value_rank - min(value_rank)) / (max(value_rank) - min(value_rank)), by = c("dataset") ]
 
+  # Select subset of data.
+  if (subset == "numeric") {
+    feature_data <- .get_ml_experiment_feature_statistics(manuscript_dir = manuscript_dir)
+    feature_data[, "dataset_real" := gsub(pattern = "_data.RDS", replacement = "", x = dataset, fixed = TRUE)]
+    data <- data[dataset %in% feature_data$dataset_real]
+
+  } else if (subset == "numeric_high_shift") {
+    feature_data <- .get_ml_experiment_feature_statistics(manuscript_dir = manuscript_dir)
+    feature_data <- feature_data[abs(mu) > 500.0]
+    feature_data[, "dataset_real" := gsub(pattern = "_data.RDS", replacement = "", x = dataset, fixed = TRUE)]
+    data <- data[dataset %in% feature_data$dataset_real]
+  }
+
   return(data)
 }
 
+
+
+.get_ml_experiment_feature_statistics <- function(manuscript_dir) {
+
+  file_name <- file.path(manuscript_dir, "ml_experiment_feature_statistics.RDS")
+
+  if (!file.exists(file_name)) {
+    data_dir <- file.path(manuscript_dir, "ml_experiment", "data_cache")
+    data_files <- list.files(path = data_dir, full.names = FALSE, pattern = ".RDS")
+
+    feature_data <- lapply(
+      data_files,
+      function(data_file, data_dir) {
+        # Read datasets, which are familiar DataObject objects.
+        dataset <- readRDS(file.path(data_dir, data_file))
+        feature_data <- lapply(
+          familiar:::get_feature_columns(dataset),
+          function(feature, dataset) {
+            if (is.numeric(dataset@data[[feature]])) {
+              x <- power.transform::huber_estimate(dataset@data[[feature]])
+              return(data.table::data.table("feature" = feature, "mu" = x$mu, "sigma" = x$sigma))
+
+            } else {
+              return(NULL)
+            }
+          },
+          dataset = dataset
[…]

### 2024-06-19 manuscript/manuscript.Rmd
Work on real-life datasets.

@@
 
-In the previous section, we simulated data to assess shift-sensitive power transformations and their robustness against outliers. In this section, we assess these aspects using real datasets.
-
-## Datasets with outliers
+## Robustness against outliers
+
+In the previous section, we simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.
 
@@
   "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
-  "This data contains three strong outliers for highly efficient vehicles. ",
+  "The dataset contains three strong outliers for highly fuel-efficient vehicles.",
   "The left panel shows the quantile-quantile plot for the original data. ",
+  "The empirical central normality test indicates that the original data do not deviate significantly from central normality ($p=0.76$).",
   "The middle panel shows the quantile-quantile plots for data transformed using ",
   "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.55$, respectively. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust invariant transformations. ",
+  "For these, the empirical central normality test yields $p=0.03$, and $p=0.86$, respectively. ",
+  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
+)
+
+.plot_top_gear(
+  plot_theme = plot_theme
+)
+```
+
+The Top Gear dataset contains data on 297 vehicles that appeared on the BBC television show *Top Gear*.
+Within this dataset, the fuel consumption feature contains outliers due to highly fuel-efficient vehicles.
+Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively).
+Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
+
+### Ischemic stroke dataset
+
+```{r ischemic-stroke, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for the maximum arterial wall thickness in the ischemic stroke dataset [@Romano2022-gq]. ",
+  "This data contains two strong outliers. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The empirical central normality test indicates that the original data do not deviate significantly from central normality ($p=0.56$).",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
+  "For these, the empirical central normality test yields $p=0.01$, and $p=0.61$, respectively. ",
   "The right panels shows the quantile-quantile plots for data transformed using ",
   "the non-robust and robust shift-sensitive transformations. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
+  "For these, the empirical central normality test yields $p=0.02$, and $p=0.10$, respectively. ",
   "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
@@
 
-.plot_top_gear(
+.plot_ischemic_stroke(
   plot_theme = plot_theme
@@
 ```
-
-The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*.
-Fuel consumption contains outliers.
-Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$).
+The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
+These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
+The angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
+The maximum arterial wall thickness feature contains several instances with outlier values.
+Figure \ref{fig:ischemic-stroke} show that non-robust transformation fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
 
-### Housing dataset
-
-```{r housing-data, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
-  "This data contains three strong outliers for highly efficient vehicles. ",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.58$, respectively. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust shift-sensitive transformations. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.36$, respectively. ",
-  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
-)
-
-.plot_sacramento_house(
-  plot_theme = plot_theme
-)
-```
-
-modeldata::Sacramento$price
-
-Ames housing data set, year built
-
-### Other data ???
+## Invariance
+
+Datasets that where invariance has better properties in terms of lambda and resulting mean and standard deviation.
+modeldata::ischemic_stroke$age
+modeldata::ames$Longitudinal
 
@@
 We investigated the use of robust invariant power transformation as part of the machine learning process, and did not find a major positive or negative effect.
-A possible explanation is that numeric features with large location shifts ($|\mu| > 1000.0$ were uncommon.
+A possible explanation is that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
 Of the 4886 numeric features in the 232 datasets with numeric features, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.

### 2024-06-19 manuscript/plots.R

@@ -1391,11 +1391,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-.plot_sacramento_house <- function(plot_theme) {
+.plot_ischemic_stroke <- function(plot_theme) {
   require(modeldata)
 
-  data <- data.table::as.data.table(modeldata::Sacramento)
-  x <- data$price
+  x <- modeldata::ischemic_stroke$max_max_wall_thickness
 
   return(
     ..plot_real_data(
@@ -1407,7 +1406,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-..plot_real_data <- function(plot_theme, x) {
+..plot_real_data <- function(plot_theme, x, lambda_limit = c(-4.0, 6.0)) {
   # Show density and Q-Q plots for:
   # - Original data
   # - conventional transformation
@@ -1430,6 +1429,13 @@ get_annotation_settings <- function(ggtheme = NULL) {
     transformed_data <- power.transform::power_transform(
       x = x,
       transformer = transformer
+    )
+
+    summary_data <- data.table::data.table(
+      "mu" = mean(transformed_data),
+      "sigma" = sd(transformed_data),
+      "lambda" = power.transform::get_lambda(transformer),
+      "transformation" = name
     )
 
     transformed_data <- data.table::data.table(
@@ -1457,6 +1463,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     return(list(
       "transformed_data" = transformed_data,
       "residual_data" = residual_data,
+      "summary_data" = summary_data,
       "central_normality_data" = central_normality_data
     ))
   }
@@ -1472,7 +1479,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     method = "yeo_johnson",
     robust = FALSE,
-    invariant = FALSE
+    invariant = FALSE,
+    lambda = lambda_limit
   )
 
   # Raymaekers robust transformer
@@ -1481,7 +1489,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
     method = "yeo_johnson",
     robust = TRUE,
     invariant = FALSE,
-    estimation_method = "raymaekers_robust"
+    estimation_method = "raymaekers_robust",
+    lambda = lambda_limit
   )
 
   # Shift-sensitive transformer
[…]

### 2024-06-21 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
+## Invariance
+
+Invariant power transformations are intended to yield improved transformations to normality
+in the presence of large shifts in location, scales far from unity, or both.
+We previously demonstrated this property using simulations, and will now demonstrate
+it using examples from real datasets.
+
+### Age in a ...
+
+In health-related datasets patient age is a common characteristic.
+
+Show invariance to units (days vs years)
+
+### Body mass of penguins
+modeldata::penguins$body_mass_g
+
+### Longitudinal coordinates in the Ames housing dataset
+
+Datasets where invariance has better properties in terms of lambda and resulting mean and standard deviation.
+modeldata::ischemic_stroke$age
+modeldata::ames$Longitudinal
+
 ## Robustness against outliers
 
-In the previous section, we simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.
-
-### Top Gear dataset
+We previously simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.
+
+### Fuel efficiency in the Top Gear dataset
 
@@
 
-### Ischemic stroke dataset
+### Maximum arterial wall thickness in an ischemic stroke dataset
 
@@
 These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
-The angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
+Angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
 The maximum arterial wall thickness feature contains several instances with outlier values.
@@
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
-
-## Invariance
-
-Datasets that where invariance has better properties in terms of lambda and resulting mean and standard deviation.
-modeldata::ischemic_stroke$age
-modeldata::ames$Longitudinal
 

### 2024-07-02 manuscript/manuscript.Rmd
Additional examples.

@@
 in the presence of large shifts in location, scales far from unity, or both.
-We previously demonstrated this property using simulations, and will now demonstrate
-it using examples from real datasets.
-
-### Age in a ...
-
-In health-related datasets patient age is a common characteristic.
-
-Show invariance to units (days vs years)
-
-### Body mass of penguins
-modeldata::penguins$body_mass_g
+Above, this characteristic was assessed using simulations. Below, this characteristic
+will be evaluated using examples from real datasets.
+
+### Age of patients with lung cancer
+
+A common variable in health-related datasets is age.
+Here we use data on 228 patients with lung cancer that was collected and published by Loprinzi et al. [@Loprinzi1994-cd].
+The age in the cohort was $62.4 \pm 9.1$ (mean ± standard deviation) years.
+Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:lung-cancer-age}:
+no transformation (residual sum $\sum r_i = 16.5$);
+conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
+shift- and scale invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
+and robust shift- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).
+
+```{r lung-cancer-age, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for age of patients with lung cancer, expressed in years. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust invariant transformations."
+)
+
+out <- .plot_lung_cancer_age(plot_theme = plot_theme)
+out$plot
+```
+
+Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
+Moreover, conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age variable after transformation.
+The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.78$).
+
+### Penguin body mass
+
+Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
+The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams.
+Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:penguin-body-mass}:
+no transformation (residual sum $\sum r_i = 48.0$);
+conventional transformation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.102$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
+shift- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
+and robust shift- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).
+
+```{r penguin-body-mass, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for penguin body mass, measured in grams. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust invariant transformations."
+)
+
+out <- .plot_penguin_body_mass(plot_theme = plot_theme)
+out$plot
+```
 
@@
 cap <- paste0(
-  "Quantile-quantile plots for the fuel consumption data in the Top Gear dataset. ",
+  "Quantile-quantile plots for fuel consumption data in the Top Gear dataset. ",
   "The dataset contains three strong outliers for highly fuel-efficient vehicles.",
@@
 
-.plot_top_gear(
+out <- .plot_top_gear(
   plot_theme = plot_theme
 )
+
+out$plot
 ```
@@
 
-.plot_ischemic_stroke(
+out <- .plot_ischemic_stroke(
   plot_theme = plot_theme
 )
+
+out$plot
 ```

### 2024-07-02 manuscript/plots.R

@@ -1370,6 +1370,37 @@ get_annotation_settings <- function(ggtheme = NULL) {
 }
 
 
+.plot_lung_cancer_age <- function(plot_theme, lambda_limit = NULL) {
+  require(survival)
+
+  return(
+    ..plot_real_data(
+      plot_theme = plot_theme,
+      x = survival::lung$age,
+      lambda_limit = NULL
+    )
+  )
+}
+
+
+
+.plot_penguin_body_mass <- function(plot_theme, lambda_limit = NULL) {
+  require(modeldata)
+
+  x <- modeldata::penguins$body_mass_g
+  x <- x[is.finite(x)]
+
+  return(
+    ..plot_real_data(
+      plot_theme = plot_theme,
+      x = x,
+      lambda_limit = NULL
+    )
+  )
+}
+
+
+
 .plot_top_gear <- function(plot_theme) {
   require(robustHD)
 
@@ -1435,6 +1466,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
       "mu" = mean(transformed_data),
       "sigma" = sd(transformed_data),
       "lambda" = power.transform::get_lambda(transformer),
+      "shift" = power.transform::get_shift(transformer),
+      "scale" = power.transform::get_scale(transformer),
       "transformation" = name
     )
 
@@ -1528,6 +1561,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     MoreArgs = list("x" = x),
     SIMPLIFY = FALSE
   )
+  names(data) <- transformer_labels
 
   # Parse transformed data
   transformed_data <- data.table::rbindlist(
@@ -1650,7 +1684,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     heights = c(0.5, 1.0)
   )
 
-  return(p)
+  return(list("data" = data, "plot"= p))
 }
 
 

### 2024-07-02 manuscript/manuscript.Rmd
Add comment on distribution.

@@
 ```
+Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
+Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass variable after transformation.
+The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
+The distribution of the original data was not centrally normal ($p = 0.03$).
 

### 2024-07-02 manuscript/manuscript.Rmd
Latitude

@@
 Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
-The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams.
+The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
 Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:penguin-body-mass}:
@@
 The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
-The distribution of the original data was not centrally normal ($p = 0.03$).
-
-### Longitudinal coordinates in the Ames housing dataset
-
-Datasets where invariance has better properties in terms of lambda and resulting mean and standard deviation.
-modeldata::ischemic_stroke$age
-modeldata::ames$Longitudinal
+
+### Latitude in the Ames housing dataset
+
+Geospatial datasets usually contain coordinates.
+The Ames housing dataset contains data on 2930 properties that were sold between 2006 and 2010 [@De-Cock2011-jf], including their geospatial coordinates.
+The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
+Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:ames-latitude}:
+no transformation (residual sum $\sum r_i = 328$);
+conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = $, $\sigma_{YJ} = $);
+shift- and scale invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = $, $\sigma_{YJ} = $);
+and robust shift- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = $, $\sigma_{YJ} = $).
+
+```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for penguin body mass, measured in grams. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust invariant transformations."
+)
+
+out <- .plot_ames_latitude(plot_theme = plot_theme)
+out$plot
+```
 

### 2024-07-03 manuscript/manuscript.Rmd
WIP on manuscript

--- 2024-07-02 manuscript/manuscript.Rmd
@@
 
--   We devise shift and scale invariant versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
+-   We devise location- and scale-invariant versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
 
@@
 Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
-The location and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
+The location- and scale-invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
 
@@
 where $x_i - x_0 > 0$.
-Likewise, the shift and scale invariant Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is:
+Likewise, the location- and scale-invariant Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is:
 
@@
 
-The optimal weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.
+The optimal weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust, location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss for \textbf{shift-sensitive Box-Cox} power transformations.
+\caption{Optimal weighting parameters and corresponding loss for \textbf{location- and scale-invariant Box-Cox} power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
 As before, for each distribution, up to $10 \%$ of instances are replaced by outlier values, and this is repeated $10$ times.
-We then compute residuals after optimising robust shift-sensitive transformations with the empirical tapered cosine weighting method.
+We then compute residuals after optimising robust location- and scale-invariant transformations with the empirical tapered cosine weighting method.
 
@@
   "Residual error as a function of empirical probability. ",
-  "Percentiles of the error are shown for robust, shift-sensitive transformations of 10000 randomly drawn asymmetric ",
+  "Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric ",
   "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
@@
 Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80.
-In the remaining we will use the empirical central normality test statistic values defined using robust shift-sensitive Yeo-Johnson transformations, as it is slightly more strict.
+In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is slightly more strict.
 This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}
@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
-shift- and scale invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
-and robust shift- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).
+location- and scale invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
+and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).
 
@@
 
-Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
+Location- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
 Moreover, conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age variable after transformation.
@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
-shift- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
-and robust shift- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).
+location- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
+and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).
 
@@
 ```
-Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
+Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
 Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass variable after transformation.
 The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
-The distribution of the original data was not centrally normal ($p = 0.03$).
-
-### Longitudinal coordinates in the Ames housing dataset
-
-Datasets where invariance has better properties in terms of lambda and resulting mean and standard deviation.
-modeldata::ischemic_stroke$age
-modeldata::ames$Longitudinal
+
+### Latitude in the Ames housing dataset
+
+Geospatial datasets usually contain coordinates.
+The Ames housing dataset contains data on 2930 properties that were sold between 2006 and 2010 [@De-Cock2011-jf], including their geospatial coordinates.
+The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
+Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:ames-latitude}:
+no transformation (residual sum $\sum r_i = 328$);
+conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$);
+location- and scale invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
+and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$).
+
+```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for latitude coordinate in a housing dataset. ",
+  "The left panel shows the quantile-quantile plot for the original data. ",
+  "The middle panel shows the quantile-quantile plots for data transformed using ",
+  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
+  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "the non-robust and robust invariant transformations."
+)
+
+out <- .plot_ames_latitude(plot_theme = plot_theme)
+out$plot
+```
+Every transformation reduced the residual sum.
+The non-robust location- and scale-invariant transformation did not improve over conventional alternatives and yielded a data distribution lacking central normality (empirical central normality test: $p=0.05$).
+TranConventional transformations had high values for the $\lambda$ parameter.
 
@@
 
-We previously simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.
+We previously simulated data to assess invariant power transformations and their robustness against outliers.
+Here, we assess invariant power transformations in real data with outliers.
 
@@
   "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust shift-sensitive transformations. ",
+  "the non-robust and robust location- and scale-invariant transformations. ",
   "For these, the empirical central normality test yields $p=0.02$, and $p=0.10$, respectively. ",
@@
 
-In their work, Box and Cox already mention a transformation with a shift parameter, but preferred the now well-known version in Eq.
-**INSERT REFERENCE** for the theoretical analysis in their paper [@Box1964-mz].
-
-We investigated the use of robust invariant power transformation as part of the machine learning process, and did not find a major positive or negative effect.
+In their work on power transformation, Box and Cox already mention transformation with a shift parameter, but preferred the version in Eq.
+**INSERT REFERENCE** for the theoretical analysis in their paper [@Box1964-mz], which subsequently became the convention.
+Yeo and Johnson's power transformation lacks a shift parameter altogether [@Yeo2000-vw].
+We showed that these power transformations are sensitive to location and scale of data distributions.
+To mitigate this issue, we defined location- and scale-invariant variants of the Box-Cox and Yeo-Johnson transformations.
+We furthermore assessed methods for making these transformations robust to outliers, and devised an empirical test for central normality.
+We showed that Conventional transformations The location- and scale-invariant transformations removed the effect of location shifts away from 0 and scaling away from 1 on estimation of the transformation parameter $\lambda$.
+
+We investigated the use of robust invariant power transformation as part of the machine learning process, and did not find a major positive or negative effect compared to conventional transformations.
 A possible explanation is that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
@@
 
-Shift-sensitive power transformations are implemented in the `power.transform` package.
+Location- and scale-invariant power transformations are implemented in the `power.transform` package.
 This manuscript was created using R Markdown, and is available from [https://github.com/alexzwanenburg/power.transform].

### 2024-07-03 manuscript/manuscript.Rmd
WIP on the discussion.

@@
 In their work on power transformation, Box and Cox already mention transformation with a shift parameter, but preferred the version in Eq.
-**INSERT REFERENCE** for the theoretical analysis in their paper [@Box1964-mz], which subsequently became the convention.
+\ref{eqn:box-cox-original} for the theoretical analysis in their paper [@Box1964-mz], which subsequently became the convention.
 Yeo and Johnson's power transformation lacks a shift parameter altogether [@Yeo2000-vw].
@@
 We furthermore assessed methods for making these transformations robust to outliers, and devised an empirical test for central normality.
-We showed that Conventional transformations The location- and scale-invariant transformations removed the effect of location shifts away from 0 and scaling away from 1 on estimation of the transformation parameter $\lambda$.
-
-We investigated the use of robust invariant power transformation as part of the machine learning process, and did not find a major positive or negative effect compared to conventional transformations.
-A possible explanation is that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
-Of the 4886 numeric features in the 232 datasets with numeric features, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
+
+Robust location- and scale-invariant transformations form a suitable replacement for their conventional counterparts.
+They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the data distribution.
+This is particularly relevant for automated data processing, where such issues may go unnoticed.
+However, we did not find a measurable benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric variable.
+One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
+Of the 4886 numeric variables in the 232 datasets, 266 (5%) variables in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
+
+Robust location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e.
+transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$, using maximum likelihood optimisation.
+It is feasible that standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional transformations may achieve a similar effect in mitigating sensitivity to the distribution's location and scale.
 

### 2024-07-03 manuscript/manuscript_appendix.Rmd
WIP on the discussion.

@@
 
-Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parameterised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
+Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parametrised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
 This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:

### 2024-07-04 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$, using maximum likelihood optimisation.
-It is feasible that standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional transformations may achieve a similar effect in mitigating sensitivity to the distribution's location and scale.
+Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
+
+This study had the following limitations.
+First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle variables with negative values, and did not consider the Box-Cox transformation.
+Likewise, we focussed on maximum likelihood estimation to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B).
 
 # Conclusion
+
+Robust location- and scale-invariant power transformations allow for automated power transformation of numerical variables/
 

### 2024-08-21 manuscript/manuscript.Rmd
WIP on manuscript.
We are getting there!

@@
   "For the left-hand panels, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
-  "whereas for the right-hand row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
+  "whereas for the right-hand row, the distribution is $\\mathcal{N}(0, \\sigma)$. " ,
   "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
   "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and breaks down due to numerical issues for $\\mu > 5 \\cdot 10^3$ (both Box-Cox and Yeo-Johnson).",
-  "Scale did not affect Box-Cox and Yeo-Johnson transformations of variables drawn from normal distributions."
+  "and breaks down due to numerical issues for $\\mu > 5 \\cdot 10^3$. ",
+  "Here, scale did not affect Box-Cox and Yeo-Johnson transformations of variables drawn from normal distributions."
 )
@@
 \begin{equation}
-\label{eqn:box-cox-shift}
+\label{eqn:box-cox-invariant}
 \phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
@@
 \begin{equation}
-\label{eqn:yeo-johnson-shift}
+\label{eqn:yeo-johnson-invariant}
 \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
@@
 \begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
@@
 where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
-Similarly, for the invariant Yeo-Johnson log-likelihood function is:
+Similarly, the invariant Yeo-Johnson log-likelihood function is:
 
@@
 \begin{split}
-\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
-Because $\lambda$ is updated during their procedure, it can not be used simultaneously estimate $\lambda$, $x_0$ and $s$ for invariant power transformations.
+Because this procedure iteratively updates $\lambda$, it can not be used to simultaneously estimate $\lambda$, $x_0$ and $s$ for invariant power transformations.
 
 In essence, obtaining robust transformation parameters that provide central normality requires identifying outliers in the data and weighting such instances during the optimisation process.
-@Raymaekers2021-kq do this (during the final steps of their iterative procedure) through a weighted maximum likelihood estimation.
+@Raymaekers2021-kq achieve this through weighted maximum likelihood estimation.
 Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
@@
 
-```{r optimal_weighting_function_all_criteria_box_cox, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Things."
-)
-# This finds lambda values at the optimised weighting parameter settings.
-data <- .get_transformation_parameters(manuscript_dir = manuscript_dir)
-
-data <- data[method == "Box-Cox"]
-data <- data[, "lambda_error" := abs(lambda - target_lambda)]
-non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
-data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
-
-data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
-data[k == 0.0, "better_than_non_robust" := NA_integer_]
-data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
-
-pass_rate_data <- data[, list(
-  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
-  "n" = sum(!is.na(better_than_non_robust)),
-  "median_error" = stats::median(lambda_error)),
-  by=c("method", "estimation_method", "weight_method")]
-pass_rate_data <- pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)]
-
-p <- ggplot2::ggplot(
-  data = data,
-  mapping = ggplot2::aes(
-    x = weight_method,
-    y = lambda_error,
-    fill = estimation_method))
-p <- p + plot_theme
-p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
-p <- p + ggplot2::facet_wrap("estimation_method", nrow=nlevels(data$estimation_method))
-p <- p + paletteer::scale_fill_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE)
-p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
-p <- p + ggplot2::xlab("weighting method")
-p <- p + ggplot2::ggtitle("shift-sensitive Box-Cox transformation")
-p <- p + ggplot2::theme(
-  axis.text.x = ggplot2::element_text(
-    angle=30.0,
-    hjust=1.0))
-p
-```
-
-```{r optimal_weighting_function_all_criteria_yeo_johnson, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Things."
-)
-# This finds lambda values at the optimised weighting parameter settings.
-data <- .get_transformation_parameters(manuscript_dir = manuscript_dir)
-
-data <- data[method == "Yeo-Johnson"]
-data <- data[, "lambda_error" := abs(lambda - target_lambda)]
-non_robust_data <- data[weight_method == "non-robust", mget(c("method", "estimation_method", "k", "ii", "lambda_error"))]
-data[non_robust_data, on=c("method", "estimation_method", "k", "ii"), "non_robust_lambda_error" := i.lambda_error]
-
-data[, "better_than_non_robust" := lambda_error <= non_robust_lambda_error]
-data[k == 0.0, "better_than_non_robust" := NA_integer_]
-data[weight_method == "non-robust", "better_than_non_robust" := NA_integer_]
-
-pass_rate_data <- data[, list(
-  "pass_rate" = sum(better_than_non_robust, na.rm=TRUE),
-  "n" = sum(!is.na(better_than_non_robust)),
-  "median_error" = stats::median(lambda_error)),
-  by=c("method", "estimation_method", "weight_method")]
-pass_rate_data[, "pass_rate" := round(pass_rate / n * 100.0, 1)][order(pass_rate)]
-
-p <- ggplot2::ggplot(
-  data = data,
-  mapping = ggplot2::aes(
-    x = weight_method,
-    y = lambda_error,
-    fill = estimation_method))
-p <- p + plot_theme
-p <- p + ggplot2::geom_violin(draw_quantiles = 0.5)
-p <- p + ggplot2::facet_wrap("estimation_method", nrow=nlevels(data$estimation_method))
-p <- p + paletteer::scale_fill_paletteer_d(
-  palette = "ggthemes::Tableau_10",
-  drop = FALSE)
-p <- p + guides(fill = ggplot2::guide_legend(title = "optimisation criterion"))
-p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
-p <- p + ggplot2::xlab("weighting method")
-p <- p + ggplot2::ggtitle("shift-sensitive Yeo-Johnson transformation")
-p <- p + ggplot2::theme(
-  axis.text.x = ggplot2::element_text(
-    angle=30.0,
-    hjust=1.0))
-p
-
-```
-# Appendix C: Right-sided outliers
-
-```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-
-.get_optimised_weighting_function_parameters(manuscript_dir = manuscript_dir, side = "right")
-
-if(!file.exists(file.path(manuscript_dir, "robustness_comparison_rightside_outlier.RDS"))){
-  set.seed(95)
-
-  .compute_lambda <- function(x, method){
-    # Create transformer object.
-    transformer <- suppressWarnings(
-      power.transform::find_transformation_parameters(
-        x = x,
-        method = method,
-        robust = FALSE,
-        shift = TRUE))
-
-    return(transformer@lambda)
-  }
-
-  # Generate table of asymmetric generalised normal distribution parameters.
-  n_distributions <- 100
-
-  # Generate alpha, beta and n.
-  n <- stats::runif(n=n_distributions, min=2, max=4)
-  n <- ceiling(10^n)
-
-  alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
-  beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)
-
-  # Generate corresponding distributions.
-  x <- mapply(
-    power.transform::ragn,
-    n = n,
-    alpha = alpha,
-    beta = beta)
-
-  # Compute lambda values without weighting for Box-Cox.
-  target_lambda_bc <- sapply(
-    x,
-    .compute_lambda,
-    method = "box_cox")
-
-  # Compute lambda values without weighting for Yeo-Johnson
-  target_lambda_yj <- sapply(
-    x,
-    .compute_lambda,
-    method = "yeo_johnson")
-
-  # Add outliers, between 0.00 and 0.10.
-  n_outliers <- 10
-  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 20)
-
-  # Find optimised k1 and k2 values for all combinations of weight sources and
-  # weighting functions.
-  x <- mapply(
-    function(x, n, alpha, beta, target_lambda_bc, target_lambda_yj, outlier_fraction){
-
-      parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda_bc"=target_lambda_bc, "target_lambda_yj"=target_lambda_yj)
-
-      x <- lapply(
-        outlier_fraction,
-        function(k, x, parameters){
-          # Set parameters.
-          parameters$k <- k
-
-          # Compute interquartile range.
-          interquartile_range <- stats::IQR(x)
-
-          # Compute upper and lower quartiles.
-          q_lower <- stats::quantile(x, probs=0.25, names=FALSE)
-          q_upper <- stats::quantile(x, probs=0.75, names=FALSE)
-
-          # Set data where the outliers will be copied into.
-          x_outlier <- x
-
-          if(k != 0.0){
-            n_draw <- ceiling(k * length(x))
-
-            # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-            # than Q3 + 1.5 IQR.
-            x_random <- stats::runif(n_draw, min=0.0, max=2.0)
-            outlier <- q_upper + 1.5 * interquartile_range + x_random * interquartile_range
-
-            # Randomly insert outlier values.
-            x_outlier[sample(seq_along(x), size=n_draw, replace=FALSE)] <- outlier
-          }
-
-          return(list(
-            "x" = x_outlier,
-            "parameters" = parameters))
-        },
-        x = x,
-        parameters)
-
-      return(x)
-    },
-    x = x,
-    n = n,
-    alpha = alpha,
-    beta = beta,
-    target_lambda_bc = target_lambda_bc,
-    target_lambda_yj = target_lambda_yj,
-    MoreArgs=list("outlier_fraction"=outlier_fraction),
-    SIMPLIFY=FALSE,
-    USE.NAMES=FALSE)
-
-  x <- unlist(x, recursive = FALSE)
-
-  # Optimise k1, k2 so that sum of absolute differences with target-lambda is
-  # minimised for each combination of source and weighting method.
-  experiment_args <- coro::generator(
-    function(){
-
-      for(method in c("box_cox", "yeo_johnson")){
-        # Non-robust transformation.
-        fun_args <- list(
-          "method" = method,
-          "robust" = FALSE,
-          "shift" = TRUE)
-
-        opt_args <- list()
-
-        yield(list(
-          "name" = "non-robust",
-          "method" = method,
-          "fun_args" = fun_args,
-          "opt_args" = opt_args
-        ))
-      }
-
-      for(robustness_source in c("original", "transformed", "residual")){
-
-        def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))
-
-        if(robustness_source == "original"){
-          def_opt_limits$k1[2] <- 1.0
-          def_opt_limits$k2[2] <- 1.0
-
-          def_opt_init <- list("k1" = 0.80, "k2" = 0.95)
-
-        } else if(robustness_source == "transformed"){
-          def_opt_init <- list("k1" = 1.28, "k2" = 1.96)
-
-        } else if(robustness_source == "residual"){
-          def_opt_init <- list("k1" = 0.50, "k2" = 1.0)
-        }
-
-        for(method in c("box_cox", "yeo_johnson")){
-          for(robustness_weighting_function in c("step", "triangle", "cosine")){
-
-            fun_args <- list(
-              "method" = method,
-              "robust" = TRUE,
-              "shift" = TRUE,
-              "weight_method" = paste0(robustness_source, "_", robustness_weighting_function),
-              "backup_use_default" = FALSE)
-
-            opt_limits <- def_opt_limits
-            opt_init <- def_opt_init
-
-            # Step-weighting only has k1 as a parameter.
-            if(robustness_weighting_function == "step"){
-              opt_limits$k2 <- NULL
-              opt_init$k2 <- NULL
-            }
-
-            yield(list(
-              "name" = paste0(robustness_source, "-", robustness_weighting_function),
-              "method" = method,
-              "fun_args" = fun_args,
-              "opt_args" = list("initial" = opt_init, "limits" = opt_limits)
-            ))
-          }
-        }
-      }
-    })
-
-  # Function for computing transformer parameters under optimisation constraints.
-  compute_robust_lambda <- function(
-    x,
-    fun_args,
-    opt_args,
-    cl_internal=NULL){
-
-    # Custom parser.
-    ..outlier_parser <- function(
-    x,
-    fun_args,
-    opt_args
-    ){
-      # Create transformer.
-      transformer <- suppressWarnings(do.call(
-        power.transform::find_transformation_parameters,
-        args=c(
-          list("x" = x$x),
-          fun_args,
-          opt_args)))
-
-      parameter_data <- c(
-        x$parameters,
-        opt_args,
-        list(
-          "method" = fun_args$method,
-          "weight_method" = fun_args$weight_method,
-          "lambda" = transformer@lambda,
-          "shift" = transformer@shift)
-      )
-
-      return(parameter_data)
-    }
-
-    if(is.null(cl_internal)){
-      parameter_data <- lapply(
-        x,
-        ..outlier_parser,
-        fun_args = fun_args,
-        opt_args = opt_args)
-
-    } else {
-      parameter_data <- parallel::parLapply(
-        cl = cl_internal,
-        X = x,
-        fun = ..outlier_parser,
-        fun_args = fun_args,
-        opt_args = opt_args)
-    }
-
-    return(parameter_data)
-  }
-
-  # Inner optimisation function that sets the loss.
-  ..optimisation_function <- function(
-    opt_param,
-    x,
-    fun_args,
-    cl_internal = NULL){
-
-    # Parse options.
-    opt_args <- list()
-    if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
-    if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))
-
-    parameter_data <- compute_robust_lambda(
-      x = x,
-      fun_args = fun_args,
-      opt_args = opt_args,
-      cl_internal = cl_internal)
-
-    if(fun_args$method == "box_cox"){
-      lambda_error <- sapply(
-        parameter_data,
-        function(x) (abs(x$lambda - x$target_lambda_bc)))
-
-    } else {
-      # Yeo-Johnson.
-      lambda_error <- sapply(
-        parameter_data,
-        function(x) (abs(x$lambda - x$target_lambda_yj)))
-    }
-
-    return(sum(lambda_error))
-  }
-
-  # Outer optimisation function
-  .optimisation_function <- function(
-    experiment,
-    data,
-    cl_internal = NULL){
-
-    set.seed(9)
-
-    # Run optimiser.
-    if(length(experiment$opt_args) > 0){
-
-      x0 <- unlist(experiment$opt_args$initial)
-
-      lower <- experiment$opt_args$limits$k1[1]
-      upper <- experiment$opt_args$limits$k1[2]
-
-      if(length(experiment$opt_args$limits) > 1){
-        lower <- c(lower, experiment$opt_args$limits$k2[1])
-        upper <- c(upper, experiment$opt_args$limits$k2[2])
-      }
-
-      h <- nloptr::sbplx(
-        x0 = x0,
-        fn = .optimisation_inner,
-        lower = lower,
-        upper = upper,
-        control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
-        x = data,
-        fun_args = experiment$fun_args,
-        cl_internal = cl_internal)
-
-      if(length(h$par) == 1){
-        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par, "value" = h$value))
-
-      } else {
-        return(list("name"=experiment$name, "method"=experiment$method, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
-      }
-
-    } else {
-      value <- ..optimisation_function(
-        opt_param=list(),
-        x = data,
-        fun_args=experiment$fun_args,
-        cl = cl_internal)
-
-      return(list("name"=experiment$name, "method"=experiment$method, "value" = value))
-    }
-  }
-
-  # Start cluster
-  cl <- parallel::makeCluster(18L)
-
-  parallel::clusterExport(
-    cl=cl,
-    varlist = c("compute_robust_lambda", "..optimisation_function")
-  )
-
-  experiment_results <- parallel::parLapplyLB(
-    cl = cl,
-    X = coro::collect(experiment_args()),
-    fun = .optimisation_function,
-    data = x,
-    chunk.size = 1L)
-
-  # Stop cluster.
-  parallel::stopCluster(cl)
-
-  saveRDS(experiment_results, file = file.path(manuscript_dir, "robustness_comparison_rightside_outlier.RDS"))
-}
-```
-
-# Appendix D: Generalisability of empirical goodness of fit test
-
-The empirical goodness of fit test was characterised using the empirical distribution of type I errors related to the mean residual error in the central portion ($\kappa = 0.80$) of the transformed feature distribution, in the presence of outliers.
-The corresponding power transformations were both robust and shift sensitive, and used maximum likelihood estimation (MLE) for optimisation of transformation parameters.
-An important concern is how well the test performs for conventional power transformations (in absence of outliers), and for optimisation criteria that are objectively better than MLE, i.e.
-the Cramér-von Mises criterion.
-
-For Figure *insert* determined type 1 error rate as before, but for several variants.
-In addition to robust, shift sensitive transformations using MLE, conventional transformations and robust, shift-sensitive using the Crámer-von Mises criterion were assessed.
-Conventional transformations only used randomly drawn asymmetric generalised normal distributions without added outliers.
-This figure suggests two things: If conventional transformations were used instead of robust, shift sensitive variants, the test characteristics would be highly similar.
-However, the robust, shift sensitive variant shows that the test characteristics are even stable in the presence of outliers.
-Secondly, the optimisation criterion does directly matter, and compared to MLE, the Crámer-von Mises criterion results in transformed distributions that better approximate central normality.
-In this case, one could argue whether characteristics of the empirical goodness of fit test should be based on the Crámer-von Mises criterion instead of MLE.
-We would argue that MLE-based test characteristics form a good foundation, for two reasons.
-One, MLE is the most prevalent optimisation criterion for power transformations, and two, the resulting transformations still approach central normality.
-
-```{r gof_test_type_1_error_rate_appendix, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Type I error rates for power transformation variants. ",
-  "The robust, shift sensitive power transformation using maximum likelihood estimation (MLE) is shown as comparison. ",
-  "Optimisation of transformation parameters using the Crámer-von Mises (C-vM) criterion leads to lower residual errors compared to MLE, ",
-  "particularly for Box-Cox transformations, i.e. they lead to transformations that are closer to central normality. ",
-  "Conventional transformations (in the absence of outliers) yield curves that are comparable to their robust, shift sensitive variants."
-)
-
-.plot_type_1_error_rate_appendix(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir)
-
-```
-
-```{r include = FALSE, eval = FALSE}
-# CANDIDATE FOR REMOVAL !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
-data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
-
-# Convert to data matrix. First aggregate within each experiment because the
-# Wilcoxon signed rank test is not for repeated measurements.
-data <- data.table::dcast(
-  data,
-  dataset ~ learner + transformation_method + normalisation_method, value.var="value_rank",
-  fun.aggregate = mean
-)
-data[, "dataset" := NULL]
-
-# Compute p-values through pair-wise comparison.
-config_pairs <- utils::combn(colnames(data), m = 2L)
-test_data <- apply(
-  config_pairs,
-  MARGIN = 2,
-  function(config_pair, data) {
-    h <- stats::wilcox.test(
-      x = data[[config_pair[1L]]],
-      y = data[[config_pair[2L]]],
-      paired = TRUE,
-      alternative = "two.sided"
-    )
-
-    return(data.table::data.table(
-      "config_1" = config_pair[1L],
-      "config_2" = config_pair[2L],
-      "raw_p_value" = h$p.value
-    ))
-  },
-  data = data
-)
-test_data <- data.table::rbindlist(test_data)
-
-# Perform multiple testing correction.
-test_data[, "p_value_corrected" := stats::p.adjust(raw_p_value)]
-```
-
 # References

### 2024-08-21 manuscript/manuscript.Rmd
Additional changes.

@@
 
-There are a few options based on which weights may be set. Here, we investigate three:
+Weights may be set using several methods. Here, we investigate three:
 
@@
 
-Based on either of the above, weights are set using a weighting function.
+Weights are set using a weighting function for all three methods.
 We will investigate the following functions, where $\dot{x}_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:
@@
 ```
-Here $\alpha \in (0,1)$ is a skewness parameter.
+Here, $\alpha \in (0,1)$ is a skewness parameter.
 $\alpha > 0.5$ creates a distribution with a negative skew, i.e.
@@
 Deviations from normality can be detected by normality tests, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers or a combination of the two.
+In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers, or both.
 Here we develop an empirical test for central normality.
@@
 The $\lambda$ parameter for conventional power transformations (Eqn.
-\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn.
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn.
 \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-Appendix B describes other criteria, and Appendix C the results of simulations using these criteria.
 
@@
 
-To assess whether the proposed power transformations lead to $\lambda$ values that are invariant to location and scale of the distribution, we simulate three different distributions.
+To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulate three different distributions.
 The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.

### 2024-08-23 manuscript/manuscript.Rmd
Changes to manuscript.

@@
 
-# Methods
+# Theory
+
+In this section, we will first introduce location and scale invariant versions of the Box-Cox and Yeo-Johnson transformations.
+Subsequently, we define weighted location and scale invariant transformations and weighting methods that allow for defining robust transformations.
+For the purpose of simulation we then define the quantile function for asymmetric generalised normal distributions.
+Finally, we define the overall framework for the empirical central normality test.
 
@@
 \begin{equation}
+\label{eqn:box-cox-weighted-invariant-log-likelihood}
 \begin{split}
@@
 \begin{equation}
+\label{eqn:yeo-johnson-weighted-invariant-log-likelihood}
 \begin{split}
@@
 
-In this section we simulate data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
+In this section we used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
 The $\lambda$ parameter for conventional power transformations (Eqn.
 \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn.
-\ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
+\ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
 
@@
 
-To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulate three different distributions.
-The first distribution is a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
-We first randomly draw $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
-The second distribution is a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
-The third distribution is a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then compute transformation parameter $\hat{\lambda}$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
-To assess location invariance, a positive value $d_{\text{shift}}$ is added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
-Similarly, to assess scale invariance, each distribution is multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.
+To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulated three different distributions.
+The first distribution was a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
+We first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
+To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
+Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.
 
@@
 The result is shown in Figure \ref{fig:shifted-distributions}.
-For each distribution, transformation parameter $\hat{\lambda}$ varies with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions.
-In contrast, estimation of $\lambda$ using invariant power transformations was invariant to $d_{\text{shift}}$ and $d_{\text{scale}}$. 
-While the effect of location on the estimated value of $\lambda$ using conventional power transformations is noticeable, scale also affects $\lambda$ for features that are not normally distributed.
-Figure \ref{fig:shifted-distributions} also shows that estimated $\lambda$ for Box-Cox power transformations in left-skewed distributions diverges from that found for invariant Box-Cox power transformations.
-
-## Empirical distribution yields consistent robust transformations
-
-Outliers may be presented in the data, and affect estimation of power transformation parameters.
-The log-likelihood function is weighted to assign less weight to outlier instances.
-Weighting is based on either probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
-A weight is then assigned to each instance using one of three weighting functions: step, triangle and tapered cosine.
+For each distribution, transformation parameter $\lambda$ varied with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions.
+The effect of location on the estimated $\lambda$ value for conventional power transformations is noticeable for all three distributions.
+However scale only affected $\lambda$ for features that were sampled from a skewed distribution.
+In contrast, estimation of $\lambda$ using invariant power transformations was invariant to both $d_{\text{shift}}$ and $d_{\text{scale}}$.
+
+## Robust transformations
+
+Outliers may be present in data and affect estimation of power transformation parameters.
+The log-likelihood function can be weighted to assign less weight to outlier instances, see equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}.
+We proposed three weighting methods: probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
+The weight itself is then assigned to each instance using one of three weighting functions: step, triangle and tapered cosine.
 The weighting functions have one (step) or two (triangle, tapered cosine) parameters.
-Here we investigate how these parameters should be set to find transformation parameter $\hat{\lambda}^r$ that are robust to outliers.
-
-Let us randomly draw $m_d=100$ asymmetric generalised normal distributions.
-Each distribution is parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-Location and scale parameters are set as $\mu = 0$ and $\sigma = 1$, respectively.
-$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(2, 4\right)$, i.e., between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
-
-Outlier values are then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
-This is repeated $m_{out} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
-Thus up to 10 percent of the values may be replaced by outliers.
-Outlier values are set according to @Tukey1977-xm.
+Here we investigate how these parameters should be set to find robust transformation parameter $\lambda_r$.
+
+We randomly drew $m_d=100$ asymmetric generalised normal distributions.
+Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
+$n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, with $\gamma \sim U\left(2, 4\right)$, i.e., between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
+
+Outlier values were then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
+This was repeated $m_{out} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
+Thus up to 10 percent of the values could be replaced by outliers.
+Outlier values were set according to @Tukey1977-xm, as follows.
 Let $x^{*} \sim U\left(-2, 2\right)$.
-Then the corresponding outlier value is:
+Then the corresponding outlier value was:
 
@@
 
-Here, $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
-Outlier values randomly replace values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$.
-
-To find the optimal values for the weighting function parameters $k_1$ and $k_2$ (if applicable), we minimise the absolute difference between the $\hat{\lambda}^r$ parameter obtained for transformation in the presence of outliers, and the $\hat{\lambda}_0$ parameter obtained using the non-robust modified transformation in absence of outliers:
-
-```{=latex}
-\begin{equation}
-\left\{ \hat{k}_1, \hat{k}_2 \right\} = \argmin_{k_1, k_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \hat{\lambda}^r \left(\mathbf{X}_{i, j}; k_1, k_2 \right) - \hat{\lambda}_{0} \left(\mathbf{X}_i \right) \right|
-\end{equation}
-```
-The double sum in right-hand side of the above equation can be considered as the loss that should be minimised.
-As mentioned earlier, transformation parameters are estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-
-The optimal weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust, location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
+$Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
+Outlier values randomly replaced values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$.
+
+To find the optimal values for the weighting function parameters $k_1$ and $k_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
+
+```{=latex}
+\begin{equation}
+\label{eqn:minimisation-weighting}
+\left\{ \hat{k}_1, \hat{k}_2 \right\} = \argmin_{k_1, k_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; k_1, k_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
+\end{equation}
+```
+
+Minimisation was conducted using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb].
+The resulting weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location and scale invariant Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 
-Figure \ref{fig:optimised-weighting-function-parameters} shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters.
-In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation.
-For both transformations, weighting based on empirical probabilities yields the most consistent $\lambda$ parameter estimates in the presence of outliers.
-Other methods did not outperform the non-robust power transformation method.
+Figure \ref{fig:optimised-weighting-function-parameters} shows the distribution of errors $\left| \lambda_r - \lambda_0 \right|$ for non-robust and robust transformations using the optimal weighting function parameters.
+In the presence of outliers, the Yeo-Johnson transformation yielded smaller errors than the Box-Cox transformation.
+For both transformations, weighting based on empirical probabilities yielded the most consistent $\lambda$ parameter estimates in the presence of outliers.
+Other methods did not outperform the non-robust transformation method.
 

### 2024-08-27 manuscript/manuscript.Rmd
Fixes to plots.

@@
 # Base font size.
-base_size <- 10
+base_size <- 9
 
@@
 plot_theme <- ggplot2::theme_light(base_size=base_size)
-plot_theme$plot.title$size <- ggplot2::rel(0.8)
+plot_theme$plot.title$size <- ggplot2::rel(1.0)
 plot_theme$plot.title$hjust <- 0.5
@@
 plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
+plot_theme$legend.title
 
@@
 Subsequently, we define weighted location and scale invariant transformations and weighting methods that allow for defining robust transformations.
-For the purpose of simulation we then define the quantile function for asymmetric generalised normal distributions.
+We then define the quantile function for asymmetric generalised normal distributions to enable random sampling.
 Finally, we define the overall framework for the empirical central normality test.
@@
 
-```{r marginal-effect-plot, echo=FALSE, fig.cap=cap, fig.height=12.0, warning=FALSE}
+```{r marginal-effect-plot, echo=FALSE, fig.cap=cap, fig.height=7.5, warning=FALSE}
 cap <- paste0(

### 2024-08-27 manuscript/manuscript.Rmd
Updates to text.

@@
 title: "Location and Scale-Invariant Power Transformations for Transforming Data to Normality"
-author: "Alex Zwanenburg"
+author: "Alex Zwanenburg, Steffen Löck"
 date: "`r Sys.Date()`"
@@
 
-Numerical variables in datasets may strongly deviate from normal distributions, e.g.
+Many statistical and several machine learning methods assume normality of the underlying data.
+However, numerical features in datasets may strongly deviate from normal distributions, e.g.
 by being skewed.
-This may complicate further analysis.
-Power transformations [@Tukey1957-rt] can help improve normality of such features.
+Power transformations aim to stabilise variance and improve normality of such features [@Bartlett1947-rx, @Tukey1957-rt].
 The two most commonly used transformations are that of @Box1964-mz and @Yeo2000-vw.
@@
 The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
-As noted by Raymaekers and Rousseeuw, this approach is very sensitive to the presence of outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].
+As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].
 
 Power transformation does not guarantee that transformed variables are normally distributed.
-In fact, depending on location and scale of the variable, power transformation may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
-This is problematic for automated power transformations, for three reasons.
-First, normality is often desired in machine learning applications.
+Depending on location and scale of a feature and the presence of outliers, power transformation may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
+This is problematic when normality of the transformed feature is not checked, e.g. in automated power transformation in machine learning workflows.
+Several issues may arise.
+First, some machine learning methods assume normality.
 Second, a large negative or positive $\lambda$-parameter may lead to numeric issues.
-Third, we currently do not have a test to automatically reject poor transformations.
+Third, we currently do not have a test to automatically reject poor transformations to normality.
 Statistical tests for normality exist, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-However, given sufficiently large sample sizes, these tests can detect trivial deviations from normality and are therefore not well-suited for applications of automated power transformations in practice.
-
-```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=4.0, warning=FALSE}
+However, given sufficiently large sample sizes such tests can detect trivial deviations from normality.
+
+```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
 cap <- paste0(
@@
 -   We devise location- and scale-invariant versions of the Box-Cox and Yeo-Johnson transformation, including versions robust to outliers.
+
+-   We derive the maximum likelihood criterion for location- and scale-invariant Box-Cox and Yeo-Johnson transformations to allow for optimising transformation parameters.
 

### 2024-08-28 manuscript/manuscript.Rmd
Updates to manuscript.

@@
     latex_engine: lualatex
+papersize: a4
 bibliography: refs.bib
@@
 
-Power transformation does not guarantee that transformed variables are normally distributed.
+Power transformation does not guarantee that transformed features are normally distributed.
 Depending on location and scale of a feature and the presence of outliers, power transformation may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
-This is problematic when normality of the transformed feature is not checked, e.g. in automated power transformation in machine learning workflows.
-Several issues may arise.
-First, some machine learning methods assume normality.
-Second, a large negative or positive $\lambda$-parameter may lead to numeric issues.
-Third, we currently do not have a test to automatically reject poor transformations to normality.
-Statistical tests for normality exist, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-However, given sufficiently large sample sizes such tests can detect trivial deviations from normality.
-
-```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
-cap <- paste0(
-  "Power transformations may reduce normality depending on location. ",
-  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution. ",
-  "For the left-hand panels, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
-  "whereas for the right-hand row, the distribution is $\\mathcal{N}(0, \\sigma)$. " ,
-  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
-  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
-  "and breaks down due to numerical issues for $\\mu > 5 \\cdot 10^3$. ",
-  "Here, scale did not affect Box-Cox and Yeo-Johnson transformations of variables drawn from normal distributions."
+If normality of the transformed feature is not checked, e.g. in automated power transformation in machine learning workflows, several issues may arise.
+For example, statistical tests such as ANOVA may produce incorrect results due to violation of the normality assumption.
+Likewise, machine learning methods that assume normality of input features may suffer a decrease in performance.
+Also, large negative or positive $\lambda$-parameters may lead to numeric issues in any subsequent computations.
+
+Statistical tests for normality, such as the Shapiro-Wilk test [@Shapiro1965-zd], could be automatically applied to transformed features.
+However, given sufficiently large sample sizes such tests can detect trivial deviations from normality, and may lead to rejection of sufficiently good power transformations.
+
+```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=2.5, warning=FALSE}
+cap <- paste0(
+  "Effect of location shift, scale and outliers on estimation of Box-Cox and Yeo-Johnson transformation parameter $\\lambda$. ",
+  "$10000$ samples were drawn from a normal distribution: ",
+  "$\\mathcal{N}(\\mu, 1)$ for the shift dataset, $\\mathcal{N}(10, \\sigma)$ for the scale dataset and $\\mathcal{N}(0, 1)$ for the outlier dataset. ",
+  "Additionally, an outlier with value $d$ was added to the outlier dataset. ",
+  "Since samples are drawn from a normal distribution, a transformation parameter of $\\lambda = 1$ is expected. ",
+  "However, a large shift in location, a scale that is small compared to the location, or presence of large outliers ",
+  "lead to incorrectly estimated transformation parameter values. "
 )
@@
 
-Here we make the following contributions:
+To address these issues, we make the following contributions:
 
@@
 
-In this section, we will first introduce location and scale invariant versions of the Box-Cox and Yeo-Johnson transformations.
-Subsequently, we define weighted location and scale invariant transformations and weighting methods that allow for defining robust transformations.
+In this section, we will first introduce location- and scale-invariant versions of the Box-Cox and Yeo-Johnson transformations.
+Subsequently, we define weighted location- and scale-invariant transformations and weighting methods for robust transformations.
 We then define the quantile function for asymmetric generalised normal distributions to enable random sampling.
@@
 
-For both invariant transformations, $\lambda$, $x_0$ and $s$ parameters are obtained by maximising the log-likelihood function.
-A full derivation is shown in Appendix A.
-The invariant Box-Cox log-likelihood function is:
+For both invariant transformations, $\lambda$, $x_0$ and $s$ parameters can be obtained by maximising the log-likelihood function, i.e. using maximum likelihood estimation (MLE).
+A full derivation of the log-likelihood function for both transformations is shown in Appendix A.
+The location- and scale-invariant Box-Cox log-likelihood function is:
 
@@
 where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
-Similarly, the invariant Yeo-Johnson log-likelihood function is:
+Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood function is:
 
@@
 
-```{r sensitivity-to-outliers, echo=FALSE, fig.cap=cap, fig.height=2.0, warning=FALSE}
-cap <- paste0(
-  "Power transformations are affected by outliers. ",
-  "Variable $x$ consists of $1000$ values randomly drawn from a normal distribution $\\mathcal{N}(0, 1)$, ",
-  "with a single outlier with value $d$. ",
-  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
-  "The figure shows that instead, $\\lambda$ is strongly affected by the presence of the outlier at $d$.")
-
-.plot_sensitivity_to_outlier(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
-
-```
-
 Real-world data may contain outliers, to which maximum likelihood estimation can be sensitive.
-Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:sensitivity-to-outliers}.
-As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e.
-achieve *central normality*.
+Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:decreased-normality}.
+As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*.
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
-Because this procedure iteratively updates $\lambda$, it can not be used to simultaneously estimate $\lambda$, $x_0$ and $s$ for invariant power transformations.
-
-In essence, obtaining robust transformation parameters that provide central normality requires identifying outliers in the data and weighting such instances during the optimisation process.
+Briefly, this process requires identifying outliers in the data and weighting such instances during the optimisation process.
 @Raymaekers2021-kq achieve this through weighted maximum likelihood estimation.
+However, because this procedure iteratively estimates and updates $\lambda$, it can not be used here to simultaneously estimate $\lambda$, $x_0$ and $s$ for location- and scale-invariant power transformations.
+Nonetheless, as a procedure, weighted MLE can be used for estimating the transformation, shift and scale parameters.
+
 Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
 Compared to @Raymaekers2021-kq, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
-The weighted invariant Box-Cox log-likelihood function is:
+The weighted location- and scale-invariant Box-Cox log-likelihood function is:
 
@@
 
-Analogously, the weighted invariant Yeo-Johnson log-likelihood function is:
+Analogously, the weighted location- and scale-invariant Yeo-Johnson log-likelihood function is:
 
@@
 
-Other criteria aside from MLE can be used for finding robust transformation parameters, see Appendix B. For the remainder, we will use MLE as the optimisation criterion.
-
-Weights may be set using several methods. Here, we investigate three:
-
-- Set weights based on probabilities of the empirical distribution of the original feature $\mathbf{X}$.
-Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $\mathbf{X}$.
-
-- Set weights based on the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
-After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
-Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
-
-- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
-
-Weights are set using a weighting function for all three methods.
-We will investigate the following functions, where $\dot{x}_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:
+The weights $w_i$ in equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood} can be set using several weighting functions.
+Here, we investigate three:
 
@@
 The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.
-Examples of the three types of weighting function are shown in Figure \ref{fig:weighting-functions}.
+Examples of these weighting function are shown in Figure \ref{fig:weighting-functions}.
 
@@
 ```
+
+Each weighting function has an argument $\dot{x}$ that is related to the (transformed) feature in one of several ways:
+
+- The weighting function uses empirical probabilities of the distribution of the original feature $\mathbf{X}$.
+  Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $\mathbf{X}$.
+  Then $\dot{x}_i = p^{*}_i=2 \left( p_i - 0.5\right)$, so that argument is zero-centered.
+
+- The weighting function uses the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
+  After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
+  Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
+  Then $\dot{x}_i = z_i$.
+
+- The weighting function uses the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+  Then $\dot{x}_i = r_i$.
 
@@
 The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}(p)$ thus incorporates the quantile function $F_{\beta}^{-1}(p^{*})$ of the symmetric generalised normal distribution.
-$F_{\beta}^{-1}(p^{*})$ was derived by Griffin to be [@Griffin2018-bf]:
-
-```{=latex}
-\begin{equation}
-F_{\beta}^{-1}(p^{*}) = \sgn\left(p^{*} - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p^{*} - 0.5 \right|; 1 / \beta \right)
+$F_{\beta}^{-1}(p)$ was derived by Griffin to be [@Griffin2018-bf]:
+
+```{=latex}
+\begin{equation}
+F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
 \end{equation}
@@
 
-A common variable in health-related datasets is age.
+A common feature in health-related datasets is age.
 Here we use data on 228 patients with lung cancer that was collected and published by Loprinzi et al. [@Loprinzi1994-cd].
@@
 Location- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
-Moreover, conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age variable after transformation.
+Moreover, conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
 The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.78$).
@@
 Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
-Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass variable after transformation.
+Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
 The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
@@
 This is particularly relevant for automated data processing, where such issues may go unnoticed.
-However, we did not find a measurable benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric variable.
+However, we did not find a measurable benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
-Of the 4886 numeric variables in the 232 datasets, 266 (5%) variables in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
+Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
 
@@
 
-Robust location- and scale-invariant power transformations allow for automated power transformation of numerical variables/
+Robust location- and scale-invariant power transformations allow for automated power transformation of numerical features.
 

### 2024-08-28 manuscript/plots.R

@@ -78,7 +78,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   require(ggplot2)
 
   # Prevent warnings due to non-standard evaluation.
-  estimation_method <- distribution <- method <- version <- NULL
+  data_type <- method <- NULL
 
   # Process / read data.
   data <- .get_data_problematic_transformations(
@@ -100,7 +100,18 @@ get_annotation_settings <- function(ggtheme = NULL) {
     linetype = "longdash",
     colour = "grey40"
   )
-  p <- p + ggplot2::annotate(
+  p <- p + paletteer::scale_color_paletteer_d(
+    palette = "ggthemes::Tableau_10",
+    name = "transformation",
+    drop = FALSE
+  )
+
+  # Branch into shift and scale-specific plots
+  p_shift <- p + ggplot2::scale_x_continuous(
+    name = latex2exp::TeX("$\\mu$"),
+    labels = scales::math_format()
+  )
+  p_shift <- p_shift + ggplot2::annotate(
     geom = "text",
     x = 6,
     y = 1,
@@ -112,30 +123,50 @@ get_annotation_settings <- function(ggtheme = NULL) {
     vjust = -1.0,
     hjust = 1.0
   )
-  p <- p + paletteer::scale_color_paletteer_d(
-    palette = "ggthemes::Tableau_10",
-    name = "method",
-    drop = FALSE
-  )
-  p <- p + ggplot2::scale_y_continuous(
-    name = latex2exp::TeX("$\\lambda$"),
-    limits = c(0.0, 6.0)
-  )
 
-  # Branch into shift and scale-specific plots
-  p_shift <- p + ggplot2::scale_x_continuous(
-    name = latex2exp::TeX("$\\mu$"),
-    labels = scales::math_format()
-  )
   p_scale <- p + ggplot2::scale_x_continuous(
     name = latex2exp::TeX("$\\sigma$"),
     labels = scales::math_format()
   )
+  p_scale <- p_scale + ggplot2::annotate(
+    geom = "text",
+    x = -6,
+    y = 1,
+    label = "expected",
+    colour = annotation_settings$colour,
+    family = annotation_settings$family,
+    fontface = annotation_settings$face,
+    size = annotation_settings$geom_text_size,
+    vjust = -1.0,
+    hjust = 0.0
+  )
+
[…]

### 2024-08-28 manuscript/simulations.R

@@ -3,63 +3,135 @@
   set.seed(19L)
 
   file_name <- file.path(manuscript_dir, "problematic_transformations_plot.RDS")
-  estimation_methods <- "mle"
 
-  x <- power.transform::ragn(10000L, location = 0, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
+  if (!file.exists(file_name)) {
+    x <- power.transform::ragn(
+      10000L,
+      location = 0.0,
+      scale = 1.0 / sqrt(2.0),
+      alpha = 0.5,
+      beta = 2.0
+    )
 
-  # generator ----------------------------------------------------------------
-  generate_experiment_data <- coro::generator(
-    function(x) {
-      shift_range <- 10^seq(from = 0, to = 6, by = 0.1)
-      scale_range <- 10^seq(from = -6, to = 0, by = 0.1)
-      outlier_range <- 10^seq(from = -1, to = 6, by = 0.1)
+    # generator ----------------------------------------------------------------
+    generate_experiment_data <- coro::generator(
+      function(x) {
+        shift_range <- 10^seq(from = 0, to = 6, by = 0.1)
+        scale_range <- 10^seq(from = -6, to = 0, by = 0.1)
+        outlier_range <- 10^seq(from = -1, to = 6, by = 0.1)
 
-      for (method in c("box_cox", "yeo_johnson")) {
-        # Iterate over shift range.
-        for (d in shift_range) {
-          coro::yield(list(
-            "x" = x + d,
-            "d" = d,
-            "s" = NA_real_,
-            "distribution" = distribution,
-            "method" = method,
-            "invariant" = invariant,
-            "robust" = robust,
-            "version" = version,
-            "estimation_method" = estimation_method,
-            "data_type" = "shifted"
-          ))
+        for (method in c("box_cox", "yeo_johnson")) {
+          # Iterate over shift range.
+          for (d in shift_range) {
+            coro::yield(list(
+              "x" = x + d,
+              "y" = d,
+              "method" = method,
+              "invariant" = FALSE,
+              "robust" = FALSE,
+              "estimation_method" = "mle",
+              "data_type" = "shift"
+            ))
+          }
+
+          # Iterate over scale range.
+          for (s in scale_range) {
+            coro::yield(list(
+              "x" = x * s + 10.0,
+              "y" = s,
+              "method" = method,
+              "invariant" = FALSE,
+              "robust" = FALSE,
+              "estimation_method" = "mle",
[…]

### 2024-08-29 manuscript/manuscript.Rmd
Updated code availability.

@@
 
-Location- and scale-invariant power transformations are implemented in the `power.transform` package.
-This manuscript was created using R Markdown, and is available from [https://github.com/alexzwanenburg/power.transform].
+Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository.
+The manuscript was created using R Markdown and is also available from the power.transform GitHub repository.
 

### 2024-08-29 manuscript/manuscript.Rmd
Replaced k_1 and k_2 by delta_1 and delta_2.

@@
 
-- A step function, with $k_1 \geq 0$ as threshold parameter:
+- A step function, with $\delta_1 \geq 0$ as threshold parameter:
 ```{=latex}
@@
 \begin{cases}
-1 & \text{if } \left| \dot{x}_i \right| \leq k_1\\
-0 & \text{if } \left| \dot{x}_i \right| > k_1
+1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_1
 \end{cases}
@@
 
-- A triangle function (or generalised Huber weight), with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
+- A triangle function (or generalised Huber weight), with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
 ```{=latex}
@@
 \begin{cases}
-1 & \text{if } \left| \dot{x}_i \right| < k_1\\
-1 - \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
-0 & \text{if } \left| \dot{x}_i \right| > k_2
+1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
+1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_2
 \end{cases}
@@
 
-- A tapered cosine function [@Tukey1967-eb], with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
+- A tapered cosine function [@Tukey1967-eb], with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
 ```{=latex}
@@
 \begin{cases}
-1 & \text{if } \left| \dot{x}_i \right| < k_1\\
-0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
-0 & \text{if } \left| \dot{x}_i \right| > k_2
+1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
+0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_2
 \end{cases}
@@
 
-All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< k_1$, instances are fully weighted, i.e.
-when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top.
-The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.
+All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< \delta_1$, instances are fully weighted, i.e.
+when $\delta_1 > 0$ the weighting functions are symmetric window functions with a flat top.
+The triangle and tapered cosine functions then gradually down-weight instances with $\delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2$, and assign no weight to instances $\left| x_i \right| > \delta_2$.
 Examples of these weighting function are shown in Figure \ref{fig:weighting-functions}.
 
-```{r weighting-functions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+```{r weighting-functions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
 cap <- paste0(
   "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
-  "The step function was parameterised with $k_1 = 0.60$. ",
-  "The triangle and tapered cosine functions were both parameterised with $k_1 = 0.30$ and $k_2 = 0.90$.")
+  "The step function was parameterised with $\\delta_1 = 0.60$. ",
+  "The triangle and tapered cosine functions were both parameterised with $\\delta_1 = 0.30$ and $\\delta_2 = 0.90$.")
 
@@
 
-To find the optimal values for the weighting function parameters $k_1$ and $k_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
+To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
 
@@
 \label{eqn:minimisation-weighting}
-\left\{ \hat{k}_1, \hat{k}_2 \right\} = \argmin_{k_1, k_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; k_1, k_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
+\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
 \end{equation}
@@
 \toprule
-method & \multicolumn{3}{c}{$k_1$} & \multicolumn{3}{c}{$k_2$} & loss \\
+method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
 & initial & limits & optimal & initial & limits & optimal & \\
@@
 
-# Code availability
-
-Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository.
+# Data and code availability
+
+Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE].
 The manuscript was created using R Markdown and is also available from the power.transform GitHub repository.
+Data and results for the machine learning experiment are separately available from Zenodo [INSERT LINK HERE].
 

### 2024-08-29 manuscript/manuscript.Rmd
Further updates to figures and text.

@@
   eval = TRUE,
-  fig.align= "center")
+  fig.align = "center",
+  dev = "cairo_pdf"
+  )
 
@@
 
-## Location and scale invariant power transformation
+## Location- and scale-invariant power transformation
 
@@
 
-## Robust location and scale invariant power transformations
+## Robust location- and scale-invariant power transformations
 
@@
 
-In this section we used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
+We used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
 The $\lambda$ parameter for conventional power transformations (Eqn.
-\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location and scale invariant power transformations (Eqn.
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location- and scale-invariant power transformations (Eqn.
 \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
@@
 The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location and scale invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
+We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
 To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
@@
 
-```{r shifted-distributions, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r shifted-distributions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.5}
 cap <- paste0(
@@
 The result is shown in Figure \ref{fig:shifted-distributions}.
-For each distribution, transformation parameter $\lambda$ varied with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions.
-The effect of location on the estimated $\lambda$ value for conventional power transformations is noticeable for all three distributions.
-However scale only affected $\lambda$ for features that were sampled from a skewed distribution.
-In contrast, estimation of $\lambda$ using invariant power transformations was invariant to both $d_{\text{shift}}$ and $d_{\text{scale}}$.
+For each distribution, transformation parameter $\lambda$ varied with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated for conventional transformations.
+In contrast, estimation of $\lambda$ for invariant power transformations was invariant to both $d_{\text{shift}}$ and $d_{\text{scale}}$.
 
@@
 
-Outliers may be present in data and affect estimation of power transformation parameters.
+Outliers may be present in data and affect estimation of transformation parameters.
 The log-likelihood function can be weighted to assign less weight to outlier instances, see equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}.
-We proposed three weighting methods: probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
-The weight itself is then assigned to each instance using one of three weighting functions: step, triangle and tapered cosine.
-The weighting functions have one (step) or two (triangle, tapered cosine) parameters.
-Here we investigate how these parameters should be set to find robust transformation parameter $\lambda_r$.
-
-We randomly drew $m_d=100$ asymmetric generalised normal distributions.
+We proposed three weighting function: step, triangle and tapered cosine, that have one, two and two parameters, respectively.
+Each weighting function then takes one of three values as input: probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
+
+To determine the weighting function parameters for each of the nine combinations, $m_d=100$ asymmetric generalised normal distributions were drawn.
 Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
@@
 
-To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
+To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for robust transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
 
@@
 Minimisation was conducted using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb].
-The resulting weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location and scale invariant Box-Cox and Yeo-Johnson transformations, respectively.
+The resulting weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss for \textbf{invariant Yeo-Johnson} power transformations.
+\caption{Optimal weighting parameters and corresponding loss for \textbf{location- and scale-invariant Yeo-Johnson} power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
 
-```{r optimised-weighting-function-parameters, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r optimised-weighting-function-parameters, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.5}
 cap <- paste0(
   "Robustness of power transformations after optimising weighting function parameters. ",
-  "The distribution of errors, i.e. the difference between robustly fitted $\\hat{\\lambda}^r$ ",
-  "and expected $\\hat{\\lambda}_{0}$ found in the absence of outliers, is shown for 1000 randomly ",
+  "The distribution of errors, i.e. the difference between robustly fitted $\\lambda^r$ ",
+  "and expected $\\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly ",
   "parametrised asymmetric generalised normal distributions with up to 10% outliers on either side ",
@@
 
-To develop an empirical test for central normality we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{gof}}$.
+To develop an empirical test for central normality we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{ecn}}$.
 We will first define the central portion $\kappa$.
@@
 As expected, the largest deviations from normality appear on the extremities of this range.
-For Box-Cox, the effect of outliers on the lower end of the distribution leads to noticeable asymmetry.
+For Box-Cox transformations, the effect of outliers on the lower end of the distribution leads to noticeable asymmetry.
 Between very low and very high empirical probabilities, residual errors are constrained and relatively flat.
@@
 
-```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
 cap <- paste0(
@@
 ```
-We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic.
+We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic $\tau_{\text{ecn}}$.
 Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}.
@@
 
-```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
 cap <- paste0(
   "Type 1 error rate of transformed asymmetric generalised normal distributions as function of ",
-  "the test statistic for the central portion $\\kappa$ of the distribution.")
+  "the test statistic $\\tau_{\\text{ecn}}$ for the central portion $\\kappa$ of the distribution.")
 
@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
-location- and scale invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
+location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
 and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).
@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
-location- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
 and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).
@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$);
-location- and scale invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
+location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
 and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$).

### 2024-08-29 manuscript/plots.R

@@ -929,7 +929,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p <- p + ggplot2::facet_wrap("method", nrow = 2)
   p <- p + paletteer::scale_fill_paletteer_d(
     palette = "ggthemes::Tableau_10",
-    drop = FALSE
+    drop = FALSE,
+    name = "transformation"
   )
   p <- p + geom_text(
     data = pass_rate_data,
@@ -944,7 +945,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     size = annotation_settings$geom_text_size,
     vjust = 0.5
   )
-  p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\hat{\\lambda}^r - \\hat{\\lambda}_{0}|$"))
+  p <- p + ggplot2::scale_y_sqrt(name = latex2exp::TeX("$| \\lambda_r - \\lambda_0 |$"))
   p <- p + ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))
   p <- p + ggplot2::xlab("weighting method")
 
@@ -1013,6 +1014,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_bc <- p_bc + plot_theme
+  p_bc <- p_bc + ggplot2::theme(legend.key.size = grid::unit(0.3, 'lines'))
   p_bc <- p_bc + ggplot2::geom_line()
   p_bc <- p_bc + ggplot2::geom_line(
     data = outlier_free_data[method == "Box-Cox"],
@@ -1024,7 +1026,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     linetype = "dashed"
   )
   p_bc <- p_bc + ggplot2::scale_colour_discrete(
-    name = "percentile (Box-Cox)",
+    name = "percentile\n(Box-Cox)",
     type = c("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E")
   )
   p_bc <- p_bc + ggplot2::xlab("empirical probability")
@@ -1043,6 +1045,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_yj <- p_yj + plot_theme
[…]

### 2024-08-30 manuscript/manuscript.Rmd
Main fixes to manuscript done.

@@
 plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
-plot_theme$legend.title
+plot_theme$legend.key.size <- grid::unit(0.3, 'lines')
 
@@
 # Abstract
+
+To-do
 
@@
 cap <- paste0(
-  "Effect of location shift, scale and outliers on estimation of Box-Cox and Yeo-Johnson transformation parameter $\\lambda$. ",
+  "Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\\lambda$. ",
   "$10000$ samples were drawn from a normal distribution: ",
@@
   "However, a large shift in location, a scale that is small compared to the location, or presence of large outliers ",
-  "lead to incorrectly estimated transformation parameter values. "
+  "lead to incorrectly estimated transformation parameter values."
 )
@@
   "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
-  "The step function was parameterised with $\\delta_1 = 0.60$. ",
-  "The triangle and tapered cosine functions were both parameterised with $\\delta_1 = 0.30$ and $\\delta_2 = 0.90$.")
+  "In this example, the step function was parameterised with $\\delta_1 = 0.60$. ",
+  "The triangle and tapered cosine functions were both parameterised with $\\delta_1 = 0.30$ and $\\delta_2 = 0.90$."
+)
 
@@
 Deviations from normality can be detected by normality tests, such as the Shapiro-Wilk test [@Shapiro1965-zd].
-In practice, normality tests may be too stringent to assess goodness of fit in presence of large sample sizes, outliers, or both.
+In practice, normality tests may be too stringent with large sample sizes, outliers, or both.
 Here we develop an empirical test for central normality.
@@
 cap <- paste0(
-  "Invariant power transformation produces transformation parameters that are invariant to location and shape. ",
+  "Invariant power transformation produces transformation parameters that are invariant to location and scale. ",
   "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\\text{shift}}$ or multiplication by $d_{\\text{scale}}$. ",
-  "Estimates of transformation parameter $\\lambda$ with the conventional power transformations (circle) show strong dependency on the overall location and scale of the distribution, ",
-  "whereas estimates with the invariant power transformations (triangle) are constant."
+  "Estimates of the transformation parameter $\\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution, ",
+  "whereas estimates obtained for the location- and scale-invariant power transformations are constant."
 )
@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss for \textbf{location- and scale-invariant Box-Cox} power transformations.
+\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
   "parametrised asymmetric generalised normal distributions with up to 10% outliers on either side ",
-  "of the distribution. The top and bottom panel show errors for the invariant Box-Cox ",
-  "and invariant Yeo-Johnson transformations, respectively. ",
+  "of the distribution. The top and bottom panel show errors for the location- and scale-invariant Box-Cox ",
+  "and Yeo-Johnson transformations, respectively. ",
   "In each panel, the error distribution for the non-robust transformation is shown to the left ",
   "for comparison with different weighting methods. ",
-  "Note that a square root transformation was used along the $y$-axis. Median errors are shown above each distribution."
+  "Note that a square root transformation was applied along the $y$-axis for display purposes. ",
+  "Median errors are shown above each distribution."
 )
@@
 
-```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.0}
 cap <- paste0(
@@
   "Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric ",
-  "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
-  "(up to 10% of samples) each. Larger errors occur for more extreme values at the edges of each distribution. ",
-  "For comparison, the 99th percentile of residual errors for transformations of outlier-free distributions are shown (grey dashed line).")
+  "generalised normal distributions, for each of which outliers were randomly drawn 10 times ",
+  "(up to 10% of samples). Larger errors occur at the edges of each distribution."
+)
 
@@
   plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir)
+  manuscript_dir = manuscript_dir
+)
 
@@
 
-```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.0}
 cap <- paste0(
@@
 ```{r include = FALSE, eval = FALSE}
-# This writes the lookup table for the goodness of fit test to the
+# This writes the lookup table for the empirical central normality test to the
 # power.transform package.
@@
 \begin{center}
-\caption{Test statistic for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
+\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
 \label{tab:empirical-central-normality}
@@
 
-```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Bi-modal distributions and empirical goodness of fit test results. ",
+```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Bi-modal distributions and empirical central normality test results. ",
   "The feature (black) is a mixture of two identical sample sets (blue and orange) ",
   "drawn from normal distributions that are offset by a distance $d$. ",
-  "We use the empirical goodness of fit test to compute p-values for the hypothesis that ",
+  "We use the empirical centrally normality test to compute the probability for the hypothesis that ",
   "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
@@
 
-Invariant power transformations are intended to yield improved transformations to normality
-in the presence of large shifts in location, scales far from unity, or both.
-Above, this characteristic was assessed using simulations. Below, this characteristic
-will be evaluated using examples from real datasets.
+Location- and scale-invariant power transformations are intended to yield improved transformations to normality in the presence of large shifts in location, distributions that due to location and scale are not centered near zero, or both.
+Earlier, we assessed these transformations using simulated data.
+In the following, they are evaluated using examples from real datasets.
+We focus on the Yeo-Johnson transformations because
 
@@
 
-```{r lung-cancer-age, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r lung-cancer-age, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
 cap <- paste0(
   "Quantile-quantile plots for age of patients with lung cancer, expressed in years. ",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust invariant transformations."
+  "The left panel shows the quantile-quantile plot for the original feature ",
+  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
+  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
+  "The right panel shows the quantile-quantile plots for the feature transformed using ",
+  "the non-robust and robust location- and-scale invariant transformations."
 )
@@
 
-```{r penguin-body-mass, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r penguin-body-mass, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
 cap <- paste0(
   "Quantile-quantile plots for penguin body mass, measured in grams. ",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust invariant transformations."
+  "The left panel shows the quantile-quantile plot for the original feature ",
+  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
+  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
+  "The right panel shows the quantile-quantile plots for the feature transformed using ",
+  "the non-robust and robust location- and-scale invariant transformations."
 )
@@
 
-```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
 cap <- paste0(
   "Quantile-quantile plots for latitude coordinate in a housing dataset. ",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust invariant transformations."
+  "The left panel shows the quantile-quantile plot for the original feature ",
+  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
+  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
+  "The right panel shows the quantile-quantile plots for the feature transformed using ",
+  "the non-robust and robust location- and-scale invariant transformations."
 )
@@
 The non-robust location- and scale-invariant transformation did not improve over conventional alternatives and yielded a data distribution lacking central normality (empirical central normality test: $p=0.05$).
-TranConventional transformations had high values for the $\lambda$ parameter.
+However, conventional transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
 
@@
 
-```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap}
+```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
 cap <- paste0(
   "Quantile-quantile plots for fuel consumption data in the Top Gear dataset. ",
-  "The dataset contains three strong outliers for highly fuel-efficient vehicles.",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The empirical central normality test indicates that the original data do not deviate significantly from central normality ($p=0.76$).",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
+  "The feature contains three strong outliers related to highly fuel-efficient vehicles.",
+  "The left panel shows the quantile-quantile plot for the original feature. ",
+  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.76$).",
+  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
+  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
   "For these, the empirical central normality test yields $p=0.01$, and $p=0.55$, respectively. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
-  "the non-robust and robust invariant transformations. ",
+  "The right panel shows the quantile-quantile plots for the feature transformed using ",
+  "the non-robust and robust location- and scale-invariant transformations. ",
   "For these, the empirical central normality test yields $p=0.03$, and $p=0.86$, respectively. ",
@@
 
-```{r ischemic-stroke, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Quantile-quantile plots for the maximum arterial wall thickness in the ischemic stroke dataset [@Romano2022-gq]. ",
-  "This data contains two strong outliers. ",
-  "The left panel shows the quantile-quantile plot for the original data. ",
-  "The empirical central normality test indicates that the original data do not deviate significantly from central normality ($p=0.56$).",
-  "The middle panel shows the quantile-quantile plots for data transformed using ",
-  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2021-kq]. ",
+```{r ischemic-stroke, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
+cap <- paste0(
+  "Quantile-quantile plots for the maximum arterial wall thickness in patients with ischemic stroke. ",
+  "This feature contains two strong outliers. ",
+  "The left panel shows the quantile-quantile plot for the original feature. ",
+  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.56$). ",
+  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
+  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
   "For these, the empirical central normality test yields $p=0.01$, and $p=0.61$, respectively. ",
-  "The right panels shows the quantile-quantile plots for data transformed using ",
+  "The right panel show the quantile-quantile plots for the feature transformed using ",
   "the non-robust and robust location- and scale-invariant transformations. ",
@@
 
-To estimate the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] (version 0.16.0); 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
+To estimate the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] version 0.16.0): 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
 The estimated marginal effects are shown in Figure \ref{fig:marginal-effect-plot}.
@@
 cap <- paste0(
-  "Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 22800 machine learning experiments on 285 datasets. ",
+  "Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets. ",
   "The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). ",
@@
   "Note that the ranges of the $x$-axes of the three main panels differ. ",
-  "ECTN: empirical central normality test."
+  "ECNT: empirical central normality test."
 )
@@
 
-Robust location- and scale-invariant transformations form a suitable replacement for their conventional counterparts.
-They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the data distribution.
+Robust location- and scale-invariant transformations are a suitable replacement for their conventional counterparts.
+They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the distribution of a feature.
 This is particularly relevant for automated data processing, where such issues may go unnoticed.
-However, we did not find a measurable benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature.
+However, we did not find a meaningful benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
@@
 
-Robust location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e.
-transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$, using maximum likelihood optimisation.
+Location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
 Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
@@
 This study had the following limitations.
-First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle variables with negative values, and did not consider the Box-Cox transformation.
-Likewise, we focussed on maximum likelihood estimation to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B).
+First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
+Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B).
 
@@
 
-Robust location- and scale-invariant power transformations allow for automated power transformation of numerical features.
+Compared to their conventional versions, robust location- and scale-invariant Box-Cox and Yeo-Johnson transformations reduce sensitivity to outliers and the location and scale of features.
+An empirical central normality test can assess the quality of transformation of features to normal distributions.
+The combination of both facilitate the use of power transformations in automated data analysis workflows.
 

### 2024-08-30 manuscript/plots.R

@@ -1014,7 +1014,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_bc <- p_bc + plot_theme
-  p_bc <- p_bc + ggplot2::theme(legend.key.size = grid::unit(0.3, 'lines'))
   p_bc <- p_bc + ggplot2::geom_line()
   p_bc <- p_bc + ggplot2::geom_line(
     data = outlier_free_data[method == "Box-Cox"],
@@ -1045,7 +1044,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_yj <- p_yj + plot_theme
-  p_yj <- p_yj + ggplot2::theme(legend.key.size = grid::unit(0.3, 'lines'))
   p_yj <- p_yj + ggplot2::geom_line()
   p_yj <- p_yj + ggplot2::geom_line(
     data = outlier_free_data[method == "Yeo-Johnson"],
@@ -1143,7 +1141,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_bc <- p_bc + plot_theme
-  p_bc <- p_bc + ggplot2::theme(legend.key.size = grid::unit(0.3, 'lines'))
   p_bc <- p_bc + ggplot2::geom_step()
   p_bc <- p_bc + ggplot2::scale_colour_discrete(
     name = "central portion κ\n(Box-Cox)",
@@ -1169,7 +1166,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
   p_yj <- p_yj + plot_theme
-  p_yj <- p_yj + ggplot2::theme(legend.key.size = grid::unit(0.3, 'lines'))
   p_yj <- p_yj + ggplot2::geom_step()
   p_yj <- p_yj + ggplot2::scale_colour_discrete(
     name = "central portion κ\n(Yeo-Johnson)",
@@ -1182,7 +1178,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
       "100 %" = "#904109"
     )
   )
-  p_yj <- p_yj + ggplot2::xlab(latex2exp::TeX("test statistic $\tau_{\text{ecn}}$"))
+  p_yj <- p_yj + ggplot2::xlab(latex2exp::TeX("test statistic $\\tau_{\\text{ecn}}$"))
   p_yj <- p_yj + ggplot2::coord_cartesian(xlim = c(0, 0.25))
   p_yj <- p_yj + ggplot2::theme(
[…]

### 2024-08-30 manuscript/manuscript.Rmd
Updates to appendix.

@@
 ---
-title: "Location and Scale-Invariant Power Transformations for Transforming Data to Normality"
+title: "Location and Scale-Invariant Power Transformations for Transforming Data to
+  Normality"
 author: "Alex Zwanenburg, Steffen Löck"
@@
 knitr::opts_chunk$set(
+  dev = "cairo_pdf",
   echo = FALSE,
   eval = TRUE,
-  fig.align = "center",
-  dev = "cairo_pdf"
+  fig.align = "center"
 )
@@
 To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulated three different distributions.
-The first distribution was a normal distribution, as previously shown in Figure \ref{fig:decreased-normality}.
 We first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.

### 2024-08-30 manuscript/manuscript_appendix.Rmd
Updates to appendix.

@@
 title: "Appendix"
-author: "Alex Zwanenburg"
+author: "Alex Zwanenburg, Steffen Löck"
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  pdf_document:
+    latex_engine: lualatex
+papersize: a4
+bibliography: refs.bib
 header-includes:
@@
   eval = TRUE,
-  fig.align= "center")
+  fig.align = "center",
+  dev = "cairo_pdf"
+)
 
@@
 require(robustHD)
-require(modeldata)
 
@@
 plot_theme$plot.title$size <- ggplot2::rel(1.0)
+plot_theme$plot.title$hjust <- 0.5
 plot_theme$strip.background <- ggplot2::element_blank()
@@
 plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
+plot_theme$legend.key.size <- grid::unit(0.3, 'lines')
 
@@
 }
+
 ```
@@
 
-## Box-Cox transformation
+## Location- and scale-invariant Box-Cox transformation
 
-For the Box-Cox transformation the partial derivative is:
+For the location- and scale-invariant Box-Cox transformation the partial derivative is:
 ```{=latex}
@@
 
-This leads to the following log-likelihood for the Box-Cox transformation:
+This leads to the following log-likelihood:
 ```{=latex}
@@
 
-We introduced per-sample weights $w_i$ to facilitate robust power transformations. The log likelihood of the robust Box-Cox transformation is:
+We introduced per-sample weights $w_i$ to facilitate robust power transformations. The weighted log-likelihood of the location- and scale-invariant Box-Cox transformation is:
 ```{=latex}
@@
 
-## Yeo-Johnson transformation
+## Location- and scale-invariant Yeo-Johnson transformation
 
-For the Yeo-Johnson transformation, the partial derivative is:
+For the location- and scale-invariant Yeo-Johnson transformation, the partial derivative is:
 ```{=latex}
@@
 
-This lead to the following log-likelihood for the Yeo-Johnson transformation:
+This leads to the following log-likelihood:
 ```{=latex}
@@
 
-The log likelihood of the robust Yeo-Johnson transformation is:
+The weighted log-likelihood for location- and scale-invariant Yeo-Johnson transformation is:
 ```{=latex}
@@
 
-Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\psi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0} =\psi^{\lambda, x_0} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
+Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
 Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
@@
 For non-robust power transformation, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
-For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].
+For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
 
@@
 
-The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$. We define this criterion as follows:
+The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$.
+We define this criterion as follows:
 
@@
 \begin{equation}
-U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right) }
+U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right) }
 \end{equation}
@@
 
-Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters. 
 For non-robust power transformations, all $w_i = 1$.
-Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include any extraneous weights $w_i$.
+Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include weights.
 The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.
@@
 
-The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
+The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$.
+We define the Cramér-von Mises criterion as follows:
 ```{=latex}
 \begin{equation}
-U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2
+U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2
 \end{equation}
@@
 
-Here $\mu$ and $\sigma$ are location and scale parameters as described above, and $w_i$ are weights as described in section **Insert**.
+Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters.
 For non-robust power transformations, all $w_i = 1$.
-The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of extraneous weights.
+The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of weights.
 This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
@@
 
-For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
+For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
 
@@
 \begin{equation}
-\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0} \left(x_i \right)} {\sum_{i=1}^n w_i}
+\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0, s} \left(x_i \right)} {\sum_{i=1}^n w_i}
 \end{equation}
@@
 
-The location, or mean, is weighted using weights $w_i$, as described in section **Insert**. For non-robust transformations, $w_i = 1$. Then, $sigma^2$ is defined as:
+The location, or mean, is weighted using weights $w_i$.
+For non-robust transformations, $w_i = 1$.
+Then, $sigma^2$ is defined as:
 ```{=latex}
 \begin{equation}
-\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
+\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
 \end{equation}
@@
 \begin{equation}
-s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
+s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
 \end{equation}
@@
 \begin{equation}
-k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
+k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
 \end{equation}
@@
 
-The second criterion based on skewness and kurtosis is the Jarque-Bera criterion. It is relatively simple to compute, compared to the D'Agostino criterion:
+The second criterion based on skewness and kurtosis is the Jarque-Bera criterion.
+It is relatively simple to compute compared to the D'Agostino criterion:
 ```{=latex}
@@
 
-# Appendix B: Simulations with other optimisation criteria
+# Appendix C: Simulations with other optimisation criteria
+
+Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in (Appendix B)[Appendix B: Optimisation of transformation parameters].
+This follows the simulation in the main manuscript, where MLE was used for optimization.
+In short, we first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
+To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
+Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.
+
+The results are shown in Figure \ref{fig:shifted-distributions-appendix}.
 
@@
 ```
+
+# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation

### 2024-08-30 manuscript/manuscript_appendix.Rmd
Further changes to the appendix.

@@
 cap <- paste0(
-  "Invariant power transformation produces transformation parameters that are invariant to location and shape. ",
+  "Invariant power transformation produces transformation parameters that are invariant to location and scale. ",
   "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\\text{shift}}$ or multiplication by $d_{\\text{scale}}$. ",
-  "Estimates of transformation parameter $\\lambda$ with the conventional power transformations (circle) show strong dependency on the overall location and scale of the distribution, ",
-  "whereas estimates with the invariant power transformations (triangle) are constant."
+  "Estimates of the transformation parameter $\\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution and the optimisation criterion, ",
+  "whereas estimates obtained for the location- and scale-invariant power transformations are constant."
 )

### 2024-09-03 manuscript/manuscript.Rmd
Added combined plot.

@@
 plot_theme$legend.key.size <- grid::unit(0.3, 'lines')
+plot_theme$plot.tag <- ggplot2::element_text(
+  size = ggplot2::rel(1.0),
+  hjust = 0.0,
+  vjust = 0.0,
+  inherit.blank = FALSE
+)
+plot_theme$plot.tag.position <- c(0, 1)
 

### 2024-09-03 manuscript/manuscript_appendix.Rmd
Added combined plot.

@@
 # Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation
+
+```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Blabla"
+)
+
+data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
+data$p
+```

### 2024-09-03 manuscript/plots.R

@@ -1740,7 +1740,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p1_qq <- p1_qq + ggplot2::theme(
     axis.title.y = ggplot2::element_blank(),
-    axis.ticks.y = ggplot2::element_blank()
+    axis.text.y = ggplot2::element_blank()
   )
   p1_d <- p_d + ggplot2::geom_density(
     data = transformed_data[transformation %in% c("conventional", "Raymaekers-Rousseeuw")]
@@ -1753,7 +1753,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p2_qq <- p2_qq + ggplot2::theme(
     axis.title.y = ggplot2::element_blank(),
-    axis.ticks.y = ggplot2::element_blank()
+    axis.text.y = ggplot2::element_blank()
   )
   p2_d <- p_d + ggplot2::geom_density(
     data = transformed_data[transformation %in% c("invariant", "robust invariant")]
@@ -1767,6 +1767,344 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
 
   return(list("data" = data, "plot"= p))
+}
+
+
+
+.plot_experimental_invariance <- function(plot_theme, lambda_limit = NULL, method = "yeo_johnson") {
+
+  # Age
+  p_age <- ..plot_combined_real_data(
+    plot_theme = plot_theme,
+    x = survival::lung$age,
+    data_name = "age",
+    lambda_limit = lambda_limit,
+    method = method,
+    first_set = TRUE
+  )
+
+  # Penguin body mass
+  x <- modeldata::penguins$body_mass_g
+  x <- x[is.finite(x)]
+  p_pen <- ..plot_combined_real_data(
+    plot_theme = plot_theme,
+    x = x,
+    data_name = "penguin body mass",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  # Housing latitude
+  p_lat <- ..plot_combined_real_data(
+    plot_theme = plot_theme,
+    x = modeldata::ames$Latitude,
+    data_name = "latitude",
+    lambda_limit = lambda_limit,
+    method = method,
+    last_set = TRUE
+  )
+
+  p <- p_age$density[[1]] + p_age$density[[2]] + p_age$density[[3]] +
+    p_age$qq[[1]] + p_age$qq[[2]] + p_age$qq[[3]] +
[…]

### 2024-09-03 manuscript/manuscript.Rmd
Fixed a citation error.

@@
 
-Many statistical and several machine learning methods assume normality of the underlying data.
+Many statistical and several machine learning methods assume normality of the underlying data, e.g.
+analysis of variance or .
 However, numerical features in datasets may strongly deviate from normal distributions, e.g.
 by being skewed.
-Power transformations aim to stabilise variance and improve normality of such features [@Bartlett1947-rx, @Tukey1957-rt].
+Power transformations aim to stabilise variance and improve normality of such features [@Bartlett1947-rx; @Tukey1957-rt].
 The two most commonly used transformations are that of @Box1964-mz and @Yeo2000-vw.

### 2024-09-05 manuscript/manuscript.Rmd
Additional improvements to manuscript.

@@
 
-Many statistical and several machine learning methods assume normality of the underlying data, e.g.
-analysis of variance or .
-However, numerical features in datasets may strongly deviate from normal distributions, e.g.
-by being skewed.
+Many statistical and some machine learning methods assume normality of the underlying data, e.g. analysis of variance and linear discriminant analysis.
+However, numerical features in datasets may strongly deviate from normal distributions, e.g. by being skewed.
 Power transformations aim to stabilise variance and improve normality of such features [@Bartlett1947-rx; @Tukey1957-rt].
@@
 
-Power transformation does not guarantee that transformed features are normally distributed.
-Depending on location and scale of a feature and the presence of outliers, power transformation may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
+Applying a power transformation does not guarantee that transformed features are normally distributed.
+Depending on location and scale of a feature and the presence of outliers, power transformations may decrease normality, as shown in Figure \ref{fig:decreased-normality}.
 If normality of the transformed feature is not checked, e.g. in automated power transformation in machine learning workflows, several issues may arise.
@@
 Likewise, machine learning methods that assume normality of input features may suffer a decrease in performance.
-Also, large negative or positive $\lambda$-parameters may lead to numeric issues in any subsequent computations.
+Moreover, large negative or positive $\lambda$-parameters may lead to numeric issues in any subsequent computations.
 
@@
 ```
-where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+subject to $x_i - x_0 > 0$.
+$\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood function is:
@@
 ```
-where $\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+where $\mu$ and $\sigma^2$ are the mean and variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 
@@
 \begin{split}
-\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
@@
 \begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
 \end{equation}
@@
 \begin{split}
-\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 \begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
 \end{equation}
@@
 The weights $w_i$ in equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood} can be set using several weighting functions. 
-Here, we investigate three:
+Using $\dot{x}_i$ as an argument that will be defined later, we investigate three weighting functions:
 
@@
 - The weighting function uses empirical probabilities of the distribution of the original feature $\mathbf{X}$.
-  Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $\mathbf{X}$.
+  After sorting $\mathbf{X}$ in ascending order, probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $\mathbf{X}$.
   Then $\dot{x}_i = p^{*}_i=2 \left( p_i - 0.5\right)$, so that argument is zero-centered.
@@
 
-- The weighting function uses the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
+- After sorting $\mathbf{X}$ in ascending order, the weighting function uses the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.
   Then $\dot{x}_i = r_i$.
@@
 \begin{equation}
-f_{\beta}(x) = \frac{\beta}{2\Gamma \left(1 / \beta \right)} e^{-\left| x \right|^\beta}
+f_{\beta}(x) = \frac{\beta}{2\Gamma\left(1 / \beta \right)} e^{-\left| x \right|^\beta}
 \end{equation}
@@
 ```
-The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}(p)$ thus incorporates the quantile function $F_{\beta}^{-1}(p^{*})$ of the symmetric generalised normal distribution.
-$F_{\beta}^{-1}(p)$ was derived by Griffin to be [@Griffin2018-bf]:
+The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}$ thus incorporates the quantile function $F_{\beta}^{-1}$ of the symmetric generalised normal distribution.
+$F_{\beta}^{-1}$ was derived by Griffin to be [@Griffin2018-bf]:
 
@@
 
-Let central normality be defined as normality of central portion $\kappa$ of the data, i.e.
+Let central normality be defined as the normality of central portion $\kappa$ of the data, i.e.
 $\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$, with $p_i$ probabilities of the empirical distribution, as previously.
@@
 ```
-Here $[\quad]$ denotes an Iverson bracket. The test statistic is equal to the mean of the residual errors of the central portion of the data.
+Here $[\quad]$ denotes an Iverson bracket.
+The test statistic is equal to the mean of the residual errors of the central portion of the data.
+
+To use the test statistic, the central portion of the data needs to be defined and the Type 1 error rates determined.
+We will do so in the [Simulation](#simulation) section.
 
@@
 \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
+The required algorithms were implemented in the `power.transform` R software package **ADD REFERENCE**.
+Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.
 
@@
 We first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
-The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
-The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
+The second distribution was a right-skewed generalised normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution was a left-skewed generalised normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+We then computed transformation parameter $\lambda$ using the original definitions (Eqn.
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (Eqn.
+\ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution.
 To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
@@
 Outlier values were then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
-This was repeated $m_{out} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
+This was repeated $m_{\text{out}} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
 Thus up to 10 percent of the values could be replaced by outliers.
@@
 $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
-Outlier values randomly replaced values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$.
+Outlier values randomly replaced values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$, with $j \in \left\{ 1, \ldots, m_{\text{out}} \right\}$
 
@@
 \label{eqn:minimisation-weighting}
-\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
+\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
 \end{equation}

### 2024-09-05 manuscript/manuscript.Rmd
Added comment on weighting methods and lack of good results.

@@
 \begin{center}
-\caption{Optimal weighting parameters and corresponding loss for \textbf{location- and scale-invariant Yeo-Johnson} power transformations.
+\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
 $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
@@
 
+We assessed several methods for robust power transformation.
+Methods using the z-score of the transformed feature or the residual error yielded worse results than the non-robust method.
+Different initial values, closer to the upper limits ($\delta_1 = 10.0; \delta_2 = 10.0$), led to a reduced loss.
+However, these values effectively correspond to a non-robust transformation, where all instances receive the same weight.
+Underperformance of these methods could be explained by their use of transformed features for setting weights.
+In essence, parameter optimisation with MLE also affects weights.
+This may lead to different optimal values for transformation parameter $\lambda$, which increases the presented loss.
+
 This study had the following limitations.

### 2024-09-05 manuscript/manuscript_appendix.Rmd
Preparation for comparing type I error rates.

@@
 
-# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation
+# Appendix D: Empirical central normality test
+
+The empirical central normality test was derived using data sampled from asymmetric generalised normal distributions, including outliers, to resemble more realistic datasets.
+Here we assess the type I error rate of two, less realistic, sets of data:
+
+1. Data sampled from asymmetric generalised normal distributions without outliers.
+2. Data sampled from normal distributions without outliers, without any power transformation applied.
+
+Other aspects of the experiment remain the same.
+Thus, we first drew $m_d=10000$ random distributions.
+For asymmetric generalised normal distributions, each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+For fully normal distributions, skewness parameter $\alpha = 0.5$ and shape parameter $\beta = 2.0$ were fixed.
+Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
+$n = \lceil 10^\gamma \rceil$ values were then randomly drawn, with $\gamma \sim U\left(1.47, 3.00\right)$, which led to between $30$ and $1000$ values being drawn to create $\mathbf{X}_i$.
+Residuals were then computed after performing robust location- and scale-invariant transformations with the empirical tapered cosine weighting method for the dataset with asymmetric generalised normal distributions,
+and without any transformation for the dataset with only normal distributions.
+
+```{r empirical-central-normality-test-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Blabla"
+)
+
+```
+
+# Appendix E: Experimental results using location- and scale-invariant Box-Cox transformation
 

### 2024-09-06 manuscript/plots.R
Added new plot.

@@ -1201,7 +1201,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
       "100 %" = "#324b67"
     )
   )
-  p_bc <- p_bc + ggplot2::xlab("test statistic")
+  p_bc <- p_bc + ggplot2::xlab(latex2exp::TeX("test statistic $\\tau_{ecn}$"))
   p_bc <- p_bc + ggplot2::ylab("type I error rate")
   p_bc <- p_bc + ggplot2::coord_cartesian(xlim = c(0, 0.25))
 
@@ -1226,7 +1226,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
       "100 %" = "#904109"
     )
   )
-  p_yj <- p_yj + ggplot2::xlab(latex2exp::TeX("test statistic $\\tau_{\\text{ecn}}$"))
+  p_yj <- p_yj + ggplot2::xlab(latex2exp::TeX("test statistic $\\tau_{ecn}$"))
   p_yj <- p_yj + ggplot2::coord_cartesian(xlim = c(0, 0.25))
   p_yj <- p_yj + ggplot2::theme(
     axis.title.y = ggplot2::element_blank(),
@@ -1243,94 +1243,192 @@ get_annotation_settings <- function(ggtheme = NULL) {
 .plot_type_1_error_rate_appendix <- function(plot_theme, manuscript_dir) {
   # Prevent warnings due to non-standard evaluation.
   residual <- rejected <- p <- n <- mare <- method <- NULL
+  outlier_id <- type <- NULL
 
-  # Get data
-  data <- .get_test_statistics_data_appendix(manuscript_dir = manuscript_dir)
+  file_name <- file.path(manuscript_dir, "type_1_error_rate_plot_appendix.RDS")
 
-  central_width <- c(0.60, 0.70, 0.80, 0.90, 0.95, 1.00)
+  if (!file.exists(file_name)) {
+    # Get data. We need three datasets:
+    # - Dataset presented in main manuscript.
+    # -
+    data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
+    data_no_outlier <- data.table::copy(data[outlier_id == 1L])
+    data_no_outlier$method <- factor(
+      data_no_outlier$method,
+      levels = c("Box-Cox", "Yeo-Johnson"),
+      labels = c("Box-Cox (no outlier)", "Yeo-Johnson (no outlier)")
+    )
[…]

### 2024-09-06 manuscript/simulations.R

@@ -1576,18 +1576,18 @@
   if (!is.null(residual_fun)) external_fun <- TRUE
 
 
-  .compute_residuals <- function(x, n_sample = 200L) {
+  .compute_residuals <- function(x, ii, n_sample = 200L) {
     # Set interpolation points.
     p_sample <- (seq_len(n_sample) - 1 / 3) / (n_sample + 1 / 3)
 
     # Determine residuals for Box-Cox transformations.
     transformer <- suppressWarnings(power.transform::find_transformation_parameters(
-      x = x$x,
+      x = x,
       method = "none"
     ))
 
     residual_data <- power.transform::get_residuals(
-      x = x$x,
+      x = x,
       transformer = transformer
     )
 
@@ -1600,8 +1600,8 @@
     )$y
 
     data <- data.table::data.table(
-      "distribution_id" = x$parameter$ii,
-      "outlier_id" = x$parameters$jj,
+      "distribution_id" = ii,
+      "outlier_id" = 1L,
       "p" = seq_along(p_sample),
       "residual" = residual,
       "method" = "none"
@@ -1643,25 +1643,13 @@
       beta = beta
     )
 
-    if (parallel) {
-      # Start cluster
-      cl <- parallel::makeCluster(18L)
-
-      # Compute all data in parallel.
-      data <- parallel::parLapply(
-        cl = cl,
-        X = x,
-        fun = residual_fun
-      )
-
-      # Stop cluster.
-      parallel::stopCluster(cl)
-    } else {
-      data <- lapply(
-        X = x,
-        FUN = residual_fun
-      )
-    }
+    data <- mapply(
+      FUN = residual_fun,
+      x = x,
+      ii = seq_along(x),
+      SIMPLIFY = FALSE,
+      USE.NAMES = FALSE
+    )
 
     data <- data.table::rbindlist(data)
 

### 2024-09-06 manuscript/manuscript.Rmd
Added new plot.
Updated appendix E.

@@
 Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80.
-In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is slightly more strict.
+In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is more conservative (see Appendix E).
 This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}
@@
 \label{tab:empirical-central-normality}
-\begin{tabular}{c c c c c c c}
+\begin{tabular}{l | c c c c c c c}
 
 \toprule
-0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
 
 \midrule
-0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
 \bottomrule
@@
 
+The empirical test for central normality differs from other tests for normality in that the test statistic is independent of the number of samples.
+This makes this test more practical for assessing central normality for larger sample numbers, where other test may detect trivial deviations from normality.
+
 This study had the following limitations.
 First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
-Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B).
+Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-D).
 

### 2024-09-06 manuscript/manuscript_appendix.Rmd
Added new plot.
Updated appendix E.

@@
 
-# Appendix D: Empirical central normality test
+# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation
+
+```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Blabla"
+)
+
+data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
+data$p
+```
+
+# Appendix E: Empirical central normality test
 
@@
 cap <- paste0(
-  "Blabla"
+  "Type 1 error rate as function of the test statistic $\\tau_{\\text{ecn}}$ ",
+  "for five datasets, with central portion $\\kappa=0.80$. ",
+  "The type 1 error rate is computed from $m_d=10000$ randomly sampled features, ",
+  "These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), ",
+  "or normal distributions with outliers (none). ",
+  "The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none)."
 )
 
+data <- .plot_type_1_error_rate_appendix(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
+data$p
 ```
 
-# Appendix E: Experimental results using location- and scale-invariant Box-Cox transformation
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
+\label{tab:empirical-central-normality-appendix}
+\begin{tabular}{l | c c c c c c c}
 
-```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Blabla"
-)
+\toprule
+type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
 
-data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
-data$p
+\midrule
+Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
+Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
+Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
+normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 ```
+
+The results are shown in Figure \ref{fig:empirical-central-normality-test-appendix} and Table \ref {tab:empirical-central-normality-appendix}.
+These indicate that the test behaves similarly for the different datasets.
+For low type 1 error rates the test statistic proposed in the main manuscript is more conservative than alternatives 
+based on residuals after Box-Cox transformations of asymmetric generalised normally distributed features or on residuals from strictly normally distributed features.

### 2024-09-06 manuscript/manuscript.Rmd
Further improvements.

@@
 In the following, they are evaluated using examples from real datasets.
-We focus on the Yeo-Johnson transformations because
+We focus on the Yeo-Johnson transformation because of its ability to handle features with negative values.
+Results for Box-Cox transformations are shown in Appendix D.
 
@@
 Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:lung-cancer-age}:
-no transformation (residual sum $\sum r_i = 16.5$);
+no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
 conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
@@
 
-Location- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
-Moreover, conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
+Location- and scale-invariant transformation led to a lower overall residual sum, indicating a better transformation.
+Robust location- and scale-invariant transformation had a higher residual sum compared to the non-robust variant, which may be due to the lack of outliers in the data.
+Conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
 The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.78$).
@@
 The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
-Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:penguin-body-mass}:
+Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:penguin-body-mass}:
 no transformation (residual sum $\sum r_i = 48.0$);
@@
   "Quantile-quantile plots for penguin body mass, measured in grams. ",
-  "The left panel shows the quantile-quantile plot for the original feature ",
+  "The left panel shows the quantile-quantile plot for the original feature. ",
   "The middle panel shows the quantile-quantile plots for the feature transformed using ",
@@
 First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
-Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-D).
+Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-C).
 

### 2024-09-09 manuscript/manuscript_appendix.Rmd
Further improvements.
Work on improving real world example data.

@@
 )
-plot_theme$plot.tag.position <- c(0, 1)
+plot_theme$plot.tag.position <- c(0.2, 1)
 
@@
 
-Other aspects of the experiment remain the same.
+Other aspects of the experiment remained the same.
 Thus, we first drew $m_d=10000$ random distributions.

### 2024-09-11 manuscript/manuscript.Rmd
Updated data and code availability

@@
 
-Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE].
-The manuscript was created using R Markdown and is also available from the power.transform GitHub repository.
-Data and results for the machine learning experiment are separately available from Zenodo [INSERT LINK HERE].
+Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE].
+The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository.
+Data and results for the machine learning experiment are separately available from Zenodo [https://doi.org/10.5281/zenodo.13736671].
 

### 2024-09-11 manuscript/manuscript.Rmd
Updated real-world datasets analysis.

@@
 
+```{r experimental-results-invariance, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=5.0}
+cap <- paste0(
+  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
+  "penguin body mass (middle row); ",
+  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
+  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
+  "the feature transformed using the conventional Yeo-Johnson transformation ",
+  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
+  "and the feature transformed using the non-robust and robust location- and-scale ",
+  "invariant Yeo-Johnson transformations (right column)."
+)
+
+data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "yeo_johnson")
+data$p
+```
+
 ### Age of patients with lung cancer
@@
 The age in the cohort was $62.4 \pm 9.1$ (mean ± standard deviation) years.
-Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:lung-cancer-age}:
+Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
 no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
@@
 
-```{r lung-cancer-age, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
-cap <- paste0(
-  "Quantile-quantile plots for age of patients with lung cancer, expressed in years. ",
-  "The left panel shows the quantile-quantile plot for the original feature ",
-  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
-  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
-  "The right panel shows the quantile-quantile plots for the feature transformed using ",
-  "the non-robust and robust location- and-scale invariant transformations."
-)
-
-out <- .plot_lung_cancer_age(plot_theme = plot_theme)
-out$plot
-```
-
 Location- and scale-invariant transformation led to a lower overall residual sum, indicating a better transformation.
@@
 The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
-Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:penguin-body-mass}:
+Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
 no transformation (residual sum $\sum r_i = 48.0$);
-conventional transformation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.102$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
 location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
@@
 
-```{r penguin-body-mass, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
-cap <- paste0(
-  "Quantile-quantile plots for penguin body mass, measured in grams. ",
-  "The left panel shows the quantile-quantile plot for the original feature. ",
-  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
-  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
-  "The right panel shows the quantile-quantile plots for the feature transformed using ",
-  "the non-robust and robust location- and-scale invariant transformations."
-)
-
-out <- .plot_penguin_body_mass(plot_theme = plot_theme)
-out$plot
-```
 Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
@@
 The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
-Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:ames-latitude}:
+Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
 no transformation (residual sum $\sum r_i = 328$);
@@
 
-```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
-cap <- paste0(
-  "Quantile-quantile plots for latitude coordinate in a housing dataset. ",
-  "The left panel shows the quantile-quantile plot for the original feature ",
-  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
-  "the conventional Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation. ",
-  "The right panel shows the quantile-quantile plots for the feature transformed using ",
-  "the non-robust and robust location- and-scale invariant transformations."
-)
-
-out <- .plot_ames_latitude(plot_theme = plot_theme)
-out$plot
-```
 Every transformation reduced the residual sum.
@@
 
+```{r experimental-results-outlier-robustness, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.5}
+cap <- paste0(
+  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
+  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
+  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
+  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
+  "the feature transformed using the conventional Yeo-Johnson transformation ",
+  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
+  "and the feature transformed using the non-robust and robust location- and-scale ",
+  "invariant Yeo-Johnson transformations (right column). ",
+  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
+)
+
+data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "yeo_johnson")
+data$p
+```
+
 ### Fuel efficiency in the Top Gear dataset
 
-```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
-cap <- paste0(
-  "Quantile-quantile plots for fuel consumption data in the Top Gear dataset. ",
-  "The feature contains three strong outliers related to highly fuel-efficient vehicles.",
-  "The left panel shows the quantile-quantile plot for the original feature. ",
-  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.76$).",
-  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
-  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.55$, respectively. ",
-  "The right panel shows the quantile-quantile plots for the feature transformed using ",
-  "the non-robust and robust location- and scale-invariant transformations. ",
-  "For these, the empirical central normality test yields $p=0.03$, and $p=0.86$, respectively. ",
-  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
-)
-
-out <- .plot_top_gear(
-  plot_theme = plot_theme
-)
-
-out$plot
-```
-
-The Top Gear dataset contains data on 297 vehicles that appeared on the BBC television show *Top Gear*.
+The Top Gear dataset contains data on 297 vehicles that appeared on the BBC television show *Top Gear* [@Alfons2021-kc].
 Within this dataset, the fuel consumption feature contains outliers due to highly fuel-efficient vehicles.
-Figure \ref{fig:top-gear} shows that the outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively).
+
+Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
+no transformation (residual sum $\sum r_i = 110$, $p=0.76$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 137$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.61$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.13$);
+and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.10$).
+
+Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
@@
 
-```{r ischemic-stroke, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
-cap <- paste0(
-  "Quantile-quantile plots for the maximum arterial wall thickness in patients with ischemic stroke. ",
-  "This feature contains two strong outliers. ",
-  "The left panel shows the quantile-quantile plot for the original feature. ",
-  "The empirical central normality test indicates that the original feature does not deviate significantly from central normality ($p=0.56$). ",
-  "The middle panel shows the quantile-quantile plots for the feature transformed using ",
-  "the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation. ",
-  "For these, the empirical central normality test yields $p=0.01$, and $p=0.61$, respectively. ",
-  "The right panel show the quantile-quantile plots for the feature transformed using ",
-  "the non-robust and robust location- and scale-invariant transformations. ",
-  "For these, the empirical central normality test yields $p=0.02$, and $p=0.10$, respectively. ",
-  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
-)
-
-out <- .plot_ischemic_stroke(
-  plot_theme = plot_theme
-)
-
-out$plot
-```
 The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
@@
 The maximum arterial wall thickness feature contains several instances with outlier values.
-Figure \ref{fig:ischemic-stroke} show that non-robust transformation fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
+Figure \ref{fig:experimental-results-outlier-robustness} shows that non-robust transformations fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
@@
 invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
-normality test ($p = 0.01$)), and normalisation method (none,
+normality test (rejecting transformations with $p \leq 0.01$), and normalisation method (none,
 $z$-standardisation), yielding 16 distinct configurations.
@@
 
-This study had the following limitations.
-First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
-Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-C).
+As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (appendix B).
 

### 2024-09-11 manuscript/manuscript_appendix.Rmd
Added plotting function for outlier data.
Updated real-world datasets analysis.

@@
 
+The effect of using location- and scale-invariant transformations was investigated using real-world datasets.
+
+## Invariance
+
+Results for Box-Cox transformations of features without outliers are shown in Figure \ref{fig:experimental-results-invariance-appendix}.
+
 ```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
 cap <- paste0(
-  "Blabla"
+  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
+  "penguin body mass (middle row); ",
+  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
+  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
+  "the feature transformed using the conventional Box-Cox transformation ",
+  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
+  "and the feature transformed using the non-robust and robust location- and-scale ",
+  "invariant Box-Cox transformations (right column)."
 )
@@
 ```
+
+### Age of patients with lung cancer
+
+Applying conventional and invariant Box-Cox transformations to age of patients with lung cancer [@Loprinzi1994-cd] yielded the following results:
+no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
+conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
+location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$);
+and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$).
+
+Compared to location- and scale-invariant Yeo-Johnson transformations, the Box-Cox transformations do not reduce residuals compared to conventional variants.
+
+### Penguin body mass
+
+Applying conventional and invariant Box-Cox transformations to the body mass of penguins [@Gorman2014-eo] yielded the following results:
+no transformation (residual sum $\sum r_i = 48.0$);
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$);
+and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$).
+
+Just as for location- and scale-invariant Yeo-Johnson transformations, Box-Cox transformations produced a lower overall residual sum compared to their conventional counterparts.
+Similarly, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
+
+### Latitude in the Ames housing dataset
+
+Applying conventional and invariant Box-Cox transformations to the latitude of houses in the Ames housing dataset [@De-Cock2011-jf] yielded the following results:
+no transformation (residual sum $\sum r_i = 328$);
+conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$);
+location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$);
+and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$.
+
+Similar to conventional Yeo-Johnson transformations (non-robust and robust), conventional Box-Cox transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
+Location- and scale-invariant Box-Cox transformations did not suffer from this issue.
+
+## Robustness against outliers
+
+Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:results-outlier-robustness-appendix}.
+
+```{r experimental-results-outlier-robustness-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
+cap <- paste0(
+  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
+  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
+  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
+  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
+  "the feature transformed using the conventional Box-Cox transformation ",
+  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
+  "and the feature transformed using the non-robust and robust location- and-scale ",
+  "invariant Box-Cox transformations (right column). ",
+  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
+)
+
+data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "box_cox")
+data$p
+```
+
+### Fuel efficiency in the Top Gear dataset
+
+The Top Gear dataset contains data on 297 vehicles, with outliers related highly fuel-efficient vehicles [@Alfons2021-kc].
+Applying conventional and invariant Box-Cox transformations to the fuel consumption feature yielded the following results:
+no transformation (residual sum $\sum r_i = 54$, $p=0.76$);
+conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.55$);
+location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{BC} = 0.5$, $\sigma_{BC} = 0.1$, $p=0.03$);
+and robust location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 49$, $\mu_{BC} = 2.0$, $\sigma_{BC} = 2.3$, $p=0.58$).
+
+### Maximum arterial wall thickness in an ischemic stroke dataset
+
+The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
+Applying conventional and invariant Box-Cox transformations to the maximum arterial wall thickness feature yielded the following results:
+no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.1$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{BC} = 7.2$, $\sigma_{BC} = 14$, $p=0.61$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{BC} = -11.8$, $\sigma_{BC} = 6.9$, $p=0.13$);
+and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.10$).
 

### 2024-09-11 manuscript/manuscript.Rmd
Added data on fits for outlier example data.

@@
 Within this dataset, the fuel consumption feature contains outliers due to highly fuel-efficient vehicles.
-
 Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 110$, $p=0.76$);
-conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 137$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.61$);
-location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.13$);
-and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.10$).
+no transformation (residual sum $\sum r_i = 54$, $p=0.56$);
+conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.61$);
+location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p=0.13$);
+and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.10$).
 

### 2024-09-12 manuscript/manuscript.Rmd
Added details on data with outliers.

@@
 conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.61$);
-location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p=0.13$);
-and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.10$).
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.55$);
+location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p=0.03$);
+and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.58$).
 
@@
 The maximum arterial wall thickness feature contains several instances with outlier values.
-Figure \ref{fig:experimental-results-outlier-robustness} shows that non-robust transformations fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
+Applying conventional and invariant Yeo-Johnson transformations to this feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
+no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.61$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.13$);
+and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.10$).
+
+Non-robust transformations failed to produce a centrally normal distribution (empirical central normality test $p=0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
@@
 This is particularly relevant for automated data processing, where such issues may go unnoticed.
-However, we did not find a meaningful benefit for -- nor detriment to -- model performance in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature.
+Compared to non-robust location- and scale-invariant transformations, real-world examples with outliers showed higher residual errors of transformed features. 
+Robust transformations seek to minimise residual errors for the central part of the distribution, instead of the entire distribution, including outliers.
+The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
+However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit of for -- nor detriment to -- model performance for location- and scale-invariant power transformations.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.

### 2024-09-13 manuscript/manuscript.Rmd
Fixed missing passing of method argument.

@@
 Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 54$, $p=0.56$);
+no transformation (residual sum $\sum r_i = 54$, $p=0.76$);
 conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p=0.01$);

### 2024-09-13 manuscript/manuscript_appendix.Rmd
Fixed missing passing of method argument.

@@
 Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.55$);
-location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{BC} = 0.5$, $\sigma_{BC} = 0.1$, $p=0.03$);
-and robust location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 49$, $\mu_{BC} = 2.0$, $\sigma_{BC} = 2.3$, $p=0.58$).
+location- and scale-invariant transformation ($\lambda = -0.7$, $\sum r_i = 44$, $\mu_{BC} = 0.6$, $\sigma_{BC} = 0.2$, $p=0.02$);
+and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 59$, $\mu_{BC} = 2.4$, $\sigma_{BC} = 1.8$, $p=0.83$).
 
@@
 no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
-conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.1$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{BC} = 7.2$, $\sigma_{BC} = 14$, $p=0.61$);
-location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{BC} = -11.8$, $\sigma_{BC} = 6.9$, $p=0.13$);
-and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.10$).
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 30$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.60$);
+location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 12$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.01$);
+and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.35$).
 
@@
 based on residuals after Box-Cox transformations of asymmetric generalised normally distributed features or on residuals from strictly normally distributed features.
+
+# Appendix F: Normalisation before transformation
+
+```{r include = FALSE, eval = TRUE}
+
+data <- .assess_standardisation_before_transformation()
+```

### 2024-09-13 manuscript/plots.R

@@ -2022,6 +2022,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = x,
     data_name = "fuel efficiency",
     lambda_limit = lambda_limit,
+    method = method,
     first_set = TRUE
   )
 
@@ -2031,6 +2032,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
     x = modeldata::ischemic_stroke$max_max_wall_thickness,
     data_name = "arterial wall thickness",
     lambda_limit = lambda_limit,
+    method = method,
     last_set = TRUE
   )
 

### 2024-09-13 manuscript/manuscript_appendix.Rmd
Corrected values.

@@
 no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 30$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.60$);
-location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 12$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.01$);
-and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.35$).
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 33$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 127$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.60$);
+location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 28$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.01$);
+and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.35$).
 

### 2024-09-13 manuscript/manuscript_appendix.Rmd
Work on assessing the effect of normalisation prior to standardisation.

@@
 data <- .assess_standardisation_before_transformation()
+data <- data[transformation %in% c("none", "conventional", "conventional (z-score norm.)", "conventional (robust scaling)", "invariant"), mget(c("residuals", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data.table::dcast(data, dataset ~ transformation, value.var = "residuals")
 ```

### 2024-09-13 manuscript/simulations.R

@@ -2119,3 +2119,182 @@
 
   return(feature_data)
 }
+
+
+
+.assess_standardisation_before_transformation <- function(
+    lambda_limit = NULL,
+    method = "yeo_johnson"
+) {
+  data <- list()
+
+  # Age of lung cancer patients
+  data[[1L]] <- ..standardisation_before_transformation(
+    x = survival::lung$age,
+    data_name = "age",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  # Penguin body mass
+  x <- modeldata::penguins$body_mass_g
+  x <- x[is.finite(x)]
+  data[[2L]] <- ..standardisation_before_transformation(
+    x = x,
+    data_name = "penguin body mass",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  # Housing latitude
+  data[[3L]] <- ..standardisation_before_transformation(
+    x = modeldata::ames$Latitude,
+    data_name = "latitude",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  # Fuel efficiency
+  utils::data("TopGear", package = "robustHD", envir = environment())
+  tg_data <- data.table::as.data.table(TopGear)
+  x <- tg_data$MPG
+  x <- x[!is.na(x)]
+  data[[4L]] <- ..standardisation_before_transformation(
+    x = x,
+    data_name = "fuel efficiency",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  # Maximum arterial wall thickness
+  data[[5L]] <- ..standardisation_before_transformation(
+    x = modeldata::ischemic_stroke$max_max_wall_thickness,
+    data_name = "arterial wall thickness",
+    lambda_limit = lambda_limit,
+    method = method
+  )
+
+  data <- rbindlist(data, use.names = TRUE)
+  return(data)
+}
+
+
+..standardisation_before_transformation <- function(
+    x,
+    data_name,
+    lambda_limit = c(-4.0, 6.0),
+    method = "yeo_johnson"
+) {
+
+  ..parse_data <- function(
+    transformation_method,
+    x
+  ) {
+
+    if (endsWith(transformation_method, "(z-score norm.)")) {
+      x <- (x - mean(x)) / stats::sd(x)
+
+    } else if (endsWith(transformation_method, "(robust scaling)")) {
+      x <- (x - stats::median(x)) / stats::IQR(x)
+    }
+
+    if (startsWith(transformation_method, "none")) {
+      # No transformer
[…]

### 2024-09-16 manuscript/manuscript.Rmd
Added more details regarding datasets containing mostly numeric features that display a large shift.
Updated software references.

@@
 \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-The required algorithms were implemented in the `power.transform` R software package **ADD REFERENCE**.
+The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq].
 Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.
@@
 Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
+For the latter two datasets, the transformation method does not show significant difference between groups (Friedman test; $p > 0.05$).
+```{r include = FALSE, eval = TRUE}
+
+data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
+data <- data[dataset %in% c("Hill_Valley_with_noise", "Hill_Valley_without_noise")]
+
+# Convert to data matrix. First aggregate within each experiment because the
+# Friedman test is not for repeated measurements.
+data <- data.table::dcast(
+  data,
+  dataset ~ transformation_method, value.var="value_rank",
+  fun.aggregate = mean
+)
+
+h_transformation <- friedman.test(as.matrix(data[, "dataset" := NULL]))
+```
 
@@
 
-Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE].
+Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [GitHub](https://github.com/oncoray/power.transform) and the [CRAN repository](https://cran.r-project.org/package=power.transform).
 The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository.
-Data and results for the machine learning experiment are separately available from Zenodo [https://doi.org/10.5281/zenodo.13736671].
+Data and results for the machine learning experiment are separately available from [Zenodo](https://doi.org/10.5281/zenodo.13736671).
 

### 2024-09-16 manuscript/manuscript.Rmd
Added note normalisation prior to transformation.

@@
 Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
+While this alternative helps prevent numerical issues -- provided that normalisation does not lead to negative values for Box-Cox transformation -- location- and scale-invariant transformations seem to provide an overall better transformation to normality (Appendix F).
 
@@
 
-As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (appendix B).
+As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (Appendix B).
 

### 2024-09-16 manuscript/manuscript_appendix.Rmd
Added note normalisation prior to transformation.

--- 2024-09-13 manuscript/manuscript_appendix.Rmd
@@
 
-```{r include = FALSE, eval = TRUE}
+An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
+Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors, after transformation to normality, of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
+In these examples location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
+
+```{r include = FALSE, eval = FALSE}
 
 data <- .assess_standardisation_before_transformation()
+data <- data[transformation %in% c("none", "conventional", "conventional (z-score norm.)", "conventional (robust scaling)", "invariant"), mget(c("residuals", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data.table::dcast(data, dataset ~ transformation, value.var = "residuals")
 ```
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
+\label{tab:normalisation-before-transformation-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
+
+\midrule
+age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
+penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
+latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
+fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
+arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
+# References

### 2024-09-17 manuscript/manuscript.Rmd
Revised discussion.

--- 2024-09-16 manuscript/manuscript.Rmd
@@
 \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq].
+The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq] (version 1.0.0).
 Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.
@@
 Location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
+We derived the log-likelihood function to facilitate optimisation using MLE.
 Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
+While this alternative helps prevent these issues -- provided that normalisation does not lead to negative values for Box-Cox transformation -- location- and scale-invariant transformations seem to provide an overall better transformation to normality (Appendix F).
 
 We assessed several methods for robust power transformation.
-Methods using the z-score of the transformed feature or the residual error yielded worse results than the non-robust method.
-Different initial values, closer to the upper limits ($\delta_1 = 10.0; \delta_2 = 10.0$), led to a reduced loss.
+Methods that relied on the z-score of the transformed feature or the residual error yielded worse results than the non-robust method.
+This is partly due to the initial choice of threshold parameters.
+Using different initial values, closer to the upper limits ($\delta_1 = 10.0; \delta_2 = 10.0$), led to a reduced loss.
 However, these values effectively correspond to a non-robust transformation, where all instances receive the same weight.
 Underperformance of these methods could be explained by their use of transformed features for setting weights.
-In essence, parameter optimisation with MLE also affects weights.
-This may lead to different optimal values for transformation parameter $\lambda$, which increases the presented loss.
-
-The empirical test for central normality differs from other tests for normality in that the test statistic is independent of the number of samples.
-This makes this test more practical for assessing central normality for larger sample numbers, where other test may detect trivial deviations from normality.
-
-As a limitation, we did observe several numerical issues for optimisation criteria other than MLE (appendix B).
+Consequently, the weights change at each iteration in the MLE optimisation process.
+This increases local variance in the log-likelihood function and create local optima that the optimiser may not handle well.
+As a consequence, optimal values for transformation parameter $\lambda$ might differ, which increases the presented loss for optimising threshold parameters.
+Methods that relied on the empirical probability did not suffer from this issue, as weights remained fixed during MLE.
+
+We introduced an empirical test for central normality to assess whether distributions deviate from normality in a way that might require closer inspection prior to further processing.
+The empirical test for central normality differs from other tests for normality, such as the Shapiro-Wilk test [@Shapiro1965-zd], in that the test statistic is independent of the number of samples.
+This makes this test more practical for assessing central normality for larger sample numbers, where other test may detect inconsequential deviations from normality.
+
+This work has the following limitation.
+Firstly, we did observe several numerical stability issues for optimisation criteria other than MLE (Appendix B).
+These appear in regions where transformation parameters would lead to very large or small numbers when using conventional power transformations.
+For MLE stability issues were not observed.
+Secondly, the empirical central normality test is based on simulations instead of statistical theory, and relies on a somewhat arbitrary definition of the central portion of a distribution.
+Thus, while the test may asses whether data is sufficiently normally distributed for practical purposes, it should not be used as a strict test for normality.
 

### 2024-09-18 manuscript_rmarkdown/manuscript.Rmd
Moved rmarkdown to its own folder.

--- 2024-09-17 manuscript/manuscript.Rmd
@@
 
-manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")
+manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript_rmarkdown")
 source(file.path(manuscript_dir, "simulations.R"))

### 2024-09-19 manuscript_latex/manuscript.tex
Further work on transition to LaTeX.

%%
%% Copyright 2007-2024 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.3 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
\usepackage{amsmath}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Computational Statistics and Data Analysis}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={},
%%            city={},
%%            postcode={},
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{Location and Scale-Invariant Power Transformations for Transforming Data to Normality} %% Article title

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[affil_1,affil_2]{Alex Zwanenburg} %% Author name

%% Author affiliation
\affiliation[affil_1]{
  organization={
    National Center for Tumor Diseases Dresden (NCT/UCC):
    German Cancer Research Center (DKFZ), Heidelberg, Germany;
    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, Dresden, Germany;
    Helmholtz-Zentrum Dresden-Rossendorf (HZDR), Dresden, Germany
  },
  addressline={Fetscherstraße 74/PF 64},
  city={Dresden},
  postcode={01307},
  country={Germany}
}
\affiliation[affil_2]{
  organization={
    OncoRay – National Center for Radiation Research in Oncology,
    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology,
    Helmholtz-Zentrum Dresden-Rossendorf
  },
  addressline={Fetscherstraße 74/PF 41},
  city={Dresden},
  postcode={01307},
  country={Germany}
}
\author[affil_2,affil_3,affil_4]{Steffen Löck} %% Author name

%% Author affiliation
\affiliation[affil_3]{
  organization={
    Department of Radiotherapy and Radiation Oncology,
    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology
  },
  addressline={Fetscherstraße 74/PF 50},
  city={Dresden},
  postcode={01307},
  country={Germany}
}
\affiliation[affil_4]{
  organization={
   German Cancer Consortium (DKTK), Partner Site Dresden,
   and German Cancer Research Center (DKFZ)
  },
  addressline={Im Neuenheimer Feld 280},
  city={Heidelberg},
  postcode={69192},
  country={Germany}
}

%% Abstract
\begin{abstract}
%% Text of abstract
Abstract text.
\end{abstract}

%%Graphical abstract
\begin{graphicalabstract}
%\includegraphics{grabs}
\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item Research highlight 1
\item Research highlight 2
\end{highlights}

%% Keywords
\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% Add \usepackage{lineno} before \begin{document} and uncomment
%% following line to enable line numbers
%% \linenumbers

%% main text
%%

%% Use \section commands to start a section
\section{Example Section}
\label{sec1}
%% Labels are used to cross-reference an item using \ref command.

Section text. See Subsection \ref{subsec1}.

%% Use \subsection commands to start a subsection.
\subsection{Example Subsection}
\label{subsec1}

Subsection text.

%% Use \subsubsection, \paragraph, \subparagraph commands to
%% start 3rd, 4th and 5th level sections.
%% Refer following link for more details.
%% https://en.wikibooks.org/wiki/LaTeX/Document_Structure#Sectioning_commands

\subsubsection{Mathematics}
%% Inline mathematics is tagged between $ symbols.
This is an example for the symbol $\alpha$ tagged as inline mathematics.

%% Displayed equations can be tagged using various environments.
%% Single line equations can be tagged using the equation environment.
\begin{equation}
f(x) = (x+a)(x+b)
\end{equation}

%% Unnumbered equations are tagged using starred versions of the environment.
%% amsmath package needs to be loaded for the starred version of equation environment.
\begin{equation*}
f(x) = (x+a)(x+b)
\end{equation*}

%% align or eqnarray environments can be used for multi line equations.
%% & is used to mark alignment points in equations.
%% \\ is used to end a row in a multiline equation.
\begin{align}
 f(x) &= (x+a)(x+b) \\
      &= x^2 + (a+b)x + ab
\end{align}

\begin{eqnarray}
 f(x) &=& (x+a)(x+b) \nonumber\\ %% If equation numbering is not needed for a row use \nonumber.
      &=& x^2 + (a+b)x + ab
\end{eqnarray}

%% Unnumbered versions of align and eqnarray
\begin{align*}
 f(x) &= (x+a)(x+b) \\
      &= x^2 + (a+b)x + ab
\end{align*}

\begin{eqnarray*}
 f(x)&=& (x+a)(x+b) \\
     &=& x^2 + (a+b)x + ab
\end{eqnarray*}

%% Refer following link for more details.
%% https://en.wikibooks.org/wiki/LaTeX/Mathematics
%% https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics

%% Use a table environment to create tables.
%% Refer following link for more details.
%% https://en.wikibooks.org/wiki/LaTeX/Tables
\begin{table}[t]%% placement specifier
%% Use tabular environment to tag the tabular data.
%% https://en.wikibooks.org/wiki/LaTeX/Tables#The_tabular_environment
\centering%% For centre alignment of tabular.
\begin{tabular}{l c r}%% Table column specifiers
%% Tabular cells are separated by &
  1 & 2 & 3 \\ %% A tabular row ends with \\
  4 & 5 & 6 \\
  7 & 8 & 9 \\
\end{tabular}
%% Use \caption command for table caption and label.
\caption{Table Caption}\label{fig1}
\end{table}

%% Use figure environment to create figures
%% Refer following link for more details.
%% https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions
\begin{figure}[t]%% placement specifier
%% Use \includegraphics command to insert graphic files. Place graphics files in
%% working directory.
\centering%% For centre alignment of image.
\includegraphics{example-image-a}
%% Use \caption command for figure caption and label.
\caption{Figure Caption}\label{fig1}
%% https://en.wikibooks.org/wiki/LaTeX/Importing_Graphics#Importing_external_graphics
\end{figure}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix
\section{Example Appendix Section}
\label{app1}

Appendix text.

%% For citations use:
%%       \citet{<label>} ==> Lamport (1994)
%%       \citep{<label>} ==> (Lamport, 1994)
%%
Example citation, See \citet{lamport94}.

%% If you have bib database file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-harv}
%%  \bibliography{<your bibdatabase>}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%% Refer following link for more details about bibliography and citations.
%% https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management

\begin{thebibliography}{00}

%% For authoryear reference style
%% \bibitem[Author(year)]{label}
%% Text of bibliographic item

\bibitem[Lamport(1994)]{lamport94}
  Leslie Lamport,
  \textit{\LaTeX: a document preparation system},
  Addison Wesley, Massachusetts,
  2nd edition,
  1994.

\end{thebibliography}
\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.

### 2024-09-19 manuscript_rmarkdown/manuscript.Rmd
Further work on transition to LaTeX.

@@
     latex_engine: lualatex
+    keep_tex: true
 papersize: a4

### 2024-09-19 manuscript_rmarkdown/manuscript.tex
Further work on transition to LaTeX.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{amsmath}
\usepackage{booktabs}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\sgn}{sgn}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Location and Scale-Invariant Power Transformations for Transforming Data to Normality},
  pdfauthor={Alex Zwanenburg, Steffen Löck},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Location and Scale-Invariant Power Transformations for
Transforming Data to Normality}
\author{Alex Zwanenburg, Steffen Löck}
\date{2024-09-19}

\begin{document}
\maketitle

\section{Abstract}\label{abstract}

To-do

\section{Introduction}\label{introduction}

Many statistical and some machine learning methods assume normality of
the underlying data, e.g.~analysis of variance and linear discriminant
analysis. However, numerical features in datasets may strongly deviate
from normal distributions, e.g.~by being skewed. Power transformations
aim to stabilise variance and improve normality of such features
(Bartlett 1947; John W. Tukey 1957). The two most commonly used
transformations are that of Box and Cox (1964) and Yeo and Johnson
(2000). The Box-Cox transformation of a feature value \(x_i\) of feature
\(\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}\) under the
transformation parameter \(\lambda\) is defined as:

\begin{equation}
\label{eqn:box-cox-original}
\phi_{\text{BC}}^\lambda (x_i) = 
\begin{cases}
\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x_i) & \text{if } \lambda = 0
\end{cases}
\end{equation}

One limitation of the Box-Cox transformation is that it is only defined
for \(x_i > 0\). In contrast, the Yeo-Johnson transformation under the
transformation parameter \(\lambda\) is defined for any
\(x_i \in \mathbb{R}\):

\begin{equation}
\label{eqn:yeo-johnson-original}
\phi_{\text{YJ}}^\lambda (x_i) = 
\begin{cases}
\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
\end{cases}
\end{equation}

The \(\lambda\)-parameter is typically optimised using maximum
likelihood estimation under the assumption that the transformed feature
is normally distributed. As noted by Raymaekers and Rousseeuw, this
approach is sensitive to outliers, and robust versions of Box-Cox and
Yeo-Johnson transformations were devised (Raymaekers and Rousseeuw
2021).

Applying a power transformation does not guarantee that transformed
features are normally distributed. Depending on location and scale of a
feature and the presence of outliers, power transformations may decrease
normality, as shown in Figure \ref{fig:decreased-normality}. If
normality of the transformed feature is not checked, e.g.~in automated
power transformation in machine learning workflows, several issues may
arise. For example, statistical tests such as ANOVA may produce
incorrect results due to violation of the normality assumption.
Likewise, machine learning methods that assume normality of input
features may suffer a decrease in performance. Moreover, large negative
or positive \(\lambda\)-parameters may lead to numeric issues in any
subsequent computations.

Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
and Wilk 1965), could be automatically applied to transformed features.
However, given sufficiently large sample sizes such tests can detect
trivial deviations from normality, and may lead to rejection of
sufficiently good power transformations.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/decreased-normality-1}

}

\caption{Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\lambda$. $10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the shift dataset, $\mathcal{N}(10, \sigma)$ for the scale dataset and $\mathcal{N}(0, 1)$ for the outlier dataset. Additionally, an outlier with value $d$ was added to the outlier dataset. Since samples are drawn from a normal distribution, a transformation parameter of $\lambda = 1$ is expected. However, a large shift in location, a scale that is small compared to the location, or presence of large outliers lead to incorrectly estimated transformation parameter values.}\label{fig:decreased-normality}
\end{figure}

To address these issues, we make the following contributions:

\begin{itemize}
\item
  We devise location- and scale-invariant versions of the Box-Cox and
  Yeo-Johnson transformation, including versions robust to outliers.
\item
  We derive the maximum likelihood criterion for location- and
  scale-invariant Box-Cox and Yeo-Johnson transformations to allow for
  optimising transformation parameters.
\item
  We define an empirical central normality test for detecting cases
  where power transformations fail to yield an approximately normally
  distributed transformed feature.
\item
  We assess the effect of power transformations on the performance of
  machine learning models.
\end{itemize}

\section{Theory}\label{theory}

In this section, we will first introduce location- and scale-invariant
versions of the Box-Cox and Yeo-Johnson transformations. Subsequently,
we define weighted location- and scale-invariant transformations and
weighting methods for robust transformations. We then define the
quantile function for asymmetric generalised normal distributions to
enable random sampling. Finally, we define the overall framework for the
empirical central normality test.

\subsection{Location- and scale-invariant power
transformation}\label{location--and-scale-invariant-power-transformation}

Box-Cox and Yeo-Johnson transformations are modified by introducing
shift parameter \(x_0\) and scale parameter \(s\) into equations
\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}. The
location- and scale-invariant Box-Cox transformation of a feature value
\(x_i\) of feature \(\mathbf{X}\) under transformation parameter
\(\lambda\), shift parameter \(x_0\) and scale parameter \(s\) is then:

\begin{equation}
\label{eqn:box-cox-invariant}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
\end{cases}
\end{equation}

where \(x_i - x_0 > 0\). Likewise, the location- and scale-invariant
Yeo-Johnson transformation of a feature value \(x_i\) under
transformation parameter \(\lambda\), shift parameter \(x_0\) and scale
parameter \(s\) is:

\begin{equation}
\label{eqn:yeo-johnson-invariant}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}

For both invariant transformations, \(\lambda\), \(x_0\) and \(s\)
parameters can be obtained by maximising the log-likelihood function,
i.e.~using maximum likelihood estimation (MLE). A full derivation of the
log-likelihood function for both transformations is shown in Appendix A.
The location- and scale-invariant Box-Cox log-likelihood function is:

\begin{equation}
\label{eqn:box-cox-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

subject to \(x_i - x_0 > 0\). \(\mu\) and \(\sigma^2\) are the mean and
variance of the Box-Cox transformed feature
\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood
function is:

\begin{equation}
\label{eqn:yeo-johnson-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

where \(\mu\) and \(\sigma^2\) are the mean and variance of the
Yeo-Johnson transformed feature
\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.

\subsection{Robust location- and scale-invariant power
transformations}\label{robust-location--and-scale-invariant-power-transformations}

Real-world data may contain outliers, to which maximum likelihood
estimation can be sensitive. Their presence may lead to poor
transformations to normality, as shown in Figure
\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
(2021), the general aim of power transformations should be to transform
non-outlier data to normality, i.e.~achieve \emph{central normality}. To
achieve this, they devised an iterative procedure to find a robust
estimate of the transformation parameter \(\lambda\). Briefly, this
process requires identifying outliers in the data and weighting such
instances during the optimisation process. Raymaekers and Rousseeuw
(2021) achieve this through weighted maximum likelihood estimation.
However, because this procedure iteratively estimates and updates
\(\lambda\), it can not be used here to simultaneously estimate
\(\lambda\), \(x_0\) and \(s\) for location- and scale-invariant power
transformations. Nonetheless, as a procedure, weighted MLE can be used
for estimating the transformation, shift and scale parameters.

Here, weighted maximum likelihood estimation is based on equations
\ref{eqn:box-cox-invariant-log-likelihood} and
\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
and Rousseeuw (2021), these log-likelihood functions includes additional
terms to accommodate estimation of \(x_0\) and \(s\). The weighted
location- and scale-invariant Box-Cox log-likelihood function is:

\begin{equation}
\label{eqn:box-cox-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Box-Cox transformed feature
\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\):

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

Analogously, the weighted location- and scale-invariant Yeo-Johnson
log-likelihood function is:

\begin{equation}
\label{eqn:yeo-johnson-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Yeo-Johnson transformed feature
\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

The weights \(w_i\) in equations
\ref{eqn:box-cox-weighted-invariant-log-likelihood} and
\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood} can be set using
several weighting functions. Using \(\dot{x}_i\) as an argument that
will be defined later, we investigate three weighting functions:

\begin{itemize}
\tightlist
\item
  A step function, with \(\delta_1 \geq 0\) as threshold parameter:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
0 & \text{if } \left| \dot{x}_i \right| > \delta_1
\end{cases}
\end{equation}

\begin{itemize}
\tightlist
\item
  A triangle function (or generalised Huber weight), with
  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
  parameters:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}

\begin{itemize}
\tightlist
\item
  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
  and \(\delta_2 \geq \delta_1\) as threshold parameters:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}

All weighting functions share the characteristic that for
\(\left| \dot{x}_i \right|< \delta_1\), instances are fully weighted,
i.e.~when \(\delta_1 > 0\) the weighting functions are symmetric window
functions with a flat top. The triangle and tapered cosine functions
then gradually down-weight instances with
\(\delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2\), and assign no
weight to instances \(\left| x_i \right| > \delta_2\). Examples of these
weighting function are shown in Figure \ref{fig:weighting-functions}.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/weighting-functions-1}

}

\caption{Weighting functions investigated in this study to make power transformations more robust against outliers. In this example, the step function was parameterised with $\delta_1 = 0.60$. The triangle and tapered cosine functions were both parameterised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
\end{figure}

Each weighting function has an argument \(\dot{x}\) that is related to
the (transformed) feature in one of several ways:

\begin{itemize}
\item
  The weighting function uses empirical probabilities of the
  distribution of the original feature \(\mathbf{X}\). After sorting
  \(\mathbf{X}\) in ascending order, probabilities are determined as
  \(p_i = \frac{i - 1/3}{n + 1/3}\), with \(i = 1, 2, \ldots n\), with
  \(n\) the number of instances of feature \(\mathbf{X}\). Then
  \(\dot{x}_i = p^{*}_i=2 \left( p_i - 0.5\right)\), so that argument is
  zero-centered.
\item
  The weighting function uses the z-score of the transformed feature
  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
  Rousseeuw (2021),
  \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
  \(\mu_M\) and \(\sigma_M\) are robust Huber M-estimates of location
  and scale of the transformed feature
  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981). Then
  \(\dot{x}_i = z_i\).
\item
  After sorting \(\mathbf{X}\) in ascending order, the weighting
  function uses the residual error between the z-score of the
  transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
  theoretical z-score from a standard normal distribution:
  \(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
  with \(\mu_M\), \(\sigma_M\) and \(p_i\) as defined above. Then
  \(\dot{x}_i = r_i\).
\end{itemize}

\subsection{Asymmetric generalised normal
distributions}\label{asymmetric-generalised-normal-distributions}

Modifications intended to make power transformations invariant to
location and scale of a feature and methods to improve their robustness
against outliers need to be assessed using data drawn from a range of
different distributions. Since the power transformations are intended
for use with unimodal distributions, the generalised normal distribution
(Subbotin 1923; Nadarajah 2005) is a suitable option for simulating
realistic feature distributions. This distribution has the following
probability density function \(f_{\beta}\) for a value
\(x \in \mathbb{R}\):

\begin{equation}
f_{\beta}(x) = \frac{\beta}{2\Gamma\left(1 / \beta \right)} e^{-\left| x \right|^\beta}
\end{equation}

Here, \(\Gamma\) is the gamma function, and \(\beta\) is a strictly
positive shape parameter. For \(\beta = 1\), the probability density
function describes a Laplace distribution. A normal distribution is
found for \(\beta=2\), and for large \(\beta\), the distribution
approaches a uniform distribution. We will refrain from introducing
scale and location parameters here directly.

Realistic feature distributions may be skewed. Gijbels et al.~describe a
recipe for introducing skewness into the otherwise symmetric generalised
normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
the following probability density function:

\begin{equation}
f_{\alpha}(x; \mu, \sigma, \beta) = \frac{2 \alpha \left(1 - \alpha\right)}{\sigma}
\begin{cases}
f_{\beta}\left( \left(1 - \alpha \right) \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x \leq \mu \\
f_{\beta}\left( \alpha \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x > \mu
\end{cases}
\end{equation}

Here, \(\alpha \in (0,1)\) is a skewness parameter. \(\alpha > 0.5\)
creates a distribution with a negative skew, i.e.~a left-skewed
distribution. A right-skewed distribution is created for
\(\alpha < 0.5\). \(\mu \in \mathcal{R}\) and \(\sigma \in (0, \infty)\)
are location and scale parameters, respectively. \(f_{\alpha}\) thus
describes the probability density function of an asymmetric generalised
normal distribution, which we will refer to here and parametrise as
\(\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)\).

We require a quantile function (or an approximation thereof) to draw
random values from an asymmetric generalised normal distribution using
inverse transform sampling. Gijbels et al.~derived the following
quantile function \(F_{\alpha}^{-1}(p)\):

\begin{equation}
F_{\alpha}^{-1}(p; \mu, \sigma, \beta) =
\begin{cases}
\mu + \frac{\sigma}{1 - \alpha} F_{\beta}^{-1} \left( \frac{p}{2 \alpha}\right) & \text{, } p \leq \alpha \\
\mu + \frac{\sigma}{\alpha} F_{\beta}^{-1} \left( \frac{1 + p - 2 \alpha}{2 \left(1 - \alpha \right)} \right) & \text{, } p > \alpha
\end{cases}
\end{equation}

The quantile function for the asymmetric generalised normal distribution
\(F_{\alpha}^{-1}\) thus incorporates the quantile function
\(F_{\beta}^{-1}\) of the symmetric generalised normal distribution.
\(F_{\beta}^{-1}\) was derived by Griffin to be (Griffin 2018):

\begin{equation}
F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
\end{equation}

Here, \(F_{\Gamma}^{-1}\) is the quantile function of the gamma
distribution with shape \(1 / \beta\), which can be numerically
approximated.

\subsection{Empirical central normality
test}\label{empirical-central-normality-test}

Power transformations aim to transform features to a normal
distribution. However, this may not always be successful or possible.
Deviations from normality can be detected by normality tests, such as
the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
tests may be too stringent with large sample sizes, outliers, or both.
Here we develop an empirical test for central normality. The null
hypothesis \(H_0\) is that the distribution is centrally normal. The
alternative hypothesis \(H_1\) is that the distribution is not centrally
normal.

Let central normality be defined as the normality of central portion
\(\kappa\) of the data,
i.e.~\(\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\),
with \(p_i\) probabilities of the empirical distribution, as previously.
We then compute the residual errors between the z-scores of the
transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
expected z-scores from a standard normal distribution:
\(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
with \(\mu_M\) and \(\sigma_M\) robust Huber M-estimates of location and
scale of the transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
(Huber 1981). The set of residual errors for the central portion of the
data is then
\(\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\).

The test statistic \(\tau_{\text{ecn}}\) is then defined as:

\begin{equation}
\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
\end{equation}

Here \([\quad]\) denotes an Iverson bracket. The test statistic is equal
to the mean of the residual errors of the central portion of the data.

To use the test statistic, the central portion of the data needs to be
defined and the Type 1 error rates determined. We will do so in the
\hyperref[simulation]{Simulation} section.

\section{Simulation}\label{simulation}

We used simulated data to assess invariance to location and scale of the
proposed power transformations, weighting for robust transformations,
and to develop the empirical central normality test. The \(\lambda\)
parameter for conventional power transformations (Eqn.
\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well
as \(\lambda\), \(x_0\) and \(s\) parameters for location- and
scale-invariant power transformations (Eqn. \ref{eqn:box-cox-invariant}
and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
algorithm for derivative-free bound constraint optimisation (Powell
2009) through maximum likelihood estimation. The required algorithms
were implemented in the \texttt{power.transform} R software package
(Zwanenburg and Löck 2024b) (version 1.0.0). Of note, the
\texttt{power.transform} package shifts feature values into the positive
domain if negative or zero values are present for Box-Cox power
transformations.

\subsection{Invariance to location and
scale}\label{invariance-to-location-and-scale}

To assess whether the proposed power transformations lead to values of
\(\lambda\) that are invariant to location and scale of the
distribution, we simulated three different distributions. We first
randomly drew \(10000\) values from a normal distribution:
\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
or equivalently
\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)\).
The second distribution was a right-skewed generalised normal
distribution
\(\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)\).
The third distribution was a left-skewed generalised normal distribution
\(\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)\).
We then computed transformation parameter \(\lambda\) using the original
definitions (Eqn. \ref{eqn:box-cox-original} and
\ref{eqn:yeo-johnson-original}) and the location- and scale-invariant
definitions (Eqn. \ref{eqn:box-cox-invariant} and
\ref{eqn:yeo-johnson-invariant}) for each distribution. To assess
location invariance, a positive value \(d_{\text{shift}}\) was added to
each distribution with \(d_{\text{shift}} \in [1, 10^6]\). Similarly, to
assess scale invariance, each distribution was multiplied by a positive
value \(d_{\text{scale}}\), where \(d_{\text{scale}} \in [1, 10^6]\).

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/shifted-distributions-1}

}

\caption{Invariant power transformation produces transformation parameters that are invariant to location and scale. Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\text{shift}}$ or multiplication by $d_{\text{scale}}$. Estimates of the transformation parameter $\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution, whereas estimates obtained for the location- and scale-invariant power transformations are constant.}\label{fig:shifted-distributions}
\end{figure}

The result is shown in Figure \ref{fig:shifted-distributions}. For each
distribution, transformation parameter \(\lambda\) varied with
\(d_{\text{shift}}\) and \(d_{\text{scale}}\) when estimated for
conventional transformations. In contrast, estimation of \(\lambda\) for
invariant power transformations was invariant to both
\(d_{\text{shift}}\) and \(d_{\text{scale}}\).

\subsection{Robust transformations}\label{robust-transformations}

Outliers may be present in data and affect estimation of transformation
parameters. The log-likelihood function can be weighted to assign less
weight to outlier instances, see equations
\ref{eqn:box-cox-weighted-invariant-log-likelihood} and
\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}. We proposed
three weighting function: step, triangle and tapered cosine, that have
one, two and two parameters, respectively. Each weighting function then
takes one of three values as input: probabilities of the empirical
distribution of the original feature, the z-score of the transformed
feature values, or the residual error between the z-score of the
transformed feature values and their expected z-score based on the
normal distribution.

To determine the weighting function parameters for each of the nine
combinations, \(m_d=100\) asymmetric generalised normal distributions
were drawn. Each distribution was parametrised with a randomly chosen
skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
parameter \(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale
parameters were set as \(\mu = 0\) and \(\sigma = 1\), respectively.
\(n = \lceil 10^\gamma \rceil\) instances were then randomly drawn, with
\(\gamma \sim U\left(2, 4\right)\), i.e., between \(100\) and \(10000\)
values are drawn to create \(\mathbf{X}_i\).

Outlier values were then drawn to randomly replace a fraction of the
values of \(\mathbf{X}_i\). This was repeated \(m_{\text{out}} = 10\)
times, with outlier fractions regularly spaced in \([0.00, 0.10]\). Thus
up to 10 percent of the values could be replaced by outliers. Outlier
values were set according to John Wilder Tukey (1977), as follows. Let
\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
was:

\begin{equation}
x_{out} =
\begin{cases}
Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
\end{cases}
\end{equation}

\(Q_1\), \(Q_3\) and \(\text{IQR}\) are the first quartile, third
quartile and interquartile range of \(\mathbf{X}_i\), respectively.
Outlier values randomly replaced values in \(\mathbf{X}_i\) to create
\(\mathbf{X}_{i,j}\), with
\(j \in \left\{ 1, \ldots, m_{\text{out}} \right\}\)

To find the optimal values for the weighting function parameters
\(\delta_1\) and \(\delta_2\) (if applicable), we minimised the absolute
difference between the \(\lambda_{r}\) parameter obtained for robust
transformation in the presence of outliers, and the \(\lambda_0\)
parameter obtained using the non-robust transformations in absence of
outliers:

\begin{equation}
\label{eqn:minimisation-weighting}
\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
\end{equation}

Minimisation was conducted using the BOBYQA algorithm for
derivative-free bound constraint optimisation (Powell 2009). The
resulting weighting function parameters for weighted MLE are shown in
Tables \ref{tab:optimal-weighting-parameters-box-cox} and
\ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location-
and scale-invariant Box-Cox and Yeo-Johnson transformations,
respectively.

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations. $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. The {optimal} column shows the optimal parameter values. The \textit{loss} column shows the loss achieved by each method, under optimised parameters. Lower loss indicates better robustness against outliers.}
\label{tab:optimal-weighting-parameters-box-cox}
\begin{tabular}{l r r r r r r r}

\toprule
method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
& initial & limits & optimal & initial & limits & optimal & \\

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations. $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. The {optimal} column shows the optimal parameter values. The \textit{loss} column shows the loss achieved by each method, under optimised parameters. Lower loss indicates better robustness against outliers.}
\label{tab:optimal-weighting-parameters-yeo-johnson}
\begin{tabular}{l r r r r r r r}

\toprule
method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
& initial & limits & optimal & initial & limits & optimal & \\

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

Figure \ref{fig:optimised-weighting-function-parameters} shows the
distribution of errors \(\left| \lambda_r - \lambda_0 \right|\) for
non-robust and robust transformations using the optimal weighting
function parameters. In the presence of outliers, the Yeo-Johnson
transformation yielded smaller errors than the Box-Cox transformation.
For both transformations, weighting based on empirical probabilities
yielded the most consistent \(\lambda\) parameter estimates in the
presence of outliers. Other methods did not outperform the non-robust
transformation method.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/optimised-weighting-function-parameters-1}

}

\caption{Robustness of power transformations after optimising weighting function parameters. The distribution of errors, i.e. the difference between robustly fitted $\lambda^r$ and expected $\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly parametrised asymmetric generalised normal distributions with up to 10\% outliers on either side of the distribution. The top and bottom panel show errors for the location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively. In each panel, the error distribution for the non-robust transformation is shown to the left for comparison with different weighting methods. Note that a square root transformation was applied along the $y$-axis for display purposes. Median errors are shown above each distribution.}\label{fig:optimised-weighting-function-parameters}
\end{figure}

\subsection{Empirical central normality
test}\label{empirical-central-normality-test-1}

To develop an empirical test for central normality we need to consider
two parameters: the central portion \(\kappa\) as a fixed parameter, and
test statistic \(\tau_{\text{ecn}}\). We will first define the central
portion \(\kappa\).

First we draw \(m_d=10000\) random asymmetric generalised normal
distributions. As before, each distribution is parametrised with a
randomly chosen skewness parameter
\(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
are set as \(\mu = 0\) and \(\sigma = 1\), respectively.
\(n = \lceil 10^\gamma \rceil\) values are then randomly drawn, with
\(\gamma \sim U\left(1.47, 3.00\right)\), which leads to between \(30\)
and \(1000\) values being drawn to create \(\mathbf{X}_i\). As before,
for each distribution, up to \(10 \%\) of instances are replaced by
outlier values, and this is repeated \(10\) times. We then compute
residuals after optimising robust location- and scale-invariant
transformations with the empirical tapered cosine weighting method.

Figure \ref{fig:empirical-central-normality-residual-error} shows the
residual errors of the transformed distribution as function of the
empirical probability. As expected, the largest deviations from
normality appear on the extremities of this range. For Box-Cox
transformations, the effect of outliers on the lower end of the
distribution leads to noticeable asymmetry. Between very low and very
high empirical probabilities, residual errors are constrained and
relatively flat. This is the candidate range for the central portion of
the data.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-residual-error-1}

}

\caption{Residual error as a function of empirical probability. Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric generalised normal distributions, for each of which outliers were randomly drawn 10 times (up to 10\% of samples). Larger errors occur at the edges of each distribution.}\label{fig:empirical-central-normality-residual-error}
\end{figure}

We then consider the empirical probability of type I errors: the
probability of incorrectly classifying a centrally normal distribution
as being non-normal based on the test statistic \(\tau_{\text{ecn}}\).
Under the assumption that the asymmetric generalised normal
distributions are centrally normal after robust transformation, this
produces the relationship shown in Figure
\ref{fig:empirical-central-normality-type-1-error-rate}. Since for
\(\kappa \leq 0.80\) error curves for both transformation methods are
similar, we fix the value of the central portion \(\kappa\) to 0.80. In
the remaining we will use the empirical central normality test statistic
values defined using robust location- and scale-invariant Yeo-Johnson
transformations, as it is more conservative (see Appendix E). This leads
to test statistic values listed in Table
\ref{tab:empirical-central-normality}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-type-1-error-rate-1}

}

\caption{Type 1 error rate of transformed asymmetric generalised normal distributions as function of the test statistic $\tau_{\text{ecn}}$ for the central portion $\kappa$ of the distribution.}\label{fig:empirical-central-normality-type-1-error-rate}
\end{figure}

\begin{table}
\begin{center}
\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
\label{tab:empirical-central-normality}
\begin{tabular}{l | c c c c c c c}

\toprule
type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

In Figure \ref{fig:empirical-central-normality-examples} we apply the
empirical central normality test to assess central normality of features
that are composed of a mixture of samples drawn from two normal
distributions. With increased separation of the underlying normal
distributions, the probability of the feature being centrally normal
decreases, as expected.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-examples-1}

}

\caption{Bi-modal distributions and empirical central normality test results. The feature (black) is a mixture of two identical sample sets (blue and orange) drawn from normal distributions that are offset by a distance $d$. We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal. As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases. Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
\end{figure}

\section{Experimental Results}\label{experimental-results}

\subsection{Invariance}\label{invariance}

Location- and scale-invariant power transformations are intended to
yield improved transformations to normality in the presence of large
shifts in location, distributions that due to location and scale are not
centered near zero, or both. Earlier, we assessed these transformations
using simulated data. In the following, they are evaluated using
examples from real datasets. We focus on the Yeo-Johnson transformation
because of its ability to handle features with negative values. Results
for Box-Cox transformations are shown in Appendix D.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-invariance-1}

}

\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row). Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Yeo-Johnson transformations (right column).}\label{fig:experimental-results-invariance}
\end{figure}

\subsubsection{Age of patients with lung
cancer}\label{age-of-patients-with-lung-cancer}

A common feature in health-related datasets is age. Here we use data on
228 patients with lung cancer that was collected and published by
Loprinzi et al. (Loprinzi et al. 1994). The age in the cohort was
\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
conventional and invariant Yeo-Johnson transformations to patient age
yielded the following results, see Figure
\ref{fig:experimental-results-invariance}: no transformation (sum of
residuals with normal distribution \(\sum r_i = 16.5\)); conventional
transformation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\));
Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 2.0\),
\(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
\(\sigma_{YJ} = 0.5 \cdot 10^3\)); location- and scale-invariant
transformation (\(\lambda = 0.9\), \(\sum r_i = 8.8\),
\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.1\)); and robust location- and
scale-invariant transformation (\(\lambda = 0.8\), \(\sum r_i = 10.6\),
\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.0\)).

Location- and scale-invariant transformation led to a lower overall
residual sum, indicating a better transformation. Robust location- and
scale-invariant transformation had a higher residual sum compared to the
non-robust variant, which may be due to the lack of outliers in the
data. Conventional transformations inflated the mean \(\mu_{YJ}\) and
standard deviation \(\sigma_{YJ}\) of the age feature after
transformation. The empirical central normality test did not detect any
statistically significant deviations from central normality for any
transformation (all \(p \geq 0.78\)).

\subsubsection{Penguin body mass}\label{penguin-body-mass}

Gorman, Williams and Fraser recorded body mass (in grams) of 342
penguins of three different species (Gorman, Williams, and Fraser 2014).
The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
deviation) grams, and not centrally normal (\(p = 0.03\)). Applying
conventional and invariant Yeo-Johnson transformations to body mass
yielded the following results, see Figure
\ref{fig:experimental-results-invariance}: no transformation (residual
sum \(\sum r_i = 48.0\)); conventional transformation
(\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); Raymaekers and Rousseeuw's robust
adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); location- and scale-invariant
transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\)); and robust location- and
scale-invariant transformation (\(\lambda = 0.4\), \(\sum r_i = 23.1\),
\(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.8\)).

Location- and scale-invariant transformation produced a lower overall
residual sum, indicating a better transformation. Moreover, conventional
transformations led to low standard deviation \(\sigma_{YJ}\) of the
body mass feature after transformation. The empirical central normality
test did not detect any statistically significant deviations from
central normality for any transformation (all \(p \geq 0.38\)).

\subsubsection{Latitude in the Ames housing
dataset}\label{latitude-in-the-ames-housing-dataset}

Geospatial datasets usually contain coordinates. The Ames housing
dataset contains data on 2930 properties that were sold between 2006 and
2010 (De Cock 2011), including their geospatial coordinates. The
latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
conventional and invariant Yeo-Johnson transformations to latitude
yielded the following results, see Figure
\ref{fig:experimental-results-invariance}: no transformation (residual
sum \(\sum r_i = 328\)); conventional transformation
(\(\lambda = 62.1\), \(\sum r_i = 319\),
\(\mu_{YJ} = 4.8 \cdot 10^{99}\), \(\sigma_{YJ} = 0.1 \cdot 10^{99}\));
Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 95.4\),
\(\sum r_i = 319\), \(\mu_{YJ} = 6.4 \cdot 10^{153}\),
\(\sigma_{YJ} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
transformation (\(\lambda = 1.5\), \(\sum r_i = 326\),
\(\mu_{YJ} = -1.2\), \(\sigma_{YJ} = 0.8\)); and robust location- and
scale-invariant transformation (\(\lambda = 1.4\), \(\sum r_i = 311\),
\(\mu_{YJ} = -1.3\), \(\sigma_{YJ} = 0.9\)).

Every transformation reduced the residual sum. The non-robust location-
and scale-invariant transformation did not improve over conventional
alternatives and yielded a data distribution lacking central normality
(empirical central normality test: \(p=0.05\)). However, conventional
transformations had high values for the \(\lambda\) parameter, which
could lead to numerical issues.

\subsection{Robustness against
outliers}\label{robustness-against-outliers}

We previously simulated data to assess invariant power transformations
and their robustness against outliers. Here, we assess invariant power
transformations in real data with outliers.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-outlier-robustness-1}

}

\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row). Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Yeo-Johnson transformations (right column). Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness}
\end{figure}

\subsubsection{Fuel efficiency in the Top Gear
dataset}\label{fuel-efficiency-in-the-top-gear-dataset}

The Top Gear dataset contains data on 297 vehicles that appeared on the
BBC television show \emph{Top Gear} (Alfons 2021). Within this dataset,
the fuel consumption feature contains outliers due to highly
fuel-efficient vehicles. Applying conventional and invariant Yeo-Johnson
transformations to the fuel consumption feature yielded the following
results, see Figure \ref{fig:experimental-results-outlier-robustness}:
no transformation (residual sum \(\sum r_i = 54\), \(p=0.76\));
conventional transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
\(\mu_{YJ} = 3.0\), \(\sigma_{YJ} = 0.3\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.55\)); location- and
scale-invariant transformation (\(\lambda = -1.3\), \(\sum r_i = 44\),
\(\mu_{YJ} = 0.5\), \(\sigma_{YJ} = 0.1\), \(p=0.03\)); and robust
location- and scale-invariant transformation (\(\lambda = -0.9\),
\(\sum r_i = 50\), \(\mu_{YJ} = 2.0\), \(\sigma_{YJ} = 2.3\),
\(p=0.58\)).

Outliers cause non-robust transformations to fail to transform the data
to a centrally normal distribution (empirical central normality test
\(p = 0.01\) and \(p=0.03\) for conventional and invariant
transformations, respectively). Robust transformations produce
distributions that are centrally normal (empirical central normality
test \(p > 0.05\)).

\subsubsection{Maximum arterial wall thickness in an ischemic stroke
dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}

The ischemic stroke dataset contains historic data from 126 patients
with risk at ischemic stroke (Kuhn and Johnson 2019). These patients
underwent Computed Tomography Angiography to characterize the carotid
artery blockages. Angiography imaging was then assessed, and various
characteristics related to the blood vessels and the disease are
measured. The maximum arterial wall thickness feature contains several
instances with outlier values. Applying conventional and invariant
Yeo-Johnson transformations to this feature yielded the following
results, see Figure \ref{fig:experimental-results-outlier-robustness}:
no transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
conventional transformation (\(\lambda = -0.7\), \(\sum r_i = 30\),
\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 136\),
\(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14\), \(p=0.61\)); location- and
scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 12\),
\(\mu_{YJ} = -11.8\), \(\sigma_{YJ} = 6.9\), \(p=0.13\)); and robust
location- and scale-invariant transformation (\(\lambda = -0.6\),
\(\sum r_i = 27\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.1\),
\(p=0.10\)).

Non-robust transformations failed to produce a centrally normal
distribution (empirical central normality test \(p=0.01\) and \(p=0.02\)
for conventional and invariant transformations, respectively). Robust
transformations produce distributions that are centrally normal
(empirical central normality test \(p > 0.05\)).

\subsection{Integration into end-to-end machine
learning}\label{integration-into-end-to-end-machine-learning}

We used 285 datasets from the Penn Machine Learning Benchmarks
collection (Romano et al. 2022). In this collection, 122 datasets
correspond to regression tasks and 163 datasets to classification tasks.
Using the familiar auto-machine learning library (Zwanenburg and Löck
2024a) (version 1.5.0), each dataset was used to train a model for each
of 16 process configurations. Each process configuration specifies the
learner (generalised linear model or random forest), transformation
method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson,
robust invariant Yeo-Johnson with empirical central normality test
(rejecting transformations with \(p \leq 0.01\)), and normalisation
method (none, \(z\)-standardisation), yielding 16 distinct
configurations. Before each experiment, each dataset was randomly split
into a training (70\%) and holdout test (30\%) set five times. Thus, a
total of 22800 models were created. Each model was then evaluated using
the holdout test set using one of two metrics, i.e.~the root relative
squared error (RRSE) for regression tasks and the area under the
receiver operating characteristic curve (AUC) for classification tasks.

For the purpose of assessing the effect of the difficulty of the task,
we computed the median performance score over all models for each
dataset and assigned one the following categories:

\begin{itemize}
\tightlist
\item
  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (87
  datasets)
\item
  easy: \(0.90 > \text{AUC} \geq 0.80\) or
  \(0.30 \geq \text{RRSE} > 0.10\) (46 datasets)
\item
  intermediate: \(0.80 > \text{AUC} \geq 0.70\) or
  \(0.60 \geq \text{RRSE} > 0.30\) (72 datasets)
\item
  difficult: \(0.70 > \text{AUC} \geq 0.60\) or
  \(0.80 \geq \text{RRSE} > 0.60\) (60 datasets)
\item
  very difficult: \(0.60 > \text{AUC} \geq 0.50\) or
  \(1.00 \geq \text{RRSE} > 0.80\) (12 datasets)
\item
  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (8
  datasets)
\end{itemize}

To remove the effect of the dataset, and allow for comparing metrics, we
ranked all performance scores for each dataset so that a higher rank
corresponds to better performance. Experiments yielding the same score
received the same, average, rank. Subsequently ranks were normalised to
the \([0.0, 1.0]\) range.

Significant differences exist between process configurations (Friedman
test: \(p < 10^{-8})\).

Here we focus on the subset of 232 datasets that contain numeric
features. Considering single process parameters, the choice of learner
(Wilcoxon signed rank test: \(p < 10^{-8})\)) and normalisation method
(Wilcoxon signed rank test: \(p = 0.007\)) had a significant impact (at
significance level \(p = 0.050\)), but transformation method (Friedman
test: \(p = 0.054\)) did not.

To estimate the marginal effects of process parameters, including
transformation method, we first fit a regression random forest (ranger
package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, minimum
node size 2, other hyperparameters default) with process parameters and
task difficulty as predictors and normalised rank as response variable.
The estimated marginal effects are shown in Figure
\ref{fig:marginal-effect-plot}. On the scale of normalised ranks
(\([0.0, 1.0]\)), the overall estimated marginal effect of using a
random forest instead of generalised linear model was \(0.272\). The
overall marginal effect of using z-standardisation to normalise features
was \(0.008\). Transformation methods had the following marginal
effects: \(0.003\) for using conventional Yeo-Johnson transformation
instead of no transformation; \(-0.007\) for using robust invariant
Yeo-Johnson transformation instead of no transformation; \(-0.009\) for
using robust invariant Yeo-Johnson transformation with empirical central
normality test instead of no transformation; \(-0.010\) for using robust
invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson
transformation; and \(-0.012\) for using using robust invariant
Yeo-Johnson transformation with empirical central normality test instead
of conventional Yeo-Johnson transformation.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/marginal-effect-plot-1}

}

\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets. The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). Random forests outperform GLM models for all task difficulties. The top-right panel shows the marginal effect of feature normalisation methods, i.e. no normalisation and z-standardisation. z-standardisation is generally beneficial, but the estimated effect is neglible. The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method. There is no consistent behaviour, and estimated effects are neglible. Note that the ranges of the $x$-axes of the three main panels differ. ECNT: empirical central normality test.}\label{fig:marginal-effect-plot}
\end{figure}

\section{Discussion}\label{discussion}

In their work on power transformation, Box and Cox already mention
transformation with a shift parameter, but preferred the version in Eq.
\ref{eqn:box-cox-original} for the theoretical analysis in their paper
(Box and Cox 1964), which subsequently became the convention. Yeo and
Johnson's power transformation lacks a shift parameter altogether (Yeo
and Johnson 2000). We showed that these power transformations are
sensitive to location and scale of data distributions. To mitigate this
issue, we defined location- and scale-invariant variants of the Box-Cox
and Yeo-Johnson transformations. We furthermore assessed methods for
making these transformations robust to outliers, and devised an
empirical test for central normality.

Robust location- and scale-invariant transformations are a suitable
replacement for their conventional counterparts. They demonstrated
robustness against outliers and prevent inaccurate transformations and
potential numerical issues due to location and scale of the distribution
of a feature. This is particularly relevant for automated data
processing, where such issues may go unnoticed. Compared to non-robust
location- and scale-invariant transformations, real-world examples with
outliers showed higher residual errors of transformed features. Robust
transformations seek to minimise residual errors for the central part of
the distribution, instead of the entire distribution, including
outliers. The empirical central normality test indicated that robust
transformations are better able to achieve central normality in the
presence of outliers. However, in a machine learning experiment of 232
real-world datasets that contained at least one numeric feature, we did
not find a meaningful benefit of for -- nor detriment to -- model
performance for location- and scale-invariant power transformations. One
reason may be that numeric features with large location shifts
(\(|\mu| > 1000.0\)) were uncommon. Of the 4886 numeric features in the
232 datasets, 266 (5\%) features in 34 datasets had large location
shifts, of which 200 appeared in just 2 datasets. For the latter two
datasets, the transformation method does not show significant difference
between groups (Friedman test; \(p > 0.05\)).

Location- and scale-invariant transformations are realised by
simultaneously optimising three parameters, i.e.~transformation
parameter \(\lambda\), shift parameter \(x_0\) and scale parameter
\(s\). We derived the log-likelihood function to facilitate optimisation
using MLE. Alternatively, standardisation of a numeric variable (e.g.,
through subtracting its median value and division by its interquartile
range) prior to conventional power transformations may achieve a similar
effect in reducing sensitivity to the distribution's location and scale.
While this alternative helps prevent these issues -- provided that
normalisation does not lead to negative values for Box-Cox
transformation -- location- and scale-invariant transformations seem to
provide an overall better transformation to normality (Appendix F).

We assessed several methods for robust power transformation. Methods
that relied on the z-score of the transformed feature or the residual
error yielded worse results than the non-robust method. This is partly
due to the initial choice of threshold parameters. Using different
initial values, closer to the upper limits
(\(\delta_1 = 10.0; \delta_2 = 10.0\)), led to a reduced loss. However,
these values effectively correspond to a non-robust transformation,
where all instances receive the same weight. Underperformance of these
methods could be explained by their use of transformed features for
setting weights. Consequently, the weights change at each iteration in
the MLE optimisation process. This increases local variance in the
log-likelihood function and create local optima that the optimiser may
not handle well. As a consequence, optimal values for transformation
parameter \(\lambda\) might differ, which increases the presented loss
for optimising threshold parameters. Methods that relied on the
empirical probability did not suffer from this issue, as weights
remained fixed during MLE.

We introduced an empirical test for central normality to assess whether
distributions deviate from normality in a way that might require closer
inspection prior to further processing. The empirical test for central
normality differs from other tests for normality, such as the
Shapiro-Wilk test (Shapiro and Wilk 1965), in that the test statistic is
independent of the number of samples. This makes this test more
practical for assessing central normality for larger sample numbers,
where other test may detect inconsequential deviations from normality.

This work has the following limitation. Firstly, we did observe several
numerical stability issues for optimisation criteria other than MLE
(Appendix B). These appear in regions where transformation parameters
would lead to very large or small numbers when using conventional power
transformations. For MLE stability issues were not observed. Secondly,
the empirical central normality test is based on simulations instead of
statistical theory, and relies on a somewhat arbitrary definition of the
central portion of a distribution. Thus, while the test may asses
whether data is sufficiently normally distributed for practical
purposes, it should not be used as a strict test for normality.

\section{Conclusion}\label{conclusion}

Compared to their conventional versions, robust location- and
scale-invariant Box-Cox and Yeo-Johnson transformations reduce
sensitivity to outliers and the location and scale of features. An
empirical central normality test can assess the quality of
transformation of features to normal distributions. The combination of
both facilitate the use of power transformations in automated data
analysis workflows.

\section{Data and code availability}\label{data-and-code-availability}

Location- and scale-invariant power transformations were implemented in
the \texttt{power.transform} package for R, which is available from
\href{https://github.com/oncoray/power.transform}{GitHub} and the
\href{https://cran.r-project.org/package=power.transform}{CRAN
repository}. The manuscript was created using R Markdown and is likewise
available from the \texttt{power.transform} GitHub repository. Data and
results for the machine learning experiment are separately available
from \href{https://doi.org/10.5281/zenodo.13736671}{Zenodo}.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Alfons2021-kc}
Alfons, Andreas. 2021. {``{robustHD}: An {R} Package for Robust
Regression with High-Dimensional Data.''} \emph{J. Open Source Softw.} 6
(67): 3786. \url{https://doi.org/10.21105/joss.03786}.

\bibitem[\citeproctext]{ref-Bartlett1947-rx}
Bartlett, M S. 1947. {``The Use of Transformations.''} \emph{Biometrics}
3 (1): 39--52. \url{https://doi.org/10.2307/3001536}.

\bibitem[\citeproctext]{ref-Box1964-mz}
Box, G E P, and D R Cox. 1964. {``An Analysis of Transformations.''}
\emph{J. R. Stat. Soc. Series B Stat. Methodol.} 26 (2): 211--52.

\bibitem[\citeproctext]{ref-De-Cock2011-jf}
De Cock, Dean. 2011. {``Ames, Iowa: Alternative to the Boston Housing
Data as an End of Semester Regression Project.''} \emph{J. Stat. Educ.}
19 (3). \url{https://doi.org/10.1080/10691898.2011.11889627}.

\bibitem[\citeproctext]{ref-Gijbels2019-te}
Gijbels, Irène, Rezaul Karim, and Anneleen Verhasselt. 2019. {``Quantile
Estimation in a Generalized Asymmetric Distributional Setting.''} In
\emph{Stochastic Models, Statistics and Their Applications}, 13--40.
Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-28665-1/_2}.

\bibitem[\citeproctext]{ref-Gorman2014-eo}
Gorman, Kristen B, Tony D Williams, and William R Fraser. 2014.
{``Ecological Sexual Dimorphism and Environmental Variability Within a
Community of Antarctic Penguins (Genus Pygoscelis).''} \emph{PLoS One} 9
(3): e90081. \url{https://doi.org/10.1371/journal.pone.0090081}.

\bibitem[\citeproctext]{ref-Griffin2018-bf}
Griffin, Maryclare. 2018. \emph{Gnorm: Generalized Normal/Exponential
Power Distribution}. \url{https://CRAN.R-project.org/package=gnorm}.

\bibitem[\citeproctext]{ref-Huber1981-su}
Huber, Peter J. 1981. \emph{Robust Statistics}. John Wiley \& Sons.
\url{https://doi.org/10.1002/0471725250}.

\bibitem[\citeproctext]{ref-Kuhn2019-kt}
Kuhn, Max, and Kjell Johnson. 2019. \emph{Feature Engineering and
Selection: A Practical Approach for Predictive Models}. Chapman \&
Hall/CRC Data Science Series. Chapman; Hall/CRC.
\url{https://doi.org/10.1201/9781315108230}.

\bibitem[\citeproctext]{ref-Loprinzi1994-cd}
Loprinzi, C L, J A Laurie, H S Wieand, J E Krook, P J Novotny, J W
Kugler, J Bartel, M Law, M Bateman, and N E Klatt. 1994. {``Prospective
Evaluation of Prognostic Variables from Patient-Completed
Questionnaires. North Central Cancer Treatment Group.''} \emph{J. Clin.
Oncol.} 12 (3): 601--7. \url{https://doi.org/10.1200/JCO.1994.12.3.601}.

\bibitem[\citeproctext]{ref-Nadarajah2005-xe}
Nadarajah, Saralees. 2005. {``A Generalized Normal Distribution.''}
\emph{J. Appl. Stat.} 32 (7): 685--94.
\url{https://doi.org/10.1080/02664760500079464}.

\bibitem[\citeproctext]{ref-Powell2009-zb}
Powell, M J D. 2009. {``The {BOBYQA} Algorithm for Bound Constrained
Optimization Without Derivatives.''} Cambridge NA Report NA2009/06.
University of Cambridge.

\bibitem[\citeproctext]{ref-Raymaekers2021-kq}
Raymaekers, Jakob, and Peter J Rousseeuw. 2021. {``Transforming
Variables to Central Normality.''} \emph{Mach. Learn.}, March.
\url{https://doi.org/10.1007/s10994-021-05960-5}.

\bibitem[\citeproctext]{ref-Romano2022-gq}
Romano, Joseph D, Trang T Le, William La Cava, John T Gregg, Daniel J
Goldberg, Praneel Chakraborty, Natasha L Ray, Daniel Himmelstein,
Weixuan Fu, and Jason H Moore. 2022. {``{PMLB} V1.0: An Open-Source
Dataset Collection for Benchmarking Machine Learning Methods.''}
\emph{Bioinformatics} 38 (3): 878--80.
\url{https://doi.org/10.1093/bioinformatics/btab727}.

\bibitem[\citeproctext]{ref-Shapiro1965-zd}
Shapiro, S S, and M B Wilk. 1965. {``An Analysis of Variance Test for
Normality (Complete Samples).''} \emph{Biometrika} 52 (3/4): 591--611.
\url{https://doi.org/10.2307/2333709}.

\bibitem[\citeproctext]{ref-Subbotin1923-qk}
Subbotin, M Th. 1923. {``On the Law of Frequency of Error.''}
\emph{Математический Сборник} 31 (2): 296--301.

\bibitem[\citeproctext]{ref-Tukey1967-eb}
Tukey, J W. 1967. {``An Introduction to the Calculations of Numerical
Spectrum Analysis.''} In \emph{Advanced Seminar on Spectral Analysis of
Time Series}, edited by B Harris. New York: John Wiley; Sons, Inc.

\bibitem[\citeproctext]{ref-Tukey1957-rt}
Tukey, John W. 1957. {``On the Comparative Anatomy of
Transformations.''} \emph{Ann. Math. Stat.} 28 (3): 602--32.

\bibitem[\citeproctext]{ref-Tukey1977-xm}
Tukey, John Wilder. 1977. \emph{Exploratory Data Analysis}.
Addison-Wesley Publishing Company.

\bibitem[\citeproctext]{ref-Wright2017-rf}
Wright, Marvin N., and Andreas Ziegler. 2017. {``{ranger}: A Fast
Implementation of Random Forests for High Dimensional Data in {C++} and
{R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
\url{https://doi.org/10.18637/jss.v077.i01}.

\bibitem[\citeproctext]{ref-Yeo2000-vw}
Yeo, In‐kwon, and Richard A Johnson. 2000. {``A New Family of Power
Transformations to Improve Normality or Symmetry.''} \emph{Biometrika}
87 (4): 954--59. \url{https://doi.org/10.1093/biomet/87.4.954}.

\bibitem[\citeproctext]{ref-Zwanenburg2021-so}
Zwanenburg, Alex, and Steffen Löck. 2024a. {``Familiar: End-to-End
Automated Machine Learning and Model Evaluation.''} The R Foundation.
\url{https://doi.org/10.32614/cran.package.familiar}.

\bibitem[\citeproctext]{ref-Zwanenburg2024-kq}
---------. 2024b. {``Power.transform: Location and Scale Invariant Power
Transformations.''} The R Foundation.
\url{https://doi.org/10.32614/cran.package.power.transform}.

\end{CSLReferences}

\end{document}

### 2024-09-19 manuscript_latex/manuscript.tex
Correct parsing of authors.

@@
 
-\author[affil_1,affil_2]{Alex Zwanenburg} %% Author name
+%% Alex Zwanenburg
+\author[affil_1,affil_2]{Alex Zwanenburg\corref{cor_note}}
+\ead{alexander.zwanenburg@nct-dresden.de}
+\cortext[cor_note]{corresponding author}
 
-%% Author affiliation
 \affiliation[affil_1]{
@@
     Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, Dresden, Germany;
-    Helmholtz-Zentrum Dresden-Rossendorf (HZDR), Dresden, Germany
-  },
+    Helmholtz-Zentrum Dresden-Rossendorf (HZDR), Dresden, Germany},
   addressline={Fetscherstraße 74/PF 64},
@@
     Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology,
-    Helmholtz-Zentrum Dresden-Rossendorf
-  },
+    Helmholtz-Zentrum Dresden-Rossendorf},
   addressline={Fetscherstraße 74/PF 41},
@@
 }
-\author[affil_2,affil_3,affil_4]{Steffen Löck} %% Author name
 
-%% Author affiliation
+%% Steffen Löck
+\author[affil_2,affil_3,affil_4]{Steffen Löck}
+
 \affiliation[affil_3]{
@@
     Department of Radiotherapy and Radiation Oncology,
-    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology
-  },
+    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology},
   addressline={Fetscherstraße 74/PF 50},
@@
    German Cancer Consortium (DKTK), Partner Site Dresden,
-   and German Cancer Research Center (DKFZ)
-  },
+   and German Cancer Research Center (DKFZ)},
   addressline={Im Neuenheimer Feld 280},
@@
 %%Graphical abstract
-\begin{graphicalabstract}
-%\includegraphics{grabs}
-\end{graphicalabstract}
+% \begin{graphicalabstract}
+% \includegraphics{grabs}
+% \end{graphicalabstract}
 

### 2024-10-16 manuscript_rmarkdown/manuscript.Rmd
Fixed several typos.

@@
 The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
-However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit of for -- nor detriment to -- model performance for location- and scale-invariant power transformations.
+However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit -- nor detriment -- to model performance for location- and scale-invariant power transformations.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
 Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
-For the latter two datasets, the transformation method does not show significant difference between groups (Friedman test; $p > 0.05$).
+For the latter two datasets, the transformation method did not show significant difference between groups (Friedman test; $p > 0.05$).
 ```{r include = FALSE, eval = TRUE}
@@
 Consequently, the weights change at each iteration in the MLE optimisation process.
-This increases local variance in the log-likelihood function and create local optima that the optimiser may not handle well.
+This increases local variance in the log-likelihood function and creates local optima that the optimiser may not handle well.
 As a consequence, optimal values for transformation parameter $\lambda$ might differ, which increases the presented loss for optimising threshold parameters.
@@
 The empirical test for central normality differs from other tests for normality, such as the Shapiro-Wilk test [@Shapiro1965-zd], in that the test statistic is independent of the number of samples.
-This makes this test more practical for assessing central normality for larger sample numbers, where other test may detect inconsequential deviations from normality.
-
-This work has the following limitation.
+This makes this test more practical for assessing central normality for larger sample numbers, where other tests may detect inconsequential deviations from normality.
+
+This work has the following limitations.
 Firstly, we did observe several numerical stability issues for optimisation criteria other than MLE (Appendix B).

### 2024-10-16 manuscript_rmarkdown/manuscript.Rmd
Added missing doi to references.

@@
 The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
-As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2021-kq].
+As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2024-zf].
 
@@
 Their presence may lead to poor transformations to normality, as shown in Figure \ref{fig:decreased-normality}.
-As indicated by @Raymaekers2021-kq, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*.
+As indicated by @Raymaekers2024-zf, the general aim of power transformations should be to transform non-outlier data to normality, i.e. achieve *central normality*.
 To achieve this, they devised an iterative procedure to find a robust estimate of the transformation parameter $\lambda$.
 Briefly, this process requires identifying outliers in the data and weighting such instances during the optimisation process.
-@Raymaekers2021-kq achieve this through weighted maximum likelihood estimation.
+@Raymaekers2024-zf achieve this through weighted maximum likelihood estimation.
 However, because this procedure iteratively estimates and updates $\lambda$, it can not be used here to simultaneously estimate $\lambda$, $x_0$ and $s$ for location- and scale-invariant power transformations.
@@
 Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
-Compared to @Raymaekers2021-kq, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
+Compared to @Raymaekers2024-zf, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
 The weighted location- and scale-invariant Box-Cox log-likelihood function is:
@@
 - The weighting function uses the z-score of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$.
-  After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
+  After @Raymaekers2024-zf, $z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}$.
   Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].

### 2024-10-16 manuscript_rmarkdown/manuscript_appendix.Rmd
Changes to appendix figure 1

@@
 ```
-Where $x_i - x_0 > 0$.
+where $x_i - x_0 > 0$.
 The location and scale-invariant Yeo-Johnson transformation is:
@@
 
-We introduced per-sample weights $w_i$ to facilitate robust power transformations. The weighted log-likelihood of the location- and scale-invariant Box-Cox transformation is:
+Similarly to @Raymaekers2024-zf, sample weights $w_i$ are introduced to facilitate robust power transformations.
+The weighted log-likelihood of the location- and scale-invariant Box-Cox transformation is:
 ```{=latex}
@@
 The normal distribution is parametrised by location parameter $\mu$ and scale parameter $\sigma$, both of which have to be estimated from the data. 
-For non-robust power transformation, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
+For non-robust power transformations, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
 For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
@@
 For non-robust transformations, $w_i = 1$.
-Then, $sigma^2$ is defined as:
+Then, $\sigma^2$ is defined as:
 ```{=latex}
@@
 
-For the kurtosis test statistic we first compute [@DAgostino1990-kp, @Anscombe1983-nz]:
+For the kurtosis test statistic we first compute [@DAgostino1990-kp; @Anscombe1983-nz]:
 
@@
 
-Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in (Appendix B)[Appendix B: Optimisation of transformation parameters].
+Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
 This follows the simulation in the main manuscript, where MLE was used for optimization.
@@
 location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$);
-and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$.
+and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$).
 
@@
 
-Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:results-outlier-robustness-appendix}.
+Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.
 
@@
 cap <- paste0(
-  "Type 1 error rate as function of the test statistic $\\tau_{\\text{ecn}}$ ",
+  "Type I error rate as function of the test statistic $\\tau_{\\text{ecn}}$ ",
   "for five datasets, with central portion $\\kappa=0.80$. ",
-  "The type 1 error rate is computed from $m_d=10000$ randomly sampled features, ",
+  "The type I error rate is computed from $m_d=10000$ randomly sampled features, ",
   "These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), ",
@@
 These indicate that the test behaves similarly for the different datasets.
-For low type 1 error rates the test statistic proposed in the main manuscript is more conservative than alternatives 
+For low type I error rates, the test statistic proposed in the main manuscript is more conservative than alternatives 
 based on residuals after Box-Cox transformations of asymmetric generalised normally distributed features or on residuals from strictly normally distributed features.

### 2024-10-16 manuscript_rmarkdown/plots.R

@@ -664,6 +664,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
   # Prevent warnings due to non-standard evaluation.
   distribution <- method <- data_type <- version <- NULL
 
+  # Update plot theme
+  plot_theme$plot.tag.position <- "top"
+  plot_theme$plot.tag$margin <- grid::unit(c(0, 0, 2, 0), "points")
+
   ..create_plot <- function(
     p,
     distr,
@@ -870,6 +874,11 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p_left_scale_conv_yj <- ..create_plot(p_left, "left-skewed", "Yeo-Johnson", "conventional", "scaled")
   p_left_scale_invar_yj <- ..create_plot(p_left, "left-skewed", "Yeo-Johnson", "invariant", "scaled", last_row = TRUE)
 
+  # Add tags
+  p_normal_shift_conv_bc <- p_normal_shift_conv_bc + ggplot2::labs(tag = "conventional Box-Cox")
+  p_normal_shift_invar_bc <- p_normal_shift_invar_bc + ggplot2::labs(tag = "invariant Box-Cox")
+  p_normal_shift_conv_yj <- p_normal_shift_conv_yj + ggplot2::labs(tag = "conventional Yeo-Johnson")
+  p_normal_shift_invar_yj <- p_normal_shift_invar_yj + ggplot2::labs(tag = "invariant Yeo-Johnson")
 
   # Patch all the plots together.
   p <-
@@ -2343,7 +2352,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
   # Without transformation -----------------------------------------------------
   p0_qq <- p_qq + ggplot2::geom_point(
-    data = residual_data[transformation == "none"]
+    data = residual_data[transformation == "none"],
+    show.legend = c(colour = TRUE, size = TRUE, shape = FALSE)
   )
   p0_d <- p_d + ggplot2::geom_density(
     data = transformed_data[transformation == "none"]
@@ -2355,7 +2365,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
   # Standard transformation ----------------------------------------------------
   p1_qq <- p_qq + ggplot2::geom_point(
-    data  = residual_data[transformation %in% c("conventional", "Raymaekers-Rousseeuw")]
+    data  = residual_data[transformation %in% c("conventional", "Raymaekers-Rousseeuw")],
[…]

### 2024-10-17 manuscript_rmarkdown/manuscript.tex
Update figures to work with ggplot2 and patchwork

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2024-09-19}
+\date{2024-10-17}
 
@@
 Yeo-Johnson transformations were devised (Raymaekers and Rousseeuw
-2021).
+2024).
 
@@
 \ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
-(2021), the general aim of power transformations should be to transform
+(2024), the general aim of power transformations should be to transform
 non-outlier data to normality, i.e.~achieve \emph{central normality}. To
@@
 instances during the optimisation process. Raymaekers and Rousseeuw
-(2021) achieve this through weighted maximum likelihood estimation.
+(2024) achieve this through weighted maximum likelihood estimation.
 However, because this procedure iteratively estimates and updates
@@
 \ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
-and Rousseeuw (2021), these log-likelihood functions includes additional
+and Rousseeuw (2024), these log-likelihood functions includes additional
 terms to accommodate estimation of \(x_0\) and \(s\). The weighted
@@
   \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
-  Rousseeuw (2021),
+  Rousseeuw (2024),
   \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
@@
 up to 10 percent of the values could be replaced by outliers. Outlier
-values were set according to John Wilder Tukey (1977), as follows. Let
+values were set according to John W. Tukey (1977), as follows. Let
 \(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
@@
 real-world datasets that contained at least one numeric feature, we did
-not find a meaningful benefit of for -- nor detriment to -- model
-performance for location- and scale-invariant power transformations. One
-reason may be that numeric features with large location shifts
-(\(|\mu| > 1000.0\)) were uncommon. Of the 4886 numeric features in the
-232 datasets, 266 (5\%) features in 34 datasets had large location
-shifts, of which 200 appeared in just 2 datasets. For the latter two
-datasets, the transformation method does not show significant difference
-between groups (Friedman test; \(p > 0.05\)).
+not find a meaningful benefit -- nor detriment -- to model performance
+for location- and scale-invariant power transformations. One reason may
+be that numeric features with large location shifts (\(|\mu| > 1000.0\))
+were uncommon. Of the 4886 numeric features in the 232 datasets, 266
+(5\%) features in 34 datasets had large location shifts, of which 200
+appeared in just 2 datasets. For the latter two datasets, the
+transformation method did not show significant difference between groups
+(Friedman test; \(p > 0.05\)).
 
@@
 the MLE optimisation process. This increases local variance in the
-log-likelihood function and create local optima that the optimiser may
+log-likelihood function and creates local optima that the optimiser may
 not handle well. As a consequence, optimal values for transformation
@@
 practical for assessing central normality for larger sample numbers,
-where other test may detect inconsequential deviations from normality.
-
-This work has the following limitation. Firstly, we did observe several
+where other tests may detect inconsequential deviations from normality.
+
+This work has the following limitations. Firstly, we did observe several
 numerical stability issues for optimisation criteria other than MLE
@@
 \emph{J. R. Stat. Soc. Series B Stat. Methodol.} 26 (2): 211--52.
+\url{https://doi.org/10.1111/J.2517-6161.1964.TB00553.X}.
 
@@
 \bibitem[\citeproctext]{ref-Griffin2018-bf}
-Griffin, Maryclare. 2018. \emph{Gnorm: Generalized Normal/Exponential
-Power Distribution}. \url{https://CRAN.R-project.org/package=gnorm}.
+Griffin, Maryclare. 2018. {``Gnorm: Generalized Normal/Exponential Power
+Distribution.''} The R Foundation.
+\url{https://doi.org/10.32614/cran.package.gnorm}.
 
@@
 
-\bibitem[\citeproctext]{ref-Raymaekers2021-kq}
-Raymaekers, Jakob, and Peter J Rousseeuw. 2021. {``Transforming
-Variables to Central Normality.''} \emph{Mach. Learn.}, March.
-\url{https://doi.org/10.1007/s10994-021-05960-5}.
+\bibitem[\citeproctext]{ref-Raymaekers2024-zf}
+Raymaekers, Jakob, and Peter J Rousseeuw. 2024. {``Transforming
+Variables to Central Normality.''} \emph{Mach. Learn.} 113 (8):
+4953--75. \url{https://doi.org/10.1007/s10994-021-05960-5}.
 
@@
 \bibitem[\citeproctext]{ref-Subbotin1923-qk}
-Subbotin, M Th. 1923. {``On the Law of Frequency of Error.''}
-\emph{Математический Сборник} 31 (2): 296--301.
+Subbotin, M Th. 1923. {``On the Law of Frequency of Error.''} \emph{Mat.
+Sb.} 31 (2): 296--301.
 
@@
 Transformations.''} \emph{Ann. Math. Stat.} 28 (3): 602--32.
+\url{https://doi.org/10.1214/AOMS/1177706875}.
 
 \bibitem[\citeproctext]{ref-Tukey1977-xm}
-Tukey, John Wilder. 1977. \emph{Exploratory Data Analysis}.
-Addison-Wesley Publishing Company.
+---------. 1977. \emph{Exploratory Data Analysis}. Addison-Wesley
+Publishing Company.
 

### 2024-10-17 manuscript_rmarkdown/manuscript_appendix.Rmd
Update figures to work with ggplot2 and patchwork

@@
   "Estimates of the transformation parameter $\\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution and the optimisation criterion, ",
-  "whereas estimates obtained for the location- and scale-invariant power transformations are constant."
+  "whereas estimates obtained for the location- and scale-invariant power transformations are constant. ",
+  "For location- and scale-invariant power transformations, the Anderson-Darling criterion leads to unstable estimates of $\\lambda$ for skewed distributions, possibly due to large weights being assigned to samples at the upper and lower ends of the distribution."
 )

### 2024-10-17 manuscript_rmarkdown/manuscript_appendix.tex
Update figures to work with ggplot2 and patchwork

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{amsmath}
\usepackage{booktabs}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\sgn}{sgn}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Appendix},
  pdfauthor={Alex Zwanenburg, Steffen Löck},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2024-10-17}

\begin{document}
\maketitle

\section{Appendix A: Log-likelihood functions for location and scale
invariant power
transformation}\label{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}

Location and scale-invariant Box-Cox and Yeo-Johnson transformations are
parametrised using location \(x_0\) and scale \(s\) parameters, in
addition to transformation parameter \(\lambda\). This leads to the
following transformations. The location and scale-invariant Box-Cox
transformation is:

\begin{equation}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
\end{cases}
\end{equation}

where \(x_i - x_0 > 0\). The location and scale-invariant Yeo-Johnson
transformation is:

\begin{equation}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}

The parameters of these power transformations can be optimised based by
maximising the log-likelihood function, under the assumption that the
transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) follows a
normal distribution. The log-likelihood functions for conventional
Box-Cox and Yeo-Johnson transformations are well-known. However, the
introduction of scaling parameter \(s\) prevents their direct use. Here,
we first derive the general form of the log-likelihood functions, and
then derive their power-transformation specific definitions.

Let \(f(x_1, \ldots, x_n)\) be the probability density function of
feature \(\mathbf{X} = \{ x_1, \ldots, x_n\}\), and
\(f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))\)
be the probability density function of the transformed feature
\(\phi^{\lambda, x_0, s} (\mathbf{X})\), that is assumed to follow a
normal distribution.

The two probability density functions are related as follows:

\begin{equation}
f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
\end{equation}

Where, \(\left|\mathbf{J}\right|\) is the determinant of Jacobian
\(\mathbf{J}\). The Jacobian takes the following form, with off-diagonal
elements \(0\):

\begin{equation}
\mathbf{J} =
\begin{bmatrix}
    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
    \vdots & \vdots  & \ddots &  \vdots \\
    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
\end{bmatrix}
\end{equation}

Thus,
\(\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\).

Since in our situation \(\{x_1, \ldots, x_n\}\) in
\(f^{\lambda, x_0, s}(x_1, \ldots, x_n)\) are considered fixed (i.e.,
known), \(f^{\lambda, x_0, s}(x_1, \ldots, x_n)\) may be considered a
likelihood function. The log-likelihood function
\(\mathcal{l}^{\lambda, x_0, s}\) is then:

\begin{equation}
\begin{split}
\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
 & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
\end{split}
\end{equation}

With \(\mu\) the average of \(\phi^{\lambda, x_0, s}(\mathbf{X})\) and
\(\sigma^2\) its variance. The first two terms derive directly from the
log-likelihood function of a normal distribution, and are not specific
to the type of power transformation used. However, the final term
differs between Box-Cox and Yeo-Johnson transformations.

\subsection{Location- and scale-invariant Box-Cox
transformation}\label{location--and-scale-invariant-box-cox-transformation}

For the location- and scale-invariant Box-Cox transformation the partial
derivative is:

\begin{equation}
\begin{split}
\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
 & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
\end{split}
\end{equation}

Thus the final term in \(\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}\) is:

\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

This leads to the following log-likelihood:

\begin{equation}
\begin{split}
\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

Similarly to Raymaekers and Rousseeuw (2024), sample weights \(w_i\) are
introduced to facilitate robust power transformations. The weighted
log-likelihood of the location- and scale-invariant Box-Cox
transformation is:

\begin{equation}
\begin{split}
\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Box-Cox transformed feature
\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\), respectively:

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

\subsection{Location- and scale-invariant Yeo-Johnson
transformation}\label{location--and-scale-invariant-yeo-johnson-transformation}

For the location- and scale-invariant Yeo-Johnson transformation, the
partial derivative is:

\begin{equation}
\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
\begin{cases}
\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
\end{cases}
\end{equation}

Thus the final term in \(\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}\) is:

\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

This leads to the following log-likelihood:

\begin{equation}
\begin{split}
\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

The weighted log-likelihood for location- and scale-invariant
Yeo-Johnson transformation is:

\begin{equation}
\begin{split}
\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Yeo-Johnson transformed feature
\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

\section{Appendix B: Optimisation of transformation
parameters}\label{appendix-b-optimisation-of-transformation-parameters}

Maximum likelihood estimation (MLE) is commonly used to optimise
parameters for power transformation. Generally, optimisation requires
minimisation or maximisation of a criterion. In MLE, the maximised
criterion is the log-likelihood function of the normal distribution.
Here, we investigate power transformation using optimisation criteria
that are closely related to test statistics for normality tests.

Let \(\mathbf{X}\) be a feature with ordered feature values, and
\(\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)\) and
\(\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)\)
its transformed values using conventional and shift and scale invariant
power transformations, respectively. Since power transformations are
monotonic, \(\mathbf{Y}\) will likewise be ordered.

Below we will focus on criteria based on the empirical density function
and those based on skewness and kurtosis of the transformed featured.
Other potential criteria, such as the Shapiro-Wilk test statistic
(Shapiro and Wilk 1965) are not investigated here. In the case of the
Shapiro-Wilk test statistic this is because of lack of scalability to
features with many (\(> 5000\)) instances, and because adapting the test
statistic to include weights is not straightforward.

\subsection{Empirical density function-based
criteria}\label{empirical-density-function-based-criteria}

The first class of criteria is based on the empirical distribution
function (EDF). Transformation parameters are then fit through
minimisation of the distance between the empirical distribution function
\(F_{\epsilon}\) and the cumulative density function (CDF) of the normal
distribution \(F_{\mathcal{N}}\). Let
\(F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}\) be the
empirical probability of instance \(i\). The normal distribution is
parametrised by location parameter \(\mu\) and scale parameter
\(\sigma\), both of which have to be estimated from the data. For
non-robust power transformations, \(\mu\) and \(\sigma\) are sample mean
and sample standard deviation, respectively. For robust power
transformations, we estimate \(\mu\) and \(\sigma\) as Huber M-estimates
of location and scale of the transformed feature
\(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981).

\subsubsection{Anderson-Darling
criterion}\label{anderson-darling-criterion}

The Anderson-Darling criterion is based on the empirical distribution
function of \(\mathbf{X}\). We define this criterion as follows:

\begin{equation}
U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right) }
\end{equation}

Here \(w_i\) are weights, and \(\mu\) and \(\sigma\) are location and
scale parameters. For non-robust power transformations, all \(w_i = 1\).
Note that this criterion is not the same as the Anderson-Darling test
statistic (Anderson and Darling 1952), which involves solving (or
approximating) an integral function, contains an extra scalar
multiplication term, and does not include weights. The Anderson-Darling
criterion seeks to minimise the squared Euclidean distance between the
EDF and the normal CDF, with differences at the upper and lower end of
the normal CDF receiving more weight than those at the the centre of the
CDF.

\subsubsection{Cramér-von Mises
criterion}\label{cramuxe9r-von-mises-criterion}

The Cramér-von Mises criterion is also based on the empirical
distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
criterion as follows:

\begin{equation}
U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2
\end{equation}

Here \(w_i\) are weights, and \(\mu\) and \(\sigma\) are location and
scale parameters. For non-robust power transformations, all \(w_i = 1\).
The criterion is similar to the Cramér-von Mises test statistic Mises
(1928), aside from a additive scalar value and the introduction of
weights. This criterion, like the Anderson-Darling criterion, seeks to
minimise the squared Euclidean distance between the EDF and the normal
CDF. Unlike the Anderson-Darling criterion, this criterion weights all
instances equally.

For conventional power transformations with a fixed shift parameter, the
transformation \(\phi^{\lambda, x_0, s} (\mathbf{X})\) may be
substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
Cramér-von Mises criterion.

\subsection{Skewness-kurtosis-based
criteria}\label{skewness-kurtosis-based-criteria}

The second class of criteria seeks to reduce skewness and (excess)
kurtosis of the transformed feature \(\mathbf{Y}\). We will first define
the location \(\mu\) and scale \(\sigma\) of the the transformed as
these are required for computing skewness and kurtosis. Here, \(\mu\) is
defined as:

\begin{equation}
\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0, s} \left(x_i \right)} {\sum_{i=1}^n w_i}
\end{equation}

The location, or mean, is weighted using weights \(w_i\). For non-robust
transformations, \(w_i = 1\). Then, \(\sigma^2\) is defined as:

\begin{equation}
\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
\end{equation}

Skewness is defined as:

\begin{equation}
s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
\end{equation}

Kurtosis is defined as:

\begin{equation}
k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
\end{equation}

\subsubsection{D'Agostino criterion}\label{dagostino-criterion}

The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
test statistic (D'Agostino and Belanger 1990). This test statistic is
composed of two separate test statistics, one of which is related to
skewness, and the other to kurtosis. Both test statistics are computed
in several steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for
non-robust power transformations, \(\nu = n\).

For the skewness test statistic we first compute (D'Agostino and
Belanger 1990):

\begin{equation}
\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
\end{equation}

\begin{equation}
\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
\end{equation}

\begin{equation}
\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
\end{equation}

\begin{equation}
\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
\end{equation}

The skewness test statistic is then:

\begin{equation}
Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
\end{equation}

For the kurtosis test statistic we first compute (D'Agostino and
Belanger 1990; Anscombe and Glynn 1983):

\begin{equation}
\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
\end{equation}

\begin{equation}
\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
\end{equation}

\begin{equation}
\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
\end{equation}

\begin{equation}
\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
\end{equation}

\begin{equation}
\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
\end{equation}

The kurtosis test statistic is then:

\begin{equation}
Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
\end{equation}

The D'Agostino \(K^2\) test statistic and our criterion are the same,
and are defined as:

\begin{equation}
U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
\end{equation}

The main difference between the test statistic as originally formulated,
and the criterion proposed here is the presence of weights for robust
power transformation.

\subsubsection{Jarque-Bera criterion}\label{jarque-bera-criterion}

The second criterion based on skewness and kurtosis is the Jarque-Bera
criterion. It is relatively simple to compute compared to the D'Agostino
criterion:

\begin{equation}
U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
\end{equation}

The main difference between the above criterion and the Jarque-Bera test
statistic (Jarque and Bera 1980) is that a scalar multiplication is
absent.

\subsection{Optimisation using non-MLE
criteria}\label{optimisation-using-non-mle-criteria}

Each of the above criteria can be used for optimisation, i.e.:

\begin{equation}
\left\{ \hat{\lambda}, \hat{x}_0, \hat{s}_0 \right\} = \argmin_{\lambda, x_0, s} U\left(\mathbf{X}, \lambda, x_0, s \right)
\end{equation}

For conventional power transformations with fixed location and scale
parameters, the transformation \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
may be substituted by \(\phi^{\lambda} (\mathbf{X})\), or equivalently,
\(x_0\) and \(s\) may be fixed:

\begin{equation}
\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0, s \right)
\end{equation}

\section{Appendix C: Simulations with other optimisation
criteria}\label{appendix-c-simulations-with-other-optimisation-criteria}

Invariance of location- and scale-invariant power transformations was
assessed using the optimisation criteria in
\href{Appendix\%20B:\%20Optimisation\%20of\%20transformation\%20parameters}{Appendix
B}. This follows the simulation in the main manuscript, where MLE was
used for optimization. In short, we first randomly drew \(10000\) values
from a normal distribution:
\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
or equivalently
\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)\).
The second distribution was a right-skewed normal distribution
\(\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)\).
The third distribution was a left-skewed normal distribution
\(\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)\).
We then computed transformation parameter \(\lambda\) using the original
definitions (equations \ref{eqn:box-cox-original} and
\ref{eqn:yeo-johnson-original}) and the location- and scale-invariant
definitions (equations \ref{eqn:box-cox-invariant} and
\ref{eqn:yeo-johnson-invariant}) for each distribution using different
optimisation criteria. To assess location invariance, a positive value
\(d_{\text{shift}}\) was added to each distribution with
\(d_{\text{shift}} \in [1, 10^6]\). Similarly, to assess scale
invariance, each distribution was multiplied by a positive value
\(d_{\text{scale}}\), where \(d_{\text{scale}} \in [1, 10^6]\).

The results are shown in Figure
\ref{fig:shifted-distributions-appendix}.

\begin{figure}

{\centering \includegraphics{manuscript_appendix_files/figure-latex/shifted-distributions-appendix-1}

}

\caption{Invariant power transformation produces transformation parameters that are invariant to location and scale. Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\text{shift}}$ or multiplication by $d_{\text{scale}}$. Estimates of the transformation parameter $\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution and the optimisation criterion, whereas estimates obtained for the location- and scale-invariant power transformations are constant. For location- and scale-invariant power transformations, the Anderson-Darling criterion leads to unstable estimates of $\lambda$ for skewed distributions, possibly due to large weights being assigned to samples at the upper and lower ends of the distribution.}\label{fig:shifted-distributions-appendix}
\end{figure}

\section{Appendix D: Experimental results using location- and
scale-invariant Box-Cox
transformation}\label{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}

The effect of using location- and scale-invariant transformations was
investigated using real-world datasets.

\subsection{Invariance}\label{invariance}

Results for Box-Cox transformations of features without outliers are
shown in Figure \ref{fig:experimental-results-invariance-appendix}.

\begin{figure}

{\centering \includegraphics{manuscript_appendix_files/figure-latex/experimental-results-invariance-appendix-1}

}

\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row). Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).}\label{fig:experimental-results-invariance-appendix}
\end{figure}

\subsubsection{Age of patients with lung
cancer}\label{age-of-patients-with-lung-cancer}

Applying conventional and invariant Box-Cox transformations to age of
patients with lung cancer (Loprinzi et al. 1994) yielded the following
results: no transformation (sum of residuals with normal distribution
\(\sum r_i = 16.5\)); conventional transformation (\(\lambda = 1.9\),
\(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
\(\sigma_{BC} = 0.4 \cdot 10^3\)); Raymaekers and Rousseeuw's robust
adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\));
location- and scale-invariant transformation (\(\lambda = 1.7\),
\(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\), \(\sigma_{BC} = 0.8\)); and
robust location- and scale-invariant transformation (\(\lambda = 1.5\),
\(\sum r_i = 11.6\), \(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\)).

Compared to location- and scale-invariant Yeo-Johnson transformations,
the Box-Cox transformations do not reduce residuals compared to
conventional variants.

\subsubsection{Penguin body mass}\label{penguin-body-mass}

Applying conventional and invariant Box-Cox transformations to the body
mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
following results: no transformation (residual sum \(\sum r_i = 48.0\));
conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); location- and
scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 27.3\),
\(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\)); and robust location- and
scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 25.2\),
\(\mu_{BC} = 0.4\), \(\sigma_{BC} = 0.6\)).

Just as for location- and scale-invariant Yeo-Johnson transformations,
Box-Cox transformations produced a lower overall residual sum compared
to their conventional counterparts. Similarly, conventional
transformations led to low standard deviation \(\sigma_{YJ}\) of the
body mass feature after transformation.

\subsubsection{Latitude in the Ames housing
dataset}\label{latitude-in-the-ames-housing-dataset}

Applying conventional and invariant Box-Cox transformations to the
latitude of houses in the Ames housing dataset (De Cock 2011) yielded
the following results: no transformation (residual sum
\(\sum r_i = 328\)); conventional transformation (\(\lambda = 62.1\),
\(\sum r_i = 319\), \(\mu_{BC} = 1.1 \cdot 10^{99}\),
\(\sigma_{BC} = 0.0 \cdot 10^{99}\)); Raymaekers and Rousseeuw's robust
adaptation (\(\lambda = 96.0\), \(\sum r_i = 319\),
\(\mu_{BC} = 6.2 \cdot 10^{153}\),
\(\sigma_{BC} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
transformation (\(\lambda = 1.9\), \(\sum r_i = 312\),
\(\mu_{BC} = 2.3\), \(\sigma_{BC} = 0.9\)); and robust location- and
scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\)).

Similar to conventional Yeo-Johnson transformations (non-robust and
robust), conventional Box-Cox transformations had high values for the
\(\lambda\) parameter, which could lead to numerical issues. Location-
and scale-invariant Box-Cox transformations did not suffer from this
issue.

\subsection{Robustness against
outliers}\label{robustness-against-outliers}

Results for Box-Cox transformations of features with outliers are shown
in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.

\begin{figure}

{\centering \includegraphics{manuscript_appendix_files/figure-latex/experimental-results-outlier-robustness-appendix-1}

}

\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row). Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column). Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness-appendix}
\end{figure}

\subsubsection{Fuel efficiency in the Top Gear
dataset}\label{fuel-efficiency-in-the-top-gear-dataset}

The Top Gear dataset contains data on 297 vehicles, with outliers
related highly fuel-efficient vehicles (Alfons 2021). Applying
conventional and invariant Box-Cox transformations to the fuel
consumption feature yielded the following results: no transformation
(residual sum \(\sum r_i = 54\), \(p=0.76\)); conventional
transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.55\)); location- and
scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p=0.02\)); and robust
location- and scale-invariant transformation (\(\lambda = 1.1\),
\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
\(p=0.83\)).

\subsubsection{Maximum arterial wall thickness in an ischemic stroke
dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}

The ischemic stroke dataset contains historic data from 126 patients
with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
conventional and invariant Box-Cox transformations to the maximum
arterial wall thickness feature yielded the following results: no
transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.60\)); location- and
scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.01\)); and robust
location- and scale-invariant transformation (\(\lambda = 0.5\),
\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
\(p=0.35\)).

\section{Appendix E: Empirical central normality
test}\label{appendix-e-empirical-central-normality-test}

The empirical central normality test was derived using data sampled from
asymmetric generalised normal distributions, including outliers, to
resemble more realistic datasets. Here we assess the type I error rate
of two, less realistic, sets of data:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Data sampled from asymmetric generalised normal distributions without
  outliers.
\item
  Data sampled from normal distributions without outliers, without any
  power transformation applied.
\end{enumerate}

Other aspects of the experiment remained the same. Thus, we first drew
\(m_d=10000\) random distributions. For asymmetric generalised normal
distributions, each distribution was parametrised with a randomly chosen
skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
parameter \(\beta \sim U\left(1.00, 5.00 \right)\). For fully normal
distributions, skewness parameter \(\alpha = 0.5\) and shape parameter
\(\beta = 2.0\) were fixed. Location and scale parameters were set as
\(\mu = 0\) and \(\sigma = 1\), respectively.
\(n = \lceil 10^\gamma \rceil\) values were then randomly drawn, with
\(\gamma \sim U\left(1.47, 3.00\right)\), which led to between \(30\)
and \(1000\) values being drawn to create \(\mathbf{X}_i\). Residuals
were then computed after performing robust location- and scale-invariant
transformations with the empirical tapered cosine weighting method for
the dataset with asymmetric generalised normal distributions, and
without any transformation for the dataset with only normal
distributions.

\begin{figure}

{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-test-appendix-1}

}

\caption{Type I error rate as function of the test statistic $\tau_{\text{ecn}}$ for five datasets, with central portion $\kappa=0.80$. The type I error rate is computed from $m_d=10000$ randomly sampled features, These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none). The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none).}\label{fig:empirical-central-normality-test-appendix}
\end{figure}

\begin{table}
\begin{center}
\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
\label{tab:empirical-central-normality-appendix}
\begin{tabular}{l | c c c c c c c}

\toprule
type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

The results are shown in Figure
\ref{fig:empirical-central-normality-test-appendix} and Table
\ref {tab:empirical-central-normality-appendix}. These indicate that the
test behaves similarly for the different datasets. For low type I error
rates, the test statistic proposed in the main manuscript is more
conservative than alternatives based on residuals after Box-Cox
transformations of asymmetric generalised normally distributed features
or on residuals from strictly normally distributed features.

\section{Appendix F: Normalisation before
transformation}\label{appendix-f-normalisation-before-transformation}

An alternative to location- and scale-invariant transformations is
normalising feature distributions prior to conventional transformations.
Table \ref{tab:normalisation-before-transformation-appendix} shows
residual errors, after transformation to normality, of the five features
from real-world datasets presented previously in Appendix D and the main
manuscript. In these examples location- and scale-invariant
transformations have similar or lower residual errors compared to errors
resulting from normalisation prior to transformation.

\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
\label{tab:normalisation-before-transformation-appendix}
\begin{tabular}{l | c c c c c}

\toprule
feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\

\midrule
age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Alfons2021-kc}
Alfons, Andreas. 2021. {``{robustHD}: An {R} Package for Robust
Regression with High-Dimensional Data.''} \emph{J. Open Source Softw.} 6
(67): 3786. \url{https://doi.org/10.21105/joss.03786}.

\bibitem[\citeproctext]{ref-Anderson1952-gz}
Anderson, T W, and D A Darling. 1952. {``Asymptotic Theory of Certain
{`Goodness of Fit'} Criteria Based on Stochastic Processes.''}
\emph{Annals of Mathematical Statistics} 23 (2): 193--212.
\url{https://doi.org/10.1214/aoms/1177729437}.

\bibitem[\citeproctext]{ref-Anscombe1983-nz}
Anscombe, F J, and William J Glynn. 1983. {``Distribution of the
Kurtosis Statistic B2 for Normal Samples.''} \emph{Biometrika} 70 (1):
227--34. \url{https://doi.org/10.1093/biomet/70.1.227}.

\bibitem[\citeproctext]{ref-Cramer1928-rc}
Cramér, Harald. 1928. {``On the Composition of Elementary Errors.''}
\emph{Scand. Actuar. J.} 1928 (1): 13--74.
\url{https://doi.org/10.1080/03461238.1928.10416862}.

\bibitem[\citeproctext]{ref-DAgostino1990-kp}
D'Agostino, Ralph B, and Albert Belanger. 1990. {``A Suggestion for
Using Powerful and Informative Tests of Normality.''} \emph{Am. Stat.}
44 (4): 316--21. \url{https://doi.org/10.2307/2684359}.

\bibitem[\citeproctext]{ref-De-Cock2011-jf}
De Cock, Dean. 2011. {``Ames, Iowa: Alternative to the Boston Housing
Data as an End of Semester Regression Project.''} \emph{J. Stat. Educ.}
19 (3). \url{https://doi.org/10.1080/10691898.2011.11889627}.

\bibitem[\citeproctext]{ref-Gorman2014-eo}
Gorman, Kristen B, Tony D Williams, and William R Fraser. 2014.
{``Ecological Sexual Dimorphism and Environmental Variability Within a
Community of Antarctic Penguins (Genus Pygoscelis).''} \emph{PLoS One} 9
(3): e90081. \url{https://doi.org/10.1371/journal.pone.0090081}.

\bibitem[\citeproctext]{ref-Huber1981-su}
Huber, Peter J. 1981. \emph{Robust Statistics}. John Wiley \& Sons.
\url{https://doi.org/10.1002/0471725250}.

\bibitem[\citeproctext]{ref-Jarque1980-hw}
Jarque, Carlos M, and Anil K Bera. 1980. {``Efficient Tests for
Normality, Homoscedasticity and Serial Independence of Regression
Residuals.''} \emph{Econ. Lett.} 6 (3): 255--59.
\url{https://doi.org/10.1016/0165-1765(80)90024-5}.

\bibitem[\citeproctext]{ref-Kuhn2019-kt}
Kuhn, Max, and Kjell Johnson. 2019. \emph{Feature Engineering and
Selection: A Practical Approach for Predictive Models}. Chapman \&
Hall/CRC Data Science Series. Chapman; Hall/CRC.
\url{https://doi.org/10.1201/9781315108230}.

\bibitem[\citeproctext]{ref-Loprinzi1994-cd}
Loprinzi, C L, J A Laurie, H S Wieand, J E Krook, P J Novotny, J W
Kugler, J Bartel, M Law, M Bateman, and N E Klatt. 1994. {``Prospective
Evaluation of Prognostic Variables from Patient-Completed
Questionnaires. North Central Cancer Treatment Group.''} \emph{J. Clin.
Oncol.} 12 (3): 601--7. \url{https://doi.org/10.1200/JCO.1994.12.3.601}.

\bibitem[\citeproctext]{ref-Von_Mises1928-ef}
Mises, Richard von. 1928. \emph{Wahrscheinlichkeit Statistik Und
Wahrheit}. Schriften Zur Wissenschaftlichen Weltauffassung.
Springer-Verlag Berlin, Heidelberg.
\url{https://doi.org/10.1007/978-3-662-36230-3}.

\bibitem[\citeproctext]{ref-Raymaekers2024-zf}
Raymaekers, Jakob, and Peter J Rousseeuw. 2024. {``Transforming
Variables to Central Normality.''} \emph{Mach. Learn.} 113 (8):
4953--75. \url{https://doi.org/10.1007/s10994-021-05960-5}.

\bibitem[\citeproctext]{ref-Shapiro1965-zd}
Shapiro, S S, and M B Wilk. 1965. {``An Analysis of Variance Test for
Normality (Complete Samples).''} \emph{Biometrika} 52 (3/4): 591--611.
\url{https://doi.org/10.2307/2333709}.

\end{CSLReferences}

\end{document}

### 2024-10-17 manuscript_latex/manuscript.tex
Further manuscript preparations.
Manuscript main body
It compiles!

To do:
- add appendix
- replace citations

@@
 %% \usepackage{lineno}
+
+%% The booktabs package provides table makeup macros.
+\usepackage{booktabs}
+
+%% The hyperref package allows for typesetting hyperlinks.
+%% \usepackage{hyperref}
+
+%% Custom math operators
+\DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator{\sgn}{sgn}
 
@@
 %% Abstract
+% You are required to provide a concise and factual abstract which does not exceed
+% 250 words. The abstract should briefly state the purpose of your research,
+% principal results and major conclusions. Some guidelines:
+%
+% Abstracts must be able to stand alone as abstracts are often presented separately from the article.
+%
+% Avoid references. If any are essential to include, ensure that you cite the author(s) and year(s).
+%
+% Avoid non-standard or uncommon abbreviations. If any are essential to include,
+% ensure they are defined within your abstract at first mention.
+
 \begin{abstract}
-%% Text of abstract
-Abstract text.
+
 \end{abstract}
@@
 %%Research highlights
+% Highlights are a short collection of bullet points that should capture the novel
+% results of your research as well as any new methods used during your study.
+% Highlights will help increase the discoverability of your article via search engines.
+% Some guidelines:
+%
+% Submit highlights as a separate editable file in the online submission system
+% with the word "highlights" included in the file name.
+%
+% Highlights should consist of 3 to 5 bullet points, each a maximum of 85 characters, including spaces.
+
 \begin{highlights}
@@
 %% Keywords
+% You are required to provide 1 to 7 keywords for indexing purposes. Keywords should be written in English.
+% Please try to avoid keywords consisting of multiple words (using "and" or "of").
+% We recommend that you only use abbreviations in keywords if they are firmly established in the field.
 \begin{keyword}
+
 %% keywords here, in the form: keyword \sep keyword
@@
 %%
+\section{Introduction}\label{introduction}
 
-%% Use \section commands to start a section
-\section{Example Section}
-\label{sec1}
-%% Labels are used to cross-reference an item using \ref command.
+Many statistical and some machine learning methods assume normality of
+the underlying data, e.g.~analysis of variance and linear discriminant
+analysis. However, numerical features in datasets may strongly deviate
+from normal distributions, e.g.~by being skewed. Power transformations
+aim to stabilise variance and improve normality of such features
+(Bartlett 1947; John W. Tukey 1957). The two most commonly used
+transformations are that of Box and Cox (1964) and Yeo and Johnson
+(2000). The Box-Cox transformation of a feature value \(x_i\) of feature
+\(\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}\) under the
+transformation parameter \(\lambda\) is defined as:
 
-Section text. See Subsection \ref{subsec1}.
-
-%% Use \subsection commands to start a subsection.
-\subsection{Example Subsection}
-\label{subsec1}
-
-Subsection text.
-
-%% Use \subsubsection, \paragraph, \subparagraph commands to
-%% start 3rd, 4th and 5th level sections.
-%% Refer following link for more details.
-%% https://en.wikibooks.org/wiki/LaTeX/Document_Structure#Sectioning_commands
-
-\subsubsection{Mathematics}
-%% Inline mathematics is tagged between $ symbols.
-This is an example for the symbol $\alpha$ tagged as inline mathematics.
-
-%% Displayed equations can be tagged using various environments.
-%% Single line equations can be tagged using the equation environment.
 \begin{equation}
-f(x) = (x+a)(x+b)
+\label{eqn:box-cox-original}
+\phi_{\text{BC}}^\lambda (x_i) = 
+\begin{cases}
+\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log(x_i) & \text{if } \lambda = 0
+\end{cases}
 \end{equation}
 
-%% Unnumbered equations are tagged using starred versions of the environment.
-%% amsmath package needs to be loaded for the starred version of equation environment.
-\begin{equation*}
-f(x) = (x+a)(x+b)
-\end{equation*}
+One limitation of the Box-Cox transformation is that it is only defined
+for \(x_i > 0\). In contrast, the Yeo-Johnson transformation under the
+transformation parameter \(\lambda\) is defined for any
+\(x_i \in \mathbb{R}\):
 
-%% align or eqnarray environments can be used for multi line equations.
-%% & is used to mark alignment points in equations.
-%% \\ is used to end a row in a multiline equation.
-\begin{align}
- f(x) &= (x+a)(x+b) \\
-      &= x^2 + (a+b)x + ab
-\end{align}
+\begin{equation}
+\label{eqn:yeo-johnson-original}
+\phi_{\text{YJ}}^\lambda (x_i) = 
+\begin{cases}
+\left( \left( 1 + x_i \right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i \geq 0\\
+\log(1 + x_i) & \text{if } \lambda = 0 \text{ and } x_i \geq 0\\
+-\left( \left( 1 - x_i\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i < 0\\
+-\log(1 - x_i) & \text{if } \lambda = 2 \text{ and } x_i < 0
+\end{cases}
+\end{equation}
 
-\begin{eqnarray}
- f(x) &=& (x+a)(x+b) \nonumber\\ %% If equation numbering is not needed for a row use \nonumber.
-      &=& x^2 + (a+b)x + ab
-\end{eqnarray}
+The \(\lambda\)-parameter is typically optimised using maximum
+likelihood estimation under the assumption that the transformed feature
+is normally distributed. As noted by Raymaekers and Rousseeuw, this
+approach is sensitive to outliers, and robust versions of Box-Cox and
+Yeo-Johnson transformations were devised (Raymaekers and Rousseeuw
+2024).
 
-%% Unnumbered versions of align and eqnarray
-\begin{align*}
- f(x) &= (x+a)(x+b) \\
-      &= x^2 + (a+b)x + ab
-\end{align*}
+Applying a power transformation does not guarantee that transformed
+features are normally distributed. Depending on location and scale of a
+feature and the presence of outliers, power transformations may decrease
+normality, as shown in Figure \ref{fig:decreased-normality}. If
+normality of the transformed feature is not checked, e.g.~in automated
+power transformation in machine learning workflows, several issues may
+arise. For example, statistical tests such as ANOVA may produce
+incorrect results due to violation of the normality assumption.
+Likewise, machine learning methods that assume normality of input
+features may suffer a decrease in performance. Moreover, large negative
+or positive \(\lambda\)-parameters may lead to numeric issues in any
+subsequent computations.
 
-\begin{eqnarray*}
- f(x)&=& (x+a)(x+b) \\
-     &=& x^2 + (a+b)x + ab
-\end{eqnarray*}
+Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
+and Wilk 1965), could be automatically applied to transformed features.
+However, given sufficiently large sample sizes such tests can detect
+trivial deviations from normality, and may lead to rejection of
+sufficiently good power transformations.
 
-%% Refer following link for more details.
-%% https://en.wikibooks.org/wiki/LaTeX/Mathematics
-%% https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
+\begin{figure}
 
-%% Use a table environment to create tables.
-%% Refer following link for more details.
-%% https://en.wikibooks.org/wiki/LaTeX/Tables
-\begin{table}[t]%% placement specifier
-%% Use tabular environment to tag the tabular data.
-%% https://en.wikibooks.org/wiki/LaTeX/Tables#The_tabular_environment
-\centering%% For centre alignment of tabular.
-\begin{tabular}{l c r}%% Table column specifiers
-%% Tabular cells are separated by &
-  1 & 2 & 3 \\ %% A tabular row ends with \\
-  4 & 5 & 6 \\
-  7 & 8 & 9 \\
+{\centering \includegraphics{figure_1}
+
+}
+
+\caption{Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\lambda$.
+$10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the shift dataset, $\mathcal{N}(10, \sigma)$ for the scale dataset and $\mathcal{N}(0, 1)$ for the outlier dataset.
+Additionally, an outlier with value $d$ was added to the outlier dataset.
+Since samples are drawn from a normal distribution, a transformation parameter of $\lambda = 1$ is expected.
+However, a large shift in location, a scale that is small compared to the location, or presence of large outliers lead to incorrectly estimated transformation parameter values.}\label{fig:decreased-normality}
+\end{figure}
+
+To address these issues, we make the following contributions:
+
+\begin{itemize}
+\item
+  We devise location- and scale-invariant versions of the Box-Cox and
+  Yeo-Johnson transformation, including versions robust to outliers.
+\item
+  We derive the maximum likelihood criterion for location- and
+  scale-invariant Box-Cox and Yeo-Johnson transformations to allow for
+  optimising transformation parameters.
+\item
+  We define an empirical central normality test for detecting cases
+  where power transformations fail to yield an approximately normally
+  distributed transformed feature.
+\item
+  We assess the effect of power transformations on the performance of
+  machine learning models.
+\end{itemize}
+
+\section{Theory}\label{theory}
+
+In this section, we will first introduce location- and scale-invariant
+versions of the Box-Cox and Yeo-Johnson transformations. Subsequently,
+we define weighted location- and scale-invariant transformations and
+weighting methods for robust transformations. We then define the
+quantile function for asymmetric generalised normal distributions to
+enable random sampling. Finally, we define the overall framework for the
+empirical central normality test.
+
+\subsection{Location- and scale-invariant power
+transformation}\label{location--and-scale-invariant-power-transformation}
+
+Box-Cox and Yeo-Johnson transformations are modified by introducing
+shift parameter \(x_0\) and scale parameter \(s\) into equations
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}. The
+location- and scale-invariant Box-Cox transformation of a feature value
+\(x_i\) of feature \(\mathbf{X}\) under transformation parameter
+\(\lambda\), shift parameter \(x_0\) and scale parameter \(s\) is then:
+
+\begin{equation}
+\label{eqn:box-cox-invariant}
+\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
+\end{cases}
+\end{equation}
+
+where \(x_i - x_0 > 0\). Likewise, the location- and scale-invariant
+Yeo-Johnson transformation of a feature value \(x_i\) under
+transformation parameter \(\lambda\), shift parameter \(x_0\) and scale
+parameter \(s\) is:
+
+\begin{equation}
+\label{eqn:yeo-johnson-invariant}
+\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
+\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
+-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
+\end{cases}
+\end{equation}
+
+For both invariant transformations, \(\lambda\), \(x_0\) and \(s\)
+parameters can be obtained by maximising the log-likelihood function,
+i.e.~using maximum likelihood estimation (MLE). A full derivation of the
+log-likelihood function for both transformations is shown in Appendix A.
+The location- and scale-invariant Box-Cox log-likelihood function is:
+
+\begin{equation}
+\label{eqn:box-cox-invariant-log-likelihood}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+
+subject to \(x_i - x_0 > 0\). \(\mu\) and \(\sigma^2\) are the mean and
+variance of the Box-Cox transformed feature
+\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
+Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood
+function is:
+
+\begin{equation}
+\label{eqn:yeo-johnson-invariant-log-likelihood}
+\begin{split}
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+
+where \(\mu\) and \(\sigma^2\) are the mean and variance of the
+Yeo-Johnson transformed feature
+\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
+
+\subsection{Robust location- and scale-invariant power
+transformations}\label{robust-location--and-scale-invariant-power-transformations}
+
+Real-world data may contain outliers, to which maximum likelihood
+estimation can be sensitive. Their presence may lead to poor
+transformations to normality, as shown in Figure
+\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
+(2024), the general aim of power transformations should be to transform
+non-outlier data to normality, i.e.~achieve \emph{central normality}. To
+achieve this, they devised an iterative procedure to find a robust
+estimate of the transformation parameter \(\lambda\). Briefly, this
+process requires identifying outliers in the data and weighting such
+instances during the optimisation process. Raymaekers and Rousseeuw
+(2024) achieve this through weighted maximum likelihood estimation.
+However, because this procedure iteratively estimates and updates
+\(\lambda\), it can not be used here to simultaneously estimate
+\(\lambda\), \(x_0\) and \(s\) for location- and scale-invariant power
+transformations. Nonetheless, as a procedure, weighted MLE can be used
+for estimating the transformation, shift and scale parameters.
+
+Here, weighted maximum likelihood estimation is based on equations
+\ref{eqn:box-cox-invariant-log-likelihood} and
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
+and Rousseeuw (2024), these log-likelihood functions includes additional
+terms to accommodate estimation of \(x_0\) and \(s\). The weighted
+location- and scale-invariant Box-Cox log-likelihood function is:
+
+\begin{equation}
+\label{eqn:box-cox-weighted-invariant-log-likelihood}
+\begin{split}
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+
+where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
+variance of the Box-Cox transformed feature
+\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\):
+
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+
+Analogously, the weighted location- and scale-invariant Yeo-Johnson
+log-likelihood function is:
+
+\begin{equation}
+\label{eqn:yeo-johnson-weighted-invariant-log-likelihood}
+\begin{split}
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+
+where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
+variance of the Yeo-Johnson transformed feature
+\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):
+
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+
+The weights \(w_i\) in equations
+\ref{eqn:box-cox-weighted-invariant-log-likelihood} and
+\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood} can be set using
+several weighting functions. Using \(\dot{x}_i\) as an argument that
+will be defined later, we investigate three weighting functions:
+
+\begin{itemize}
+\item
+  A step function, with \(\delta_1 \geq 0\) as threshold parameter:
+\end{itemize}
+
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_1
+\end{cases}
+\end{equation}
+
+\begin{itemize}
+\item
+  A triangle function (or generalised Huber weight), with
+  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
+  parameters:
+\end{itemize}
+
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
+1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_2
+\end{cases}
+\end{equation}
+
+\begin{itemize}
+\item
+  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
+  and \(\delta_2 \geq \delta_1\) as threshold parameters:
+\end{itemize}
+
+\begin{equation}
+w_i =
+\begin{cases}
+1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
+0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
+0 & \text{if } \left| \dot{x}_i \right| > \delta_2
+\end{cases}
+\end{equation}
+
+All weighting functions share the characteristic that for
+\(\left| \dot{x}_i \right|< \delta_1\), instances are fully weighted,
+i.e.~when \(\delta_1 > 0\) the weighting functions are symmetric window
+functions with a flat top. The triangle and tapered cosine functions
+then gradually down-weight instances with
+\(\delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2\), and assign no
+weight to instances \(\left| x_i \right| > \delta_2\). Examples of these
+weighting function are shown in Figure \ref{fig:weighting-functions}.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_2}
+
+}
+
+\caption{Weighting functions investigated in this study to make power transformations more robust against outliers.
+In this example, the step function was parameterised with $\delta_1 = 0.60$.
+The triangle and tapered cosine functions were both parameterised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
+\end{figure}
+
+Each weighting function has an argument \(\dot{x}\) that is related to
+the (transformed) feature in one of several ways:
+
+\begin{itemize}
+\item
+  The weighting function uses empirical probabilities of the
+  distribution of the original feature \(\mathbf{X}\). After sorting
+  \(\mathbf{X}\) in ascending order, probabilities are determined as
+  \(p_i = \frac{i - 1/3}{n + 1/3}\), with \(i = 1, 2, \ldots n\), with
+  \(n\) the number of instances of feature \(\mathbf{X}\). Then
+  \(\dot{x}_i = p^{*}_i=2 \left( p_i - 0.5\right)\), so that argument is
+  zero-centered.
+\item
+  The weighting function uses the z-score of the transformed feature
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
+  Rousseeuw (2024),
+  \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
+  \(\mu_M\) and \(\sigma_M\) are robust Huber M-estimates of location
+  and scale of the transformed feature
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981). Then
+  \(\dot{x}_i = z_i\).
+\item
+  After sorting \(\mathbf{X}\) in ascending order, the weighting
+  function uses the residual error between the z-score of the
+  transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
+  theoretical z-score from a standard normal distribution:
+  \(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
+  with \(\mu_M\), \(\sigma_M\) and \(p_i\) as defined above. Then
+  \(\dot{x}_i = r_i\).
+\end{itemize}
+
+\subsection{Asymmetric generalised normal
+distributions}\label{asymmetric-generalised-normal-distributions}
+
+Modifications intended to make power transformations invariant to
+location and scale of a feature and methods to improve their robustness
+against outliers need to be assessed using data drawn from a range of
+different distributions. Since the power transformations are intended
+for use with unimodal distributions, the generalised normal distribution
+(Subbotin 1923; Nadarajah 2005) is a suitable option for simulating
+realistic feature distributions. This distribution has the following
+probability density function \(f_{\beta}\) for a value
+\(x \in \mathbb{R}\):
+
+\begin{equation}
+f_{\beta}(x) = \frac{\beta}{2\Gamma\left(1 / \beta \right)} e^{-\left| x \right|^\beta}
+\end{equation}
+
+Here, \(\Gamma\) is the gamma function, and \(\beta\) is a strictly
+positive shape parameter. For \(\beta = 1\), the probability density
+function describes a Laplace distribution. A normal distribution is
+found for \(\beta=2\), and for large \(\beta\), the distribution
+approaches a uniform distribution. We will refrain from introducing
+scale and location parameters here directly.
+
+Realistic feature distributions may be skewed. Gijbels et al.~describe a
+recipe for introducing skewness into the otherwise symmetric generalised
+normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
+the following probability density function:
+
+\begin{equation}
+f_{\alpha}(x; \mu, \sigma, \beta) = \frac{2 \alpha \left(1 - \alpha\right)}{\sigma}
+\begin{cases}
+f_{\beta}\left( \left(1 - \alpha \right) \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x \leq \mu \\
+f_{\beta}\left( \alpha \frac{\left| x - \mu \right|}{\sigma} \right) & \text{, } x > \mu
+\end{cases}
+\end{equation}
+
+Here, \(\alpha \in (0,1)\) is a skewness parameter. \(\alpha > 0.5\)
+creates a distribution with a negative skew, i.e.~a left-skewed
+distribution. A right-skewed distribution is created for
+\(\alpha < 0.5\). \(\mu \in \mathcal{R}\) and \(\sigma \in (0, \infty)\)
+are location and scale parameters, respectively. \(f_{\alpha}\) thus
+describes the probability density function of an asymmetric generalised
+normal distribution, which we will refer to here and parametrise as
+\(\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)\).
+
+We require a quantile function (or an approximation thereof) to draw
+random values from an asymmetric generalised normal distribution using
+inverse transform sampling. Gijbels et al.~derived the following
+quantile function \(F_{\alpha}^{-1}(p)\):
+
+\begin{equation}
+F_{\alpha}^{-1}(p; \mu, \sigma, \beta) =
+\begin{cases}
+\mu + \frac{\sigma}{1 - \alpha} F_{\beta}^{-1} \left( \frac{p}{2 \alpha}\right) & \text{, } p \leq \alpha \\
+\mu + \frac{\sigma}{\alpha} F_{\beta}^{-1} \left( \frac{1 + p - 2 \alpha}{2 \left(1 - \alpha \right)} \right) & \text{, } p > \alpha
+\end{cases}
+\end{equation}
+
+The quantile function for the asymmetric generalised normal distribution
+\(F_{\alpha}^{-1}\) thus incorporates the quantile function
+\(F_{\beta}^{-1}\) of the symmetric generalised normal distribution.
+\(F_{\beta}^{-1}\) was derived by Griffin to be (Griffin 2018):
+
+\begin{equation}
+F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
+\end{equation}
+
+Here, \(F_{\Gamma}^{-1}\) is the quantile function of the gamma
+distribution with shape \(1 / \beta\), which can be numerically
+approximated.
+
+\subsection{Empirical central normality
+test}\label{empirical-central-normality-test}
+
+Power transformations aim to transform features to a normal
+distribution. However, this may not always be successful or possible.
+Deviations from normality can be detected by normality tests, such as
+the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
+tests may be too stringent with large sample sizes, outliers, or both.
+Here we develop an empirical test for central normality. The null
+hypothesis \(H_0\) is that the distribution is centrally normal. The
+alternative hypothesis \(H_1\) is that the distribution is not centrally
+normal.
+
+Let central normality be defined as the normality of central portion
+\(\kappa\) of the data,
+i.e.~\(\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\),
+with \(p_i\) probabilities of the empirical distribution, as previously.
+We then compute the residual errors between the z-scores of the
+transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
+expected z-scores from a standard normal distribution:
+\(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
+with \(\mu_M\) and \(\sigma_M\) robust Huber M-estimates of location and
+scale of the transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
+(Huber 1981). The set of residual errors for the central portion of the
+data is then
+\(\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\).
+
+The test statistic \(\tau_{\text{ecn}}\) is then defined as:
+
+\begin{equation}
+\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
+\end{equation}
+
+Here \([\quad]\) denotes an Iverson bracket. The test statistic is equal
+to the mean of the residual errors of the central portion of the data.
+
+To use the test statistic, the central portion of the data needs to be
+defined and the Type 1 error rates determined. We will do so in
+section \ref{simulation}.
+
+\section{Simulation}\label{simulation}
+
+We used simulated data to assess invariance to location and scale of the
+proposed power transformations, weighting for robust transformations,
+and to develop the empirical central normality test. The \(\lambda\)
+parameter for conventional power transformations (Eqn.
+\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well
+as \(\lambda\), \(x_0\) and \(s\) parameters for location- and
+scale-invariant power transformations (Eqn. \ref{eqn:box-cox-invariant}
+and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
+algorithm for derivative-free bound constraint optimisation (Powell
+2009) through maximum likelihood estimation. The required algorithms
+were implemented in the \texttt{power.transform} R software package
+(Zwanenburg and Löck 2024b) (version 1.0.0). Of note, the
+\texttt{power.transform} package shifts feature values into the positive
+domain if negative or zero values are present for Box-Cox power
+transformations.
+
+\subsection{Invariance to location and
+scale}\label{invariance-to-location-and-scale}
+
+To assess whether the proposed power transformations lead to values of
+\(\lambda\) that are invariant to location and scale of the
+distribution, we simulated three different distributions. We first
+randomly drew \(10000\) values from a normal distribution:
+\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
+or equivalently
+\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)\).
+The second distribution was a right-skewed generalised normal
+distribution
+\(\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)\).
+The third distribution was a left-skewed generalised normal distribution
+\(\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)\).
+We then computed transformation parameter \(\lambda\) using the original
+definitions (Eqn. \ref{eqn:box-cox-original} and
+\ref{eqn:yeo-johnson-original}) and the location- and scale-invariant
+definitions (Eqn. \ref{eqn:box-cox-invariant} and
+\ref{eqn:yeo-johnson-invariant}) for each distribution. To assess
+location invariance, a positive value \(d_{\text{shift}}\) was added to
+each distribution with \(d_{\text{shift}} \in [1, 10^6]\). Similarly, to
+assess scale invariance, each distribution was multiplied by a positive
+value \(d_{\text{scale}}\), where \(d_{\text{scale}} \in [1, 10^6]\).
+
+\begin{figure}
+
+{\centering \includegraphics{figure_3}
+
+}
+
+\caption{Invariant power transformation produces transformation parameters that are invariant to location and scale.
+Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\text{shift}}$ or multiplication by $d_{\text{scale}}$.
+Estimates of the transformation parameter $\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution, whereas estimates obtained for the location- and scale-invariant power transformations are constant.}\label{fig:shifted-distributions}
+\end{figure}
+
+The result is shown in Figure \ref{fig:shifted-distributions}. For each
+distribution, transformation parameter \(\lambda\) varied with
+\(d_{\text{shift}}\) and \(d_{\text{scale}}\) when estimated for
+conventional transformations. In contrast, estimation of \(\lambda\) for
+invariant power transformations was invariant to both
+\(d_{\text{shift}}\) and \(d_{\text{scale}}\).
+
+\subsection{Robust transformations}\label{robust-transformations}
+
+Outliers may be present in data and affect estimation of transformation
+parameters. The log-likelihood function can be weighted to assign less
+weight to outlier instances, see equations
+\ref{eqn:box-cox-weighted-invariant-log-likelihood} and
+\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}. We proposed
+three weighting function: step, triangle and tapered cosine, that have
+one, two and two parameters, respectively. Each weighting function then
+takes one of three values as input: probabilities of the empirical
+distribution of the original feature, the z-score of the transformed
+feature values, or the residual error between the z-score of the
+transformed feature values and their expected z-score based on the
+normal distribution.
+
+To determine the weighting function parameters for each of the nine
+combinations, \(m_d=100\) asymmetric generalised normal distributions
+were drawn. Each distribution was parametrised with a randomly chosen
+skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
+parameter \(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale
+parameters were set as \(\mu = 0\) and \(\sigma = 1\), respectively.
+\(n = \lceil 10^\gamma \rceil\) instances were then randomly drawn, with
+\(\gamma \sim U\left(2, 4\right)\), i.e., between \(100\) and \(10000\)
+values are drawn to create \(\mathbf{X}_i\).
+
+Outlier values were then drawn to randomly replace a fraction of the
+values of \(\mathbf{X}_i\). This was repeated \(m_{\text{out}} = 10\)
+times, with outlier fractions regularly spaced in \([0.00, 0.10]\). Thus
+up to 10 percent of the values could be replaced by outliers. Outlier
+values were set according to John W. Tukey (1977), as follows. Let
+\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
+was:
+
+\begin{equation}
+x_{out} =
+\begin{cases}
+Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
+Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
+\end{cases}
+\end{equation}
+
+\(Q_1\), \(Q_3\) and \(\text{IQR}\) are the first quartile, third
+quartile and interquartile range of \(\mathbf{X}_i\), respectively.
+Outlier values randomly replaced values in \(\mathbf{X}_i\) to create
+\(\mathbf{X}_{i,j}\), with
+\(j \in \left\{ 1, \ldots, m_{\text{out}} \right\}\)
+
+To find the optimal values for the weighting function parameters
+\(\delta_1\) and \(\delta_2\) (if applicable), we minimised the absolute
+difference between the \(\lambda_{r}\) parameter obtained for robust
+transformation in the presence of outliers, and the \(\lambda_0\)
+parameter obtained using the non-robust transformations in absence of
+outliers:
+
+\begin{equation}
+\label{eqn:minimisation-weighting}
+\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
+\end{equation}
+
+Minimisation was conducted using the BOBYQA algorithm for
+derivative-free bound constraint optimisation (Powell 2009). The
+resulting weighting function parameters for weighted MLE are shown in
+Tables \ref{tab:optimal-weighting-parameters-box-cox} and
+\ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location-
+and scale-invariant Box-Cox and Yeo-Johnson transformations,
+respectively.
+
+\begin{table}
+\begin{center}
+\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
+The {optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
+Lower loss indicates better robustness against outliers.}
+\label{tab:optimal-weighting-parameters-box-cox}
+\begin{tabular}{l r r r r r r r}
+
+\toprule
+method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
+& initial & limits & optimal & initial & limits & optimal & \\
+
+\midrule
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
+\bottomrule
 \end{tabular}
-%% Use \caption command for table caption and label.
-\caption{Table Caption}\label{fig1}
+\end{center}
 \end{table}
 
-%% Use figure environment to create figures
-%% Refer following link for more details.
-%% https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions
-\begin{figure}[t]%% placement specifier
-%% Use \includegraphics command to insert graphic files. Place graphics files in
-%% working directory.
-\centering%% For centre alignment of image.
-\includegraphics{example-image-a}
-%% Use \caption command for figure caption and label.
-\caption{Figure Caption}\label{fig1}
-%% https://en.wikibooks.org/wiki/LaTeX/Importing_Graphics#Importing_external_graphics
+\begin{table}
+\begin{center}
+\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
+The {optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
+Lower loss indicates better robustness against outliers.}
+\label{tab:optimal-weighting-parameters-yeo-johnson}
+\begin{tabular}{l r r r r r r r}
+
+\toprule
+method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
+& initial & limits & optimal & initial & limits & optimal & \\
+
+\midrule
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+Figure \ref{fig:optimised-weighting-function-parameters} shows the
+distribution of errors \(\left| \lambda_r - \lambda_0 \right|\) for
+non-robust and robust transformations using the optimal weighting
+function parameters. In the presence of outliers, the Yeo-Johnson
+transformation yielded smaller errors than the Box-Cox transformation.
+For both transformations, weighting based on empirical probabilities
+yielded the most consistent \(\lambda\) parameter estimates in the
+presence of outliers. Other methods did not outperform the non-robust
+transformation method.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_4}
+
+}
+
+\caption{Robustness of power transformations after optimising weighting function parameters.
+The distribution of errors, i.e.
+the difference between robustly fitted $\lambda^r$ and expected $\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly parametrised asymmetric generalised normal distributions with up to 10\% outliers on either side of the distribution.
+The top and bottom panel show errors for the location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
+In each panel, the error distribution for the non-robust transformation is shown to the left for comparison with different weighting methods.
+Note that a square root transformation was applied along the $y$-axis for display purposes.
+Median errors are shown above each distribution.}\label{fig:optimised-weighting-function-parameters}
 \end{figure}
+
+\subsection{Empirical central normality
+test}\label{empirical-central-normality-test-1}
+
+To develop an empirical test for central normality we need to consider
+two parameters: the central portion \(\kappa\) as a fixed parameter, and
+test statistic \(\tau_{\text{ecn}}\). We will first define the central
+portion \(\kappa\).
+
+First we draw \(m_d=10000\) random asymmetric generalised normal
+distributions. As before, each distribution is parametrised with a
+randomly chosen skewness parameter
+\(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
+are set as \(\mu = 0\) and \(\sigma = 1\), respectively.
+\(n = \lceil 10^\gamma \rceil\) values are then randomly drawn, with
+\(\gamma \sim U\left(1.47, 3.00\right)\), which leads to between \(30\)
+and \(1000\) values being drawn to create \(\mathbf{X}_i\). As before,
+for each distribution, up to \(10 \%\) of instances are replaced by
+outlier values, and this is repeated \(10\) times. We then compute
+residuals after optimising robust location- and scale-invariant
+transformations with the empirical tapered cosine weighting method.
+
+Figure \ref{fig:empirical-central-normality-residual-error} shows the
+residual errors of the transformed distribution as function of the
+empirical probability. As expected, the largest deviations from
+normality appear on the extremities of this range. For Box-Cox
+transformations, the effect of outliers on the lower end of the
+distribution leads to noticeable asymmetry. Between very low and very
+high empirical probabilities, residual errors are constrained and
+relatively flat. This is the candidate range for the central portion of
+the data.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_5}
+
+}
+
+\caption{Residual error as a function of empirical probability.
+Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric generalised normal distributions, for each of which outliers were randomly drawn 10 times (up to 10\% of samples).
+Larger errors occur at the edges of each distribution.}\label{fig:empirical-central-normality-residual-error}
+\end{figure}
+
+We then consider the empirical probability of type I errors: the
+probability of incorrectly classifying a centrally normal distribution
+as being non-normal based on the test statistic \(\tau_{\text{ecn}}\).
+Under the assumption that the asymmetric generalised normal
+distributions are centrally normal after robust transformation, this
+produces the relationship shown in Figure
+\ref{fig:empirical-central-normality-type-1-error-rate}. Since for
+\(\kappa \leq 0.80\) error curves for both transformation methods are
+similar, we fix the value of the central portion \(\kappa\) to 0.80. In
+the remaining we will use the empirical central normality test statistic
+values defined using robust location- and scale-invariant Yeo-Johnson
+transformations, as it is more conservative (see Appendix E). This leads
+to test statistic values listed in Table
+\ref{tab:empirical-central-normality}
+
+\begin{figure}
+
+{\centering \includegraphics{figure_6}
+
+}
+
+\caption{Type 1 error rate of transformed asymmetric generalised normal distributions as function of the test statistic $\tau_{\text{ecn}}$ for the central portion $\kappa$ of the distribution.}\label{fig:empirical-central-normality-type-1-error-rate}
+\end{figure}
+
+\begin{table}
+\begin{center}
+\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
+\label{tab:empirical-central-normality}
+\begin{tabular}{l | c c c c c c c}
+
+\toprule
+type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+
+\midrule
+$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+In Figure \ref{fig:empirical-central-normality-examples} we apply the
+empirical central normality test to assess central normality of features
+that are composed of a mixture of samples drawn from two normal
+distributions. With increased separation of the underlying normal
+distributions, the probability of the feature being centrally normal
+decreases, as expected.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_7}
+
+}
+
+\caption{Bi-modal distributions and empirical central normality test results.
+The feature (black) is a mixture of two identical sample sets (blue and orange) drawn from normal distributions that are offset by a distance $d$.
+We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
+As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
+Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
+\end{figure}
+
+\section{Experimental Results}\label{experimental-results}
+
+\subsection{Invariance}\label{invariance}
+
+Location- and scale-invariant power transformations are intended to
+yield improved transformations to normality in the presence of large
+shifts in location, distributions that due to location and scale are not
+centered near zero, or both. Earlier, we assessed these transformations
+using simulated data. In the following, they are evaluated using
+examples from real datasets. We focus on the Yeo-Johnson transformation
+because of its ability to handle features with negative values. Results
+for Box-Cox transformations are shown in Appendix D.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_8}
+
+}
+
+\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row).
+Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Yeo-Johnson transformations (right column).}\label{fig:experimental-results-invariance}
+\end{figure}
+
+\subsubsection{Age of patients with lung
+cancer}\label{age-of-patients-with-lung-cancer}
+
+A common feature in health-related datasets is age. Here we use data on
+228 patients with lung cancer that was collected and published by
+Loprinzi et al. (Loprinzi et al. 1994). The age in the cohort was
+\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
+conventional and invariant Yeo-Johnson transformations to patient age
+yielded the following results, see Figure
+\ref{fig:experimental-results-invariance}: no transformation (sum of
+residuals with normal distribution \(\sum r_i = 16.5\)); conventional
+transformation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
+\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\));
+Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 2.0\),
+\(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
+\(\sigma_{YJ} = 0.5 \cdot 10^3\)); location- and scale-invariant
+transformation (\(\lambda = 0.9\), \(\sum r_i = 8.8\),
+\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.1\)); and robust location- and
+scale-invariant transformation (\(\lambda = 0.8\), \(\sum r_i = 10.6\),
+\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.0\)).
+
+Location- and scale-invariant transformation led to a lower overall
+residual sum, indicating a better transformation. Robust location- and
+scale-invariant transformation had a higher residual sum compared to the
+non-robust variant, which may be due to the lack of outliers in the
+data. Conventional transformations inflated the mean \(\mu_{YJ}\) and
+standard deviation \(\sigma_{YJ}\) of the age feature after
+transformation. The empirical central normality test did not detect any
+statistically significant deviations from central normality for any
+transformation (all \(p \geq 0.78\)).
+
+\subsubsection{Penguin body mass}\label{penguin-body-mass}
+
+Gorman, Williams and Fraser recorded body mass (in grams) of 342
+penguins of three different species (Gorman, Williams, and Fraser 2014).
+The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
+deviation) grams, and not centrally normal (\(p = 0.03\)). Applying
+conventional and invariant Yeo-Johnson transformations to body mass
+yielded the following results, see Figure
+\ref{fig:experimental-results-invariance}: no transformation (residual
+sum \(\sum r_i = 48.0\)); conventional transformation
+(\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); Raymaekers and Rousseeuw's robust
+adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); location- and scale-invariant
+transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
+\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\)); and robust location- and
+scale-invariant transformation (\(\lambda = 0.4\), \(\sum r_i = 23.1\),
+\(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.8\)).
+
+Location- and scale-invariant transformation produced a lower overall
+residual sum, indicating a better transformation. Moreover, conventional
+transformations led to low standard deviation \(\sigma_{YJ}\) of the
+body mass feature after transformation. The empirical central normality
+test did not detect any statistically significant deviations from
+central normality for any transformation (all \(p \geq 0.38\)).
+
+\subsubsection{Latitude in the Ames housing
+dataset}\label{latitude-in-the-ames-housing-dataset}
+
+Geospatial datasets usually contain coordinates. The Ames housing
+dataset contains data on 2930 properties that were sold between 2006 and
+2010 (De Cock 2011), including their geospatial coordinates. The
+latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
+conventional and invariant Yeo-Johnson transformations to latitude
+yielded the following results, see Figure
+\ref{fig:experimental-results-invariance}: no transformation (residual
+sum \(\sum r_i = 328\)); conventional transformation
+(\(\lambda = 62.1\), \(\sum r_i = 319\),
+\(\mu_{YJ} = 4.8 \cdot 10^{99}\), \(\sigma_{YJ} = 0.1 \cdot 10^{99}\));
+Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 95.4\),
+\(\sum r_i = 319\), \(\mu_{YJ} = 6.4 \cdot 10^{153}\),
+\(\sigma_{YJ} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
+transformation (\(\lambda = 1.5\), \(\sum r_i = 326\),
+\(\mu_{YJ} = -1.2\), \(\sigma_{YJ} = 0.8\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.4\), \(\sum r_i = 311\),
+\(\mu_{YJ} = -1.3\), \(\sigma_{YJ} = 0.9\)).
+
+Every transformation reduced the residual sum. The non-robust location-
+and scale-invariant transformation did not improve over conventional
+alternatives and yielded a data distribution lacking central normality
+(empirical central normality test: \(p=0.05\)). However, conventional
+transformations had high values for the \(\lambda\) parameter, which
+could lead to numerical issues.
+
+\subsection{Robustness against
+outliers}\label{robustness-against-outliers}
+
+We previously simulated data to assess invariant power transformations
+and their robustness against outliers. Here, we assess invariant power
+transformations in real data with outliers.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_9}
+
+}
+
+\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row).
+Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Yeo-Johnson transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Yeo-Johnson transformations (right column).
+Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness}
+\end{figure}
+
+\subsubsection{Fuel efficiency in the Top Gear
+dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
+
+The Top Gear dataset contains data on 297 vehicles that appeared on the
+BBC television show \emph{Top Gear} (Alfons 2021). Within this dataset,
+the fuel consumption feature contains outliers due to highly
+fuel-efficient vehicles. Applying conventional and invariant Yeo-Johnson
+transformations to the fuel consumption feature yielded the following
+results, see Figure \ref{fig:experimental-results-outlier-robustness}:
+no transformation (residual sum \(\sum r_i = 54\), \(p=0.76\));
+conventional transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
+\(\mu_{YJ} = 3.0\), \(\sigma_{YJ} = 0.3\), \(p=0.01\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
+\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.55\)); location- and
+scale-invariant transformation (\(\lambda = -1.3\), \(\sum r_i = 44\),
+\(\mu_{YJ} = 0.5\), \(\sigma_{YJ} = 0.1\), \(p=0.03\)); and robust
+location- and scale-invariant transformation (\(\lambda = -0.9\),
+\(\sum r_i = 50\), \(\mu_{YJ} = 2.0\), \(\sigma_{YJ} = 2.3\),
+\(p=0.58\)).
+
+Outliers cause non-robust transformations to fail to transform the data
+to a centrally normal distribution (empirical central normality test
+\(p = 0.01\) and \(p=0.03\) for conventional and invariant
+transformations, respectively). Robust transformations produce
+distributions that are centrally normal (empirical central normality
+test \(p > 0.05\)).
+
+\subsubsection{Maximum arterial wall thickness in an ischemic stroke
+dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
+
+The ischemic stroke dataset contains historic data from 126 patients
+with risk at ischemic stroke (Kuhn and Johnson 2019). These patients
+underwent Computed Tomography Angiography to characterize the carotid
+artery blockages. Angiography imaging was then assessed, and various
+characteristics related to the blood vessels and the disease are
+measured. The maximum arterial wall thickness feature contains several
+instances with outlier values. Applying conventional and invariant
+Yeo-Johnson transformations to this feature yielded the following
+results, see Figure \ref{fig:experimental-results-outlier-robustness}:
+no transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
+conventional transformation (\(\lambda = -0.7\), \(\sum r_i = 30\),
+\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p=0.01\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 136\),
+\(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14\), \(p=0.61\)); location- and
+scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 12\),
+\(\mu_{YJ} = -11.8\), \(\sigma_{YJ} = 6.9\), \(p=0.13\)); and robust
+location- and scale-invariant transformation (\(\lambda = -0.6\),
+\(\sum r_i = 27\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.1\),
+\(p=0.10\)).
+
+Non-robust transformations failed to produce a centrally normal
+distribution (empirical central normality test \(p=0.01\) and \(p=0.02\)
+for conventional and invariant transformations, respectively). Robust
+transformations produce distributions that are centrally normal
+(empirical central normality test \(p > 0.05\)).
+
+\subsection{Integration into end-to-end machine
+learning}\label{integration-into-end-to-end-machine-learning}
+
+We used 285 datasets from the Penn Machine Learning Benchmarks
+collection (Romano et al. 2022). In this collection, 122 datasets
+correspond to regression tasks and 163 datasets to classification tasks.
+Using the familiar auto-machine learning library (Zwanenburg and Löck
+2024a) (version 1.5.0), each dataset was used to train a model for each
+of 16 process configurations. Each process configuration specifies the
+learner (generalised linear model or random forest), transformation
+method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson,
+robust invariant Yeo-Johnson with empirical central normality test
+(rejecting transformations with \(p \leq 0.01\)), and normalisation
+method (none, \(z\)-standardisation), yielding 16 distinct
+configurations. Before each experiment, each dataset was randomly split
+into a training (70\%) and holdout test (30\%) set five times. Thus, a
+total of 22800 models were created. Each model was then evaluated using
+the holdout test set using one of two metrics, i.e.~the root relative
+squared error (RRSE) for regression tasks and the area under the
+receiver operating characteristic curve (AUC) for classification tasks.
+
+For the purpose of assessing the effect of the difficulty of the task,
+we computed the median performance score over all models for each
+dataset and assigned one the following categories:
+
+\begin{itemize}
+\item
+  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (87
+  datasets)
+\item
+  easy: \(0.90 > \text{AUC} \geq 0.80\) or
+  \(0.30 \geq \text{RRSE} > 0.10\) (46 datasets)
+\item
+  intermediate: \(0.80 > \text{AUC} \geq 0.70\) or
+  \(0.60 \geq \text{RRSE} > 0.30\) (72 datasets)
+\item
+  difficult: \(0.70 > \text{AUC} \geq 0.60\) or
+  \(0.80 \geq \text{RRSE} > 0.60\) (60 datasets)
+\item
+  very difficult: \(0.60 > \text{AUC} \geq 0.50\) or
+  \(1.00 \geq \text{RRSE} > 0.80\) (12 datasets)
+\item
+  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (8
+  datasets)
+\end{itemize}
+
+To remove the effect of the dataset, and allow for comparing metrics, we
+ranked all performance scores for each dataset so that a higher rank
+corresponds to better performance. Experiments yielding the same score
+received the same, average, rank. Subsequently ranks were normalised to
+the \([0.0, 1.0]\) range.
+
+Significant differences exist between process configurations (Friedman
+test: \(p < 10^{-8})\).
+
+Here we focus on the subset of 232 datasets that contain numeric
+features. Considering single process parameters, the choice of learner
+(Wilcoxon signed rank test: \(p < 10^{-8})\)) and normalisation method
+(Wilcoxon signed rank test: \(p = 0.007\)) had a significant impact (at
+significance level \(p = 0.050\)), but transformation method (Friedman
+test: \(p = 0.054\)) did not.
+
+To estimate the marginal effects of process parameters, including
+transformation method, we first fit a regression random forest (ranger
+package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, minimum
+node size 2, other hyperparameters default) with process parameters and
+task difficulty as predictors and normalised rank as response variable.
+The estimated marginal effects are shown in Figure
+\ref{fig:marginal-effect-plot}. On the scale of normalised ranks
+(\([0.0, 1.0]\)), the overall estimated marginal effect of using a
+random forest instead of generalised linear model was \(0.272\). The
+overall marginal effect of using z-standardisation to normalise features
+was \(0.008\). Transformation methods had the following marginal
+effects: \(0.003\) for using conventional Yeo-Johnson transformation
+instead of no transformation; \(-0.007\) for using robust invariant
+Yeo-Johnson transformation instead of no transformation; \(-0.009\) for
+using robust invariant Yeo-Johnson transformation with empirical central
+normality test instead of no transformation; \(-0.010\) for using robust
+invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson
+transformation; and \(-0.012\) for using using robust invariant
+Yeo-Johnson transformation with empirical central normality test instead
+of conventional Yeo-Johnson transformation.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_10}
+
+}
+
+\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets.
+The top-left panel shows the marginal effect of learners, i.e.
+random forests and generalised linear models (GLM).
+Random forests outperform GLM models for all task difficulties.
+The top-right panel shows the marginal effect of feature normalisation methods, i.e.
+no normalisation and z-standardisation.
+z-standardisation is generally beneficial, but the estimated effect is neglible.
+The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method.
+There is no consistent behaviour, and estimated effects are neglible.
+Note that the ranges of the $x$-axes of the three main panels differ.
+ECNT: empirical central normality test.}\label{fig:marginal-effect-plot}
+\end{figure}
+
+\section{Discussion}\label{discussion}
+
+In their work on power transformation, Box and Cox already mention
+transformation with a shift parameter, but preferred the version in Eq.
+\ref{eqn:box-cox-original} for the theoretical analysis in their paper
+(Box and Cox 1964), which subsequently became the convention. Yeo and
+Johnson's power transformation lacks a shift parameter altogether (Yeo
+and Johnson 2000). We showed that these power transformations are
+sensitive to location and scale of data distributions. To mitigate this
+issue, we defined location- and scale-invariant variants of the Box-Cox
+and Yeo-Johnson transformations. We furthermore assessed methods for
+making these transformations robust to outliers, and devised an
+empirical test for central normality.
+
+Robust location- and scale-invariant transformations are a suitable
+replacement for their conventional counterparts. They demonstrated
+robustness against outliers and prevent inaccurate transformations and
+potential numerical issues due to location and scale of the distribution
+of a feature. This is particularly relevant for automated data
+processing, where such issues may go unnoticed. Compared to non-robust
+location- and scale-invariant transformations, real-world examples with
+outliers showed higher residual errors of transformed features. Robust
+transformations seek to minimise residual errors for the central part of
+the distribution, instead of the entire distribution, including
+outliers. The empirical central normality test indicated that robust
+transformations are better able to achieve central normality in the
+presence of outliers. However, in a machine learning experiment of 232
+real-world datasets that contained at least one numeric feature, we did
+not find a meaningful benefit -- nor detriment -- to model performance
+for location- and scale-invariant power transformations. One reason may
+be that numeric features with large location shifts (\(|\mu| > 1000.0\))
+were uncommon. Of the 4886 numeric features in the 232 datasets, 266
+(5\%) features in 34 datasets had large location shifts, of which 200
+appeared in just 2 datasets. For the latter two datasets, the
+transformation method did not show significant difference between groups
+(Friedman test; \(p > 0.05\)).
+
+Location- and scale-invariant transformations are realised by
+simultaneously optimising three parameters, i.e.~transformation
+parameter \(\lambda\), shift parameter \(x_0\) and scale parameter
+\(s\). We derived the log-likelihood function to facilitate optimisation
+using MLE. Alternatively, standardisation of a numeric variable (e.g.,
+through subtracting its median value and division by its interquartile
+range) prior to conventional power transformations may achieve a similar
+effect in reducing sensitivity to the distribution's location and scale.
+While this alternative helps prevent these issues -- provided that
+normalisation does not lead to negative values for Box-Cox
+transformation -- location- and scale-invariant transformations seem to
+provide an overall better transformation to normality (Appendix F).
+
+We assessed several methods for robust power transformation. Methods
+that relied on the z-score of the transformed feature or the residual
+error yielded worse results than the non-robust method. This is partly
+due to the initial choice of threshold parameters. Using different
+initial values, closer to the upper limits
+(\(\delta_1 = 10.0; \delta_2 = 10.0\)), led to a reduced loss. However,
+these values effectively correspond to a non-robust transformation,
+where all instances receive the same weight. Underperformance of these
+methods could be explained by their use of transformed features for
+setting weights. Consequently, the weights change at each iteration in
+the MLE optimisation process. This increases local variance in the
+log-likelihood function and creates local optima that the optimiser may
+not handle well. As a consequence, optimal values for transformation
+parameter \(\lambda\) might differ, which increases the presented loss
+for optimising threshold parameters. Methods that relied on the
+empirical probability did not suffer from this issue, as weights
+remained fixed during MLE.
+
+We introduced an empirical test for central normality to assess whether
+distributions deviate from normality in a way that might require closer
+inspection prior to further processing. The empirical test for central
+normality differs from other tests for normality, such as the
+Shapiro-Wilk test (Shapiro and Wilk 1965), in that the test statistic is
+independent of the number of samples. This makes this test more
+practical for assessing central normality for larger sample numbers,
+where other tests may detect inconsequential deviations from normality.
+
+This work has the following limitations. Firstly, we did observe several
+numerical stability issues for optimisation criteria other than MLE
+(Appendix B). These appear in regions where transformation parameters
+would lead to very large or small numbers when using conventional power
+transformations. For MLE stability issues were not observed. Secondly,
+the empirical central normality test is based on simulations instead of
+statistical theory, and relies on a somewhat arbitrary definition of the
+central portion of a distribution. Thus, while the test may asses
+whether data is sufficiently normally distributed for practical
+purposes, it should not be used as a strict test for normality.
+
+\section{Conclusion}\label{conclusion}
+
+Compared to their conventional versions, robust location- and
+scale-invariant Box-Cox and Yeo-Johnson transformations reduce
+sensitivity to outliers and the location and scale of features. An
+empirical central normality test can assess the quality of
+transformation of features to normal distributions. The combination of
+both facilitate the use of power transformations in automated data
+analysis workflows.
+
+\section{Data and code availability}\label{data-and-code-availability}
+
+Location- and scale-invariant power transformations were implemented in
+the \texttt{power.transform} package for R, which is available from
+GitHub (https://github.com/oncoray/power.transform) and the CRAN
+repository (https://cran.r-project.org/package=power.transform).
+The manuscript was created using R Markdown and is likewise
+available from the \texttt{power.transform} GitHub repository. Data and
+results for the machine learning experiment are separately available
+from Zenodo (https://doi.org/10.5281/zenodo.13736671).
 

### 2024-10-17 manuscript_latex/manuscript.tex
Added appendix.

@@
 \appendix
-\section{Example Appendix Section}
-\label{app1}
-
-Appendix text.
+\section{Log-likelihood functions for location and scale
+invariant power transformation}\label{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}
+
+Location and scale-invariant Box-Cox and Yeo-Johnson transformations are
+parametrised using location \(x_0\) and scale \(s\) parameters, in
+addition to transformation parameter \(\lambda\). This leads to the
+following transformations. The location and scale-invariant Box-Cox
+transformation is:
+
+\begin{equation}
+\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
+\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
+\end{cases}
+\end{equation}
+
+where \(x_i - x_0 > 0\). The location and scale-invariant Yeo-Johnson
+transformation is:
+
+\begin{equation}
+\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
+\begin{cases}
+\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
+\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
+-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
+-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
+\end{cases}
+\end{equation}
+
+The parameters of these power transformations can be optimised based by
+maximising the log-likelihood function, under the assumption that the
+transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) follows a
+normal distribution. The log-likelihood functions for conventional
+Box-Cox and Yeo-Johnson transformations are well-known. However, the
+introduction of scaling parameter \(s\) prevents their direct use. Here,
+we first derive the general form of the log-likelihood functions, and
+then derive their power-transformation specific definitions.
+
+Let \(f(x_1, \ldots, x_n)\) be the probability density function of
+feature \(\mathbf{X} = \{ x_1, \ldots, x_n\}\), and
+\(f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))\)
+be the probability density function of the transformed feature
+\(\phi^{\lambda, x_0, s} (\mathbf{X})\), that is assumed to follow a
+normal distribution.
+
+The two probability density functions are related as follows:
+
+\begin{equation}
+f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
+\end{equation}
+
+Where, \(\left|\mathbf{J}\right|\) is the determinant of Jacobian
+\(\mathbf{J}\). The Jacobian takes the following form, with off-diagonal
+elements \(0\):
+
+\begin{equation}
+\mathbf{J} =
+\begin{bmatrix}
+    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
+    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
+    \vdots & \vdots  & \ddots &  \vdots \\
+    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
+\end{bmatrix}
+\end{equation}
+
+Thus,
+\(\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\).
+
+Since in our situation \(\{x_1, \ldots, x_n\}\) in
+\(f^{\lambda, x_0, s}(x_1, \ldots, x_n)\) are considered fixed (i.e.,
+known), \(f^{\lambda, x_0, s}(x_1, \ldots, x_n)\) may be considered a
+likelihood function. The log-likelihood function
+\(\mathcal{l}^{\lambda, x_0, s}\) is then:
+
+\begin{equation}
+\begin{split}
+\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
+ & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
+ & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
+\end{split}
+\end{equation}
+
+With \(\mu\) the average of \(\phi^{\lambda, x_0, s}(\mathbf{X})\) and
+\(\sigma^2\) its variance. The first two terms derive directly from the
+log-likelihood function of a normal distribution, and are not specific
+to the type of power transformation used. However, the final term
+differs between Box-Cox and Yeo-Johnson transformations.
+
+\subsection{Location- and scale-invariant Box-Cox
+transformation}\label{location--and-scale-invariant-box-cox-transformation}
+
+For the location- and scale-invariant Box-Cox transformation the partial
+derivative is:
+
+\begin{equation}
+\begin{split}
+\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
+ & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
+\end{split}
+\end{equation}
+
+Thus the final term in \(\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}\) is:
+
+\begin{equation}
+\begin{split}
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
+& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
+& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+
+This leads to the following log-likelihood:
+
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+
+Similarly to Raymaekers and Rousseeuw (2024), sample weights \(w_i\) are
+introduced to facilitate robust power transformations. The weighted
+log-likelihood of the location- and scale-invariant Box-Cox
+transformation is:
+
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
+\end{split}
+\end{equation}
+
+where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
+variance of the Box-Cox transformed feature
+\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\), respectively:
+
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+
+\subsection{Location- and scale-invariant Yeo-Johnson
+transformation}\label{location--and-scale-invariant-yeo-johnson-transformation}
+
+For the location- and scale-invariant Yeo-Johnson transformation, the
+partial derivative is:
+
+\begin{equation}
+\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
+\begin{cases}
+\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
+\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
+\end{cases}
+\end{equation}
+
+Thus the final term in \(\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}\) is:
+
+\begin{equation}
+\begin{split}
+\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+
+This leads to the following log-likelihood:
+
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+
+The weighted log-likelihood for location- and scale-invariant
+Yeo-Johnson transformation is:
+
+\begin{equation}
+\begin{split}
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\end{split}
+\end{equation}
+
+where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
+variance of the Yeo-Johnson transformed feature
+\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):
+
+\begin{equation}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation}
+
+\section{Optimisation of transformation parameters}\label{appendix-b-optimisation-of-transformation-parameters}
+
+Maximum likelihood estimation (MLE) is commonly used to optimise
+parameters for power transformation. Generally, optimisation requires
+minimisation or maximisation of a criterion. In MLE, the maximised
+criterion is the log-likelihood function of the normal distribution.
+Here, we investigate power transformation using optimisation criteria
+that are closely related to test statistics for normality tests.
+
+Let \(\mathbf{X}\) be a feature with ordered feature values, and
+\(\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)\) and
+\(\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)\)
+its transformed values using conventional and shift and scale invariant
+power transformations, respectively. Since power transformations are
+monotonic, \(\mathbf{Y}\) will likewise be ordered.
+
+Below we will focus on criteria based on the empirical density function
+and those based on skewness and kurtosis of the transformed featured.
+Other potential criteria, such as the Shapiro-Wilk test statistic
+(Shapiro and Wilk 1965) are not investigated here. In the case of the
+Shapiro-Wilk test statistic this is because of lack of scalability to
+features with many (\(> 5000\)) instances, and because adapting the test
+statistic to include weights is not straightforward.
+
+\subsection{Empirical density function-based
+criteria}\label{empirical-density-function-based-criteria}
+
+The first class of criteria is based on the empirical distribution
+function (EDF). Transformation parameters are then fit through
+minimisation of the distance between the empirical distribution function
+\(F_{\epsilon}\) and the cumulative density function (CDF) of the normal
+distribution \(F_{\mathcal{N}}\). Let
+\(F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}\) be the
+empirical probability of instance \(i\). The normal distribution is
+parametrised by location parameter \(\mu\) and scale parameter
+\(\sigma\), both of which have to be estimated from the data. For
+non-robust power transformations, \(\mu\) and \(\sigma\) are sample mean
+and sample standard deviation, respectively. For robust power
+transformations, we estimate \(\mu\) and \(\sigma\) as Huber M-estimates
+of location and scale of the transformed feature
+\(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981).
+
+\subsubsection{Anderson-Darling
+criterion}\label{anderson-darling-criterion}
+
+The Anderson-Darling criterion is based on the empirical distribution
+function of \(\mathbf{X}\). We define this criterion as follows:
+
+\begin{equation}
+U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right) }
+\end{equation}
+
+Here \(w_i\) are weights, and \(\mu\) and \(\sigma\) are location and
+scale parameters. For non-robust power transformations, all \(w_i = 1\).
+Note that this criterion is not the same as the Anderson-Darling test
+statistic (Anderson and Darling 1952), which involves solving (or
+approximating) an integral function, contains an extra scalar
+multiplication term, and does not include weights. The Anderson-Darling
+criterion seeks to minimise the squared Euclidean distance between the
+EDF and the normal CDF, with differences at the upper and lower end of
+the normal CDF receiving more weight than those at the the centre of the
+CDF.
+
+\subsubsection{Cramér-von Mises
+criterion}\label{cramuxe9r-von-mises-criterion}
+
+The Cramér-von Mises criterion is also based on the empirical
+distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
+criterion as follows:
+
+\begin{equation}
+U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2
+\end{equation}
+
+Here \(w_i\) are weights, and \(\mu\) and \(\sigma\) are location and
+scale parameters. For non-robust power transformations, all \(w_i = 1\).
+The criterion is similar to the Cramér-von Mises test statistic Mises
+(1928), aside from a additive scalar value and the introduction of
+weights. This criterion, like the Anderson-Darling criterion, seeks to
+minimise the squared Euclidean distance between the EDF and the normal
+CDF. Unlike the Anderson-Darling criterion, this criterion weights all
+instances equally.
+
+For conventional power transformations with a fixed shift parameter, the
+transformation \(\phi^{\lambda, x_0, s} (\mathbf{X})\) may be
+substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
+Cramér-von Mises criterion.
+
+\subsection{Skewness-kurtosis-based
+criteria}\label{skewness-kurtosis-based-criteria}
+
+The second class of criteria seeks to reduce skewness and (excess)
+kurtosis of the transformed feature \(\mathbf{Y}\). We will first define
+the location \(\mu\) and scale \(\sigma\) of the the transformed as
+these are required for computing skewness and kurtosis. Here, \(\mu\) is
+defined as:
+
+\begin{equation}
+\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0, s} \left(x_i \right)} {\sum_{i=1}^n w_i}
+\end{equation}
+
+The location, or mean, is weighted using weights \(w_i\). For non-robust
+transformations, \(w_i = 1\). Then, \(\sigma^2\) is defined as:
+
+\begin{equation}
+\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
+\end{equation}
+
+Skewness is defined as:
+
+\begin{equation}
+s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
+\end{equation}
+
+Kurtosis is defined as:
+
+\begin{equation}
+k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
+\end{equation}
+
+\subsubsection{D'Agostino criterion}\label{dagostino-criterion}
+
+The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
+test statistic (D'Agostino and Belanger 1990). This test statistic is
+composed of two separate test statistics, one of which is related to
+skewness, and the other to kurtosis. Both test statistics are computed
+in several steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for
+non-robust power transformations, \(\nu = n\).
+
+For the skewness test statistic we first compute (D'Agostino and
+Belanger 1990):
+
+\begin{equation}
+\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
+\end{equation}
+
+\begin{equation}
+\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
+\end{equation}
+
+\begin{equation}
+\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
+\end{equation}
+
+\begin{equation}
+\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
+\end{equation}
+
+The skewness test statistic is then:
+
+\begin{equation}
+Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
+\end{equation}
+
+For the kurtosis test statistic we first compute (D'Agostino and
+Belanger 1990; Anscombe and Glynn 1983):
+
+\begin{equation}
+\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
+\end{equation}
+
+\begin{equation}
+\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
+\end{equation}
+
+\begin{equation}
+\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
+\end{equation}
+
+\begin{equation}
+\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
+\end{equation}
+
+\begin{equation}
+\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
+\end{equation}
+
+The kurtosis test statistic is then:
+
+\begin{equation}
+Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
+\end{equation}
+
+The D'Agostino \(K^2\) test statistic and our criterion are the same,
+and are defined as:
+
+\begin{equation}
+U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
+\end{equation}
+
+The main difference between the test statistic as originally formulated,
+and the criterion proposed here is the presence of weights for robust
+power transformation.
+
+\subsubsection{Jarque-Bera criterion}\label{jarque-bera-criterion}
+
+The second criterion based on skewness and kurtosis is the Jarque-Bera
+criterion. It is relatively simple to compute compared to the D'Agostino
+criterion:
+
+\begin{equation}
+U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
+\end{equation}
+
+The main difference between the above criterion and the Jarque-Bera test
+statistic (Jarque and Bera 1980) is that a scalar multiplication is
+absent.
+
+\subsection{Optimisation using non-MLE
+criteria}\label{optimisation-using-non-mle-criteria}
+
+Each of the above criteria can be used for optimisation, i.e.:
+
+\begin{equation}
+\left\{ \hat{\lambda}, \hat{x}_0, \hat{s}_0 \right\} = \argmin_{\lambda, x_0, s} U\left(\mathbf{X}, \lambda, x_0, s \right)
+\end{equation}
+
+For conventional power transformations with fixed location and scale
+parameters, the transformation \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
+may be substituted by \(\phi^{\lambda} (\mathbf{X})\), or equivalently,
+\(x_0\) and \(s\) may be fixed:
+
+\begin{equation}
+\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0, s \right)
+\end{equation}
+
+\section{Simulations with other optimisation criteria}\label{appendix-c-simulations-with-other-optimisation-criteria}
+
+Invariance of location- and scale-invariant power transformations was
+assessed using the optimisation criteria in section
+\ref{appendix-b-optimisation-of-transformation-parameters}. This follows the simulation in the main manuscript, where MLE was
+used for optimization. In short, we first randomly drew \(10000\) values
+from a normal distribution:
+\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
+or equivalently
+\(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)\).
+The second distribution was a right-skewed normal distribution
+\(\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)\).
+The third distribution was a left-skewed normal distribution
+\(\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)\).
+We then computed transformation parameter \(\lambda\) using the original
+definitions (equations \ref{eqn:box-cox-original} and
+\ref{eqn:yeo-johnson-original}) and the location- and scale-invariant
+definitions (equations \ref{eqn:box-cox-invariant} and
+\ref{eqn:yeo-johnson-invariant}) for each distribution using different
+optimisation criteria. To assess location invariance, a positive value
+\(d_{\text{shift}}\) was added to each distribution with
+\(d_{\text{shift}} \in [1, 10^6]\). Similarly, to assess scale
+invariance, each distribution was multiplied by a positive value
+\(d_{\text{scale}}\), where \(d_{\text{scale}} \in [1, 10^6]\).
+
+The results are shown in Figure
+\ref{fig:shifted-distributions-appendix}.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_appendix_1}
+
+}
+
+\caption{Invariant power transformation produces transformation parameters that are invariant to location and scale.
+Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\text{shift}}$ or multiplication by $d_{\text{scale}}$.
+Estimates of the transformation parameter $\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution and the optimisation criterion, whereas estimates obtained for the location- and scale-invariant power transformations are constant.
+For location- and scale-invariant power transformations, the Anderson-Darling criterion leads to unstable estimates of $\lambda$ for skewed distributions, possibly due to large weights being assigned to samples at the upper and lower ends of the distribution.}\label{fig:shifted-distributions-appendix}
+\end{figure}
+
+\section{Experimental results using location- and scale-invariant Box-Cox transformation}
+\label{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}
+
+The effect of using location- and scale-invariant transformations was
+investigated using real-world datasets.
+
+\subsection{Invariance}\label{invariance}
+
+Results for Box-Cox transformations of features without outliers are
+shown in Figure \ref{fig:experimental-results-invariance-appendix}.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_appendix_2}
+
+}
+
+\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row).
+Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).}\label{fig:experimental-results-invariance-appendix}
+\end{figure}
+
+\subsubsection{Age of patients with lung
+cancer}\label{age-of-patients-with-lung-cancer}
+
+Applying conventional and invariant Box-Cox transformations to age of
+patients with lung cancer (Loprinzi et al. 1994) yielded the following
+results: no transformation (sum of residuals with normal distribution
+\(\sum r_i = 16.5\)); conventional transformation (\(\lambda = 1.9\),
+\(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
+\(\sigma_{BC} = 0.4 \cdot 10^3\)); Raymaekers and Rousseeuw's robust
+adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
+\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\));
+location- and scale-invariant transformation (\(\lambda = 1.7\),
+\(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\), \(\sigma_{BC} = 0.8\)); and
+robust location- and scale-invariant transformation (\(\lambda = 1.5\),
+\(\sum r_i = 11.6\), \(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\)).
+
+Compared to location- and scale-invariant Yeo-Johnson transformations,
+the Box-Cox transformations do not reduce residuals compared to
+conventional variants.
+
+\subsubsection{Penguin body mass}\label{penguin-body-mass}
+
+Applying conventional and invariant Box-Cox transformations to the body
+mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
+following results: no transformation (residual sum \(\sum r_i = 48.0\));
+conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); location- and
+scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 27.3\),
+\(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\)); and robust location- and
+scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 25.2\),
+\(\mu_{BC} = 0.4\), \(\sigma_{BC} = 0.6\)).
+
+Just as for location- and scale-invariant Yeo-Johnson transformations,
+Box-Cox transformations produced a lower overall residual sum compared
+to their conventional counterparts. Similarly, conventional
+transformations led to low standard deviation \(\sigma_{YJ}\) of the
+body mass feature after transformation.
+
+\subsubsection{Latitude in the Ames housing
+dataset}\label{latitude-in-the-ames-housing-dataset}
+
+Applying conventional and invariant Box-Cox transformations to the
+latitude of houses in the Ames housing dataset (De Cock 2011) yielded
+the following results: no transformation (residual sum
+\(\sum r_i = 328\)); conventional transformation (\(\lambda = 62.1\),
+\(\sum r_i = 319\), \(\mu_{BC} = 1.1 \cdot 10^{99}\),
+\(\sigma_{BC} = 0.0 \cdot 10^{99}\)); Raymaekers and Rousseeuw's robust
+adaptation (\(\lambda = 96.0\), \(\sum r_i = 319\),
+\(\mu_{BC} = 6.2 \cdot 10^{153}\),
+\(\sigma_{BC} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
+transformation (\(\lambda = 1.9\), \(\sum r_i = 312\),
+\(\mu_{BC} = 2.3\), \(\sigma_{BC} = 0.9\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
+\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\)).
+
+Similar to conventional Yeo-Johnson transformations (non-robust and
+robust), conventional Box-Cox transformations had high values for the
+\(\lambda\) parameter, which could lead to numerical issues. Location-
+and scale-invariant Box-Cox transformations did not suffer from this
+issue.
+
+\subsection{Robustness against
+outliers}\label{robustness-against-outliers}
+
+Results for Box-Cox transformations of features with outliers are shown
+in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_appendix_3}
+
+}
+
+\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row).
+Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).
+Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness-appendix}
+\end{figure}
+
+\subsubsection{Fuel efficiency in the Top Gear
+dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
+
+The Top Gear dataset contains data on 297 vehicles, with outliers
+related highly fuel-efficient vehicles (Alfons 2021). Applying
+conventional and invariant Box-Cox transformations to the fuel
+consumption feature yielded the following results: no transformation
+(residual sum \(\sum r_i = 54\), \(p=0.76\)); conventional
+transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
+\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p=0.01\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
+\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.55\)); location- and
+scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
+\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p=0.02\)); and robust
+location- and scale-invariant transformation (\(\lambda = 1.1\),
+\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
+\(p=0.83\)).
+
+\subsubsection{Maximum arterial wall thickness in an ischemic stroke
+dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
+
+The ischemic stroke dataset contains historic data from 126 patients
+with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
+conventional and invariant Box-Cox transformations to the maximum
+arterial wall thickness feature yielded the following results: no
+transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
+conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
+\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
+\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.60\)); location- and
+scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
+\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.01\)); and robust
+location- and scale-invariant transformation (\(\lambda = 0.5\),
+\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
+\(p=0.35\)).
+
+\section{Empirical central normality test}\label{appendix-e-empirical-central-normality-test}
+
+The empirical central normality test was derived using data sampled from
+asymmetric generalised normal distributions, including outliers, to
+resemble more realistic datasets. Here we assess the type I error rate
+of two, less realistic, sets of data:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Data sampled from asymmetric generalised normal distributions without
+  outliers.
+\item
+  Data sampled from normal distributions without outliers, without any
+  power transformation applied.
+\end{enumerate}
+
+Other aspects of the experiment remained the same. Thus, we first drew
+\(m_d=10000\) random distributions. For asymmetric generalised normal
+distributions, each distribution was parametrised with a randomly chosen
+skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
+parameter \(\beta \sim U\left(1.00, 5.00 \right)\). For fully normal
+distributions, skewness parameter \(\alpha = 0.5\) and shape parameter
+\(\beta = 2.0\) were fixed. Location and scale parameters were set as
+\(\mu = 0\) and \(\sigma = 1\), respectively.
+\(n = \lceil 10^\gamma \rceil\) values were then randomly drawn, with
+\(\gamma \sim U\left(1.47, 3.00\right)\), which led to between \(30\)
+and \(1000\) values being drawn to create \(\mathbf{X}_i\). Residuals
+were then computed after performing robust location- and scale-invariant
+transformations with the empirical tapered cosine weighting method for
+the dataset with asymmetric generalised normal distributions, and
+without any transformation for the dataset with only normal
+distributions.
+
+\begin{figure}
+
+{\centering \includegraphics{figure_appendix_4}
+
+}
+
+\caption{Type I error rate as function of the test statistic $\tau_{\text{ecn}}$ for five datasets, with central portion $\kappa=0.80$.
+The type I error rate is computed from $m_d=10000$ randomly sampled features, These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none).
+The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none).}\label{fig:empirical-central-normality-test-appendix}
+\end{figure}
+
+\begin{table}
+\begin{center}
+\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
+\label{tab:empirical-central-normality-appendix}
+\begin{tabular}{l | c c c c c c c}
+
+\toprule
+type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+
+\midrule
+Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
+Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
+Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
+normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+The results are shown in Figure
+\ref{fig:empirical-central-normality-test-appendix} and Table
+\ref {tab:empirical-central-normality-appendix}. These indicate that the
+test behaves similarly for the different datasets. For low type I error
+rates, the test statistic proposed in the main manuscript is more
+conservative than alternatives based on residuals after Box-Cox
+transformations of asymmetric generalised normally distributed features
+or on residuals from strictly normally distributed features.
+
+\section{Normalisation before transformation}\label{appendix-f-normalisation-before-transformation}
+
+An alternative to location- and scale-invariant transformations is
+normalising feature distributions prior to conventional transformations.
+Table \ref{tab:normalisation-before-transformation-appendix} shows
+residual errors, after transformation to normality, of the five features
+from real-world datasets presented previously section \ref{experimental-results}
+and \ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.
+In these examples location- and scale-invariant
+transformations have similar or lower residual errors compared to errors
+resulting from normalisation prior to transformation.
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
+\label{tab:normalisation-before-transformation-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
+
+\midrule
+age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
+penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
+latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
+fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
+arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 

### 2024-10-18 manuscript_latex/manuscript.tex
Replaced cross-references to main manuscript and appendix

@@
 i.e.~using maximum likelihood estimation (MLE). A full derivation of the
-log-likelihood function for both transformations is shown in Appendix A.
+log-likelihood function for both transformations is shown in \ref{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}.
 The location- and scale-invariant Box-Cox log-likelihood function is:
@@
 values defined using robust location- and scale-invariant Yeo-Johnson
-transformations, as it is more conservative (see Appendix E). This leads
+transformations, as it is more conservative (see \ref{appendix-e-empirical-central-normality-test}). This leads
 to test statistic values listed in Table
@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in Appendix D.
+for Box-Cox transformations are shown in 
+\ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.
 
@@
 transformation -- location- and scale-invariant transformations seem to
-provide an overall better transformation to normality (Appendix F).
+provide an overall better transformation to normality 
+(\ref{appendix-f-normalisation-before-transformation}).
 
@@
 numerical stability issues for optimisation criteria other than MLE
-(Appendix B). These appear in regions where transformation parameters
+(\ref{appendix-b-optimisation-of-transformation-parameters}).
+These appear in regions where transformation parameters
 would lead to very large or small numbers when using conventional power
@@
 Invariance of location- and scale-invariant power transformations was
-assessed using the optimisation criteria in section
-\ref{appendix-b-optimisation-of-transformation-parameters}. This follows the simulation in the main manuscript, where MLE was
+assessed using the optimisation criteria in 
+\ref{appendix-b-optimisation-of-transformation-parameters}. This follows the simulation in section \ref{invariance-to-location-and-scale}, where MLE was
 used for optimization. In short, we first randomly drew \(10000\) values
@@
 test behaves similarly for the different datasets. For low type I error
-rates, the test statistic proposed in the main manuscript is more
+rates, the test statistic proposed in section \ref{empirical-central-normality-test-1} is more
 conservative than alternatives based on residuals after Box-Cox

### 2024-10-18 manuscript_latex/manuscript.tex
Update duplicate labels.

@@
 
-\subsection{Invariance}\label{invariance}
+\subsection{Invariance}\label{sec:invariance}
 
@@
 \subsubsection{Age of patients with lung
-cancer}\label{age-of-patients-with-lung-cancer}
+cancer}\label{sec:age-of-patients-with-lung-cancer}
 
@@
 
-\subsubsection{Penguin body mass}\label{penguin-body-mass}
+\subsubsection{Penguin body mass}\label{sec:penguin-body-mass}
 
@@
 \subsubsection{Latitude in the Ames housing
-dataset}\label{latitude-in-the-ames-housing-dataset}
+dataset}\label{sec:latitude-in-the-ames-housing-dataset}
 
@@
 \subsection{Robustness against
-outliers}\label{robustness-against-outliers}
+outliers}\label{sec:robustness-against-outliers}
 
@@
 \subsubsection{Fuel efficiency in the Top Gear
-dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
+dataset}\label{sec:fuel-efficiency-in-the-top-gear-dataset}
 
@@
 \subsubsection{Maximum arterial wall thickness in an ischemic stroke
-dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
+dataset}\label{sec:maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
 
@@
 
-\subsection{Invariance}\label{invariance}
+\subsection{Invariance}\label{sec-app:invariance}
 
@@
 \subsubsection{Age of patients with lung
-cancer}\label{age-of-patients-with-lung-cancer}
+cancer}\label{sec-app:age-of-patients-with-lung-cancer}
 
@@
 
-\subsubsection{Penguin body mass}\label{penguin-body-mass}
+\subsubsection{Penguin body mass}\label{sec-app:penguin-body-mass}
 
@@
 \subsubsection{Latitude in the Ames housing
-dataset}\label{latitude-in-the-ames-housing-dataset}
+dataset}\label{sec-app:latitude-in-the-ames-housing-dataset}
 
@@
 \subsection{Robustness against
-outliers}\label{robustness-against-outliers}
+outliers}\label{sec-app:robustness-against-outliers}
 
@@
 \subsubsection{Fuel efficiency in the Top Gear
-dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
+dataset}\label{sec-app:fuel-efficiency-in-the-top-gear-dataset}
 
@@
 \subsubsection{Maximum arterial wall thickness in an ischemic stroke
-dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
+dataset}\label{sec-app:maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
 

### 2024-10-18 manuscript_latex/manuscript.tex
Add citations in Bibtex format.

@@
 \usepackage{booktabs}
-
-%% The hyperref package allows for typesetting hyperlinks.
-%% \usepackage{hyperref}
 
@@
 aim to stabilise variance and improve normality of such features
-(Bartlett 1947; John W. Tukey 1957). The two most commonly used
-transformations are that of Box and Cox (1964) and Yeo and Johnson
-(2000). The Box-Cox transformation of a feature value \(x_i\) of feature
+\citep{Bartlett1947-rx, Tukey1957-rt}. The two most commonly used
+transformations are that of \citet{Box1964-mz} and \citet{Yeo2000-vw}.
+The Box-Cox transformation of a feature value \(x_i\) of feature
 \(\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}\) under the
@@
 approach is sensitive to outliers, and robust versions of Box-Cox and
-Yeo-Johnson transformations were devised (Raymaekers and Rousseeuw
-2024).
+Yeo-Johnson transformations were devised \citep{Raymaekers2024-zf}.
 
@@
 
-Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
-and Wilk 1965), could be automatically applied to transformed features.
+Statistical tests for normality, such as the Shapiro-Wilk test
+\citep{Shapiro1965-zd}, could be automatically applied to transformed features.
 However, given sufficiently large sample sizes such tests can detect
@@
 
-{\centering \includegraphics{figure_1}
+{\centering \includegraphics[width=1\linewidth]{figure_1}
 
@@
 transformations to normality, as shown in Figure
-\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
-(2024), the general aim of power transformations should be to transform
+\ref{fig:decreased-normality}. As indicated by \citet{Raymaekers2024-zf},
+the general aim of power transformations should be to transform
 non-outlier data to normality, i.e.~achieve \emph{central normality}. To
@@
 process requires identifying outliers in the data and weighting such
-instances during the optimisation process. Raymaekers and Rousseeuw
-(2024) achieve this through weighted maximum likelihood estimation.
+instances during the optimisation process. \citet{Raymaekers2024-zf}
+achieve this through weighted maximum likelihood estimation.
 However, because this procedure iteratively estimates and updates
@@
 \ref{eqn:box-cox-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
-and Rousseeuw (2024), these log-likelihood functions includes additional
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to \citet{Raymaekers2024-zf},
+these log-likelihood functions includes additional
 terms to accommodate estimation of \(x_0\) and \(s\). The weighted
@@
 \item
-  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
+  A tapered cosine function \citep{Tukey1967-eb}, with \(\delta_1 \geq 0\)
   and \(\delta_2 \geq \delta_1\) as threshold parameters:
@@
   The weighting function uses the z-score of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
-  Rousseeuw (2024),
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After \citep{Raymaekers2024-zf},
   \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
@@
   and scale of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981). Then
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) \citep{Huber1981-su}. Then
   \(\dot{x}_i = z_i\).
@@
 for use with unimodal distributions, the generalised normal distribution
-(Subbotin 1923; Nadarajah 2005) is a suitable option for simulating
+\citep{Subbotin1923-qk, Nadarajah2005-xe} is a suitable option for simulating
 realistic feature distributions. This distribution has the following
@@
 recipe for introducing skewness into the otherwise symmetric generalised
-normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
+normal distribution \citep{Gijbels2019-te}, leading to
 the following probability density function:
@@
 \(F_{\beta}^{-1}\) of the symmetric generalised normal distribution.
-\(F_{\beta}^{-1}\) was derived by Griffin to be (Griffin 2018):
+\(F_{\beta}^{-1}\) was derived by Griffin to be \citep{Griffin2018-bf}:
 
@@
 Deviations from normality can be detected by normality tests, such as
-the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
+the Shapiro-Wilk test \citep{Shapiro1965-zd}. In practice, normality
 tests may be too stringent with large sample sizes, outliers, or both.
@@
 scale of the transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
-(Huber 1981). The set of residual errors for the central portion of the
+\citep{Huber1981-su}. The set of residual errors for the central portion of the
 data is then
@@
 and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
-algorithm for derivative-free bound constraint optimisation (Powell
-2009) through maximum likelihood estimation. The required algorithms
+algorithm for derivative-free bound constraint optimisation \citep{Powell2009-zb}
+through maximum likelihood estimation. The required algorithms
 were implemented in the \texttt{power.transform} R software package
-(Zwanenburg and Löck 2024b) (version 1.0.0). Of note, the
+(version 1.0.0) \citep{Zwanenburg2024-kq}. Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 up to 10 percent of the values could be replaced by outliers. Outlier
-values were set according to John W. Tukey (1977), as follows. Let
+values were set according to \citet{Tukey1977-xm}, as follows. Let
 \(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
@@
 Minimisation was conducted using the BOBYQA algorithm for
-derivative-free bound constraint optimisation (Powell 2009). The
+derivative-free bound constraint optimisation \citep{Powell2009-zb}. The
 resulting weighting function parameters for weighted MLE are shown in
@@
 228 patients with lung cancer that was collected and published by
-Loprinzi et al. (Loprinzi et al. 1994). The age in the cohort was
+Loprinzi et al. \citep{Loprinzi1994-cd}. The age in the cohort was
 \(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
@@
 Gorman, Williams and Fraser recorded body mass (in grams) of 342
-penguins of three different species (Gorman, Williams, and Fraser 2014).
+penguins of three different species \citep{Gorman2014-eo}.
 The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
@@
 dataset contains data on 2930 properties that were sold between 2006 and
-2010 (De Cock 2011), including their geospatial coordinates. The
+2010 \citep{De-Cock2011-jf} including their geospatial coordinates. The
 latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
@@
 The Top Gear dataset contains data on 297 vehicles that appeared on the
-BBC television show \emph{Top Gear} (Alfons 2021). Within this dataset,
+BBC television show \emph{Top Gear} \citep{Alfons2021-kc}. Within this dataset,
 the fuel consumption feature contains outliers due to highly
@@
 The ischemic stroke dataset contains historic data from 126 patients
-with risk at ischemic stroke (Kuhn and Johnson 2019). These patients
+with risk at ischemic stroke \citep{Kuhn2019-kt}. These patients
 underwent Computed Tomography Angiography to characterize the carotid
@@
 We used 285 datasets from the Penn Machine Learning Benchmarks
-collection (Romano et al. 2022). In this collection, 122 datasets
+collection \citep{Romano2022-gq}. In this collection, 122 datasets
 correspond to regression tasks and 163 datasets to classification tasks.
-Using the familiar auto-machine learning library (Zwanenburg and Löck
-2024a) (version 1.5.0), each dataset was used to train a model for each
+Using the familiar auto-machine learning library (version 1.5.0) \citep{Zwanenburg2021-so},
+each dataset was used to train a model for each
 of 16 process configurations. Each process configuration specifies the
@@
 transformation method, we first fit a regression random forest (ranger
-package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, minimum
+package version 0.16.0, \citep{Wright2017-rf}): 2000 trees, minimum
 node size 2, other hyperparameters default) with process parameters and
@@
 \ref{eqn:box-cox-original} for the theoretical analysis in their paper
-(Box and Cox 1964), which subsequently became the convention. Yeo and
-Johnson's power transformation lacks a shift parameter altogether (Yeo
-and Johnson 2000). We showed that these power transformations are
+\citep{Box1964-mz}, which subsequently became the convention. Yeo and
+Johnson's power transformation lacks a shift parameter altogether 
+\citep{Yeo2000-vw}. We showed that these power transformations are
 sensitive to location and scale of data distributions. To mitigate this
@@
 normality differs from other tests for normality, such as the
-Shapiro-Wilk test (Shapiro and Wilk 1965), in that the test statistic is
+Shapiro-Wilk test \citep{Shapiro1965-zd}, in that the test statistic is
 independent of the number of samples. This makes this test more
@@
 
-Similarly to Raymaekers and Rousseeuw (2024), sample weights \(w_i\) are
+Similarly to \citet{Raymaekers2024-zf}, sample weights \(w_i\) are
 introduced to facilitate robust power transformations. The weighted
@@
 Other potential criteria, such as the Shapiro-Wilk test statistic
-(Shapiro and Wilk 1965) are not investigated here. In the case of the
+\citep{Shapiro1965-zd} are not investigated here. In the case of the
 Shapiro-Wilk test statistic this is because of lack of scalability to
@@
 of location and scale of the transformed feature
-\(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981).
+\(\phi^{\lambda, x_0, s} (\mathbf{X})\) \citep{Huber1981-su}.
 
@@
 Note that this criterion is not the same as the Anderson-Darling test
-statistic (Anderson and Darling 1952), which involves solving (or
+statistic \citep{Anderson1952-gz}, which involves solving (or
 approximating) an integral function, contains an extra scalar
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cramér-von Mises test statistic Mises
-(1928), aside from a additive scalar value and the introduction of
+The criterion is similar to the Cramér-von Mises test statistic \citep{Cramer1928-rc, Von_Mises1928-ef},
+aside from a additive scalar value and the introduction of
 weights. This criterion, like the Anderson-Darling criterion, seeks to
@@
 The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
-test statistic (D'Agostino and Belanger 1990). This test statistic is
+test statistic \citep{DAgostino1990-kp}. This test statistic is
 composed of two separate test statistics, one of which is related to
@@
 
-For the skewness test statistic we first compute (D'Agostino and
-Belanger 1990):
+For the skewness test statistic we first compute \citep{DAgostino1990-kp}:
 
@@
 
-For the kurtosis test statistic we first compute (D'Agostino and
-Belanger 1990; Anscombe and Glynn 1983):
+For the kurtosis test statistic we first compute \citep{Anscombe1983-nz, DAgostino1990-kp}:
 
@@
 The main difference between the above criterion and the Jarque-Bera test
-statistic (Jarque and Bera 1980) is that a scalar multiplication is
-absent.
+statistic \citep{Jarque1980-hw} is that a scalar multiplication is absent.
 
@@
 Applying conventional and invariant Box-Cox transformations to age of
-patients with lung cancer (Loprinzi et al. 1994) yielded the following
+patients with lung cancer \citep{Loprinzi1994-cd} yielded the following
 results: no transformation (sum of residuals with normal distribution
@@
 Applying conventional and invariant Box-Cox transformations to the body
-mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
+mass of penguins \citep{Gorman2014-eo} yielded the
 following results: no transformation (residual sum \(\sum r_i = 48.0\));
@@
 Applying conventional and invariant Box-Cox transformations to the
-latitude of houses in the Ames housing dataset (De Cock 2011) yielded
+latitude of houses in the Ames housing dataset \citep{De-Cock2011-jf} yielded
 the following results: no transformation (residual sum
@@
 The Top Gear dataset contains data on 297 vehicles, with outliers
-related highly fuel-efficient vehicles (Alfons 2021). Applying
+related highly fuel-efficient vehicles \citep{Alfons2021-kc}. Applying
 conventional and invariant Box-Cox transformations to the fuel
@@
 The ischemic stroke dataset contains historic data from 126 patients
-with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
+with risk at ischemic stroke \citep{Kuhn2019-kt}. Applying
 conventional and invariant Box-Cox transformations to the maximum
@@
 %%
-Example citation, See \citet{lamport94}.
+%% Example citation, See \citet{lamport94}.
 
@@
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-%% Refer following link for more details about bibliography and citations.
-%% https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management
-
-\begin{thebibliography}{00}
-
-%% For authoryear reference style
-%% \bibitem[Author(year)]{label}
-%% Text of bibliographic item
-
-\bibitem[Lamport(1994)]{lamport94}
-  Leslie Lamport,
-  \textit{\LaTeX: a document preparation system},
-  Addison Wesley, Massachusetts,
-  2nd edition,
-  1994.
-
-\end{thebibliography}
+\bibliographystyle{elsarticle-harv}
+\bibliography{refs}
+
 \end{document}

### 2024-10-18 manuscript_latex/manuscript.tex
Make figures fit in width.
figures were created for pages with smaller margins than elsarticle
Set width of figures.
Added abstract and highlights.

@@
 \begin{abstract}
-
+Power transformations are used to stabilize variance and achieve
+normality of features, especially in methods assuming normal 
+distributions such as ANOVA and linear discriminant analysis. 
+However, the commonly used Box-Cox and Yeo-Johnson power transformation 
+methods are sensitive to the location, scale, and presence of 
+outliers in the data.
+
+Here we present location- and scale-invariant Box-Cox and Yeo-Johnson
+transformations to mitigate these issues. We derive maximum likelihood
+estimation criteria for optimizing transformation parameters and propose
+robust adaptations that reduce the influence of outliers. 
+We also introduce an empirical test for assessing central normality
+of transformed features.
+
+In simulations and real-world datasets, robust location- and scale-invariant 
+transformations outperform conventional variants, resulting in better 
+transformations to central normality. In a machine learning experiment
+with 232 datasets with numerical features, integrating robust 
+ocation- and scale-invariant power transformations into an automated data 
+processing and machine learning pipeline did not result in a meaningful
+improvement or detriment in model performance compared to conventional variants.
+
+In conclusion, robust location- and scale-invariant power
+transformations can replace conventional variants.
 \end{abstract}
@@
 \begin{highlights}
-\item Research highlight 1
-\item Research highlight 2
+\item Location and scale of feature distributions may cause wrong power transformations.
+\item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
+\item Robust power transformations succesfully reduced negative effects of outliers.
+\item New invariant robust power transformations can replace conventional variants.
 \end{highlights}
@@
 
+power transformation \sep invariant transformation \sep robust transformation \sep normality test
 %% keywords here, in the form: keyword \sep keyword
@@
 
-{\centering \includegraphics{figure_2}
+{\centering \includegraphics[width=1\linewidth]{figure_2}
 
@@
 
-{\centering \includegraphics{figure_3}
+{\centering \includegraphics[width=1\linewidth]{figure_3}
 
@@
 
-{\centering \includegraphics{figure_4}
+{\centering \includegraphics[width=1\linewidth]{figure_4}
 
@@
 
-{\centering \includegraphics{figure_5}
+{\centering \includegraphics[width=1\linewidth]{figure_5}
 
@@
 
-{\centering \includegraphics{figure_6}
+{\centering \includegraphics[width=1\linewidth]{figure_6}
 
@@
 
-{\centering \includegraphics{figure_7}
+{\centering \includegraphics[width=1\linewidth]{figure_7}
 
@@
 
-{\centering \includegraphics{figure_8}
+{\centering \includegraphics[width=1\linewidth]{figure_8}
 
@@
 
-{\centering \includegraphics{figure_9}
+{\centering \includegraphics[width=1\linewidth]{figure_9}
 
@@
 
-{\centering \includegraphics{figure_10}
+{\centering \includegraphics[width=1\linewidth]{figure_10}
 
@@
 
-{\centering \includegraphics{figure_appendix_1}
+{\centering \includegraphics[width=1\linewidth]{figure_appendix_1}
 
@@
 
-{\centering \includegraphics{figure_appendix_2}
+{\centering \includegraphics[width=1\linewidth]{figure_appendix_2}
 
@@
 
-{\centering \includegraphics{figure_appendix_3}
+{\centering \includegraphics[width=1\linewidth]{figure_appendix_3}
 
@@
 
-{\centering \includegraphics{figure_appendix_4}
+{\centering \includegraphics[width=1\linewidth]{figure_appendix_4}
 
@@
 \bottomrule
+
 \end{tabular}
@@
 \toprule
-feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
+feature & none & conv. & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
 

### 2024-10-21 manuscript_latex/manuscript.tex
Add acknowledgements.
Fixed spelling error.

@@
 no normalisation and z-standardisation.
-z-standardisation is generally beneficial, but the estimated effect is neglible.
+z-standardisation is generally beneficial, but the estimated effect is negligible.
 The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method.
-There is no consistent behaviour, and estimated effects are neglible.
+There is no consistent behaviour, and estimated effects are negligible.
 Note that the ranges of the $x$-axes of the three main panels differ.
@@
 
+\section{Acknowledgements}\label{sec:acknowledgments}
+
+This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
+
 %% The Appendices part is started with the command \appendix;

### 2024-10-21 manuscript_latex/manuscript.tex
Fixed typo.

@@
 \caption{Type I error rate as function of the test statistic $\tau_{\text{ecn}}$ for five datasets, with central portion $\kappa=0.80$.
-The type I error rate is computed from $m_d=10000$ randomly sampled features, These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none).
+The type I error rate is computed from $m_d=10000$ randomly sampled features.
+These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none).
 The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none).}\label{fig:empirical-central-normality-test-appendix}

### 2024-10-21 manuscript_latex/manuscript.tex
Fixed spelling errors.

@@
 with 232 datasets with numerical features, integrating robust 
-ocation- and scale-invariant power transformations into an automated data 
+location- and scale-invariant power transformations into an automated data 
 processing and machine learning pipeline did not result in a meaningful
@@
 \item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
-\item Robust power transformations succesfully reduced negative effects of outliers.
+\item Robust power transformations successfully reduced negative effects of outliers.
 \item New invariant robust power transformations can replace conventional variants.
@@
 \caption{Weighting functions investigated in this study to make power transformations more robust against outliers.
-In this example, the step function was parameterised with $\delta_1 = 0.60$.
-The triangle and tapered cosine functions were both parameterised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
+In this example, the step function was parametrised with $\delta_1 = 0.60$.
+The triangle and tapered cosine functions were both parametrised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
 \end{figure}
@@
 shifts in location, distributions that due to location and scale are not
-centered near zero, or both. Earlier, we assessed these transformations
+centred near zero, or both. Earlier, we assessed these transformations
 using simulated data. In the following, they are evaluated using
@@
 random forests and generalised linear models (GLM).
-Random forests outperform GLM models for all task difficulties.
+Random forests outperform GLMs for all task difficulties.
 The top-right panel shows the marginal effect of feature normalisation methods, i.e.

### 2024-10-23 manuscript_latex/_manuscript.tex
Update to springer-nature format.

--- 2024-10-21 manuscript_latex/manuscript.tex
@@
 processing and machine learning pipeline did not result in a meaningful
-improvement or detriment in model performance compared to conventional variants.
+improvement or detriment to model performance compared to conventional variants.
 
@@
 \begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{L}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
@@
 \begin{split}
-\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 \begin{split}
-\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{L}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
@@
 \begin{split}
-\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{L}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 likelihood function. The log-likelihood function
-\(\mathcal{l}^{\lambda, x_0, s}\) is then:
+\(\mathcal{L}^{\lambda, x_0, s}\) is then:
 
@@
 \begin{split}
-\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
+\mathcal{L}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
  & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
@@
 
-Thus the final term in \(\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}\) is:
+Thus the final term in \(\mathcal{L}_{\text{BC}}^{\lambda, x_0, s}\) is:
 
@@
 \begin{split}
-\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{L}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
@@
 \begin{split}
-\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{L}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
@@
 
-Thus the final term in \(\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}\) is:
+Thus the final term in \(\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s}\) is:
 
@@
 \begin{split}
-\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 \begin{split}
-\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{L}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]

### 2024-10-23 manuscript_latex/manuscript.tex
Update to springer-nature format.

@@
-%%
-%% Copyright 2007-2024 Elsevier Ltd
-%%
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%%
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
-
-\documentclass[preprint,12pt,authoryear]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsmath package provides various useful equation environments.
-\usepackage{amsmath}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
-
-%% The booktabs package provides table makeup macros.
-\usepackage{booktabs}
-
-%% Custom math operators
+%Version 3 October 2023
+% See section 11 of the User Manual for version history
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
+
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
+
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
+
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%Note: the following reference styles support Namedate and Numbered referencing.
+By default the style follows the most common style.
+To switch between the options you can add or remove “Numbered” in the optional parenthesis.
+%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst%
+
+%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+%%\documentclass[sn-mathphys-num]{sn-jnl}% Math and Physical Sciences Numbered Reference Style 
+%%\documentclass[sn-mathphys-ay]{sn-jnl}% Math and Physical Sciences Author Year Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
+\documentclass[pdflatex,sn-apa]{sn-jnl}% APA Reference Style 
+%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+%% \usepackage{amsmath,amssymb,amsfonts}%
+%% \usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+%% \usepackage{algorithm}%
+%% \usepackage{algorithmicx}%
+%% \usepackage{algpseudocode}%
+\usepackage{listings}%
+%%%%
+
 \DeclareMathOperator*{\argmax}{argmax}
@@
 
-\journal{Computational Statistics and Data Analysis}
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published 
+%%%%  by Springer Nature. The guidance has been prepared in partnership with 
+%%%%  production teams to conform to Springer Nature technical requirements. 
+%%%%  Editorial and presentation requirements differ among journal portfolios and 
+%%%%  research disciplines. You may find sections in this template are irrelevant 
+%%%%  to your work and are empowered to omit any such section if allowed by the 
+%%%%  journal you intend to submit to. The submission guidelines and policies 
+%%%%  of the journal take precedence. A detailed User Manual is available in the 
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+%% as per the requirement new theorem styles can be included as shown below
+%%\theoremstyle{thmstyleone}%
+%%\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+%%\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+%%\theoremstyle{thmstyletwo}%
+%%\newtheorem{example}{Example}%
+%%\newtheorem{remark}{Remark}%
+
+%%\theoremstyle{thmstylethree}%
+%%\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
 
@@
 
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%            addressline={},
-%%            city={},
-%%            postcode={},
-%%            state={},
-%%            country={}}
-%% \fntext[label3]{}
-
-\title{Location and Scale-Invariant Power Transformations for Transforming Data to Normality} %% Article title
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-
-%% Alex Zwanenburg
-\author[affil_1,affil_2]{Alex Zwanenburg\corref{cor_note}}
-\ead{alexander.zwanenburg@nct-dresden.de}
-\cortext[cor_note]{corresponding author}
-
-\affiliation[affil_1]{
-  organization={
-    National Center for Tumor Diseases Dresden (NCT/UCC):
-    German Cancer Research Center (DKFZ), Heidelberg, Germany;
-    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, Dresden, Germany;
-    Helmholtz-Zentrum Dresden-Rossendorf (HZDR), Dresden, Germany},
-  addressline={Fetscherstraße 74/PF 64},
-  city={Dresden},
-  postcode={01307},
-  country={Germany}
+\title[Location and Scale-Invariant Power Transformations for Transforming Data to Normality]{Location and Scale-Invariant Power Transformations for Transforming Data to Normality}
+
+%%=============================================================%%
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% \author*[1,2]{\fnm{Joergen W.} \spfx{van der} \sur{Ploeg} 
+%%  \sfx{IV}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\author*[1,2]{\fnm{Alex} \sur{Zwanenburg}}\email{alexander.zwanenburg@nct-dresden.de}
+
+\author[2,3,4]{\fnm{Steffen} \sur{L{\"o}ck}}\email{steffen.loeck@oncoray.de}
+
+\affil*[1]{
+	\orgdiv{National Center for Tumor Diseases Dresden (NCT/UCC)},
+	\orgname{German Cancer Research Center (DKFZ), Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, Helmholtz-Zentrum Dresden-Rossendorf (HZDR)},
+	\orgaddress{\street{Fetscherstra{\ss}e 74/PF 64}, \city{Dresden}, \postcode{01307}, \country{Germany}}
 }
-\affiliation[affil_2]{
-  organization={
-    OncoRay – National Center for Radiation Research in Oncology,
-    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology,
+
+\affil[2]{
+	\orgdiv{OncoRay - National Center for Radiation Research in Oncology}
+	\orgname{Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology,
     Helmholtz-Zentrum Dresden-Rossendorf},
-  addressline={Fetscherstraße 74/PF 41},
-  city={Dresden},
-  postcode={01307},
-  country={Germany}
+	\orgaddress{\street{Fetscherstra{\ss}e 74/PF 41}, \city{Dresden}, \postcode{01307}, \country{Germany}}
 }
 
-%% Steffen Löck
-\author[affil_2,affil_3,affil_4]{Steffen Löck}
-
-\affiliation[affil_3]{
-  organization={
-    Department of Radiotherapy and Radiation Oncology,
-    Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology},
-  addressline={Fetscherstraße 74/PF 50},
-  city={Dresden},
-  postcode={01307},
-  country={Germany}
+\affil[3]{
+	\orgdiv{Department of Radiotherapy and Radiation Oncology},
+	\orgname{Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology},
+	\orgaddress{\street{Fetscherstra{\ss}e 74/PF 50}, \city{Dresden}, \postcode{01307}, \country{Germany}}
 }
-\affiliation[affil_4]{
-  organization={
-   German Cancer Consortium (DKTK), Partner Site Dresden,
-   and German Cancer Research Center (DKFZ)},
-  addressline={Im Neuenheimer Feld 280},
-  city={Heidelberg},
-  postcode={69192},
-  country={Germany}
+
+\affil[4]{
+	\orgdiv{German Cancer Consortium (DKTK), Partner Site Dresden},
+	\orgname{German Cancer Research Center (DKFZ)},
+	\orgaddress{\street{Im Neuenheimer Feld 280}, \city{Heidelberg}, \postcode{69192}, \country{Germany}}
 }
 
-%% Abstract
-% You are required to provide a concise and factual abstract which does not exceed
-% 250 words. The abstract should briefly state the purpose of your research,
-% principal results and major conclusions. Some guidelines:
-%
-% Abstracts must be able to stand alone as abstracts are often presented separately from the article.
-%
-% Avoid references. If any are essential to include, ensure that you cite the author(s) and year(s).
-%
-% Avoid non-standard or uncommon abbreviations. If any are essential to include,
-% ensure they are defined within your abstract at first mention.
-
-\begin{abstract}
-Power transformations are used to stabilize variance and achieve
+%%==================================%%
+%% Sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{Power transformations are used to stabilize variance and achieve
 normality of features, especially in methods assuming normal 
@@
 methods are sensitive to the location, scale, and presence of 
-outliers in the data.
-
-Here we present location- and scale-invariant Box-Cox and Yeo-Johnson
+outliers in the data. Here we present location- and scale-invariant Box-Cox and Yeo-Johnson
 transformations to mitigate these issues. We derive maximum likelihood
@@
 We also introduce an empirical test for assessing central normality
-of transformed features.
-
-In simulations and real-world datasets, robust location- and scale-invariant 
+of transformed features. In simulations and real-world datasets, robust location- and scale-invariant 
 transformations outperform conventional variants, resulting in better 
@@
 processing and machine learning pipeline did not result in a meaningful
-improvement or detriment in model performance compared to conventional variants.
-
+improvement or detriment to model performance compared to conventional variants.
 In conclusion, robust location- and scale-invariant power
-transformations can replace conventional variants.
-\end{abstract}
-
-%%Graphical abstract
-% \begin{graphicalabstract}
-% \includegraphics{grabs}
-% \end{graphicalabstract}
-
-%%Research highlights
-% Highlights are a short collection of bullet points that should capture the novel
-% results of your research as well as any new methods used during your study.
-% Highlights will help increase the discoverability of your article via search engines.
-% Some guidelines:
+transformations can replace conventional variants.}
+
+%%================================%%
+%% Sample for structured abstract %%
+%%================================%%
+
+% \abstract{\textbf{Purpose:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
 % 
-% Submit highlights as a separate editable file in the online submission system
-% with the word "highlights" included in the file name.
+% \textbf{Methods:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
 % 
-% Highlights should consist of 3 to 5 bullet points, each a maximum of 85 characters, including spaces.
-
-\begin{highlights}
-\item Location and scale of feature distributions may cause wrong power transformations.
-\item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
-\item Robust power transformations successfully reduced negative effects of outliers.
-\item New invariant robust power transformations can replace conventional variants.
-\end{highlights}
-
-%% Keywords
-% You are required to provide 1 to 7 keywords for indexing purposes. Keywords should be written in English.
-% Please try to avoid keywords consisting of multiple words (using "and" or "of").
-% We recommend that you only use abbreviations in keywords if they are firmly established in the field.
-\begin{keyword}
-
-power transformation \sep invariant transformation \sep robust transformation \sep normality test
-%% keywords here, in the form: keyword \sep keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
-
-\end{frontmatter}
-
-%% Add \usepackage{lineno} before \begin{document} and uncomment
-%% following line to enable line numbers
-%% \linenumbers
-
-%% main text
-%%
+% \textbf{Results:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.
+% 
+% \textbf{Conclusion:} The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+The abstract must not include subheadings (unless expressly permitted in the journal's Instructions to Authors), equations or citations.
+As a guide the abstract should not exceed 200 words.
+Most journals do not set a hard limit however authors are advised to check the author instructions for the journal they are submitting to.}
+
+\keywords{Data preprocessing, Feature transformation, Outliers, Power transformation}
+
+%%\pacs[JEL Classification]{D8, H51}
+
+%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}
+
+\maketitle
+
 \section{Introduction}\label{introduction}
@@
 \begin{table}
-\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
@@
 $r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
-\bottomrule
+\botrule
+
 \end{tabular}
-\end{center}
 \end{table}
@@
 \begin{table}
-\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
@@
 \end{tabular}
-\end{center}
 \end{table}
@@
 \begin{table}
-\begin{center}
 \caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
@@
 Loprinzi et al. \citep{Loprinzi1994-cd}. The age in the cohort was
-\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
+\(62.4 \pm 9.1\) (mean \(\pm\) standard deviation) years. Applying
 conventional and invariant Yeo-Johnson transformations to patient age
@@
 penguins of three different species \citep{Gorman2014-eo}.
-The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
+The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean \(\pm\) standard
 deviation) grams, and not centrally normal (\(p = 0.03\)). Applying
@@
 2010 \citep{De-Cock2011-jf} including their geospatial coordinates. The
-latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
+latitude was \(42.03 \pm 0.02)\) (mean \(\pm\) standard deviation). Applying
 conventional and invariant Yeo-Johnson transformations to latitude
@@
 
-\section{Data and code availability}\label{data-and-code-availability}
-
-Location- and scale-invariant power transformations were implemented in
+\backmatter
+
+\section*{Declarations}
+
+\begin{itemize}
+\item Funding: No funding was received for conducting this study.
+\item Competing interests: The authors have no relevant financial or non-financial interests to disclose.
+\item Ethics approval and consent to participate: Not applicable
+\item Consent for publication: Not applicable
+\item Data availability: Data and results for the machine learning experiment are available
+from Zenodo (https://doi.org/10.5281/zenodo.13736671). The manuscript was created using R Markdown with integrated code and is
+available from the \texttt{power.transform} GitHub repository, together with intermediate analysis results
+for computationally intensive experiments.
+\item Materials availability: Not applicable
+\item Code availability: Location- and scale-invariant power transformations were implemented in
 the \texttt{power.transform} package for R, which is available from
 GitHub (https://github.com/oncoray/power.transform) and the CRAN
-repository (https://cran.r-project.org/package=power.transform).
-The manuscript was created using R Markdown and is likewise
-available from the \texttt{power.transform} GitHub repository. Data and
-results for the machine learning experiment are separately available
-from Zenodo (https://doi.org/10.5281/zenodo.13736671).
-
-\section{Acknowledgements}\label{sec:acknowledgments}
-
-This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-\appendix
+repository (https://cran.r-project.org/package=power.transform). The manuscript was created using R Markdown with integrated code, and is available from the \texttt{power.transform} GitHub repository.
+\item Author contribution: Conceptualization: Alex Zwanenburg, Steffen L{\"o}ck; Methodology: Alex Zwanenburg; Formal analysis and investigation: Alex Zwanenburg; Writing - original draft preparation: Alex Zwanenburg; Writing - review and editing: Alex Zwanenburg, Steffen L{\"o}ck; Resources: Alex Zwanenburg, Steffen L{\"o}ck; Supervision: Steffen L{\"o}ck.
+\end{itemize}
+
+\begin{appendices}
+
 \section{Log-likelihood functions for location and scale
@@
 
-\subsubsection{Cramér-von Mises
-criterion}\label{cramuxe9r-von-mises-criterion}
-
-The Cramér-von Mises criterion is also based on the empirical
-distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
+\subsubsection{Cram{\'e}r-von Mises
+criterion}\label{cramer-von-mises-criterion}
+
+The Cram{\'e}r-von Mises criterion is also based on the empirical
+distribution function of \(\mathbf{X}\). We define the Cram{\'e}r-von Mises
 criterion as follows:
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cramér-von Mises test statistic \citep{Cramer1928-rc, Von_Mises1928-ef},
+The criterion is similar to the Cram{\'e}r-von Mises test statistic \citep{Cramer1928-rc, Von_Mises1928-ef},
 aside from a additive scalar value and the introduction of
@@
 substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
-Cramér-von Mises criterion.
+Cram{\'e}r-von Mises criterion.
 
@@
 \begin{table}
-\begin{center}
 \caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
@@
 \end{tabular}
-\end{center}
 \end{table}
 
-%% For citations use:
-%%       \citet{<label>} ==> Lamport (1994)
-%%       \citep{<label>} ==> (Lamport, 1994)
-%%
-%% Example citation, See \citet{lamport94}.
-
-%% If you have bib database file and want bibtex to generate the
-%% bibitems, please use
-%%
-%%  \bibliographystyle{elsarticle-harv}
-%%  \bibliography{<your bibdatabase>}
-
-\bibliographystyle{elsarticle-harv}
-\bibliography{refs}
+\end{appendices}
+
+%%===========================================================================================%%
+%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
+%% system, please include the references within the manuscript file itself. You may do this  %%
+%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
+%% file, and delete the associated \verb+\bibliography+ commands.                            %%
+%%===========================================================================================%%
+
+\bibliography{refs}% common bib file
+%% if required, the content of .bbl file can be included here once bbl is generated
+%%\input sn-article.bbl
 
 \end{document}
-
-\endinput
-%%
-%% End of file `elsarticle-template-harv.tex'.

### 2025-02-18 manuscript_rmarkdown/manuscript_appendix.Rmd
Revise appendix F.

@@
 An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
-Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors, after transformation to normality, of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
-In these examples location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
+We assessed two normalisation methods were used:
+1. z-standardisation, i.e. $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of feature $\mathbf{X}$; 
+and 2.
+robust scaling, i.e.
+$x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$, with $\text{IQR}$ representing the interquartile range of feature $\mathbf{X}$.
+
+Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors after transformation to normality of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
+Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$.
+Normalisation prior to transformation prevents unexpectedly high $\lambda$ values in the presence of large location shifts, as can be observed for the latitude feature.
+Normalisation prior to transformation, however, does not consistently improve residual errors after transformation. 
+Interestingly, use of normalisation prior to transformation with robust power transformation using the method by Raymaekers and Rousseeuw [@Raymaekers2024-zf] led to higher residual errors for arterial wall thickness, fuel efficiency and latitude features.
+This is similar to observations by @Raymaekers2024-zf, who observed that their method can emphasize values of outliers after transformation.
+Location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
 
@@
 data <- .assess_standardisation_before_transformation()
-data <- data[transformation %in% c("none", "conventional", "conventional (z-score norm.)", "conventional (robust scaling)", "invariant"), mget(c("residuals", "transformation", "dataset"))]
+data <- data[, mget(c("residuals", "lambda", "transformation", "dataset"))]
 data[, "residuals" := round(residuals, digits = 1L)]
-data.table::dcast(data, dataset ~ transformation, value.var = "residuals")
+data[, "lambda" := round(lambda, digits = 1L)]
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
 ```
+
 ```{=latex}
@@
 \begin{center}
-\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-appendix}
@@
 \toprule
-feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
+feature & age & AWT & FE & latitude & PBM \\
 
 \midrule
-age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
-penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
-latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
-fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
-arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
+none & 16.5 & 110.1	& 54.5 & 328.4 & 48.0 \\
+conventional & 11.5 & 30.0 & 55.3 & 319.0 & 32.2 \\
+conventional (z-score norm.) & 11.5 & 19.3 & 49.0 & 326.2 & 33.3 \\
+conventional (robust scaling) & 11.3 & 31.8 & 53.3 & 324.5 & 32.2 \\
+Raymaekers-Rousseeuw & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
+Raymaekers-Rousseeuw (z-score norm.) & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+invariant & 8.8 & 12.2 & 44.0 & 326.4 & 26.8 \\
+robust invariant & 10.6 & 27.1 & 49.5 & 310.6 & 23.1 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality.
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:normalisation-before-transformation-lambda-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+conventional & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.) & 1.2 & -1.7 & -0.1 & 1.3 & 0.6 \\
+conventional (robust scaling) & 1.3 & 0.0 & 0.2 & 1.4 & 0.4 \\
+Raymaekers-Rousseeuw & 2.0 & 1.1 & 0.8 & 95.4 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.) & 1.3 & 1.4 & 1.0 & 1.3 & 0.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.4 & 1.3 & 1.0 & 1.4 & 0.4 \\
+invariant & 1.3 & 0.2 & -1.3 & 1.5 & 0.5 \\
+robust invariant & 1.4 & -0.6 & 0.9 & 1.4 & 0.4 \\
 \bottomrule

### 2025-02-18 manuscript_rmarkdown/manuscript_appendix.Rmd
Add table with p-values for ecn test.

@@
 Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors after transformation to normality of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
-Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$.
+Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$, 
+\ref{tab:normalisation-before-transformation-p-value-appendix} shows p-values of the empirical central normality test.
 Normalisation prior to transformation prevents unexpectedly high $\lambda$ values in the presence of large location shifts, as can be observed for the latitude feature.
@@
 Location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
+Not surprisingly, normalisation before transformation does not noticeably improve central normality, as assessed using the empirical central normality test.
 
@@
 data <- .assess_standardisation_before_transformation()
-data <- data[, mget(c("residuals", "lambda", "transformation", "dataset"))]
+data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
 data[, "residuals" := round(residuals, digits = 1L)]
 data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
 data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
 data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
 ```
@@
 \begin{center}
-\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-appendix}
@@
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
@@
 
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:normalisation-before-transformation-p-value-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none & 0.685 & 0.558 & 0.764 & 0.354 & 0.030 \\
+conventional & 0.886 & 0.010 & 0.012 & 0.317 & 0.381 \\
+conventional (z-score norm.) & 0.842 & 0.036 & 0.011 & 0.075 & 0.378 \\
+conventional (robust scaling) & 0.855 & 0.002 & 0.007 & 0.077 & 0.416 \\
+Raymaekers-Rousseeuw & 0.886 & 0.610 & 0.553 & 0.296 & 0.381 \\
+Raymaekers-Rousseeuw (z-score norm.) & 0.685 & 0.632 & 0.812 & 0.075 & 0.378 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.826 & 0.609 & 0.813 & 0.077 & 0.434 \\
+invariant & 0.907 & 0.134 & 0.026 & 0.049 & 0.549 \\
+robust invariant & 0.781 & 0.102 & 0.583 & 0.085 & 0.717 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
 # References

### 2025-02-18 manuscript_rmarkdown/manuscript_appendix.tex
Add table with p-values for ecn test.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2024-10-17}
+\date{2025-02-18}
 
@@
 normalising feature distributions prior to conventional transformations.
+We assessed two normalisation methods were used: 1. z-standardisation,
+i.e.~\(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
+and \(\sigma\) the mean and standard deviation of feature
+\(\mathbf{X}\); and 2. robust scaling,
+i.e.~\(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
+with \(\text{IQR}\) representing the interquartile range of feature
+\(\mathbf{X}\).
+
 Table \ref{tab:normalisation-before-transformation-appendix} shows
-residual errors, after transformation to normality, of the five features
+residual errors after transformation to normality of the five features
 from real-world datasets presented previously in Appendix D and the main
-manuscript. In these examples location- and scale-invariant
-transformations have similar or lower residual errors compared to errors
-resulting from normalisation prior to transformation.
+manuscript. Table
+\ref{tab:normalisation-before-transformation-lambda-appendix} shows the
+corresponding values for transformation parameter \(\lambda\),
+\ref{tab:normalisation-before-transformation-p-value-appendix} shows
+p-values of the empirical central normality test. Normalisation prior to
+transformation prevents unexpectedly high \(\lambda\) values in the
+presence of large location shifts, as can be observed for the latitude
+feature. Normalisation prior to transformation, however, does not
+consistently improve residual errors after transformation.
+Interestingly, use of normalisation prior to transformation with robust
+power transformation using the method by Raymaekers and Rousseeuw
+(Raymaekers and Rousseeuw 2024) led to higher residual errors for
+arterial wall thickness, fuel efficiency and latitude features. This is
+similar to observations by Raymaekers and Rousseeuw (2024), who observed
+that their method can emphasize values of outliers after transformation.
+Location- and scale-invariant transformations have similar or lower
+residual errors compared to errors resulting from normalisation prior to
+transformation. Not surprisingly, normalisation before transformation
+does not noticeably improve central normality, as assessed using the
+empirical central normality test.
 
@@
 \begin{center}
-\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-appendix}
@@
 \toprule
-feature & none & conventional & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
+feature & age & AWT & FE & latitude & PBM \\
 
 \midrule
-age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
-penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
-latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
-fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
-arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
+none & 16.5 & 110.1 & 54.5 & 328.4 & 48.0 \\
+conventional & 11.5 & 30.0 & 55.3 & 319.0 & 32.2 \\
+conventional (z-score norm.) & 11.5 & 19.3 & 49.0 & 326.2 & 33.3 \\
+conventional (robust scaling) & 11.3 & 31.8 & 53.3 & 324.5 & 32.2 \\
+Raymaekers-Rousseeuw & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
+Raymaekers-Rousseeuw (z-score norm.) & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+invariant & 8.8 & 12.2 & 44.0 & 326.4 & 26.8 \\
+robust invariant & 10.6 & 27.1 & 49.5 & 310.6 & 23.1 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:normalisation-before-transformation-lambda-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+conventional & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.) & 1.2 & -1.7 & -0.1 & 1.3 & 0.6 \\
+conventional (robust scaling) & 1.3 & 0.0 & 0.2 & 1.4 & 0.4 \\
+Raymaekers-Rousseeuw & 2.0 & 1.1 & 0.8 & 95.4 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.) & 1.3 & 1.4 & 1.0 & 1.3 & 0.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.4 & 1.3 & 1.0 & 1.4 & 0.4 \\
+invariant & 1.3 & 0.2 & -1.3 & 1.5 & 0.5 \\
+robust invariant & 1.4 & -0.6 & 0.9 & 1.4 & 0.4 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:normalisation-before-transformation-p-value-appendix}
+\begin{tabular}{l | c c c c c}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none & 0.685 & 0.558 & 0.764 & 0.354 & 0.030 \\
+conventional & 0.886 & 0.010 & 0.012 & 0.317 & 0.381 \\
+conventional (z-score norm.) & 0.842 & 0.036 & 0.011 & 0.075 & 0.378 \\
+conventional (robust scaling) & 0.855 & 0.002 & 0.007 & 0.077 & 0.416 \\
+Raymaekers-Rousseeuw & 0.886 & 0.610 & 0.553 & 0.296 & 0.381 \\
+Raymaekers-Rousseeuw (z-score norm.) & 0.685 & 0.632 & 0.812 & 0.075 & 0.378 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.826 & 0.609 & 0.813 & 0.077 & 0.434 \\
+invariant & 0.907 & 0.134 & 0.026 & 0.049 & 0.549 \\
+robust invariant & 0.781 & 0.102 & 0.583 & 0.085 & 0.717 \\
 \bottomrule

### 2025-02-19 manuscript_rmarkdown/simulations.R
WIP on simulating special cases.

@@ -2298,3 +2298,429 @@
 
   return(data)
 }
+
+
+
+.get_transformation_parameters_special_cases <- function(
+    manuscript_dir,
+    data_generator
+) {
+  file_name <- file.path(
+    manuscript_dir,
+    paste0("robustness_comparison_special_cases_", data_generator, ".RDS")
+  )
+
+  # transformer generator ------------------------------------------------------
+  experiment_args <- coro::generator(
+    function(
+      transformation_methods = c("box_cox", "yeo_johnson")
+    ) {
+      for (transformation_method in transformation_methods) {
+        # Conventional transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "robust" = FALSE,
+          "invariant" = FALSE,
+        )
+
+        coro::yield(list(
+          "name" = "conventional",
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "fun_args" = fun_args
+        ))
+
+        # Invariant transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "robust" = FALSE,
+          "invariant" = TRUE
+        )
+
+        coro::yield(list(
+          "name" = "non-robust",
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "fun_args" = fun_args
+        ))
+
+        # Robust invariant transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "robust" = TRUE,
+          "invariant" = TRUE,
+        )
+
+        coro::yield(list(
+          "name" = "robust",
+          "method" = transformation_method,
+          "estimation_method" = "mle",
+          "fun_args" = fun_args
+        ))
+      }
+    }
+  )
+
+
+  data_generator <- coro::generator(
+    function(
+      name,
+      transformation_method,
+      generating_function = "ragn",
+      mu = NULL,
+      sigma = NULL,
+      n_samples = NULL,
+      alpha = NULL,
+      beta = NULL,
+      lambda = NULL,
+      n_distrs = 100L,
+      n_outliers = 10L,
+      side = "both"
+    ) {
+      if (is.null(mu)) {
+        mu <- (10^stats::runif(n_distrs, min = 0.0, max = 4.0) - 1.0) *
[…]

### 2025-02-26 manuscript_rmarkdown/manuscript.Rmd
WIP on simulating special cases.
Added abstract.

@@
 
-To-do
+Power transformations are used to stabilize variance and achieve normality of features, especially in methods assuming normal distributions such as ANOVA and linear discriminant analysis.
+However, the commonly used Box-Cox and Yeo-Johnson power transformation methods are sensitive to the location, scale, and presence of outliers in the data.
+Here we present location- and scale-invariant Box-Cox and Yeo-Johnson transformations to mitigate these issues.
+We derive maximum likelihood estimation criteria for optimizing transformation parameters and propose robust adaptations that reduce the influence of outliers.
+We also introduce an empirical test for assessing central normality of transformed features.
+In simulations and real-world datasets, robust location- and scale-invariant transformations outperform conventional variants, resulting in better transformations to central normality.
+In a machine learning experiment with 231 datasets with numerical features, 
+integrating robust location- and scale-invariant power transformations into an automated data processing and machine learning pipeline did not result in a meaningful improvement or detriment to model performance compared to conventional variants.
+In conclusion, robust location- and scale-invariant power transformations can replace conventional variants.
 
@@
 In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is more conservative (see Appendix E).
-This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}
+This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}.
 

### 2025-02-26 manuscript_rmarkdown/manuscript.Rmd
WIP on additional ML models.

@@
 
-We used 285 datasets from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
-In this collection, 122 datasets correspond to regression tasks and 163 datasets
+We used 231 datasets containing at least one numeric feature from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
+In this collection, 114 datasets correspond to regression tasks and 117 datasets
 to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
-each dataset was used to train a model for each of 16 process configurations.
-Each process configuration specifies the learner (generalised linear model or
+each dataset was used to train a model for each of 32 process configurations.
+Each process configuration specifies the learner (generalised linear model, L1-regularised linear models (Lasso), gradient boosted linear model, or
 random forest), transformation method (none, conventional Yeo-Johnson, robust
@@
 normality test (rejecting transformations with $p \leq 0.01$), and normalisation method (none,
-$z$-standardisation), yielding 16 distinct configurations.
+$z$-standardisation), yielding 32 distinct configurations.
 Before each experiment, each dataset was randomly split into a training (70%)
-and holdout test (30%) set five times. Thus, a total of 22800 models were
+and holdout test (30%) set five times. Thus, a total of 27720 models were
 created. Each model was then evaluated using the holdout test set using one of
@@
 
-* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (87 datasets)
-* easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (46 datasets)
-* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (72 datasets)
-* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (60 datasets)
-* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (12 datasets)
-* unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (8 datasets)
+* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (59 datasets)
+* easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (40 datasets)
+* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (50 datasets)
+* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (32 datasets)
+* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (49 datasets)
+* unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (1 dataset)
 
@@
 # Convert to data matrix. First aggregate within each experiment because the
-# Wilcoxon signed rank test is not for repeated measurements.
+# Friedman test is not for repeated measurements.
 data <- data.table::dcast(
@@
 
-h_learner <- stats::wilcox.test(
-    x = data[["random_forest_ranger"]],
-    y = data[["glm"]],
-    paired = TRUE,
-    alternative = "two.sided"
-)
+h_learner <- friedman.test(as.matrix(data[, "dataset" := NULL]))
+
+# h_learner <- stats::wilcox.test(
+#     x = data[["random_forest_ranger"]],
+#     y = data[["glm"]],
+#     paired = TRUE,
+#     alternative = "two.sided"
+# )
 ```
@@
 
-Here we focus on the subset of 232 datasets that contain numeric features. 
-Considering single process parameters, the choice of learner (Wilcoxon signed rank test: $p < 10^{-8})$) and normalisation method (Wilcoxon signed rank test: $p = 0.007$) had a significant impact (at significance level $p = 0.050$), but transformation method (Friedman test: $p = 0.054$) did not.
+Considering single process parameters, the choice of learner (Friedman test: $p < 10^{-8})$), normalisation method (Wilcoxon signed rank test: $p < 10^{-5}$), and transformation method $p = 0.006$, all had a significant impact (at significance level $p = 0.05$).
 

### 2025-02-26 manuscript_rmarkdown/plots.R

@@ -2426,6 +2426,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
   # Get ML experiment data.
   data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir, subset = subset)
 
+  # Drop single dataset classified as "unsolvable".
+  data <- data[task_difficulty != "unsolvable"]
+  data$task_difficulty <- droplevels(data$task_difficulty)
+
   # Build a random forest to create latent representations of the data.
   model_data <- data[, mget(c("learner", "transformation_method", "normalisation_method", "task_difficulty", "value_rank"))]
   model <- ranger::ranger(
@@ -2453,25 +2457,62 @@ get_annotation_settings <- function(ggtheme = NULL) {
     ordered=TRUE
   )
 
-  difficulty_levels <- c("all", "very_easy", "easy", "intermediate", "difficult", "very_difficult", "unsolvable")
-  difficulty_labels <- c("overall", "very easy", "easy", "intermediate", "difficult", "very difficult", "unsolvable")
+  # difficulty_levels <- c("all", "very_easy", "easy", "intermediate", "difficult", "very_difficult", "unsolvable")
+  # difficulty_labels <- c("overall", "very easy", "easy", "intermediate", "difficult", "very difficult", "unsolvable")
+  difficulty_levels <- c("all", "very_easy", "easy", "intermediate", "difficult", "very_difficult")
+  difficulty_labels <- c("overall", "very easy", "easy", "intermediate", "difficult", "very difficult")
 
   # Effect of learner ----------------------------------------------------------
   result_list <- list()
-  x <- predict(model, data = abstract_data[learner == "random_forest_ranger"])$predictions -
-    predict(model, data = abstract_data[learner == "glm"])$predictions
-  result_list <- c(result_list, list(data.table::data.table("value" = mean(x), "task_difficulty" = "all")))
+  learners <- levels(data$learner)
 
-  for (difficulty in levels(abstract_data$task_difficulty)) {
-    x <- predict(model, data = abstract_data[learner == "random_forest_ranger" & task_difficulty == difficulty])$predictions -
-      predict(model, data = abstract_data[learner == "glm" & task_difficulty == difficulty])$predictions
-    result_list <- c(result_list, list(data.table::data.table("value" = mean(x), "task_difficulty" = difficulty)))
+  for (learner_a in learners) {
+    for (learner_b in learners) {
+      if (learner_a == learner_b) next
+
[…]

### 2025-03-09 manuscript_rmarkdown/manuscript.Rmd
Update text for new ML experiments.

@@
 
-* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (59 datasets)
+* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (57 datasets)
 * easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (40 datasets)
-* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (50 datasets)
-* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (32 datasets)
-* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (49 datasets)
+* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (52 datasets)
+* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (33 datasets)
+* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (48 datasets)
 * unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (1 dataset)
@@
 
-Considering single process parameters, the choice of learner (Friedman test: $p < 10^{-8})$), normalisation method (Wilcoxon signed rank test: $p < 10^{-5}$), and transformation method $p = 0.006$, all had a significant impact (at significance level $p = 0.05$).
+Considering single process parameters, the choice of learner (Friedman test: $p < 10^{-8})$), normalisation method (Wilcoxon signed rank test: $p < 10^{-8}$), and transformation method $p = 0.007$, all had a significant impact (at significance level $p = 0.05$).
 
@@
 The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository.
-Data and results for the machine learning experiment are separately available from [Zenodo](https://doi.org/10.5281/zenodo.13736671).
+Data and results for the machine learning experiment are separately available from [Zenodo](https://doi.org/10.5281/zenodo.14986689).
 

### 2025-03-13 manuscript_rmarkdown/manuscript_appendix.Rmd
Work on empirical central normality test.

@@
 - \usepackage{amsmath}
+- \usepackage{amsthm}
 - \usepackage{booktabs}
@@
 
-The empirical central normality test was derived using data sampled from asymmetric generalised normal distributions, including outliers, to resemble more realistic datasets.
-Here we assess the type I error rate of two, less realistic, sets of data:
+\begin{definition}[empirical central normality test]
 
-1. Data sampled from asymmetric generalised normal distributions without outliers.
-2. Data sampled from normal distributions without outliers, without any power transformation applied.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n\right\} \in \mathbb{R}^n$ be a vector of length $n$,
+sorted so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each value in $\mathbf{X}$. For tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group.
+Let $z_i = F^{-1}_{\mathcal{N}}(p_i)$ be the corresponding z-score, derived from the quantile function of the normal distribution.
 
-Other aspects of the experiment remained the same.
-Thus, we first drew $m_d=10000$ random distributions.
-For asymmetric generalised normal distributions, each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-For fully normal distributions, skewness parameter $\alpha = 0.5$ and shape parameter $\beta = 2.0$ were fixed.
-Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
-$n = \lceil 10^\gamma \rceil$ values were then randomly drawn, with $\gamma \sim U\left(1.47, 3.00\right)$, which led to between $30$ and $1000$ values being drawn to create $\mathbf{X}_i$.
-Residuals were then computed after performing robust location- and scale-invariant transformations with the empirical tapered cosine weighting method for the dataset with asymmetric generalised normal distributions,
-and without any transformation for the dataset with only normal distributions.
+\end{definition}
 
-```{r empirical-central-normality-test-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Type I error rate as function of the test statistic $\\tau_{\\text{ecn}}$ ",
-  "for five datasets, with central portion $\\kappa=0.80$. ",
-  "The type I error rate is computed from $m_d=10000$ randomly sampled features, ",
-  "These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), ",
-  "or normal distributions with outliers (none). ",
-  "The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none)."
-)
-
-data <- .plot_type_1_error_rate_appendix(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
-data$p
-```
+Critical values for the test statistics of the empirical central normality test 
+are found in Table \ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 \begin{center}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
-\label{tab:empirical-central-normality-appendix}
-\begin{tabular}{l | c c c c c c c}
+\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n}$ for
+empirical central normality at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The presented values should be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
 
 \toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+$n$ \textbackslash $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
 
 \midrule
-Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
-Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
-Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
-Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
-normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
+    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
+   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
+   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
+   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
+  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
+  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
+  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
+ 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
+ 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
+ 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
+10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
 \bottomrule
@@
 ```
-
-The results are shown in Figure \ref{fig:empirical-central-normality-test-appendix} and Table \ref {tab:empirical-central-normality-appendix}.
-These indicate that the test behaves similarly for the different datasets.
-For low type I error rates, the test statistic proposed in the main manuscript is more conservative than alternatives 
-based on residuals after Box-Cox transformations of asymmetric generalised normally distributed features or on residuals from strictly normally distributed features.
 

### 2025-03-13 manuscript_rmarkdown/manuscript_appendix.tex
Work on empirical central normality test.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-02-18}
+\date{2025-03-13}
 
@@
 
-The empirical central normality test was derived using data sampled from
-asymmetric generalised normal distributions, including outliers, to
-resemble more realistic datasets. Here we assess the type I error rate
-of two, less realistic, sets of data:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Data sampled from asymmetric generalised normal distributions without
-  outliers.
-\item
-  Data sampled from normal distributions without outliers, without any
-  power transformation applied.
-\end{enumerate}
-
-Other aspects of the experiment remained the same. Thus, we first drew
-\(m_d=10000\) random distributions. For asymmetric generalised normal
-distributions, each distribution was parametrised with a randomly chosen
-skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
-parameter \(\beta \sim U\left(1.00, 5.00 \right)\). For fully normal
-distributions, skewness parameter \(\alpha = 0.5\) and shape parameter
-\(\beta = 2.0\) were fixed. Location and scale parameters were set as
-\(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) values were then randomly drawn, with
-\(\gamma \sim U\left(1.47, 3.00\right)\), which led to between \(30\)
-and \(1000\) values being drawn to create \(\mathbf{X}_i\). Residuals
-were then computed after performing robust location- and scale-invariant
-transformations with the empirical tapered cosine weighting method for
-the dataset with asymmetric generalised normal distributions, and
-without any transformation for the dataset with only normal
-distributions.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-test-appendix-1}
-
-}
-
-\caption{Type I error rate as function of the test statistic $\tau_{\text{ecn}}$ for five datasets, with central portion $\kappa=0.80$.
-The type I error rate is computed from $m_d=10000$ randomly sampled features, These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none).
-The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none).}\label{fig:empirical-central-normality-test-appendix}
-\end{figure}
-
 \begin{table}
 \begin{center}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
-\label{tab:empirical-central-normality-appendix}
-\begin{tabular}{l | c c c c c c c}
+\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n}$ for
+empirical central normality at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The presented values should be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
 
 \toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+$n$ \ $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
 
 \midrule
-Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
-Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
-Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
-Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
-normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
+    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
+   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
+   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
+   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
+  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
+  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
+  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
+ 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
+ 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
+ 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
+10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
 \bottomrule
@@
 \end{table}
-
-The results are shown in Figure
-\ref{fig:empirical-central-normality-test-appendix} and Table
-\ref {tab:empirical-central-normality-appendix}. These indicate that the
-test behaves similarly for the different datasets. For low type I error
-rates, the test statistic proposed in the main manuscript is more
-conservative than alternatives based on residuals after Box-Cox
-transformations of asymmetric generalised normally distributed features
-or on residuals from strictly normally distributed features.
 

### 2025-03-13 R/R/GoodnessOfFit.R

@@ -12,6 +12,8 @@ NULL
 #' distributed. The null-hypothesis is that the transformed distribution is
 #' centrally normal.
 #'
+#' This function is a wrapper around `ecn.test`.
+#'
 #' @param transformer A transformer object created using
 #'   `find_transformation_parameters`.
 #' @param verbose Sets verbosity of the fubction.
@@ -205,9 +207,10 @@ setMethod(
   x_0 <- x_interp[[x_name]][1L]
   y_0 <- x_interp[[y_name]][1L]
 
-  # Find nearest points to interpolation position..
-  x <- unique(sort(data[[x_name]]))
-  y <- unique(sort(data[[y_name]]))
+  # Find nearest points to interpolation position. Using native data.table
+  # routines is faster.
+  x <- unique(data[, mget(x_name)][order(get(x_name))])[[x_name]]
+  y <- unique(data[, mget(y_name)][order(get(y_name))])[[y_name]]
 
   # Lower and upper values of x.
   x_l <- tail(x[x < x_0], n = 1L)

### 2025-03-13 manuscript_rmarkdown/simulations.R

@@ -2846,20 +2846,46 @@
 
 
 
-.get_test_statistic_lookup_table <- function(manuscript_dir, k = 0.70) {
-  alpha_levels <- c(0.10, 0.20, 0.50, 0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
-  n <- c(5, 10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000)
+.get_test_statistic_lookup_table <- function(manuscript_dir, k = 0.70, for_manuscript = TRUE) {
+  if (for_manuscript) {
+    alpha_levels <- c(0.10, 0.20, 0.50, 0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
+    n <- c(5, 10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000)
+  } else {
+    alpha_levels <- c(
+      0.01, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30, 0.35, 0.40, 0.45, 0.50, 0.55,
+      0.60, 0.65, 0.70, 0.75, 0.80, 0.85, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95,
+      0.96, 0.97, 0.975, 0.98, 0.9825, 0.985, 0.9875, 0.99,
+      0.991, 0.992, 0.993, 0.994, 0.995, 0.996, 0.997, 0.998, 0.999)
+    n <- c(
+      5, 10, 15, 20, 30, 40, 50, 60, 70, 80, 90, 100, 120, 140, 160, 180, 200,
+      225, 250, 275, 300, 340, 380, 420, 460, 500,
+      560, 620, 680, 740, 800, 900, 1000, 1100, 1200, 1500,
+      2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000)
+  }
+
 
   data <- .get_test_statistics_data(manuscript_dir)
   data <- data[kappa == k]
+  data[, "kappa" := NULL]
 
   # Compute alpha
   data[, "alpha" := 1.0 - seq_len(.N)/.N, by = c("n")]
 
   # Define all combinations.
-  interp_list <- apply(expand.grid(list("alpha" = alpha_levels, "n" = n)), function(ii) as.list(ii), MARGIN = 1L, simplify = FALSE)
-  tau <- lapply(interp_list, power.transform:::.interpolate_2d, data = data)
+  new_data <- data.table::as.data.table(expand.grid(list("alpha" = alpha_levels, "n" = n)))
[…]

### 2025-03-14 manuscript_rmarkdown/manuscript_appendix.Rmd
Further progress on empirical central normality test.
Almost there :)

@@
 - \usepackage{booktabs}
+- \newtheorem{definition}{Definition}
 - \DeclareMathOperator*{\argmax}{argmax}
@@
 
-\begin{definition}[empirical central normality test]
+\begin{definition}[central portion of a sequence]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group, i.e.
+if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
+$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
+Furthermore, let central portion $\kappa \in (0,1)$.
 
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n\right\} \in \mathbb{R}^n$ be a vector of length $n$,
-sorted so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
-Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
-corresponding to each value in $\mathbf{X}$. For tied values in $\mathbf{X}$,
-percentile values are replaced by the average in their group.
-Let $z_i = F^{-1}_{\mathcal{N}}(p_i)$ be the corresponding z-score, derived from the quantile function of the normal distribution.
+Then the central portion of $\mathbf{X}$ is 
+$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
 
@@
 
-Critical values for the test statistics of the empirical central normality test 
+\begin{definition}[Residual errors of the central portion of a sequence]
+
+Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+
+Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
+$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Central normality of a sequence]
+
+The central portion of a sequence is normally distributed if the sum of residual
+errors is equal to zero:
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+
+\end{definition}
+
+In practice, a finite sequence sampled from a normal distribution $\mathcal{N}(\mu, \sigma)$ will have a non-zero sum of residual errors.
+
+\begin{definition}[Empirical central normality test]
+$\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+
+$\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+\end{definition}
+
+The null-hypothesis $\mathcal{H}_0$ is that the central portion of sequence $\mathbf{X}$
+is normally distributed, with the alternative hypothesis $\mathcal{H}_1$ that it
+is not normally distributed. To test this hypothesis, a test statistic is computed
+and compared against a critical value.
+
+\begin{definition}[Empirical central normality test statistic]
+Let $\alpha$ be the significance level for the empirical central normality test.
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+
+The test statistic is then $\tau_{\text{ecn}, \alpha, n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+
+\end{definition}
+
+Critical values for the test statistic of the empirical central normality test 
 are found in Table \ref{tab:empirical-central-normality-critical-statistic}. 
+The null-hypothesis should be rejected at significance level $\alpha$ if 
+$\tau_{\text{ecn}, \alpha, n, \kappa} \geq \tau_{\text{ecn}, \alpha, n, \kappa, \text{critical}}$.
+
+```{r include = FALSE, eval = FALSE}
+# This writes the lookup table for critical test statistic.
+load(file = "./R/sysdata.rda")
+
+gof_lookup_table <- .get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.70,
+  for_manuscript = FALSE
+)
+
+usethis::use_data(
+  gof_lookup_table, 
+  two_sided_function_parameters, 
+  internal = TRUE, 
+  overwrite = TRUE)
+```
+
+```{r include = FALSE, eval = FALSE}
+# This generates the table shown below.
+.get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.70,
+  for_manuscript = TRUE
+)
+
+```
 
@@
 \midrule
-    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
-   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
-   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
-   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
-  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
-  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
-  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
- 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
- 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
- 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
-10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
+5     & 25.28 & 22.17 & 17.46 & 13.15 & 10.48 & 8.32 & 6.62 & 4.93 & 2.24 \\
+10    & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 & 7.59 & 6.74 & 5.89 & 4.12 \\
+20    & 14.93 & 12.96 &  9.97 &  7.71 &  6.77 & 6.11 & 5.56 & 5.02 & 3.97 \\
+50    &  9.69 &  8.49 &  6.59 &  5.20 &  4.63 & 4.22 & 3.91 & 3.58 & 2.99 \\
+100   &  6.92 &  6.06 &  4.71 &  3.73 &  3.32 & 3.04 & 2.83 & 2.59 & 2.20 \\
+200   &  4.89 &  4.29 &  3.35 &  2.66 &  2.37 & 2.17 & 2.02 & 1.87 & 1.60 \\
+500   &  3.12 &  2.73 &  2.12 &  1.69 &  1.51 & 1.39 & 1.29 & 1.19 & 1.02 \\
+1000  &  2.19 &  1.92 &  1.50 &  1.20 &  1.07 & 0.98 & 0.92 & 0.85 & 0.73 \\
+2000  &  1.56 &  1.37 &  1.06 &  0.85 &  0.76 & 0.70 & 0.65 & 0.60 & 0.52 \\
+5000  &  0.99 &  0.86 &  0.67 &  0.54 &  0.48 & 0.44 & 0.41 & 0.38 & 0.32 \\
+10000 &  0.69 &  0.61 &  0.48 &  0.38 &  0.34 & 0.31 & 0.29 & 0.27 & 0.23 \\
 \bottomrule

### 2025-03-14 manuscript_rmarkdown/manuscript_appendix.tex
Further progress on empirical central normality test.
Almost there :)

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{booktabs}
+\newtheorem{definition}{Definition}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-03-13}
+\date{2025-03-14}
 
@@
 
+\begin{definition}[central portion of a sequence]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group, i.e.
+if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
+$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
+Furthermore, let central portion $\kappa \in (0,1)$.
+
+Then the central portion of $\mathbf{X}$ is 
+$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Residual errors of the central portion of a sequence]
+
+Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+
+Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
+$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Central normality of a sequence]
+
+The central portion of a sequence is normally distributed if the sum of residual
+errors is equal to zero:
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+
+\end{definition}
+
+In practice, a finite sequence sampled from a normal distribution
+\(\mathcal{N}(\mu, \sigma)\) will have a non-zero sum of residual
+errors.
+
+\begin{definition}[Empirical central normality test]
+$\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+
+$\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+\end{definition}
+
+The null-hypothesis \(\mathcal{H}_0\) is that the central portion of
+sequence \(\mathbf{X}\) is normally distributed, with the alternative
+hypothesis \(\mathcal{H}_1\) that it is not normally distributed. To
+test this hypothesis, a test statistic is computed and compared against
+a critical value.
+
+\begin{definition}[Empirical central normality test statistic]
+Let $\alpha$ be the significance level for the empirical central normality test.
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+
+The test statistic is then $\tau_{\text{ecn}, \alpha, n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+
+\end{definition}
+
+Critical values for the test statistic of the empirical central
+normality test are found in Table
+\ref{tab:empirical-central-normality-critical-statistic}. The
+null-hypothesis should be rejected at significance level \(\alpha\) if
+\(\tau_{\text{ecn}, \alpha, n, \kappa} \geq \tau_{\text{ecn}, \alpha, n, \kappa, \text{critical}}\).
+
 \begin{table}
@@
 \toprule
-$n$ \ $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
+$n$ \textbackslash \~ $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
 
 \midrule
-    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
-   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
-   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
-   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
-  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
-  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
-  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
- 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
- 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
- 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
-10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
+5     & 25.28 & 22.17 & 17.46 & 13.15 & 10.48 & 8.32 & 6.62 & 4.93 & 2.24 \\
+10    & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 & 7.59 & 6.74 & 5.89 & 4.12 \\
+20    & 14.93 & 12.96 &  9.97 &  7.71 &  6.77 & 6.11 & 5.56 & 5.02 & 3.97 \\
+50    &  9.69 &  8.49 &  6.59 &  5.20 &  4.63 & 4.22 & 3.91 & 3.58 & 2.99 \\
+100   &  6.92 &  6.06 &  4.71 &  3.73 &  3.32 & 3.04 & 2.83 & 2.59 & 2.20 \\
+200   &  4.89 &  4.29 &  3.35 &  2.66 &  2.37 & 2.17 & 2.02 & 1.87 & 1.60 \\
+500   &  3.12 &  2.73 &  2.12 &  1.69 &  1.51 & 1.39 & 1.29 & 1.19 & 1.02 \\
+1000  &  2.19 &  1.92 &  1.50 &  1.20 &  1.07 & 0.98 & 0.92 & 0.85 & 0.73 \\
+2000  &  1.56 &  1.37 &  1.06 &  0.85 &  0.76 & 0.70 & 0.65 & 0.60 & 0.52 \\
+5000  &  0.99 &  0.86 &  0.67 &  0.54 &  0.48 & 0.44 & 0.41 & 0.38 & 0.32 \\
+10000 &  0.69 &  0.61 &  0.48 &  0.38 &  0.34 & 0.31 & 0.29 & 0.27 & 0.23 \\
 \bottomrule

### 2025-03-14 R/R/GoodnessOfFit.R

@@ -44,8 +44,8 @@ assess_transformation <- function(
 
   if (is.na(h$p_value)) {
     message("p-value could not be determined.")
-  } else if (h$p_value == 0.0 && verbose) {
-    message("p-value is smaller than 10^-4")
+  } else if (h$p_value > 0.0001 && verbose) {
+    message("p-value is smaller than 0.0001")
   }
 
   return(h$p_value)
@@ -75,7 +75,7 @@ ecn.test <- function(x, transformer = NULL) {
   if (is.null(transformer)) {
     transformer <- find_transformation_parameters(x = x, method = "none")
   }
-
+  browser()
   # Compute residual data.
   residual_data <- get_residuals(
     x = x,
@@ -88,22 +88,28 @@ ecn.test <- function(x, transformer = NULL) {
   # Compute mean residual error for the central portion.
   test_statistic_value <- mean(abs(residual_data$residual))
 
-  if (test_statistic_value < min(gof_lookup_table$test_statistic)) {
-    p_value <- 1.0
+  # Lookup critical tau at alpha = 0.95.
+  test_statistic_critical <- .interpolate_2d(
+    list("n" = length(x), "alpha" = 0.95),
+    data = gof_lookup_table
+  )
 
-  } else if (test_statistic_value > gof_lookup_table[alpha == 0.0001]$test_statistic) {
-    p_value <- 0.0
+  # Lookup p-value.
+  p_value <- 1.0 - .interpolate_alpha(
+    list("n" = length(x), "tau" = test_statistic_value),
+    data = gof_lookup_table,
+    limited = TRUE
+  )
 
-  } else {
-    p_value <- stats::spline(
-      x = gof_lookup_table$test_statistic,
-      y = gof_lookup_table$alpha,
-      method = "hyman",
-      xout = test_statistic_value
-    )$y
+  if (length(x) < 5L) {
+    p_value <- NA_real_
   }
 
-  return(list("tau" = test_statistic_value, "p_value" = p_value))
+  return(list(
+    "tau" = test_statistic_value,
+    "tau_critical" = test_statistic_critical,
+    "p_value" = p_value
+  ))
 }
 
 
@@ -197,7 +203,7 @@ setMethod(
   # organised as a 3D grid. x_interp: named list with two elements of length 1.
   # limited: whether values outside the grid should be limited to the extremes
   # of the grid.
-
+browser()
   # Get names corresponding to x and y variables.
   x_name <- names(x_interp)[1L]
[…]

### 2025-03-14 manuscript_rmarkdown/simulations.R

@@ -1672,332 +1672,332 @@
 
 
 
-.get_test_statistics_data <- function(manuscript_dir, as_table = TRUE, reduce = FALSE) {
-  # Non-standard evaluation in data.table.
-  residual <- p <- n <- test_statistic <- method <- alpha <- NULL
-
-  ..down_sample_data <- function(x, alpha, upper, lower, n) {
-    # Find the x-values.
-    alpha_out <- mapply(
-      function(upper, lower, n) {
-        # Linear placement of interpolation points. The starting point (upper)
-        # is not included.
-        x <- rev(seq(from = lower, to = upper, length.out = n + 1L))
-        return(tail(x, n = n))
-      },
-      upper = upper,
-      lower = lower,
-      n = n,
-      SIMPLIFY = FALSE,
-      USE.NAMES = FALSE
-    )
-
-    alpha_out <- unlist(alpha_out, use.names = FALSE)
-
-    # Find the corresponding
-    x_out <- stats::spline(
-      x = alpha,
-      y = x,
-      method = "hyman",
-      xout = alpha_out
-    )$y
-
-    # Add start and end points.
-    x_out <- c(min(x), x_out, max(x))
-    alpha_out <- c(1.0, alpha_out, 0.0)
-
-    return(list("alpha" = alpha_out, "test_statistic" = x_out))
-  }
-
-  # Compute test statistic values.
-  data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
-
-  central_width <- 0.80
-
-  significance_values <- c(0.50, 0.20, 0.10, 0.05, 0.02, 0.01, 0.001)
-
-  p_lower <- 0.50 - central_width / 2
-  p_upper <- 0.50 + central_width / 2
-
-  # Clip empirical probabilities to the center.
-  data <- data[p >= p_lower & p <= p_upper]
-
-  # Compute mean absolute residual error per feature.
-  data <- data[, list("test_statistic" = mean(residual)), by = c("distribution_id", "outlier_id", "method")]
-  data <- data[, list("n" = .N), by = c("test_statistic", "method")][order(test_statistic, method)]
-  data[, "alpha" := 1.0 - cumsum(n) / sum(n), by = "method"]
-
-  if (reduce) {
-    # Reduce complexity of the data to save storage size.
-
-    # 5 points between 1.00 and 0.95
-    # 10 points between 0.95 and 0.75
-    # 10 points between 0.75 and 0.25
[…]

### 2025-03-14 manuscript_rmarkdown/manuscript_appendix.Rmd
Test works (technically)
However, alpha has been defined the wrong way around?

@@
 \begin{definition}[Empirical central normality test]
-$\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
 
-$\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+THe alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
 \end{definition}
@@
 \begin{center}
-\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n}$ for
+\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n, \kappa = 0.70}$ for
 empirical central normality at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The presented values should be divided by $100$.}
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
 \label{tab:empirical-central-normality-critical-statistic}

### 2025-03-18 manuscript_rmarkdown/manuscript_appendix.Rmd
Update central normality test.

@@
 
-\begin{definition}[Empirical central normality test]
+\begin{definition}[Central normality test]
 The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
@@
 
-\begin{definition}[Empirical central normality test statistic]
+\begin{definition}[Central normality test statistic]
 Let $\alpha$ be the significance level for the empirical central normality test.
@@
 
-The test statistic is then $\tau_{\text{ecn}, \alpha, n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+The test statistic is then $\tau_{\alpha, n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
 
@@
 
-Critical values for the test statistic of the empirical central normality test 
+Critical values for the test statistic of the central normality test 
 are found in Table \ref{tab:empirical-central-normality-critical-statistic}. 
 The null-hypothesis should be rejected at significance level $\alpha$ if 
-$\tau_{\text{ecn}, \alpha, n, \kappa} \geq \tau_{\text{ecn}, \alpha, n, \kappa, \text{critical}}$.
+$\tau_{\alpha, n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
 
@@
 
-gof_lookup_table <- .get_test_statistic_lookup_table(
+central_normality_lookup_table <- .get_test_statistic_lookup_table(
   manuscript_dir = manuscript_dir,
@@
 
+empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.70,
+  for_manuscript = FALSE,
+  with_outliers = TRUE
+)
+
 usethis::use_data(
-  gof_lookup_table, 
+  central_normality_lookup_table,
+  empirical_central_normality_lookup_table,
   two_sided_function_parameters, 
@@
 \begin{center}
-\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n, \kappa = 0.70}$ for
-empirical central normality at $\kappa = 0.70$, as a function of significance 
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+central normality at $\kappa = 0.70$, as a function of significance 
 level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:empirical-central-normality-critical-statistic}
+\label{tab:central-normality-critical-statistic}
 \begin{tabular}{l | r r r r r r r r r}
@@
 \midrule
-5     & 25.28 & 22.17 & 17.46 & 13.15 & 10.48 & 8.32 & 6.62 & 4.93 & 2.24 \\
-10    & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 & 7.59 & 6.74 & 5.89 & 4.12 \\
-20    & 14.93 & 12.96 &  9.97 &  7.71 &  6.77 & 6.11 & 5.56 & 5.02 & 3.97 \\
-50    &  9.69 &  8.49 &  6.59 &  5.20 &  4.63 & 4.22 & 3.91 & 3.58 & 2.99 \\
-100   &  6.92 &  6.06 &  4.71 &  3.73 &  3.32 & 3.04 & 2.83 & 2.59 & 2.20 \\
-200   &  4.89 &  4.29 &  3.35 &  2.66 &  2.37 & 2.17 & 2.02 & 1.87 & 1.60 \\
-500   &  3.12 &  2.73 &  2.12 &  1.69 &  1.51 & 1.39 & 1.29 & 1.19 & 1.02 \\
-1000  &  2.19 &  1.92 &  1.50 &  1.20 &  1.07 & 0.98 & 0.92 & 0.85 & 0.73 \\
-2000  &  1.56 &  1.37 &  1.06 &  0.85 &  0.76 & 0.70 & 0.65 & 0.60 & 0.52 \\
-5000  &  0.99 &  0.86 &  0.67 &  0.54 &  0.48 & 0.44 & 0.41 & 0.38 & 0.32 \\
-10000 &  0.69 &  0.61 &  0.48 &  0.38 &  0.34 & 0.31 & 0.29 & 0.27 & 0.23 \\
+5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
+10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
+20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
+50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
+100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
+200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
+500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
+1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
+2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
+5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
+10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
 \bottomrule
@@
 \end{table}
+```
+
+```{r include = FALSE, eval = FALSE}
+# This generates the table shown below.
+.get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.70,
+  for_manuscript = TRUE,
+  with_outliers = TRUE
+)
+
 ```

### 2025-03-18 R/R/GoodnessOfFit.R

@@ -40,7 +40,7 @@ assess_transformation <- function(
     ...) {
 
   # Perform test.
-  h <- ecn.test(x = x, transformer = transformer)
+  h <- cn.test(x = x, transformer = transformer)
 
   if (is.na(h$p_value)) {
     message("p-value could not be determined.")
@@ -53,19 +53,23 @@ assess_transformation <- function(
 
 
 
-#' Empirical central normality test
+#' Central normality test
 #'
-#' Assesses empirical central normality of input data
+#' Assesses central normality of input data.
 #'
 #' @param x vector of input data, of at least length 5.
 #' @param transformer A transformer object created using
 #'   `find_transformation_parameters`. Optional, if present residuals are
 #'   determined from x after transformation.
+#' @param robust Determines the use of the strict test for central normality
+#'   (`FALSE`) or the more robust version where test statistics where determined
+#'   from normally distributed data with 10% outliers (`TRUE`).
 #'
-#' @return list with mean absolute error (`tau`) and p-value (`p_value`) for
-#'   empirical central normality test.
+#' @return list with mean absolute error (`tau`), critical value (at
+#'   significance level = 0.95) of the test statistic (`tau_critical`) and
+#'   p-value (`p_value`) for (empirical) central normality test.
 #' @export
-ecn.test <- function(x, transformer = NULL) {
+cn.test <- function(x, transformer = NULL, robust = TRUE) {
 
   # Prevent CRAN NOTE due to non-standard use of variables by data.table.
   p_observed <- alpha <- NULL
@@ -82,6 +86,14 @@ ecn.test <- function(x, transformer = NULL) {
     transformer = transformer
   )
 
+  # Select look-up table.
+  if (!robust) {
+    lookup_table <- central_normality_lookup_table
+
+  } else {
+    lookup_table <- empirical_central_normality_lookup_table
+  }
+
   # The test uses a central portion kappa = 0.70.
   residual_data <- residual_data[p_observed >= 0.15 & p_observed <= 0.85]
 
@@ -91,14 +103,14 @@ ecn.test <- function(x, transformer = NULL) {
   # Lookup critical tau at alpha = 0.95.
   test_statistic_critical <- .interpolate_2d(
     list("n" = length(x), "alpha" = 0.95),
-    data = gof_lookup_table
+    data = lookup_table
   )
 
   # Lookup p-value.
   p_value <- 1.0 - .interpolate_alpha(
[…]

### 2025-03-18 manuscript_rmarkdown/manuscript_appendix.Rmd
Insert critical values for empirical central normality test.

@@
 \begin{definition}[Central normality test statistic]
-Let $\alpha$ be the significance level for the empirical central normality test.
+Let $\alpha$ be the significance level for the central normality test.
 Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
@@
 Critical values for the test statistic of the central normality test 
-are found in Table \ref{tab:empirical-central-normality-critical-statistic}. 
-The null-hypothesis should be rejected at significance level $\alpha$ if 
+are found in Table \ref{tab:central-normality-critical-statistic}. These are 
+determined by repeatedly and randomly sampling data (i.e. $\mathbf{X}$) from
+a normal distribution $\mathcal{N}(0,1)$. The null-hypothesis should be rejected
+at significance level $\alpha$ if 
 $\tau_{\alpha, n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
+
+The empirical central normality test is a central normality test, in the
+definition provided above. However, critical values of the test statistic are
+not determined directly from data drawn randomly from $\mathcal{N}(0,1)$.
+Rather, after $\mathbf{X}$ is drawn, 10% of the values are randomly replaced by
+random outlier values. The resulting distribution is centrally normal, but
+not fully normal.
+
+Critical values for the empirical variant of the central normality test are
+found in Table \ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 \caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-central normality at $\kappa = 0.70$, as a function of significance 
+\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
 level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
@@
 
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
+
+\midrule
+5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
+10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
+20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
+50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
+100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
+200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
+500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
+1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
+2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
+5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
+10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 ```

### 2025-03-21 manuscript_rmarkdown/manuscript.Rmd
Update p-values for empirical tests.

@@
 
-```{=latex}
 \begin{equation}
@@
 \end{equation}
-```
+
 Here, $\alpha \in (0,1)$ is a skewness parameter.
@@
 \end{equation}
-```
+
 The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}$ thus incorporates the quantile function $F_{\beta}^{-1}$ of the symmetric generalised normal distribution.
@@
 \end{equation}
-```
+
 Here, $F_{\Gamma}^{-1}$ is the quantile function of the gamma distribution with shape $1 / \beta$, which can be numerically approximated.
 
-## Empirical central normality test
+## Central normality test
 
@@
 In practice, normality tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop an empirical test for central normality.
-The null hypothesis $H_0$ is that the distribution is centrally normal.
-The alternative hypothesis $H_1$ is that the distribution is not centrally normal.
-
-Let central normality be defined as the normality of central portion $\kappa$ of the data, i.e.
-$\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$, with $p_i$ probabilities of the empirical distribution, as previously.
-We then compute the residual errors between the z-scores of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ and the expected z-scores from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].
-The set of residual errors for the central portion of the data is then $\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
-
-The test statistic $\tau_{\text{ecn}}$ is then defined as:
-```{=latex}
-\begin{equation}
-\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
-\end{equation}
-```
-Here $[\quad]$ denotes an Iverson bracket.
-The test statistic is equal to the mean of the residual errors of the central portion of the data.
-
-To use the test statistic, the central portion of the data needs to be defined and the Type 1 error rates determined.
+Here we develop a test for central normality.
+
+\begin{definition}[central portion of a sequence]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group, i.e.
+if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
+$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
+Furthermore, let central portion $\kappa \in (0,1)$.
+
+Then the central portion of $\mathbf{X}$ is 
+$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Residual errors of the central portion of a sequence]
+
+Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+
+Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
+$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Central normality of a sequence]
+
+The central portion of a sequence is normally distributed if the sum of residual
+errors is equal to zero:
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+
+\end{definition}
+
+In practice, a finite sequence sampled from a normal distribution $\mathcal{N}(\mu, \sigma)$ will have a non-zero sum of residual errors.
+
+\begin{definition}[Central normality test]
+The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+
+THe alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+\end{definition}
+
+The null-hypothesis $\mathcal{H}_0$ is that the central portion of sequence $\mathbf{X}$
+is normally distributed, with the alternative hypothesis $\mathcal{H}_1$ that it
+is not normally distributed. To test this hypothesis, a test statistic is computed
+and compared against a critical value.
+
+\begin{definition}[Central normality test statistic]
+Let $\alpha$ be the significance level for the central normality test.
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+
+The test statistic is then $\tau_{n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+
+\end{definition}
+
+The null-hypothesis should be rejected at significance level $\alpha$ if 
+$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
+
+The central portion of the data needs to be defined and the Type 1 error rates determined to provide critical test statistics.
 We will do so in the [Simulation](#simulation) section.
@@
 usethis::use_data(
-  gof_lookup_table, 
+  central_normality_lookup_table,
+  empirical_central_normality_lookup_table,
   two_sided_function_parameters, 
   internal = TRUE, 
-  overwrite = TRUE)
+  overwrite = TRUE
+)
 ```
@@
 
-## Empirical central normality test
-
-To develop an empirical test for central normality we need to consider two parameters: the central portion $\kappa$ as a fixed parameter, and test statistic $\tau_{\text{ecn}}$.
-We will first define the central portion $\kappa$.
-
-First we draw $m_d=10000$ random asymmetric generalised normal distributions.
-As before, each distribution is parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-Location and scale parameters are set as $\mu = 0$ and $\sigma = 1$, respectively.
-$n = \lceil 10^\gamma \rceil$ values are then randomly drawn, with $\gamma \sim U\left(1.47, 3.00\right)$, which leads to between $30$ and $1000$ values being drawn to create $\mathbf{X}_i$.
-As before, for each distribution, up to $10 \%$ of instances are replaced by outlier values, and this is repeated $10$ times.
-We then compute residuals after optimising robust location- and scale-invariant transformations with the empirical tapered cosine weighting method.
-
-Figure \ref{fig:empirical-central-normality-residual-error} shows the residual errors of the transformed distribution as function of the empirical probability.
-As expected, the largest deviations from normality appear on the extremities of this range.
-For Box-Cox transformations, the effect of outliers on the lower end of the distribution leads to noticeable asymmetry.
-Between very low and very high empirical probabilities, residual errors are constrained and relatively flat.
-This is the candidate range for the central portion of the data.
-
-```{r empirical-central-normality-residual-error, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.0}
+## Central normality and empirical central normality test
+
+To develop a test for central normality we need to consider two
+parameters: the central portion $\kappa$ as a fixed parameter, and test
+statistic $\tau_{\text{ecn}}$. We will first define the central portion
+$\kappa$.
+
+For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
+we randomly draw $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
+This dataset is used to compute the critical test statistic values for the central normality test.
+Additionally, we add $10 \%$ outliers by randomly replacing elements of each ordered sequence.
+The dataset with outliers is used to compute the critical test statistic values for the empirical variant of the central normality test.
+For each sequence we then compute $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
+
+The dataset with outliers is used to select a suitable value for $\kappa$.
+Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
+as a function of $n$ for different values of $\kappa$. With decreasing
+$\kappa$, the test statistic curve decreases. The curves
+for $\kappa = 0.60$ and $\kappa = 0.70$ are similar, whereas curves for 
+$\kappa \geq 0.90$ are affected by outliers, as can be observed by comparing curves
+of the central normality test with those of the empirical variant.
+
+Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
+we use $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
+normality tests are shown in Appendix E.
+
+```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
 cap <- paste0(
-  "Residual error as a function of empirical probability. ",
-  "Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric ",
-  "generalised normal distributions, for each of which outliers were randomly drawn 10 times ",
-  "(up to 10% of samples). Larger errors occur at the edges of each distribution."
-)
-
-.plot_residuals(
+  "Critical test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$ of the (empirical) ",
+  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
+  "The critical test statistics for central normality test are determined using fully normal data, ",
+  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
+  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
+  "emp: empirical"
+)
+
+.plot_test_statistic_centrality(
   plot_theme = plot_theme,
@@
 ```
-We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic $\tau_{\text{ecn}}$.
-Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}.
-Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80.
-In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is more conservative (see Appendix E).
-This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}.
-
-```{r empirical-central-normality-type-1-error-rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.0}
-cap <- paste0(
-  "Type 1 error rate of transformed asymmetric generalised normal distributions as function of ",
-  "the test statistic $\\tau_{\\text{ecn}}$ for the central portion $\\kappa$ of the distribution.")
-
-.plot_type_1_error_rate(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir)
-
-```
-
-```{r include = FALSE, eval = FALSE}
-# This writes the lookup table for the empirical central normality test to the
-# power.transform package.
-load(file = "./R/sysdata.rda")
-
-gof_lookup_table <- .get_test_statistics_data(
-  manuscript_dir = manuscript_dir,
-  reduce = TRUE,
-  as_table = FALSE)
-gof_lookup_table[, "method" := NULL]
-
-usethis::use_data(
-  gof_lookup_table, 
-  two_sided_function_parameters, 
-  internal = TRUE, 
-  overwrite = TRUE)
-```
-
-```{r include = FALSE, eval = FALSE}
-# This generates the data for the table below.
-table_data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
-\label{tab:empirical-central-normality}
-\begin{tabular}{l | c c c c c c c}
-
-\toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
-
-\midrule
-$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical central normality test to assess central normality of features that are composed of a mixture of samples drawn from two normal distributions.
-With increased separation of the underlying normal distributions, the probability of the feature being centrally normal decreases, as expected.
+
+In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
+central normality test to assess central normality of features that are composed
+of a mixture of samples drawn from two normal distributions ($n = 100$ each).
+With increased separation of the underlying normal distributions, the probability of the
+feature being centrally normal decreases, as expected.
 
@@
   "Bi-modal distributions and empirical central normality test results. ",
-  "The feature (black) is a mixture of two identical sample sets (blue and orange) ",
+  "The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) ",
   "drawn from normal distributions that are offset by a distance $d$. ",
@@
 data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "yeo_johnson")
-data$p
+data$plot
 ```
@@
 Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
-no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
-conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$);
-location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$);
-and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$).
+no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$, $p = 0.71$);
+conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.94$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.94$);
+location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$, $p = 0.95$);
+and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$, $p = 0.80$).
 
@@
 Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
-no transformation (residual sum $\sum r_i = 48.0$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
-location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
-and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).
+no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.06$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.06$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$, $p = 0.18$);
+and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$, $p = 0.40$).
 
@@
 Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
-no transformation (residual sum $\sum r_i = 328$);
-conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$);
-location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$);
-and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$).
+no transformation (residual sum $\sum r_i = 328$, $p < 0.001$);
+conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$, $p < 0.001$);
+location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$, $p < 0.001$);
+and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$, $p < 0.001$).
 
@@
 data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "yeo_johnson")
-data$p
+data$plot
 ```
@@
 Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 54$, $p=0.76$);
-conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.55$);
-location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p=0.03$);
-and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.58$).
-
-Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.03$ for conventional and invariant transformations, respectively).
+no transformation (residual sum $\sum r_i = 54$, $p=0.58$);
+conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.32$);
+location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p < 0.001$);
+and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.35$).
+
+Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p < 0.001$ for conventional and invariant transformations).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
@@
 Applying conventional and invariant Yeo-Johnson transformations to this feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
-conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.61$);
-location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.13$);
-and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.10$).
-
-Non-robust transformations failed to produce a centrally normal distribution (empirical central normality test $p=0.01$ and $p=0.02$ for conventional and invariant transformations, respectively).
+no transformation (residual sum $\sum r_i = 110$, $p=0.86$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.02$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.91$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.23$);
+and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.18$).
+
+The conventional non-robust transformation failed to produce a centrally normal distribution (empirical central normality test $p=0.02$).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

### 2025-03-21 manuscript_rmarkdown/manuscript_appendix.Rmd
Update p-values for empirical tests.

@@
 data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
-data$p
+data$plot
 ```
@@
 Applying conventional and invariant Box-Cox transformations to age of patients with lung cancer [@Loprinzi1994-cd] yielded the following results:
-no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$);
-conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$);
-location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$);
-and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$).
+no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$, $p = 0.71$);
+conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
+location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$, $p = 0.94$);
+and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$, $p = 0.93$).
 
@@
 Applying conventional and invariant Box-Cox transformations to the body mass of penguins [@Gorman2014-eo] yielded the following results:
-no transformation (residual sum $\sum r_i = 48.0$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$);
-location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$);
-and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$).
+no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$, $p = 0.10$);
+and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$, $p = 0.44$).
 
@@
 Applying conventional and invariant Box-Cox transformations to the latitude of houses in the Ames housing dataset [@De-Cock2011-jf] yielded the following results:
-no transformation (residual sum $\sum r_i = 328$);
-conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$);
-location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$);
-and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$).
+no transformation (residual sum $\sum r_i = 328$, $p < 0.001$);
+conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$, $p < 0.001$);
+location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$, $p < 0.001$);
+and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$, $p < 0.001$).
 
@@
 data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "box_cox")
-data$p
+data$plot
 ```
@@
 Applying conventional and invariant Box-Cox transformations to the fuel consumption feature yielded the following results:
-no transformation (residual sum $\sum r_i = 54$, $p=0.76$);
-conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.55$);
-location- and scale-invariant transformation ($\lambda = -0.7$, $\sum r_i = 44$, $\mu_{BC} = 0.6$, $\sigma_{BC} = 0.2$, $p=0.02$);
-and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 59$, $\mu_{BC} = 2.4$, $\sigma_{BC} = 1.8$, $p=0.83$).
+no transformation (residual sum $\sum r_i = 54$, $p=0.58$);
+conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.32$);
+location- and scale-invariant transformation ($\lambda = -0.7$, $\sum r_i = 44$, $\mu_{BC} = 0.6$, $\sigma_{BC} = 0.2$, $p < 0.001$);
+and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 59$, $\mu_{BC} = 2.4$, $\sigma_{BC} = 1.8$, $p=0.67$).
 
@@
 Applying conventional and invariant Box-Cox transformations to the maximum arterial wall thickness feature yielded the following results:
-no transformation (residual sum $\sum r_i = 110$, $p=0.56$);
+no transformation (residual sum $\sum r_i = 110$, $p=0.86$);
 conventional transformation ($\lambda = -0.5$, $\sum r_i = 33$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 127$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.60$);
-location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 28$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.01$);
-and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.35$).
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 127$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.90$);
+location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 28$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.02$);
+and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.60$).
 
@@
 
-\begin{definition}[central portion of a sequence]
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
-ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
-Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
-corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
-percentile values are replaced by the average in their group, i.e.
-if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
-$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
-Furthermore, let central portion $\kappa \in (0,1)$.
-
-Then the central portion of $\mathbf{X}$ is 
-$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
-
-\end{definition}
-
-\begin{definition}[Residual errors of the central portion of a sequence]
-
-Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
-with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
-and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
-
-Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
-$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
-
-\end{definition}
-
-\begin{definition}[Central normality of a sequence]
-
-The central portion of a sequence is normally distributed if the sum of residual
-errors is equal to zero:
-$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
-
-\end{definition}
-
-In practice, a finite sequence sampled from a normal distribution $\mathcal{N}(\mu, \sigma)$ will have a non-zero sum of residual errors.
-
-\begin{definition}[Central normality test]
-The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
-
-THe alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
-\end{definition}
-
-The null-hypothesis $\mathcal{H}_0$ is that the central portion of sequence $\mathbf{X}$
-is normally distributed, with the alternative hypothesis $\mathcal{H}_1$ that it
-is not normally distributed. To test this hypothesis, a test statistic is computed
-and compared against a critical value.
-
-\begin{definition}[Central normality test statistic]
-Let $\alpha$ be the significance level for the central normality test.
-Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
-
-The test statistic is then $\tau_{\alpha, n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
-
-\end{definition}
-
-Critical values for the test statistic of the central normality test 
-are found in Table \ref{tab:central-normality-critical-statistic}. These are 
-determined by repeatedly and randomly sampling data (i.e. $\mathbf{X}$) from
-a normal distribution $\mathcal{N}(0,1)$. The null-hypothesis should be rejected
-at significance level $\alpha$ if 
-$\tau_{\alpha, n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
-
-The empirical central normality test is a central normality test, in the
-definition provided above. However, critical values of the test statistic are
-not determined directly from data drawn randomly from $\mathcal{N}(0,1)$.
-Rather, after $\mathbf{X}$ is drawn, 10% of the values are randomly replaced by
-random outlier values. The resulting distribution is centrally normal, but
-not fully normal.
-
-Critical values for the empirical variant of the central normality test are
-found in Table \ref{tab:empirical-central-normality-critical-statistic}.
+Critical values for the central normality test and its empirical variant are
+found in Table \ref{tab:central-normality-critical-statistic} and Table 
+\ref{tab:empirical-central-normality-critical-statistic}, respectively.
 
@@
 
+```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.70}$ of the (empirical) ",
+  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
+  "The critical test statistics for central normality test are determined using fully normal data, ",
+  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
+  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
+  "emp: empirical"
+)
+
+.plot_test_statistic_tau(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
 # Appendix F: Normalisation before transformation
@@
 \label{tab:normalisation-before-transformation-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 
@@
 \label{tab:normalisation-before-transformation-lambda-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 
@@
 \label{tab:normalisation-before-transformation-p-value-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 
@@
 \midrule
-none & 0.685 & 0.558 & 0.764 & 0.354 & 0.030 \\
-conventional & 0.886 & 0.010 & 0.012 & 0.317 & 0.381 \\
-conventional (z-score norm.) & 0.842 & 0.036 & 0.011 & 0.075 & 0.378 \\
-conventional (robust scaling) & 0.855 & 0.002 & 0.007 & 0.077 & 0.416 \\
-Raymaekers-Rousseeuw & 0.886 & 0.610 & 0.553 & 0.296 & 0.381 \\
-Raymaekers-Rousseeuw (z-score norm.) & 0.685 & 0.632 & 0.812 & 0.075 & 0.378 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.826 & 0.609 & 0.813 & 0.077 & 0.434 \\
-invariant & 0.907 & 0.134 & 0.026 & 0.049 & 0.549 \\
-robust invariant & 0.781 & 0.102 & 0.583 & 0.085 & 0.717 \\
+none                                  & 0.712 & 0.864 & 0.577     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.936 & 0.017 & $< 0.001$ & $< 0.001$ & 0.058 \\
+conventional (z-score norm.)          & 0.860 & 0.058 & $< 0.001$ & $< 0.001$ & 0.054 \\
+conventional (robust scaling)         & 0.881 & 0.001 & $< 0.001$ & $< 0.001$ & 0.076 \\
+Raymaekers-Rousseeuw                  & 0.936 & 0.905 & 0.320     & $< 0.001$ & 0.058 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.605 & 0.922 & 0.635     & $< 0.001$ & 0.054 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.830 & 0.948 & 0.637     & $< 0.001$ & 0.085 \\
+invariant                             & 0.948 & 0.231 & $< 0.001$ & $< 0.001$ & 0.184 \\
+robust invariant                      & 0.797 & 0.182 & 0.352     & $< 0.001$ & 0.397 \\
 \bottomrule

### 2025-03-21 manuscript_rmarkdown/plots.R

@@ -1542,7 +1542,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
     # Normal distribution.
     x_1 <- power.transform::ragn(
-      10000L,
+      100L,
       location = -location_shift / 2,
       scale = 1 / sqrt(2),
       alpha = 0.5,

### 2025-03-21 manuscript_rmarkdown/manuscript.Rmd
Correct use of alpha.

@@
 
-Considering single process parameters, the choice of learner (Friedman test: $p < 10^{-8})$), normalisation method (Wilcoxon signed rank test: $p < 10^{-8}$), and transformation method $p = 0.007$, all had a significant impact (at significance level $p = 0.05$).
-
-To estimate the marginal effects of process parameters, including transformation method, we first fit a regression random forest (ranger package [@Wright2017-rf] version 0.16.0): 2000 trees, minimum node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
+Considering single process parameters, 
+the choice of learner (Friedman test: $p < 10^{-8})$),
+normalisation method (Wilcoxon signed rank test: $p = 4 \cdot 10^{-8}$), 
+and transformation method (Friedman test:$p = 0.007$,
+all had a significant impact (at $p = 0.05$).
+
+To estimate the marginal effects of process parameters, including transformation method, 
+we first fit a regression random forest (ranger package [@Wright2017-rf] version 0.16.0): 2000 trees, 
+node size 2, other hyperparameters default) with process parameters and task difficulty as predictors and normalised rank as response variable.
 The estimated marginal effects are shown in Figure \ref{fig:marginal-effect-plot}.

### 2025-03-21 manuscript_rmarkdown/plots.R

@@ -2810,14 +2810,14 @@ get_annotation_settings <- function(ggtheme = NULL) {
   data <- rbind(data_w_outliers, data_wo_outliers)
 
   # Compute alpha
-  data[, "alpha" := (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
+  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
 
   # Compute alpha = 0.95
   data <- data[, list("tau" = stats::spline(
     x = alpha,
     y = mean_residual_error,
     method = "fmm",
-    xout = 0.95
+    xout = 0.05
   )$y),
   by = c("n", "kappa", "outlier")]
 
@@ -2836,7 +2836,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p <- p + plot_theme
   p <- p + ggplot2::scale_x_log10(name = "n")
-  p <- p + ggplot2::ylab(latex2exp::TeX("test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$"))
+  p <- p + ggplot2::ylab(latex2exp::TeX("test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$"))
   p <- p + ggplot2::scale_colour_discrete(
     name = "central portion κ",
     type = c(
@@ -2890,12 +2890,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
   data <- rbind(data_w_outliers, data_wo_outliers)
   data <- data[kappa == k]
 
-  alpha_levels <- c(0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
+  alpha_levels <- 1.0 - c(0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
 
   # Compute alpha
-  data[, "alpha" := (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
+  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
 
-  # Compute alpha = 0.95
+  # Compute tau at the specified confidence levels.
   data <- data[
     ,
     list(
@@ -2912,7 +2912,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   data$alpha <- factor(
     x = data$alpha,
[…]

### 2025-03-21 manuscript_rmarkdown/simulations.R

@@ -2865,14 +2865,15 @@
     with_outliers = FALSE
 ) {
   if (for_manuscript) {
-    alpha_levels <- c(0.10, 0.20, 0.50, 0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
+    alpha_levels <- rev(1.0 - c(0.10, 0.20, 0.50, 0.80, 0.90, 0.95, 0.975, 0.99, 0.999))
     n <- c(5, 10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000)
   } else {
-    alpha_levels <- c(
+    alpha_levels <- rev(1.0 - c(
       0.00, 0.01, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30, 0.35, 0.40, 0.45, 0.50, 0.55,
       0.60, 0.65, 0.70, 0.75, 0.80, 0.85, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95,
       0.96, 0.97, 0.975, 0.98, 0.9825, 0.985, 0.9875, 0.99,
-      0.991, 0.992, 0.993, 0.994, 0.995, 0.996, 0.997, 0.998, 0.999, 0.9995)
+      0.991, 0.992, 0.993, 0.994, 0.995, 0.996, 0.997, 0.998, 0.999, 0.9995
+    ))
     n <- c(
       5, 10, 15, 20, 30, 40, 50, 60, 70, 80, 90, 100, 120, 140, 160, 180, 200,
       225, 250, 275, 300, 340, 380, 420, 460, 500,
@@ -2885,7 +2886,7 @@
   data[, "kappa" := NULL]
 
   # Compute alpha. Higher alpha yields higher tau (with n constant).
-  data[, "alpha" := (seq_len(.N) - 1L) / (.N - 1L), by = c("n")]
+  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n")]
 
   # Define all combinations.
   new_data <- data.table::as.data.table(expand.grid(list("alpha" = alpha_levels, "n" = n)))

### 2025-03-21 manuscript_rmarkdown/manuscript.Rmd
Add type I error rate plots.

@@
 For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
-we randomly draw $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
-This dataset is used to compute the critical test statistic values for the central normality test.
-Additionally, we add $10 \%$ outliers by randomly replacing elements of each ordered sequence.
-The dataset with outliers is used to compute the critical test statistic values for the empirical variant of the central normality test.
-For each sequence we then compute $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
-
-The dataset with outliers is used to select a suitable value for $\kappa$.
+we randomly drew $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
+This dataset was used to compute the critical test statistic values for the central normality test.
+Additionally, we added $10 \%$ outliers by randomly replacing elements of each ordered sequence.
+The dataset with outliers was used to compute the critical test statistic values for the empirical variant of the central normality test.
+For each sequence we then computed $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
+
 Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
@@
 Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
-we use $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
+we used $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
 normality tests are shown in Appendix E.
@@
 .plot_test_statistic_centrality(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
+To assess type I error rates for the (empirical) central normality test and compare
+these to the Shapiro-Wilk test for normality, we randomly drew another 1000 sequences from a
+standard normal distribution ($\mathcal{N}(0,1)$ for each 
+$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
+A second dataset was created by replacing elements with randomly drawn outliers, as described above.
+For each sequence, the p-value for the respective test was used to reject the null
+hypothesis that the sequence is (centrally) normal. The null hypothesis was rejected if $p \leq 0.05$.
+The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
+Figure \ref{empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
+centrally normal data (with tails of the distribution not being normally distributed),
+whereas the central normality test and Shapiro-Wilk test overestimated type I error rates.
+
+```{r empirical-central-normality-error_rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers. ",
+  "ECN: empirical central normality; CN: central normality"
+)
+
+.plot_test_dependency_sample_size(
   plot_theme = plot_theme,

### 2025-03-27 manuscript_rmarkdown/manuscript.Rmd
WIP on simulations with clean data.

@@
 ```
+## Simulated data
+
+[Generated data: clean, n = 30, n = 100, n = 1000, lambda = -1.00, 1.00, 3.00]
+
+```{r include = FALSE, eval = FALSE}
+
+data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))
+data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
+```
 

### 2025-03-27 manuscript_rmarkdown/manuscript_appendix.Rmd
Correct use of alpha.
WIP on simulations with clean data.

@@
 - \usepackage{booktabs}
+- \usepackage{multirow}
 - \newtheorem{definition}{Definition}
@@
 
+# Appendix D: Simulations on clean data
+
+Robust transformations are hypothesised to have a cost in efficiency for data
+without outliers, i.e. clean data. Here we randomly draw data from a standard normal
+distribution $\mathcal{N}(0,1)$. Subsequently, we perform an inverse transformation
+$\left(\phi_{\text{YJ}}^{\lambda, 0, 1}\right)^{-1}$, with $\lambda \in \left(0, 2\right)$.
+
+An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
+We assessed two normalisation methods were used:
+1. z-standardisation, i.e. $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of feature $\mathbf{X}$; 
+and 2.
+robust scaling, i.e.
+$x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$, with $\text{IQR}$ representing the interquartile range of feature $\mathbf{X}$.
+
+[Generated data: clean, n = 30, n = 100, n = 1000, lambda = 0.1, 1.0, 1.9]
+
+```{r include = FALSE, eval = FALSE}
+
+data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))
+data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-residuals-appendix}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 4.0 & 3.4 & 4.6 & 40.1 & 8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
+conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 & 8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
+conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 & 8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
+conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 & 8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
+Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 & 8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 & 8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 & 8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
+invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 & 8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
+robust invariant                      & & 3.0 & 3.3 & 3.6 &  5.2 & 8.3 &  5.9 &  19.9 & 13.6 &  32.9 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-lambda-appendix}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
+conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
+conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
+Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
+invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
+robust invariant                      & & 0.4 & 1.3 & 1.5 &  0.0 & 0.8 & 1.7 &  0.0 & 1.1 & 2.0 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+```
+
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-p-value-appendix}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.986 & 0.614 & 0.323 & $<0.001$ & 0.063 & 0.016 & $<0.001$ & 0.997 & $<0.001$ \\
+conventional                          & & 0.986 & 0.464 & 0.488 &   0.675  & 0.139 & 0.620 &   0.902  & 0.999 &   0.698 \\
+conventional (z-score norm.)          & & 0.979 & 0.440 & 0.477 &   0.200  & 0.141 & 0.770 &   0.419  & 0.999 &   0.723 \\
+conventional (robust scaling)         & & 0.982 & 0.439 & 0.484 &   0.487  & 0.130 & 0.683 &   0.883  & 0.999 &   0.817 \\
+Raymaekers-Rousseeuw                  & & 0.989 & 0.464 & 0.488 &   0.675  & 0.446 & 0.620 &   0.876  & 0.992 &   0.821 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.992 & 0.440 & 0.477 &   0.200  & 0.448 & 0.770 &   0.460  & 0.991 &   0.719 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.993 & 0.439 & 0.484 &   0.487  & 0.422 & 0.683 &   0.887  & 0.991 &   0.848 \\
+invariant                             & & 0.993 & 0.355 & 0.461 &   0.720  & 0.153 & 0.490 &   0.910  & 0.999 &   0.682 \\
+robust invariant                      & & 0.994 & 0.407 & 0.484 &   0.642  & 0.291 & 0.489 &   0.848  & 0.996 &   0.628 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+```
+
 # Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation
@@
 \toprule
-$n$ \textbackslash $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
+$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
 
@@
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
@@
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
@@
 p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}

### 2025-03-27 manuscript_rmarkdown/manuscript_appendix.tex
WIP on simulations with clean data.

--- 2025-03-13 manuscript_rmarkdown/manuscript_appendix.tex
@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{booktabs}
+\usepackage{multirow}
+\newtheorem{definition}{Definition}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-03-13}
+\date{2025-03-27}
 
@@
 
+\section{Appendix D: Simulations on clean
+data}\label{appendix-d-simulations-on-clean-data}
+
+An alternative to location- and scale-invariant transformations is
+normalising feature distributions prior to conventional transformations.
+We assessed two normalisation methods were used: 1. z-standardisation,
+i.e.~\(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
+and \(\sigma\) the mean and standard deviation of feature
+\(\mathbf{X}\); and 2. robust scaling,
+i.e.~\(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
+with \(\text{IQR}\) representing the interquartile range of feature
+\(\mathbf{X}\).
+
+{[}Generated data: clean, n = 30, n = 100, n = 1000, lambda = 0.1, 1.0,
+1.9{]}
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-residuals-appendix}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 4.0 & 3.4 & 4.6 & 40.1 & 8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
+conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 & 8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
+conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 & 8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
+conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 & 8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
+Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 & 8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 & 8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 & 8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
+invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 & 8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
+robust invariant                      & & 3.0 & 3.3 & 3.6 &  5.2 & 8.3 &  5.9 &  19.9 & 13.6 &  32.9 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-lambda-appendix}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
+conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
+conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
+Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
+invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
+robust invariant                      & & 0.4 & 1.3 & 1.5 &  0.0 & 0.8 & 1.7 &  0.0 & 1.1 & 2.0 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-p-value-appendix}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.986 & 0.614 & 0.323 & $<0.001$ & 0.063 & 0.016 & $<0.001$ & 0.997 & $<0.001$ \\
+conventional                          & & 0.986 & 0.464 & 0.488 &   0.675  & 0.139 & 0.620 &   0.902  & 0.999 &   0.698 \\
+conventional (z-score norm.)          & & 0.979 & 0.440 & 0.477 &   0.200  & 0.141 & 0.770 &   0.419  & 0.999 &   0.723 \\
+conventional (robust scaling)         & & 0.982 & 0.439 & 0.484 &   0.487  & 0.130 & 0.683 &   0.883  & 0.999 &   0.817 \\
+Raymaekers-Rousseeuw                  & & 0.989 & 0.464 & 0.488 &   0.675  & 0.446 & 0.620 &   0.876  & 0.992 &   0.821 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.992 & 0.440 & 0.477 &   0.200  & 0.448 & 0.770 &   0.460  & 0.991 &   0.719 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.993 & 0.439 & 0.484 &   0.487  & 0.422 & 0.683 &   0.887  & 0.991 &   0.848 \\
+invariant                             & & 0.993 & 0.355 & 0.461 &   0.720  & 0.153 & 0.490 &   0.910  & 0.999 &   0.682 \\
+robust invariant                      & & 0.994 & 0.407 & 0.484 &   0.642  & 0.291 & 0.489 &   0.848  & 0.996 &   0.628 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
 \section{Appendix D: Experimental results using location- and
@@
 results: no transformation (sum of residuals with normal distribution
-\(\sum r_i = 16.5\)); conventional transformation (\(\lambda = 1.9\),
-\(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
-\(\sigma_{BC} = 0.4 \cdot 10^3\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
-\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\));
-location- and scale-invariant transformation (\(\lambda = 1.7\),
-\(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\), \(\sigma_{BC} = 0.8\)); and
-robust location- and scale-invariant transformation (\(\lambda = 1.5\),
-\(\sum r_i = 11.6\), \(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\)).
+\(\sum r_i = 16.5\), \(p = 0.71\)); conventional transformation
+(\(\lambda = 1.9\), \(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
+\(\sigma_{BC} = 0.4 \cdot 10^3\), \(p = 0.94\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
+\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\),
+\(p = 0.94\)); location- and scale-invariant transformation
+(\(\lambda = 1.7\), \(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\),
+\(\sigma_{BC} = 0.8\), \(p = 0.94\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.5\), \(\sum r_i = 11.6\),
+\(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\), \(p = 0.93\)).
 
@@
 mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
-following results: no transformation (residual sum \(\sum r_i = 48.0\));
-conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); Raymaekers and
+following results: no transformation (residual sum \(\sum r_i = 48.0\),
+\(p < 0.001\)); conventional transformation (\(\lambda = -0.5\),
+\(\sum r_i = 32.2\), \(\mu_{BC} = 2.1\),
+\(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\)); Raymaekers and
 Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); location- and
-scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 27.3\),
-\(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 25.2\),
-\(\mu_{BC} = 0.4\), \(\sigma_{BC} = 0.6\)).
+\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\));
+location- and scale-invariant transformation (\(\lambda = 0.5\),
+\(\sum r_i = 27.3\), \(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\),
+\(p = 0.10\)); and robust location- and scale-invariant transformation
+(\(\lambda = 0.2\), \(\sum r_i = 25.2\), \(\mu_{BC} = 0.4\),
+\(\sigma_{BC} = 0.6\), \(p = 0.44\)).
 
@@
 the following results: no transformation (residual sum
-\(\sum r_i = 328\)); conventional transformation (\(\lambda = 62.1\),
-\(\sum r_i = 319\), \(\mu_{BC} = 1.1 \cdot 10^{99}\),
-\(\sigma_{BC} = 0.0 \cdot 10^{99}\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = 96.0\), \(\sum r_i = 319\),
-\(\mu_{BC} = 6.2 \cdot 10^{153}\),
-\(\sigma_{BC} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
-transformation (\(\lambda = 1.9\), \(\sum r_i = 312\),
-\(\mu_{BC} = 2.3\), \(\sigma_{BC} = 0.9\)); and robust location- and
+\(\sum r_i = 328\), \(p < 0.001\)); conventional transformation
+(\(\lambda = 62.1\), \(\sum r_i = 319\),
+\(\mu_{BC} = 1.1 \cdot 10^{99}\), \(\sigma_{BC} = 0.0 \cdot 10^{99}\),
+\(p < 0.001\)); Raymaekers and Rousseeuw's robust adaptation
+(\(\lambda = 96.0\), \(\sum r_i = 319\),
+\(\mu_{BC} = 6.2 \cdot 10^{153}\), \(\sigma_{BC} = 0.3 \cdot 10^{153}\),
+\(p < 0.001\)); location- and scale-invariant transformation
+(\(\lambda = 1.9\), \(\sum r_i = 312\), \(\mu_{BC} = 2.3\),
+\(\sigma_{BC} = 0.9\), \(p < 0.001\)); and robust location- and
 scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\)).
+\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\), \(p < 0.001\)).
 
@@
 consumption feature yielded the following results: no transformation
-(residual sum \(\sum r_i = 54\), \(p=0.76\)); conventional
+(residual sum \(\sum r_i = 54\), \(p=0.58\)); conventional
 transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
-\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p=0.01\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
-\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.55\)); location- and
+\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p < 0.001\)); Raymaekers
+and Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
+\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.32\)); location- and
 scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
-\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p=0.02\)); and robust
+\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p < 0.001\)); and robust
 location- and scale-invariant transformation (\(\lambda = 1.1\),
 \(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
-\(p=0.83\)).
+\(p=0.67\)).
 
@@
 arterial wall thickness feature yielded the following results: no
-transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
+transformation (residual sum \(\sum r_i = 110\), \(p=0.86\));
 conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
@@
 Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.60\)); location- and
+\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.90\)); location- and
 scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
-\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.01\)); and robust
+\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.02\)); and robust
 location- and scale-invariant transformation (\(\lambda = 0.5\),
 \(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
-\(p=0.35\)).
+\(p=0.60\)).
 
@@
 
+Critical values for the central normality test and its empirical variant
+are found in Table \ref{tab:central-normality-critical-statistic} and
+Table \ref{tab:empirical-central-normality-critical-statistic},
+respectively.
+
 \begin{table}
 \begin{center}
-\caption{Critical values of test statistic $\tau_{\text{ecn},\alpha, n}$ for
-empirical central normality at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The presented values should be divided by $100$.}
-\label{tab:empirical-central-normality-critical-statistic}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:central-normality-critical-statistic}
 \begin{tabular}{l | r r r r r r r r r}
@@
 \toprule
-$n$ \ $\alpha$ & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 & 0.95 & 0.975 & 0.99 & 0.999 \\
+$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
 
 \midrule
-    5 & 24.02 & 21.26 & 17.37 & 14.02 & 12.04 &  9.57 &  7.64 &  5.54 & 2.35 \\
-   10 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 &  7.65 &  6.79 &  5.87 & 4.23 \\
-   20 & 15.12 & 13.17 & 10.03 &  7.74 &  6.80 &  6.13 &  5.58 &  5.04 & 4.07 \\
-   50 &  9.99 &  8.70 &  6.69 &  5.26 &  4.67 &  4.26 &  3.92 &  3.59 & 3.02 \\
-  100 &  7.13 &  6.21 &  4.83 &  3.79 &  3.38 &  3.08 &  2.87 &  2.63 & 2.24 \\ 
-  200 &  5.19 &  4.54 &  3.51 &  2.77 &  2.47 &  2.26 &  2.09 &  1.92 & 1.65 \\ 
-  500 &  3.47 &  3.06 &  2.38 &  1.86 &  1.65 &  1.51 &  1.40 &  1.28 & 1.08 \\
- 1000 &  2.65 &  2.35 &  1.84 &  1.44 &  1.27 &  1.16 &  1.07 &  0.98 & 0.81 \\
- 2000 &  2.13 &  1.91 &  1.53 &  1.20 &  1.06 &  0.96 &  0.88 &  0.79 & 0.65 \\
- 5000 &  1.69 &  1.56 &  1.30 &  1.07 &  0.95 &  0.87 &  0.80 &  0.72 & 0.58 \\
-10000 &  1.50 &  1.41 &  1.23 &  1.05 &  0.96 &  0.89 &  0.83 &  0.77 & 0.64 \\
+5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
+10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
+20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
+50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
+100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
+200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
+500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
+1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
+2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
+5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
+10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
 \bottomrule
@@
 \end{table}
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
+
+\midrule
+5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
+10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
+20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
+50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
+100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
+200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
+500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
+1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
+2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
+5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
+10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-critical-statistics-1}
+
+}
+
+\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.70}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
+The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
+with fully normal data where 10\% of elements are replaced by outliers.
+emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
+\end{figure}
 
@@
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 
@@
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-lambda-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 
@@
 p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
 \label{tab:normalisation-before-transformation-p-value-appendix}
-\begin{tabular}{l | c c c c c}
+\begin{tabular}{l | r r r r r}
 

### 2025-03-27 manuscript_rmarkdown/plots.R

@@ -1665,11 +1665,10 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p_2 <- .create_density_and_qq_plot(location_shift = 2.0, plot_theme = plot_theme)
   p_3 <- .create_density_and_qq_plot(location_shift = 3.0, plot_theme = plot_theme)
   p_4 <- .create_density_and_qq_plot(location_shift = 4.0, plot_theme = plot_theme)
-  p_5 <- .create_density_and_qq_plot(location_shift = 5.0, plot_theme = plot_theme)
 
-  p <- p_1$density + p_2$density + p_3$density + p_4$density + p_5$density +
-    p_1$qq + p_2$qq + p_3$qq + p_4$qq + p_5$qq +
-    patchwork::plot_layout(ncol = 5, heights = c(1.3, 1))
+  p <- p_1$density + p_2$density + p_3$density + p_4$density +
+    p_1$qq + p_2$qq + p_3$qq + p_4$qq +
+    patchwork::plot_layout(ncol = 4, heights = c(1.3, 1))
 
   return(p)
 }

### 2025-03-28 manuscript_rmarkdown/manuscript_appendix.Rmd
Added table for comparing estimators on 10000 randomly sampled datasets.

@@
 
+These transformed data
+
 An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
@@
 \end{table}
+```
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# Add rank based on residual error for central normality.
+data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
+```
+```{=latex}
+\begin{table}
+\begin{center}
+\caption{
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:comparison_methods_simulations}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 8.51          & 6.33          & 7.31          & 5.55          & 7.42          & 6.39 \\
+conventional                          & & 3.72          & 3.78          & 3.83          & 4.74          & 6.47          & 5.47 \\
+conventional (z-score norm.)          & & 4.72          & 2.89          & 6.37          & 3.43          & 5.42          & 3.50 \\
+conventional (robust scaling)         & & 4.07          & 4.31          & 5.36          & 4.23          & 4.41          & 4.32 \\
+Raymaekers-Rousseeuw                  & & 4.50          & 7.03          & 3.97          & 8.12          & 6.25          & 5.57 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 5.26          & 6.36          & 6.22          & 6.42          & 5.26          & 6.92 \\
+Raymaekers-Rousseeuw (robust scaling) & & 4.71          & 7.84          & 5.31          & 7.06          & 4.35          & 7.58 \\
+invariant                             & & \textbf{3.15} & 3.72          & \textbf{3.27} & 3.14          & \textbf{2.69} & 3.05 \\
+robust invariant                      & & 6.37          & \textbf{2.73} & 3.36          & \textbf{2.30} & 2.72          & \textbf{2.20} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+```
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean", method = "box_cox")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty", method = "box_cox")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted", method = "box_cox")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# Add rank based on residual error for central normality.
+data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
 ```

### 2025-03-28 manuscript_rmarkdown/manuscript_appendix.tex
Added table for comparing estimators on 10000 randomly sampled datasets.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-03-27}
+\date{2025-03-28}
 
@@
 
+Robust transformations are hypothesised to have a cost in efficiency for
+data without outliers, i.e.~clean data. Here we randomly draw data from
+a standard normal distribution \(\mathcal{N}(0,1)\). Subsequently, we
+perform an inverse transformation
+\(\left(\phi_{\text{YJ}}^{\lambda, 0, 1}\right)^{-1}\), with
+\(\lambda \in \left(0, 2\right)\).
+
+These transformed data
+
 An alternative to location- and scale-invariant transformations is
@@
 }
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:comparison_methods_simulations}
+\begin{tabular}{l | r r r r r r}
+
+\toprule
+& \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+outliers & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & 8.51          & 6.33          & 7.31          & 5.55          & 7.42          & 6.39 \\
+conventional                          & 3.72          & 3.78          & 3.83          & 4.74          & 6.47          & 5.47 \\
+conventional (z-score norm.)          & 4.72          & 2.89          & 6.37          & 3.43          & 5.42          & 3.50 \\
+conventional (robust scaling)         & 4.07          & 4.31          & 5.36          & 4.23          & 4.41          & 4.32 \\
+Raymaekers-Rousseeuw                  & 4.50          & 7.03          & 3.97          & 8.12          & 6.25          & 5.57 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 5.26          & 6.36          & 6.22          & 6.42          & 5.26          & 6.92 \\
+Raymaekers-Rousseeuw (robust scaling) & 4.71          & 7.84          & 5.31          & 7.06          & 4.35          & 7.58 \\
+invariant                             & \textbf{3.15} & 3.72          & \textbf{3.27} & 3.14          & \textbf{2.69} & 3.05 \\
+robust invariant                      & 6.37          & \textbf{2.73} & 3.36          & \textbf{2.30} & 2.72          & \textbf{2.20} \\
+
+\bottomrule
+\end{tabular}
 \end{center}

### 2025-03-28 manuscript_rmarkdown/simulations.R

@@ -2124,6 +2124,190 @@
 
 
 
+.assess_simulated_data <- function(
+    manuscript_dir,
+    method = "yeo_johnson",
+    data_type = "clean",
+    k = 0.10
+) {
+  file_name <- paste0(
+    "simulated_data_",
+    data_type,
+    "_", method, ".RDS"
+  )
+  file_name <- file.path(manuscript_dir, file_name)
+
+  set.seed(95L)
+
+  if (file.exists(file_name)) return(readRDS(file_name))
+
+  # Generate distributions to assess.
+  n_distributions <- 10000L
+
+  # Generate alpha, beta and n.
+  n <- stats::runif(n = n_distributions, min = 1.0, max = 4.0)
+  n <- ceiling(10^n)
+
+  # Parameters for asymmetric generalised normal distributions.
+  alpha <- stats::runif(n = n_distributions, min = 0.01, max = 0.99)
+  beta <- stats::runif(n = n_distributions, min = 1.00, max = 5.00)
+
+  # Parameters for normal distributed data that is inversely transformed.
+  # Note that Yeo-Johnson is only directly invertible for lambda between 0 and
+  # 2, and Box-Cox only if lambda > 0.
+  lambda <- stats::runif(n = n_distributions, min = 0.00, max = 2.00)
+
+  ..compute <- function(
+    ii,
+    n,
+    alpha,
+    beta,
+    conf_lambda,
+    data_type,
+    method,
+    k
+  ) {
+    set.seed(ii)
+    if (data_type == "clean") {
+      # Clean data is produced by first inverting the power transformation (with
+      # known lambda) for normally distributed data. This means that the
+      # transformation parameters can be exactly determined - aside from sampling
+      # effects.
+
+      x <- stats::rnorm(
+        n = n,
+        mean = 0.0,
+        sd = 1.0
+      )
+
+      # Ensure that x is entirely positive
+      if (method == "box_cox") x <- x + 1E-4 - min(x)
+
+      # Perform inverse transform.
+      x <- power.transform::revert_power_transform(
+        y = x,
+        transformer = power.transform::create_transformer_skeleton(
+          method = method,
+          lambda = conf_lambda
+        )
+      )
+
+    } else if (data_type == "dirty") {
+      # Dirty data are produced from asymmetric generalised normal distributions.
+      # These are dirty in the sense that the parameters of the power
+      # transformation cannot be directly derived from the distribution.
[…]

### 2025-03-30 manuscript_rmarkdown/manuscript.Rmd
Reworked robust transformations.

@@
 
-Modifications intended to make power transformations invariant to location and scale of a feature and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions.
-Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option for simulating realistic feature distributions.
+Modifications intended to make power transformations invariant to location and 
+scale of a feature and methods to improve their robustness against outliers need 
+to be assessed using data drawn from a range of different distributions. 
+Since the power transformations are intended for use with unimodal distributions, 
+the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option for simulating realistic feature distributions.
 This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:
 
-```{=latex}
+\begin{definition}[Standard generalised normal distribution probability density function]
+Let $x \in \mathcal{R}$.
+Let $\beta > 0$ be a shape parameter.
+Let $\Gamma$ be the gamma function.
+
+Then the standard generalised normal distribution probability density function, without scale and location parameters, is:
 \begin{equation}
@@
 \end{equation}
-```
-Here, $\Gamma$ is the gamma function, and $\beta$ is a strictly positive shape parameter.
+\end{definition}
+
 For $\beta = 1$, the probability density function describes a Laplace distribution.
-A normal distribution is found for $\beta=2$, and for large $\beta$, the distribution approaches a uniform distribution.
-We will refrain from introducing scale and location parameters here directly.
+A normal distribution is found for $\beta=2$, and for large $\beta$,
+the distribution approaches a uniform distribution.
 
 Realistic feature distributions may be skewed. 
-Gijbels et al.
-describe a recipe for introducing skewness into the otherwise symmetric generalised normal distribution [@Gijbels2019-te], leading to the following probability density function:
-
+Gijbels et al. describe a recipe for introducing skewness into the otherwise symmetric generalised normal distribution [@Gijbels2019-te], 
+leading to the following probability density function:
+
+\begin{definion}[Asymmetric generalised normal distribution probability density function]
+Let $\alpha \in (0,1)$ be a skewness parameter.
+Let $\mu \in \mathcal{R}$ be a location parameter.
+Let $\sigma > 0$ be a shape parameter.
+
+Then the probability density function of the asymmetric generalised normal distribution is:
 \begin{equation}
@@
 \end{equation}
-
-Here, $\alpha \in (0,1)$ is a skewness parameter.
-$\alpha > 0.5$ creates a distribution with a negative skew, i.e.
-a left-skewed distribution.
+\end{definition}
+
+$\alpha > 0.5$ creates a distribution with a negative skew, i.e. a left-skewed distribution.
 A right-skewed distribution is created for $\alpha < 0.5$.
-$\mu \in \mathcal{R}$ and $\sigma \in (0, \infty)$ are location and scale parameters, respectively.
-$f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution, which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.
-
-We require a quantile function (or an approximation thereof) to draw random values from an asymmetric generalised normal distribution using inverse transform sampling.
-Gijbels et al.
-derived the following quantile function $F_{\alpha}^{-1}(p)$:
-
+$f_{\alpha}$ thus describes the probability density function of an asymmetric generalised normal distribution,
+which we will refer to here and parametrise as $\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)$.
+
+We require a quantile function (or an approximation thereof) to draw random values
+from an asymmetric generalised normal distribution using inverse transform sampling.
+Gijbels et al. derived the quantile function $F_{\alpha}^{-1}(p)$, which incorporates
+the quantile function of the symmetric generalised normal distribution derived by Griffin [@Griffin2018-bf]:
+
+\begin{definition}[Asymmetric generalised normal distribution quantile function]
+Let $p \in \left[0, 1\right]$ be a probability.
+Let  $F_{\Gamma}^{-1}$ is the quantile function of the gamma distribution with shape $1 / \beta$, which can be numerically approximated. 
+Let $F_{\beta}^{-1}$ be the quantile function of the generalised normal distribution:
+\begin{equation}
+F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
+\end{equation}
+
+Then the quantile function of the asymmetric generalised normal distribution is:
 \begin{equation}
@@
 \end{equation}
-
-The quantile function for the asymmetric generalised normal distribution $F_{\alpha}^{-1}$ thus incorporates the quantile function $F_{\beta}^{-1}$ of the symmetric generalised normal distribution.
-$F_{\beta}^{-1}$ was derived by Griffin to be [@Griffin2018-bf]:
-
-\begin{equation}
-F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
-\end{equation}
-
-Here, $F_{\Gamma}^{-1}$ is the quantile function of the gamma distribution with shape $1 / \beta$, which can be numerically approximated.
+\end{definition}
 
@@
 
-THe alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
 \end{definition}
@@
 \begin{definition}[Central normality test statistic]
-Let $\alpha$ be the significance level for the central normality test.
 Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
@@
 
-We used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
-The $\lambda$ parameter for conventional power transformations (Eqn.
-\ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location- and scale-invariant power transformations (Eqn.
-\ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq] (version 1.0.0).
+We used simulated data to assess invariance to location and scale of the proposed power transformations,
+to develop the empirical central normality test, and to determine weighting for robust transformations.
+The $\lambda$ parameter for conventional power transformations (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}),
+as well as $\lambda$, $x_0$ and $s$ parameters for location- and scale-invariant power transformations 
+(Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm 
+for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
+The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq] (version 1.0.1).
 Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.
@@
 
-To assess whether the proposed power transformations lead to values of $\lambda$ that are invariant to location and scale of the distribution, we simulated three different distributions.
-We first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
-The second distribution was a right-skewed generalised normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
-The third distribution was a left-skewed generalised normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+To assess whether the proposed power transformations lead to values of $\lambda$ 
+that are invariant to location and scale of the distribution, we simulated three different sequences
+We first randomly drew $10000$ values from a normal distribution: 
+$\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, 
+or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution was a right-skewed generalised normal distribution 
+$\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution was a left-skewed generalised normal distribution 
+$\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
 We then computed transformation parameter $\lambda$ using the original definitions (Eqn.
@@
 
+## Central normality and empirical central normality test
+
+To develop a test for central normality we need to consider two
+parameters: the central portion $\kappa$ as a fixed parameter, and test
+statistic $\tau_{\text{ecn}}$. We will first define the central portion
+$\kappa$.
+
+For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
+we randomly drew $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
+This dataset was used to compute the critical test statistic values for the central normality test.
+Additionally, we added $10 \%$ outliers by randomly replacing elements of each ordered sequence.
+The dataset with outliers was used to compute the critical test statistic values for the empirical variant of the central normality test.
+For each sequence we then computed $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
+
+Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
+as a function of $n$ for different values of $\kappa$. With decreasing
+$\kappa$, the test statistic curve decreases. The curves
+for $\kappa = 0.60$ and $\kappa = 0.70$ are similar, whereas curves for 
+$\kappa \geq 0.90$ are affected by outliers, as can be observed by comparing curves
+of the central normality test with those of the empirical variant.
+
+Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
+we used $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
+normality tests are shown in Appendix E.
+
+```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Critical test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$ of the (empirical) ",
+  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
+  "The critical test statistics for central normality test are determined using fully normal data, ",
+  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
+  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
+  "emp: empirical"
+)
+
+.plot_test_statistic_centrality(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
+To assess type I error rates for the (empirical) central normality test and compare
+these to the Shapiro-Wilk test for normality, we randomly drew another 1000 sequences from a
+standard normal distribution ($\mathcal{N}(0,1)$ for each 
+$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
+A second dataset was created by replacing elements with randomly drawn outliers, as described above.
+For each sequence, the p-value for the respective test was used to reject the null
+hypothesis that the sequence is (centrally) normal. The null hypothesis was rejected if $p \leq 0.05$.
+The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
+Figure \ref{empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
+centrally normal data (with tails of the distribution not being normally distributed),
+whereas the central normality test and Shapiro-Wilk test overestimated type I error rates.
+
+```{r empirical-central-normality-error_rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers. ",
+  "ECN: empirical central normality; CN: central normality"
+)
+
+.plot_test_dependency_sample_size(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
+In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
+central normality test to assess central normality of features that are composed
+of a mixture of samples drawn from two normal distributions ($n = 100$ each).
+With increased separation of the underlying normal distributions, the probability of the
+feature being centrally normal decreases, as expected.
+
+```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Bi-modal distributions and empirical central normality test results. ",
+  "The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) ",
+  "drawn from normal distributions that are offset by a distance $d$. ",
+  "We use the empirical centrally normality test to compute the probability for the hypothesis that ",
+  "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
+  "the probability that the feature is centrally normal decreases. ",
+  "Quantile-quantile plots are drawn below each distribution."
+)
+
+.plot_bimodal_distribution_test(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
 ## Robust transformations
@@
 two_sided_function_parameters <- .get_optimised_weighting_function_parameters(
-  manuscript_dir = manuscript_dir,
-  side = "both"
+  manuscript_dir = manuscript_dir
 )
@@
 
-## Central normality and empirical central normality test
-
-To develop a test for central normality we need to consider two
-parameters: the central portion $\kappa$ as a fixed parameter, and test
-statistic $\tau_{\text{ecn}}$. We will first define the central portion
-$\kappa$.
-
-For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
-we randomly drew $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
-This dataset was used to compute the critical test statistic values for the central normality test.
-Additionally, we added $10 \%$ outliers by randomly replacing elements of each ordered sequence.
-The dataset with outliers was used to compute the critical test statistic values for the empirical variant of the central normality test.
-For each sequence we then computed $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
-
-Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
-as a function of $n$ for different values of $\kappa$. With decreasing
-$\kappa$, the test statistic curve decreases. The curves
-for $\kappa = 0.60$ and $\kappa = 0.70$ are similar, whereas curves for 
-$\kappa \geq 0.90$ are affected by outliers, as can be observed by comparing curves
-of the central normality test with those of the empirical variant.
-
-Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
-we used $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
-normality tests are shown in Appendix E.
-
-```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
-cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$ of the (empirical) ",
-  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
-  "The critical test statistics for central normality test are determined using fully normal data, ",
-  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
-  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
-  "emp: empirical"
-)
-
-.plot_test_statistic_centrality(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-
-```
-
-To assess type I error rates for the (empirical) central normality test and compare
-these to the Shapiro-Wilk test for normality, we randomly drew another 1000 sequences from a
-standard normal distribution ($\mathcal{N}(0,1)$ for each 
-$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
-A second dataset was created by replacing elements with randomly drawn outliers, as described above.
-For each sequence, the p-value for the respective test was used to reject the null
-hypothesis that the sequence is (centrally) normal. The null hypothesis was rejected if $p \leq 0.05$.
-The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
-Figure \ref{empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
-centrally normal data (with tails of the distribution not being normally distributed),
-whereas the central normality test and Shapiro-Wilk test overestimated type I error rates.
-
-```{r empirical-central-normality-error_rate, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
-cap <- paste0(
-  "Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers. ",
-  "ECN: empirical central normality; CN: central normality"
-)
-
-.plot_test_dependency_sample_size(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-
-```
-
-In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
-central normality test to assess central normality of features that are composed
-of a mixture of samples drawn from two normal distributions ($n = 100$ each).
-With increased separation of the underlying normal distributions, the probability of the
-feature being centrally normal decreases, as expected.
-
-```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
-cap <- paste0(
-  "Bi-modal distributions and empirical central normality test results. ",
-  "The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) ",
-  "drawn from normal distributions that are offset by a distance $d$. ",
-  "We use the empirical centrally normality test to compute the probability for the hypothesis that ",
-  "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
-  "the probability that the feature is centrally normal decreases. ",
-  "Quantile-quantile plots are drawn below each distribution."
-)
-
-.plot_bimodal_distribution_test(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-
-```
 ## Simulated data

### 2025-03-30 manuscript_rmarkdown/manuscript_appendix.Rmd
Reworked robust transformations.

@@
 
+# data[outlier == FALSE, "error" := residuals]
+# data[outlier == TRUE, "error" := residuals_central]
+# data_baseline <- data[transformation == "conventional"][, mget(c("dataset", "outlier", "experiment_id", "error"))]
+# data.table::setnames(data_baseline, "error", "reference_error")
+# data <- data[data_baseline, on = c("dataset", "outlier", "experiment_id")]
+# data[, "fractional_error" := error / reference_error]
+
 # Add rank based on residual error for central normality.
-data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
 
@@
 ```
+
 ```{=latex}
@@
 \caption{
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst.
+}
 \label{tab:comparison_methods_simulations}
@@
 
-none                                  & & 8.51          & 6.33          & 7.31          & 5.55          & 7.42          & 6.39 \\
-conventional                          & & 3.72          & 3.78          & 3.83          & 4.74          & 6.47          & 5.47 \\
-conventional (z-score norm.)          & & 4.72          & 2.89          & 6.37          & 3.43          & 5.42          & 3.50 \\
-conventional (robust scaling)         & & 4.07          & 4.31          & 5.36          & 4.23          & 4.41          & 4.32 \\
-Raymaekers-Rousseeuw                  & & 4.50          & 7.03          & 3.97          & 8.12          & 6.25          & 5.57 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 5.26          & 6.36          & 6.22          & 6.42          & 5.26          & 6.92 \\
-Raymaekers-Rousseeuw (robust scaling) & & 4.71          & 7.84          & 5.31          & 7.06          & 4.35          & 7.58 \\
-invariant                             & & \textbf{3.15} & 3.72          & \textbf{3.27} & 3.14          & \textbf{2.69} & 3.05 \\
-robust invariant                      & & 6.37          & \textbf{2.73} & 3.36          & \textbf{2.30} & 2.72          & \textbf{2.20} \\
+none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
+conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
+conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
+conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
+Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
+invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
+robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\
 

### 2025-03-30 manuscript_rmarkdown/manuscript_appendix.tex
Reworked robust transformations.

--- 2025-03-27 manuscript_rmarkdown/manuscript_appendix.tex
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-03-27}
+\date{2025-03-30}
 
@@
 }
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst.
+}
+\label{tab:comparison_methods_simulations}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
+conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
+conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
+conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
+Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
+invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
+robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\
+
+\bottomrule
+\end{tabular}
 \end{center}

### 2025-03-30 R/R/ParameterEstimators.R

@@ -366,7 +366,7 @@ setMethod(
   function(transformer, estimator, ...) {
 
     if (transformer@robust) {
-      return("empirical_probability_triangle")
+      return("empirical_probability_cosine")
 
     } else {
       return("none")

### 2025-03-30 manuscript_rmarkdown/simulations.R

@@ -435,63 +435,61 @@
 
 
 
-.get_optimised_weighting_function_parameters <- function(manuscript_dir, side = "both") {
+.get_optimised_weighting_function_parameters <- function(
+    manuscript_dir
+) {
+
   # generator ------------------------------------------------------------------
   experiment_args <- coro::generator(
     function(
         manuscript_dir,
-        side,
-        transformation_methods = NULL,
-        estimation_methods = NULL) {
-      if (side == "both") {
-        file_dir <- "robustness_comparison"
-      } else if (side == "right") {
-        file_dir <- "robustness_comparison_right"
-      }
+        transformation_method = NULL,
+        estimation_methods = NULL
+    ) {
 
-      if (is.null(transformation_methods)) {
-        transformation_methods <- c("box_cox", "yeo_johnson")
-      }
+      file_dir <- "robustness_parameters"
 
       if (is.null(estimation_methods)) {
-        estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
+        estimation_methods <- setdiff(
+          power.transform:::..estimators_all(),
+          power.transform:::..estimators_raymaekers_robust())
+        # estimation_methods <- "mle"
       }
 
       for (estimation_method in estimation_methods) {
-        for (transformation_method in transformation_methods) {
-          # Non-robust transformation.
-          fun_args <- list(
+
+        # Non-robust transformation.
+        fun_args <- list(
+          "method" = transformation_method,
+          "estimation_method" = estimation_method,
+          "robust" = FALSE,
+          "invariant" = TRUE
+        )
+
+        opt_args <- list()
+
+        file_name <- file.path(
+          manuscript_dir,
+          file_dir,
+          paste0(
+            transformation_method, "_",
+            estimation_method, "_",
+            "non_robust.RDS"
+          )
+        )
+
+        if (!file.exists(file_name)) {
+          # Only yield arguments if the file does not exist.
+          yield(list(
+            "name" = "non-robust",
             "method" = transformation_method,
             "estimation_method" = estimation_method,
-            "robust" = FALSE,
-            "invariant" = TRUE
-          )
-
-          opt_args <- list()
-
-          file_name <- file.path(
-            manuscript_dir,
-            file_dir,
-            paste0(
-              transformation_method, "_",
-              estimation_method, "_",
-              "non_robust.RDS"
-            )
-          )
-
[…]

### 2025-03-30 manuscript_rmarkdown/manuscript.Rmd
Remove distribution of errors (as that no longer makes sense).

@@
 
-Figure \ref{fig:optimised-weighting-function-parameters} shows the distribution of errors $\left| \lambda_r - \lambda_0 \right|$ for non-robust and robust transformations using the optimal weighting function parameters.
-In the presence of outliers, the Yeo-Johnson transformation yielded smaller errors than the Box-Cox transformation.
-For both transformations, weighting based on empirical probabilities yielded the most consistent $\lambda$ parameter estimates in the presence of outliers.
-Other methods did not outperform the non-robust transformation method.
-
-```{r optimised-weighting-function-parameters, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=3.5}
-cap <- paste0(
-  "Robustness of power transformations after optimising weighting function parameters. ",
-  "The distribution of errors, i.e. the difference between robustly fitted $\\lambda^r$ ",
-  "and expected $\\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly ",
-  "parametrised asymmetric generalised normal distributions with up to 10% outliers on either side ",
-  "of the distribution. The top and bottom panel show errors for the location- and scale-invariant Box-Cox ",
-  "and Yeo-Johnson transformations, respectively. ",
-  "In each panel, the error distribution for the non-robust transformation is shown to the left ",
-  "for comparison with different weighting methods. ",
-  "Note that a square root transformation was applied along the $y$-axis for display purposes. ",
-  "Median errors are shown above each distribution."
-)
-
-.plot_optimised_weighting_function_parameters(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-```
-
 ## Simulated data

### 2025-03-31 manuscript_rmarkdown/manuscript.Rmd
Inserted new values.

@@
 Outliers may be present in data and affect estimation of transformation parameters.
-The log-likelihood function can be weighted to assign less weight to outlier instances, see equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}.
+The log-likelihood function can be weighted to assign less weight to outlier instances,
+see equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}.
 We proposed three weighting function: step, triangle and tapered cosine, that have one, two and two parameters, respectively.
-Each weighting function then takes one of three values as input: probabilities of the empirical distribution of the original feature, the z-score of the transformed feature values, or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
-
-To determine the weighting function parameters for each of the nine combinations, $m_d=100$ asymmetric generalised normal distributions were drawn.
-Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+Each weighting function then takes one of three values as input:
+probabilities of the empirical distribution of the original feature,
+the z-score of the transformed feature values,
+or the residual error between the z-score of the transformed feature values and their expected z-score based on the normal distribution.
+
+To determine the weighting function parameters for each of the nine combinations,
+$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
+drawn from randomly parametrised asymmetric generalised normal distributions.
+Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
+and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
 Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
-$n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, with $\gamma \sim U\left(2, 4\right)$, i.e., between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.
-
-Outlier values were then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
-This was repeated $m_{\text{out}} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
-Thus up to 10 percent of the values could be replaced by outliers.
+To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
+with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.
+
+Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
 Outlier values were set according to @Tukey1977-xm, as follows.
@@
 \end{equation}
-```
 
 $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
-Outlier values randomly replaced values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$, with $j \in \left\{ 1, \ldots, m_{\text{out}} \right\}$
-
-To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for robust transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:
-
-```{=latex}
-\begin{equation}
-\label{eqn:minimisation-weighting}
-\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
-\end{equation}
-```
+Outlier values randomly replaced elements in $\mathbf{X}_i$.
+
+To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
+we minimised $\sum_{i=1}^{m_d = 100}\tau_{n, \kappa = 0.70, i}$, i.e. the mean of residual 
+errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.
 
 Minimisation was conducted using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb].
-The resulting weighting function parameters for weighted MLE are shown in Tables \ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} for robust location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
+The resulting weighting function parameters for weighted MLE are shown in Tables
+\ref{tab:optimal-weighting-parameters-box-cox} and \ref{tab:optimal-weighting-parameters-yeo-johnson} 
+for robust location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
 
@@
 
-```{=latex}
 \begin{table}
@@
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
-The {optimal} column shows the optimal parameter values.
+The \textit{optimal} column shows the optimal parameter values.
 The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+Lower loss indicates better transformation to central normality in the presence of outliers.}
 \label{tab:optimal-weighting-parameters-box-cox}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 4.98 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.75 & ---  & ---       & ---  & 2.83 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.82 & 2.63 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.89 & 2.63 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 5.04 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.20 & 1.96 & $(0, 10]$ & 4.90 & 4.96 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.22 & 1.96 & $(0, 10]$ & 5.12 & 5.23 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 5.07 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 5.18 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 5.11 \\
 \bottomrule
@@
 \end{table}
-```
-
-```{=latex}
+
 \begin{table}
@@
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
-The {optimal} column shows the optimal parameter values.
+The \textit{optimal} column shows the optimal parameter values. 
 The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+Lower loss indicates better transformation to central normality in the presence of outliers.}
 \label{tab:optimal-weighting-parameters-yeo-johnson}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 5.09 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 2.85 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.83 & 2.46 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.04 & 0.95 & $(0, 1]$  & 0.88 & 2.45 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 4.50 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.42 & 1.96 & $(0, 10]$ & 4.35 & 4.08 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 4.18 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 5.79 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.09 & 5.89 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.07 & 1.00 & $(0, 10]$ & 1.08 & 5.81 \\
 \bottomrule
@@
 \end{table}
-```
 

### 2025-03-31 manuscript_rmarkdown/manuscript.Rmd
Added  simulation data to the main manuscript.

@@
 To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
-we minimised $\sum_{i=1}^{m_d = 100}\tau_{n, \kappa = 0.70, i}$, i.e. the mean of residual 
+we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.70, i}$, i.e. the mean of residual 
 errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.
@@
 
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# Add rank based on residual error for central normality.
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
+```
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
+conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
+conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
+conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
+Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
+invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
+robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+```{r include = FALSE, eval = FALSE}
+
 data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))

### 2025-03-31 manuscript_rmarkdown/manuscript_appendix.Rmd
Added  simulation data to the main manuscript.

--- 2025-03-27 manuscript_rmarkdown/manuscript_appendix.Rmd
@@
 \end{table}
+```
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean", method = "box_cox")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty", method = "box_cox")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted", method = "box_cox")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# Add rank based on residual error for central normality.
+data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
 ```

### 2025-04-02 manuscript_rmarkdown/manuscript.Rmd
Expand description of text.

@@
 drawn from randomly parametrised asymmetric generalised normal distributions.
-Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
+Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
 and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
@@
 
-## Simulated data
-
-[Generated data: clean, n = 30, n = 100, n = 1000, lambda = -1.00, 1.00, 3.00]
+## Assessing transformation using simulated data
+
+Three datasets with 10000 sequences each were created to assess transformation
+to normality. For each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
+elements were randomly drawn, with $\gamma \sim U\left(1, 4\right)$, 
+resulting in sequences with between $10$ and $10000$ elements.
+
+- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
+  transformed using an inverse power transformation with randomly drawn 
+  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
+  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
+  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
+  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
+  were shifted into the positive domain, with minimum value $1 \cdot 10^{-4}$.
+
+- A \textit{dirty} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
+  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
+  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+
+- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)$,
+  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
+  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+
+A dataset with outliers was created for each of the above datasets by replacing
+10 percent of elements in each sequence, as described earlier.
+
+In addition to no power transformation and location and scale-invariant power transformations,
+conventional and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2024-zf] were assessed.
+For the latter two methods, normalisation before standardisation using was additionally assessed using
+the following two methods:
+
+1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of sequence $\mathbf{X}$.
+
+2. robust scaling: $x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$,
+   with $\text{IQR}$ representing the interquartile range of $\mathbf{X}$.
+
+This results in nine types of power transformation. Transformation parameters
+were optimised for each sequence. Subsequently the sum of residual errors 
+and sum of residual errors of the central portion ($\kappa = 0.70$) were computed
+for each sequence after transformation. Then, each method was ranked according 
+to the sum of residual errors (data without outliers) or the sum of residual
+errors of the central portion (data with outliers). The average rank of each
+transformation method was computed over all 10000 sequences in each dataset.
+For Yeo-Johnson transformations average ranks are shown in Table \ref{tab:comparison_methods_simulation_yeo_johnson}.
 
@@
 \caption{
-Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
 (with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
@@
 }
-\label{tab:comparison_methods_simulations}
+\label{tab:comparison_methods_simulations_yeo_johnson}
 \begin{tabular}{l | l r r r r r r}

### 2025-04-03 manuscript_rmarkdown/manuscript.Rmd
Switch central portion to 80%

@@
 $\kappa$, the test statistic curve decreases. The curves
-for $\kappa = 0.60$ and $\kappa = 0.70$ are similar, whereas curves for 
+for $\kappa = 0.60$, $\kappa = 0.70$ and $\kappa = 0.80$ are similar, whereas curves for 
 $\kappa \geq 0.90$ are affected by outliers, as can be observed by comparing curves
@@
 Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
-we used $\kappa = 0.70$. Lookup tables and curves for critical statistic values for central normality and empirical central
+we used $\kappa = 0.80$. Lookup tables and curves for critical statistic values for central normality and empirical central
 normality tests are shown in Appendix E.
@@
   "Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers. ",
-  "ECN: empirical central normality; CN: central normality"
+  "The empirical central normality (ECN) test uses $\\kappa = 0.80$. CN: central normality"
 )
@@
 To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
-we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.70, i}$, i.e. the mean of residual 
-errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.
+we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.80, i}$, i.e. the mean of residual 
+errors of the central 80% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.
 

### 2025-04-03 manuscript_rmarkdown/manuscript_appendix.Rmd
Switch central portion to 80%

@@
   manuscript_dir = manuscript_dir,
-  k = 0.70,
+  k = 0.80,
   for_manuscript = FALSE
@@
   manuscript_dir = manuscript_dir,
-  k = 0.70,
+  k = 0.80,
   for_manuscript = FALSE,
@@
   manuscript_dir = manuscript_dir,
-  k = 0.70,
+  k = 0.80,
   for_manuscript = TRUE
@@
 data <- .assess_standardisation_before_transformation()
-data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
-data[, "residuals" := round(residuals, digits = 1L)]
+data <- data[, mget(c("residuals_central", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals_central" := round(residuals_central, digits = 1L)]
 data[, "lambda" := round(lambda, digits = 1L)]
 data[, "p_value" := round(p_value, digits = 3L)]
-data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
 data.table::dcast(data, transformation ~ dataset, value.var = "lambda")

### 2025-04-03 R/R/GoodnessOfFit.R

@@ -94,8 +94,8 @@ cn.test <- function(x, transformer = NULL, robust = TRUE) {
     lookup_table <- empirical_central_normality_lookup_table
   }
 
-  # The test uses a central portion kappa = 0.70.
-  residual_data <- residual_data[p_observed >= 0.15 & p_observed <= 0.85]
+  # The test uses a central portion kappa = 0.80.
+  residual_data <- residual_data[p_observed >= 0.1 & p_observed <= 0.9]
 
   # Compute mean residual error for the central portion.
   test_statistic_value <- mean(abs(residual_data$residual))

### 2025-04-03 manuscript_rmarkdown/plots.R

@@ -1067,466 +1067,6 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
-# .plot_residuals <- function(plot_theme, manuscript_dir) {
-#   # Prevent warnings due to non-standard evaluation.
-#   has_outliers <- residual <- threshold <- residual_error <- method <- p <- NULL
-#
-#   if (!file.exists(file.path(manuscript_dir, "manuscript_residual_error.RDS"))) {
-#     # Get data
-#     data <- .get_goodness_of_fit_data(manuscript_dir = manuscript_dir)
-#
-#     outlier_free_data <- data[
-#       has_outliers == FALSE,
-#       list("residual_error" = stats::quantile(residual, probs = 0.99)),
-#       by = c("p", "method")
-#     ]
-#
-#     data <- data[, list(
-#       "residual_50" = stats::quantile(residual, probs = 0.50),
-#       "residual_90" = stats::quantile(residual, probs = 0.90),
-#       "residual_95" = stats::quantile(residual, probs = 0.95),
-#       "residual_99" = stats::quantile(residual, probs = 0.99),
-#       "residual_max" = max(residual)
-#     ),
-#     by = c("p", "method")
-#     ]
-#
-#     data <- data.table::melt(
-#       data = data,
-#       id.vars = c("p", "method"),
-#       variable.name = "threshold",
-#       value.name = "residual_error"
-#     )
-#
-#     data$threshold <- factor(
-#       x = data$threshold,
-#       levels = c("residual_50", "residual_90", "residual_95", "residual_99", "residual_max"),
-#       labels = c("50 %", "90 %", "95 %", "99 %", "100 %")
-#     )
-#
-#     saveRDS(
-#       data,
-#       file.path(manuscript_dir, "manuscript_residual_error.RDS")
-#     )
-#   } else {
-#     data <- readRDS(file.path(manuscript_dir, "manuscript_residual_error.RDS"))
-#
-#     outlier_free_data <- data[
-#       has_outliers == FALSE,
-#       list("residual_error" = stats::quantile(residual, probs = 0.99)),
-#       by = c("p", "method")
-#     ]
-#   }
-#
[…]

### 2025-04-03 manuscript_rmarkdown/manuscript.Rmd
Update weights in table.

@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 4.98 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.75 & ---  & ---       & ---  & 2.83 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.82 & 2.63 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.89 & 2.63 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 5.04 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.20 & 1.96 & $(0, 10]$ & 4.90 & 4.96 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.22 & 1.96 & $(0, 10]$ & 5.12 & 5.23 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 5.07 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 5.18 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 5.11 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.06 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.16 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.86 & 2.93 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.90 & 2.96 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 6.21 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.09 & 1.96 & $(0, 10]$ & 4.92 & 6.15 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.88 & 1.96 & $(0, 10]$ & 3.43 & 6.68 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 6.25 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 6.47 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 6.28 \\
 \bottomrule
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 5.09 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 2.85 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.83 & 2.46 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.04 & 0.95 & $(0, 1]$  & 0.88 & 2.45 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 4.50 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.42 & 1.96 & $(0, 10]$ & 4.35 & 4.08 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 4.18 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 5.79 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.09 & 5.89 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.07 & 1.00 & $(0, 10]$ & 1.08 & 5.81 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.19 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.13 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.87 & 2.73 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.14 & 0.95 & $(0, 1]$  & 0.91 & 2.80 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 5.50 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.30 & 1.96 & $(0, 10]$ & 5.91 & 5.78 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 5.16 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 7.51 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.08 & 7.61 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.17 & 1.00 & $(0, 10]$ & 1.54 & 8.80 \\
 \bottomrule

### 2025-04-04 manuscript_rmarkdown/manuscript_appendix.Rmd
Rework optimisation of weighting parameters to use a combined loss
Prevents myopic focus on reducing residuals.

@@
 data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))
-data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
+data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
 data[, "residuals" := round(residuals, digits = 1L)]
+data[, "residuals_central" := round(residuals_central, digits = 1L)]
 data[, "lambda" := round(lambda, digits = 1L)]
@@
 data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
 data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
@@
 
-# Yeo-Johnson on clean and dirty data.
+# Box-Cox on clean and dirty data.
 clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean", method = "box_cox")
@@
 # Add rank based on residual error for central normality.
-data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
 
@@
 
+```
+
+```{r include = FALSE, eval = FALSE}
+
+data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0), method = "box_cox")
+data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "residuals_central" := round(residuals, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
 ```

### 2025-04-04 R/R/ParameterEstimators.R

@@ -366,7 +366,9 @@ setMethod(
   function(transformer, estimator, ...) {
 
     if (transformer@robust) {
-      return("empirical_probability_triangle")
+      if (transformer@method == "box_cox") return("empirical_probability_step")
+
+      return("empirical_probability_cosine")
 
     } else {
       return("none")

### 2025-04-04 manuscript_rmarkdown/simulations.R

@@ -618,8 +618,42 @@
           x_outlier[sample(seq_along(x), size = n_draw, replace = FALSE)] <- outlier
         }
 
+        # Determine transformation parameter lambda for clean data.
+        clean_lambda_yeo_johnson <- power.transform::get_lambda(
+          power.transform::find_transformation_parameters(
+            x = x,
+            method = "yeo_johnson",
+            robust = FALSE,
+            invariant = TRUE
+          )
+        )
+
+        # For Box-Cox, shift all data to positive.
+        if (any(x_outlier <= 0.0)) {
+          x_min <- min(x_outlier)
+          x_outlier_bc <- x_outlier - x_min + 1.0
+          x_bc <- x - x_min + 1.0
+        } else {
+          x_outlier_bc <- x_outlier_bc
+          x_bc <- x
+        }
+
+        # Determine transformation parameter lambda for clean data.
+        clean_lambda_box_cox <- power.transform::get_lambda(
+          power.transform::find_transformation_parameters(
+            x = x_bc,
+            method = "box_cox",
+            robust = FALSE,
+            invariant = TRUE
+          )
+        )
+
+
         return(list(
-          "x" = x_outlier,
+          "x_bc" = x_outlier_bc,
+          "x_yj" = x_outlier,
+          "lambda_clean_bc" = clean_lambda_box_cox,
+          "lambda_clean_yj" = clean_lambda_yeo_johnson,
           "parameters" = parameters
         ))
       },
@@ -633,7 +667,7 @@
 
   # Helper function for computing transformer parameters under optimisation
   # constraints.
-  .compute_robust_error <- function(
+  .compute_loss <- function(
     x,
     fun_args,
     opt_args
@@ -647,11 +681,17 @@
       # Prevent notes
       method <- estimation_method <- NULL
 
+      if (fun_args$method == "box_cox") {
+        data <- x$x_bc
+      } else {
+        data <- x$x_yj
+      }
+
       # Create transformer.
       transformer <- suppressWarnings(do.call(
         power.transform::find_transformation_parameters,
         args = c(
-          list("x" = x$x),
+          list("x" = data),
           list("weighting_function_parameters" = opt_args),
           fun_args
         )
@@ -660,21 +700,36 @@
       # Transform data
       transformed_data <- tryCatch(
[…]

### 2025-04-08 manuscript_rmarkdown/manuscript.Rmd
Update to text and tables.

@@
 
+```{r include = FALSE, eval = FALSE}
+# This creates and sets the lookup table for critical test statistic.
+load(file = "./R/sysdata.rda")
+
+central_normality_lookup_table <- .get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.80,
+  for_manuscript = FALSE
+)
+
+empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.80,
+  for_manuscript = FALSE,
+  with_outliers = TRUE
+)
+
+usethis::use_data(
+  central_normality_lookup_table,
+  empirical_central_normality_lookup_table,
+  two_sided_function_parameters, 
+  internal = TRUE, 
+  overwrite = TRUE
+)
+```
+
 Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
@@
 Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
-we used $\kappa = 0.80$. Lookup tables and curves for critical statistic values for central normality and empirical central
-normality tests are shown in Appendix E.
+we used $\kappa = 0.80$. Critical statistic values for central normality and empirical central
+normality tests are shown are shown in Table \ref{tab:central-normality-critical-statistic} and Table 
+\ref{tab:empirical-central-normality-critical-statistic}, respectively.
 
@@
 cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$ of the (empirical) ",
+  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$ of the (empirical) ",
   "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
@@
 To determine the weighting function parameters for each of the nine combinations,
-$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
+$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
 drawn from randomly parametrised asymmetric generalised normal distributions.
@@
 To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
-with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.
+with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
 
@@
 To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
-we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.80, i}$, i.e. the mean of residual 
-errors of the central 80% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.
+we minimised a composite loss $L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}$.
+The composite loss consisted of two components: a loss term $L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}$, i.e. the mean of residual 
+errors of the central 80% of elements of each sequence $\mathbf{X}_i$, which aimed at optimising central normality; 
+and a loss term $L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)$,
+with $\lambda_{0,i}$ and $\lambda_{i}$ transformation parameters found
+for sequence $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
+and tolerance parameter $\xi = 0.5$ for Box-Cox and $\xi = 0.3$ for Yeo-Johnson power transformations.
+This second term aimed to prevent solutions that provide small improvements in central normality
+at the cost of a poor fit of the tails of sequences.
 
@@
 The \textit{optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better transformation to central normality in the presence of outliers.}
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
 \label{tab:optimal-weighting-parameters-box-cox}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.06 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.16 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.86 & 2.93 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.90 & 2.96 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 6.21 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.09 & 1.96 & $(0, 10]$ & 4.92 & 6.15 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.88 & 1.96 & $(0, 10]$ & 3.43 & 6.68 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 6.25 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 6.47 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 6.28 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 49.6 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.80 & ---  & ---       & ---  & 38.5 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.01 & 1.00 & $(0, 1]$  & 0.86 & 43.7 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 1.00 & $(0, 1]$  & 0.90 & 42.7 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.39 & ---  & ---       & ---  & 52.2 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 2.39 & 2.40 & $(0, 10]$ & 4.92 & 52.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.07 & 3.63 & $(0, 10]$ & 3.43 & 57.4 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.83 & ---  & ---       & ---  & 53.1 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.80 & 0.81 & $(0, 10]$ & 0.84 & 54.4 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.76 & 0.77 & $(0, 10]$ & 0.76 & 53.2 \\
 \bottomrule
@@
 The \textit{optimal} column shows the optimal parameter values. 
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better transformation to central normality in the presence of outliers.}
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
 \label{tab:optimal-weighting-parameters-yeo-johnson}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.19 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.13 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.87 & 2.73 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.14 & 0.95 & $(0, 1]$  & 0.91 & 2.80 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 5.50 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.30 & 1.96 & $(0, 10]$ & 5.91 & 5.78 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 5.16 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 7.51 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.08 & 7.61 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.17 & 1.00 & $(0, 10]$ & 1.54 & 8.80 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 42.1 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.78 & ---  & ---       & ---  & 35.1 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.20 & 0.95 & $(0, 1]$  & 1.00 & 34.0 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.54 & 0.95 & $(0, 1]$  & 1.00 & 32.6 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.32 & ---  & ---       & ---  & 43.0 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.28 & 1.96 & $(0, 10]$ & 3.43 & 49.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.41 & 1.96 & $(0, 10]$ & 3.75 & 51.3 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.92 & ---  & ---       & ---  & 65.4 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.87 & 1.00 & $(0, 10]$ & 0.87 & 66.0 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.06 & 1.00 & $(0, 10]$ & 1.07 & 67.1 \\
 \bottomrule
@@
 
-## Assessing transformation using simulated data
+## Assessing transformations using simulated data
 
@@
   respectively. Prior to inverse transformation, sequences for Box-Cox transformations
-  were shifted into the positive domain, with minimum value $1 \cdot 10^{-4}$.
+  were shifted into the positive domain, with minimum value $1$.
 
@@
 were optimised for each sequence. Subsequently the sum of residual errors 
-and sum of residual errors of the central portion ($\kappa = 0.70$) were computed
+and sum of residual errors of the central portion ($\kappa = 0.80$) were computed
 for each sequence after transformation. Then, each method was ranked according 
@@
 transformation method was computed over all 10000 sequences in each dataset.
-For Yeo-Johnson transformations average ranks are shown in Table \ref{tab:comparison_methods_simulation_yeo_johnson}.
+Average ranks for Yeo-Johnson transformations are shown in Table
+\ref{tab:comparison_methods_simulation_yeo_johnson}.
+Location and shift-invariant Yeo-Johnson transformation ranked best for
+all datasets without outliers. The robust variant ranked best in dirty and shifted
+datasets with outliers, but not in the clean dataset.
+Results for Box-Cox transformations are shown in Table \ref{tab:comparison_methods_simulations_box_cox}.
 
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
@@
 
-none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
-conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
-conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
-conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
-Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
-Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
-invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
-robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\
+none                                  & &         8.51  &         7.60  &         7.33  &         6.15  &         7.44  &         6.60 \\
+conventional                          & &         3.70  &         5.70  &         3.87  &         7.06  &         6.49  &         5.93 \\
+conventional (z-score norm.)          & &         4.72  &         5.68  &         6.38  &         5.01  &         5.43  &         4.87 \\
+conventional (robust scaling)         & &         4.06  &         6.77  &         5.37  &         5.75  &         4.41  &         5.63 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.20  &         4.01  &         4.13  &         6.27  &         5.65 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.27  & \textbf{2.66} &         6.23  &         3.59  &         5.28  &         3.48 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.95  &         5.32  &         3.77  &         4.36  &         3.66 \\
+invariant                             & & \textbf{3.13} &         6.77  & \textbf{3.22} &         6.12  & \textbf{2.64} &         6.03 \\
+robust invariant                      & &         6.41  &         3.66  &         3.27  & \textbf{3.41} &         2.68  & \textbf{3.10} \\
 
@@
 
-Location- and scale-invariant power transformations are intended to yield improved transformations to normality in the presence of large shifts in location, distributions that due to location and scale are not centered near zero, or both.
+Location- and scale-invariant power transformations are intended to yield improved
+transformations to normality in the presence of large shifts in location, distributions that due to location and scale are not centered near zero, or both.
 Earlier, we assessed these transformations using simulated data.

### 2025-04-08 manuscript_rmarkdown/manuscript_appendix.Rmd
Update to text and tables.

@@
 
+# Appendix D: Central normality test and empirical central normality test critical values
+
+Critical values for the central normality test and its empirical variant are
+found in Table \ref{tab:central-normality-critical-statistic} and Table 
+\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+
+```{r include = FALSE, eval = FALSE}
+# This generates the table shown below.
+.get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.80,
+  for_manuscript = TRUE
+)
+
+```
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
+\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
+
+\midrule
+    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
+   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
+   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
+   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
+  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
+  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
+  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
+ 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
+ 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
+ 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+```{r include = FALSE, eval = FALSE}
+# This generates the table shown below.
+.get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir,
+  k = 0.80,
+  for_manuscript = TRUE,
+  with_outliers = TRUE
+)
+
+```
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
+
+\midrule
+    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
+   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
+   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
+   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
+  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
+  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
+  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
+ 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
+ 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
+ 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
+10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+cap <- paste0(
+  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80}$ of the (empirical) ",
+  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
+  "The critical test statistics for central normality test are determined using fully normal data, ",
+  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
+  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
+  "emp: empirical"
+)
+
+.plot_test_statistic_tau(
+  plot_theme = plot_theme,
+  manuscript_dir = manuscript_dir
+)
+
+```
+
 # Appendix D: Simulations on clean data
@@
 ```
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+If necessary, each sequence was shifted so that every element had a strictly positive value. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_box_cox}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
+conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
+conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
+conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
+Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
+invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
+robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.60$).
-
-# Appendix E: Empirical central normality test
-
-Critical values for the central normality test and its empirical variant are
-found in Table \ref{tab:central-normality-critical-statistic} and Table 
-\ref{tab:empirical-central-normality-critical-statistic}, respectively.
-
-```{r include = FALSE, eval = FALSE}
-# This writes the lookup table for critical test statistic.
-load(file = "./R/sysdata.rda")
-
-central_normality_lookup_table <- .get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = FALSE
-)
-
-empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = FALSE,
-  with_outliers = TRUE
-)
-
-usethis::use_data(
-  central_normality_lookup_table,
-  empirical_central_normality_lookup_table,
-  two_sided_function_parameters, 
-  internal = TRUE, 
-  overwrite = TRUE)
-```
-
-```{r include = FALSE, eval = FALSE}
-# This generates the table shown below.
-.get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = TRUE
-)
-
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
-
-\midrule
-5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
-10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
-20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
-50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
-100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
-200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
-500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
-1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
-2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
-5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
-10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-```{r include = FALSE, eval = FALSE}
-# This generates the table shown below.
-.get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.70,
-  for_manuscript = TRUE,
-  with_outliers = TRUE
-)
-
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:empirical-central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
-
-\midrule
-5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
-10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
-20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
-50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
-100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
-200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
-500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
-1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
-2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
-5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
-10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
-cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.70}$ of the (empirical) ",
-  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
-  "The critical test statistics for central normality test are determined using fully normal data, ",
-  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
-  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
-  "emp: empirical"
-)
-
-.plot_test_statistic_tau(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-
-```
 

### 2025-04-08 manuscript_rmarkdown/manuscript.Rmd
Update experimental results.

@@
 Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
-no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$, $p = 0.71$);
-conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.94$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.94$);
-location- and scale-invariant transformation ($\lambda = 0.9$, $\sum r_i = 8.8$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.1$, $p = 0.95$);
-and robust location- and scale-invariant transformation ($\lambda = 0.8$, $\sum r_i = 10.6$, $\mu_{YJ} = 1.2$, $\sigma_{YJ} = 1.0$, $p = 0.80$).
-
-Location- and scale-invariant transformation led to a lower overall residual sum, indicating a better transformation.
-Robust location- and scale-invariant transformation had a higher residual sum compared to the non-robust variant, which may be due to the lack of outliers in the data.
+no transformation ($\sum r_i = 16.5$, $p = 0.69$);
+conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.96$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.96$);
+location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 8.8$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 1.1$, $p = 0.98$);
+and robust location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 9.3$, $\mu_{YJ} = -1.0$, $\sigma_{YJ} = 1.1$, $p = 0.93$).
+
+Location- and scale-invariant transformation led to a lower overall residual errors, indicating a better transformation.
 Conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
-The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.78$).
+The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.93$).
 
@@
 Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
-The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
+The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p < 0.001$).
 Applying conventional and invariant Yeo-Johnson transformations to body mass yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
 no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.06$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.06$);
-location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$, $p = 0.18$);
-and robust location- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$, $p = 0.40$).
-
-Location- and scale-invariant transformation produced a lower overall residual sum, indicating a better transformation.
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.10$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.10$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$, $p = 0.28$);
+and robust location- and scale-invariant transformation ($\lambda = 0.3$, $\sum r_i = 22.0$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.9$, $p = 0.69$).
+
+Location- and scale-invariant transformation produced a lower overall residual errors, indicating a better transformation.
 Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
-The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
+The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.10$).
 
@@
 conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$, $p < 0.001$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 315$, $\mu_{YJ} = 6.4 \cdot 10^{153}$, $\sigma_{YJ} = 0.3 \cdot 10^{153}$, $p < 0.001$);
 location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 0.8$, $p < 0.001$);
-and robust location- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 0.9$, $p < 0.001$).
+and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 308$, $\mu_{YJ} = -1.3$, $\sigma_{YJ} = 1.2$, $p < 0.001$).
 
 Every transformation reduced the residual sum.
-The non-robust location- and scale-invariant transformation did not improve over conventional alternatives and yielded a data distribution lacking central normality (empirical central normality test: $p=0.05$).
-However, conventional transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
+None of the transformations yielded a centrally normal distribution.
+Conventional transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
 
@@
 Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 54$, $p=0.58$);
+no transformation (residual sum $\sum r_i = 54$, $p=0.72$);
 conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p < 0.001$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.32$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.37$);
 location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p < 0.001$);
-and robust location- and scale-invariant transformation ($\lambda = -0.9$, $\sum r_i = 50$, $\mu_{YJ} = 2.0$, $\sigma_{YJ} = 2.3$, $p=0.35$).
-
-Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution (empirical central normality test $p < 0.001$ for conventional and invariant transformations).
+and robust location- and scale-invariant transformation ($\lambda = 1.0$, $\sum r_i = 56$, $\mu_{YJ} = 1.7$, $\sigma_{YJ} = 1.0$, $p=0.76$).
+
+Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution
+(empirical central normality test $p < 0.001$ for conventional and invariant transformations).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
@@
 Applying conventional and invariant Yeo-Johnson transformations to this feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 110$, $p=0.86$);
-conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.02$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14$, $p=0.91$);
-location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.23$);
-and robust location- and scale-invariant transformation ($\lambda = -0.6$, $\sum r_i = 27$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.1$, $p=0.18$).
-
-The conventional non-robust transformation failed to produce a centrally normal distribution (empirical central normality test $p=0.02$).
+no transformation (residual sum $\sum r_i = 110$, $p=0.83$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.003$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14.3$, $p=0.88$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.15$);
+and robust location- and scale-invariant transformation ($\lambda = -0.3$, $\sum r_i = 30$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.2$, $p=0.18$).
+
+The conventional non-robust transformation failed to produce a centrally normal distribution (empirical central normality test $p=0.003$).
 Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

### 2025-04-08 manuscript_rmarkdown/manuscript_appendix.Rmd
Revised experimental results in appendix.

@@
 
-# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation
+# Appendix D: Experimental results
 
-The effect of using location- and scale-invariant transformations was investigated using real-world datasets.
+The effect of using location- and scale-invariant transformations was investigated
+using real-world datasets, as described in the main manuscript.
 
-## Invariance
+## Yeo-Johnson transformation
 
-Results for Box-Cox transformations of features without outliers are shown in Figure \ref{fig:experimental-results-invariance-appendix}.
+Additional results for Yeo-Johnson transformations are shown in Tables
+\ref{tab:experimental-results-appendix-residuals}, 
+\ref{tab:experimental-results-appendix-lambda},
+\ref{tab:experimental-results-appendix-p-value}.
 
-```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
-  "penguin body mass (middle row); ",
-  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
-  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
-  "the feature transformed using the conventional Box-Cox transformation ",
-  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
-  "and the feature transformed using the non-robust and robust location- and-scale ",
-  "invariant Box-Cox transformations (right column)."
-)
+```{r include = FALSE, eval = FALSE}
 
-data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
-data$plot
+data <- .assess_standardisation_before_transformation()
+data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "residuals_central" := round(residuals_central, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
 ```
 
-### Age of patients with lung cancer
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-residuals}
+\begin{tabular}{l | r r r r r}
 
-Applying conventional and invariant Box-Cox transformations to age of patients with lung cancer [@Loprinzi1994-cd] yielded the following results:
-no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$, $p = 0.71$);
-conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
-location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$, $p = 0.94$);
-and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$, $p = 0.93$).
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
 
-Compared to location- and scale-invariant Yeo-Johnson transformations, the Box-Cox transformations do not reduce residuals compared to conventional variants.
+\midrule
+none                                  & 16.5 & 110.1 & 54.5 & 328.4 & 48.0 \\
+conventional                          & 11.5 &  30.0 & 55.3 & 319.0 & 32.2 \\
+conventional (z-score norm.)          & 11.5 &  19.3 & 49.0 & 326.2 & 33.3 \\
+conventional (robust scaling)         & 11.3 &  31.8 & 53.3 & 324.5 & 32.2 \\
+Raymaekers-Rousseeuw                  & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\
+robust invariant                      &  9.3 &  30.2 & 55.8 & 308.1 & 22.0 \\
 
-### Penguin body mass
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-Applying conventional and invariant Box-Cox transformations to the body mass of penguins [@Gorman2014-eo] yielded the following results:
-no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
-location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$, $p = 0.10$);
-and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$, $p = 0.44$).
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-lambda}
+\begin{tabular}{l | r r r r r}
 
-Just as for location- and scale-invariant Yeo-Johnson transformations, Box-Cox transformations produced a lower overall residual sum compared to their conventional counterparts.
-Similarly, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
 
-### Latitude in the Ames housing dataset
+\midrule
+conventional                          & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.)          & 1.2 & -1.7 & -0.1 &  1.3 &  0.6 \\
+conventional (robust scaling)         & 1.3 &  0.0 &  0.2 &  1.4 &  0.4 \\
+Raymaekers-Rousseeuw                  & 2.0 &  1.1 &  0.8 & 95.4 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 1.3 &  1.4 &  1.0 &  1.3 &  0.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.4 &  1.3 &  1.0 &  1.4 &  0.4 \\
+invariant                             & 1.3 &  0.2 & -1.3 &  1.5 &  0.5 \\
+robust invariant                      & 1.3 & -0.3 &  1.0 &  1.1 &  0.3 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-Applying conventional and invariant Box-Cox transformations to the latitude of houses in the Ames housing dataset [@De-Cock2011-jf] yielded the following results:
-no transformation (residual sum $\sum r_i = 328$, $p < 0.001$);
-conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$, $p < 0.001$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$, $p < 0.001$);
-location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$, $p < 0.001$);
-and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$, $p < 0.001$).
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-p-value}
+\begin{tabular}{l | r r r r r}
 
-Similar to conventional Yeo-Johnson transformations (non-robust and robust), conventional Box-Cox transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
-Location- and scale-invariant Box-Cox transformations did not suffer from this issue.
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
 
-## Robustness against outliers
+\midrule
 
-Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.
+none                                  & 0.694 & 0.832     & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.003     & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.919 & 0.021     & $< 0.001$ & $< 0.001$ & 0.099 \\
+conventional (robust scaling)         & 0.933 & $< 0.001$ & $< 0.001$ & $< 0.001$ & 0.129 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.881     & 0.373     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.694 & 0.899     & 0.809     & $< 0.001$ & 0.099 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.900 & 0.880     & 0.811     & $< 0.001$ & 0.144 \\
+invariant                             & 0.976 & 0.147     & $< 0.001$ & $< 0.001$ & 0.276 \\
+robust invariant                      & 0.933 & 0.178     & 0.759     & $< 0.001$ & 0.688 \\
 
-```{r experimental-results-outlier-robustness-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
-cap <- paste0(
-  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
-  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
-  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
-  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
-  "the feature transformed using the conventional Box-Cox transformation ",
-  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
-  "and the feature transformed using the non-robust and robust location- and-scale ",
-  "invariant Box-Cox transformations (right column). ",
-  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
-)
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
-data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "box_cox")
-data$plot
+## Box-Cox transformation
+
+Results for Box-Cox transformations are shown in Tables
+
+```{r include = FALSE, eval = FALSE}
+
+data <- .assess_standardisation_before_transformation(method = "box_cox")
+data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "residuals_central" := round(residuals_central, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
 ```
-
-### Fuel efficiency in the Top Gear dataset
-
-The Top Gear dataset contains data on 297 vehicles, with outliers related highly fuel-efficient vehicles [@Alfons2021-kc].
-Applying conventional and invariant Box-Cox transformations to the fuel consumption feature yielded the following results:
-no transformation (residual sum $\sum r_i = 54$, $p=0.58$);
-conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p < 0.001$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.32$);
-location- and scale-invariant transformation ($\lambda = -0.7$, $\sum r_i = 44$, $\mu_{BC} = 0.6$, $\sigma_{BC} = 0.2$, $p < 0.001$);
-and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 59$, $\mu_{BC} = 2.4$, $\sigma_{BC} = 1.8$, $p=0.67$).
-
-### Maximum arterial wall thickness in an ischemic stroke dataset
-
-The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
-Applying conventional and invariant Box-Cox transformations to the maximum arterial wall thickness feature yielded the following results:
-no transformation (residual sum $\sum r_i = 110$, $p=0.86$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 33$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 127$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.90$);
-location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 28$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.02$);
-and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.60$).
 
@@
 
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
-AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-appendix}
-\begin{tabular}{l | r r r r r}
-
-\toprule
-feature & age & AWT & FE & latitude & PBM \\
-
-\midrule
-none & 16.5 & 110.1	& 54.5 & 328.4 & 48.0 \\
-conventional & 11.5 & 30.0 & 55.3 & 319.0 & 32.2 \\
-conventional (z-score norm.) & 11.5 & 19.3 & 49.0 & 326.2 & 33.3 \\
-conventional (robust scaling) & 11.3 & 31.8 & 53.3 & 324.5 & 32.2 \\
-Raymaekers-Rousseeuw & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
-Raymaekers-Rousseeuw (z-score norm.) & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
-Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
-invariant & 8.8 & 12.2 & 44.0 & 326.4 & 26.8 \\
-robust invariant & 10.6 & 27.1 & 49.5 & 310.6 & 23.1 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
-AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-lambda-appendix}
-\begin{tabular}{l | r r r r r}
-
-\toprule
-feature & age & AWT & FE & latitude & PBM \\
-
-\midrule
-conventional & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
-conventional (z-score norm.) & 1.2 & -1.7 & -0.1 & 1.3 & 0.6 \\
-conventional (robust scaling) & 1.3 & 0.0 & 0.2 & 1.4 & 0.4 \\
-Raymaekers-Rousseeuw & 2.0 & 1.1 & 0.8 & 95.4 & -0.5 \\
-Raymaekers-Rousseeuw (z-score norm.) & 1.3 & 1.4 & 1.0 & 1.3 & 0.6 \\
-Raymaekers-Rousseeuw (robust scaling) & 1.4 & 1.3 & 1.0 & 1.4 & 0.4 \\
-invariant & 1.3 & 0.2 & -1.3 & 1.5 & 0.5 \\
-robust invariant & 1.4 & -0.6 & 0.9 & 1.4 & 0.4 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{
-p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
-AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-p-value-appendix}
-\begin{tabular}{l | r r r r r}
-
-\toprule
-feature & age & AWT & FE & latitude & PBM \\
-
-\midrule
-none                                  & 0.712 & 0.864 & 0.577     & $< 0.001$ & $< 0.001$ \\
-conventional                          & 0.936 & 0.017 & $< 0.001$ & $< 0.001$ & 0.058 \\
-conventional (z-score norm.)          & 0.860 & 0.058 & $< 0.001$ & $< 0.001$ & 0.054 \\
-conventional (robust scaling)         & 0.881 & 0.001 & $< 0.001$ & $< 0.001$ & 0.076 \\
-Raymaekers-Rousseeuw                  & 0.936 & 0.905 & 0.320     & $< 0.001$ & 0.058 \\
-Raymaekers-Rousseeuw (z-score norm.)  & 0.605 & 0.922 & 0.635     & $< 0.001$ & 0.054 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.830 & 0.948 & 0.637     & $< 0.001$ & 0.085 \\
-invariant                             & 0.948 & 0.231 & $< 0.001$ & $< 0.001$ & 0.184 \\
-robust invariant                      & 0.797 & 0.182 & 0.352     & $< 0.001$ & 0.397 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-```
-
 # References

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.Rmd
Add experimental results for Box-Cox.

@@
 
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-residuals}
+\begin{tabular}{l | r r r r r-bc}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none                                  & 16.5 & 110.1 &  54.5 & 328.4 & 48.0 \\
+conventional                          & 11.5 &  33.5 &  55.7 & 318.9 & 32.2 \\
+conventional (z-score norm.)          & 12.9 &  44.5 &  58.8 & 305.3 & 27.3 \\
+conventional (robust scaling)         & 12.9 &  44.7 &  59.0 & 305.3 & 27.3 \\
+Raymaekers-Rousseeuw                  & 11.5 & 127.1 &  47.6 & 314.8 & 32.2 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & 13.1 &  54.8 & 127.5 & 510.2 & 23.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 18.9 & 100.2 & 113.7 & 510.2 & 23.6 \\
+invariant                             & 11.6 &  28.0 &  48.4 & 311.8 & 27.3 \\
+robust invariant                      & 12.8 & 150.1 &  60.5 & 646.0 & 27.4 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-lambda-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+conventional                          & 1.9 & -0.5 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.)          & 1.2 &  0.1 &  0.2 &  1.2 &  0.5 \\
+conventional (robust scaling)         & 1.2 &  0.1 &  0.1 &  1.2 &  0.5 \\
+Raymaekers-Rousseeuw                  & 1.9 &  1.1 &  0.8 & 95.9 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 1.2 &  0.6 & -0.5 &  0.6 &  0.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.9 &  1.0 & -0.5 &  0.6 &  0.3 \\
+invariant                             & 1.7 & -1.0 & -0.7 &  1.9 &  0.5 \\
+robust invariant                      & 1.3 &  1.2 &  1.1 &  0.4 &  0.1 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-p-value-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+none                                  & 0.694 & 0.832 & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.002 & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+conventional (robust scaling)         & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.871 & 0.365     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.891 & 0.268 & $< 0.001$ & $< 0.001$ &	0.495 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.562 & 0.789 & $< 0.001$ & $< 0.001$ &	0.492 \\
+invariant                             & 0.962 & 0.004 & $< 0.001$ & $< 0.001$ &	0.154 \\
+robust invariant                      & 0.899 & 0.890 & 0.866     & $< 0.001$ & 0.622 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
 # Appendix F: Normalisation before transformation

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.Rmd
Rearrange appendix sections.

@@
 
-For conventional power transformations with fixed location and scale parameters, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, or equivalently, $x_0$ and $s$ may be fixed:
+For conventional power transformations with fixed location and scale parameters, 
+the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, 
+or equivalently, $x_0$ and $s$ may be fixed:
 ```{=latex}
@@
 
-# Appendix C: Simulations with other optimisation criteria
-
-Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
+## Simulations with other optimisation criteria
+
+Invariance of location- and scale-invariant power transformations was assessed 
+using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
 This follows the simulation in the main manuscript, where MLE was used for optimization.
-In short, we first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
-The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
-The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
-We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
+In short, we first randomly drew $10000$ values from a normal distribution:
+$\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$,
+or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
+The second distribution was a right-skewed normal distribution 
+$\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
+The third distribution was a left-skewed normal distribution 
+$\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
+
+We then computed transformation parameter $\lambda$ using the original definitions
+(equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) 
+and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} 
+and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
 To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
@@
 
-# Appendix D: Simulations on clean data
+# Appendix E: Simulations on clean data
 
@@
 
-# Appendix D: Experimental results
+# Appendix F: Experimental results
 
@@
 
-Results for Box-Cox transformations are shown in Tables
+Results for Box-Cox transformations are shown in Tables \ref{tab:experimental-results-appendix-residuals-bc},
+\ref{tab:experimental-results-appendix-lambda-bc} and \ref{tab:experimental-results-appendix-p-value-bc}.
 
@@
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:experimental-results-appendix-residuals}
-\begin{tabular}{l | r r r r r-bc}
+\label{tab:experimental-results-appendix-residuals-bc}
+\begin{tabular}{l | r r r r r}
 
@@
 
-# Appendix F: Normalisation before transformation
-
-An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
-We assessed two normalisation methods were used:
-1. z-standardisation, i.e. $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of feature $\mathbf{X}$; 
-and 2.
-robust scaling, i.e.
-$x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$, with $\text{IQR}$ representing the interquartile range of feature $\mathbf{X}$.
-
-Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors after transformation to normality of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
-Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$, 
-\ref{tab:normalisation-before-transformation-p-value-appendix} shows p-values of the empirical central normality test.
-Normalisation prior to transformation prevents unexpectedly high $\lambda$ values in the presence of large location shifts, as can be observed for the latitude feature.
-Normalisation prior to transformation, however, does not consistently improve residual errors after transformation. 
-Interestingly, use of normalisation prior to transformation with robust power transformation using the method by Raymaekers and Rousseeuw [@Raymaekers2024-zf] led to higher residual errors for arterial wall thickness, fuel efficiency and latitude features.
-This is similar to observations by @Raymaekers2024-zf, who observed that their method can emphasize values of outliers after transformation.
-Location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
-Not surprisingly, normalisation before transformation does not noticeably improve central normality, as assessed using the empirical central normality test.
-
-```{r include = FALSE, eval = FALSE}
-
-data <- .assess_standardisation_before_transformation()
-data <- data[, mget(c("residuals_central", "lambda", "p_value", "transformation", "dataset"))]
-data[, "residuals_central" := round(residuals_central, digits = 1L)]
-data[, "lambda" := round(lambda, digits = 1L)]
-data[, "p_value" := round(p_value, digits = 3L)]
-data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
-data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
-data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
-```
-
 # References

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.Rmd
Main changes to appendix are now complete.

@@
 - \usepackage{multirow}
+- \usepackage{placeins}
 - \newtheorem{definition}{Definition}
@@
 
-# Appendix E: Simulations on clean data
-
-Robust transformations are hypothesised to have a cost in efficiency for data
-without outliers, i.e. clean data. Here we randomly draw data from a standard normal
-distribution $\mathcal{N}(0,1)$. Subsequently, we perform an inverse transformation
-$\left(\phi_{\text{YJ}}^{\lambda, 0, 1}\right)^{-1}$, with $\lambda \in \left(0, 2\right)$.
-
-These transformed data
-
-An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
-We assessed two normalisation methods were used:
-1. z-standardisation, i.e. $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of feature $\mathbf{X}$; 
-and 2.
-robust scaling, i.e.
-$x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$, with $\text{IQR}$ representing the interquartile range of feature $\mathbf{X}$.
-
-[Generated data: clean, n = 30, n = 100, n = 1000, lambda = 0.1, 1.0, 1.9]
-
-```{r include = FALSE, eval = FALSE}
-
-data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))
-data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
-data[, "residuals" := round(residuals, digits = 1L)]
-data[, "residuals_central" := round(residuals_central, digits = 1L)]
-data[, "lambda" := round(lambda, digits = 1L)]
-data[, "p_value" := round(p_value, digits = 3L)]
-data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
-data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
-data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
-data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-residuals-appendix}
-\small{
-\begin{tabular}{l | l r r r r r r r r r}
-
-\toprule
-& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
-transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
-
-\midrule
-
-none                                  & & 4.0 & 3.4 & 4.6 & 40.1 & 8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
-conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 & 8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
-conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 & 8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
-conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 & 8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
-Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 & 8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 & 8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
-Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 & 8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
-invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 & 8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
-robust invariant                      & & 3.0 & 3.3 & 3.6 &  5.2 & 8.3 &  5.9 &  19.9 & 13.6 &  32.9 \\
-\bottomrule
-\end{tabular}
-}
-\end{center}
-\end{table}
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-lambda-appendix}
-\small{
-\begin{tabular}{l | l r r r r r r r r r}
-
-\toprule
-& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
-transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
-
-\midrule
-
-conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
-conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
-conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
-Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
-invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
-robust invariant                      & & 0.4 & 1.3 & 1.5 &  0.0 & 0.8 & 1.7 &  0.0 & 1.1 & 2.0 \\
-\bottomrule
-\end{tabular}
-}
-\end{center}
-\end{table}
-```
-
-```{=latex}
-\begin{table}
-\begin{center}
-\caption{
-p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
-Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-p-value-appendix}
-\small{
-\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
-
-\toprule
-& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
-transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
-
-\midrule
-
-none                                  & & 0.986 & 0.614 & 0.323 & $<0.001$ & 0.063 & 0.016 & $<0.001$ & 0.997 & $<0.001$ \\
-conventional                          & & 0.986 & 0.464 & 0.488 &   0.675  & 0.139 & 0.620 &   0.902  & 0.999 &   0.698 \\
-conventional (z-score norm.)          & & 0.979 & 0.440 & 0.477 &   0.200  & 0.141 & 0.770 &   0.419  & 0.999 &   0.723 \\
-conventional (robust scaling)         & & 0.982 & 0.439 & 0.484 &   0.487  & 0.130 & 0.683 &   0.883  & 0.999 &   0.817 \\
-Raymaekers-Rousseeuw                  & & 0.989 & 0.464 & 0.488 &   0.675  & 0.446 & 0.620 &   0.876  & 0.992 &   0.821 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.992 & 0.440 & 0.477 &   0.200  & 0.448 & 0.770 &   0.460  & 0.991 &   0.719 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.993 & 0.439 & 0.484 &   0.487  & 0.422 & 0.683 &   0.887  & 0.991 &   0.848 \\
-invariant                             & & 0.993 & 0.355 & 0.461 &   0.720  & 0.153 & 0.490 &   0.910  & 0.999 &   0.682 \\
-robust invariant                      & & 0.994 & 0.407 & 0.484 &   0.642  & 0.291 & 0.489 &   0.848  & 0.996 &   0.628 \\
-\bottomrule
-\end{tabular}
-}
-\end{center}
-\end{table}
-```
+\FloatBarrier
+
+# Appendix E: Assessing transformations using simulated data
+
+## Ranking Box-Cox transformations
+
+Box-Cox transformations were assessed in the same manner as Yeo-Johnson transformations.
+The results are shown in Table \ref{tab:comparison_methods_simulations_box_cox}.
 
@@
 
+\FloatBarrier
+
+## Examples using clean data
+
+Robust transformations are hypothesised to have a cost in efficiency for data
+without outliers, i.e. clean data. Here we draw nine sequences with elements randomly drawn from a standard normal
+distribution $\mathcal{N}(0,1)$. Subsequently, we perform an inverse transformation
+$\left(\phi^{\lambda, 0, 1}\right)^{-1}$, with $\lambda \in \left(0, 2\right)$.
+
+The sequences drawn resulted from the permutations of the number of elements of each
+sequence ($n \in \{30, 100, 500\}$) and transformation parameter for the
+inverse transformation $\lambda \in \{0.1, 1.0, 1.9\}$. Each sequence then underwent
+power transformation using Box-Cox and Yeo-Johnson transformations. For Box-Cox
+transformations, each sequence was shifted prior to inverse transformation
+so that every element after inverse transformation would be strictly positive.
+
+Results for Box-Cox transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-bc},
+\ref{tab:clean-transformation-appendix-lambda-bc},
+and \ref{tab:clean-transformation-appendix-p-value-bc}.
+Additionally, results for Yeo-Johnson transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-yj},
+\ref{tab:clean-transformation-appendix-lambda-yj},
+and \ref{tab:clean-transformation-appendix-p-value-yj}.
+As may be observed, in these examples robust location and shift-invariant 
+transformations have higher residual errors because of low weights being
+assigned to tails of a distribution. This is also shown in 
+Tables \ref{tab:comparison_methods_simulations_yeo_johnson}
+and \ref{tab:comparison_methods_simulations_box_cox} for clean data without outliers.
+
+```{r include = FALSE, eval = FALSE}
+
+data <- .assess_standardisation_before_transformation_simulation(lambda_limit = c(-4.0, 6.0))
+data <- data[, mget(c("residuals", "residuals_central", "lambda", "p_value", "transformation", "dataset"))]
+data[, "residuals" := round(residuals, digits = 1L)]
+data[, "residuals_central" := round(residuals_central, digits = 1L)]
+data[, "lambda" := round(lambda, digits = 1L)]
+data[, "p_value" := round(p_value, digits = 3L)]
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
+data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
+data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
+data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
+```
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-residuals-yj}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 4.0 & 3.4 & 4.6 & 40.1 &  8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
+conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 &  8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
+conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 &  8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
+conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 &  8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
+Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 &  8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 &  8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 &  8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
+invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 &  8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
+robust invariant                      & & 3.0 & 3.4 & 3.6 &  5.5 & 10.1 &  6.2 &  18.7 & 13.2 &  32.8 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-lambda-yj}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
+conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
+conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
+Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
+invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
+robust invariant                      & & 0.5 & 1.1 & 1.5 & -0.1 & 0.7 & 1.7 &  0.1 & 1.1 & 1.9 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-p-value-yj}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.981 & 0.816 & 0.576 & $<0.001$ & 0.099 & 0.011 & $<0.001$ & 0.993 & $<0.001$ \\
+conventional                          & & 0.979 & 0.614 & 0.617 &   0.722  & 0.221 & 0.782 &   0.946  & 0.997 &   0.637 \\
+conventional (z-score norm.)          & & 0.970 & 0.594 & 0.645 &   0.251  & 0.224 & 0.895 &   0.529  & 0.998 &   0.828 \\
+conventional (robust scaling)         & & 0.974 & 0.593 & 0.640 &   0.538  & 0.207 & 0.837 &   0.933  & 0.998 &   0.810 \\
+Raymaekers-Rousseeuw                  & & 0.983 & 0.614 & 0.617 &   0.722  & 0.622 & 0.782 &   0.927  & 0.985 &   0.831 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.985 & 0.594 & 0.645 &   0.251  & 0.623 & 0.895 &   0.564  & 0.984 &   0.807 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.987 & 0.593 & 0.640 &   0.538  & 0.599 & 0.837 &   0.935  & 0.984 &   0.875 \\
+invariant                             & & 0.987 & 0.487 & 0.573 &   0.778  & 0.241 & 0.683 &   0.963  & 0.997 &   0.865 \\
+robust invariant                      & & 0.992 & 0.753 & 0.667 &   0.909  & 0.739 & 0.628 &   0.968  & 0.994 &   0.680 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
 ```{r include = FALSE, eval = FALSE}
@@
 
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-residuals-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 5.3 & 3.4 & 3.8 & 45.7 &  8.9 & 6.0 & 221.8 & 10.1 & 29.7 \\
+conventional                          & & 2.4 & 3.2 & 3.5 &  5.9 &  8.4 & 5.5 &  14.3 &  9.8 & 16.1 \\
+conventional (z-score norm.)          & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+conventional (robust scaling)         & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+Raymaekers-Rousseeuw                  & & 2.4 & 3.2 & 3.5 &  5.9 & 13.3 & 5.7 &  16.1 & 13.6 & 16.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  19.1 & 19.0 & 19.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  18.4 & 19.0 & 19.9 \\
+invariant                             & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  13.7 &  9.9 & 16.0 \\
+robust invariant                      & & 2.4 & 4.3 & 4.7 &  8.9 & 20.4 & 6.7 &  14.1 & 15.8 & 37.6 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-lambda-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.4 & 1.3 & 0.5 & 0.0 &  0.9 & 0.7 & 0.1 & 1.1 & 2.0 \\
+conventional (z-score norm.)          & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+conventional (robust scaling)         & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+Raymaekers-Rousseeuw                  & & 0.4 & 1.3 & 0.5 & 0.0 & -0.1 & 0.5 & 0.1 & 0.9 & 1.8 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+invariant                             & & 0.4 & 1.6 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.1 & 1.9 \\
+robust invariant                      & & 0.5 & 0.8 & 0.4 & 0.0 &  0.2 & 0.7 & 0.1 & 0.9 & 0.9 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-p-value-bc}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.940 & 0.816 & 0.498 & $<0.001$ & 0.098 & 0.714 & $<0.001$ & 0.993 & 0.928 \\
+conventional                          & & 0.948 & 0.695 & 0.603 &   0.798  & 0.163 & 0.791 &   0.931  & 0.997 & 0.658 \\
+conventional (z-score norm.)          & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+conventional (robust scaling)         & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+Raymaekers-Rousseeuw                  & & 0.948 & 0.695 & 0.603 &   0.798  & 0.768 & 0.806 &   0.816  & 0.962 & 0.782 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.775  & 0.866 & 0.921 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.807  & 0.866 & 0.921 \\
+invariant                             & & 0.947 & 0.661 & 0.624 &   0.611  & 0.182 & 0.789 &   0.956  & 0.997 & 0.664 \\
+robust invariant                      & & 0.969 & 0.824 & 0.750 &   0.972  & 0.779 & 0.775 &   0.956  & 0.934 & 0.840 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
 # Appendix F: Experimental results
@@
 
+\FloatBarrier
+
 ## Box-Cox transformation
@@
 
+\FloatBarrier
+
 # References

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.tex
Main changes to appendix are now complete.

--- 2025-03-27 manuscript_rmarkdown/manuscript_appendix.tex
@@
 \usepackage{multirow}
+\usepackage{placeins}
 \newtheorem{definition}{Definition}
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-03-27}
+\date{2025-04-09}
 
@@
 
-\section{Appendix C: Simulations with other optimisation
-criteria}\label{appendix-c-simulations-with-other-optimisation-criteria}
+\subsection{Simulations with other optimisation
+criteria}\label{simulations-with-other-optimisation-criteria}
 
@@
 \(\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)\).
+
 We then computed transformation parameter \(\lambda\) using the original
@@
 
-\section{Appendix D: Simulations on clean
-data}\label{appendix-d-simulations-on-clean-data}
-
-An alternative to location- and scale-invariant transformations is
-normalising feature distributions prior to conventional transformations.
-We assessed two normalisation methods were used: 1. z-standardisation,
-i.e.~\(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
-and \(\sigma\) the mean and standard deviation of feature
-\(\mathbf{X}\); and 2. robust scaling,
-i.e.~\(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
-with \(\text{IQR}\) representing the interquartile range of feature
-\(\mathbf{X}\).
-
-{[}Generated data: clean, n = 30, n = 100, n = 1000, lambda = 0.1, 1.0,
-1.9{]}
+\section{Appendix D: Central normality test and empirical central
+normality test critical
+values}\label{appendix-d-central-normality-test-and-empirical-central-normality-test-critical-values}
+
+Critical values for the central normality test and its empirical variant
+are found in Table \ref{tab:central-normality-critical-statistic} and
+Table \ref{tab:empirical-central-normality-critical-statistic},
+respectively.
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
+\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
+
+\midrule
+    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
+   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
+   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
+   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
+  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
+  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
+  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
+ 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
+ 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
+ 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
+
+\midrule
+    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
+   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
+   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
+   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
+  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
+  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
+  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
+ 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
+ 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
+ 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
+10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-critical-statistics-1}
+
+}
+
+\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.80}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
+The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
+with fully normal data where 10\% of elements are replaced by outliers.
+emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
+\end{figure}
+
+\FloatBarrier
+
+\section{Appendix E: Assessing transformations using simulated
+data}\label{appendix-e-assessing-transformations-using-simulated-data}
+
+\subsection{Ranking Box-Cox
+transformations}\label{ranking-box-cox-transformations}
+
+Box-Cox transformations were assessed in the same manner as Yeo-Johnson
+transformations. The results are shown in Table
+\ref{tab:comparison_methods_simulations_box_cox}.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+If necessary, each sequence was shifted so that every element had a strictly positive value. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_box_cox}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
+conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
+conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
+conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
+Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
+invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
+robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Examples using clean data}\label{examples-using-clean-data}
+
+Robust transformations are hypothesised to have a cost in efficiency for
+data without outliers, i.e.~clean data. Here we draw nine sequences with
+elements randomly drawn from a standard normal distribution
+\(\mathcal{N}(0,1)\). Subsequently, we perform an inverse transformation
+\(\left(\phi^{\lambda, 0, 1}\right)^{-1}\), with
+\(\lambda \in \left(0, 2\right)\).
+
+The sequences drawn resulted from the permutations of the number of
+elements of each sequence (\(n \in \{30, 100, 500\}\)) and
+transformation parameter for the inverse transformation
+\(\lambda \in \{0.1, 1.0, 1.9\}\). Each sequence then underwent power
+transformation using Box-Cox and Yeo-Johnson transformations. For
+Box-Cox transformations, each sequence was shifted prior to inverse
+transformation so that every element after inverse transformation would
+be strictly positive.
+
+Results for Box-Cox transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-bc},
+\ref{tab:clean-transformation-appendix-lambda-bc}, and
+\ref{tab:clean-transformation-appendix-p-value-bc}. Additionally,
+results for Yeo-Johnson transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-yj},
+\ref{tab:clean-transformation-appendix-lambda-yj}, and
+\ref{tab:clean-transformation-appendix-p-value-yj}. As may be observed,
+in these examples robust location and shift-invariant transformations
+have higher residual errors because of low weights being assigned to
+tails of a distribution. This is also shown in Tables
+\ref{tab:comparison_methods_simulations_yeo_johnson} and
+\ref{tab:comparison_methods_simulations_box_cox} for clean data without
+outliers.
 
@@
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-residuals-appendix}
+\label{tab:clean-transformation-appendix-residuals-yj}
 \small{
@@
 invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 &  8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
-robust invariant                      & & 3.0 & 3.3 & 3.6 &  5.2 & 8.3 &  5.9 &  19.9 & 13.6 &  32.9 \\
+robust invariant                      & & 3.0 & 3.4 & 3.6 &  5.5 & 10.1 &  6.2 &  18.7 & 13.2 &  32.8 \\
+
 \bottomrule
@@
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-lambda-appendix}
+\label{tab:clean-transformation-appendix-lambda-yj}
 \small{
@@
 invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
-robust invariant                      & & 0.4 & 1.3 & 1.5 &  0.0 & 0.8 & 1.7 &  0.0 & 1.1 & 2.0 \\
+robust invariant                      & & 0.5 & 1.1 & 1.5 & -0.1 & 0.7 & 1.7 &  0.1 & 1.1 & 1.9 \\
 \bottomrule
@@
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
-\label{tab:clean-transformation-p-value-appendix}
+\label{tab:clean-transformation-appendix-p-value-yj}
 \small{
@@
 
-none                                  & & 0.986 & 0.614 & 0.323 & $<0.001$ & 0.063 & 0.016 & $<0.001$ & 0.997 & $<0.001$ \\
-conventional                          & & 0.986 & 0.464 & 0.488 &   0.675  & 0.139 & 0.620 &   0.902  & 0.999 &   0.698 \\
-conventional (z-score norm.)          & & 0.979 & 0.440 & 0.477 &   0.200  & 0.141 & 0.770 &   0.419  & 0.999 &   0.723 \\
-conventional (robust scaling)         & & 0.982 & 0.439 & 0.484 &   0.487  & 0.130 & 0.683 &   0.883  & 0.999 &   0.817 \\
-Raymaekers-Rousseeuw                  & & 0.989 & 0.464 & 0.488 &   0.675  & 0.446 & 0.620 &   0.876  & 0.992 &   0.821 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.992 & 0.440 & 0.477 &   0.200  & 0.448 & 0.770 &   0.460  & 0.991 &   0.719 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.993 & 0.439 & 0.484 &   0.487  & 0.422 & 0.683 &   0.887  & 0.991 &   0.848 \\
-invariant                             & & 0.993 & 0.355 & 0.461 &   0.720  & 0.153 & 0.490 &   0.910  & 0.999 &   0.682 \\
-robust invariant                      & & 0.994 & 0.407 & 0.484 &   0.642  & 0.291 & 0.489 &   0.848  & 0.996 &   0.628 \\
+none                                  & & 0.981 & 0.816 & 0.576 & $<0.001$ & 0.099 & 0.011 & $<0.001$ & 0.993 & $<0.001$ \\
+conventional                          & & 0.979 & 0.614 & 0.617 &   0.722  & 0.221 & 0.782 &   0.946  & 0.997 &   0.637 \\
+conventional (z-score norm.)          & & 0.970 & 0.594 & 0.645 &   0.251  & 0.224 & 0.895 &   0.529  & 0.998 &   0.828 \\
+conventional (robust scaling)         & & 0.974 & 0.593 & 0.640 &   0.538  & 0.207 & 0.837 &   0.933  & 0.998 &   0.810 \\
+Raymaekers-Rousseeuw                  & & 0.983 & 0.614 & 0.617 &   0.722  & 0.622 & 0.782 &   0.927  & 0.985 &   0.831 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.985 & 0.594 & 0.645 &   0.251  & 0.623 & 0.895 &   0.564  & 0.984 &   0.807 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.987 & 0.593 & 0.640 &   0.538  & 0.599 & 0.837 &   0.935  & 0.984 &   0.875 \\
+invariant                             & & 0.987 & 0.487 & 0.573 &   0.778  & 0.241 & 0.683 &   0.963  & 0.997 &   0.865 \\
+robust invariant                      & & 0.992 & 0.753 & 0.667 &   0.909  & 0.739 & 0.628 &   0.968  & 0.994 &   0.680 \\
+
 \bottomrule
@@
 
-\section{Appendix D: Experimental results using location- and
-scale-invariant Box-Cox
-transformation}\label{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-residuals-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 5.3 & 3.4 & 3.8 & 45.7 &  8.9 & 6.0 & 221.8 & 10.1 & 29.7 \\
+conventional                          & & 2.4 & 3.2 & 3.5 &  5.9 &  8.4 & 5.5 &  14.3 &  9.8 & 16.1 \\
+conventional (z-score norm.)          & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+conventional (robust scaling)         & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+Raymaekers-Rousseeuw                  & & 2.4 & 3.2 & 3.5 &  5.9 & 13.3 & 5.7 &  16.1 & 13.6 & 16.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  19.1 & 19.0 & 19.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  18.4 & 19.0 & 19.9 \\
+invariant                             & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  13.7 &  9.9 & 16.0 \\
+robust invariant                      & & 2.4 & 4.3 & 4.7 &  8.9 & 20.4 & 6.7 &  14.1 & 15.8 & 37.6 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-lambda-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.4 & 1.3 & 0.5 & 0.0 &  0.9 & 0.7 & 0.1 & 1.1 & 2.0 \\
+conventional (z-score norm.)          & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+conventional (robust scaling)         & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+Raymaekers-Rousseeuw                  & & 0.4 & 1.3 & 0.5 & 0.0 & -0.1 & 0.5 & 0.1 & 0.9 & 1.8 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+invariant                             & & 0.4 & 1.6 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.1 & 1.9 \\
+robust invariant                      & & 0.5 & 0.8 & 0.4 & 0.0 &  0.2 & 0.7 & 0.1 & 0.9 & 0.9 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-p-value-bc}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.940 & 0.816 & 0.498 & $<0.001$ & 0.098 & 0.714 & $<0.001$ & 0.993 & 0.928 \\
+conventional                          & & 0.948 & 0.695 & 0.603 &   0.798  & 0.163 & 0.791 &   0.931  & 0.997 & 0.658 \\
+conventional (z-score norm.)          & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+conventional (robust scaling)         & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+Raymaekers-Rousseeuw                  & & 0.948 & 0.695 & 0.603 &   0.798  & 0.768 & 0.806 &   0.816  & 0.962 & 0.782 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.775  & 0.866 & 0.921 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.807  & 0.866 & 0.921 \\
+invariant                             & & 0.947 & 0.661 & 0.624 &   0.611  & 0.182 & 0.789 &   0.956  & 0.997 & 0.664 \\
+robust invariant                      & & 0.969 & 0.824 & 0.750 &   0.972  & 0.779 & 0.775 &   0.956  & 0.934 & 0.840 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\section{Appendix F: Experimental
+results}\label{appendix-f-experimental-results}
 
 The effect of using location- and scale-invariant transformations was
-investigated using real-world datasets.
-
-\subsection{Invariance}\label{invariance}
-
-Results for Box-Cox transformations of features without outliers are
-shown in Figure \ref{fig:experimental-results-invariance-appendix}.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/experimental-results-invariance-appendix-1}
-
-}
-
-\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row).
-Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).}\label{fig:experimental-results-invariance-appendix}
-\end{figure}
-
-\subsubsection{Age of patients with lung
-cancer}\label{age-of-patients-with-lung-cancer}
-
-Applying conventional and invariant Box-Cox transformations to age of
-patients with lung cancer (Loprinzi et al. 1994) yielded the following
-results: no transformation (sum of residuals with normal distribution
-\(\sum r_i = 16.5\), \(p = 0.71\)); conventional transformation
-(\(\lambda = 1.9\), \(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
-\(\sigma_{BC} = 0.4 \cdot 10^3\), \(p = 0.94\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
-\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\),
-\(p = 0.94\)); location- and scale-invariant transformation
-(\(\lambda = 1.7\), \(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\),
-\(\sigma_{BC} = 0.8\), \(p = 0.94\)); and robust location- and
-scale-invariant transformation (\(\lambda = 1.5\), \(\sum r_i = 11.6\),
-\(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\), \(p = 0.93\)).
-
-Compared to location- and scale-invariant Yeo-Johnson transformations,
-the Box-Cox transformations do not reduce residuals compared to
-conventional variants.
-
-\subsubsection{Penguin body mass}\label{penguin-body-mass}
-
-Applying conventional and invariant Box-Cox transformations to the body
-mass of penguins (Gorman, Williams, and Fraser 2014) yielded the
-following results: no transformation (residual sum \(\sum r_i = 48.0\),
-\(p < 0.001\)); conventional transformation (\(\lambda = -0.5\),
-\(\sum r_i = 32.2\), \(\mu_{BC} = 2.1\),
-\(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\), \(p = 0.06\));
-location- and scale-invariant transformation (\(\lambda = 0.5\),
-\(\sum r_i = 27.3\), \(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\),
-\(p = 0.10\)); and robust location- and scale-invariant transformation
-(\(\lambda = 0.2\), \(\sum r_i = 25.2\), \(\mu_{BC} = 0.4\),
-\(\sigma_{BC} = 0.6\), \(p = 0.44\)).
-
-Just as for location- and scale-invariant Yeo-Johnson transformations,
-Box-Cox transformations produced a lower overall residual sum compared
-to their conventional counterparts. Similarly, conventional
-transformations led to low standard deviation \(\sigma_{YJ}\) of the
-body mass feature after transformation.
-
-\subsubsection{Latitude in the Ames housing
-dataset}\label{latitude-in-the-ames-housing-dataset}
-
-Applying conventional and invariant Box-Cox transformations to the
-latitude of houses in the Ames housing dataset (De Cock 2011) yielded
-the following results: no transformation (residual sum
-\(\sum r_i = 328\), \(p < 0.001\)); conventional transformation
-(\(\lambda = 62.1\), \(\sum r_i = 319\),
-\(\mu_{BC} = 1.1 \cdot 10^{99}\), \(\sigma_{BC} = 0.0 \cdot 10^{99}\),
-\(p < 0.001\)); Raymaekers and Rousseeuw's robust adaptation
-(\(\lambda = 96.0\), \(\sum r_i = 319\),
-\(\mu_{BC} = 6.2 \cdot 10^{153}\), \(\sigma_{BC} = 0.3 \cdot 10^{153}\),
-\(p < 0.001\)); location- and scale-invariant transformation
-(\(\lambda = 1.9\), \(\sum r_i = 312\), \(\mu_{BC} = 2.3\),
-\(\sigma_{BC} = 0.9\), \(p < 0.001\)); and robust location- and
-scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\), \(p < 0.001\)).
-
-Similar to conventional Yeo-Johnson transformations (non-robust and
-robust), conventional Box-Cox transformations had high values for the
-\(\lambda\) parameter, which could lead to numerical issues. Location-
-and scale-invariant Box-Cox transformations did not suffer from this
-issue.
-
-\subsection{Robustness against
-outliers}\label{robustness-against-outliers}
-
-Results for Box-Cox transformations of features with outliers are shown
-in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/experimental-results-outlier-robustness-appendix-1}
-
-}
-
-\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row).
-Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).
-Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness-appendix}
-\end{figure}
-
-\subsubsection{Fuel efficiency in the Top Gear
-dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
-
-The Top Gear dataset contains data on 297 vehicles, with outliers
-related highly fuel-efficient vehicles (Alfons 2021). Applying
-conventional and invariant Box-Cox transformations to the fuel
-consumption feature yielded the following results: no transformation
-(residual sum \(\sum r_i = 54\), \(p=0.58\)); conventional
-transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
-\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p < 0.001\)); Raymaekers
-and Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
-\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.32\)); location- and
-scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
-\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p < 0.001\)); and robust
-location- and scale-invariant transformation (\(\lambda = 1.1\),
-\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
-\(p=0.67\)).
-
-\subsubsection{Maximum arterial wall thickness in an ischemic stroke
-dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
-
-The ischemic stroke dataset contains historic data from 126 patients
-with risk at ischemic stroke (Kuhn and Johnson 2019). Applying
-conventional and invariant Box-Cox transformations to the maximum
-arterial wall thickness feature yielded the following results: no
-transformation (residual sum \(\sum r_i = 110\), \(p=0.86\));
-conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
-\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.90\)); location- and
-scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
-\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.02\)); and robust
-location- and scale-invariant transformation (\(\lambda = 0.5\),
-\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
-\(p=0.60\)).
-
-\section{Appendix E: Empirical central normality
-test}\label{appendix-e-empirical-central-normality-test}
-
-Critical values for the central normality test and its empirical variant
-are found in Table \ref{tab:central-normality-critical-statistic} and
-Table \ref{tab:empirical-central-normality-critical-statistic},
-respectively.
-
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
-
-\midrule
-5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
-10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
-20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
-50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
-100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
-200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
-500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
-1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
-2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
-5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
-10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:empirical-central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\
-
-\midrule
-5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
-10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
-20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
-50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
-100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
-200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
-500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
-1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
-2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
-5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
-10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-critical-statistics-1}
-
-}
-
-\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.70}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
-The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
-with fully normal data where 10\% of elements are replaced by outliers.
-emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
-\end{figure}
-
-\section{Appendix F: Normalisation before
-transformation}\label{appendix-f-normalisation-before-transformation}
-
-An alternative to location- and scale-invariant transformations is
-normalising feature distributions prior to conventional transformations.
-We assessed two normalisation methods were used: 1. z-standardisation,
-i.e.~\(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
-and \(\sigma\) the mean and standard deviation of feature
-\(\mathbf{X}\); and 2. robust scaling,
-i.e.~\(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
-with \(\text{IQR}\) representing the interquartile range of feature
-\(\mathbf{X}\).
-
-Table \ref{tab:normalisation-before-transformation-appendix} shows
-residual errors after transformation to normality of the five features
-from real-world datasets presented previously in Appendix D and the main
-manuscript. Table
-\ref{tab:normalisation-before-transformation-lambda-appendix} shows the
-corresponding values for transformation parameter \(\lambda\),
-\ref{tab:normalisation-before-transformation-p-value-appendix} shows
-p-values of the empirical central normality test. Normalisation prior to
-transformation prevents unexpectedly high \(\lambda\) values in the
-presence of large location shifts, as can be observed for the latitude
-feature. Normalisation prior to transformation, however, does not
-consistently improve residual errors after transformation.
-Interestingly, use of normalisation prior to transformation with robust
-power transformation using the method by Raymaekers and Rousseeuw
-(Raymaekers and Rousseeuw 2024) led to higher residual errors for
-arterial wall thickness, fuel efficiency and latitude features. This is
-similar to observations by Raymaekers and Rousseeuw (2024), who observed
-that their method can emphasize values of outliers after transformation.
-Location- and scale-invariant transformations have similar or lower
-residual errors compared to errors resulting from normalisation prior to
-transformation. Not surprisingly, normalisation before transformation
-does not noticeably improve central normality, as assessed using the
-empirical central normality test.
+investigated using real-world datasets, as described in the main
+manuscript.
+
+\subsection{Yeo-Johnson
+transformation}\label{yeo-johnson-transformation}
+
+Additional results for Yeo-Johnson transformations are shown in Tables
+\ref{tab:experimental-results-appendix-residuals},
+\ref{tab:experimental-results-appendix-lambda},
+\ref{tab:experimental-results-appendix-p-value}.
 
@@
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-appendix}
+\label{tab:experimental-results-appendix-residuals}
 \begin{tabular}{l | r r r r r}
@@
 invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\
-robust invariant & 10.6 & 27.1 & 49.5 & 310.6 & 23.1 \\
+robust invariant                      &  9.3 &  30.2 & 55.8 & 308.1 & 22.0 \\
+
 \bottomrule
@@
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-lambda-appendix}
+\label{tab:experimental-results-appendix-lambda}
 \begin{tabular}{l | r r r r r}
@@
 invariant                             & 1.3 &  0.2 & -1.3 &  1.5 &  0.5 \\
-robust invariant & 1.4 & -0.6 & 0.9 & 1.4 & 0.4 \\
+robust invariant                      & 1.3 & -0.3 &  1.0 &  1.1 &  0.3 \\
 \bottomrule
@@
 AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
-\label{tab:normalisation-before-transformation-p-value-appendix}
+\label{tab:experimental-results-appendix-p-value}
 \begin{tabular}{l | r r r r r}
@@
 \midrule
-none                                  & 0.712 & 0.864 & 0.577     & $< 0.001$ & $< 0.001$ \\
-conventional                          & 0.936 & 0.017 & $< 0.001$ & $< 0.001$ & 0.058 \\
-conventional (z-score norm.)          & 0.860 & 0.058 & $< 0.001$ & $< 0.001$ & 0.054 \\
-conventional (robust scaling)         & 0.881 & 0.001 & $< 0.001$ & $< 0.001$ & 0.076 \\
-Raymaekers-Rousseeuw                  & 0.936 & 0.905 & 0.320     & $< 0.001$ & 0.058 \\
-Raymaekers-Rousseeuw (z-score norm.)  & 0.605 & 0.922 & 0.635     & $< 0.001$ & 0.054 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.830 & 0.948 & 0.637     & $< 0.001$ & 0.085 \\
-invariant                             & 0.948 & 0.231 & $< 0.001$ & $< 0.001$ & 0.184 \\
-robust invariant                      & 0.797 & 0.182 & 0.352     & $< 0.001$ & 0.397 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
+
+none                                  & 0.694 & 0.832     & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.003     & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.919 & 0.021     & $< 0.001$ & $< 0.001$ & 0.099 \\
+conventional (robust scaling)         & 0.933 & $< 0.001$ & $< 0.001$ & $< 0.001$ & 0.129 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.881     & 0.373     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.694 & 0.899     & 0.809     & $< 0.001$ & 0.099 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.900 & 0.880     & 0.811     & $< 0.001$ & 0.144 \\
+invariant                             & 0.976 & 0.147     & $< 0.001$ & $< 0.001$ & 0.276 \\
+robust invariant                      & 0.933 & 0.178     & 0.759     & $< 0.001$ & 0.688 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Box-Cox transformation}\label{box-cox-transformation}
+
+Results for Box-Cox transformations are shown in Tables
+\ref{tab:experimental-results-appendix-residuals-bc},
+\ref{tab:experimental-results-appendix-lambda-bc} and
+\ref{tab:experimental-results-appendix-p-value-bc}.
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-residuals-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none                                  & 16.5 & 110.1 &  54.5 & 328.4 & 48.0 \\
+conventional                          & 11.5 &  33.5 &  55.7 & 318.9 & 32.2 \\
+conventional (z-score norm.)          & 12.9 &  44.5 &  58.8 & 305.3 & 27.3 \\
+conventional (robust scaling)         & 12.9 &  44.7 &  59.0 & 305.3 & 27.3 \\
+Raymaekers-Rousseeuw                  & 11.5 & 127.1 &  47.6 & 314.8 & 32.2 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & 13.1 &  54.8 & 127.5 & 510.2 & 23.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 18.9 & 100.2 & 113.7 & 510.2 & 23.6 \\
+invariant                             & 11.6 &  28.0 &  48.4 & 311.8 & 27.3 \\
+robust invariant                      & 12.8 & 150.1 &  60.5 & 646.0 & 27.4 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-lambda-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+conventional                          & 1.9 & -0.5 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.)          & 1.2 &  0.1 &  0.2 &  1.2 &  0.5 \\
+conventional (robust scaling)         & 1.2 &  0.1 &  0.1 &  1.2 &  0.5 \\
+Raymaekers-Rousseeuw                  & 1.9 &  1.1 &  0.8 & 95.9 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 1.2 &  0.6 & -0.5 &  0.6 &  0.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.9 &  1.0 & -0.5 &  0.6 &  0.3 \\
+invariant                             & 1.7 & -1.0 & -0.7 &  1.9 &  0.5 \\
+robust invariant                      & 1.3 &  1.2 &  1.1 &  0.4 &  0.1 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-p-value-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+none                                  & 0.694 & 0.832 & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.002 & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+conventional (robust scaling)         & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.871 & 0.365     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.891 & 0.268 & $< 0.001$ & $< 0.001$ & 0.495 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.562 & 0.789 & $< 0.001$ & $< 0.001$ & 0.492 \\
+invariant                             & 0.962 & 0.004 & $< 0.001$ & $< 0.001$ & 0.154 \\
+robust invariant                      & 0.899 & 0.890 & 0.866     & $< 0.001$ & 0.622 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
 
@@
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-Alfons2021-kc}
-Alfons, Andreas. 2021. {``{robustHD}: An {R} Package for Robust
-Regression with High-Dimensional Data.''} \emph{J. Open Source Softw.} 6
-(67): 3786. \url{https://doi.org/10.21105/joss.03786}.
-
 \bibitem[\citeproctext]{ref-Anderson1952-gz}
@@
 
-\bibitem[\citeproctext]{ref-De-Cock2011-jf}
-De Cock, Dean. 2011. {``Ames, Iowa: Alternative to the Boston Housing
-Data as an End of Semester Regression Project.''} \emph{J. Stat. Educ.}
-19 (3). \url{https://doi.org/10.1080/10691898.2011.11889627}.
-
-\bibitem[\citeproctext]{ref-Gorman2014-eo}
-Gorman, Kristen B, Tony D Williams, and William R Fraser. 2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a
-Community of Antarctic Penguins (Genus Pygoscelis).''} \emph{PLoS One} 9
-(3): e90081. \url{https://doi.org/10.1371/journal.pone.0090081}.
-
 \bibitem[\citeproctext]{ref-Huber1981-su}
@@
 
-\bibitem[\citeproctext]{ref-Kuhn2019-kt}
-Kuhn, Max, and Kjell Johnson. 2019. \emph{Feature Engineering and
-Selection: A Practical Approach for Predictive Models}. Chapman \&
-Hall/CRC Data Science Series. Chapman; Hall/CRC.
-\url{https://doi.org/10.1201/9781315108230}.
-
-\bibitem[\citeproctext]{ref-Loprinzi1994-cd}
-Loprinzi, C L, J A Laurie, H S Wieand, J E Krook, P J Novotny, J W
-Kugler, J Bartel, M Law, M Bateman, and N E Klatt. 1994. {``Prospective
-Evaluation of Prognostic Variables from Patient-Completed
-Questionnaires. North Central Cancer Treatment Group.''} \emph{J. Clin.
-Oncol.} 12 (3): 601--7. \url{https://doi.org/10.1200/JCO.1994.12.3.601}.
-
 \bibitem[\citeproctext]{ref-Von_Mises1928-ef}

### 2025-04-09 manuscript_rmarkdown/simulations.R

@@ -1296,6 +1296,8 @@
         sd = 1.0
       )
 
+      if (method == "box_cox" && any(x <= -1.0 / lambda[jj])) x <- x + 1 - min(x) - 1.0 / lambda[jj]
+
       # Perform inverse transform.
       transformer <- power.transform::create_transformer_skeleton(
         method = method,

### 2025-04-09 manuscript_rmarkdown/manuscript.Rmd
Updated machine learning results.

@@
 Before each experiment, each dataset was randomly split into a training (70%)
-and holdout test (30%) set five times. Thus, a total of 27720 models were
+and holdout test (30%) set five times. Thus, a total of 36960 models were
 created. Each model was then evaluated using the holdout test set using one of
@@
 
-For the purpose of assessing the effect of the difficulty of the task, we computed the median performance score over all models for each dataset and assigned one the following categories:
+For the purpose of assessing the effect of the difficulty of the task, 
+we computed the median performance score over all models for each dataset and assigned one the following categories:
 
@@
 The estimated marginal effects are shown in Figure \ref{fig:marginal-effect-plot}.
-On the scale of normalised ranks ($[0.0, 1.0]$), the overall estimated marginal effect of using a random forest instead of generalised linear model was $0.272$.
-The overall marginal effect of using z-standardisation to normalise features was $0.008$.
+On the scale of normalised ranks ($[0.0, 1.0]$), only the random forest performed better than the average of all learners (rank difference $0.173$).
+The overall marginal effect of using z-standardisation to normalise features was $0.011$.
 Transformation methods had the following marginal effects:
-$0.003$ for using conventional Yeo-Johnson transformation instead of no transformation;
-$-0.007$ for using robust invariant Yeo-Johnson transformation instead of no transformation; 
-$-0.009$ for using robust invariant Yeo-Johnson transformation with empirical central normality test instead of no transformation; 
-$-0.010$ for using robust invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson transformation;
+$-0.001$ for using conventional Yeo-Johnson transformation instead of no transformation;
+$-0.012$ for using robust invariant Yeo-Johnson transformation instead of no transformation; 
+$-0.012$ for using robust invariant Yeo-Johnson transformation with empirical central normality test instead of no transformation; 
+$-0.012$ for using robust invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson transformation;
 and $-0.012$ for using using robust invariant Yeo-Johnson transformation with empirical central normality test instead of conventional Yeo-Johnson transformation.
@@
 cap <- paste0(
-  "Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets. ",
-  "The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). ",
-  "Random forests outperform GLM models for all task difficulties. ",
+  "Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 36960 machine learning experiments on 231 datasets. ",
+  "The four top-left panels shows the marginal effect of learners vs. the average, i.e. generalised linear models (GLM), L1-regularised LM (Lasso), gradient-boosted LM and random forest. ",
+  "Random forests outperform the average. ",
   "The top-right panel shows the marginal effect of feature normalisation methods, i.e. no normalisation and z-standardisation. ",
-  "z-standardisation is generally beneficial, but the estimated effect is neglible. ",
-  "The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method. ",
-  "There is no consistent behaviour, and estimated effects are neglible. ",
+  "z-standardisation is generally beneficial, but the estimated effect is marginal. ",
+  "The bottom panel shows the marginal effects of different transformation methods, split by normalisation method. ",
+  "The estimated effects are similar in size to the effect of normalisation and marginal. ",
   "Note that the ranges of the $x$-axes of the three main panels differ. ",
@@
 The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
-However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit -- nor detriment -- to model performance for location- and scale-invariant power transformations.
+However, in a machine learning experiment of 231 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit -- nor detriment -- to model performance for location- and scale-invariant power transformations.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
-Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
+Of the 4886 numeric features in the 231 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
 For the latter two datasets, the transformation method did not show significant difference between groups (Friedman test; $p > 0.05$).
@@
 
-Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [GitHub](https://github.com/oncoray/power.transform) and the [CRAN repository](https://cran.r-project.org/package=power.transform).
+Location- and scale-invariant power transformations were implemented in the `power.transform` 
+package for R, which is available from [GitHub](https://github.com/oncoray/power.transform) 
+and the [CRAN repository](https://cran.r-project.org/package=power.transform). 
 The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository. 

### 2025-04-09 manuscript_rmarkdown/plots.R

@@ -1996,46 +1996,39 @@ get_annotation_settings <- function(ggtheme = NULL) {
     ordered=TRUE
   )
 
-  # difficulty_levels <- c("all", "very_easy", "easy", "intermediate", "difficult", "very_difficult", "unsolvable")
-  # difficulty_labels <- c("overall", "very easy", "easy", "intermediate", "difficult", "very difficult", "unsolvable")
-  difficulty_levels <- c("all", "very_easy", "easy", "intermediate", "difficult", "very_difficult")
+  difficulty_levels <- c("overall", "very_easy", "easy", "intermediate", "difficult", "very_difficult")
   difficulty_labels <- c("overall", "very easy", "easy", "intermediate", "difficult", "very difficult")
+
+  learner_levels <- c("glm", "lasso", "xgboost_lm", "random_forest_ranger")
+  learner_labels <- c("GLM", "Lasso", "gradient-boosted LM", "random forest")
+
+  abstract_data[, "value" := predict(model, data = abstract_data)$predictions]
 
   # Effect of learner ----------------------------------------------------------
   result_list <- list()
   learners <- levels(data$learner)
 
   for (learner_a in learners) {
-    for (learner_b in learners) {
-      if (learner_a == learner_b) next
+    result_list <- c(
+      result_list,
+      list(data.table::data.table(
+        "value" = mean(abstract_data[learner == learner_a]$value) - mean(abstract_data$value),
+        "task_difficulty" = "overall",
+        "learner_1" = learner_a
+      ))
+    )
 
-      x <- predict(model, data = abstract_data[learner == learner_a])$predictions -
-        predict(model, data = abstract_data[learner == learner_b])$predictions
-
+    for (difficulty in levels(abstract_data$task_difficulty)) {
       result_list <- c(
         result_list,
         list(data.table::data.table(
-          "value" = mean(x),
-          "task_difficulty" = "all",
-          "learner_1" = learner_a,
-          "learner_2" = learner_b
+          "value" =
+            mean(abstract_data[learner == learner_a & task_difficulty == difficulty]$value) -
+            mean(abstract_data[task_difficulty == difficulty]$value),
+          "task_difficulty" = difficulty,
+          "learner_1" = learner_a
         ))
       )
-
-      for (difficulty in levels(abstract_data$task_difficulty)) {
-        x <- predict(model, data = abstract_data[learner == learner_a & task_difficulty == difficulty])$predictions -
[…]

### 2025-04-09 manuscript_rmarkdown/manuscript.Rmd
Update discussion.

@@
 
-In addition to no power transformation and location and scale-invariant power transformations,
+In addition to no power transformation and location- and scale-invariant power transformations,
 conventional and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2024-zf] were assessed.
@@
 
-In their work on power transformation, Box and Cox already mention transformation with a shift parameter, but preferred the version in Eq.
-\ref{eqn:box-cox-original} for the theoretical analysis in their paper [@Box1964-mz], which subsequently became the convention.
+In their work on power transformation, Box and Cox already mention transformation with a shift parameter,
+but preferred the version in Eq. \ref{eqn:box-cox-original} for the theoretical analysis in their paper [@Box1964-mz], which subsequently became the convention.
 Yeo and Johnson's power transformation lacks a shift parameter altogether [@Yeo2000-vw].
@@
 Robust location- and scale-invariant transformations are a suitable replacement for their conventional counterparts.
-They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the distribution of a feature.
+They demonstrated robustness against outliers and prevent inaccurate
+transformations and potential numerical issues due to location and scale of the distribution of a feature.
 This is particularly relevant for automated data processing, where such issues may go unnoticed.
-Compared to non-robust location- and scale-invariant transformations, real-world examples with outliers showed higher residual errors of transformed features. 
-Robust transformations seek to minimise residual errors for the central part of the distribution, instead of the entire distribution, including outliers.
-The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
-However, in a machine learning experiment of 231 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit -- nor detriment -- to model performance for location- and scale-invariant power transformations.
+
+In simulation, robust location- and scale-invariant transformations ranked best in datasets with outliers,
+and ranked highly in datasets without outliers, except for data without outliers that were 
+directly derived from strictly normal distributions.
+This could be observed in real-world examples, where robust location-and scale-invariant transformations
+achieved central normality for when the non-robust variant could not.
+
+However, in a machine learning experiment of 231 real-world datasets that contained at least one numeric feature, 
+we did not find a meaningful benefit -- nor detriment -- to model performance for location- and scale-invariant power transformations.
 One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
@@
 
-Location- and scale-invariant transformations are realised by simultaneously optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
+Location- and scale-invariant transformations are realised by simultaneously 
+optimising three parameters, i.e. transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$.
 We derived the log-likelihood function to facilitate optimisation using MLE.
-Alternatively, standardisation of a numeric variable (e.g., through subtracting its median value and division by its interquartile range) prior to conventional power transformations may achieve a similar effect in reducing sensitivity to the distribution's location and scale.
-While this alternative helps prevent these issues -- provided that normalisation does not lead to negative values for Box-Cox transformation -- location- and scale-invariant transformations seem to provide an overall better transformation to normality (Appendix F).
-
-We assessed several methods for robust power transformation.
-Methods that relied on the z-score of the transformed feature or the residual error yielded worse results than the non-robust method.
-This is partly due to the initial choice of threshold parameters.
-Using different initial values, closer to the upper limits ($\delta_1 = 10.0; \delta_2 = 10.0$), led to a reduced loss.
-However, these values effectively correspond to a non-robust transformation, where all instances receive the same weight.
-Underperformance of these methods could be explained by their use of transformed features for setting weights.
-Consequently, the weights change at each iteration in the MLE optimisation process.
+Alternatively, standardisation of a numeric variable 
+(e.g., through subtracting its median value and division by its interquartile range) 
+prior to conventional power transformations can achieve a similar effect in reducing sensitivity
+to the distribution's location and scale, as demonstrated by our findings.
+While this alternative helps prevent these issues 
+-- provided that normalisation does not lead to negative values for Box-Cox transformations -- 
+location- and scale-invariant transformations are able to achieve better transformations
+to normality, demonstrated by a lower residual error.
+
+We assessed several weighting methods to achieve robust power transformations.
+Methods that relied on the z-score of the transformed feature or the residual error 
+yielded worse results than the non-robust method.
+Underperformance of these weighting methods could be explained by their reliance on transformed feature values for setting weights.
+Consequently, their weights change at each iteration in the MLE optimisation process.
 This increases local variance in the log-likelihood function and creates local optima that the optimiser may not handle well.
-As a consequence, optimal values for transformation parameter $\lambda$ might differ, which increases the presented loss for optimising threshold parameters.
 Methods that relied on the empirical probability did not suffer from this issue, as weights remained fixed during MLE.
 
-We introduced an empirical test for central normality to assess whether distributions deviate from normality in a way that might require closer inspection prior to further processing.
-The empirical test for central normality differs from other tests for normality, such as the Shapiro-Wilk test [@Shapiro1965-zd], in that the test statistic is independent of the number of samples.
-This makes this test more practical for assessing central normality for larger sample numbers, where other tests may detect inconsequential deviations from normality.
+We introduced an empirical test for central normality to assess whether sequences
+deviate from normality in a way that might require closer inspection prior to further processing.
+The empirical test for central normality differs from other tests for normality, 
+such as the Shapiro-Wilk test [@Shapiro1965-zd], as it assesses normality of the central portion
+of a feature, instead of the entire feature.
 
@@
 For MLE stability issues were not observed.
-Secondly, the empirical central normality test is based on simulations instead of statistical theory, and relies on a somewhat arbitrary definition of the central portion of a distribution.
-Thus, while the test may asses whether data is sufficiently normally distributed for practical purposes, it should not be used as a strict test for normality.
 
@@
 
-Compared to their conventional versions, robust location- and scale-invariant Box-Cox and Yeo-Johnson transformations reduce sensitivity to outliers and the location and scale of features.
+Compared to their conventional versions, robust location- and scale-invariant 
+Box-Cox and Yeo-Johnson transformations reduce sensitivity to outliers and the location and scale of features.
 An empirical central normality test can assess the quality of transformation of features to normal distributions.

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.Rmd
Harmonise styling.

@@
 
-# Appendix A: Log-likelihood functions for location and scale invariant power transformation
-
-Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parametrised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
-This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:
+# Appendix A: Log-likelihood functions for location- and scale-invariant power transformation
+
+Location- and scale-invariant Box-Cox and Yeo-Johnson transformations are parametrised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
+This leads to the following transformations. The location- and scale-invariant Box-Cox transformation is:
 
@@
 where $x_i - x_0 > 0$.
-The location and scale-invariant Yeo-Johnson transformation is:
+The location- and scale-invariant Yeo-Johnson transformation is:
 
@@
 and \ref{tab:clean-transformation-appendix-p-value-yj}.
-As may be observed, in these examples robust location and shift-invariant 
+As may be observed, in these examples robust location- and shift-invariant 
 transformations have higher residual errors because of low weights being

### 2025-04-09 manuscript_rmarkdown/manuscript.Rmd
Changed text to definitions.

@@
 - \usepackage{booktabs}
+- \newtheorem*{definition}{Definition}
 - \DeclareMathOperator*{\argmax}{argmax}
@@
 
-Many statistical and some machine learning methods assume normality of the underlying data, e.g. analysis of variance and linear discriminant analysis.
+Many statistical and some machine learning methods assume normality of the underlying data,
+e.g. analysis of variance and linear discriminant analysis.
 However, numerical features in datasets may strongly deviate from normal distributions, e.g. by being skewed.
@@
 The two most commonly used transformations are that of @Box1964-mz and @Yeo2000-vw.
-The Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}$ under the transformation parameter $\lambda$ is defined as:
-
-```{=latex}
+
+The Box-Cox transformation is defined as:
+
+\begin{definition}[Box-Cox power transformation]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a finite sequence of length $n > 0$.
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Then the Box-Cox power transformation of element $x_i$ is defined as:
 \begin{equation}
@@
 \end{equation}
-```
+\end{definition}
+
 One limitation of the Box-Cox transformation is that it is only defined for $x_i > 0$. 
-In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$:
-
-```{=latex}
+In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$, as follows:
+
+\begin{definition}[Yeo-Johnson power transformation]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a finite sequence of length $n > 0$.
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
 \begin{equation}
@@
 \end{equation}
-```
+\end{definition}
+
 The $\lambda$-parameter is typically optimised using maximum likelihood estimation under the assumption that the transformed feature is normally distributed.
-As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2024-zf].
+As noted by Raymaekers and Rousseeuw, this approach is sensitive to outliers, 
+and robust versions of Box-Cox and Yeo-Johnson transformations were devised [@Raymaekers2024-zf].
 
@@
 Statistical tests for normality, such as the Shapiro-Wilk test [@Shapiro1965-zd], could be automatically applied to transformed features.
-However, given sufficiently large sample sizes such tests can detect trivial deviations from normality, and may lead to rejection of sufficiently good power transformations.
+However, given sufficiently large sample sizes such tests can detect trivial
+deviations from normality, and may lead to rejection of sufficiently good power transformations.
 
@@
 
-Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
-The location- and scale-invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:
-
-```{=latex}
+Box-Cox and Yeo-Johnson transformations are modified by introducing
+shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.
+The location- and scale-invariant Box-Cox transformation of a feature value 
+$x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ 
+is then defined as:
+
+\begin{definition}[Location- and scale- invariant Box-Cox power transformation]
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i - x_0 > 0 \right\}$ be a finite sequence of length $n > 0$.
+
+Then the location- and scale-invariant Box-Cox power transformation of element $x_i$ is defined as:
+
 \begin{equation}
@@
 \end{equation}
-```
-where $x_i - x_0 > 0$.
-Likewise, the location- and scale-invariant Yeo-Johnson transformation of a feature value $x_i$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is:
-
-```{=latex}
+\end{definition}
+
+Likewise, the location- and scale-invariant Yeo-Johnson transformation of a feature value $x_i$ 
+under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is defined as:
+
+\begin{definition}[Location- and scale- invariant Yeo-Johnson power transformation]
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathcal{R} \right\}$ be a finite sequence of length $n > 0$.
+
+Then the location- and scale-invariant Yeo-Johnson power transformation of element $x_i$ is defined as:
+
 \begin{equation}
@@
 \end{equation}
-```
-
-For both invariant transformations, $\lambda$, $x_0$ and $s$ parameters can be obtained by maximising the log-likelihood function, i.e. using maximum likelihood estimation (MLE).
+\end{definition}
+
+For both invariant transformations, $\lambda$, $x_0$ and $s$ parameters
+can be obtained by maximising the log-likelihood function, i.e. using maximum likelihood estimation (MLE).
 A full derivation of the log-likelihood function for both transformations is shown in Appendix A.
@@
 
-```{=latex}
+\begin{definition} [Log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Box-Cox power transformation is defined as:
 \begin{equation}
@@
 \end{equation}
-```
-subject to $x_i - x_0 > 0$.
-$\mu$ and $\sigma^2$ are the mean and variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+\end{definition}
+
 Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood function is:
 
-```{=latex}
+\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is defined as:
 \begin{equation}
@@
 \end{equation}
-```
-where $\mu$ and $\sigma^2$ are the mean and variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+\end{definition}
 
@@
 
-\begin{definion}[Asymmetric generalised normal distribution probability density function]
+\begin{definition}[Asymmetric generalised normal distribution probability density function]
 Let $\alpha \in (0,1)$ be a skewness parameter.
@@
 
-\begin{definition}[central portion of a sequence]
+\begin{definition}[Central portion of a sequence]
 Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
@@
 To assess whether the proposed power transformations lead to values of $\lambda$ 
-that are invariant to location and scale of the distribution, we simulated three different sequences
+that are invariant to location and scale of the distribution, we simulated three different sequences.
 We first randomly drew $10000$ values from a normal distribution: 

### 2025-04-09 manuscript_rmarkdown/manuscript.tex
Changed text to definitions.

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{booktabs}
+\newtheorem*{definition}{Definition}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2024-10-17}
+\date{2025-04-09}
 
@@
 
-To-do
+Power transformations are used to stabilize variance and achieve
+normality of features, especially in methods assuming normal
+distributions such as ANOVA and linear discriminant analysis. However,
+the commonly used Box-Cox and Yeo-Johnson power transformation methods
+are sensitive to the location, scale, and presence of outliers in the
+data. Here we present location- and scale-invariant Box-Cox and
+Yeo-Johnson transformations to mitigate these issues. We derive maximum
+likelihood estimation criteria for optimizing transformation parameters
+and propose robust adaptations that reduce the influence of outliers. We
+also introduce an empirical test for assessing central normality of
+transformed features. In simulations and real-world datasets, robust
+location- and scale-invariant transformations outperform conventional
+variants, resulting in better transformations to central normality. In a
+machine learning experiment with 231 datasets with numerical features,
+integrating robust location- and scale-invariant power transformations
+into an automated data processing and machine learning pipeline did not
+result in a meaningful improvement or detriment to model performance
+compared to conventional variants. In conclusion, robust location- and
+scale-invariant power transformations can replace conventional variants.
 
@@
 transformations are that of Box and Cox (1964) and Yeo and Johnson
-(2000). The Box-Cox transformation of a feature value \(x_i\) of feature
-\(\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}\) under the
-transformation parameter \(\lambda\) is defined as:
-
+(2000).
+
+The Box-Cox transformation is defined as:
+
+\begin{definition}[Box-Cox power transformation]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a finite sequence of length $n > 0$.
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Then the Box-Cox power transformation of element $x_i$ is defined as:
 \begin{equation}
@@
 \end{equation}
+\end{definition}
 
@@
 transformation parameter \(\lambda\) is defined for any
-\(x_i \in \mathbb{R}\):
-
+\(x_i \in \mathbb{R}\), as follows:
+
+\begin{definition}[Yeo-Johnson power transformation]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a finite sequence of length $n > 0$.
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
 \begin{equation}
@@
 \(x_i\) of feature \(\mathbf{X}\) under transformation parameter
-\(\lambda\), shift parameter \(x_0\) and scale parameter \(s\) is then:
+\(\lambda\), shift parameter \(x_0\) and scale parameter \(s\) is then
+defined as:
+
+\begin{definition}[Location- and scale- invariant Box-Cox power transformation]
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i - x_0 > 0 \right\}$ be a finite sequence of length $n > 0$.
+
+Then the location- and scale-invariant Box-Cox power transformation of element $x_i$ is defined as:
 
@@
 \end{equation}
-
-where \(x_i - x_0 > 0\). Likewise, the location- and scale-invariant
-Yeo-Johnson transformation of a feature value \(x_i\) under
-transformation parameter \(\lambda\), shift parameter \(x_0\) and scale
-parameter \(s\) is:
+\end{definition}
+
+Likewise, the location- and scale-invariant Yeo-Johnson transformation
+of a feature value \(x_i\) under transformation parameter \(\lambda\),
+shift parameter \(x_0\) and scale parameter \(s\) is defined as:
+
+\begin{definition}[Location- and scale- invariant Yeo-Johnson power transformation]
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathcal{R} \right\}$ be a finite sequence of length $n > 0$.
+
+Then the location- and scale-invariant Yeo-Johnson power transformation of element $x_i$ is defined as:
 
@@
 
+\begin{definition} [Log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Box-Cox power transformation is defined as:
 \begin{equation}
@@
 \end{equation}
-
-subject to \(x_i - x_0 > 0\). \(\mu\) and \(\sigma^2\) are the mean and
-variance of the Box-Cox transformed feature
-\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
+\end{definition}
+
 Similarly, the location- and scale-invariant Yeo-Johnson log-likelihood
@@
 
+\begin{definition}
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is defined as:
 \begin{equation}
@@
 \mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
-& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
 \end{split}
 \end{equation}
-
-where \(\mu\) and \(\sigma^2\) are the mean and variance of the
-Yeo-Johnson transformed feature
-\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
+\end{definition}
 
@@
 
+\begin{definition}[Standard generalised normal distribution probability density function]
+Let $x \in \mathcal{R}$.
+Let $\beta > 0$ be a shape parameter.
+Let $\Gamma$ be the gamma function.
+
+Then the standard generalised normal distribution probability density function, without scale and location parameters, is:
 \begin{equation}
@@
 \end{equation}
-
-Here, \(\Gamma\) is the gamma function, and \(\beta\) is a strictly
-positive shape parameter. For \(\beta = 1\), the probability density
-function describes a Laplace distribution. A normal distribution is
-found for \(\beta=2\), and for large \(\beta\), the distribution
-approaches a uniform distribution. We will refrain from introducing
-scale and location parameters here directly.
+\end{definition}
+
+For \(\beta = 1\), the probability density function describes a Laplace
+distribution. A normal distribution is found for \(\beta=2\), and for
+large \(\beta\), the distribution approaches a uniform distribution.
 
@@
 
+\begin{definition}[Asymmetric generalised normal distribution probability density function]
+Let $\alpha \in (0,1)$ be a skewness parameter.
+Let $\mu \in \mathcal{R}$ be a location parameter.
+Let $\sigma > 0$ be a shape parameter.
+
+Then the probability density function of the asymmetric generalised normal distribution is:
 \begin{equation}
@@
 \end{equation}
-
-Here, \(\alpha \in (0,1)\) is a skewness parameter. \(\alpha > 0.5\)
-creates a distribution with a negative skew, i.e.~a left-skewed
-distribution. A right-skewed distribution is created for
-\(\alpha < 0.5\). \(\mu \in \mathcal{R}\) and \(\sigma \in (0, \infty)\)
-are location and scale parameters, respectively. \(f_{\alpha}\) thus
-describes the probability density function of an asymmetric generalised
-normal distribution, which we will refer to here and parametrise as
+\end{definition}
+
+\(\alpha > 0.5\) creates a distribution with a negative skew, i.e.~a
+left-skewed distribution. A right-skewed distribution is created for
+\(\alpha < 0.5\). \(f_{\alpha}\) thus describes the probability density
+function of an asymmetric generalised normal distribution, which we will
+refer to here and parametrise as
 \(\mathcal{AGN}\left(\mu, \sigma, \alpha, \beta \right)\).
@@
 random values from an asymmetric generalised normal distribution using
-inverse transform sampling. Gijbels et al.~derived the following
-quantile function \(F_{\alpha}^{-1}(p)\):
-
+inverse transform sampling. Gijbels et al.~derived the quantile function
+\(F_{\alpha}^{-1}(p)\), which incorporates the quantile function of the
+symmetric generalised normal distribution derived by Griffin (Griffin
+2018):
+
+\begin{definition}[Asymmetric generalised normal distribution quantile function]
+Let $p \in \left[0, 1\right]$ be a probability.
+Let  $F_{\Gamma}^{-1}$ is the quantile function of the gamma distribution with shape $1 / \beta$, which can be numerically approximated. 
+Let $F_{\beta}^{-1}$ be the quantile function of the generalised normal distribution:
+\begin{equation}
+F_{\beta}^{-1}(p) = \mathop{\mathrm{sgn}}\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
+\end{equation}
+
+Then the quantile function of the asymmetric generalised normal distribution is:
 \begin{equation}
@@
 \end{equation}
-
-The quantile function for the asymmetric generalised normal distribution
-\(F_{\alpha}^{-1}\) thus incorporates the quantile function
-\(F_{\beta}^{-1}\) of the symmetric generalised normal distribution.
-\(F_{\beta}^{-1}\) was derived by Griffin to be (Griffin 2018):
-
-\begin{equation}
-F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
-\end{equation}
-
-Here, \(F_{\Gamma}^{-1}\) is the quantile function of the gamma
-distribution with shape \(1 / \beta\), which can be numerically
-approximated.
-
-\subsection{Empirical central normality
-test}\label{empirical-central-normality-test}
+\end{definition}
+
+\subsection{Central normality test}\label{central-normality-test}
 
@@
 tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop an empirical test for central normality. The null
-hypothesis \(H_0\) is that the distribution is centrally normal. The
-alternative hypothesis \(H_1\) is that the distribution is not centrally
-normal.
-
-Let central normality be defined as the normality of central portion
-\(\kappa\) of the data,
-i.e.~\(\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\),
-with \(p_i\) probabilities of the empirical distribution, as previously.
-We then compute the residual errors between the z-scores of the
-transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
-expected z-scores from a standard normal distribution:
-\(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
-with \(\mu_M\) and \(\sigma_M\) robust Huber M-estimates of location and
-scale of the transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
-(Huber 1981). The set of residual errors for the central portion of the
-data is then
-\(\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\).
-
-The test statistic \(\tau_{\text{ecn}}\) is then defined as:
-
-\begin{equation}
-\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
-\end{equation}
-
-Here \([\quad]\) denotes an Iverson bracket. The test statistic is equal
-to the mean of the residual errors of the central portion of the data.
-
-To use the test statistic, the central portion of the data needs to be
-defined and the Type 1 error rates determined. We will do so in the
-\hyperref[simulation]{Simulation} section.
+Here we develop a test for central normality.
+
+\begin{definition}[Central portion of a sequence]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group, i.e.
+if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
+$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
+Furthermore, let central portion $\kappa \in (0,1)$.
+
+Then the central portion of $\mathbf{X}$ is 
+$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Residual errors of the central portion of a sequence]
+
+Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+
+Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
+$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Central normality of a sequence]
+
+The central portion of a sequence is normally distributed if the sum of residual
+errors is equal to zero:
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+
+\end{definition}
+
+In practice, a finite sequence sampled from a normal distribution
+\(\mathcal{N}(\mu, \sigma)\) will have a non-zero sum of residual
+errors.
+
+\begin{definition}[Central normality test]
+The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+
+The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+\end{definition}
+
+The null-hypothesis \(\mathcal{H}_0\) is that the central portion of
+sequence \(\mathbf{X}\) is normally distributed, with the alternative
+hypothesis \(\mathcal{H}_1\) that it is not normally distributed. To
+test this hypothesis, a test statistic is computed and compared against
+a critical value.
+
+\begin{definition}[Central normality test statistic]
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+
+The test statistic is then $\tau_{n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+
+\end{definition}
+
+The null-hypothesis should be rejected at significance level \(\alpha\)
+if \(\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}\).
+
+The central portion of the data needs to be defined and the Type 1 error
+rates determined to provide critical test statistics. We will do so in
+the \hyperref[simulation]{Simulation} section.
 
@@
 We used simulated data to assess invariance to location and scale of the
-proposed power transformations, weighting for robust transformations,
-and to develop the empirical central normality test. The \(\lambda\)
-parameter for conventional power transformations (Eqn.
+proposed power transformations, to develop the empirical central
+normality test, and to determine weighting for robust transformations.
+The \(\lambda\) parameter for conventional power transformations (Eqn.
 \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well
@@
 were implemented in the \texttt{power.transform} R software package
-(Zwanenburg and Löck 2024b) (version 1.0.0). Of note, the
+(Zwanenburg and Löck 2024b) (version 1.0.1). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 \(\lambda\) that are invariant to location and scale of the
-distribution, we simulated three different distributions. We first
-randomly drew \(10000\) values from a normal distribution:
+distribution, we simulated three different sequences. We first randomly
+drew \(10000\) values from a normal distribution:
 \(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
@@
 
+\subsection{Central normality and empirical central normality
+test}\label{central-normality-and-empirical-central-normality-test}
+
+To develop a test for central normality we need to consider two
+parameters: the central portion \(\kappa\) as a fixed parameter, and
+test statistic \(\tau_{\text{ecn}}\). We will first define the central
+portion \(\kappa\).
+
+For each number of samples
+\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}\)
+we randomly drew \(m_d = 30000\) ordered sequences \(\mathbf{X}\) from
+\(\mathcal{N}(0,1)\). This dataset was used to compute the critical test
+statistic values for the central normality test. Additionally, we added
+\(10 \%\) outliers by randomly replacing elements of each ordered
+sequence. The dataset with outliers was used to compute the critical
+test statistic values for the empirical variant of the central normality
+test. For each sequence we then computed \(\tau_{n, \kappa}\) for
+\(\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}\).
+
+Figure \ref{empirical-central-normality-kappa} shows
+\(\tau_{\alpha = 0.95, n, \kappa}\) as a function of \(n\) for different
+values of \(\kappa\). With decreasing \(\kappa\), the test statistic
+curve decreases. The curves for \(\kappa = 0.60\), \(\kappa = 0.70\) and
+\(\kappa = 0.80\) are similar, whereas curves for \(\kappa \geq 0.90\)
+are affected by outliers, as can be observed by comparing curves of the
+central normality test with those of the empirical variant.
+
+Since the (empirical) central normality tests assesses whether the
+central part of a sequence is normally distributed, we used
+\(\kappa = 0.80\). Critical statistic values for central normality and
+empirical central normality tests are shown are shown in Table
+\ref{tab:central-normality-critical-statistic} and Table
+\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-kappa-1}
+
+}
+
+\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the (empirical) central normality test as function of $n$ for several values of central portion $\kappa$.
+The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
+with fully normal data where 10\% of elements are replaced by outliers.
+emp: empirical}\label{fig:empirical-central-normality-kappa}
+\end{figure}
+
+To assess type I error rates for the (empirical) central normality test
+and compare these to the Shapiro-Wilk test for normality, we randomly
+drew another 1000 sequences from a standard normal distribution
+(\(\mathcal{N}(0,1)\) for each
+\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
+A second dataset was created by replacing elements with randomly drawn
+outliers, as described above. For each sequence, the p-value for the
+respective test was used to reject the null hypothesis that the sequence
+is (centrally) normal. The null hypothesis was rejected if
+\(p \leq 0.05\). The type I error rate was determined by computing the
+fraction of centrally normal sequences rejected this way. Figure
+\ref{empirical-central-normality-error_rate} shows that the empirical
+central normality test had the expected type I error rate for centrally
+normal data (with tails of the distribution not being normally
+distributed), whereas the central normality test and Shapiro-Wilk test
+overestimated type I error rates.
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
+
+}
+
+\caption{Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers.
+The empirical central normality (ECN) test uses $\kappa = 0.80$.
+CN: central normality}\label{fig:empirical-central-normality-error_rate}
+\end{figure}
+
+In Figure \ref{fig:empirical-central-normality-examples} we apply the
+empirical central normality test to assess central normality of features
+that are composed of a mixture of samples drawn from two normal
+distributions (\(n = 100\) each). With increased separation of the
+underlying normal distributions, the probability of the feature being
+centrally normal decreases, as expected.
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-examples-1}
+
+}
+
+\caption{Bi-modal distributions and empirical central normality test results.
+The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) drawn from normal distributions that are offset by a distance $d$.
+We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
+As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
+Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
+\end{figure}
+
 \subsection{Robust transformations}\label{robust-transformations}
@@
 To determine the weighting function parameters for each of the nine
-combinations, \(m_d=100\) asymmetric generalised normal distributions
-were drawn. Each distribution was parametrised with a randomly chosen
-skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
-parameter \(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale
-parameters were set as \(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) instances were then randomly drawn, with
-\(\gamma \sim U\left(2, 4\right)\), i.e., between \(100\) and \(10000\)
-values are drawn to create \(\mathbf{X}_i\).
-
-Outlier values were then drawn to randomly replace a fraction of the
-values of \(\mathbf{X}_i\). This was repeated \(m_{\text{out}} = 10\)
-times, with outlier fractions regularly spaced in \([0.00, 0.10]\). Thus
-up to 10 percent of the values could be replaced by outliers. Outlier
-values were set according to John W. Tukey (1977), as follows. Let
-\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
-was:
+combinations, \(m_d=500\) sequences \(\mathbf{X}_i\)
+(\(i \in \{1, 2, \ldots, m_d\}\)) were randomly drawn from randomly
+parametrised asymmetric generalised normal distributions. Each
+distribution was parametrised with a random skewness parameter
+\(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
+were set as \(\mu = 0\) and \(\sigma = 1\), respectively. To form each
+sequence \(\mathbf{X}_i\), \(n = \lceil 10^\gamma \rceil\) instances
+were then randomly drawn, with
+\(\gamma \sim U\left(\log_{10}50, 3\right)\), resulting in a set of
+sequences with between \(50\) and \(1000\) elements each.
+
+Outlier values were then drawn to randomly replace 10 percent of the
+elements of \(\mathbf{X}_i\). Outlier values were set according to John
+W. Tukey (1977), as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
+Then the corresponding outlier value was:
 
@@
 quartile and interquartile range of \(\mathbf{X}_i\), respectively.
-Outlier values randomly replaced values in \(\mathbf{X}_i\) to create
-\(\mathbf{X}_{i,j}\), with
-\(j \in \left\{ 1, \ldots, m_{\text{out}} \right\}\)
+Outlier values randomly replaced elements in \(\mathbf{X}_i\).
 
 To find the optimal values for the weighting function parameters
-\(\delta_1\) and \(\delta_2\) (if applicable), we minimised the absolute
-difference between the \(\lambda_{r}\) parameter obtained for robust
-transformation in the presence of outliers, and the \(\lambda_0\)
-parameter obtained using the non-robust transformations in absence of
-outliers:
-
-\begin{equation}
-\label{eqn:minimisation-weighting}
-\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
-\end{equation}
+\(\delta_1\) and \(\delta_2\) (if applicable), we minimised a composite
+loss
+\(L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}\).
+The composite loss consisted of two components: a loss term
+\(L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}\), i.e.~the mean of
+residual errors of the central 80\% of elements of each sequence
+\(\mathbf{X}_i\), which aimed at optimising central normality; and a
+loss term
+\(L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)\),
+with \(\lambda_{0,i}\) and \(\lambda_{i}\) transformation parameters
+found for sequence \(\mathbf{X}_i\) prior to and after adding outliers,
+respectively, and tolerance parameter \(\xi = 0.5\) for Box-Cox and
+\(\xi = 0.3\) for Yeo-Johnson power transformations. This second term
+aimed to prevent solutions that provide small improvements in central
+normality at the cost of a poor fit of the tails of sequences.
 
@@
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
-The {optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+The \textit{optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
 \label{tab:optimal-weighting-parameters-box-cox}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 49.6 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.80 & ---  & ---       & ---  & 38.5 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.01 & 1.00 & $(0, 1]$  & 0.86 & 43.7 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 1.00 & $(0, 1]$  & 0.90 & 42.7 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.39 & ---  & ---       & ---  & 52.2 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 2.39 & 2.40 & $(0, 10]$ & 4.92 & 52.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.07 & 3.63 & $(0, 10]$ & 3.43 & 57.4 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.83 & ---  & ---       & ---  & 53.1 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.80 & 0.81 & $(0, 10]$ & 0.84 & 54.4 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.76 & 0.77 & $(0, 10]$ & 0.76 & 53.2 \\
 \bottomrule
@@
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
-The {optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+The \textit{optimal} column shows the optimal parameter values. 
+The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
 \label{tab:optimal-weighting-parameters-yeo-johnson}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 42.1 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.78 & ---  & ---       & ---  & 35.1 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.20 & 0.95 & $(0, 1]$  & 1.00 & 34.0 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.54 & 0.95 & $(0, 1]$  & 1.00 & 32.6 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.32 & ---  & ---       & ---  & 43.0 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.28 & 1.96 & $(0, 10]$ & 3.43 & 49.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.41 & 1.96 & $(0, 10]$ & 3.75 & 51.3 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.92 & ---  & ---       & ---  & 65.4 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.87 & 1.00 & $(0, 10]$ & 0.87 & 66.0 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.06 & 1.00 & $(0, 10]$ & 1.07 & 67.1 \\
 \bottomrule
@@
 
-Figure \ref{fig:optimised-weighting-function-parameters} shows the
-distribution of errors \(\left| \lambda_r - \lambda_0 \right|\) for
-non-robust and robust transformations using the optimal weighting
-function parameters. In the presence of outliers, the Yeo-Johnson
-transformation yielded smaller errors than the Box-Cox transformation.
-For both transformations, weighting based on empirical probabilities
-yielded the most consistent \(\lambda\) parameter estimates in the
-presence of outliers. Other methods did not outperform the non-robust
-transformation method.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_files/figure-latex/optimised-weighting-function-parameters-1}
-
-}
-
-\caption{Robustness of power transformations after optimising weighting function parameters.
-The distribution of errors, i.e.
-the difference between robustly fitted $\lambda^r$ and expected $\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly parametrised asymmetric generalised normal distributions with up to 10\% outliers on either side of the distribution.
-The top and bottom panel show errors for the location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
-In each panel, the error distribution for the non-robust transformation is shown to the left for comparison with different weighting methods.
-Note that a square root transformation was applied along the $y$-axis for display purposes.
-Median errors are shown above each distribution.}\label{fig:optimised-weighting-function-parameters}
-\end{figure}
-
-\subsection{Empirical central normality
-test}\label{empirical-central-normality-test-1}
-
-To develop an empirical test for central normality we need to consider
-two parameters: the central portion \(\kappa\) as a fixed parameter, and
-test statistic \(\tau_{\text{ecn}}\). We will first define the central
-portion \(\kappa\).
-
-First we draw \(m_d=10000\) random asymmetric generalised normal
-distributions. As before, each distribution is parametrised with a
-randomly chosen skewness parameter
+\subsection{Assessing transformations using simulated
+data}\label{assessing-transformations-using-simulated-data}
+
+Three datasets with 10000 sequences each were created to assess
+transformation to normality. For each sequence \(\mathbf{X}_i\),
+\(n = \lceil 10^\gamma \rceil\) elements were randomly drawn, with
+\(\gamma \sim U\left(1, 4\right)\), resulting in sequences with between
+\(10\) and \(10000\) elements.
+
+\begin{itemize}
+\item
+  A \textit{clean} dataset with each sequence drawn from
+  \(\mathcal{N}(0, 1)\) and transformed using an inverse power
+  transformation with randomly drawn transformation parameter
+  \(\lambda \sim U\left(0.00, 2.00 \right)\).
+  \(\left(\phi_{\text{BC}}^\lambda\right)^{-1}\) and
+  \(\left(\phi_{\text{YJ}}^\lambda\right)^{-1}\) were used as inverse
+  transformations for assessing Box-Cox and Yeo-Johnson transformations,
+  respectively. Prior to inverse transformation, sequences for Box-Cox
+  transformations were shifted into the positive domain, with minimum
+  value \(1\).
+\item
+  A \textit{dirty} dataset with each sequence drawn from
+  \(\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)\),
+  with randomly drawn skewness parameter
   \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
-are set as \(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) values are then randomly drawn, with
-\(\gamma \sim U\left(1.47, 3.00\right)\), which leads to between \(30\)
-and \(1000\) values being drawn to create \(\mathbf{X}_i\). As before,
-for each distribution, up to \(10 \%\) of instances are replaced by
-outlier values, and this is repeated \(10\) times. We then compute
-residuals after optimising robust location- and scale-invariant
-transformations with the empirical tapered cosine weighting method.
-
-Figure \ref{fig:empirical-central-normality-residual-error} shows the
-residual errors of the transformed distribution as function of the
-empirical probability. As expected, the largest deviations from
-normality appear on the extremities of this range. For Box-Cox
-transformations, the effect of outliers on the lower end of the
-distribution leads to noticeable asymmetry. Between very low and very
-high empirical probabilities, residual errors are constrained and
-relatively flat. This is the candidate range for the central portion of
-the data.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-residual-error-1}
-
-}
-
-\caption{Residual error as a function of empirical probability.
-Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric generalised normal distributions, for each of which outliers were randomly drawn 10 times (up to 10\% of samples).
-Larger errors occur at the edges of each distribution.}\label{fig:empirical-central-normality-residual-error}
-\end{figure}
-
-We then consider the empirical probability of type I errors: the
-probability of incorrectly classifying a centrally normal distribution
-as being non-normal based on the test statistic \(\tau_{\text{ecn}}\).
-Under the assumption that the asymmetric generalised normal
-distributions are centrally normal after robust transformation, this
-produces the relationship shown in Figure
-\ref{fig:empirical-central-normality-type-1-error-rate}. Since for
-\(\kappa \leq 0.80\) error curves for both transformation methods are
-similar, we fix the value of the central portion \(\kappa\) to 0.80. In
-the remaining we will use the empirical central normality test statistic
-values defined using robust location- and scale-invariant Yeo-Johnson
-transformations, as it is more conservative (see Appendix E). This leads
-to test statistic values listed in Table
-\ref{tab:empirical-central-normality}
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-type-1-error-rate-1}
-
-}
-
-\caption{Type 1 error rate of transformed asymmetric generalised normal distributions as function of the test statistic $\tau_{\text{ecn}}$ for the central portion $\kappa$ of the distribution.}\label{fig:empirical-central-normality-type-1-error-rate}
-\end{figure}
+  \(\beta \sim U\left(1.00, 5.00 \right)\).
+\item
+  A \textit{shifted} dataset with each sequence drawn from
+  \(\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)\),
+  with randomly drawn skewness parameter
+  \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+  \(\beta \sim U\left(1.00, 5.00 \right)\).
+\end{itemize}
+
+A dataset with outliers was created for each of the above datasets by
+replacing 10 percent of elements in each sequence, as described earlier.
+
+In addition to no power transformation and location- and scale-invariant
+power transformations, conventional and Raymaekers and Rousseeuw's
+robust adaptation (Raymaekers and Rousseeuw 2024) were assessed. For the
+latter two methods, normalisation before standardisation using was
+additionally assessed using the following two methods:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  z-standardisation:
+  \(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
+  and \(\sigma\) the mean and standard deviation of sequence
+  \(\mathbf{X}\).
+\item
+  robust scaling:
+  \(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
+  with \(\text{IQR}\) representing the interquartile range of
+  \(\mathbf{X}\).
+\end{enumerate}
+
+This results in nine types of power transformation. Transformation
+parameters were optimised for each sequence. Subsequently the sum of
+residual errors and sum of residual errors of the central portion
+(\(\kappa = 0.80\)) were computed for each sequence after
+transformation. Then, each method was ranked according to the sum of
+residual errors (data without outliers) or the sum of residual errors of
+the central portion (data with outliers). The average rank of each
+transformation method was computed over all 10000 sequences in each
+dataset. Average ranks for Yeo-Johnson transformations are shown in
+Table \ref{tab:comparison_methods_simulation_yeo_johnson}. Location and
+shift-invariant Yeo-Johnson transformation ranked best for all datasets
+without outliers. The robust variant ranked best in dirty and shifted
+datasets with outliers, but not in the clean dataset. Results for
+Box-Cox transformations are shown in Table
+\ref{tab:comparison_methods_simulations_box_cox}.
 
@@
 \begin{center}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
-\label{tab:empirical-central-normality}
-\begin{tabular}{l | c c c c c c c}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson}
+\begin{tabular}{l | l r r r r r r}
 
 \toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
 
 \midrule
-$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+
+none                                  & &         8.51  &         7.60  &         7.33  &         6.15  &         7.44  &         6.60 \\
+conventional                          & &         3.70  &         5.70  &         3.87  &         7.06  &         6.49  &         5.93 \\
+conventional (z-score norm.)          & &         4.72  &         5.68  &         6.38  &         5.01  &         5.43  &         4.87 \\
+conventional (robust scaling)         & &         4.06  &         6.77  &         5.37  &         5.75  &         4.41  &         5.63 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.20  &         4.01  &         4.13  &         6.27  &         5.65 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.27  & \textbf{2.66} &         6.23  &         3.59  &         5.28  &         3.48 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.95  &         5.32  &         3.77  &         4.36  &         3.66 \\
+invariant                             & & \textbf{3.13} &         6.77  & \textbf{3.22} &         6.12  & \textbf{2.64} &         6.03 \\
+robust invariant                      & &         6.41  &         3.66  &         3.27  & \textbf{3.41} &         2.68  & \textbf{3.10} \\
+
 \bottomrule
@@
 \end{table}
-
-In Figure \ref{fig:empirical-central-normality-examples} we apply the
-empirical central normality test to assess central normality of features
-that are composed of a mixture of samples drawn from two normal
-distributions. With increased separation of the underlying normal
-distributions, the probability of the feature being centrally normal
-decreases, as expected.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-examples-1}
-
-}
-
-\caption{Bi-modal distributions and empirical central normality test results.
-The feature (black) is a mixture of two identical sample sets (blue and orange) drawn from normal distributions that are offset by a distance $d$.
-We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
-As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
-Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
-\end{figure}
 
@@
 yielded the following results, see Figure
-\ref{fig:experimental-results-invariance}: no transformation (sum of
-residuals with normal distribution \(\sum r_i = 16.5\)); conventional
-transformation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
-\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\));
-Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 2.0\),
-\(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
-\(\sigma_{YJ} = 0.5 \cdot 10^3\)); location- and scale-invariant
-transformation (\(\lambda = 0.9\), \(\sum r_i = 8.8\),
-\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.1\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.8\), \(\sum r_i = 10.6\),
-\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.0\)).
+\ref{fig:experimental-results-invariance}: no transformation
+(\(\sum r_i = 16.5\), \(p = 0.69\)); conventional transformation
+(\(\lambda = 2.0\), \(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
+\(\sigma_{YJ} = 0.5 \cdot 10^3\), \(p = 0.96\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
+\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\),
+\(p = 0.96\)); location- and scale-invariant transformation
+(\(\lambda = 1.3\), \(\sum r_i = 8.8\), \(\mu_{YJ} = -1.2\),
+\(\sigma_{YJ} = 1.1\), \(p = 0.98\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.3\), \(\sum r_i = 9.3\),
+\(\mu_{YJ} = -1.0\), \(\sigma_{YJ} = 1.1\), \(p = 0.93\)).
 
 Location- and scale-invariant transformation led to a lower overall
-residual sum, indicating a better transformation. Robust location- and
-scale-invariant transformation had a higher residual sum compared to the
-non-robust variant, which may be due to the lack of outliers in the
-data. Conventional transformations inflated the mean \(\mu_{YJ}\) and
-standard deviation \(\sigma_{YJ}\) of the age feature after
-transformation. The empirical central normality test did not detect any
-statistically significant deviations from central normality for any
-transformation (all \(p \geq 0.78\)).
+residual errors, indicating a better transformation. Conventional
+transformations inflated the mean \(\mu_{YJ}\) and standard deviation
+\(\sigma_{YJ}\) of the age feature after transformation. The empirical
+central normality test did not detect any statistically significant
+deviations from central normality for any transformation (all
+\(p \geq 0.93\)).
 
@@
 The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
-deviation) grams, and not centrally normal (\(p = 0.03\)). Applying
+deviation) grams, and not centrally normal (\(p < 0.001\)). Applying
 conventional and invariant Yeo-Johnson transformations to body mass
@@
 \ref{fig:experimental-results-invariance}: no transformation (residual
-sum \(\sum r_i = 48.0\)); conventional transformation
+sum \(\sum r_i = 48.0\), \(p < 0.001\)); conventional transformation
 (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); location- and scale-invariant
-transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
-\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.4\), \(\sum r_i = 23.1\),
-\(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.8\)).
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\));
+location- and scale-invariant transformation (\(\lambda = 0.5\),
+\(\sum r_i = 26.8\), \(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\),
+\(p = 0.28\)); and robust location- and scale-invariant transformation
+(\(\lambda = 0.3\), \(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\),
+\(\sigma_{YJ} = 0.9\), \(p = 0.69\)).
 
 Location- and scale-invariant transformation produced a lower overall
-residual sum, indicating a better transformation. Moreover, conventional
-transformations led to low standard deviation \(\sigma_{YJ}\) of the
-body mass feature after transformation. The empirical central normality
-test did not detect any statistically significant deviations from
-central normality for any transformation (all \(p \geq 0.38\)).
+residual errors, indicating a better transformation. Moreover,
+conventional transformations led to low standard deviation
+\(\sigma_{YJ}\) of the body mass feature after transformation. The
+empirical central normality test did not detect any statistically
+significant deviations from central normality for any transformation
+(all \(p \geq 0.10\)).
 
@@
 \ref{fig:experimental-results-invariance}: no transformation (residual
-sum \(\sum r_i = 328\)); conventional transformation
+sum \(\sum r_i = 328\), \(p < 0.001\)); conventional transformation
 (\(\lambda = 62.1\), \(\sum r_i = 319\),
-\(\mu_{YJ} = 4.8 \cdot 10^{99}\), \(\sigma_{YJ} = 0.1 \cdot 10^{99}\));
-Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 95.4\),
-\(\sum r_i = 319\), \(\mu_{YJ} = 6.4 \cdot 10^{153}\),
-\(\sigma_{YJ} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
-transformation (\(\lambda = 1.5\), \(\sum r_i = 326\),
-\(\mu_{YJ} = -1.2\), \(\sigma_{YJ} = 0.8\)); and robust location- and
-scale-invariant transformation (\(\lambda = 1.4\), \(\sum r_i = 311\),
-\(\mu_{YJ} = -1.3\), \(\sigma_{YJ} = 0.9\)).
-
-Every transformation reduced the residual sum. The non-robust location-
-and scale-invariant transformation did not improve over conventional
-alternatives and yielded a data distribution lacking central normality
-(empirical central normality test: \(p=0.05\)). However, conventional
+\(\mu_{YJ} = 4.8 \cdot 10^{99}\), \(\sigma_{YJ} = 0.1 \cdot 10^{99}\),
+\(p < 0.001\)); Raymaekers and Rousseeuw's robust adaptation
+(\(\lambda = 95.4\), \(\sum r_i = 315\),
+\(\mu_{YJ} = 6.4 \cdot 10^{153}\), \(\sigma_{YJ} = 0.3 \cdot 10^{153}\),
+\(p < 0.001\)); location- and scale-invariant transformation
+(\(\lambda = 1.5\), \(\sum r_i = 326\), \(\mu_{YJ} = -1.2\),
+\(\sigma_{YJ} = 0.8\), \(p < 0.001\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.1\), \(\sum r_i = 308\),
+\(\mu_{YJ} = -1.3\), \(\sigma_{YJ} = 1.2\), \(p < 0.001\)).
+
+Every transformation reduced the residual sum. None of the
+transformations yielded a centrally normal distribution. Conventional
 transformations had high values for the \(\lambda\) parameter, which
@@
 results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum \(\sum r_i = 54\), \(p=0.76\));
+no transformation (residual sum \(\sum r_i = 54\), \(p=0.72\));
 conventional transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
-\(\mu_{YJ} = 3.0\), \(\sigma_{YJ} = 0.3\), \(p=0.01\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
-\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.55\)); location- and
+\(\mu_{YJ} = 3.0\), \(\sigma_{YJ} = 0.3\), \(p < 0.001\)); Raymaekers
+and Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
+\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.37\)); location- and
 scale-invariant transformation (\(\lambda = -1.3\), \(\sum r_i = 44\),
-\(\mu_{YJ} = 0.5\), \(\sigma_{YJ} = 0.1\), \(p=0.03\)); and robust
-location- and scale-invariant transformation (\(\lambda = -0.9\),
-\(\sum r_i = 50\), \(\mu_{YJ} = 2.0\), \(\sigma_{YJ} = 2.3\),
-\(p=0.58\)).
+\(\mu_{YJ} = 0.5\), \(\sigma_{YJ} = 0.1\), \(p < 0.001\)); and robust
+location- and scale-invariant transformation (\(\lambda = 1.0\),
+\(\sum r_i = 56\), \(\mu_{YJ} = 1.7\), \(\sigma_{YJ} = 1.0\),
+\(p=0.76\)).
 
@@
 to a centrally normal distribution (empirical central normality test
-\(p = 0.01\) and \(p=0.03\) for conventional and invariant
-transformations, respectively). Robust transformations produce
-distributions that are centrally normal (empirical central normality
-test \(p > 0.05\)).
+\(p < 0.001\) for conventional and invariant transformations). Robust
+transformations produce distributions that are centrally normal
+(empirical central normality test \(p > 0.05\)).
 
@@
 results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
+no transformation (residual sum \(\sum r_i = 110\), \(p=0.83\));
 conventional transformation (\(\lambda = -0.7\), \(\sum r_i = 30\),
-\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p=0.01\)); Raymaekers and
+\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p=0.003\)); Raymaekers and
 Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 136\),
-\(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14\), \(p=0.61\)); location- and
+\(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14.3\), \(p=0.88\)); location- and
 scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 12\),
-\(\mu_{YJ} = -11.8\), \(\sigma_{YJ} = 6.9\), \(p=0.13\)); and robust
-location- and scale-invariant transformation (\(\lambda = -0.6\),
-\(\sum r_i = 27\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.1\),
-\(p=0.10\)).
-
-Non-robust transformations failed to produce a centrally normal
-distribution (empirical central normality test \(p=0.01\) and \(p=0.02\)
-for conventional and invariant transformations, respectively). Robust
-transformations produce distributions that are centrally normal
+\(\mu_{YJ} = -11.8\), \(\sigma_{YJ} = 6.9\), \(p=0.15\)); and robust
+location- and scale-invariant transformation (\(\lambda = -0.3\),
+\(\sum r_i = 30\), \(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.2\),
+\(p=0.18\)).
+
+The conventional non-robust transformation failed to produce a centrally
+normal distribution (empirical central normality test \(p=0.003\)).
+Robust transformations produce distributions that are centrally normal
 (empirical central normality test \(p > 0.05\)).
@@
 
-We used 285 datasets from the Penn Machine Learning Benchmarks
-collection (Romano et al. 2022). In this collection, 122 datasets
-correspond to regression tasks and 163 datasets to classification tasks.
-Using the familiar auto-machine learning library (Zwanenburg and Löck
-2024a) (version 1.5.0), each dataset was used to train a model for each
-of 16 process configurations. Each process configuration specifies the
-learner (generalised linear model or random forest), transformation
-method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson,
-robust invariant Yeo-Johnson with empirical central normality test
-(rejecting transformations with \(p \leq 0.01\)), and normalisation
-method (none, \(z\)-standardisation), yielding 16 distinct
-configurations. Before each experiment, each dataset was randomly split
-into a training (70\%) and holdout test (30\%) set five times. Thus, a
-total of 22800 models were created. Each model was then evaluated using
-the holdout test set using one of two metrics, i.e.~the root relative
-squared error (RRSE) for regression tasks and the area under the
-receiver operating characteristic curve (AUC) for classification tasks.
+We used 231 datasets containing at least one numeric feature from the
+Penn Machine Learning Benchmarks collection (Romano et al. 2022). In
+this collection, 114 datasets correspond to regression tasks and 117
+datasets to classification tasks. Using the familiar auto-machine
+learning library (Zwanenburg and Löck 2024a) (version 1.5.0), each
+dataset was used to train a model for each of 32 process configurations.
+Each process configuration specifies the learner (generalised linear
+model, L1-regularised linear models (Lasso), gradient boosted linear
+model, or random forest), transformation method (none, conventional
+Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson
+with empirical central normality test (rejecting transformations with
+\(p \leq 0.01\)), and normalisation method (none,
+\(z\)-standardisation), yielding 32 distinct configurations. Before each
+experiment, each dataset was randomly split into a training (70\%) and
+holdout test (30\%) set five times. Thus, a total of 36960 models were
+created. Each model was then evaluated using the holdout test set using
+one of two metrics, i.e.~the root relative squared error (RRSE) for
+regression tasks and the area under the receiver operating
+characteristic curve (AUC) for classification tasks.
 
@@
 \item
-  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (87
+  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (57
   datasets)
@@
   easy: \(0.90 > \text{AUC} \geq 0.80\) or
-  \(0.30 \geq \text{RRSE} > 0.10\) (46 datasets)
+  \(0.30 \geq \text{RRSE} > 0.10\) (40 datasets)
 \item
   intermediate: \(0.80 > \text{AUC} \geq 0.70\) or
-  \(0.60 \geq \text{RRSE} > 0.30\) (72 datasets)
+  \(0.60 \geq \text{RRSE} > 0.30\) (52 datasets)
 \item
   difficult: \(0.70 > \text{AUC} \geq 0.60\) or
-  \(0.80 \geq \text{RRSE} > 0.60\) (60 datasets)
+  \(0.80 \geq \text{RRSE} > 0.60\) (33 datasets)
 \item
   very difficult: \(0.60 > \text{AUC} \geq 0.50\) or
-  \(1.00 \geq \text{RRSE} > 0.80\) (12 datasets)
-\item
-  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (8
-  datasets)
+  \(1.00 \geq \text{RRSE} > 0.80\) (48 datasets)
+\item
+  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (1
+  dataset)
 \end{itemize}
@@
 
-Here we focus on the subset of 232 datasets that contain numeric
-features. Considering single process parameters, the choice of learner
-(Wilcoxon signed rank test: \(p < 10^{-8})\)) and normalisation method
-(Wilcoxon signed rank test: \(p = 0.007\)) had a significant impact (at
-significance level \(p = 0.050\)), but transformation method (Friedman
-test: \(p = 0.054\)) did not.
+Considering single process parameters, the choice of learner (Friedman
+test: \(p < 10^{-8})\)), normalisation method (Wilcoxon signed rank
+test: \(p = 4 \cdot 10^{-8}\)), and transformation method (Friedman
+test:\(p = 0.007\), all had a significant impact (at \(p = 0.05\)).
 
@@
 transformation method, we first fit a regression random forest (ranger
-package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, minimum
-node size 2, other hyperparameters default) with process parameters and
-task difficulty as predictors and normalised rank as response variable.
-The estimated marginal effects are shown in Figure
+package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
+2, other hyperparameters default) with process parameters and task
+difficulty as predictors and normalised rank as response variable. The
+estimated marginal effects are shown in Figure
 \ref{fig:marginal-effect-plot}. On the scale of normalised ranks
-(\([0.0, 1.0]\)), the overall estimated marginal effect of using a
-random forest instead of generalised linear model was \(0.272\). The
-overall marginal effect of using z-standardisation to normalise features
-was \(0.008\). Transformation methods had the following marginal
-effects: \(0.003\) for using conventional Yeo-Johnson transformation
-instead of no transformation; \(-0.007\) for using robust invariant
-Yeo-Johnson transformation instead of no transformation; \(-0.009\) for
-using robust invariant Yeo-Johnson transformation with empirical central
-normality test instead of no transformation; \(-0.010\) for using robust
-invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson
+(\([0.0, 1.0]\)), only the random forest performed better than the
+average of all learners (rank difference \(0.173\)). The overall
+marginal effect of using z-standardisation to normalise features was
+\(0.011\). Transformation methods had the following marginal effects:
+\(-0.001\) for using conventional Yeo-Johnson transformation instead of
+no transformation; \(-0.012\) for using robust invariant Yeo-Johnson
+transformation instead of no transformation; \(-0.012\) for using robust
+invariant Yeo-Johnson transformation with empirical central normality
+test instead of no transformation; \(-0.012\) for using robust invariant
+Yeo-Johnson transformation instead of conventional Yeo-Johnson
 transformation; and \(-0.012\) for using using robust invariant
@@
 
-\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets.
-The top-left panel shows the marginal effect of learners, i.e.
-random forests and generalised linear models (GLM).
-Random forests outperform GLM models for all task difficulties.
+\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 36960 machine learning experiments on 231 datasets.
+The four top-left panels shows the marginal effect of learners vs.
+the average, i.e.
+generalised linear models (GLM), L1-regularised LM (Lasso), gradient-boosted LM and random forest.
+Random forests outperform the average.
 The top-right panel shows the marginal effect of feature normalisation methods, i.e.
 no normalisation and z-standardisation.
-z-standardisation is generally beneficial, but the estimated effect is neglible.
-The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method.
-There is no consistent behaviour, and estimated effects are neglible.
+z-standardisation is generally beneficial, but the estimated effect is marginal.
+The bottom panel shows the marginal effects of different transformation methods, split by normalisation method.
+The estimated effects are similar in size to the effect of normalisation and marginal.
 Note that the ranges of the $x$-axes of the three main panels differ.
@@
 of a feature. This is particularly relevant for automated data
-processing, where such issues may go unnoticed. Compared to non-robust
-location- and scale-invariant transformations, real-world examples with
-outliers showed higher residual errors of transformed features. Robust
-transformations seek to minimise residual errors for the central part of
-the distribution, instead of the entire distribution, including
-outliers. The empirical central normality test indicated that robust
-transformations are better able to achieve central normality in the
-presence of outliers. However, in a machine learning experiment of 232
-real-world datasets that contained at least one numeric feature, we did
-not find a meaningful benefit -- nor detriment -- to model performance
-for location- and scale-invariant power transformations. One reason may
-be that numeric features with large location shifts (\(|\mu| > 1000.0\))
-were uncommon. Of the 4886 numeric features in the 232 datasets, 266
+processing, where such issues may go unnoticed.
+
+In simulation, robust location- and scale-invariant transformations
+ranked best in datasets with outliers, and ranked highly in datasets
+without outliers, except for data without outliers that were directly
+derived from strictly normal distributions. This could be observed in
+real-world examples, where robust location-and scale-invariant
+transformations achieved central normality for when the non-robust
+variant could not.
+
+However, in a machine learning experiment of 231 real-world datasets
+that contained at least one numeric feature, we did not find a
+meaningful benefit -- nor detriment -- to model performance for
+location- and scale-invariant power transformations. One reason may be
+that numeric features with large location shifts (\(|\mu| > 1000.0\))
+were uncommon. Of the 4886 numeric features in the 231 datasets, 266
 (5\%) features in 34 datasets had large location shifts, of which 200
@@
 through subtracting its median value and division by its interquartile
-range) prior to conventional power transformations may achieve a similar
-effect in reducing sensitivity to the distribution's location and scale.
-While this alternative helps prevent these issues -- provided that
-normalisation does not lead to negative values for Box-Cox
-transformation -- location- and scale-invariant transformations seem to
-provide an overall better transformation to normality (Appendix F).
-
-We assessed several methods for robust power transformation. Methods
-that relied on the z-score of the transformed feature or the residual
-error yielded worse results than the non-robust method. This is partly
-due to the initial choice of threshold parameters. Using different
-initial values, closer to the upper limits
-(\(\delta_1 = 10.0; \delta_2 = 10.0\)), led to a reduced loss. However,
-these values effectively correspond to a non-robust transformation,
-where all instances receive the same weight. Underperformance of these
-methods could be explained by their use of transformed features for
-setting weights. Consequently, the weights change at each iteration in
-the MLE optimisation process. This increases local variance in the
+range) prior to conventional power transformations can achieve a similar
+effect in reducing sensitivity to the distribution's location and scale,
+as demonstrated by our findings. While this alternative helps prevent
+these issues -- provided that normalisation does not lead to negative
+values for Box-Cox transformations -- location- and scale-invariant
+transformations are able to achieve better transformations to normality,
+demonstrated by a lower residual error.
+
+We assessed several weighting methods to achieve robust power
+transformations. Methods that relied on the z-score of the transformed
+feature or the residual error yielded worse results than the non-robust
+method. Underperformance of these weighting methods could be explained
+by their reliance on transformed feature values for setting weights.
+Consequently, their weights change at each iteration in the MLE
+optimisation process. This increases local variance in the
 log-likelihood function and creates local optima that the optimiser may
-not handle well. As a consequence, optimal values for transformation
-parameter \(\lambda\) might differ, which increases the presented loss
-for optimising threshold parameters. Methods that relied on the
-empirical probability did not suffer from this issue, as weights
-remained fixed during MLE.
+not handle well. Methods that relied on the empirical probability did
+not suffer from this issue, as weights remained fixed during MLE.
 
 We introduced an empirical test for central normality to assess whether
-distributions deviate from normality in a way that might require closer
+sequences deviate from normality in a way that might require closer
 inspection prior to further processing. The empirical test for central
 normality differs from other tests for normality, such as the
-Shapiro-Wilk test (Shapiro and Wilk 1965), in that the test statistic is
-independent of the number of samples. This makes this test more
-practical for assessing central normality for larger sample numbers,
-where other tests may detect inconsequential deviations from normality.
+Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
+the central portion of a feature, instead of the entire feature.
 
@@
 would lead to very large or small numbers when using conventional power
-transformations. For MLE stability issues were not observed. Secondly,
-the empirical central normality test is based on simulations instead of
-statistical theory, and relies on a somewhat arbitrary definition of the
-central portion of a distribution. Thus, while the test may asses
-whether data is sufficiently normally distributed for practical
-purposes, it should not be used as a strict test for normality.
+transformations. For MLE stability issues were not observed.
 
@@
 results for the machine learning experiment are separately available
-from \href{https://doi.org/10.5281/zenodo.13736671}{Zenodo}.
+from \href{https://doi.org/10.5281/zenodo.14986689}{Zenodo}.
 

### 2025-04-09 manuscript_rmarkdown/plots.R

@@ -2250,7 +2250,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   data_2_labels <- data.table::data.table(
     x = data_2_x_range,
     task_difficulty  = factor(c("overall", "overall"), levels = difficulty_labels),
-    label = c("none better", "z-stand. better")
+    label = c("worse", "z-stand. better")
   )
 
   p2 <- ggplot2::ggplot(
@@ -2295,7 +2295,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   data_1_labels <- data.table::data.table(
     x = data_1_x_range,
     task_difficulty  = factor(c("overall", "overall"), levels = difficulty_labels),
-    label = c("worse than avg.", "better than avg.")
+    label = c("worse", "better than avg.")
   )
 
   p1 <- ggplot2::ggplot(
@@ -2338,7 +2338,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p1 <- p1 + ggplot2::ggtitle("learner vs. average")
 
   p <- (p1 + p2 + patchwork::plot_layout(ncol = 2, axes = "collect_y", axis_titles = "collect_y", widths = c(0.8, 0.2))) / p3 + patchwork::plot_layout(
-    heights = c(0.3, 1.0)
+    heights = c(0.3333, 0.6667)
   )
 
   return(p)

### 2025-04-11 manuscript_rmarkdown/manuscript.Rmd
Fixed spelling, grammar and style issues.

@@
 
-Power transformations are used to stabilize variance and achieve normality of features, especially in methods assuming normal distributions such as ANOVA and linear discriminant analysis.
+Power transformations are used to stabilize variance and achieve normality of features, 
+especially in methods assuming normal distributions such as ANOVA and linear discriminant analysis.
 However, the commonly used Box-Cox and Yeo-Johnson power transformation methods are sensitive to the location, scale, and presence of outliers in the data.
 Here we present location- and scale-invariant Box-Cox and Yeo-Johnson transformations to mitigate these issues.
-We derive maximum likelihood estimation criteria for optimizing transformation parameters and propose robust adaptations that reduce the influence of outliers.
+We derive maximum likelihood estimation criteria for optimizing transformation parameters 
+and propose robust adaptations that reduce the influence of outliers.
 We also introduce an empirical test for assessing central normality of transformed features.
-In simulations and real-world datasets, robust location- and scale-invariant transformations outperform conventional variants, resulting in better transformations to central normality.
+In simulations and real-world datasets, robust location- and scale-invariant
+transformations outperform conventional variants, resulting in better transformations to central normality.
 In a machine learning experiment with 231 datasets with numerical features, 
-integrating robust location- and scale-invariant power transformations into an automated data processing and machine learning pipeline did not result in a meaningful improvement or detriment to model performance compared to conventional variants.
+integrating robust location- and scale-invariant power transformations into an automated
+data processing and machine learning pipeline did not result in a meaningful improvement or detriment to model performance compared to conventional variants.
 In conclusion, robust location- and scale-invariant power transformations can replace conventional variants.
@@
   "$10000$ samples were drawn from a normal distribution: ",
-  "$\\mathcal{N}(\\mu, 1)$ for the shift dataset, $\\mathcal{N}(10, \\sigma)$ for the scale dataset and $\\mathcal{N}(0, 1)$ for the outlier dataset. ",
+  "$\\mathcal{N}(\\mu, 1)$ for the \\textit{shift} dataset, $\\mathcal{N}(10, \\sigma)$ for the \\textit{scale} dataset and $\\mathcal{N}(0, 1)$ for the \\textit{outlier} dataset. ",
   "Additionally, an outlier with value $d$ was added to the outlier dataset. ",
@@
 Finally, we define the overall framework for the empirical central normality test.
+
+First, we define a (numeric) feature. Throughout this work, whenever reference is made to a feature,
+a numeric feature is meant unless noted otherwise.
+
+\begin{definition}[Numeric feature]
+A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
+\end{definition}
 
@@
 Here, weighted maximum likelihood estimation is based on equations \ref{eqn:box-cox-invariant-log-likelihood} and \ref{eqn:yeo-johnson-invariant-log-likelihood}.
-Compared to @Raymaekers2024-zf, these log-likelihood functions includes additional terms to accommodate estimation of $x_0$ and $s$.
+Compared to @Raymaekers2024-zf, these log-likelihood functions include additional terms to accommodate estimation of $x_0$ and $s$.
 The weighted location- and scale-invariant Box-Cox log-likelihood function is:
 
-```{=latex}
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation is:
+
 \begin{equation}
@@
 \end{equation}
-```
-where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
+\end{definition}
 
@@
 
-```{=latex}
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
+
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is:
+
 \begin{equation}
@@
 \end{equation}
-```
-where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:
-
-```{=latex}
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
-```
+\end{definition}
 
@@
 - A step function, with $\delta_1 \geq 0$ as threshold parameter:
-```{=latex}
-\begin{equation}
+\begin{equation*}
 w_i =
@@
 \end{cases}
-\end{equation}
-```
+\end{equation*}
 
 - A triangle function (or generalised Huber weight), with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
-```{=latex}
-\begin{equation}
+\begin{equation*}
 w_i =
@@
 \end{cases}
-\end{equation}
-```
+\end{equation*}
 
 - A tapered cosine function [@Tukey1967-eb], with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
-```{=latex}
-\begin{equation}
+\begin{equation*}
 w_i =
@@
 \end{cases}
-\end{equation}
-```
-
-All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< \delta_1$, instances are fully weighted, i.e.
-when $\delta_1 > 0$ the weighting functions are symmetric window functions with a flat top.
+\end{equation*}
+
+All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< \delta_1$, 
+instances are fully weighted, i.e. when $\delta_1 > 0$ the weighting functions are symmetric window functions with a flat top.
 The triangle and tapered cosine functions then gradually down-weight instances with $\delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2$, and assign no weight to instances $\left| x_i \right| > \delta_2$.
@@
 
-Figure \ref{empirical-central-normality-kappa} shows $\tau_{\alpha = 0.95, n, \kappa}$
+Figure \ref{fig:empirical-central-normality-kappa} shows $\tau_{\alpha = 0.05, n, \kappa}$
 as a function of $n$ for different values of $\kappa$. With decreasing
@@
 The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
-Figure \ref{empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
+Figure \ref{fig:empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
 centrally normal data (with tails of the distribution not being normally distributed),
@@
 see equations \ref{eqn:box-cox-weighted-invariant-log-likelihood} and \ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}.
-We proposed three weighting function: step, triangle and tapered cosine, that have one, two and two parameters, respectively.
-Each weighting function then takes one of three values as input:
+We propose three weighting functions: step, triangle and tapered cosine, that have one, two and two parameters, respectively.
+Each weighting function then uses one of the following inputs:
 probabilities of the empirical distribution of the original feature,
@@
 
-Location- and scale-invariant transformation led to a lower overall residual errors, indicating a better transformation.
+Location- and scale-invariant transformation led to a lower overall residual error, indicating a better transformation to normality.
 Conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
@@
 normalisation method (Wilcoxon signed rank test: $p = 4 \cdot 10^{-8}$), 
-and transformation method (Friedman test:$p = 0.007$,
+and transformation method (Friedman test: $p = 0.007$),
 all had a significant impact (at $p = 0.05$).
@@
 On the scale of normalised ranks ($[0.0, 1.0]$), only the random forest performed better than the average of all learners (rank difference $0.173$).
-The overall marginal effect of using z-standardisation to normalise features was $0.011$.
+The marginal improvement in performance from z-standardisation was $0.011$.
 Transformation methods had the following marginal effects:
@@
 In simulation, robust location- and scale-invariant transformations ranked best in datasets with outliers,
-and ranked highly in datasets without outliers, except for data without outliers that were 
-directly derived from strictly normal distributions.
-This could be observed in real-world examples, where robust location-and scale-invariant transformations
-achieved central normality for when the non-robust variant could not.
+and ranked highly in datasets without outliers, except for features without outliers that were 
+directly sampled from strictly normal distributions.
+In real-world examples, robust location-and scale-invariant transformations
+achieved central normality when the non-robust variant could not.
 
@@
 prior to conventional power transformations can achieve a similar effect in reducing sensitivity
-to the distribution's location and scale, as demonstrated by our findings.
+to the feature's location and scale.
 While this alternative helps prevent these issues 
--- provided that normalisation does not lead to negative values for Box-Cox transformations -- 
+-- provided that standardisation does not lead to negative values for Box-Cox transformations -- 
 location- and scale-invariant transformations are able to achieve better transformations
-to normality, demonstrated by a lower residual error.
+to normality, as demonstrated by lower residual errors in simulation experiments.
 
@@
 of a feature, instead of the entire feature. 
-
-This work has the following limitations.
-Firstly, we did observe several numerical stability issues for optimisation criteria other than MLE (Appendix B).
-These appear in regions where transformation parameters would lead to very large or small numbers when using conventional power transformations.
+Compared to the central normality test and the Shapiro-Wilk test, the empirical central
+normality test remains consistent in the presence of outliers, although the former
+tests are more powerful when outliers are absent.
+
+This work has the following limitation: We observed several numerical stability
+issues for optimisation criteria other than MLE (Appendix B). These appear in regions
+where transformation parameters would lead to very large or small numbers when using conventional power transformations.
 For MLE stability issues were not observed.

### 2025-04-11 manuscript_rmarkdown/manuscript.tex
Fixed spelling, grammar and style issues.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-04-09}
+\date{2025-04-11}
 
@@
 \caption{Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\lambda$.
-$10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the shift dataset, $\mathcal{N}(10, \sigma)$ for the scale dataset and $\mathcal{N}(0, 1)$ for the outlier dataset.
+$10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the \textit{shift} dataset, $\mathcal{N}(10, \sigma)$ for the \textit{scale} dataset and $\mathcal{N}(0, 1)$ for the \textit{outlier} dataset.
 Additionally, an outlier with value $d$ was added to the outlier dataset.
@@
 empirical central normality test.
+
+First, we define a (numeric) feature. Throughout this work, whenever
+reference is made to a feature, a numeric feature is meant unless noted
+otherwise.
+
+\begin{definition}[Numeric feature]
+A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
+\end{definition}
 
@@
 
-\begin{definition}
+\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
 Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
@@
 \ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
-and Rousseeuw (2024), these log-likelihood functions includes additional
+and Rousseeuw (2024), these log-likelihood functions include additional
 terms to accommodate estimation of \(x_0\) and \(s\). The weighted
 location- and scale-invariant Box-Cox log-likelihood function is:
+
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation is:
 
@@
 \end{equation}
-
-where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
-variance of the Box-Cox transformed feature
-\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\):
-
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
+\end{definition}
 
@@
 log-likelihood function is:
+
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
+
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is:
 
@@
 \mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
-& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
 \end{split}
 \end{equation}
-
-where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
-variance of the Yeo-Johnson transformed feature
-\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):
-
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
+\end{definition}
 
@@
 \begin{itemize}
-\tightlist
 \item
   A step function, with \(\delta_1 \geq 0\) as threshold parameter:
-\end{itemize}
-
-\begin{equation}
+  \begin{equation*}
   w_i =
@@
   \end{cases}
-\end{equation}
-
-\begin{itemize}
-\tightlist
+  \end{equation*}
 \item
@@
   \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
-  parameters:
-\end{itemize}
-
-\begin{equation}
+  parameters: \begin{equation*}
   w_i =
@@
   and \(\delta_2 \geq \delta_1\) as threshold parameters:
-\end{itemize}
-
-\begin{equation}
+  \begin{equation*}
   w_i =
@@
   \end{cases}
-\end{equation}
+  \end{equation*}
+\end{itemize}
 
@@
 
-Figure \ref{empirical-central-normality-kappa} shows
-\(\tau_{\alpha = 0.95, n, \kappa}\) as a function of \(n\) for different
+Figure \ref{fig:empirical-central-normality-kappa} shows
+\(\tau_{\alpha = 0.05, n, \kappa}\) as a function of \(n\) for different
 values of \(\kappa\). With decreasing \(\kappa\), the test statistic
@@
 fraction of centrally normal sequences rejected this way. Figure
-\ref{empirical-central-normality-error_rate} shows that the empirical
-central normality test had the expected type I error rate for centrally
-normal data (with tails of the distribution not being normally
+\ref{fig:empirical-central-normality-error_rate} shows that the
+empirical central normality test had the expected type I error rate for
+centrally normal data (with tails of the distribution not being normally
 distributed), whereas the central normality test and Shapiro-Wilk test
@@
 \ref{eqn:box-cox-weighted-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}. We proposed
-three weighting function: step, triangle and tapered cosine, that have
+\ref{eqn:yeo-johnson-weighted-invariant-log-likelihood}. We propose
+three weighting functions: step, triangle and tapered cosine, that have
 one, two and two parameters, respectively. Each weighting function then
-takes one of three values as input: probabilities of the empirical
+uses one of the following inputs: probabilities of the empirical
 distribution of the original feature, the z-score of the transformed
@@
 Location- and scale-invariant transformation led to a lower overall
-residual errors, indicating a better transformation. Conventional
-transformations inflated the mean \(\mu_{YJ}\) and standard deviation
-\(\sigma_{YJ}\) of the age feature after transformation. The empirical
-central normality test did not detect any statistically significant
-deviations from central normality for any transformation (all
-\(p \geq 0.93\)).
+residual error, indicating a better transformation to normality.
+Conventional transformations inflated the mean \(\mu_{YJ}\) and standard
+deviation \(\sigma_{YJ}\) of the age feature after transformation. The
+empirical central normality test did not detect any statistically
+significant deviations from central normality for any transformation
+(all \(p \geq 0.93\)).
 
@@
 test: \(p = 4 \cdot 10^{-8}\)), and transformation method (Friedman
-test:\(p = 0.007\), all had a significant impact (at \(p = 0.05\)).
+test: \(p = 0.007\)), all had a significant impact (at \(p = 0.05\)).
 
@@
 (\([0.0, 1.0]\)), only the random forest performed better than the
-average of all learners (rank difference \(0.173\)). The overall
-marginal effect of using z-standardisation to normalise features was
-\(0.011\). Transformation methods had the following marginal effects:
-\(-0.001\) for using conventional Yeo-Johnson transformation instead of
-no transformation; \(-0.012\) for using robust invariant Yeo-Johnson
+average of all learners (rank difference \(0.173\)). The marginal
+improvement in performance from z-standardisation was \(0.011\).
+Transformation methods had the following marginal effects: \(-0.001\)
+for using conventional Yeo-Johnson transformation instead of no
+transformation; \(-0.012\) for using robust invariant Yeo-Johnson
 transformation instead of no transformation; \(-0.012\) for using robust
@@
 ranked best in datasets with outliers, and ranked highly in datasets
-without outliers, except for data without outliers that were directly
-derived from strictly normal distributions. This could be observed in
-real-world examples, where robust location-and scale-invariant
-transformations achieved central normality for when the non-robust
-variant could not.
+without outliers, except for features without outliers that were
+directly sampled from strictly normal distributions. In real-world
+examples, robust location-and scale-invariant transformations achieved
+central normality when the non-robust variant could not.
 
@@
 range) prior to conventional power transformations can achieve a similar
-effect in reducing sensitivity to the distribution's location and scale,
-as demonstrated by our findings. While this alternative helps prevent
-these issues -- provided that normalisation does not lead to negative
-values for Box-Cox transformations -- location- and scale-invariant
-transformations are able to achieve better transformations to normality,
-demonstrated by a lower residual error.
+effect in reducing sensitivity to the feature's location and scale.
+While this alternative helps prevent these issues -- provided that
+standardisation does not lead to negative values for Box-Cox
+transformations -- location- and scale-invariant transformations are
+able to achieve better transformations to normality, as demonstrated by
+lower residual errors in simulation experiments.
 
@@
 the central portion of a feature, instead of the entire feature.
-
-This work has the following limitations. Firstly, we did observe several
-numerical stability issues for optimisation criteria other than MLE
-(Appendix B). These appear in regions where transformation parameters
-would lead to very large or small numbers when using conventional power
+Compared to the central normality test and the Shapiro-Wilk test, the
+empirical central normality test remains consistent in the presence of
+outliers, although the former tests are more powerful when outliers are
+absent.
+
+This work has the following limitation: We observed several numerical
+stability issues for optimisation criteria other than MLE (Appendix B).
+These appear in regions where transformation parameters would lead to
+very large or small numbers when using conventional power
 transformations. For MLE stability issues were not observed.

### 2025-04-11 manuscript_rmarkdown/manuscript_appendix.Rmd
Fixed spelling, grammar and style issues.

@@
 
-The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
+The parameters of these power transformations can be optimised by maximising the
+log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
 The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
@@
 
-Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively.
+Let $\mathbf{X}$ be a feature with ordered feature values,
+and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and
+$\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ 
+its transformed values using conventional and location- and scale-invariant power transformations, respectively.
 Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.
 
-Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
+Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed feature.
 Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
@@
 For non-robust power transformations, all $w_i = 1$.
-Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include weights.
-The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.
+Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz],
+which involves solving (or approximating) an integral function,
+contains an extra scalar multiplication term, and does not include weights.
+The Anderson-Darling criterion seeks to minimise the squared Euclidean distance 
+between the EDF and the normal CDF, with differences at the upper and lower end
+of the normal CDF receiving more weight than those at the centre of the CDF.
 
@@
 For non-robust power transformations, all $w_i = 1$.
-The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of weights.
+The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef],
+aside from an additive scalar value and the introduction of weights.
 This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
@@
 
-For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
+For conventional power transformations with a fixed shift parameter,
+the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted 
+by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.
 
@@
 The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
-We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
+We will first define the location $\mu$ and scale $\sigma$ of the transformed feature
+as these are required for computing skewness and kurtosis.
 Here, $\mu$ is defined as:
@@
 
-The location, or mean, is weighted using weights $w_i$.
-For non-robust transformations, $w_i = 1$.
+The location, or mean, is weighted by $w_i$, which equals $1$ for non-robust transformations.
+
 Then, $\sigma^2$ is defined as:
@@
 \textbf{central normality} at $\kappa = 0.80$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
 \label{tab:central-normality-critical-statistic}
@@
 \textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+level $\alpha$ and number of instances $n$.The values in this table should be divided by $100$ before use.}
 \label{tab:empirical-central-normality-critical-statistic}
@@
 transformations, each sequence was shifted prior to inverse transformation
-so that every element after inverse transformation would be strictly positive.
+ensuring all elements were strictly positive after inverse transformation.
 

### 2025-04-11 manuscript_rmarkdown/manuscript_appendix.tex
Fixed spelling, grammar and style issues.

@@
 
-\section{Appendix A: Log-likelihood functions for location and scale
-invariant power
-transformation}\label{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}
-
-Location and scale-invariant Box-Cox and Yeo-Johnson transformations are
-parametrised using location \(x_0\) and scale \(s\) parameters, in
+\section{Appendix A: Log-likelihood functions for location- and
+scale-invariant power
+transformation}\label{appendix-a-log-likelihood-functions-for-location--and-scale-invariant-power-transformation}
+
+Location- and scale-invariant Box-Cox and Yeo-Johnson transformations
+are parametrised using location \(x_0\) and scale \(s\) parameters, in
 addition to transformation parameter \(\lambda\). This leads to the
-following transformations. The location and scale-invariant Box-Cox
+following transformations. The location- and scale-invariant Box-Cox
 transformation is:
@@
 
-where \(x_i - x_0 > 0\). The location and scale-invariant Yeo-Johnson
+where \(x_i - x_0 > 0\). The location- and scale-invariant Yeo-Johnson
 transformation is:
@@
 \ref{tab:clean-transformation-appendix-p-value-yj}. As may be observed,
-in these examples robust location and shift-invariant transformations
+in these examples robust location- and shift-invariant transformations
 have higher residual errors because of low weights being assigned to

### 2025-04-11 manuscript_rmarkdown/manuscript.Rmd
Further improvements based on reviewer comments.

@@
 The \textit{optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
+The \textit{loss} column shows the composite loss achieved by each method under optimised parameters.
+This loss is based on residual errors of transformed features and deviations in $\lambda$ parameters
+found compared to the $\lambda$ parameters found prior to inserting outliers.
+The loss metric can be compared between different weighting methods.
+}
 \label{tab:optimal-weighting-parameters-box-cox}
@@
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
-The \textit{optimal} column shows the optimal parameter values. 
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
+The \textit{loss} column shows the composite loss achieved by each method under optimised parameters.
+This loss is based on residual errors of transformed features and deviations in $\lambda$ parameters
+found compared to the $\lambda$ parameters found prior to inserting outliers.
+The loss metric can be compared between different weighting methods.
+}
 \label{tab:optimal-weighting-parameters-yeo-johnson}
@@
 We derived the log-likelihood function to facilitate optimisation using MLE.
-Alternatively, standardisation of a numeric variable 
+Alternatively, standardisation of a numeric feature 
 (e.g., through subtracting its median value and division by its interquartile range) 
@@
 We assessed several weighting methods to achieve robust power transformations.
+Robust power transformations should satisfy two conflicting aims: they should minimise
+residual errors after transformation to central normality, and minimise the overall
+effect of outliers on estimation of transformation parameter $\lambda$.
+These aims are reflected in the composite loss used to optimise weighting parameters.
+For Yeo-Johnson transformation, empirical probabilities with tapered cosine weighting
+resulted in the lowest loss.
+The optimal parameters led to weights that gradually decline towards the tails of empirical probabilities, i.e. elements with low and high values
+receive less or no weight.
+Similarly, for Box-Cox transformation, empirical probabilities with step weighting
+resulted in the lowest loss. In this case the central 80% of the elements are used,
+and the remaining 10% in each tail are ignored.
 Methods that relied on the z-score of the transformed feature or the residual error 
-yielded worse results than the non-robust method.
+yielded worse loss than the non-robust method or those based on empirical probabilities.
 Underperformance of these weighting methods could be explained by their reliance on transformed feature values for setting weights.

### 2025-04-11 manuscript_latex/manuscript.tex
Updated tex file of main manuscript

@@
 %% \usepackage{amsmath,amssymb,amsfonts}%
-%% \usepackage{amsthm}%
+\usepackage{amsthm}%
 \usepackage{mathrsfs}%
@@
 %%%%
-
+\newtheorem*{definition}{Definition}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 \affil*[1]{
-	\orgdiv{National Center for Tumor Diseases Dresden (NCT/UCC)},
-	\orgname{German Cancer Research Center (DKFZ), Faculty of Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, Helmholtz-Zentrum Dresden-Rossendorf (HZDR)},
+	\orgdiv{National Center for Tumor Diseases (NCT), NCT/UCC Dresden },
+	\orgname{a partnership between DKFZ, Faculty of
+	Medicine and University Hospital Carl Gustav Carus, TUD Dresden University of Technology, and
+	Helmholtz-Zentrum Dresden-Rossendorf (HZDR)},
 	\orgaddress{\street{Fetscherstra{\ss}e 74/PF 64}, \city{Dresden}, \postcode{01307}, \country{Germany}}
@@
 normality of features, especially in methods assuming normal
-distributions such as ANOVA and linear discriminant analysis. 
-However, the commonly used Box-Cox and Yeo-Johnson power transformation 
-methods are sensitive to the location, scale, and presence of 
-outliers in the data. Here we present location- and scale-invariant Box-Cox and Yeo-Johnson
-transformations to mitigate these issues. We derive maximum likelihood
-estimation criteria for optimizing transformation parameters and propose
-robust adaptations that reduce the influence of outliers. 
-We also introduce an empirical test for assessing central normality
-of transformed features. In simulations and real-world datasets, robust location- and scale-invariant 
-transformations outperform conventional variants, resulting in better 
-transformations to central normality. In a machine learning experiment
-with 232 datasets with numerical features, integrating robust 
-location- and scale-invariant power transformations into an automated data 
-processing and machine learning pipeline did not result in a meaningful
-improvement or detriment to model performance compared to conventional variants.
-In conclusion, robust location- and scale-invariant power
-transformations can replace conventional variants.}
+distributions such as ANOVA and linear discriminant analysis. However,
+the commonly used Box-Cox and Yeo-Johnson power transformation methods
+are sensitive to the location, scale, and presence of outliers in the
+data. Here we present location- and scale-invariant Box-Cox and
+Yeo-Johnson transformations to mitigate these issues. We derive maximum
+likelihood estimation criteria for optimizing transformation parameters
+and propose robust adaptations that reduce the influence of outliers. We
+also introduce an empirical test for assessing central normality of
+transformed features. In simulations and real-world datasets, robust
+location- and scale-invariant transformations outperform conventional
+variants, resulting in better transformations to central normality. In a
+machine learning experiment with 231 datasets with numerical features,
+integrating robust location- and scale-invariant power transformations
+into an automated data processing and machine learning pipeline did not
+result in a meaningful improvement or detriment to model performance
+compared to conventional variants. In conclusion, robust location- and
+scale-invariant power transformations can replace conventional variants.}
 
@@
 aim to stabilise variance and improve normality of such features
-\citep{Bartlett1947-rx, Tukey1957-rt}. The two most commonly used
-transformations are that of \citet{Box1964-mz} and \citet{Yeo2000-vw}.
-The Box-Cox transformation of a feature value \(x_i\) of feature
-\(\mathbf{X}=\left\{x_1, x_2, \ldots, x_n \right\}\) under the
-transformation parameter \(\lambda\) is defined as:
-
+(Bartlett 1947; John W. Tukey 1957). The two most commonly used
+transformations are that of Box and Cox (1964) and Yeo and Johnson
+(2000).
+
+The Box-Cox transformation is defined as:
+
+\begin{definition}[Box-Cox power transformation]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a finite sequence of length $n > 0$.
+Let $\lambda \in \mathcal{R}$ be a transformation parameter.
+Then the Box-Cox power transformation of element $x_i$ is defined as:
 \begin{equation}
@@
 approach is sensitive to outliers, and robust versions of Box-Cox and
-Yeo-Johnson transformations were devised \citep{Raymaekers2024-zf}.
+Yeo-Johnson transformations were devised (Raymaekers and Rousseeuw
+2024).
 
@@
 
-Statistical tests for normality, such as the Shapiro-Wilk test
-\citep{Shapiro1965-zd}, could be automatically applied to transformed features.
+Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
+and Wilk 1965), could be automatically applied to transformed features.
 However, given sufficiently large sample sizes such tests can detect
@@
 
+First, we define a (numeric) feature. Throughout this work, whenever
+reference is made to a feature, a numeric feature is meant unless noted
+otherwise.
+
+\begin{definition}[Numeric feature]
+A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
+\end{definition}
+
 \subsection{Location- and scale-invariant power
@@
 i.e.~using maximum likelihood estimation (MLE). A full derivation of the
-log-likelihood function for both transformations is shown in \ref{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}.
+log-likelihood function for both transformations is shown in Appendix A.
 The location- and scale-invariant Box-Cox log-likelihood function is:
 
+\begin{definition} [Log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Box-Cox power transformation is defined as:
 \begin{equation}
@@
 \begin{split}
-\mathcal{L}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{BC}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
@@
 
+\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+
+Then the log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is defined as:
 \begin{equation}
@@
 \begin{split}
-\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
-& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+& - n \log s + (\lambda - 1) \sum_{i=1}^n \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
 \end{split}
 \end{equation}
-
-where \(\mu\) and \(\sigma^2\) are the mean and variance of the
-Yeo-Johnson transformed feature
-\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\), respectively.
+\end{definition}
 
@@
 transformations to normality, as shown in Figure
-\ref{fig:decreased-normality}. As indicated by \citet{Raymaekers2024-zf},
-the general aim of power transformations should be to transform
+\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
+(2024), the general aim of power transformations should be to transform
 non-outlier data to normality, i.e.~achieve \emph{central normality}. To
@@
 process requires identifying outliers in the data and weighting such
-instances during the optimisation process. \citet{Raymaekers2024-zf}
-achieve this through weighted maximum likelihood estimation.
+instances during the optimisation process. Raymaekers and Rousseeuw
+(2024) achieve this through weighted maximum likelihood estimation.
 However, because this procedure iteratively estimates and updates
@@
 \ref{eqn:box-cox-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to \citet{Raymaekers2024-zf},
-these log-likelihood functions includes additional
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
+and Rousseeuw (2024), these log-likelihood functions include additional
 terms to accommodate estimation of \(x_0\) and \(s\). The weighted
@@
 
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation is:
+
 \begin{equation}
@@
 \begin{split}
-\mathcal{L}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
@@
 
+\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
+
+Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
+Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
+Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
+\end{equation*}
+
+Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
+\begin{equation*}
+\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
+\end{equation*}
+
+Then, the weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is:
+
 \begin{equation}
@@
 \begin{split}
-\mathcal{L}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
-& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
+\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
 \end{split}
 \end{equation}
-
-where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
-variance of the Yeo-Johnson transformed feature
-\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):
-
-\begin{equation}
-\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
-\end{equation}
+\end{definition}
 
@@
   A step function, with \(\delta_1 \geq 0\) as threshold parameter:
-\end{itemize}
-
-\begin{equation}
+  \begin{equation*}
   w_i =
@@
   \end{cases}
-\end{equation}
-
-\begin{itemize}
+  \end{equation*}
 \item
@@
   \end{cases}
-\end{equation}
-
-\begin{itemize}
+  \end{equation*}
 \item
-  A tapered cosine function \citep{Tukey1967-eb}, with \(\delta_1 \geq 0\)
+  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
   and \(\delta_2 \geq \delta_1\) as threshold parameters:
-\end{itemize}
-
-\begin{equation}
+  \begin{equation*}
   w_i =
@@
 \caption{Weighting functions investigated in this study to make power transformations more robust against outliers.
-In this example, the step function was parametrised with $\delta_1 = 0.60$.
-The triangle and tapered cosine functions were both parametrised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
+In this example, the step function was parameterised with $\delta_1 = 0.60$.
+The triangle and tapered cosine functions were both parameterised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
 \end{figure}
@@
   The weighting function uses the z-score of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After \citep{Raymaekers2024-zf},
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
+  Rousseeuw (2024),
   \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
@@
   and scale of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) \citep{Huber1981-su}. Then
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981). Then
   \(\dot{x}_i = z_i\).
@@
 for use with unimodal distributions, the generalised normal distribution
-\citep{Subbotin1923-qk, Nadarajah2005-xe} is a suitable option for simulating
+(Subbotin 1923; Nadarajah 2005) is a suitable option for simulating
 realistic feature distributions. This distribution has the following
@@
 recipe for introducing skewness into the otherwise symmetric generalised
-normal distribution \citep{Gijbels2019-te}, leading to
+normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
 the following probability density function:
 
+\begin{definition}[Asymmetric generalised normal distribution probability density function]
+Let $\alpha \in (0,1)$ be a skewness parameter.
+Let $\mu \in \mathcal{R}$ be a location parameter.
+Let $\sigma > 0$ be a shape parameter.
+
+Then the probability density function of the asymmetric generalised normal distribution is:
 \begin{equation}
@@
 \end{equation}
-
-The quantile function for the asymmetric generalised normal distribution
-\(F_{\alpha}^{-1}\) thus incorporates the quantile function
-\(F_{\beta}^{-1}\) of the symmetric generalised normal distribution.
-\(F_{\beta}^{-1}\) was derived by Griffin to be \citep{Griffin2018-bf}:
-
-\begin{equation}
-F_{\beta}^{-1}(p) = \sgn\left(p - 0.5 \right) F_{\Gamma}^{-1}\left(2 \left|p - 0.5 \right|; 1 / \beta \right)
-\end{equation}
-
-Here, \(F_{\Gamma}^{-1}\) is the quantile function of the gamma
-distribution with shape \(1 / \beta\), which can be numerically
-approximated.
-
-\subsection{Empirical central normality
-test}\label{empirical-central-normality-test}
+\end{definition}
+
+\subsection{Central normality test}\label{central-normality-test}
 
@@
 Deviations from normality can be detected by normality tests, such as
-the Shapiro-Wilk test \citep{Shapiro1965-zd}. In practice, normality
+the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
 tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop an empirical test for central normality. The null
-hypothesis \(H_0\) is that the distribution is centrally normal. The
-alternative hypothesis \(H_1\) is that the distribution is not centrally
-normal.
-
-Let central normality be defined as the normality of central portion
-\(\kappa\) of the data,
-i.e.~\(\mathbf{X}_{\text{central}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\),
-with \(p_i\) probabilities of the empirical distribution, as previously.
-We then compute the residual errors between the z-scores of the
-transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\) and the
-expected z-scores from a standard normal distribution:
-\(r_i =\left| \left( \phi^{\lambda, x_0, s}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|\),
-with \(\mu_M\) and \(\sigma_M\) robust Huber M-estimates of location and
-scale of the transformed feature \(\phi^{\lambda, x_0, s} (\mathbf{X})\)
-\citep{Huber1981-su}. The set of residual errors for the central portion of the
-data is then
-\(\mathbf{R}_{\text{central}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}\).
-
-The test statistic \(\tau_{\text{ecn}}\) is then defined as:
-
-\begin{equation}
-\tau_{\text{ecn}} = \frac{\sum_{i=1}^{N} r_i \left[r_i \in \mathbf{R}_{\text{central}}\right]}{\sum_{i=1}^N \left[r_i \in \mathbf{R}_{\text{central}}\right]} 
-\end{equation}
-
-Here \([\quad]\) denotes an Iverson bracket. The test statistic is equal
-to the mean of the residual errors of the central portion of the data.
-
-To use the test statistic, the central portion of the data needs to be
-defined and the Type 1 error rates determined. We will do so in
-section \ref{simulation}.
+Here we develop a test for central normality.
+
+\begin{definition}[Central portion of a sequence]
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
+Let $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$ be the percentile value
+corresponding to each element in $\mathbf{X}$. For elements with tied values in $\mathbf{X}$,
+percentile values are replaced by the average in their group, i.e.
+if $x_j = x_{j+1} = \ldots = x_{j+m}$, then 
+$p_j' = p_{j+1}' = \ldots = p_{j+m}' = 1/(m+1)\sum_j^{j+m}p_j$.
+Furthermore, let central portion $\kappa \in (0,1)$.
+
+Then the central portion of $\mathbf{X}$ is 
+$\mathbf{X}_{\text{c}} = \left\{x_i \in \mathbf{X} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Residual errors of the central portion of a sequence]
+
+Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+
+Then, the set of residual errors of the central portion of $\mathbf{X}$ (i.e. $\mathbf{X}_{\text{c}}$) is
+$\mathbf{R}_{\text{c}} = \left\{ r_i \in \left\{ r_1, r_2, \ldots, r_n\right\} \, | \,  \frac{1-\kappa}{2} \leq  p_i \leq \frac{1 + \kappa}{2}\right\}$.
+
+\end{definition}
+
+\begin{definition}[Central normality of a sequence]
+
+The central portion of a sequence is normally distributed if the sum of residual
+errors is equal to zero:
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+
+\end{definition}
+
+In practice, a finite sequence sampled from a normal distribution
+\(\mathcal{N}(\mu, \sigma)\) will have a non-zero sum of residual
+errors.
+
+\begin{definition}[Central normality test]
+The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
+
+The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+\end{definition}
+
+The null-hypothesis \(\mathcal{H}_0\) is that the central portion of
+sequence \(\mathbf{X}\) is normally distributed, with the alternative
+hypothesis \(\mathcal{H}_1\) that it is not normally distributed. To
+test this hypothesis, a test statistic is computed and compared against
+a critical value.
+
+\begin{definition}[Central normality test statistic]
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+
+The test statistic is then $\tau_{n, \kappa} = \frac{1}{m}\sum_{r_j \in \mathbf{R}_{\text{c}}}r_j$.
+
+\end{definition}
+
+The null-hypothesis should be rejected at significance level \(\alpha\)
+if \(\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}\).
+
+The central portion of the data needs to be defined and the Type 1 error
+rates determined to provide critical test statistics. We will do so in
+the \hyperref[simulation]{Simulation} section.
 
@@
 and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
-algorithm for derivative-free bound constraint optimisation \citep{Powell2009-zb}
-through maximum likelihood estimation. The required algorithms
+algorithm for derivative-free bound constraint optimisation (Powell
+2009) through maximum likelihood estimation. The required algorithms
 were implemented in the \texttt{power.transform} R software package
-(version 1.0.0) \citep{Zwanenburg2024-kq}. Of note, the
+(Zwanenburg and Löck 2024b) (version 1.0.1). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 
+\subsection{Central normality and empirical central normality
+test}\label{central-normality-and-empirical-central-normality-test}
+
+To develop a test for central normality we need to consider two
+parameters: the central portion \(\kappa\) as a fixed parameter, and
+test statistic \(\tau_{\text{ecn}}\). We will first define the central
+portion \(\kappa\).
+
+For each number of samples
+\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}\)
+we randomly drew \(m_d = 30000\) ordered sequences \(\mathbf{X}\) from
+\(\mathcal{N}(0,1)\). This dataset was used to compute the critical test
+statistic values for the central normality test. Additionally, we added
+\(10 \%\) outliers by randomly replacing elements of each ordered
+sequence. The dataset with outliers was used to compute the critical
+test statistic values for the empirical variant of the central normality
+test. For each sequence we then computed \(\tau_{n, \kappa}\) for
+\(\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}\).
+
+Figure \ref{fig:empirical-central-normality-kappa} shows
+\(\tau_{\alpha = 0.05, n, \kappa}\) as a function of \(n\) for different
+values of \(\kappa\). With decreasing \(\kappa\), the test statistic
+curve decreases. The curves for \(\kappa = 0.60\), \(\kappa = 0.70\) and
+\(\kappa = 0.80\) are similar, whereas curves for \(\kappa \geq 0.90\)
+are affected by outliers, as can be observed by comparing curves of the
+central normality test with those of the empirical variant.
+
+Since the (empirical) central normality tests assesses whether the
+central part of a sequence is normally distributed, we used
+\(\kappa = 0.80\). Critical statistic values for central normality and
+empirical central normality tests are shown are shown in Table
+\ref{tab:central-normality-critical-statistic} and Table
+\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figure_4}
+
+}
+
+\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the (empirical) central normality test as function of $n$ for several values of central portion $\kappa$.
+The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
+with fully normal data where 10\% of elements are replaced by outliers.
+emp: empirical}\label{fig:empirical-central-normality-kappa}
+\end{figure}
+
+To assess type I error rates for the (empirical) central normality test
+and compare these to the Shapiro-Wilk test for normality, we randomly
+drew another 1000 sequences from a standard normal distribution
+(\(\mathcal{N}(0,1)\) for each
+\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
+A second dataset was created by replacing elements with randomly drawn
+outliers, as described above. For each sequence, the p-value for the
+respective test was used to reject the null hypothesis that the sequence
+is (centrally) normal. The null hypothesis was rejected if
+\(p \leq 0.05\). The type I error rate was determined by computing the
+fraction of centrally normal sequences rejected this way. Figure
+\ref{fig:empirical-central-normality-error_rate} shows that the
+empirical central normality test had the expected type I error rate for
+centrally normal data (with tails of the distribution not being normally
+distributed), whereas the central normality test and Shapiro-Wilk test
+overestimated type I error rates.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figure_5}
+
+}
+
+\caption{Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers.
+The empirical central normality (ECN) test uses $\kappa = 0.80$.
+CN: central normality}\label{fig:empirical-central-normality-error_rate}
+\end{figure}
+
+In Figure \ref{fig:empirical-central-normality-examples} we apply the
+empirical central normality test to assess central normality of features
+that are composed of a mixture of samples drawn from two normal
+distributions (\(n = 100\) each). With increased separation of the
+underlying normal distributions, the probability of the feature being
+centrally normal decreases, as expected.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figure_6}
+
+}
+
+\caption{Bi-modal distributions and empirical central normality test results.
+The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) drawn from normal distributions that are offset by a distance $d$.
+We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
+As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
+Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
+\end{figure}
+
 \subsection{Robust transformations}\label{robust-transformations}
@@
 To determine the weighting function parameters for each of the nine
-combinations, \(m_d=100\) asymmetric generalised normal distributions
-were drawn. Each distribution was parametrised with a randomly chosen
-skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
-parameter \(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale
-parameters were set as \(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) instances were then randomly drawn, with
-\(\gamma \sim U\left(2, 4\right)\), i.e., between \(100\) and \(10000\)
-values are drawn to create \(\mathbf{X}_i\).
-
-Outlier values were then drawn to randomly replace a fraction of the
-values of \(\mathbf{X}_i\). This was repeated \(m_{\text{out}} = 10\)
-times, with outlier fractions regularly spaced in \([0.00, 0.10]\). Thus
-up to 10 percent of the values could be replaced by outliers. Outlier
-values were set according to \citet{Tukey1977-xm}, as follows. Let
-\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
-was:
+combinations, \(m_d=500\) sequences \(\mathbf{X}_i\)
+(\(i \in \{1, 2, \ldots, m_d\}\)) were randomly drawn from randomly
+parametrised asymmetric generalised normal distributions. Each
+distribution was parametrised with a random skewness parameter
+\(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
+were set as \(\mu = 0\) and \(\sigma = 1\), respectively. To form each
+sequence \(\mathbf{X}_i\), \(n = \lceil 10^\gamma \rceil\) instances
+were then randomly drawn, with
+\(\gamma \sim U\left(\log_{10}50, 3\right)\), resulting in a set of
+sequences with between \(50\) and \(1000\) elements each.
+
+Outlier values were then drawn to randomly replace 10 percent of the
+elements of \(\mathbf{X}_i\). Outlier values were set according to John
+W. Tukey (1977), as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
+Then the corresponding outlier value was:
 
@@
 quartile and interquartile range of \(\mathbf{X}_i\), respectively.
-Outlier values randomly replaced values in \(\mathbf{X}_i\) to create
-\(\mathbf{X}_{i,j}\), with
-\(j \in \left\{ 1, \ldots, m_{\text{out}} \right\}\)
+Outlier values randomly replaced elements in \(\mathbf{X}_i\).
 
 To find the optimal values for the weighting function parameters
-\(\delta_1\) and \(\delta_2\) (if applicable), we minimised the absolute
-difference between the \(\lambda_{r}\) parameter obtained for robust
-transformation in the presence of outliers, and the \(\lambda_0\)
-parameter obtained using the non-robust transformations in absence of
-outliers:
-
-\begin{equation}
-\label{eqn:minimisation-weighting}
-\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
-\end{equation}
+\(\delta_1\) and \(\delta_2\) (if applicable), we minimised a composite
+loss
+\(L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}\).
+The composite loss consisted of two components: a loss term
+\(L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}\), i.e.~the mean of
+residual errors of the central 80\% of elements of each sequence
+\(\mathbf{X}_i\), which aimed at optimising central normality; and a
+loss term
+\(L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)\),
+with \(\lambda_{0,i}\) and \(\lambda_{i}\) transformation parameters
+found for sequence \(\mathbf{X}_i\) prior to and after adding outliers,
+respectively, and tolerance parameter \(\xi = 0.5\) for Box-Cox and
+\(\xi = 0.3\) for Yeo-Johnson power transformations. This second term
+aimed to prevent solutions that provide small improvements in central
+normality at the cost of a poor fit of the tails of sequences.
 
 Minimisation was conducted using the BOBYQA algorithm for
-derivative-free bound constraint optimisation \citep{Powell2009-zb}. The
+derivative-free bound constraint optimisation (Powell 2009). The
 resulting weighting function parameters for weighted MLE are shown in
@@
 \begin{table}
+\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
-The {optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+The \textit{optimal} column shows the optimal parameter values.
+The \textit{loss} column shows the composite loss achieved by each method under optimised parameters.
+This loss is based on residual errors of transformed features and deviations in $\lambda$ parameters
+found compared to the $\lambda$ parameters found prior to inserting outliers.
+The loss metric can be compared between different weighting methods.
+}
 \label{tab:optimal-weighting-parameters-box-cox}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
-\botrule
-
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 49.6 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.80 & ---  & ---       & ---  & 38.5 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.01 & 1.00 & $(0, 1]$  & 0.86 & 43.7 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 1.00 & $(0, 1]$  & 0.90 & 42.7 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.39 & ---  & ---       & ---  & 52.2 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 2.39 & 2.40 & $(0, 10]$ & 4.92 & 52.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.07 & 3.63 & $(0, 10]$ & 3.43 & 57.4 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.83 & ---  & ---       & ---  & 53.1 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.80 & 0.81 & $(0, 10]$ & 0.84 & 54.4 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.76 & 0.77 & $(0, 10]$ & 0.76 & 53.2 \\
+\bottomrule
 \end{tabular}
+\end{center}
 \end{table}
@@
 \begin{table}
+\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
-$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution.
+$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
+and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
 The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
-The {optimal} column shows the optimal parameter values.
-The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
-Lower loss indicates better robustness against outliers.}
+The \textit{loss} column shows the composite loss achieved by each method under optimised parameters.
+This loss is based on residual errors of transformed features and deviations in $\lambda$ parameters
+found compared to the $\lambda$ parameters found prior to inserting outliers.
+The loss metric can be compared between different weighting methods.
+}
 \label{tab:optimal-weighting-parameters-yeo-johnson}
@@
 \midrule
-non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
-$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
-$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
-$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
-$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
-$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
-$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
-$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
-$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
-$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
+non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 42.1 \\
+$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.78 & ---  & ---       & ---  & 35.1 \\
+$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.20 & 0.95 & $(0, 1]$  & 1.00 & 34.0 \\
+$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.54 & 0.95 & $(0, 1]$  & 1.00 & 32.6 \\
+$z$ (step)               & 1.28 & $(0, 10]$ & 2.32 & ---  & ---       & ---  & 43.0 \\
+$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.28 & 1.96 & $(0, 10]$ & 3.43 & 49.7 \\
+$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.41 & 1.96 & $(0, 10]$ & 3.75 & 51.3 \\
+$r$ (step)               & 0.50 & $(0, 10]$ & 0.92 & ---  & ---       & ---  & 65.4 \\
+$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.87 & 1.00 & $(0, 10]$ & 0.87 & 66.0 \\
+$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.06 & 1.00 & $(0, 10]$ & 1.07 & 67.1 \\
 \bottomrule
 \end{tabular}
+\end{center}
 \end{table}
 
-Figure \ref{fig:optimised-weighting-function-parameters} shows the
-distribution of errors \(\left| \lambda_r - \lambda_0 \right|\) for
-non-robust and robust transformations using the optimal weighting
-function parameters. In the presence of outliers, the Yeo-Johnson
-transformation yielded smaller errors than the Box-Cox transformation.
-For both transformations, weighting based on empirical probabilities
-yielded the most consistent \(\lambda\) parameter estimates in the
-presence of outliers. Other methods did not outperform the non-robust
-transformation method.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_4}
-
+\subsection{Assessing transformations using simulated
+data}\label{assessing-transformations-using-simulated-data}
+
+Three datasets with 10000 sequences each were created to assess
+transformation to normality. For each sequence \(\mathbf{X}_i\),
+\(n = \lceil 10^\gamma \rceil\) elements were randomly drawn, with
+\(\gamma \sim U\left(1, 4\right)\), resulting in sequences with between
+\(10\) and \(10000\) elements.
+
+\begin{itemize}
+\item
+  A \textit{clean} dataset with each sequence drawn from
+  \(\mathcal{N}(0, 1)\) and transformed using an inverse power
+  transformation with randomly drawn transformation parameter
+  \(\lambda \sim U\left(0.00, 2.00 \right)\).
+  \(\left(\phi_{\text{BC}}^\lambda\right)^{-1}\) and
+  \(\left(\phi_{\text{YJ}}^\lambda\right)^{-1}\) were used as inverse
+  transformations for assessing Box-Cox and Yeo-Johnson transformations,
+  respectively. Prior to inverse transformation, sequences for Box-Cox
+  transformations were shifted into the positive domain, with minimum
+  value \(1\).
+\item
+  A \textit{dirty} dataset with each sequence drawn from
+  \(\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)\),
+  with randomly drawn skewness parameter
+  \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+  \(\beta \sim U\left(1.00, 5.00 \right)\).
+\item
+  A \textit{shifted} dataset with each sequence drawn from
+  \(\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)\),
+  with randomly drawn skewness parameter
+  \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
+  \(\beta \sim U\left(1.00, 5.00 \right)\).
+\end{itemize}
+
+A dataset with outliers was created for each of the above datasets by
+replacing 10 percent of elements in each sequence, as described earlier.
+
+In addition to no power transformation and location- and scale-invariant
+power transformations, conventional and Raymaekers and Rousseeuw's
+robust adaptation (Raymaekers and Rousseeuw 2024) were assessed. For the
+latter two methods, normalisation before standardisation using was
+additionally assessed using the following two methods:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  z-standardisation:
+  \(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
+  and \(\sigma\) the mean and standard deviation of sequence
+  \(\mathbf{X}\).
+\item
+  robust scaling:
+  \(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
+  with \(\text{IQR}\) representing the interquartile range of
+  \(\mathbf{X}\).
+\end{enumerate}
+
+This results in nine types of power transformation. Transformation
+parameters were optimised for each sequence. Subsequently the sum of
+residual errors and sum of residual errors of the central portion
+(\(\kappa = 0.80\)) were computed for each sequence after
+transformation. Then, each method was ranked according to the sum of
+residual errors (data without outliers) or the sum of residual errors of
+the central portion (data with outliers). The average rank of each
+transformation method was computed over all 10000 sequences in each
+dataset. Average ranks for Yeo-Johnson transformations are shown in
+Table \ref{tab:comparison_methods_simulation_yeo_johnson}. Location and
+shift-invariant Yeo-Johnson transformation ranked best for all datasets
+without outliers. The robust variant ranked best in dirty and shifted
+datasets with outliers, but not in the clean dataset. Results for
+Box-Cox transformations are shown in Table
+\ref{tab:comparison_methods_simulations_box_cox}.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
 }
-
-\caption{Robustness of power transformations after optimising weighting function parameters.
-The distribution of errors, i.e.
-the difference between robustly fitted $\lambda^r$ and expected $\lambda_{0}$ found in the absence of outliers, is shown for 1000 randomly parametrised asymmetric generalised normal distributions with up to 10\% outliers on either side of the distribution.
-The top and bottom panel show errors for the location- and scale-invariant Box-Cox and Yeo-Johnson transformations, respectively.
-In each panel, the error distribution for the non-robust transformation is shown to the left for comparison with different weighting methods.
-Note that a square root transformation was applied along the $y$-axis for display purposes.
-Median errors are shown above each distribution.}\label{fig:optimised-weighting-function-parameters}
-\end{figure}
-
-\subsection{Empirical central normality
-test}\label{empirical-central-normality-test-1}
-
-To develop an empirical test for central normality we need to consider
-two parameters: the central portion \(\kappa\) as a fixed parameter, and
-test statistic \(\tau_{\text{ecn}}\). We will first define the central
-portion \(\kappa\).
-
-First we draw \(m_d=10000\) random asymmetric generalised normal
-distributions. As before, each distribution is parametrised with a
-randomly chosen skewness parameter
-\(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-\(\beta \sim U\left(1.00, 5.00 \right)\). Location and scale parameters
-are set as \(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) values are then randomly drawn, with
-\(\gamma \sim U\left(1.47, 3.00\right)\), which leads to between \(30\)
-and \(1000\) values being drawn to create \(\mathbf{X}_i\). As before,
-for each distribution, up to \(10 \%\) of instances are replaced by
-outlier values, and this is repeated \(10\) times. We then compute
-residuals after optimising robust location- and scale-invariant
-transformations with the empirical tapered cosine weighting method.
-
-Figure \ref{fig:empirical-central-normality-residual-error} shows the
-residual errors of the transformed distribution as function of the
-empirical probability. As expected, the largest deviations from
-normality appear on the extremities of this range. For Box-Cox
-transformations, the effect of outliers on the lower end of the
-distribution leads to noticeable asymmetry. Between very low and very
-high empirical probabilities, residual errors are constrained and
-relatively flat. This is the candidate range for the central portion of
-the data.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_5}
-
-}
-
-\caption{Residual error as a function of empirical probability.
-Percentiles of the error are shown for robust, location- and scale-invariant transformations of 10000 randomly drawn asymmetric generalised normal distributions, for each of which outliers were randomly drawn 10 times (up to 10\% of samples).
-Larger errors occur at the edges of each distribution.}\label{fig:empirical-central-normality-residual-error}
-\end{figure}
-
-We then consider the empirical probability of type I errors: the
-probability of incorrectly classifying a centrally normal distribution
-as being non-normal based on the test statistic \(\tau_{\text{ecn}}\).
-Under the assumption that the asymmetric generalised normal
-distributions are centrally normal after robust transformation, this
-produces the relationship shown in Figure
-\ref{fig:empirical-central-normality-type-1-error-rate}. Since for
-\(\kappa \leq 0.80\) error curves for both transformation methods are
-similar, we fix the value of the central portion \(\kappa\) to 0.80. In
-the remaining we will use the empirical central normality test statistic
-values defined using robust location- and scale-invariant Yeo-Johnson
-transformations, as it is more conservative (see \ref{appendix-e-empirical-central-normality-test}). This leads
-to test statistic values listed in Table
-\ref{tab:empirical-central-normality}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_6}
-
-}
-
-\caption{Type 1 error rate of transformed asymmetric generalised normal distributions as function of the test statistic $\tau_{\text{ecn}}$ for the central portion $\kappa$ of the distribution.}\label{fig:empirical-central-normality-type-1-error-rate}
-\end{figure}
-
-\begin{table}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
-\label{tab:empirical-central-normality}
-\begin{tabular}{l | c c c c c c c}
+\label{tab:comparison_methods_simulations_yeo_johnson}
+\begin{tabular}{l | l r r r r r r}
 
 \toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
 
 \midrule
-$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
+
+none                                  & &         8.51  &         7.60  &         7.33  &         6.15  &         7.44  &         6.60 \\
+conventional                          & &         3.70  &         5.70  &         3.87  &         7.06  &         6.49  &         5.93 \\
+conventional (z-score norm.)          & &         4.72  &         5.68  &         6.38  &         5.01  &         5.43  &         4.87 \\
+conventional (robust scaling)         & &         4.06  &         6.77  &         5.37  &         5.75  &         4.41  &         5.63 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.20  &         4.01  &         4.13  &         6.27  &         5.65 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.27  & \textbf{2.66} &         6.23  &         3.59  &         5.28  &         3.48 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.95  &         5.32  &         3.77  &         4.36  &         3.66 \\
+invariant                             & & \textbf{3.13} &         6.77  & \textbf{3.22} &         6.12  & \textbf{2.64} &         6.03 \\
+robust invariant                      & &         6.41  &         3.66  &         3.27  & \textbf{3.41} &         2.68  & \textbf{3.10} \\
+
 \bottomrule
 \end{tabular}
+\end{center}
 \end{table}
 
-In Figure \ref{fig:empirical-central-normality-examples} we apply the
-empirical central normality test to assess central normality of features
-that are composed of a mixture of samples drawn from two normal
-distributions. With increased separation of the underlying normal
-distributions, the probability of the feature being centrally normal
-decreases, as expected.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_7}
-
-}
-
-\caption{Bi-modal distributions and empirical central normality test results.
-The feature (black) is a mixture of two identical sample sets (blue and orange) drawn from normal distributions that are offset by a distance $d$.
-We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
-As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
-Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
-\end{figure}
-
 \section{Experimental Results}\label{experimental-results}
 
-\subsection{Invariance}\label{sec:invariance}
+\subsection{Invariance}\label{invariance}
 
@@
 shifts in location, distributions that due to location and scale are not
-centred near zero, or both. Earlier, we assessed these transformations
+centered near zero, or both. Earlier, we assessed these transformations
 using simulated data. In the following, they are evaluated using
@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in 
-\ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.
+for Box-Cox transformations are shown in Appendix D.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_8}
+{\centering \includegraphics[width=1\linewidth]{figure_7}
 
@@
 \subsubsection{Age of patients with lung
-cancer}\label{sec:age-of-patients-with-lung-cancer}
+cancer}\label{age-of-patients-with-lung-cancer}
 
@@
 228 patients with lung cancer that was collected and published by
-Loprinzi et al. \citep{Loprinzi1994-cd}. The age in the cohort was
-\(62.4 \pm 9.1\) (mean \(\pm\) standard deviation) years. Applying
+Loprinzi et al. (Loprinzi et al. 1994). The age in the cohort was
+\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
 conventional and invariant Yeo-Johnson transformations to patient age
 yielded the following results, see Figure
-\ref{fig:experimental-results-invariance}: no transformation (sum of
-residuals with normal distribution \(\sum r_i = 16.5\)); conventional
-transformation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
-\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\));
-Raymaekers and Rousseeuw's robust adaptation (\(\lambda = 2.0\),
-\(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
-\(\sigma_{YJ} = 0.5 \cdot 10^3\)); location- and scale-invariant
-transformation (\(\lambda = 0.9\), \(\sum r_i = 8.8\),
-\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.1\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.8\), \(\sum r_i = 10.6\),
-\(\mu_{YJ} = 1.2\), \(\sigma_{YJ} = 1.0\)).
+\ref{fig:experimental-results-invariance}: no transformation
+(\(\sum r_i = 16.5\), \(p = 0.69\)); conventional transformation
+(\(\lambda = 2.0\), \(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
+\(\sigma_{YJ} = 0.5 \cdot 10^3\), \(p = 0.96\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
+\(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\),
+\(p = 0.96\)); location- and scale-invariant transformation
+(\(\lambda = 1.3\), \(\sum r_i = 8.8\), \(\mu_{YJ} = -1.2\),
+\(\sigma_{YJ} = 1.1\), \(p = 0.98\)); and robust location- and
+scale-invariant transformation (\(\lambda = 1.3\), \(\sum r_i = 9.3\),
+\(\mu_{YJ} = -1.0\), \(\sigma_{YJ} = 1.1\), \(p = 0.93\)).
 
 Location- and scale-invariant transformation led to a lower overall
-residual sum, indicating a better transformation. Robust location- and
-scale-invariant transformation had a higher residual sum compared to the
-non-robust variant, which may be due to the lack of outliers in the
-data. Conventional transformations inflated the mean \(\mu_{YJ}\) and
-standard deviation \(\sigma_{YJ}\) of the age feature after
-transformation. The empirical central normality test did not detect any
-statistically significant deviations from central normality for any
-transformation (all \(p \geq 0.78\)).
-
-\subsubsection{Penguin body mass}\label{sec:penguin-body-mass}
+residual error, indicating a better transformation to normality.
+Conventional transformations inflated the mean \(\mu_{YJ}\) and standard
+deviation \(\sigma_{YJ}\) of the age feature after transformation. The
+empirical central normality test did not detect any statistically
+significant deviations from central normality for any transformation
+(all \(p \geq 0.93\)).
+
+\subsubsection{Penguin body mass}\label{penguin-body-mass}
 
 Gorman, Williams and Fraser recorded body mass (in grams) of 342
-penguins of three different species \citep{Gorman2014-eo}.
-The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean \(\pm\) standard
-deviation) grams, and not centrally normal (\(p = 0.03\)). Applying
+penguins of three different species (Gorman, Williams, and Fraser 2014).
+The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
+deviation) grams, and not centrally normal (\(p < 0.001\)). Applying
 conventional and invariant Yeo-Johnson transformations to body mass
@@
 \ref{fig:experimental-results-invariance}: no transformation (residual
-sum \(\sum r_i = 48.0\)); conventional transformation
+sum \(\sum r_i = 48.0\), \(p < 0.001\)); conventional transformation
 (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\)); location- and scale-invariant
-transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
-\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.4\), \(\sum r_i = 23.1\),
-\(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.8\)).
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); Raymaekers and
+Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\));
+location- and scale-invariant transformation (\(\lambda = 0.5\),
+\(\sum r_i = 26.8\), \(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\),
+\(p = 0.28\)); and robust location- and scale-invariant transformation
+(\(\lambda = 0.3\), \(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\),
+\(\sigma_{YJ} = 0.9\), \(p = 0.69\)).
 
 Location- and scale-invariant transformation produced a lower overall
-residual sum, indicating a better transformation. Moreover, conventional
-transformations led to low standard deviation \(\sigma_{YJ}\) of the
-body mass feature after transformation. The empirical central normality
-test did not detect any statistically significant deviations from
-central normality for any transformation (all \(p \geq 0.38\)).
+residual errors, indicating a better transformation. Moreover,
+conventional transformations led to low standard deviation
+\(\sigma_{YJ}\) of the body mass feature after transformation. The
+empirical central normality test did not detect any statistically
+significant deviations from central normality for any transformation
+(all \(p \geq 0.10\)).
 
 \subsubsection{Latitude in the Ames housing
-dataset}\label{sec:latitude-in-the-ames-housing-dataset}
+dataset}\label{latitude-in-the-ames-housing-dataset}
 
@@
 dataset contains data on 2930 properties that were sold between 2006 and
-2010 \citep{De-Cock2011-jf} including their geospatial coordinates. The
-latitude was \(42.03 \pm 0.02)\) (mean \(\pm\) standard deviation). Applying
+2010 (De Cock 2011), including their geospatial coordinates. The
+latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
 conventional and invariant Yeo-Johnson transformations to latitude
@@
 \subsection{Robustness against
-outliers}\label{sec:robustness-against-outliers}
+outliers}\label{robustness-against-outliers}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_9}
+{\centering \includegraphics[width=1\linewidth]{figure_8}
 
@@
 \subsubsection{Fuel efficiency in the Top Gear
-dataset}\label{sec:fuel-efficiency-in-the-top-gear-dataset}
+dataset}\label{fuel-efficiency-in-the-top-gear-dataset}
 
 The Top Gear dataset contains data on 297 vehicles that appeared on the
-BBC television show \emph{Top Gear} \citep{Alfons2021-kc}. Within this dataset,
+BBC television show \emph{Top Gear} (Alfons 2021). Within this dataset,
 the fuel consumption feature contains outliers due to highly
@@
 to a centrally normal distribution (empirical central normality test
-\(p = 0.01\) and \(p=0.03\) for conventional and invariant
-transformations, respectively). Robust transformations produce
-distributions that are centrally normal (empirical central normality
-test \(p > 0.05\)).
+\(p < 0.001\) for conventional and invariant transformations). Robust
+transformations produce distributions that are centrally normal
+(empirical central normality test \(p > 0.05\)).
 
 \subsubsection{Maximum arterial wall thickness in an ischemic stroke
-dataset}\label{sec:maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
+dataset}\label{maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
 
 The ischemic stroke dataset contains historic data from 126 patients
-with risk at ischemic stroke \citep{Kuhn2019-kt}. These patients
+with risk at ischemic stroke (Kuhn and Johnson 2019). These patients
 underwent Computed Tomography Angiography to characterize the carotid
@@
 
-We used 285 datasets from the Penn Machine Learning Benchmarks
-collection \citep{Romano2022-gq}. In this collection, 122 datasets
-correspond to regression tasks and 163 datasets to classification tasks.
-Using the familiar auto-machine learning library (version 1.5.0) \citep{Zwanenburg2021-so},
-each dataset was used to train a model for each
-of 16 process configurations. Each process configuration specifies the
-learner (generalised linear model or random forest), transformation
-method (none, conventional Yeo-Johnson, robust invariant Yeo-Johnson,
-robust invariant Yeo-Johnson with empirical central normality test
-(rejecting transformations with \(p \leq 0.01\)), and normalisation
-method (none, \(z\)-standardisation), yielding 16 distinct
-configurations. Before each experiment, each dataset was randomly split
-into a training (70\%) and holdout test (30\%) set five times. Thus, a
-total of 22800 models were created. Each model was then evaluated using
-the holdout test set using one of two metrics, i.e.~the root relative
-squared error (RRSE) for regression tasks and the area under the
-receiver operating characteristic curve (AUC) for classification tasks.
+We used 231 datasets containing at least one numeric feature from the
+Penn Machine Learning Benchmarks collection (Romano et al. 2022). In
+this collection, 114 datasets correspond to regression tasks and 117
+datasets to classification tasks. Using the familiar auto-machine
+learning library (Zwanenburg and Löck 2024a) (version 1.5.0), each
+dataset was used to train a model for each of 32 process configurations.
+Each process configuration specifies the learner (generalised linear
+model, L1-regularised linear models (Lasso), gradient boosted linear
+model, or random forest), transformation method (none, conventional
+Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson
+with empirical central normality test (rejecting transformations with
+\(p \leq 0.01\)), and normalisation method (none,
+\(z\)-standardisation), yielding 32 distinct configurations. Before each
+experiment, each dataset was randomly split into a training (70\%) and
+holdout test (30\%) set five times. Thus, a total of 36960 models were
+created. Each model was then evaluated using the holdout test set using
+one of two metrics, i.e.~the root relative squared error (RRSE) for
+regression tasks and the area under the receiver operating
+characteristic curve (AUC) for classification tasks.
 
@@
 \begin{itemize}
+\tightlist
 \item
-  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (87
+  very easy: \(\text{AUC} \geq 0.90\) or \(\text{RRSE} \leq 0.10\) (57
   datasets)
@@
   easy: \(0.90 > \text{AUC} \geq 0.80\) or
-  \(0.30 \geq \text{RRSE} > 0.10\) (46 datasets)
+  \(0.30 \geq \text{RRSE} > 0.10\) (40 datasets)
 \item
   intermediate: \(0.80 > \text{AUC} \geq 0.70\) or
-  \(0.60 \geq \text{RRSE} > 0.30\) (72 datasets)
+  \(0.60 \geq \text{RRSE} > 0.30\) (52 datasets)
 \item
   difficult: \(0.70 > \text{AUC} \geq 0.60\) or
-  \(0.80 \geq \text{RRSE} > 0.60\) (60 datasets)
+  \(0.80 \geq \text{RRSE} > 0.60\) (33 datasets)
 \item
   very difficult: \(0.60 > \text{AUC} \geq 0.50\) or
-  \(1.00 \geq \text{RRSE} > 0.80\) (12 datasets)
+  \(1.00 \geq \text{RRSE} > 0.80\) (48 datasets)
 \item
-  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (8
-  datasets)
+  unsolvable: \(\text{AUC} < 0.50\) or \(\text{RRSE} > 1.00\) (1
+  dataset)
 \end{itemize}
@@
 
-Here we focus on the subset of 232 datasets that contain numeric
-features. Considering single process parameters, the choice of learner
-(Wilcoxon signed rank test: \(p < 10^{-8})\)) and normalisation method
-(Wilcoxon signed rank test: \(p = 0.007\)) had a significant impact (at
-significance level \(p = 0.050\)), but transformation method (Friedman
-test: \(p = 0.054\)) did not.
+Considering single process parameters, the choice of learner (Friedman
+test: \(p < 10^{-8})\)), normalisation method (Wilcoxon signed rank
+test: \(p = 4 \cdot 10^{-8}\)), and transformation method (Friedman
+test: \(p = 0.007\)), all had a significant impact (at \(p = 0.05\)).
 
@@
 transformation method, we first fit a regression random forest (ranger
-package version 0.16.0, \citep{Wright2017-rf}): 2000 trees, minimum
-node size 2, other hyperparameters default) with process parameters and
-task difficulty as predictors and normalised rank as response variable.
-The estimated marginal effects are shown in Figure
+package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
+2, other hyperparameters default) with process parameters and task
+difficulty as predictors and normalised rank as response variable. The
+estimated marginal effects are shown in Figure
 \ref{fig:marginal-effect-plot}. On the scale of normalised ranks
-(\([0.0, 1.0]\)), the overall estimated marginal effect of using a
-random forest instead of generalised linear model was \(0.272\). The
-overall marginal effect of using z-standardisation to normalise features
-was \(0.008\). Transformation methods had the following marginal
-effects: \(0.003\) for using conventional Yeo-Johnson transformation
-instead of no transformation; \(-0.007\) for using robust invariant
-Yeo-Johnson transformation instead of no transformation; \(-0.009\) for
-using robust invariant Yeo-Johnson transformation with empirical central
-normality test instead of no transformation; \(-0.010\) for using robust
-invariant Yeo-Johnson transformation instead of conventional Yeo-Johnson
+(\([0.0, 1.0]\)), only the random forest performed better than the
+average of all learners (rank difference \(0.173\)). The marginal
+improvement in performance from z-standardisation was \(0.011\).
+Transformation methods had the following marginal effects: \(-0.001\)
+for using conventional Yeo-Johnson transformation instead of no
+transformation; \(-0.012\) for using robust invariant Yeo-Johnson
+transformation instead of no transformation; \(-0.012\) for using robust
+invariant Yeo-Johnson transformation with empirical central normality
+test instead of no transformation; \(-0.012\) for using robust invariant
+Yeo-Johnson transformation instead of conventional Yeo-Johnson
 transformation; and \(-0.012\) for using using robust invariant
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_10}
+{\centering \includegraphics[width=1\linewidth]{figure_9}
 
@@
 
-\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets.
-The top-left panel shows the marginal effect of learners, i.e.
-random forests and generalised linear models (GLM).
-Random forests outperform GLMs for all task difficulties.
+\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 36960 machine learning experiments on 231 datasets.
+The four top-left panels shows the marginal effect of learners vs.
+the average, i.e.
+generalised linear models (GLM), L1-regularised LM (Lasso), gradient-boosted LM and random forest.
+Random forests outperform the average.
 The top-right panel shows the marginal effect of feature normalisation methods, i.e.
 no normalisation and z-standardisation.
-z-standardisation is generally beneficial, but the estimated effect is negligible.
-The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method.
-There is no consistent behaviour, and estimated effects are negligible.
+z-standardisation is generally beneficial, but the estimated effect is marginal.
+The bottom panel shows the marginal effects of different transformation methods, split by normalisation method.
+The estimated effects are similar in size to the effect of normalisation and marginal.
 Note that the ranges of the $x$-axes of the three main panels differ.
@@
 \ref{eqn:box-cox-original} for the theoretical analysis in their paper
-\citep{Box1964-mz}, which subsequently became the convention. Yeo and
-Johnson's power transformation lacks a shift parameter altogether 
-\citep{Yeo2000-vw}. We showed that these power transformations are
+(Box and Cox 1964), which subsequently became the convention. Yeo and
+Johnson's power transformation lacks a shift parameter altogether (Yeo
+and Johnson 2000). We showed that these power transformations are
 sensitive to location and scale of data distributions. To mitigate this
@@
 of a feature. This is particularly relevant for automated data
-processing, where such issues may go unnoticed. Compared to non-robust
-location- and scale-invariant transformations, real-world examples with
-outliers showed higher residual errors of transformed features. Robust
-transformations seek to minimise residual errors for the central part of
-the distribution, instead of the entire distribution, including
-outliers. The empirical central normality test indicated that robust
-transformations are better able to achieve central normality in the
-presence of outliers. However, in a machine learning experiment of 232
-real-world datasets that contained at least one numeric feature, we did
-not find a meaningful benefit -- nor detriment -- to model performance
-for location- and scale-invariant power transformations. One reason may
-be that numeric features with large location shifts (\(|\mu| > 1000.0\))
-were uncommon. Of the 4886 numeric features in the 232 datasets, 266
+processing, where such issues may go unnoticed.
+
+In simulation, robust location- and scale-invariant transformations
+ranked best in datasets with outliers, and ranked highly in datasets
+without outliers, except for features without outliers that were
+directly sampled from strictly normal distributions. In real-world
+examples, robust location-and scale-invariant transformations achieved
+central normality when the non-robust variant could not.
+
+However, in a machine learning experiment of 231 real-world datasets
+that contained at least one numeric feature, we did not find a
+meaningful benefit -- nor detriment -- to model performance for
+location- and scale-invariant power transformations. One reason may be
+that numeric features with large location shifts (\(|\mu| > 1000.0\))
+were uncommon. Of the 4886 numeric features in the 231 datasets, 266
 (5\%) features in 34 datasets had large location shifts, of which 200
@@
 \(s\). We derived the log-likelihood function to facilitate optimisation
-using MLE. Alternatively, standardisation of a numeric variable (e.g.,
+using MLE. Alternatively, standardisation of a numeric feature (e.g.,
 through subtracting its median value and division by its interquartile
-range) prior to conventional power transformations may achieve a similar
-effect in reducing sensitivity to the distribution's location and scale.
+range) prior to conventional power transformations can achieve a similar
+effect in reducing sensitivity to the feature's location and scale.
 While this alternative helps prevent these issues -- provided that
-normalisation does not lead to negative values for Box-Cox
-transformation -- location- and scale-invariant transformations seem to
-provide an overall better transformation to normality 
-(\ref{appendix-f-normalisation-before-transformation}).
-
-We assessed several methods for robust power transformation. Methods
-that relied on the z-score of the transformed feature or the residual
-error yielded worse results than the non-robust method. This is partly
-due to the initial choice of threshold parameters. Using different
-initial values, closer to the upper limits
-(\(\delta_1 = 10.0; \delta_2 = 10.0\)), led to a reduced loss. However,
-these values effectively correspond to a non-robust transformation,
-where all instances receive the same weight. Underperformance of these
-methods could be explained by their use of transformed features for
-setting weights. Consequently, the weights change at each iteration in
-the MLE optimisation process. This increases local variance in the
+standardisation does not lead to negative values for Box-Cox
+transformations -- location- and scale-invariant transformations are
+able to achieve better transformations to normality, as demonstrated by
+lower residual errors in simulation experiments.
+
+We assessed several weighting methods to achieve robust power
+transformations. Robust power transformations should satisfy two
+conflicting aims: they should minimise residual errors after
+transformation to central normality, and minimise the overall effect of
+outliers on estimation of transformation parameter \(\lambda\). These
+aims are reflected in the composite loss used to optimise weighting
+parameters. For Yeo-Johnson transformation, empirical probabilities with
+tapered cosine weighting resulted in the lowest loss. The optimal
+parameters led to weights that gradually decline towards the tails of
+empirical probabilities, i.e.~elements with low and high values receive
+less or no weight. Similarly, for Box-Cox transformation, empirical
+probabilities with step weighting resulted in the lowest loss. In this
+case the central 80\% of the elements are used, and the remaining 10\%
+in each tail are ignored. Methods that relied on the z-score of the
+transformed feature or the residual error yielded worse loss than the
+non-robust method or those based on empirical probabilities.
+Underperformance of these weighting methods could be explained by their
+reliance on transformed feature values for setting weights.
+Consequently, their weights change at each iteration in the MLE
+optimisation process. This increases local variance in the
 log-likelihood function and creates local optima that the optimiser may
-not handle well. As a consequence, optimal values for transformation
-parameter \(\lambda\) might differ, which increases the presented loss
-for optimising threshold parameters. Methods that relied on the
-empirical probability did not suffer from this issue, as weights
-remained fixed during MLE.
+not handle well. Methods that relied on the empirical probability did
+not suffer from this issue, as weights remained fixed during MLE.
 
 We introduced an empirical test for central normality to assess whether
-distributions deviate from normality in a way that might require closer
+sequences deviate from normality in a way that might require closer
 inspection prior to further processing. The empirical test for central
 normality differs from other tests for normality, such as the
-Shapiro-Wilk test \citep{Shapiro1965-zd}, in that the test statistic is
-independent of the number of samples. This makes this test more
-practical for assessing central normality for larger sample numbers,
-where other tests may detect inconsequential deviations from normality.
-
-This work has the following limitations. Firstly, we did observe several
-numerical stability issues for optimisation criteria other than MLE
-(\ref{appendix-b-optimisation-of-transformation-parameters}).
-These appear in regions where transformation parameters
-would lead to very large or small numbers when using conventional power
-transformations. For MLE stability issues were not observed. Secondly,
-the empirical central normality test is based on simulations instead of
-statistical theory, and relies on a somewhat arbitrary definition of the
-central portion of a distribution. Thus, while the test may asses
-whether data is sufficiently normally distributed for practical
-purposes, it should not be used as a strict test for normality.
+Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
+the central portion of a feature, instead of the entire feature.
+Compared to the central normality test and the Shapiro-Wilk test, the
+empirical central normality test remains consistent in the presence of
+outliers, although the former tests are more powerful when outliers are
+absent.
+
+This work has the following limitation: We observed several numerical
+stability issues for optimisation criteria other than MLE (Appendix B).
+These appear in regions where transformation parameters would lead to
+very large or small numbers when using conventional power
+transformations. For MLE stability issues were not observed.
 
@@
 \item Data availability: Data and results for the machine learning experiment are available
-from Zenodo (https://doi.org/10.5281/zenodo.13736671). The manuscript was created using R Markdown with integrated code and is
+from Zenodo (https://doi.org/10.5281/zenodo.14986689). The manuscript was created using R Markdown with integrated code and is
 available from the \texttt{power.transform} GitHub repository, together with intermediate analysis results

### 2025-04-11 manuscript_latex/manuscript.tex
Add appendix stuff.

@@
 \begin{split}
-\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
+\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi_{YJ}^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
 & - n \log s + (\lambda - 1) \sum_{i=1}^n \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 \begin{split}
-\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
+\mathcal{L}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
 & - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
@@
 were implemented in the \texttt{power.transform} R software package
-(Zwanenburg and Löck 2024b) (version 1.0.1). Of note, the
+(Zwanenburg and L\"ock 2024b) (version 1.0.1). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in Appendix D.
+for Box-Cox transformations are shown in Appendix E.
 
@@
 datasets to classification tasks. Using the familiar auto-machine
-learning library (Zwanenburg and Löck 2024a) (version 1.5.0), each
+learning library (Zwanenburg and L\"ock 2024a) (version 1.5.0), each
 dataset was used to train a model for each of 32 process configurations.
@@
 
-\section{Log-likelihood functions for location and scale
-invariant power transformation}\label{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}
-
-Location and scale-invariant Box-Cox and Yeo-Johnson transformations are
-parametrised using location \(x_0\) and scale \(s\) parameters, in
+\section{Appendix A: Log-likelihood functions for location- and
+scale-invariant power
+transformation}\label{appendix-a}
+
+Location- and scale-invariant Box-Cox and Yeo-Johnson transformations
+are parametrised using location \(x_0\) and scale \(s\) parameters, in
 addition to transformation parameter \(\lambda\). This leads to the
-following transformations. The location and scale-invariant Box-Cox
+following transformations. The location- and scale-invariant Box-Cox
 transformation is:
@@
 \subsection{Location- and scale-invariant Box-Cox
-transformation}\label{location--and-scale-invariant-box-cox-transformation}
+transformation}\label{location--and-scale-invariant-box-cox-transformation-appendix}
 
@@
 
-Similarly to \citet{Raymaekers2024-zf}, sample weights \(w_i\) are
+Similarly to Raymaekers and Rousseeuw (2024), sample weights \(w_i\) are
 introduced to facilitate robust power transformations. The weighted
@@
 \subsection{Location- and scale-invariant Yeo-Johnson
-transformation}\label{location--and-scale-invariant-yeo-johnson-transformation}
+transformation}\label{location--and-scale-invariant-yeo-johnson-transformation-appendix}
 
@@
 
-\section{Optimisation of transformation parameters}\label{appendix-b-optimisation-of-transformation-parameters}
+\section{Appendix B: Optimisation of transformation
+parameters}\label{appendix-b}
 
@@
 Other potential criteria, such as the Shapiro-Wilk test statistic
-\citep{Shapiro1965-zd} are not investigated here. In the case of the
+(Shapiro and Wilk 1965) are not investigated here. In the case of the
 Shapiro-Wilk test statistic this is because of lack of scalability to
@@
 of location and scale of the transformed feature
-\(\phi^{\lambda, x_0, s} (\mathbf{X})\) \citep{Huber1981-su}.
+\(\phi^{\lambda, x_0, s} (\mathbf{X})\) (Huber 1981).
 
@@
 Note that this criterion is not the same as the Anderson-Darling test
-statistic \citep{Anderson1952-gz}, which involves solving (or
+statistic (Anderson and Darling 1952), which involves solving (or
 approximating) an integral function, contains an extra scalar
@@
 
-\subsubsection{Cram{\'e}r-von Mises
-criterion}\label{cramer-von-mises-criterion}
-
-The Cram{\'e}r-von Mises criterion is also based on the empirical
-distribution function of \(\mathbf{X}\). We define the Cram{\'e}r-von Mises
+\subsubsection{Cramér-von Mises
+criterion}\label{cramuxe9r-von-mises-criterion}
+
+The Cramér-von Mises criterion is also based on the empirical
+distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
 criterion as follows:
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cram{\'e}r-von Mises test statistic \citep{Cramer1928-rc, Von_Mises1928-ef},
-aside from a additive scalar value and the introduction of
+The criterion is similar to the Cramér-von Mises test statistic Mises
+(1928), aside from a additive scalar value and the introduction of
 weights. This criterion, like the Anderson-Darling criterion, seeks to
@@
 substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
-Cram{\'e}r-von Mises criterion.
+Cramér-von Mises criterion.
 
@@
 The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
-test statistic \citep{DAgostino1990-kp}. This test statistic is
+test statistic (D'Agostino and Belanger 1990). This test statistic is
 composed of two separate test statistics, one of which is related to
@@
 
-For the skewness test statistic we first compute \citep{DAgostino1990-kp}:
+For the skewness test statistic we first compute (D'Agostino and
+Belanger 1990):
 
@@
 
-For the kurtosis test statistic we first compute \citep{Anscombe1983-nz, DAgostino1990-kp}:
+For the kurtosis test statistic we first compute (D'Agostino and
+Belanger 1990; Anscombe and Glynn 1983):
 
@@
 The main difference between the above criterion and the Jarque-Bera test
-statistic \citep{Jarque1980-hw} is that a scalar multiplication is absent.
+statistic (Jarque and Bera 1980) is that a scalar multiplication is
+absent.
 
@@
 
-\section{Simulations with other optimisation criteria}\label{appendix-c-simulations-with-other-optimisation-criteria}
+\subsection{Simulations with other optimisation
+criteria}\label{simulations-with-other-optimisation-criteria}
 
 Invariance of location- and scale-invariant power transformations was
-assessed using the optimisation criteria in 
-\ref{appendix-b-optimisation-of-transformation-parameters}. This follows the simulation in section \ref{invariance-to-location-and-scale}, where MLE was
+assessed using the optimisation criteria in Appendix B. 
+This follows the simulation in the main manuscript, where MLE was
 used for optimization. In short, we first randomly drew \(10000\) values
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_appendix_1}
+{\centering \includegraphics[width=1\linewidth]{appendix_figure_1}
 
@@
 
-\section{Experimental results using location- and scale-invariant Box-Cox transformation}
-\label{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}
-
-The effect of using location- and scale-invariant transformations was
-investigated using real-world datasets.
-
-\subsection{Invariance}\label{sec-app:invariance}
-
-Results for Box-Cox transformations of features without outliers are
-shown in Figure \ref{fig:experimental-results-invariance-appendix}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_appendix_2}
-
-}
-
-\caption{Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); penguin body mass (middle row); and latitude coordinates of houses sold in Ames, Iowa (bottom row).
-Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).}\label{fig:experimental-results-invariance-appendix}
-\end{figure}
-
-\subsubsection{Age of patients with lung
-cancer}\label{sec-app:age-of-patients-with-lung-cancer}
-
-Applying conventional and invariant Box-Cox transformations to age of
-patients with lung cancer \citep{Loprinzi1994-cd} yielded the following
-results: no transformation (sum of residuals with normal distribution
-\(\sum r_i = 16.5\)); conventional transformation (\(\lambda = 1.9\),
-\(\sum r_i = 11.5\), \(\mu_{BC} = 1.6 \cdot 10^3\),
-\(\sigma_{BC} = 0.4 \cdot 10^3\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = 1.9\), \(\sum r_i = 11.5\),
-\(\mu_{BC} = 1.6 \cdot 10^3\), \(\sigma_{BC} = 0.4 \cdot 10^3\));
-location- and scale-invariant transformation (\(\lambda = 1.7\),
-\(\sum r_i = 11.6\), \(\mu_{BC} = 1.9\), \(\sigma_{BC} = 0.8\)); and
-robust location- and scale-invariant transformation (\(\lambda = 1.5\),
-\(\sum r_i = 11.6\), \(\mu_{BC} = 3.6\), \(\sigma_{BC} = 1.2\)).
-
-Compared to location- and scale-invariant Yeo-Johnson transformations,
-the Box-Cox transformations do not reduce residuals compared to
-conventional variants.
-
-\subsubsection{Penguin body mass}\label{sec-app:penguin-body-mass}
-
-Applying conventional and invariant Box-Cox transformations to the body
-mass of penguins \citep{Gorman2014-eo} yielded the
-following results: no transformation (residual sum \(\sum r_i = 48.0\));
-conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{BC} = 2.1\), \(\sigma_{BC} = 4 \cdot 10^{-3}\)); location- and
-scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 27.3\),
-\(\mu_{BC} = 0.3\), \(\sigma_{BC} = 0.6\)); and robust location- and
-scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 25.2\),
-\(\mu_{BC} = 0.4\), \(\sigma_{BC} = 0.6\)).
-
-Just as for location- and scale-invariant Yeo-Johnson transformations,
-Box-Cox transformations produced a lower overall residual sum compared
-to their conventional counterparts. Similarly, conventional
-transformations led to low standard deviation \(\sigma_{YJ}\) of the
-body mass feature after transformation.
-
-\subsubsection{Latitude in the Ames housing
-dataset}\label{sec-app:latitude-in-the-ames-housing-dataset}
-
-Applying conventional and invariant Box-Cox transformations to the
-latitude of houses in the Ames housing dataset \citep{De-Cock2011-jf} yielded
-the following results: no transformation (residual sum
-\(\sum r_i = 328\)); conventional transformation (\(\lambda = 62.1\),
-\(\sum r_i = 319\), \(\mu_{BC} = 1.1 \cdot 10^{99}\),
-\(\sigma_{BC} = 0.0 \cdot 10^{99}\)); Raymaekers and Rousseeuw's robust
-adaptation (\(\lambda = 96.0\), \(\sum r_i = 319\),
-\(\mu_{BC} = 6.2 \cdot 10^{153}\),
-\(\sigma_{BC} = 0.3 \cdot 10^{153}\)); location- and scale-invariant
-transformation (\(\lambda = 1.9\), \(\sum r_i = 312\),
-\(\mu_{BC} = 2.3\), \(\sigma_{BC} = 0.9\)); and robust location- and
-scale-invariant transformation (\(\lambda = 1.2\), \(\sum r_i = 316\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 1.4\)).
-
-Similar to conventional Yeo-Johnson transformations (non-robust and
-robust), conventional Box-Cox transformations had high values for the
-\(\lambda\) parameter, which could lead to numerical issues. Location-
-and scale-invariant Box-Cox transformations did not suffer from this
-issue.
-
-\subsection{Robustness against
-outliers}\label{sec-app:robustness-against-outliers}
-
-Results for Box-Cox transformations of features with outliers are shown
-in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_appendix_3}
-
-}
-
-\caption{Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption (top row), where outliers are related to highly fuel-efficient vehicles; and maximum arterial wall thickness in patients with ischemic stroke (bottom row).
-Multiple quantile-quantile plots are shown: for the original feature (left column); the feature transformed using the conventional Box-Cox transformation and Raymaekers and Rousseeuw's robust adaptation (middle column); and the feature transformed using the non-robust and robust location- and-scale invariant Box-Cox transformations (right column).
-Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses.}\label{fig:experimental-results-outlier-robustness-appendix}
-\end{figure}
-
-\subsubsection{Fuel efficiency in the Top Gear
-dataset}\label{sec-app:fuel-efficiency-in-the-top-gear-dataset}
-
-The Top Gear dataset contains data on 297 vehicles, with outliers
-related highly fuel-efficient vehicles \citep{Alfons2021-kc}. Applying
-conventional and invariant Box-Cox transformations to the fuel
-consumption feature yielded the following results: no transformation
-(residual sum \(\sum r_i = 54\), \(p=0.76\)); conventional
-transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
-\(\mu_{BC} = 3.0\), \(\sigma_{BC} = 0.3\), \(p=0.01\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
-\(\mu_{BC} = 29\), \(\sigma_{BC} = 15\), \(p=0.55\)); location- and
-scale-invariant transformation (\(\lambda = -0.7\), \(\sum r_i = 44\),
-\(\mu_{BC} = 0.6\), \(\sigma_{BC} = 0.2\), \(p=0.02\)); and robust
-location- and scale-invariant transformation (\(\lambda = 1.1\),
-\(\sum r_i = 59\), \(\mu_{BC} = 2.4\), \(\sigma_{BC} = 1.8\),
-\(p=0.83\)).
-
-\subsubsection{Maximum arterial wall thickness in an ischemic stroke
-dataset}\label{sec-app:maximum-arterial-wall-thickness-in-an-ischemic-stroke-dataset}
-
-The ischemic stroke dataset contains historic data from 126 patients
-with risk at ischemic stroke \citep{Kuhn2019-kt}. Applying
-conventional and invariant Box-Cox transformations to the maximum
-arterial wall thickness feature yielded the following results: no
-transformation (residual sum \(\sum r_i = 110\), \(p=0.56\));
-conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 33\),
-\(\mu_{BC} = 1.0\), \(\sigma_{BC} = 0.2\), \(p=0.01\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 127\),
-\(\mu_{BC} = 5.5\), \(\sigma_{BC} = 12\), \(p=0.60\)); location- and
-scale-invariant transformation (\(\lambda = -1.0\), \(\sum r_i = 28\),
-\(\mu_{BC} = 0.7\), \(\sigma_{BC} = 0.1\), \(p=0.01\)); and robust
-location- and scale-invariant transformation (\(\lambda = 0.5\),
-\(\sum r_i = 56\), \(\mu_{BC} = 2.2\), \(\sigma_{BC} = 1.4\),
-\(p=0.35\)).
-
-\section{Empirical central normality test}\label{appendix-e-empirical-central-normality-test}
-
-The empirical central normality test was derived using data sampled from
-asymmetric generalised normal distributions, including outliers, to
-resemble more realistic datasets. Here we assess the type I error rate
-of two, less realistic, sets of data:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Data sampled from asymmetric generalised normal distributions without
-  outliers.
-\item
-  Data sampled from normal distributions without outliers, without any
-  power transformation applied.
-\end{enumerate}
-
-Other aspects of the experiment remained the same. Thus, we first drew
-\(m_d=10000\) random distributions. For asymmetric generalised normal
-distributions, each distribution was parametrised with a randomly chosen
-skewness parameter \(\alpha \sim U\left(0.01, 0.99\right)\) and shape
-parameter \(\beta \sim U\left(1.00, 5.00 \right)\). For fully normal
-distributions, skewness parameter \(\alpha = 0.5\) and shape parameter
-\(\beta = 2.0\) were fixed. Location and scale parameters were set as
-\(\mu = 0\) and \(\sigma = 1\), respectively.
-\(n = \lceil 10^\gamma \rceil\) values were then randomly drawn, with
-\(\gamma \sim U\left(1.47, 3.00\right)\), which led to between \(30\)
-and \(1000\) values being drawn to create \(\mathbf{X}_i\). Residuals
-were then computed after performing robust location- and scale-invariant
-transformations with the empirical tapered cosine weighting method for
-the dataset with asymmetric generalised normal distributions, and
-without any transformation for the dataset with only normal
-distributions.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figure_appendix_4}
-
-}
-
-\caption{Type I error rate as function of the test statistic $\tau_{\text{ecn}}$ for five datasets, with central portion $\kappa=0.80$.
-The type I error rate is computed from $m_d=10000$ randomly sampled features.
-These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), or normal distributions with outliers (none).
-The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none).}\label{fig:empirical-central-normality-test-appendix}
-\end{figure}
+\section{Appendix C: Central normality test and empirical central
+normality test critical
+values}\label{appendix-c}
+
+Critical values for the central normality test and its empirical variant
+are found in Table \ref{tab:central-normality-critical-statistic} and
+Table \ref{tab:empirical-central-normality-critical-statistic},
+respectively.
 
 \begin{table}
-\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
-\label{tab:empirical-central-normality-appendix}
-\begin{tabular}{l | c c c c c c c}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
+\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
 
 \toprule
-type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
 
 \midrule
-Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
-Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
-Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
-Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
-normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
-\bottomrule
-
-\end{tabular}
-\end{table}
-
-The results are shown in Figure
-\ref{fig:empirical-central-normality-test-appendix} and Table
-\ref {tab:empirical-central-normality-appendix}. These indicate that the
-test behaves similarly for the different datasets. For low type I error
-rates, the test statistic proposed in section \ref{empirical-central-normality-test-1} is more
-conservative than alternatives based on residuals after Box-Cox
-transformations of asymmetric generalised normally distributed features
-or on residuals from strictly normally distributed features.
-
-\section{Normalisation before transformation}\label{appendix-f-normalisation-before-transformation}
-
-An alternative to location- and scale-invariant transformations is
-normalising feature distributions prior to conventional transformations.
-Table \ref{tab:normalisation-before-transformation-appendix} shows
-residual errors, after transformation to normality, of the five features
-from real-world datasets presented previously section \ref{experimental-results}
-and \ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.
-In these examples location- and scale-invariant
-transformations have similar or lower residual errors compared to errors
-resulting from normalisation prior to transformation.
-
-\begin{table}
-\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. conv.: conventional; norm.: normalisation; rob.: robust}
-\label{tab:normalisation-before-transformation-appendix}
-\begin{tabular}{l | c c c c c}
-
-\toprule
-feature & none & conv. & conv. (z-score norm.) & conv. (rob. scaling) & invariant \\
-
-\midrule
-age                     &  16.5 &  11.5 &  11.5 &  11.3 &   8.8 \\
-penguin body mass       &  48.0 &  32.2 &  33.3 &  32.2 &  26.8 \\
-latitude                & 328.1 & 319.0 & 326.2 & 324.5 & 326.4 \\
-fuel efficiency         &  54.5 &  55.3 &  49.0 &  53.3 &  44.0 \\
-arterial wall thickness & 110.1 &  30.0 &  19.3 &  31.8 &  12.2 \\
+    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
+   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
+   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
+   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
+  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
+  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
+  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
+ 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
+ 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
+ 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
 \bottomrule
 \end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
+\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+\label{tab:empirical-central-normality-critical-statistic}
+\begin{tabular}{l | r r r r r r r r r}
+
+\toprule
+$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
+
+\midrule
+    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
+   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
+   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
+   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
+  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
+  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
+  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
+ 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
+ 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
+ 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
+10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_figure_2}
+
+}
+
+\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.80}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
+The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
+with fully normal data where 10\% of elements are replaced by outliers.
+emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
+\end{figure}
+
+\section{Appendix D: Assessing transformations using simulated
+data}\label{appendix-d}
+
+\subsection{Ranking Box-Cox
+transformations}\label{ranking-box-cox-transformations}
+
+Box-Cox transformations were assessed in the same manner as Yeo-Johnson
+transformations. The results are shown in Table
+\ref{tab:comparison_methods_simulations_box_cox}.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+If necessary, each sequence was shifted so that every element had a strictly positive value. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_box_cox}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
+conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
+conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
+conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
+Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
+invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
+robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Examples using clean data}\label{examples-using-clean-data}
+
+Robust transformations are hypothesised to have a cost in efficiency for
+data without outliers, i.e.~clean data. Here we draw nine sequences with
+elements randomly drawn from a standard normal distribution
+\(\mathcal{N}(0,1)\). Subsequently, we perform an inverse transformation
+\(\left(\phi^{\lambda, 0, 1}\right)^{-1}\), with
+\(\lambda \in \left(0, 2\right)\).
+
+The sequences drawn resulted from the permutations of the number of
+elements of each sequence (\(n \in \{30, 100, 500\}\)) and
+transformation parameter for the inverse transformation
+\(\lambda \in \{0.1, 1.0, 1.9\}\). Each sequence then underwent power
+transformation using Box-Cox and Yeo-Johnson transformations. For
+Box-Cox transformations, each sequence was shifted prior to inverse
+transformation so that every element after inverse transformation would
+be strictly positive.
+
+Results for Box-Cox transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-bc},
+\ref{tab:clean-transformation-appendix-lambda-bc}, and
+\ref{tab:clean-transformation-appendix-p-value-bc}. Additionally,
+results for Yeo-Johnson transformations are shown in Table
+\ref{tab:clean-transformation-appendix-residuals-yj},
+\ref{tab:clean-transformation-appendix-lambda-yj}, and
+\ref{tab:clean-transformation-appendix-p-value-yj}. As may be observed,
+in these examples robust location- and shift-invariant transformations
+have higher residual errors because of low weights being assigned to
+tails of a distribution. This is also shown in Tables
+\ref{tab:comparison_methods_simulations_yeo_johnson} and
+\ref{tab:comparison_methods_simulations_box_cox} for clean data without
+outliers.
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-residuals-yj}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 4.0 & 3.4 & 4.6 & 40.1 &  8.9 & 20.8 & 184.6 & 10.1 & 191.8 \\
+conventional                          & & 2.7 & 3.3 & 3.6 &  6.5 &  8.1 &  5.6 &  14.2 & 10.1 &  16.1 \\
+conventional (z-score norm.)          & & 2.5 & 3.3 & 3.7 &  9.8 &  8.1 &  5.6 &  24.1 & 10.1 &  15.9 \\
+conventional (robust scaling)         & & 2.6 & 3.3 & 3.7 &  7.4 &  8.2 &  5.6 &  15.5 & 10.1 &  14.0 \\
+Raymaekers-Rousseeuw                  & & 3.8 & 3.3 & 3.6 &  6.5 &  8.5 &  5.6 &  14.4 & 11.0 &  17.6 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.6 & 3.3 & 3.7 &  9.8 &  8.5 &  5.6 &  24.1 & 11.1 &  15.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.6 & 3.3 & 3.7 &  7.4 &  8.6 &  5.6 &  15.4 & 11.1 &  15.1 \\
+invariant                             & & 3.0 & 3.5 & 3.6 &  5.0 &  8.1 &  5.6 &  13.1 & 10.9 &  17.8 \\
+robust invariant                      & & 3.0 & 3.4 & 3.6 &  5.5 & 10.1 &  6.2 &  18.7 & 13.2 &  32.8 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-lambda-yj}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.6 & 1.2 & 1.5 &  0.0 & 0.9 & 1.6 &  0.1 & 1.0 & 1.9 \\
+conventional (z-score norm.)          & & 0.5 & 1.2 & 1.5 & -0.2 & 0.9 & 1.7 & -0.2 & 1.0 & 2.2 \\
+conventional (robust scaling)         & & 0.5 & 1.3 & 1.6 & -0.3 & 0.9 & 1.7 & -0.2 & 1.0 & 2.1 \\
+Raymaekers-Rousseeuw                  & & 1.0 & 1.2 & 1.5 &  0.0 & 0.7 & 1.6 &  0.1 & 1.0 & 1.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.7 & 1.2 & 1.5 & -0.2 & 0.7 & 1.7 & -0.2 & 1.0 & 2.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.6 & 1.3 & 1.6 & -0.3 & 0.6 & 1.7 & -0.2 & 1.0 & 2.1 \\
+invariant                             & & 0.4 & 1.4 & 1.7 &  0.1 & 0.9 & 1.8 &  0.1 & 1.1 & 3.0 \\
+robust invariant                      & & 0.5 & 1.1 & 1.5 & -0.1 & 0.7 & 1.7 &  0.1 & 1.1 & 1.9 \\
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-p-value-yj}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.981 & 0.816 & 0.576 & $<0.001$ & 0.099 & 0.011 & $<0.001$ & 0.993 & $<0.001$ \\
+conventional                          & & 0.979 & 0.614 & 0.617 &   0.722  & 0.221 & 0.782 &   0.946  & 0.997 &   0.637 \\
+conventional (z-score norm.)          & & 0.970 & 0.594 & 0.645 &   0.251  & 0.224 & 0.895 &   0.529  & 0.998 &   0.828 \\
+conventional (robust scaling)         & & 0.974 & 0.593 & 0.640 &   0.538  & 0.207 & 0.837 &   0.933  & 0.998 &   0.810 \\
+Raymaekers-Rousseeuw                  & & 0.983 & 0.614 & 0.617 &   0.722  & 0.622 & 0.782 &   0.927  & 0.985 &   0.831 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.985 & 0.594 & 0.645 &   0.251  & 0.623 & 0.895 &   0.564  & 0.984 &   0.807 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.987 & 0.593 & 0.640 &   0.538  & 0.599 & 0.837 &   0.935  & 0.984 &   0.875 \\
+invariant                             & & 0.987 & 0.487 & 0.573 &   0.778  & 0.241 & 0.683 &   0.963  & 0.997 &   0.865 \\
+robust invariant                      & & 0.992 & 0.753 & 0.667 &   0.909  & 0.739 & 0.628 &   0.968  & 0.994 &   0.680 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-residuals-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 5.3 & 3.4 & 3.8 & 45.7 &  8.9 & 6.0 & 221.8 & 10.1 & 29.7 \\
+conventional                          & & 2.4 & 3.2 & 3.5 &  5.9 &  8.4 & 5.5 &  14.3 &  9.8 & 16.1 \\
+conventional (z-score norm.)          & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+conventional (robust scaling)         & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  14.0 & 10.4 & 16.1 \\
+Raymaekers-Rousseeuw                  & & 2.4 & 3.2 & 3.5 &  5.9 & 13.3 & 5.7 &  16.1 & 13.6 & 16.9 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  19.1 & 19.0 & 19.9 \\
+Raymaekers-Rousseeuw (robust scaling) & & 2.4 & 3.2 & 4.9 &  6.3 &  8.9 & 6.2 &  18.4 & 19.0 & 19.9 \\
+invariant                             & & 2.4 & 3.2 & 3.7 &  5.4 &  8.7 & 5.5 &  13.7 &  9.9 & 16.0 \\
+robust invariant                      & & 2.4 & 4.3 & 4.7 &  8.9 & 20.4 & 6.7 &  14.1 & 15.8 & 37.6 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-lambda-bc}
+\small{
+\begin{tabular}{l | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+conventional                          & & 0.4 & 1.3 & 0.5 & 0.0 &  0.9 & 0.7 & 0.1 & 1.1 & 2.0 \\
+conventional (z-score norm.)          & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+conventional (robust scaling)         & & 0.4 & 1.1 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.0 & 1.4 \\
+Raymaekers-Rousseeuw                  & & 0.4 & 1.3 & 0.5 & 0.0 & -0.1 & 0.5 & 0.1 & 0.9 & 1.8 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.4 & 1.1 & 0.4 & 0.1 &  0.8 & 0.7 & 0.1 & 0.9 & 1.2 \\
+invariant                             & & 0.4 & 1.6 & 0.8 & 0.2 &  0.9 & 0.8 & 0.1 & 1.1 & 1.9 \\
+robust invariant                      & & 0.5 & 0.8 & 0.4 & 0.0 &  0.2 & 0.7 & 0.1 & 0.9 & 0.9 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
+\label{tab:clean-transformation-appendix-p-value-bc}
+\small{
+\begin{tabular}{p{3.5cm} | l r r r r r r r r r}
+
+\toprule
+& n: & \multicolumn{3}{c}{30} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{500} \\
+transformation & $\lambda$: & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 & 0.1 & 1.0 & 1.9 \\
+
+\midrule
+
+none                                  & & 0.940 & 0.816 & 0.498 & $<0.001$ & 0.098 & 0.714 & $<0.001$ & 0.993 & 0.928 \\
+conventional                          & & 0.948 & 0.695 & 0.603 &   0.798  & 0.163 & 0.791 &   0.931  & 0.997 & 0.658 \\
+conventional (z-score norm.)          & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+conventional (robust scaling)         & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
+Raymaekers-Rousseeuw                  & & 0.948 & 0.695 & 0.603 &   0.798  & 0.768 & 0.806 &   0.816  & 0.962 & 0.782 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.775  & 0.866 & 0.921 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.807  & 0.866 & 0.921 \\
+invariant                             & & 0.947 & 0.661 & 0.624 &   0.611  & 0.182 & 0.789 &   0.956  & 0.997 & 0.664 \\
+robust invariant                      & & 0.969 & 0.824 & 0.750 &   0.972  & 0.779 & 0.775 &   0.956  & 0.934 & 0.840 \\
+
+\bottomrule
+\end{tabular}
+}
+\end{center}
+\end{table}
+
+\section{Appendix E: Experimental
+results}\label{appendix-e}
+
+The effect of using location- and scale-invariant transformations was
+investigated using real-world datasets, as described in the main
+manuscript.
+
+\subsection{Yeo-Johnson
+transformation}\label{yeo-johnson-transformation}
+
+Additional results for Yeo-Johnson transformations are shown in Tables
+\ref{tab:experimental-results-appendix-residuals},
+\ref{tab:experimental-results-appendix-lambda},
+\ref{tab:experimental-results-appendix-p-value}.
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-residuals}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none                                  & 16.5 & 110.1 & 54.5 & 328.4 & 48.0 \\
+conventional                          & 11.5 &  30.0 & 55.3 & 319.0 & 32.2 \\
+conventional (z-score norm.)          & 11.5 &  19.3 & 49.0 & 326.2 & 33.3 \\
+conventional (robust scaling)         & 11.3 &  31.8 & 53.3 & 324.5 & 32.2 \\
+Raymaekers-Rousseeuw                  & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\
+robust invariant                      &  9.3 &  30.2 & 55.8 & 308.1 & 22.0 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-lambda}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+conventional                          & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.)          & 1.2 & -1.7 & -0.1 &  1.3 &  0.6 \\
+conventional (robust scaling)         & 1.3 &  0.0 &  0.2 &  1.4 &  0.4 \\
+Raymaekers-Rousseeuw                  & 2.0 &  1.1 &  0.8 & 95.4 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 1.3 &  1.4 &  1.0 &  1.3 &  0.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.4 &  1.3 &  1.0 &  1.4 &  0.4 \\
+invariant                             & 1.3 &  0.2 & -1.3 &  1.5 &  0.5 \\
+robust invariant                      & 1.3 & -0.3 &  1.0 &  1.1 &  0.3 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-p-value}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+none                                  & 0.694 & 0.832     & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.003     & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.919 & 0.021     & $< 0.001$ & $< 0.001$ & 0.099 \\
+conventional (robust scaling)         & 0.933 & $< 0.001$ & $< 0.001$ & $< 0.001$ & 0.129 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.881     & 0.373     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.694 & 0.899     & 0.809     & $< 0.001$ & 0.099 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.900 & 0.880     & 0.811     & $< 0.001$ & 0.144 \\
+invariant                             & 0.976 & 0.147     & $< 0.001$ & $< 0.001$ & 0.276 \\
+robust invariant                      & 0.933 & 0.178     & 0.759     & $< 0.001$ & 0.688 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\subsection{Box-Cox transformation}\label{box-cox-transformation}
+
+Results for Box-Cox transformations are shown in Tables
+\ref{tab:experimental-results-appendix-residuals-bc},
+\ref{tab:experimental-results-appendix-lambda-bc} and
+\ref{tab:experimental-results-appendix-p-value-bc}.
+
+\begin{table}
+\begin{center}
+\caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-residuals-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+none                                  & 16.5 & 110.1 &  54.5 & 328.4 & 48.0 \\
+conventional                          & 11.5 &  33.5 &  55.7 & 318.9 & 32.2 \\
+conventional (z-score norm.)          & 12.9 &  44.5 &  58.8 & 305.3 & 27.3 \\
+conventional (robust scaling)         & 12.9 &  44.7 &  59.0 & 305.3 & 27.3 \\
+Raymaekers-Rousseeuw                  & 11.5 & 127.1 &  47.6 & 314.8 & 32.2 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & 13.1 &  54.8 & 127.5 & 510.2 & 23.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 18.9 & 100.2 & 113.7 & 510.2 & 23.6 \\
+invariant                             & 11.6 &  28.0 &  48.4 & 311.8 & 27.3 \\
+robust invariant                      & 12.8 & 150.1 &  60.5 & 646.0 & 27.4 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality. 
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-lambda-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+conventional                          & 1.9 & -0.5 & -0.1 & 62.1 & -0.5 \\
+conventional (z-score norm.)          & 1.2 &  0.1 &  0.2 &  1.2 &  0.5 \\
+conventional (robust scaling)         & 1.2 &  0.1 &  0.1 &  1.2 &  0.5 \\
+Raymaekers-Rousseeuw                  & 1.9 &  1.1 &  0.8 & 95.9 & -0.5 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 1.2 &  0.6 & -0.5 &  0.6 &  0.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.9 &  1.0 & -0.5 &  0.6 &  0.3 \\
+invariant                             & 1.7 & -1.0 & -0.7 &  1.9 &  0.5 \\
+robust invariant                      & 1.3 &  1.2 &  1.1 &  0.4 &  0.1 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+p-values of empirical central normality tests for features from real-world datasets after Box-Cox transformation to normality.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
+AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
+\label{tab:experimental-results-appendix-p-value-bc}
+\begin{tabular}{l | r r r r r}
+
+\toprule
+feature & age & AWT & FE & latitude & PBM \\
+
+\midrule
+
+none                                  & 0.694 & 0.832 & 0.727     & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.961 & 0.002 & $< 0.001$ & $< 0.001$ & 0.101 \\
+conventional (z-score norm.)          & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+conventional (robust scaling)         & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
+Raymaekers-Rousseeuw                  & 0.961 & 0.871 & 0.365     & $< 0.001$ & 0.101 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.891 & 0.268 & $< 0.001$ & $< 0.001$ & 0.495 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.562 & 0.789 & $< 0.001$ & $< 0.001$ & 0.492 \\
+invariant                             & 0.962 & 0.004 & $< 0.001$ & $< 0.001$ & 0.154 \\
+robust invariant                      & 0.899 & 0.890 & 0.866     & $< 0.001$ & 0.622 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
 \end{table}

### 2025-04-11 manuscript_latex/manuscript.tex
Update references.

@@
 aim to stabilise variance and improve normality of such features
-(Bartlett 1947; John W. Tukey 1957). The two most commonly used
-transformations are that of Box and Cox (1964) and Yeo and Johnson
-(2000).
+\citep{Bartlett1947-rx, Tukey1957-rt}. The two most commonly used
+transformations are that of \citet{Box1964-mz} and \citet{Yeo2000-vw}.
 
@@
 
-Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
-and Wilk 1965), could be automatically applied to transformed features.
+Statistical tests for normality, such as the Shapiro-Wilk test \citep{Shapiro1965-zd},
+could be automatically applied to transformed features.
 However, given sufficiently large sample sizes such tests can detect
@@
 transformations to normality, as shown in Figure
-\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
-(2024), the general aim of power transformations should be to transform
+\ref{fig:decreased-normality}. As indicated by \citet{Raymaekers2024-zf},
+ the general aim of power transformations should be to transform
 non-outlier data to normality, i.e.~achieve \emph{central normality}. To
@@
 \ref{eqn:box-cox-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
-and Rousseeuw (2024), these log-likelihood functions include additional
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to \citet{Raymaekers2024-zf},
+these log-likelihood functions include additional
 terms to accommodate estimation of \(x_0\) and \(s\). The weighted
@@
 recipe for introducing skewness into the otherwise symmetric generalised
-normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
+normal distribution  \citep{Gijbels2019-te}, leading to
 the following probability density function:
@@
 \(F_{\alpha}^{-1}(p)\), which incorporates the quantile function of the
-symmetric generalised normal distribution derived by Griffin (Griffin
-2018):
+symmetric generalised normal distribution derived by Griffin \citep{Gijbels2019-te}:
 
@@
 Deviations from normality can be detected by normality tests, such as
-the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
+the Shapiro-Wilk test  \citep{Gijbels2019-te}. In practice, normality
 tests may be too stringent with large sample sizes, outliers, or both.
@@
 Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$, 
-with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su], 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ \citep{Huber1981-su}, 
 and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
@@
 and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
-algorithm for derivative-free bound constraint optimisation (Powell
-2009) through maximum likelihood estimation. The required algorithms
+algorithm for derivative-free bound constraint optimisation \citep{Powell2009-zb} 
+through maximum likelihood estimation. The required algorithms
 were implemented in the \texttt{power.transform} R software package
-(Zwanenburg and L\"ock 2024b) (version 1.0.1). Of note, the
+\citep{Zwanenburg2024-kq} (version 1.0.1). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 Outlier values were then drawn to randomly replace 10 percent of the
-elements of \(\mathbf{X}_i\). Outlier values were set according to John
-W. Tukey (1977), as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
+elements of \(\mathbf{X}_i\). Outlier values were set according to \citet{Tukey1977-xm},
+as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
 Then the corresponding outlier value was:
@@
 power transformations, conventional and Raymaekers and Rousseeuw's
-robust adaptation (Raymaekers and Rousseeuw 2024) were assessed. For the
+robust adaptation \citep{Raymaekers2024-zf} were assessed. For the
 latter two methods, normalisation before standardisation using was
@@
 dataset contains data on 2930 properties that were sold between 2006 and
-2010 (De Cock 2011), including their geospatial coordinates. The
+2010 \citep{De-Cock2011-jf}, including their geospatial coordinates. The
 latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
@@
 We used 231 datasets containing at least one numeric feature from the
-Penn Machine Learning Benchmarks collection (Romano et al. 2022). In
+Penn Machine Learning Benchmarks collection \citep{Romano2022-gq}. In
 this collection, 114 datasets correspond to regression tasks and 117
 datasets to classification tasks. Using the familiar auto-machine
-learning library (Zwanenburg and L\"ock 2024a) (version 1.5.0), each
+learning library \citep{Zwanenburg2021-so} (version 1.5.0), each
 dataset was used to train a model for each of 32 process configurations.
@@
 transformation method, we first fit a regression random forest (ranger
-package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
+package \citep{Wright2017-rf} version 0.16.0): 2000 trees, node size
 2, other hyperparameters default) with process parameters and task
@@
 \ref{eqn:box-cox-original} for the theoretical analysis in their paper
-(Box and Cox 1964), which subsequently became the convention. Yeo and
-Johnson's power transformation lacks a shift parameter altogether (Yeo
-and Johnson 2000). We showed that these power transformations are
+\citep{Box1964-mz}, which subsequently became the convention. Yeo and
+Johnson's power transformation lacks a shift parameter altogether \citep{Yeo2000-vw}.
+We showed that these power transformations are
 sensitive to location and scale of data distributions. To mitigate this
@@
 normality differs from other tests for normality, such as the
-Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
+Shapiro-Wilk test \citep{Shapiro1965-zd}, as it assesses normality of
 the central portion of a feature, instead of the entire feature.
@@
 Other potential criteria, such as the Shapiro-Wilk test statistic
-(Shapiro and Wilk 1965) are not investigated here. In the case of the
+\citet{Raymaekers2024-zf} are not investigated here. In the case of the
 Shapiro-Wilk test statistic this is because of lack of scalability to
@@
 \subsubsection{Cramér-von Mises
-criterion}\label{cramuxe9r-von-mises-criterion}
+criterion}\label{cramer-von-mises-criterion}
 
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cramér-von Mises test statistic Mises
-(1928), aside from a additive scalar value and the introduction of
+The criterion is similar to the Cramér-von Mises test statistic 
+\citep{Cramer1928-rc, Von_Mises1928-ef}, aside from a additive scalar value and the introduction of
 weights. This criterion, like the Anderson-Darling criterion, seeks to
@@
 The main difference between the above criterion and the Jarque-Bera test
-statistic (Jarque and Bera 1980) is that a scalar multiplication is
+statistic \citep{Jarque1980-hw} is that a scalar multiplication is
 absent.

### 2025-04-12 manuscript_rmarkdown/manuscript_appendix.Rmd
Add appendix stuff.
Update manuscript tex.
It compiles properly.

@@
 Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
-The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
+The log-likelihood function $\mathcal{L}^{\lambda, x_0, s}$ is then:
 ```{=latex}
@@
 
-Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
+Thus the final term in $\mathcal{L}_{\text{BC}}^{\lambda, x_0, s}$ is:
 ```{=latex}
@@
 
-Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
+Thus the final term in $\mathcal{L}_{\text{YJ}}^{\lambda, x_0, s}$ is:
 ```{=latex}
@@
 
-# Appendix D: Central normality test and empirical central normality test critical values
+# Appendix C: Central normality test and empirical central normality test critical values
 
@@
 
-# Appendix E: Assessing transformations using simulated data
+# Appendix D: Assessing transformations using simulated data
 
@@
 
-# Appendix F: Experimental results
+# Appendix E: Experimental results
 

### 2025-04-12 manuscript_latex/manuscript.tex

@@
 \usepackage{listings}%
+\usepackage{placeins}
 %%%%
@@
 i.e.~using maximum likelihood estimation (MLE). A full derivation of the
-log-likelihood function for both transformations is shown in Appendix A.
+log-likelihood function for both transformations is shown in \hyperref[appendix-a]{Appendix A}.
 The location- and scale-invariant Box-Cox log-likelihood function is:
@@
 dataset. Average ranks for Yeo-Johnson transformations are shown in
-Table \ref{tab:comparison_methods_simulation_yeo_johnson}. Location and
+Table \ref{tab:comparison_methods_simulations_yeo_johnson}. Location and
 shift-invariant Yeo-Johnson transformation ranked best for all datasets
@@
 \begin{table}
-\begin{center}
 \caption{
@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in Appendix E.
+for Box-Cox transformations are shown in \hyperref[appendix-e]{Appendix E}.
 
@@
 Loprinzi et al. \citep{Loprinzi1994-cd}. The age in the cohort was
-\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
+\(62.4 \pm 9.1\) (mean $\pm$ standard deviation) years. Applying
 conventional and invariant Yeo-Johnson transformations to patient age
@@
 penguins of three different species \citep{Gorman2014-eo}.
-The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
+The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean $\pm$ standard
 deviation) grams, and not centrally normal (\(p < 0.001\)). Applying
@@
 2010 \citep{De-Cock2011-jf}, including their geospatial coordinates. The
-latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
+latitude was \(42.03 \pm 0.02)\) (mean $\pm$ standard deviation). Applying
 conventional and invariant Yeo-Johnson transformations to latitude
@@
 \begin{itemize}
-\tightlist
 \item
@@
 \end{figure}
+
+\FloatBarrier
 
@@
 This work has the following limitation: We observed several numerical
-stability issues for optimisation criteria other than MLE (Appendix B).
+stability issues for optimisation criteria other than MLE (\hyperref[appendix-b]{Appendix B}).
 These appear in regions where transformation parameters would lead to
@@
 
-\section{Appendix A: Log-likelihood functions for location- and
+\section{Log-likelihood functions for location- and
 scale-invariant power
@@
 
-\section{Appendix B: Optimisation of transformation
+\section{Optimisation of transformation
 parameters}\label{appendix-b}
@@
 
-\subsubsection{Cramér-von Mises
+\subsubsection{Cram\'{e}r-von Mises
 criterion}\label{cramer-von-mises-criterion}
 
-The Cramér-von Mises criterion is also based on the empirical
-distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
+The Cram\'{e}r-von Mises criterion is also based on the empirical
+distribution function of \(\mathbf{X}\). We define the Cram\'{e}r-von Mises
 criterion as follows:
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cramér-von Mises test statistic 
+The criterion is similar to the Cram\'{e}r-von Mises test statistic 
 \citep{Cramer1928-rc, Von_Mises1928-ef}, aside from a additive scalar value and the introduction of
@@
 substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
-Cramér-von Mises criterion.
+Cram\'{e}r-von Mises criterion.
 
@@
 Invariance of location- and scale-invariant power transformations was
-assessed using the optimisation criteria in Appendix B. 
+assessed using the optimisation criteria in \hyperref[appendix-b]{Appendix B}. 
 This follows the simulation in the main manuscript, where MLE was
@@
 
-\section{Appendix C: Central normality test and empirical central
+\FloatBarrier
+
+\section{Central normality test and empirical central
 normality test critical
@@
 \begin{table}
-\begin{center}
 \caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
@@
 \begin{table}
-\begin{center}
 \caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
@@
 
-\section{Appendix D: Assessing transformations using simulated
+\FloatBarrier
+
+\section{Assessing transformations using simulated
 data}\label{appendix-d}
@@
 \end{tabular}
-\end{center}
 \end{table}
+
+\FloatBarrier
 
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
@@
 }
-\end{center}
 \end{table}
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
@@
 }
-\end{center}
 \end{table}
 
-\section{Appendix E: Experimental
+\FloatBarrier
+
+\section{Experimental
 results}\label{appendix-e}
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality. 

### 2025-04-12 manuscript_rmarkdown/manuscript.tex

@@
 \(s\). We derived the log-likelihood function to facilitate optimisation
-using MLE. Alternatively, standardisation of a numeric variable (e.g.,
+using MLE. Alternatively, standardisation of a numeric feature (e.g.,
 through subtracting its median value and division by its interquartile
@@
 We assessed several weighting methods to achieve robust power
-transformations. Methods that relied on the z-score of the transformed
-feature or the residual error yielded worse results than the non-robust
-method. Underperformance of these weighting methods could be explained
-by their reliance on transformed feature values for setting weights.
+transformations. Robust power transformations should satisfy two
+conflicting aims: they should minimise residual errors after
+transformation to central normality, and minimise the overall effect of
+outliers on estimation of transformation parameter \(\lambda\). These
+aims are reflected in the composite loss used to optimise weighting
+parameters. For Yeo-Johnson transformation, empirical probabilities with
+tapered cosine weighting resulted in the lowest loss. The optimal
+parameters led to weights that gradually decline towards the tails of
+empirical probabilities, i.e.~elements with low and high values receive
+less or no weight. Similarly, for Box-Cox transformation, empirical
+probabilities with step weighting resulted in the lowest loss. In this
+case the central 80\% of the elements are used, and the remaining 10\%
+in each tail are ignored. Methods that relied on the z-score of the
+transformed feature or the residual error yielded worse loss than the
+non-robust method or those based on empirical probabilities.
+Underperformance of these weighting methods could be explained by their
+reliance on transformed feature values for setting weights.
 Consequently, their weights change at each iteration in the MLE

### 2025-10-15 manuscript_rmarkdown/manuscript.Rmd
Update manuscript tex.
It compiles properly.
Replace IQR for standardisation with Qn estimator.

@@
 
-2. robust scaling: $x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$,
-   with $\text{IQR}$ representing the interquartile range of $\mathbf{X}$.
+2. robust scaling: $x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / Q_n\left(\mathbf{X}\right)$,
+   with $Q_n$ representing the robust $Q_n$ scale estimator of $\mathbf{X}$ [@Croux1992-zz; @Rousseeuw1993-it].
 

### 2025-10-15 manuscript_rmarkdown/manuscript_appendix.Rmd
Replace IQR for standardisation with Qn estimator.

@@
 conventional (z-score norm.)          & 11.5 &  19.3 & 49.0 & 326.2 & 33.3 \\
-conventional (robust scaling)         & 11.3 &  31.8 & 53.3 & 324.5 & 32.2 \\
+conventional (robust scaling)         & 11.3 &  33.3 & 55.9 & 326.4 & 31.5 \\
 Raymaekers-Rousseeuw                  & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
 Raymaekers-Rousseeuw (z-score norm.)  & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
-Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+Raymaekers-Rousseeuw (robust scaling) & 13.4 & 203.0 & 56.9 & 326.4 & 31.5 \\
 invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\
@@
 conventional (z-score norm.)          & 1.2 & -1.7 & -0.1 &  1.3 &  0.6 \\
-conventional (robust scaling)         & 1.3 &  0.0 &  0.2 &  1.4 &  0.4 \\
+conventional (robust scaling)         & 1.2 &  0.0 &  0.3 &  1.3 &  0.6 \\
 Raymaekers-Rousseeuw                  & 2.0 &  1.1 &  0.8 & 95.4 & -0.5 \\
 Raymaekers-Rousseeuw (z-score norm.)  & 1.3 &  1.4 &  1.0 &  1.3 &  0.6 \\
-Raymaekers-Rousseeuw (robust scaling) & 1.4 &  1.3 &  1.0 &  1.4 &  0.4 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.3 &  1.2 &  1.0 &  1.3 &  0.6 \\
 invariant                             & 1.3 &  0.2 & -1.3 &  1.5 &  0.5 \\
@@
 Raymaekers-Rousseeuw (z-score norm.)  & 13.1 &  54.8 & 127.5 & 510.2 & 23.6 \\
-Raymaekers-Rousseeuw (robust scaling) & 18.9 & 100.2 & 113.7 & 510.2 & 23.6 \\
+Raymaekers-Rousseeuw (robust scaling) & 13.1 & 100.2 &  44.9 & 423.2 & 23.6 \\
 invariant                             & 11.6 &  28.0 &  48.4 & 311.8 & 27.3 \\
@@
 Raymaekers-Rousseeuw (z-score norm.)  & 1.2 &  0.6 & -0.5 &  0.6 &  0.3 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.9 &  1.0 & -0.5 &  0.6 &  0.3 \\
+Raymaekers-Rousseeuw (robust scaling) & 1.2 &  1.0 &  0.8 &  0.7 &  0.3 \\
 invariant                             & 1.7 & -1.0 & -0.7 &  1.9 &  0.5 \\

### 2025-10-15 manuscript_rmarkdown/simulations.R

@@ -1400,7 +1400,8 @@
       x <- (x - mean(x)) / stats::sd(x)
 
     } else if (endsWith(transformation_method, "(robust scaling)")) {
-      x <- (x - stats::median(x)) / stats::IQR(x)
+      x <- (x - stats::median(x)) / robustbase::Qn(x)
+
     }
 
     if (startsWith(transformation_method, "none")) {

### 2025-10-17 manuscript_rmarkdown/manuscript_appendix.Rmd
Add information on subsets of ranking of Yeo-Johnson based methods.

@@
 
+## Ranking Yeo-Johnson transformations (pt. II)
+
+The main manuscript shows the overall ranking of different Yeo-Johnson-based power transformation methods across three
+sets of problems.
+To assess the effect of limited sequence lengths, we perform ranking on subsets of the experiment.
+The results for ranking the subset of sequences with 30 samples or fewer ($n = 16001$) 
+is shown in Table \ref{tab:comparison_methods_simulations_yeo_johnson_10_30}.
+In the clean dataset without outliers, where all sequences were sampled from a normal distribution and 
+then underwent an inverse Yeo-Johnson transformation,
+conventional and location- and scale-invariant are ranked similarly, with no clear advantage
+to any method. For the remaining datasets, the robust location- and scale-invariant
+method ranked best. It slightly outperformed (robustly scaled) conventional Yeo-Johnson transformations
+and (z-score normalised) robust transformations [@Raymaekers2024-zf], in datasets without and with outliers, respectively.
+
+In the subset for larger sequence lengths, with at least 1000 samples per sequence ($n = 33037$),
+the cost of making the location- and scale-invariant transformation robust against outliers
+becomes more apparent (Table \ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000}).
+In the clean dataset without outliers, down-weighting of the tails of the distribution by
+the robust location- and scale-invariant method leads to larger residuals compared 
+to other methods (Table \ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}.
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# 30 samples or fewer
+data <- data[n_samples <= 30L, ]
+
+# Add rank based on residual error for central normality.
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "norm_residuals" := residuals / n_samples]
+data[outlier == TRUE, "norm_residuals" := residuals_central / n_samples]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90),
+    "avg. norm. res." = round(mean(norm_residuals), digits = 3)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+# Table with average rank.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
+# Table with average normalised residuals.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg. norm. res.")
+
+```
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Each sequence contained 30 samples or fewer. 
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.75  &         6.81  &         7.27  &         6.40  &         7.45  &         6.60 \\
+conventional                          & &         4.19  &         5.47  &         4.48  &         5.78  &         6.55  &         5.94 \\
+conventional (z-score norm.)          & &         4.36  &         5.12  &         5.01  &         4.99  &         4.35  &         4.70 \\
+conventional (robust scaling)         & & \textbf{4.05} &         5.63  &         4.38  &         5.49  &         3.76  &         5.20 \\
+Raymaekers-Rousseeuw                  & &         5.36  &         4.19  &         5.14  &         4.62  &         6.32  &         5.83 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.56  &         4.08  &         5.67  &         4.05  &         5.03  &         3.81 \\
+Raymaekers-Rousseeuw (robust scaling) & &         5.33  &         4.14  &         5.15  &         4.15  &         4.55  &         3.94 \\
+invariant                             & &         4.11  &         5.92  &         4.12  &         5.67  &         3.62  &         5.37 \\
+robust invariant                      & &         4.28  & \textbf{3.65} & \textbf{3.79} & \textbf{3.85} & \textbf{3.37} & \textbf{3.61} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained 30 samples or fewer. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.246 & 0.139 & 0.186 & 0.125 & 0.187 & 0.125 \\
+conventional                          & & 0.139 & 0.114 & 0.142 & 0.106 & 0.187 & 0.125 \\
+conventional (z-score norm.)          & & 0.139 & 0.113 & 0.142 & 0.104 & 0.142 & 0.104 \\
+conventional (robust scaling)         & & 0.139 & 0.114 & 0.141 & 0.104 & 0.141 & 0.104 \\
+Raymaekers-Rousseeuw                  & & 0.211 & 0.121 & 0.183 & 1.212 & 0.187 & 0.125 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.129 & 0.175 & 0.125 & 0.175 & 0.125 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.209 & 0.124 & 0.172 & 0.116 & 0.172 & 0.116 \\
+invariant                             & & 0.141 & 0.115 & 0.140 & 0.106 & 0.140 & 0.106 \\
+robust invariant                      & & 0.141 & 0.100 & 0.140 & 0.095 & 0.140 & 0.090 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# 30 samples or fewer
+data <- data[n_samples >= 1000L, ]
+
+# Add rank based on residual error for central normality.
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "norm_residuals" := residuals / n_samples]
+data[outlier == TRUE, "norm_residuals" := residuals_central / n_samples]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90),
+    "avg. norm. res." = round(mean(norm_residuals), digits = 3)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+# Table with average rank.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
+# Table with average normalised residuals.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg. norm. res.")
+
+```
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 33037 sequences each. Each sequence contained between 1000 and 10000 samples.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         8.85  &         8.14  &         7.35  &         6.29  &         7.41  &         6.87 \\
+conventional                          & &         3.50  &         5.65  &         3.52  &         7.78  &         6.42  &         6.05 \\
+conventional (z-score norm.)          & &         5.24  &         5.94  &         7.34  &         4.98  &         6.26  &         4.79 \\
+conventional (robust scaling)         & &         3.53  &         6.98  &         5.97  &         5.79  &         4.87  &         5.65 \\
+Raymaekers-Rousseeuw                  & &         4.15  &         2.81  &         3.37  &         3.72  &         6.19  &         5.79 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.34  & \textbf{1.86} &         6.57  &         3.39  &         5.49  &         3.30 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.20  &         2.33  &         5.21  &         3.36  &         4.12  &         3.23 \\
+invariant                             & & \textbf{2.36} &         7.59  & \textbf{2.66} &         6.82  & \textbf{1.98} &         6.81 \\
+robust invariant                      & &         7.83  &         3.70  &         3.01  & \textbf{2.88} &         2.26  & \textbf{2.51} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained between 1000 and 10000 samples. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.186 & 0.073 & 0.106 & 0.043 & 0.106 & 0.043 \\
+conventional                          & & 0.014 & 0.060 & 0.060 & 0.046 & 0.106 & 0.043 \\
+conventional (z-score norm.)          & & 0.021 & 0.061 & 0.070 & 0.040 & 0.070 & 0.040 \\
+conventional (robust scaling)         & & 0.014 & 0.061 & 0.069 & 0.040 & 0.069 & 0.040 \\
+Raymaekers-Rousseeuw                  & & 0.015 & 0.017 & 0.060 & 0.024 & 0.106 & 0.043 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.016 & 0.070 & 0.024 & 0.070 & 0.024 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.015 & 0.016 & 0.068 & 0.024 & 0.068 & 0.024 \\
+invariant                             & & 0.013 & 0.067 & 0.059 & 0.044 & 0.059 & 0.044 \\
+robust invariant                      & & 0.034 & 0.034 & 0.061 & 0.026 & 0.061 & 0.026 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
 ## Examples using clean data

### 2025-10-17 manuscript_rmarkdown/manuscript_appendix.Rmd
Update ranking results for Box-Cox transformations.

@@
 Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
@@
 
-none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
-conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
-conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
-conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
-Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
-Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
-Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
-invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
-robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.47  &         6.78 \\
+conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.10 \\
+conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.34  &         5.84 \\
+conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.35  &         5.84 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.30  &         5.80 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.48  &         3.50 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.53  &         3.51 \\
+invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.51} &         5.58 \\
+robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} & \textbf{3.51} & \textbf{2.00} \\
 

### 2025-10-17 manuscript_rmarkdown/manuscript.Rmd
Updated ranking experiment with 10x more sequences.

@@
 
-Three datasets with 10000 sequences each were created to assess transformation
+Three datasets with 100000 sequences each were created to assess transformation
 to normality. For each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
@@
 errors of the central portion (data with outliers). The average rank of each
-transformation method was computed over all 10000 sequences in each dataset.
+transformation method was computed over all 100000 sequences in each dataset.
 Average ranks for Yeo-Johnson transformations are shown in Table
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
@@
 
-none                                  & &         8.51  &         7.60  &         7.33  &         6.15  &         7.44  &         6.60 \\
-conventional                          & &         3.70  &         5.70  &         3.87  &         7.06  &         6.49  &         5.93 \\
-conventional (z-score norm.)          & &         4.72  &         5.68  &         6.38  &         5.01  &         5.43  &         4.87 \\
-conventional (robust scaling)         & &         4.06  &         6.77  &         5.37  &         5.75  &         4.41  &         5.63 \\
-Raymaekers-Rousseeuw                  & &         4.50  &         3.20  &         4.01  &         4.13  &         6.27  &         5.65 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         5.27  & \textbf{2.66} &         6.23  &         3.59  &         5.28  &         3.48 \\
-Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.95  &         5.32  &         3.77  &         4.36  &         3.66 \\
-invariant                             & & \textbf{3.13} &         6.77  & \textbf{3.22} &         6.12  & \textbf{2.64} &         6.03 \\
-robust invariant                      & &         6.41  &         3.66  &         3.27  & \textbf{3.41} &         2.68  & \textbf{3.10} \\
+none                                  & &         8.52  &         7.62  &         7.35  &         6.17  &         7.46  &         6.62 \\
+conventional                          & &         3.82  &         5.78  &         3.86  &         7.08  &         6.51  &         5.95 \\
+conventional (z-score norm.)          & &         4.70  &         5.75  &         6.44  &         5.07  &         5.48  &         4.93 \\
+conventional (robust scaling)         & &         3.83  &         6.61  &         5.36  &         5.72  &         4.41  &         5.59 \\
+Raymaekers-Rousseeuw                  & &         4.64  &         3.22  &         4.01  &         4.11  &         6.28  &         5.67 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.24  & \textbf{2.66} &         6.25  &         3.62  &         5.29  &         3.50 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.99  &         5.22  &         3.69  &         4.27  &         3.57 \\
+invariant                             & & \textbf{3.11} &         6.74  & \textbf{3.24} &         6.13  & \textbf{2.63} &         6.03 \\
+robust invariant                      & &         6.44  &         3.64  &         3.27  & \textbf{3.40} &         2.65  & \textbf{3.14} \\
 

### 2025-10-17 manuscript_rmarkdown/manuscript_appendix.Rmd
Added data on residual errors for yeo-johnson transformation methods.

@@
 
-## Ranking Yeo-Johnson transformations (pt. II)
+## Ranking Yeo-Johnson transformations: residual errors
+
+In the main manuscript, transformation methods are compared by ranking within
+each experiment, defined by the combination of sequence and absence/presence of
+outliers. This allows for direct comparison of methods between different
+experiments.
+
+Table \ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual} shows
+the average normalised residual error for these transformation methods. This
+metric is computed from the residual error and the residual error of the central
+portion ($\kappa = 0.70$) for sequences without and with outliers, respectively.
+In each experiment, the respective residual error was normalised by dividing with
+the sequence length, and can be interpreted as the average residual error of a
+single sample. This reduces the effect of sequence length on the residual error.
+
+Table \ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff} shows
+the average normalised difference in residual error between the best method for each
+experiment and the individual methods. As above, normalisation was done by
+division by the length of the sequence corresponding to each experiment.
+
+Both tables show that the differences between best-performing methods on each
+dataset are relatively small.
+
+```{r include = FALSE, eval = FALSE}
+
+# Yeo-Johnson on clean and dirty data.
+clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
+dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
+shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")
+
+data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
+data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))
+
+# Add rank based on residual error for central normality.
+data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "norm_residuals" := residuals / n_samples]
+data[outlier == TRUE, "norm_residuals" := residuals_central / (0.8 * n_samples)]
+
+data[outlier == FALSE, "min_residual" := min(residuals), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == TRUE, "min_residual" := min(residuals_central), by = c("experiment_id", "outlier", "dataset")]
+data[outlier == FALSE, "residual_delta" := (residuals - min_residual) / n_samples]
+data[outlier == TRUE, "residual_delta" := (residuals_central - min_residual) / (0.8 * n_samples)]
+
+table_data <- data[
+  ,
+  list(
+    "avg." = round(mean(method_rank), digits = 2L),
+    "10%" = quantile(method_rank, 0.10),
+    "50%" = quantile(method_rank, 0.50),
+    "90%" = quantile(method_rank, 0.90),
+    "avg. norm. res." = round(mean(norm_residuals), digits = 3),
+    "avg. res. diff." = round(mean(residual_delta), digits = 3)
+  ),
+  by = c("transformation", "outlier", "dataset")
+]
+
+# Table with average rank --> this was already presented in the main manuscript.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
+
+# Table with average normalised residuals.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg. norm. res.")
+
+# Table with difference with best-performing method in an experiment.
+data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg. res. diff.")
+```
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.202 & 0.112 & 0.128 & 0.080 & 0.129 & 0.080 \\
+conventional                          & & 0.058 & 0.094 & 0.085 & 0.078 & 0.129 & 0.080 \\
+conventional (z-score norm.)          & & 0.061 & 0.094 & 0.091 & 0.073 & 0.091 & 0.073 \\
+conventional (robust scaling)         & & 0.058 & 0.094 & 0.090 & 0.073 & 0.090 & 0.073 \\
+Raymaekers-Rousseeuw                  & & 0.071 & 0.055 & 0.092 & 0.279 & 0.129 & 0.080 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.087 & 0.057 & 0.097 & 0.060 & 0.097 & 0.060 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.071 & 0.056 & 0.095 & 0.059 & 0.095 & 0.059 \\
+invariant                             & & 0.057 & 0.099 & 0.083 & 0.075 & 0.083 & 0.075 \\
+robust invariant                      & & 0.069 & 0.066 & 0.084 & 0.058 & 0.084 & 0.058 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average difference in residual error between Yeo-Johnson
+transformation methods and the best transformation method within each
+experiment, normalised by the corresponding sequence length. This is based on
+either the residual error (without outliers) or the residual error of the
+central portion (with outliers; $\kappa = 0.80$) over 3 datasets with 100000
+sequences each. Residual errors were normalised by sequence length to reduce the
+effect of sequence length on the residual error. The clean dataset consists of
+sequences derived through inverse Yeo-Johnson transformation of data sampled
+from a standard normal distribution. The dirty dataset contains sequences
+sampled from asymmetric generalised normal distributions, centred at 0. The
+shifted dataset also contains sequences sampled from asymmetric generalised
+normal distributions, but centred at 100, and scaled by 0.001. Several
+transformation methods include normalisation before transformation, indicated by
+z-score normalisation (norm.) or robust scaling. A lower difference is better,
+with a minimum value of 0.0. Discrepancies with the rank-based method may occur,
+as normalised differences in residual error are not fully comparable between
+experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.150 & 0.068 & 0.050 & 0.033 & 0.049 & 0.033 \\
+conventional                          & & 0.005 & 0.049 & 0.006 & 0.032 & 0.049 & 0.033 \\
+conventional (z-score norm.)          & & 0.008 & 0.049 & 0.013 & 0.026 & 0.011 & 0.025 \\
+conventional (robust scaling)         & & 0.005 & 0.050 & 0.012 & 0.026 & 0.010 & 0.026 \\
+Raymaekers-Rousseeuw                  & & 0.018 & 0.011 & 0.014 & 0.233 & 0.049 & 0.033 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.034 & 0.012 & 0.018 & 0.014 & 0.017 & 0.013 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.018 & 0.011 & 0.017 & 0.012 & 0.015 & 0.011 \\
+invariant                             & & 0.004 & 0.054 & 0.005 & 0.029 & 0.003 & 0.028 \\
+robust invariant                      & & 0.016 & 0.021 & 0.006 & 0.011 & 0.004 & 0.010 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+## Ranking Yeo-Johnson transformations: effect of sequence length
 
@@
 data[outlier == FALSE, "norm_residuals" := residuals / n_samples]
-data[outlier == TRUE, "norm_residuals" := residuals_central / n_samples]
+data[outlier == TRUE, "norm_residuals" := residuals_central / (0.8 * n_samples)]
 
@@
 
-none                                  & & 0.246 & 0.139 & 0.186 & 0.125 & 0.187 & 0.125 \\
-conventional                          & & 0.139 & 0.114 & 0.142 & 0.106 & 0.187 & 0.125 \\
-conventional (z-score norm.)          & & 0.139 & 0.113 & 0.142 & 0.104 & 0.142 & 0.104 \\
-conventional (robust scaling)         & & 0.139 & 0.114 & 0.141 & 0.104 & 0.141 & 0.104 \\
-Raymaekers-Rousseeuw                  & & 0.211 & 0.121 & 0.183 & 1.212 & 0.187 & 0.125 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.129 & 0.175 & 0.125 & 0.175 & 0.125 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.209 & 0.124 & 0.172 & 0.116 & 0.172 & 0.116 \\
-invariant                             & & 0.141 & 0.115 & 0.140 & 0.106 & 0.140 & 0.106 \\
-robust invariant                      & & 0.141 & 0.100 & 0.140 & 0.095 & 0.140 & 0.090 \\
+none                                  & & 0.246 & 0.173 & 0.186 & 0.156 & 0.187 & 0.157 \\
+conventional                          & & 0.139 & 0.142 & 0.142 & 0.132 & 0.187 & 0.157 \\
+conventional (z-score norm.)          & & 0.139 & 0.142 & 0.142 & 0.130 & 0.142 & 0.130 \\
+conventional (robust scaling)         & & 0.139 & 0.142 & 0.141 & 0.130 & 0.141 & 0.130 \\
+Raymaekers-Rousseeuw                  & & 0.211 & 0.151 & 0.183 & 1.515 & 0.187 & 0.157 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.161 & 0.175 & 0.156 & 0.175 & 0.156 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.209 & 0.155 & 0.172 & 0.144 & 0.172 & 0.144 \\
+invariant                             & & 0.141 & 0.144 & 0.140 & 0.132 & 0.140 & 0.132 \\
+robust invariant                      & & 0.141 & 0.125 & 0.140 & 0.119 & 0.140 & 0.119 \\
 
@@
 
-none                                  & & 0.186 & 0.073 & 0.106 & 0.043 & 0.106 & 0.043 \\
-conventional                          & & 0.014 & 0.060 & 0.060 & 0.046 & 0.106 & 0.043 \\
-conventional (z-score norm.)          & & 0.021 & 0.061 & 0.070 & 0.040 & 0.070 & 0.040 \\
-conventional (robust scaling)         & & 0.014 & 0.061 & 0.069 & 0.040 & 0.069 & 0.040 \\
-Raymaekers-Rousseeuw                  & & 0.015 & 0.017 & 0.060 & 0.024 & 0.106 & 0.043 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.016 & 0.070 & 0.024 & 0.070 & 0.024 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.015 & 0.016 & 0.068 & 0.024 & 0.068 & 0.024 \\
-invariant                             & & 0.013 & 0.067 & 0.059 & 0.044 & 0.059 & 0.044 \\
-robust invariant                      & & 0.034 & 0.034 & 0.061 & 0.026 & 0.061 & 0.026 \\
+none                                  & & 0.186 & 0.091 & 0.106 & 0.053 & 0.106 & 0.053 \\
+conventional                          & & 0.014 & 0.076 & 0.060 & 0.057 & 0.106 & 0.053 \\
+conventional (z-score norm.)          & & 0.021 & 0.076 & 0.070 & 0.050 & 0.070 & 0.050 \\
+conventional (robust scaling)         & & 0.014 & 0.076 & 0.069 & 0.050 & 0.069 & 0.050 \\
+Raymaekers-Rousseeuw                  & & 0.015 & 0.021 & 0.060 & 0.030 & 0.106 & 0.053 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.020 & 0.070 & 0.030 & 0.070 & 0.030 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.015 & 0.020 & 0.068 & 0.030 & 0.068 & 0.030 \\
+invariant                             & & 0.013 & 0.084 & 0.059 & 0.055 & 0.059 & 0.055 \\
+robust invariant                      & & 0.034 & 0.042 & 0.061 & 0.032 & 0.061 & 0.032 \\
 

### 2025-10-17 manuscript_rmarkdown/manuscript.Rmd
Insert sentence on presence of shift parameter in Box and Cox (1964).

@@
 
-Box-Cox and Yeo-Johnson transformations are modified by introducing
+@Box1964-mz originally included a shift parameter in their transformation, but
+for their formal analysis set it to $0$, whereas @Yeo2000-vw did not include
+a shift or scale parameter. Here, Box-Cox and Yeo-Johnson transformations are modified by introducing
 shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}.

### 2025-10-17 manuscript_rmarkdown/manuscript.Rmd
Initial work on figure 5.

@@
   k = 0.80,
+  outlier_rate = 0.00,
   for_manuscript = FALSE
@@
   k = 0.80,
-  for_manuscript = FALSE,
-  with_outliers = TRUE
+  for_manuscript = FALSE
 )

### 2025-10-20 manuscript_rmarkdown/manuscript.Rmd
Reworked the empirical-central-normality-kappa plot (Figure 4)

@@
 
-## Central normality and empirical central normality test
+## Empirical central normality test
 
 To develop a test for central normality we need to consider two
-parameters: the central portion $\kappa$ as a fixed parameter, and test
-statistic $\tau_{\text{ecn}}$. We will first define the central portion
+parameters: the central portion $\kappa$ and outlier rate $\zeta$ as fixed parameters,
+and test statistic $\tau_{\text{ecn}}$. We will first define the central portion
 $\kappa$.
@@
 This dataset was used to compute the critical test statistic values for the central normality test.
-Additionally, we added $10 \%$ outliers by randomly replacing elements of each ordered sequence.
-The dataset with outliers was used to compute the critical test statistic values for the empirical variant of the central normality test.
-For each sequence we then computed $\tau_{n, \kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
+Additionally, outliers were inserted by randomly replacing each element of an ordered sequence with a probability $\zeta$.
+We then computed the critical test statistic values for the empirical central normality test.
+For each sequence we then computed $\tau_{n, \kappa, \zeta}$ for 
+$\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$ and
+$\zeta \in \left\{0.0, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20 \right\}$.
 
@@
 
-Figure \ref{fig:empirical-central-normality-kappa} shows $\tau_{\alpha = 0.05, n, \kappa}$
-as a function of $n$ for different values of $\kappa$. With decreasing
+Figure \ref{fig:empirical-central-normality-kappa} shows $\tau_{\alpha = 0.05, n, \kappa, \zeta}$
+as a function of $n$ for different values of $\kappa$ and $\zeta$. With decreasing
 $\kappa$, the test statistic curve decreases. The curves
@@
 
-Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
+Since the empirical central normality tests assesses whether the central part of a sequence is normally distributed,
 we used $\kappa = 0.80$. Critical statistic values for central normality and empirical central
@@
 cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$ of the (empirical) ",
-  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
-  "The critical test statistics for central normality test are determined using fully normal data, ",
-  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
-  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
+  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa, \\zeta}$ of the empirical ",
+  "central normality test as function of $n$ for several values of central portion $\\kappa$ ",
+  "and outlier rates $\\zeta$. ",
   "emp: empirical"

### 2025-10-20 manuscript_rmarkdown/plots.R

@@ -2349,22 +2349,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
 .plot_test_statistic_centrality <- function(manuscript_dir, plot_theme) {
   # Plots data that allows us to select the fitting centrality parameter kappa.
 
-  data_w_outliers <- .get_test_statistics_data(
-    manuscript_dir = manuscript_dir,
-    with_outliers = TRUE
+  data <- .get_test_statistics_data(
+    manuscript_dir = manuscript_dir
   )
-  data_w_outliers[, "outlier" := TRUE]
-
-  data_wo_outliers <- .get_test_statistics_data(
-    manuscript_dir = manuscript_dir,
-    with_outliers = FALSE
-  )
-  data_wo_outliers[, "outlier" := FALSE]
-
-  data <- rbind(data_w_outliers, data_wo_outliers)
 
   # Compute alpha
-  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
+  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier_rate")]
 
   # Compute alpha = 0.95
   data <- data[, list("tau" = stats::spline(
@@ -2373,12 +2363,18 @@ get_annotation_settings <- function(ggtheme = NULL) {
     method = "fmm",
     xout = 0.05
   )$y),
-  by = c("n", "kappa", "outlier")]
+  by = c("n", "kappa", "outlier_rate")]
 
   data$kappa <- factor(
     x = data$kappa,
     levels = c(0.60, 0.70, 0.80, 0.90, 0.95, 1.00),
     labels = c("60 %", "70 %", "80 %", "90 %", "95 %", "100 %")
+  )
+
+  data$outlier_rate <- factor(
+    x = data$outlier_rate,
+    levels = c(0.00, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20),
+    labels = c("0 %", "1 %", "2 %", "5 %", "10 %", "15 %", "20 %")
   )
 
   p <- ggplot2::ggplot(
@@ -2390,9 +2386,9 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p <- p + plot_theme
   p <- p + ggplot2::scale_x_log10(name = "n")
[…]

### 2025-10-21 manuscript_rmarkdown/manuscript.Rmd
Rework of empirical central normality test to include outlier rate, and kappa.

@@
 
-central_normality_lookup_table <- .get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  outlier_rate = 0.00,
-  for_manuscript = FALSE
-)
-
-empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = FALSE
+ecn_lookup_table <- .get_test_statistic_lookup_table(
+  manuscript_dir = manuscript_dir
 )
@@
 usethis::use_data(
-  central_normality_lookup_table,
-  empirical_central_normality_lookup_table,
+  ecn_lookup_table,
   two_sided_function_parameters, 

### 2025-10-21 manuscript_rmarkdown/manuscript_appendix.Rmd
Rework of empirical central normality test to include outlier rate, and kappa.

@@
 
-Critical values for the central normality test and its empirical variant are
+Critical values for the empirical central normality test are
 found in Table \ref{tab:central-normality-critical-statistic} and Table 
@@
 cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80}$ of the (empirical) ",
+  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80, \\zeta}$ of the empirical ",
   "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
-  "The critical test statistics for central normality test are determined using fully normal data, ",
-  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
-  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
   "emp: empirical"

### 2025-10-21 R/R/GoodnessOfFit.R

@@ -12,7 +12,7 @@ NULL
 #' distributed. The null-hypothesis is that the transformed distribution is
 #' centrally normal.
 #'
-#' This function is a wrapper around `ecn.test`.
+#' This function is a wrapper around `ecn_test`.
 #'
 #' @param transformer A transformer object created using
 #'   `find_transformation_parameters`.
@@ -39,8 +39,15 @@ assess_transformation <- function(
     verbose = TRUE,
     ...) {
 
-  # Perform test.
-  h <- ecn.test(x = x, transformer = transformer)
+  # Perform test. Note that assess_transformation is interested only on the
+  # p-value of the test, hence we try to capture any errors, and handle them
+  # gracefully.
+  h <- tryCatch(
+    ecn_test(x = x, transformer = transformer, ...),
+    error = identity
+  )
+
+  if (is(h, "simpleError")) h <- list("p_value" = NA_real_)
 
   if (is.na(h$p_value)) {
     message("p-value could not be determined.")
@@ -71,81 +78,270 @@ assess_transformation <- function(
 #'   significance level = 0.95) of the test statistic (`tau_critical`) and
 #'   p-value (`p_value`) for the empirical central normality test.
 #' @export
-ecn.test <- function(x, transformer = NULL) {
-  return(cn.test(x = x, transformer = transformer, robust = TRUE))
-}
-
-
-
-
-#' Central normality test
-#'
-#' Assesses central normality of input data.
-#'
-#' @param x vector of input data, of at least length 5.
-#' @param transformer A transformer object created using
-#'   `find_transformation_parameters`. Optional, if present residuals are
-#'   determined from x after transformation.
-#' @param robust Determines the use of the strict test for central normality
-#'   (`FALSE`) or the more robust version where test statistics where determined
-#'   from normally distributed data with 10% outliers (`TRUE`).
-#'
-#' @return list with mean absolute error (`tau`), critical value (at
-#'   significance level = 0.95) of the test statistic (`tau_critical`) and
-#'   p-value (`p_value`) for (empirical) central normality test.
-#' @export
-cn.test <- function(x, transformer = NULL, robust = FALSE) {
-
[…]

### 2025-10-21 manuscript_rmarkdown/plots.R

@@ -2440,25 +2440,15 @@ get_annotation_settings <- function(ggtheme = NULL) {
 .plot_test_statistic_tau <- function(manuscript_dir, plot_theme, k = 0.80) {
   # Plots critical tau for several alpha levels.
 
-  data_w_outliers <- .get_test_statistics_data(
-    manuscript_dir = manuscript_dir,
-    with_outliers = TRUE
+  data <- .get_test_statistics_data(
+    manuscript_dir = manuscript_dir
   )
-  data_w_outliers[, "outlier" := TRUE]
-
-  data_wo_outliers <- .get_test_statistics_data(
-    manuscript_dir = manuscript_dir,
-    with_outliers = FALSE
-  )
-  data_wo_outliers[, "outlier" := FALSE]
-
-  data <- rbind(data_w_outliers, data_wo_outliers)
   data <- data[kappa == k]
 
   alpha_levels <- 1.0 - c(0.80, 0.90, 0.95, 0.975, 0.99, 0.999)
 
   # Compute alpha
-  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier")]
+  data[, "alpha" := 1.0 - (seq_len(.N) - 1L) / (.N - 1L), by = c("n", "kappa", "outlier_rate")]
 
   # Compute tau at the specified confidence levels.
   data <- data[
@@ -2471,13 +2461,19 @@ get_annotation_settings <- function(ggtheme = NULL) {
         xout = alpha_levels)$y,
       "alpha" = alpha_levels
     ),
-    by = c("n", "outlier")
+    by = c("n", "outlier_rate")
   ]
 
   data$alpha <- factor(
     x = data$alpha,
     levels = alpha_levels,
     labels = c("20.0 %", "10.0 %", "5.0 %", "2.5 %", "1.0 %", "0.1 %")
+  )
+
+  data$outlier_rate <- factor(
+    x = data$outlier_rate,
+    levels = c(0.00, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20),
+    labels = c("0 %", "1 %", "2 %", "5 %", "10 %", "15 %", "20 %")
   )
 
   p <- ggplot2::ggplot(
@@ -2489,9 +2485,9 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
   p <- p + plot_theme
[…]

### 2025-10-21 manuscript_rmarkdown/simulations.R

@@ -1716,51 +1716,41 @@
 
 
 .get_test_statistic_lookup_table <- function(
-    manuscript_dir,
-    k = 0.80,
-    outlier_rate = 0.10,
-    for_manuscript = TRUE,
-    with_outliers = FALSE
+    manuscript_dir
 ) {
-  if (for_manuscript) {
-    alpha_levels <- rev(1.0 - c(0.10, 0.20, 0.50, 0.80, 0.90, 0.95, 0.975, 0.99, 0.999))
-    n <- c(5, 10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000)
-  } else {
-    alpha_levels <- rev(1.0 - c(
-      0.00, 0.01, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30, 0.35, 0.40, 0.45, 0.50, 0.55,
-      0.60, 0.65, 0.70, 0.75, 0.80, 0.85, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95,
-      0.96, 0.97, 0.975, 0.98, 0.9825, 0.985, 0.9875, 0.99,
-      0.991, 0.992, 0.993, 0.994, 0.995, 0.996, 0.997, 0.998, 0.999, 0.9995
-    ))
-    n <- c(
-      5, 10, 15, 20, 30, 40, 50, 60, 70, 80, 90, 100, 120, 140, 160, 180, 200,
-      225, 250, 275, 300, 340, 380, 420, 460, 500,
-      560, 620, 680, 740, 800, 900, 1000, 1100, 1200, 1500,
-      2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000)
-  }
+
+  alpha_levels <- rev(1.0 - c(
+    0.00, 0.01, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30, 0.35, 0.40, 0.45, 0.50, 0.55,
+    0.60, 0.65, 0.70, 0.75, 0.80, 0.85, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95,
+    0.96, 0.97, 0.975, 0.98, 0.9825, 0.985, 0.9875, 0.99,
+    0.991, 0.992, 0.993, 0.994, 0.995, 0.996, 0.997, 0.998, 0.999, 0.9995
+  ))
 
   data <- .get_test_statistics_data(manuscript_dir)
-  data <- data[kappa == k]
-  data[, "kappa" := NULL]
 
[…]

### 2025-10-21 manuscript_rmarkdown/manuscript.Rmd
Update lookup table.

@@
 Since the empirical central normality tests assesses whether the central part of a sequence is normally distributed,
-we used $\kappa = 0.80$. Critical statistic values for central normality and empirical central
-normality tests are shown are shown in Table \ref{tab:central-normality-critical-statistic} and Table 
-\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+we used $\kappa = 0.80$. Critical statistic values for $\kappa = 0.80$ and $\zeta = 0.10$ are shown in Table 
+\ref{tab:empirical-central-normality-critical-statistic}.
 

### 2025-10-21 manuscript_rmarkdown/manuscript_appendix.Rmd
Update lookup table.

@@
 # This generates the table shown below.
-.get_test_statistic_lookup_table(
+.get_test_statistic_lookup_table_manuscript(
   manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = TRUE
+  kappa_lookup = 0.80,
+  outlier_rate_lookup = 0.10
 )
@@
 \begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
-\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80, \zeta = 0.10}$ for
+central normality at $\kappa = 0.80$ and outlier rate $\zeta = 0.10$, as a function of significance 
 level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
-\label{tab:central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
-
-\midrule
-    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
-   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
-   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
-   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
-  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
-  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
-  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
- 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
- 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
- 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
-10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-```{r include = FALSE, eval = FALSE}
-# This generates the table shown below.
-.get_test_statistic_lookup_table(
-  manuscript_dir = manuscript_dir,
-  k = 0.80,
-  for_manuscript = TRUE,
-  with_outliers = TRUE
-)
-
-```
-
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$.The values in this table should be divided by $100$ before use.}
 \label{tab:empirical-central-normality-critical-statistic}
@@
 \midrule
-    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
-   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
-   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
-   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
-  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
-  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
-  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
- 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
- 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
- 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
-10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
+    5 & 210.38 & 82.23 & 59.47 & 48.68 & 38.71 & 29.28 & 19.75 & 13.71 & 11.15 \\
+   10 &  43.22 & 33.68 & 29.26 & 26.04 & 22.96 & 19.62 & 14.37 & 10.50 &  8.89 \\
+   20 &  31.31 & 25.35 & 22.61 & 20.58 & 18.23 & 15.64 & 11.51 &  8.61 &  7.49 \\
+   50 &  19.97 & 16.53 & 14.83 & 13.33 & 11.77 & 10.11 &  7.60 &  5.85 &  5.16 \\
+  100 &  14.01 & 11.57 & 10.30 &  9.31 &  8.31 &  7.22 &  5.49 &  4.29 &  3.78 \\
+  200 &   9.87 &  8.15 &  7.39 &  6.73 &  6.02 &  5.28 &  4.06 &  3.16 &  2.80 \\
+  500 &   6.30 &  5.34 &  4.87 &  4.49 &  4.08 &  3.60 &  2.80 &  2.17 &  1.92 \\
+ 1000 &   4.72 &  4.02 &  3.70 &  3.43 &  3.15 &  2.81 &  2.21 &  1.72 &  1.51 \\
+ 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.88 &  1.49 &  1.31 \\
+ 5000 &   2.80 &  2.51 &  2.37 &  2.25 &  2.12 &  1.96 &  1.66 &  1.38 &  1.24 \\
+10000 &   2.41 &  2.20 &  2.10 &  2.02 &  1.92 &  1.81 &  1.60 &  1.39 &  1.28 \\
 \bottomrule

### 2025-10-24 manuscript_rmarkdown/manuscript.Rmd
Update section 3.2 and Figure 5

@@
 
-To develop a test for central normality we need to consider two
-parameters: the central portion $\kappa$ and outlier rate $\zeta$ as fixed parameters,
-and test statistic $\tau_{\text{ecn}}$. We will first define the central portion
-$\kappa$.
-
+The critical test statistic for the empirical central normality test depends on 
+the significance level $\alpha$, sequence length $n$, central portion $\kappa$
+and estimated outlier rate $\zeta$, and was set through simulation. 
 For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
-we randomly drew $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
-This dataset was used to compute the critical test statistic values for the central normality test.
-Additionally, outliers were inserted by randomly replacing each element of an ordered sequence with a probability $\zeta$.
-We then computed the critical test statistic values for the empirical central normality test.
-For each sequence we then computed $\tau_{n, \kappa, \zeta}$ for 
-$\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$ and
+we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
+Outliers were inserted by randomly replacing elements of an ordered sequence with probability 
 $\zeta \in \left\{0.0, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20 \right\}$. 
+Outlier values were set according to @Tukey1977-xm, as follows.
+Let $x^{*} \sim U\left(-2, 2\right)$.
+Then the corresponding outlier value was:
+
+\begin{equation}
+\label{eqn:outlier_rate}
+x_{out} =
+\begin{cases}
+Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
+Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
+\end{cases}
+\end{equation}
+
+Here, $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
+
+For each sequence we then computed $\tau_{\kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
 
@@
 
-Figure \ref{fig:empirical-central-normality-kappa} shows $\tau_{\alpha = 0.05, n, \kappa, \zeta}$
-as a function of $n$ for different values of $\kappa$ and $\zeta$. With decreasing
-$\kappa$, the test statistic curve decreases. The curves
-for $\kappa = 0.60$, $\kappa = 0.70$ and $\kappa = 0.80$ are similar, whereas curves for 
-$\kappa \geq 0.90$ are affected by outliers, as can be observed by comparing curves
-of the central normality test with those of the empirical variant.
-
-Since the empirical central normality tests assesses whether the central part of a sequence is normally distributed,
-we used $\kappa = 0.80$. Critical statistic values for $\kappa = 0.80$ and $\zeta = 0.10$ are shown in Table 
+Figure \ref{fig:empirical-central-normality-kappa} shows the critical test static
+$\tau_{\alpha = 0.05, n, \kappa, \zeta}$
+as a function of $n$ for different values of $\kappa$ and $\zeta$.
+With decreasing $\kappa$, the test statistic curve decreases.
+The curves for $\kappa = 0.60$, $\kappa = 0.70$ and $\kappa = 0.80$ are similar,
+whereas curves for $\kappa \geq 0.90$ were affected by outliers.
+
+We use $\kappa = 0.80$ and $\zeta = 0.10$ for the empirical central normality tests 
+in the remainder of the manuscript, unless noted otherwise.
+Critical statistic values of the empirical central normality test for 
+$\kappa = 0.80$ and $\zeta = 0.10$ are shown in Table 
 \ref{tab:empirical-central-normality-critical-statistic}.
@@
   "central normality test as function of $n$ for several values of central portion $\\kappa$ ",
-  "and outlier rates $\\zeta$. ",
+  "and estimated outlier rates $\\zeta$. ",
   "emp: empirical"
@@
 
-To assess type I error rates for the (empirical) central normality test and compare
-these to the Shapiro-Wilk test for normality, we randomly drew another 1000 sequences from a
+To assess type I error rates for the empirical central normality test 
+($\kappa = 0.80$ and $\zeta \in \left\{0.00, 0.05, 0.10\right\}$) and compare
+these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
 standard normal distribution ($\mathcal{N}(0,1)$ for each 
 $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
-A second dataset was created by replacing elements with randomly drawn outliers, as described above.
+Each sequence was additionally randomly populated with outliers at 5% and 10% rates, as earlier.
+
 For each sequence, the p-value for the respective test was used to reject the null
@@
 The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
-Figure \ref{fig:empirical-central-normality-error_rate} shows that the empirical central normality test had the expected type I error rate for
-centrally normal data (with tails of the distribution not being normally distributed),
-whereas the central normality test and Shapiro-Wilk test overestimated type I error rates.
+Figure \ref{fig:empirical-central-normality-error_rate} shows that when the estimated outlier rate 
+for the empirical central normality test matches the outlier rate in the data,
+the test yields an accurate Type I error rate.
+The Shapiro-Wilk test is overly conservative in the presence of any outliers. 
+Empirical central normality tests where the estimated outlier rate is lower
+than the actual outlier rate are likewise conservative, but to a lesser degree. 
+Empirical central normality tests where the estimated outlier rate is higher 
+than the actual outlier rate are optimistic.
 
@@
 cap <- paste0(
-  "Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers. ",
-  "The empirical central normality (ECN) test uses $\\kappa = 0.80$. CN: central normality"
+  "Observed type I error rates for central normality tests in the presence of data sampled from normal distributions ",
+  "without outliers, or with a 5% or 10% outlier rate. ",
+  "The empirical central normality (ECN) test used $\\kappa = 0.80$, and was further parameterised by the ",
+  "estimated outlier rate $\\zeta = 0.00$ (ECN (0 %)), $\\zeta = 0.05$ (ECN (5 %)) and $\\zeta = 0.10$ (ECN (10%))."  
 )
@@
 with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
-
 Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
-Outlier values were set according to @Tukey1977-xm, as follows.
-Let $x^{*} \sim U\left(-2, 2\right)$.
-Then the corresponding outlier value was:
-
-\begin{equation}
-x_{out} =
-\begin{cases}
-Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
-Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
-\end{cases}
-\end{equation}
-
-$Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
-Outlier values randomly replaced elements in $\mathbf{X}_i$.
 

### 2025-10-24 manuscript_rmarkdown/plots.R

@@ -2542,9 +2542,12 @@ get_annotation_settings <- function(ggtheme = NULL) {
     n
   ) {
     return(
-      lapply(
-        X = n,
-        FUN = ..compute
+      unlist(
+        lapply(
+          X = n,
+          FUN = ..compute
+        ),
+        recursive = FALSE
       )
     )
   }
@@ -2554,7 +2557,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   ) {
     set.seed(n)
 
-    n_rep <- 1000L
+    n_rep <- 10000L
     data <- list()
     for (jj in seq_len(n_rep)) {
       x <- power.transform::ragn(floor(n), location = 0, scale = 1 / sqrt(2), alpha = 0.5, beta = 2)
@@ -2569,33 +2572,38 @@ get_annotation_settings <- function(ggtheme = NULL) {
       x_outlier_10 <- x
 
       # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
-      # than Q3 + 1.5 IQR.
-      n_draw_5 <- ceiling(0.05 * length(x))
-      n_draw_10 <- ceiling(0.10 * length(x))
-      x_random_5 <- stats::runif(n_draw_5, min = -2.0, max = 2.0)
-      x_random_10 <- stats::runif(n_draw_10, min = -2.0, max = 2.0)
-
+      # than Q3 + 1.5 IQR, and use these to randomly replace instances.
+      # We do this for 5% and 10% outlier rates.
+      ii_outlier_5 <- stats::runif(length(x)) <= 0.05
+      n_draw_5 <- sum(ii_outlier_5)
       outlier_5 <- numeric(n_draw_5)
+
+      if (n_draw_5 > 0L) {
+        x_random_5 <- stats::runif(n_draw_5, min = -2.0, max = 2.0)
+        if (any(x_random_5 < 0)) {
+          outlier_5[x_random_5 < 0] <- q_lower - 1.5 * interquartile_range + x_random_5[x_random_5 < 0] * interquartile_range
+        }
+
+        if (any(x_random_5 >= 0)) {
+          outlier_5[x_random_5 >= 0] <- q_upper + 1.5 * interquartile_range + x_random_5[x_random_5 >= 0] * interquartile_range
+        }
+      }
+      x_outlier_5[ii_outlier_5] <- outlier_5
+
+      ii_outlier_10 <- stats::runif(length(x)) <= 0.10
[…]

### 2025-10-24 manuscript_rmarkdown/manuscript.Rmd
Rework section 2.4.

@@
 - \newtheorem*{definition}{Definition}
+- \newtheorem*{corollary}[Corollary]
 - \DeclareMathOperator*{\argmax}{argmax}
@@
 
-## Central normality test
+## Empirical central normality test
 
@@
 In practice, normality tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop a test for central normality.
+Here we develop an empirical test for central normality.
 
@@
 
-\begin{definition}[Central normality of a sequence]
-
-The central portion of a sequence is normally distributed if the sum of residual
-errors is equal to zero:
-$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
-
-\end{definition}
-
-In practice, a finite sequence sampled from a normal distribution $\mathcal{N}(\mu, \sigma)$ will have a non-zero sum of residual errors.
+\begin{definition}[Centrally normal distribution]
+
+A centrally normal distribution is any distribution whose quantile function is 
+defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
+with $\mathcal{N}$ any normal distribution.
+
+\end{definition}
+
+\begin{corollary}
+
+If the sum of residual errors of the central portion of a sequence is 0, i.e.,
+$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$, 
+the values of the elements in the central portion of that sequence follow a 
+centrally normal distribution.
+
+\end{corollary}
+
+By the Glivenko-Cantelli theorem, if sequence $\mathbf{X}$ was sampled from a
+normal distribution $\mathcal{N}(\mu, \sigma)$,
+$\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,
+i.e. an infinite sequence sampled from a normal distribution has a central portion
+with values sampled from a centrally normal distribution.
 
 \begin{definition}[Central normality test]
-The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
-
-The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
-\end{definition}
-
-The null-hypothesis $\mathcal{H}_0$ is that the central portion of sequence $\mathbf{X}$
-is normally distributed, with the alternative hypothesis $\mathcal{H}_1$ that it
-is not normally distributed. To test this hypothesis, a test statistic is computed
-and compared against a critical value.
+The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+from a population that has a centrally normal distribution.
+
+The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+population that lacks a centrally normal distribution.
+\end{definition}
+
+To test this hypothesis, a test statistic is computed and compared against a critical value.
 
 \begin{definition}[Central normality test statistic]
-Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$.
 
@@
 The null-hypothesis should be rejected at significance level $\alpha$ if 
-$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
-
-The central portion of the data needs to be defined and the Type 1 error rates determined to provide critical test statistics.
-We will do so in the [Simulation](#simulation) section.
+$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \zeta, \text{critical}}$,
+with $\zeta$ the population estimate of the rate of outliers. Critical test
+values are determined in the [Simulation](#simulation) section.
 
@@
 the significance level $\alpha$, sequence length $n$, central portion $\kappa$
-and estimated outlier rate $\zeta$, and was set through simulation. 
+and estimated outlier rate $\zeta$.
+Critical test statistic values were set through simulation, as described below.
+
 For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$

### 2025-10-25 manuscript_rmarkdown/manuscript.Rmd
Make definition of central normal distribution clearer.

@@
 
-\begin{definition}[Centrally normal distribution]
-
-A centrally normal distribution is any distribution whose quantile function is 
+\begin{definition}[Central normal distribution]
+
+A central normal distribution is any distribution whose quantile function is 
 defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
@@
 
-If the sum of residual errors of the central portion of a sequence is 0, i.e.,
-$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$, 
-the values of the elements in the central portion of that sequence follow a 
-centrally normal distribution.
+For $\kappa = 1.0$, the central normal distribution is a normal distribution.
 
@@
 
-By the Glivenko-Cantelli theorem, if sequence $\mathbf{X}$ was sampled from a
-normal distribution $\mathcal{N}(\mu, \sigma)$,
+\begin{corollary}
+
+By the Glivenko-Cantelli theorem, if a sequence $\mathbf{X}$ was sampled from a
+central normal distribution, then almost surely
 $\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,
-i.e. an infinite sequence sampled from a normal distribution has a central portion
-with values sampled from a centrally normal distribution.
+
+\end{corollary}
 
@@
 The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
-from a population that has a centrally normal distribution.
+from a population that is distributed according to a central normal distribution.
 
 The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
-population that lacks a centrally normal distribution.
+population that is not distributed according to a central normal distribution.
 \end{definition}
@@
 $\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \zeta, \text{critical}}$,
-with $\zeta$ the population estimate of the rate of outliers. Critical test
-values are determined in the [Simulation](#simulation) section.
+with $\zeta$ the population estimate of the rate of outliers from a normal
+distribution, to be defined in the [Simulation](#simulation) section.
 

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Figure 5 without outlier-rate.

@@
 The null-hypothesis should be rejected at significance level $\alpha$ if 
-$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \zeta, \text{critical}}$,
-with $\zeta$ the population estimate of the rate of outliers from a normal
-distribution, to be defined in the [Simulation](#simulation) section.
+$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
+Critical test statistic values are determined in 
+the [Simulation](#simulation-ecn-test) section.
 
@@
 
-## Empirical central normality test
+## Empirical central normality test {#simulation-ecn-test}
 
 The critical test statistic for the empirical central normality test depends on 
-the significance level $\alpha$, sequence length $n$, central portion $\kappa$
-and estimated outlier rate $\zeta$.
+the significance level $\alpha$, sequence length $n$ and central portion $\kappa$.
 Critical test statistic values were set through simulation, as described below.
@@
 we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
-Outliers were inserted by randomly replacing elements of an ordered sequence with probability 
-$\zeta \in \left\{0.0, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20 \right\}$. 
-Outlier values were set according to @Tukey1977-xm, as follows.
-Let $x^{*} \sim U\left(-2, 2\right)$.
-Then the corresponding outlier value was:
-
-\begin{equation}
-\label{eqn:outlier_rate}
-x_{out} =
-\begin{cases}
-Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
-Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
-\end{cases}
-\end{equation}
-
-Here, $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
-
-For each sequence we then computed $\tau_{\kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
+For each sequence we then computed $\tau_{\kappa}$ for 
+$\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
 
@@
 
-Figure \ref{fig:empirical-central-normality-kappa} shows the critical test static
-$\tau_{\alpha = 0.05, n, \kappa, \zeta}$
-as a function of $n$ for different values of $\kappa$ and $\zeta$.
-With decreasing $\kappa$, the test statistic curve decreases.
-The curves for $\kappa = 0.60$, $\kappa = 0.70$ and $\kappa = 0.80$ are similar,
-whereas curves for $\kappa \geq 0.90$ were affected by outliers.
-
-We use $\kappa = 0.80$ and $\zeta = 0.10$ for the empirical central normality tests 
-in the remainder of the manuscript, unless noted otherwise.
-Critical statistic values of the empirical central normality test for 
-$\kappa = 0.80$ and $\zeta = 0.10$ are shown in Table 
-\ref{tab:empirical-central-normality-critical-statistic}.
-
-```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
+Figure \ref{fig:empirical-central-normality-combined} shows the critical test statistic
+$\tau_{\alpha, n, \kappa}$ as a function of $n$ for different values of $\kappa$ and $\alpha$. 
+For larger central portion $\kappa$, critical test statistic values increase.
+Similarly, critical test statistic values increase for lower significance levels.
+
+```{r empirical-central-normality-combined, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
 cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa, \\zeta}$ of the empirical ",
-  "central normality test as function of $n$ for several values of central portion $\\kappa$ ",
-  "and estimated outlier rates $\\zeta$. ",
-  "emp: empirical"
-)
-
-.plot_test_statistic_centrality(
+  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$ of the empirical ",
+  "central normality test as function of $n$ for several values of central portion $\\kappa$ (left) ",
+  "and several values of significance level $\\alpha$ (right)."
+)
+
+.plot_test_statistic_combined(
   plot_theme = plot_theme,
@@
 
-To assess type I error rates for the empirical central normality test 
-($\kappa = 0.80$ and $\zeta \in \left\{0.00, 0.05, 0.10\right\}$) and compare
+To assess type I error rates for the empirical central normality test and compare
 these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
-standard normal distribution ($\mathcal{N}(0,1)$ for each 
+central normal distribution for each 
 $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
-Each sequence was additionally randomly populated with outliers at 5% and 10% rates, as earlier.
+
+Each sequence was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
+Then, the elements outside the central portion of the sequence (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
+Elements with $pi < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
+and elements with $p_i > \frac{1 + \kappa}{2}$ are replaced with values sampled from $\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)$.
 
@@
 In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
-central normality test to assess central normality of features that are composed
+central normality test (with $\kappa = 0.80$) to assess central normality of features that are composed
 of a mixture of samples drawn from two normal distributions ($n = 100$ each).
-With increased separation of the underlying normal distributions, the probability of the
-feature being centrally normal decreases, as expected.
+With increased separation of the underlying normal distributions, 
+the test's p-value decreases, as expected.
 
@@
   "drawn from normal distributions that are offset by a distance $d$. ",
-  "We use the empirical centrally normality test to compute the probability for the hypothesis that ",
-  "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
-  "the probability that the feature is centrally normal decreases. ",
+  "The empirical central normality ($\\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from ",
+  "a central normal distribution. As may be observed, with increasing offset $d$ ",
+  "the test's p-value decreases. ",
   "Quantile-quantile plots are drawn below each distribution."
@@
 ```
+
+We use $\kappa = 0.80$ for the empirical central normality tests 
+in the remainder of the manuscript, unless noted otherwise.
+Critical statistic values of the empirical central normality test for 
+$\kappa = 0.80$ are shown in Table 
+\ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
+These values were set according to @Tukey1977-xm, as follows.
+Let $x^{*} \sim U\left(-2, 2\right)$.
+Then the corresponding outlier value was:
+
+\begin{equation}
+\label{eqn:outlier_rate}
+x_{out} =
+\begin{cases}
+Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
+Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
+\end{cases}
+\end{equation}
+
+Here, $Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
 

### 2025-10-28 manuscript_rmarkdown/manuscript_appendix.Rmd
Figure 5 without outlier-rate.

@@
 
-Critical values for the empirical central normality test are
-found in Table \ref{tab:central-normality-critical-statistic} and Table 
-\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+Critical values for the empirical central normality test at $\kappa = 0.80$ are
+found in Table \ref{tab:empirical-central-normality-critical-statistic}.
 
@@
   manuscript_dir = manuscript_dir,
-  kappa_lookup = 0.80,
-  outlier_rate_lookup = 0.10
+  kappa_lookup = 0.80
 )
@@
 \begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80, \zeta = 0.10}$ for
-central normality at $\kappa = 0.80$ and outlier rate $\zeta = 0.10$, as a function of significance 
+\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
+central normality at $\kappa = 0.80$, as a function of significance 
 level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
@@
 \midrule
-    5 & 210.38 & 82.23 & 59.47 & 48.68 & 38.71 & 29.28 & 19.75 & 13.71 & 11.15 \\
-   10 &  43.22 & 33.68 & 29.26 & 26.04 & 22.96 & 19.62 & 14.37 & 10.50 &  8.89 \\
-   20 &  31.31 & 25.35 & 22.61 & 20.58 & 18.23 & 15.64 & 11.51 &  8.61 &  7.49 \\
-   50 &  19.97 & 16.53 & 14.83 & 13.33 & 11.77 & 10.11 &  7.60 &  5.85 &  5.16 \\
-  100 &  14.01 & 11.57 & 10.30 &  9.31 &  8.31 &  7.22 &  5.49 &  4.29 &  3.78 \\
-  200 &   9.87 &  8.15 &  7.39 &  6.73 &  6.02 &  5.28 &  4.06 &  3.16 &  2.80 \\
-  500 &   6.30 &  5.34 &  4.87 &  4.49 &  4.08 &  3.60 &  2.80 &  2.17 &  1.92 \\
- 1000 &   4.72 &  4.02 &  3.70 &  3.43 &  3.15 &  2.81 &  2.21 &  1.72 &  1.51 \\
- 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.88 &  1.49 &  1.31 \\
- 5000 &   2.80 &  2.51 &  2.37 &  2.25 &  2.12 &  1.96 &  1.66 &  1.38 &  1.24 \\
-10000 &   2.41 &  2.20 &  2.10 &  2.02 &  1.92 &  1.81 &  1.60 &  1.39 &  1.28 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
-cap <- paste0(
-  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80, \\zeta}$ of the empirical ",
-  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
-  "emp: empirical"
-)
-
-.plot_test_statistic_tau(
-  plot_theme = plot_theme,
-  manuscript_dir = manuscript_dir
-)
-
-```
+    5 & 200.88 & 90.25 & 64.46 & 49.04 & 36.07 & 27.70 & 19.28 & 13.41 & 10.90 \\
+   10 &  45.90 & 31.78 & 27.12 & 24.14 & 21.23 & 18.30 & 13.68 & 10.20 &  8.71 \\
+   20 &  28.88 & 22.53 & 19.96 & 18.01 & 15.99 & 13.93 & 10.67 &  8.26 &  7.27 \\
+   50 &  16.86 & 14.10 & 12.66 & 11.55 & 10.28 &  8.95 &  7.00 &  5.53 &  4.91 \\
+  100 &  12.41 & 10.00 &  9.02 &  8.21 &  7.36 &  6.45 &  5.02 &  4.00 &  3.58 \\
+  200 &   8.65 &  7.05 &  6.37 &  5.80 &  5.22 &  4.57 &  3.57 &  2.84 &  2.54 \\
+  500 &   5.54 &  4.47 &  4.03 &  3.69 &  3.30 &  2.90 &  2.27 &  1.81 &  1.63 \\
+ 1000 &   3.88 &  3.17 &  2.86 &  2.60 &  2.34 &  2.05 &  1.61 &  1.28 &  1.15 \\
+ 2000 &   2.79 &  2.24 &  2.03 &  1.84 &  1.66 &  1.45 &  1.14 &  0.91 &  0.82 \\ 
+ 5000 &   1.75 &  1.41 &  1.27 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.21 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.36 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 

### 2025-10-28 manuscript_rmarkdown/plots.R

@@ -2352,6 +2352,27 @@ get_annotation_settings <- function(ggtheme = NULL) {
 
 
 
+.plot_test_statistic_combined <- function(manuscript_dir, plot_theme) {
+  p_kappa <- .plot_test_statistic_centrality(
+    manuscript_dir = manuscript_dir,
+    plot_theme = plot_theme
+  )
+
+  p_alpha <- .plot_test_statistic_tau(
+    manuscript_dir = manuscript_dir,
+    plot_theme = plot_theme
+  )
+
+  p <- p_kappa + p_alpha + patchwork::plot_layout(
+    ncol = 2L,
+    guides = "keep"
+  )
+
+  return(p)
+}
+
+
+
 .plot_test_statistic_centrality <- function(manuscript_dir, plot_theme) {
   # Plots data that allows us to select the fitting centrality parameter kappa.
 
@@ -2400,7 +2421,8 @@ get_annotation_settings <- function(ggtheme = NULL) {
     )
   )
 
-  P <- p + ggplot2::geom_line(data = data)
+  p <- p + ggplot2::geom_line(data = data)
+  p <- p + ggplot2::ggtitle(latex2exp::TeX("central portion $\\kappa$"))
 
   return(p)
 }
@@ -2451,7 +2473,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   p <- p + ggplot2::scale_x_log10(name = "n")
   p <- p + ggplot2::ylab(latex2exp::TeX("test statistic $\\tau_{\\alpha, n, \\kappa = 0.80}$"))
   p <- p + ggplot2::scale_colour_discrete(
-    name = "significance\nlevel",
+    name = "significance\nlevel α",
     type = c(
       "20.0 %" = "#bacbde",
       "10.0 %" = "#537dac",
@@ -2463,6 +2485,7 @@ get_annotation_settings <- function(ggtheme = NULL) {
   )
 
   p <- p + ggplot2::geom_line(data = data)
+  p <- p + ggplot2::ggtitle(latex2exp::TeX("significance level $\\alpha$"))
 
   return(p)
 }
@@ -2490,78 +2513,70 @@ get_annotation_settings <- function(ggtheme = NULL) {
   ) {
     set.seed(n)
 
+    kappa <- c(1.0, 0.8, 0.6)
     n_rep <- 10000L
     data <- list()
+
     for (jj in seq_len(n_rep)) {
+      x_data <- list()
[…]

### 2025-10-28 manuscript_rmarkdown/simulations.R

@@ -1559,8 +1559,8 @@
 
 
 .get_test_statistics_data <- function(manuscript_dir) {
-  # n_rep <- 30000L  # number of distributions drawn.
-  n_rep <- 10L  # number of distributions drawn.
+  n_rep <- 30000L  # number of distributions drawn.
+  # n_rep <- 10L  # number of distributions drawn.
   kappa <- c(0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00)  # central portion of distribution.
 
   # From 5 to 10000 in equal steps (log 10 scale)
@@ -1593,20 +1593,20 @@
 
     parallel::clusterExport(cl = cl, "..get_test_statistics_data")
 
-    data <- lapply(
-      X = split(n, assignment_id),
-      FUN = .compute_wrapper,
-      n_rep = n_rep,
-      kappa = kappa
-    )
-
-    # data <- parallel::parLapply(
-    #   cl = cl,
+    # data <- lapply(
     #   X = split(n, assignment_id),
-    #   fun = .compute_wrapper,
+    #   FUN = .compute_wrapper,
     #   n_rep = n_rep,
     #   kappa = kappa
     # )
+
+    data <- parallel::parLapply(
+      cl = cl,
+      X = split(n, assignment_id),
+      fun = .compute_wrapper,
+      n_rep = n_rep,
+      kappa = kappa
+    )
 
     # Aggregate to single table.
     data <- data.table::rbindlist(unlist(data, recursive = FALSE))

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Update discussion regarding ECN.

@@
 deviate from normality in a way that might require closer inspection prior to further processing.
-The empirical test for central normality differs from other tests for normality, 
+The empirical central normality test differs from other tests for normality, 
 such as the Shapiro-Wilk test [@Shapiro1965-zd], as it assesses normality of the central portion
 of a feature, instead of the entire feature. 
-Compared to the central normality test and the Shapiro-Wilk test, the empirical central
-normality test remains consistent in the presence of outliers, although the former
-tests are more powerful when outliers are absent.
+Compared to the Shapiro-Wilk test, the empirical central
+normality test remains consistent when a feature was not sampled from a normal distribution,
+such as a central normal distribution with $\kappa < 1.0$. This enables assessing
+reasonable normality of (transformed) features in the presence of outliers.
 

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Don't parse references to internal links in tex.

@@
     latex_engine: lualatex
+    citation_package: natbib
     keep_tex: true

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Update discussion and caption for figure 5.

@@
 For each sequence, the p-value for the respective test was used to reject the null
-hypothesis that the sequence is (centrally) normal. The null hypothesis was rejected if $p \leq 0.05$.
-The type I error rate was determined by computing the fraction of centrally normal sequences rejected this way.
-Figure \ref{fig:empirical-central-normality-error_rate} shows that when the estimated outlier rate 
-for the empirical central normality test matches the outlier rate in the data,
-the test yields an accurate Type I error rate.
-The Shapiro-Wilk test is overly conservative in the presence of any outliers. 
-Empirical central normality tests where the estimated outlier rate is lower
-than the actual outlier rate are likewise conservative, but to a lesser degree. 
-Empirical central normality tests where the estimated outlier rate is higher 
-than the actual outlier rate are optimistic.
+hypothesis that sequences were derived from a population with a (central) normal distribution.
+The null hypothesis was rejected if $p \leq 0.05$.
+The type I error rate was determined by computing the fraction of sequences rejected this way.
+Figure \ref{fig:empirical-central-normality-error_rate} shows that when the
+central portion $\kappa$ of the central normal distribution is equal or greater
+than the $\kappa$ parameter for the empirical central normality test, the tests 
+are accurate. If $\kappa$ of the empirical central normality test exceeds that 
+the central portion $\kappa$ for the underlying distribution, 
+tests are conservative. The Shapiro-Wilk test behaves similar to the empirical
+central normality test with $\kappa = 1.0$, which is expected. For $n < 10$,
+the empirical central normality test can be optimistic.
 
@@
 cap <- paste0(
-  "Observed type I error rates for central normality tests in the presence of data sampled from normal distributions ",
-  "without outliers, or with a 5% or 10% outlier rate. ",
-  "The empirical central normality (ECN) test used $\\kappa = 0.80$, and was further parameterised by the ",
-  "estimated outlier rate $\\zeta = 0.00$ (ECN (0 %)), $\\zeta = 0.05$ (ECN (5 %)) and $\\zeta = 0.10$ (ECN (10%))."  
+  "Observed type I error rates for central normality tests in the presence of data sampled from ",
+  "central normal distributions with $\\kappa \\in \\left{0.6, 0.8, 1.0\\right}$. ",
+  "Three empirical central normality (ECN) tests were performed, parameterised with same value for the central portion $\\kappa$. ",
+  "Additionally, the Shapiro-Wilk test was assessed.",
+  "ECN: empirical central normality"
 )

### 2025-10-28 manuscript_rmarkdown/manuscript.tex
Update discussion and caption for figure 5.

@@
 \PassOptionsToPackage{hyphens}{url}
-%
 \documentclass[
@@
 ]{article}
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
 \usepackage{amsmath,amssymb}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{iftex}
@@
 \makeatother
-\usepackage{xcolor}
-\usepackage[margin=1in]{geometry}
 \usepackage{graphicx}
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 \usepackage{amsmath}
@@
 \newtheorem*{definition}{Definition}
+\newtheorem*{corollary}[Corollary]
 \DeclareMathOperator*{\argmax}{argmax}
@@
 \DeclareMathOperator{\sgn}{sgn}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-04-11}
+\date{2025-10-28}
 
@@
 
-Statistical tests for normality, such as the Shapiro-Wilk test (Shapiro
-and Wilk 1965), could be automatically applied to transformed features.
-However, given sufficiently large sample sizes such tests can detect
-trivial deviations from normality, and may lead to rejection of
+Statistical tests for normality, such as the Shapiro-Wilk test
+\citep{Shapiro1965-zd}, could be automatically applied to transformed
+features. However, given sufficiently large sample sizes such tests can
+detect trivial deviations from normality, and may lead to rejection of
 sufficiently good power transformations.
@@
 
+\citet{Box1964-mz} originally included a shift parameter in their
+transformation, but for their formal analysis set it to \(0\), whereas
+\citet{Yeo2000-vw} did not include a shift or scale parameter. Here,
 Box-Cox and Yeo-Johnson transformations are modified by introducing
@@
 transformations to normality, as shown in Figure
-\ref{fig:decreased-normality}. As indicated by Raymaekers and Rousseeuw
-(2024), the general aim of power transformations should be to transform
-non-outlier data to normality, i.e.~achieve \emph{central normality}. To
-achieve this, they devised an iterative procedure to find a robust
-estimate of the transformation parameter \(\lambda\). Briefly, this
-process requires identifying outliers in the data and weighting such
-instances during the optimisation process. Raymaekers and Rousseeuw
-(2024) achieve this through weighted maximum likelihood estimation.
-However, because this procedure iteratively estimates and updates
-\(\lambda\), it can not be used here to simultaneously estimate
-\(\lambda\), \(x_0\) and \(s\) for location- and scale-invariant power
-transformations. Nonetheless, as a procedure, weighted MLE can be used
-for estimating the transformation, shift and scale parameters.
+\ref{fig:decreased-normality}. As indicated by
+\citet{Raymaekers2024-zf}, the general aim of power transformations
+should be to transform non-outlier data to normality, i.e.~achieve
+\emph{central normality}. To achieve this, they devised an iterative
+procedure to find a robust estimate of the transformation parameter
+\(\lambda\). Briefly, this process requires identifying outliers in the
+data and weighting such instances during the optimisation process.
+\citet{Raymaekers2024-zf} achieve this through weighted maximum
+likelihood estimation. However, because this procedure iteratively
+estimates and updates \(\lambda\), it can not be used here to
+simultaneously estimate \(\lambda\), \(x_0\) and \(s\) for location- and
+scale-invariant power transformations. Nonetheless, as a procedure,
+weighted MLE can be used for estimating the transformation, shift and
+scale parameters.
 
@@
 \ref{eqn:box-cox-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
-and Rousseeuw (2024), these log-likelihood functions include additional
-terms to accommodate estimation of \(x_0\) and \(s\). The weighted
-location- and scale-invariant Box-Cox log-likelihood function is:
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to
+\citet{Raymaekers2024-zf}, these log-likelihood functions include
+additional terms to accommodate estimation of \(x_0\) and \(s\). The
+weighted location- and scale-invariant Box-Cox log-likelihood function
+is:
 
@@
 \item
-  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
-  and \(\delta_2 \geq \delta_1\) as threshold parameters:
-  \begin{equation*}
+  A tapered cosine function \citep{Tukey1967-eb}, with
+  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
+  parameters: \begin{equation*}
   w_i =
@@
   The weighting function uses the z-score of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After Raymaekers and
-  Rousseeuw (2024),
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After
+  \citet{Raymaekers2024-zf},
   \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
@@
 for use with unimodal distributions, the generalised normal distribution
-(Subbotin 1923; Nadarajah 2005) is a suitable option for simulating
-realistic feature distributions. This distribution has the following
-probability density function \(f_{\beta}\) for a value
+\citep{Subbotin1923-qk, Nadarajah2005-xe} is a suitable option for
+simulating realistic feature distributions. This distribution has the
+following probability density function \(f_{\beta}\) for a value
 \(x \in \mathbb{R}\):
@@
 recipe for introducing skewness into the otherwise symmetric generalised
-normal distribution (Gijbels, Karim, and Verhasselt 2019), leading to
-the following probability density function:
+normal distribution \citep{Gijbels2019-te}, leading to the following
+probability density function:
 
@@
 \(F_{\alpha}^{-1}(p)\), which incorporates the quantile function of the
-symmetric generalised normal distribution derived by Griffin (Griffin
-2018):
+symmetric generalised normal distribution derived by Griffin
+\citep{Griffin2018-bf}:
 
@@
 
-\subsection{Central normality test}\label{central-normality-test}
+\subsection{Empirical central normality
+test}\label{empirical-central-normality-test}
 
@@
 Deviations from normality can be detected by normality tests, such as
-the Shapiro-Wilk test (Shapiro and Wilk 1965). In practice, normality
+the Shapiro-Wilk test \citep{Shapiro1965-zd}. In practice, normality
 tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop a test for central normality.
+Here we develop an empirical test for central normality.
 
@@
 
-\begin{definition}[Central normality of a sequence]
-
-The central portion of a sequence is normally distributed if the sum of residual
-errors is equal to zero:
-$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
-
-\end{definition}
-
-In practice, a finite sequence sampled from a normal distribution
-\(\mathcal{N}(\mu, \sigma)\) will have a non-zero sum of residual
-errors.
+\begin{definition}[Central normal distribution]
+
+A central normal distribution is any distribution whose quantile function is 
+defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
+with $\mathcal{N}$ any normal distribution.
+
+\end{definition}
+
+\begin{corollary}
+
+For $\kappa = 1.0$, the central normal distribution is a normal distribution.
+
+\end{corollary}
+
+\begin{corollary}
+
+By the Glivenko-Cantelli theorem, if a sequence $\mathbf{X}$ was sampled from a
+central normal distribution, then almost surely
+$\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,
+
+\end{corollary}
 
 \begin{definition}[Central normality test]
-The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
-
-The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
-\end{definition}
-
-The null-hypothesis \(\mathcal{H}_0\) is that the central portion of
-sequence \(\mathbf{X}\) is normally distributed, with the alternative
-hypothesis \(\mathcal{H}_1\) that it is not normally distributed. To
-test this hypothesis, a test statistic is computed and compared against
-a critical value.
+The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+from a population that is distributed according to a central normal distribution.
+
+The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+population that is not distributed according to a central normal distribution.
+\end{definition}
+
+To test this hypothesis, a test statistic is computed and compared
+against a critical value.
 
 \begin{definition}[Central normality test statistic]
-Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$.
 
@@
 if \(\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}\).
-
-The central portion of the data needs to be defined and the Type 1 error
-rates determined to provide critical test statistics. We will do so in
-the \hyperref[simulation]{Simulation} section.
+Critical test statistic values are determined in the
+\hyperref[simulation-ecn-test]{Simulation} section.
 
@@
 and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
-algorithm for derivative-free bound constraint optimisation (Powell
-2009) through maximum likelihood estimation. The required algorithms
-were implemented in the \texttt{power.transform} R software package
-(Zwanenburg and Löck 2024b) (version 1.0.1). Of note, the
+algorithm for derivative-free bound constraint optimisation
+\citep{Powell2009-zb} through maximum likelihood estimation. The
+required algorithms were implemented in the \texttt{power.transform} R
+software package \citep{Zwanenburg2024-kq} (version 1.0.1). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 
-\subsection{Central normality and empirical central normality
-test}\label{central-normality-and-empirical-central-normality-test}
-
-To develop a test for central normality we need to consider two
-parameters: the central portion \(\kappa\) as a fixed parameter, and
-test statistic \(\tau_{\text{ecn}}\). We will first define the central
-portion \(\kappa\).
+\subsection{Empirical central normality test}\label{simulation-ecn-test}
+
+The critical test statistic for the empirical central normality test
+depends on the significance level \(\alpha\), sequence length \(n\) and
+central portion \(\kappa\). Critical test statistic values were set
+through simulation, as described below.
 
@@
 \(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}\)
-we randomly drew \(m_d = 30000\) ordered sequences \(\mathbf{X}\) from
-\(\mathcal{N}(0,1)\). This dataset was used to compute the critical test
-statistic values for the central normality test. Additionally, we added
-\(10 \%\) outliers by randomly replacing elements of each ordered
-sequence. The dataset with outliers was used to compute the critical
-test statistic values for the empirical variant of the central normality
-test. For each sequence we then computed \(\tau_{n, \kappa}\) for
-\(\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}\).
-
-Figure \ref{fig:empirical-central-normality-kappa} shows
-\(\tau_{\alpha = 0.05, n, \kappa}\) as a function of \(n\) for different
-values of \(\kappa\). With decreasing \(\kappa\), the test statistic
-curve decreases. The curves for \(\kappa = 0.60\), \(\kappa = 0.70\) and
-\(\kappa = 0.80\) are similar, whereas curves for \(\kappa \geq 0.90\)
-are affected by outliers, as can be observed by comparing curves of the
-central normality test with those of the empirical variant.
-
-Since the (empirical) central normality tests assesses whether the
-central part of a sequence is normally distributed, we used
-\(\kappa = 0.80\). Critical statistic values for central normality and
-empirical central normality tests are shown are shown in Table
-\ref{tab:central-normality-critical-statistic} and Table
-\ref{tab:empirical-central-normality-critical-statistic}, respectively.
+we randomly sampled \(m_d = 30000\) ordered sequences \(\mathbf{X}\)
+from \(\mathcal{N}(0,1)\). For each sequence we then computed
+\(\tau_{\kappa}\) for
+\(\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}\).
+
+Figure \ref{fig:empirical-central-normality-combined} shows the critical
+test statistic \(\tau_{\alpha, n, \kappa}\) as a function of \(n\) for
+different values of \(\kappa\) and \(\alpha\). For larger central
+portion \(\kappa\), critical test statistic values increase. Similarly,
+critical test statistic values increase for lower significance levels.
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-kappa-1}
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-combined-1}
 
@@
 
-\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the (empirical) central normality test as function of $n$ for several values of central portion $\kappa$.
-The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
-with fully normal data where 10\% of elements are replaced by outliers.
-emp: empirical}\label{fig:empirical-central-normality-kappa}
+\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the empirical central normality test as function of $n$ for several values of central portion $\kappa$ (left) and several values of significance level $\alpha$ (right).}\label{fig:empirical-central-normality-combined}
 \end{figure}
 
-To assess type I error rates for the (empirical) central normality test
+To assess type I error rates for the empirical central normality test
 and compare these to the Shapiro-Wilk test for normality, we randomly
-drew another 1000 sequences from a standard normal distribution
-(\(\mathcal{N}(0,1)\) for each
+drew another 10000 sequences from a central normal distribution for each
 \(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
-A second dataset was created by replacing elements with randomly drawn
-outliers, as described above. For each sequence, the p-value for the
-respective test was used to reject the null hypothesis that the sequence
-is (centrally) normal. The null hypothesis was rejected if
-\(p \leq 0.05\). The type I error rate was determined by computing the
-fraction of centrally normal sequences rejected this way. Figure
-\ref{fig:empirical-central-normality-error_rate} shows that the
-empirical central normality test had the expected type I error rate for
-centrally normal data (with tails of the distribution not being normally
-distributed), whereas the central normality test and Shapiro-Wilk test
-overestimated type I error rates.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
-
-}
-
-\caption{Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers.
-The empirical central normality (ECN) test uses $\kappa = 0.80$.
-CN: central normality}\label{fig:empirical-central-normality-error_rate}
-\end{figure}
+
+Each sequence was created as follows: first \(n\) values are sampled
+from \(\mathcal{N}(0, 1)\). Then, the elements outside the central
+portion of the sequence
+(i.e.~\(\mathbf{X} \setminus \mathbf{X}_{\text{c}}\)) are replaced.
+Elements with \(pi < \frac{1-\kappa}{2}\) are replaced with values
+sampled from \(\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))\), and
+elements with \(p_i > \frac{1 + \kappa}{2}\) are replaced with values
+sampled from \(\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)\).
+
+For each sequence, the p-value for the respective test was used to
+reject the null hypothesis that sequences were derived from a population
+with a (central) normal distribution. The null hypothesis was rejected
+if \(p \leq 0.05\). The type I error rate was determined by computing
+the fraction of sequences rejected this way. Figure
+\ref{fig:empirical-central-normality-error_rate} shows that when the
+central portion \(\kappa\) of the central normal distribution is equal
+or greater than the \(\kappa\) parameter for the empirical central
+normality test, the tests are accurate. If \(\kappa\) of the empirical
+central normality test exceeds that the central portion \(\kappa\) for
+the underlying distribution, tests are conservative. The Shapiro-Wilk
+test behaves similar to the empirical central normality test with
+\(\kappa = 1.0\), which is expected. For \(n < 10\), the empirical
+central normality test can be optimistic.
+
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
+
+\}
+
+\textbackslash caption\{Observed type I error rates for central
+normality tests in the presence of data sampled from central normal
+distributions with \$\kappa \in \left{0.6, 0.8, 1.0\right}\}. Three
+empirical central normality (ECN) tests were performed, parameterised
+with same value for the central portion \(\kappa\).Additional, the
+Shapiro-Wilk test was assessed.ECN: empirical central
+normality\}\label{fig:empirical-central-normality-error_rate}
+\textbackslash end\{figure\}
 
 In Figure \ref{fig:empirical-central-normality-examples} we apply the
-empirical central normality test to assess central normality of features
-that are composed of a mixture of samples drawn from two normal
-distributions (\(n = 100\) each). With increased separation of the
-underlying normal distributions, the probability of the feature being
-centrally normal decreases, as expected.
+empirical central normality test (with \(\kappa = 0.80\)) to assess
+central normality of features that are composed of a mixture of samples
+drawn from two normal distributions (\(n = 100\) each). With increased
+separation of the underlying normal distributions, the test's p-value
+decreases, as expected.
 
@@
 The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) drawn from normal distributions that are offset by a distance $d$.
-We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
-As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
+The empirical central normality ($\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from a central normal distribution.
+As may be observed, with increasing offset $d$ the test's p-value decreases.
 Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
 \end{figure}
+
+We use \(\kappa = 0.80\) for the empirical central normality tests in
+the remainder of the manuscript, unless noted otherwise. Critical
+statistic values of the empirical central normality test for
+\(\kappa = 0.80\) are shown in Table
+\ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 \(\gamma \sim U\left(\log_{10}50, 3\right)\), resulting in a set of
-sequences with between \(50\) and \(1000\) elements each.
-
-Outlier values were then drawn to randomly replace 10 percent of the
-elements of \(\mathbf{X}_i\). Outlier values were set according to John
-W. Tukey (1977), as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
-Then the corresponding outlier value was:
+sequences with between \(50\) and \(1000\) elements each. Outlier values
+were then drawn to randomly replace 10 percent of the elements of
+\(\mathbf{X}_i\). These values were set according to
+\citet{Tukey1977-xm}, as follows. Let
+\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
+was:
 
 \begin{equation}
+\label{eqn:outlier_rate}
 x_{out} =
@@
 
-\(Q_1\), \(Q_3\) and \(\text{IQR}\) are the first quartile, third
+Here, \(Q_1\), \(Q_3\) and \(\text{IQR}\) are the first quartile, third
 quartile and interquartile range of \(\mathbf{X}_i\), respectively.
-Outlier values randomly replaced elements in \(\mathbf{X}_i\).
 
@@
 
-Three datasets with 10000 sequences each were created to assess
+Three datasets with 100000 sequences each were created to assess
 transformation to normality. For each sequence \(\mathbf{X}_i\),
@@
   robust scaling:
-  \(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)\),
-  with \(\text{IQR}\) representing the interquartile range of
-  \(\mathbf{X}\).
+  \(x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / Q_n\left(\mathbf{X}\right)\),
+  with \(Q_n\) representing the robust \(Q_n\) scale estimator of
+  \(\mathbf{X}\) \citep{Croux1992-zz, Rousseeuw1993-it}.
 \end{enumerate}
@@
 the central portion (data with outliers). The average rank of each
-transformation method was computed over all 10000 sequences in each
+transformation method was computed over all 100000 sequences in each
 dataset. Average ranks for Yeo-Johnson transformations are shown in
@@
 Gorman, Williams and Fraser recorded body mass (in grams) of 342
-penguins of three different species (Gorman, Williams, and Fraser 2014).
-The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard
-deviation) grams, and not centrally normal (\(p < 0.001\)). Applying
-conventional and invariant Yeo-Johnson transformations to body mass
-yielded the following results, see Figure
-\ref{fig:experimental-results-invariance}: no transformation (residual
-sum \(\sum r_i = 48.0\), \(p < 0.001\)); conventional transformation
-(\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+penguins of three different species \citep{Gorman2014-eo}. The body mass
+was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard deviation) grams, and
+not centrally normal (\(p < 0.001\)). Applying conventional and
+invariant Yeo-Johnson transformations to body mass yielded the following
+results, see Figure \ref{fig:experimental-results-invariance}: no
+transformation (residual sum \(\sum r_i = 48.0\), \(p < 0.001\));
+conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
 \(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\));
-location- and scale-invariant transformation (\(\lambda = 0.5\),
-\(\sum r_i = 26.8\), \(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\),
-\(p = 0.28\)); and robust location- and scale-invariant transformation
-(\(\lambda = 0.3\), \(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\),
-\(\sigma_{YJ} = 0.9\), \(p = 0.69\)).
+Raymaekers and Rousseeuw's robust adaptation (\(\lambda = -0.5\),
+\(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); location- and
+scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
+\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\), \(p = 0.28\)); and robust
+location- and scale-invariant transformation (\(\lambda = 0.3\),
+\(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.9\),
+\(p = 0.69\)).
 
@@
 The Top Gear dataset contains data on 297 vehicles that appeared on the
-BBC television show \emph{Top Gear} (Alfons 2021). Within this dataset,
-the fuel consumption feature contains outliers due to highly
+BBC television show \emph{Top Gear} \citep{Alfons2021-kc}. Within this
+dataset, the fuel consumption feature contains outliers due to highly
 fuel-efficient vehicles. Applying conventional and invariant Yeo-Johnson
@@
 We used 231 datasets containing at least one numeric feature from the
-Penn Machine Learning Benchmarks collection (Romano et al. 2022). In
+Penn Machine Learning Benchmarks collection \citep{Romano2022-gq}. In
 this collection, 114 datasets correspond to regression tasks and 117
 datasets to classification tasks. Using the familiar auto-machine
-learning library (Zwanenburg and Löck 2024a) (version 1.5.0), each
-dataset was used to train a model for each of 32 process configurations.
-Each process configuration specifies the learner (generalised linear
-model, L1-regularised linear models (Lasso), gradient boosted linear
-model, or random forest), transformation method (none, conventional
-Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson
-with empirical central normality test (rejecting transformations with
+learning library \citep{Zwanenburg2021-so} (version 1.5.0), each dataset
+was used to train a model for each of 32 process configurations. Each
+process configuration specifies the learner (generalised linear model,
+L1-regularised linear models (Lasso), gradient boosted linear model, or
+random forest), transformation method (none, conventional Yeo-Johnson,
+robust invariant Yeo-Johnson, robust invariant Yeo-Johnson with
+empirical central normality test (rejecting transformations with
 \(p \leq 0.01\)), and normalisation method (none,
@@
 transformation method, we first fit a regression random forest (ranger
-package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
-2, other hyperparameters default) with process parameters and task
+package \citep{Wright2017-rf} version 0.16.0): 2000 trees, node size 2,
+other hyperparameters default) with process parameters and task
 difficulty as predictors and normalised rank as response variable. The
@@
 \ref{eqn:box-cox-original} for the theoretical analysis in their paper
-(Box and Cox 1964), which subsequently became the convention. Yeo and
-Johnson's power transformation lacks a shift parameter altogether (Yeo
-and Johnson 2000). We showed that these power transformations are
+\citep{Box1964-mz}, which subsequently became the convention. Yeo and
+Johnson's power transformation lacks a shift parameter altogether
+\citep{Yeo2000-vw}. We showed that these power transformations are
 sensitive to location and scale of data distributions. To mitigate this
@@
 sequences deviate from normality in a way that might require closer
-inspection prior to further processing. The empirical test for central
-normality differs from other tests for normality, such as the
-Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
-the central portion of a feature, instead of the entire feature.
-Compared to the central normality test and the Shapiro-Wilk test, the
-empirical central normality test remains consistent in the presence of
-outliers, although the former tests are more powerful when outliers are
-absent.
+inspection prior to further processing. The empirical central normality
+test differs from other tests for normality, such as the Shapiro-Wilk
+test \citep{Shapiro1965-zd}, as it assesses normality of the central
+portion of a feature, instead of the entire feature. Compared to the
+Shapiro-Wilk test, the empirical central normality test remains
+consistent when a feature was not sampled from a normal distribution,
+such as a central normal distribution with \(\kappa < 1.0\). This
+enables assessing reasonable normality of (transformed) features in the
+presence of outliers.
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-Alfons2021-kc}
-Alfons, Andreas. 2021. {``{robustHD}: An {R} Package for Robust
-Regression with High-Dimensional Data.''} \emph{J. Open Source Softw.} 6
-(67): 3786. \url{https://doi.org/10.21105/joss.03786}.
-
-\bibitem[\citeproctext]{ref-Bartlett1947-rx}
-Bartlett, M S. 1947. {``The Use of Transformations.''} \emph{Biometrics}
-3 (1): 39--52. \url{https://doi.org/10.2307/3001536}.
-
-\bibitem[\citeproctext]{ref-Box1964-mz}
-Box, G E P, and D R Cox. 1964. {``An Analysis of Transformations.''}
-\emph{J. R. Stat. Soc. Series B Stat. Methodol.} 26 (2): 211--52.
-\url{https://doi.org/10.1111/J.2517-6161.1964.TB00553.X}.
-
-\bibitem[\citeproctext]{ref-De-Cock2011-jf}
-De Cock, Dean. 2011. {``Ames, Iowa: Alternative to the Boston Housing
-Data as an End of Semester Regression Project.''} \emph{J. Stat. Educ.}
-19 (3). \url{https://doi.org/10.1080/10691898.2011.11889627}.
-
-\bibitem[\citeproctext]{ref-Gijbels2019-te}
-Gijbels, Irène, Rezaul Karim, and Anneleen Verhasselt. 2019. {``Quantile
-Estimation in a Generalized Asymmetric Distributional Setting.''} In
-\emph{Stochastic Models, Statistics and Their Applications}, 13--40.
-Springer International Publishing.
-\url{https://doi.org/10.1007/978-3-030-28665-1/_2}.
-
-\bibitem[\citeproctext]{ref-Gorman2014-eo}
-Gorman, Kristen B, Tony D Williams, and William R Fraser. 2014.
-{``Ecological Sexual Dimorphism and Environmental Variability Within a
-Community of Antarctic Penguins (Genus Pygoscelis).''} \emph{PLoS One} 9
-(3): e90081. \url{https://doi.org/10.1371/journal.pone.0090081}.
-
-\bibitem[\citeproctext]{ref-Griffin2018-bf}
-Griffin, Maryclare. 2018. {``Gnorm: Generalized Normal/Exponential Power
-Distribution.''} The R Foundation.
-\url{https://doi.org/10.32614/cran.package.gnorm}.
-
-\bibitem[\citeproctext]{ref-Huber1981-su}
-Huber, Peter J. 1981. \emph{Robust Statistics}. John Wiley \& Sons.
-\url{https://doi.org/10.1002/0471725250}.
-
-\bibitem[\citeproctext]{ref-Kuhn2019-kt}
-Kuhn, Max, and Kjell Johnson. 2019. \emph{Feature Engineering and
-Selection: A Practical Approach for Predictive Models}. Chapman \&
-Hall/CRC Data Science Series. Chapman; Hall/CRC.
-\url{https://doi.org/10.1201/9781315108230}.
-
-\bibitem[\citeproctext]{ref-Loprinzi1994-cd}
-Loprinzi, C L, J A Laurie, H S Wieand, J E Krook, P J Novotny, J W
-Kugler, J Bartel, M Law, M Bateman, and N E Klatt. 1994. {``Prospective
-Evaluation of Prognostic Variables from Patient-Completed
-Questionnaires. North Central Cancer Treatment Group.''} \emph{J. Clin.
-Oncol.} 12 (3): 601--7. \url{https://doi.org/10.1200/JCO.1994.12.3.601}.
-
-\bibitem[\citeproctext]{ref-Nadarajah2005-xe}
-Nadarajah, Saralees. 2005. {``A Generalized Normal Distribution.''}
-\emph{J. Appl. Stat.} 32 (7): 685--94.
-\url{https://doi.org/10.1080/02664760500079464}.
-
-\bibitem[\citeproctext]{ref-Powell2009-zb}
-Powell, M J D. 2009. {``The {BOBYQA} Algorithm for Bound Constrained
-Optimization Without Derivatives.''} Cambridge NA Report NA2009/06.
-University of Cambridge.
-
-\bibitem[\citeproctext]{ref-Raymaekers2024-zf}
-Raymaekers, Jakob, and Peter J Rousseeuw. 2024. {``Transforming
-Variables to Central Normality.''} \emph{Mach. Learn.} 113 (8):
-4953--75. \url{https://doi.org/10.1007/s10994-021-05960-5}.
-
-\bibitem[\citeproctext]{ref-Romano2022-gq}
-Romano, Joseph D, Trang T Le, William La Cava, John T Gregg, Daniel J
-Goldberg, Praneel Chakraborty, Natasha L Ray, Daniel Himmelstein,
-Weixuan Fu, and Jason H Moore. 2022. {``{PMLB} V1.0: An Open-Source
-Dataset Collection for Benchmarking Machine Learning Methods.''}
-\emph{Bioinformatics} 38 (3): 878--80.
-\url{https://doi.org/10.1093/bioinformatics/btab727}.
-
-\bibitem[\citeproctext]{ref-Shapiro1965-zd}
-Shapiro, S S, and M B Wilk. 1965. {``An Analysis of Variance Test for
-Normality (Complete Samples).''} \emph{Biometrika} 52 (3/4): 591--611.
-\url{https://doi.org/10.2307/2333709}.
-
-\bibitem[\citeproctext]{ref-Subbotin1923-qk}
-Subbotin, M Th. 1923. {``On the Law of Frequency of Error.''} \emph{Mat.
-Sb.} 31 (2): 296--301.
-
-\bibitem[\citeproctext]{ref-Tukey1967-eb}
-Tukey, J W. 1967. {``An Introduction to the Calculations of Numerical
-Spectrum Analysis.''} In \emph{Advanced Seminar on Spectral Analysis of
-Time Series}, edited by B Harris. New York: John Wiley; Sons, Inc.
-
-\bibitem[\citeproctext]{ref-Tukey1957-rt}
-Tukey, John W. 1957. {``On the Comparative Anatomy of
-Transformations.''} \emph{Ann. Math. Stat.} 28 (3): 602--32.
-\url{https://doi.org/10.1214/AOMS/1177706875}.
-
-\bibitem[\citeproctext]{ref-Tukey1977-xm}
----------. 1977. \emph{Exploratory Data Analysis}. Addison-Wesley
-Publishing Company.
-
-\bibitem[\citeproctext]{ref-Wright2017-rf}
-Wright, Marvin N., and Andreas Ziegler. 2017. {``{ranger}: A Fast
-Implementation of Random Forests for High Dimensional Data in {C++} and
-{R}.''} \emph{Journal of Statistical Software} 77 (1): 1--17.
-\url{https://doi.org/10.18637/jss.v077.i01}.
-
-\bibitem[\citeproctext]{ref-Yeo2000-vw}
-Yeo, In‐kwon, and Richard A Johnson. 2000. {``A New Family of Power
-Transformations to Improve Normality or Symmetry.''} \emph{Biometrika}
-87 (4): 954--59. \url{https://doi.org/10.1093/biomet/87.4.954}.
-
-\bibitem[\citeproctext]{ref-Zwanenburg2021-so}
-Zwanenburg, Alex, and Steffen Löck. 2024a. {``Familiar: End-to-End
-Automated Machine Learning and Model Evaluation.''} The R Foundation.
-\url{https://doi.org/10.32614/cran.package.familiar}.
-
-\bibitem[\citeproctext]{ref-Zwanenburg2024-kq}
----------. 2024b. {``Power.transform: Location and Scale Invariant Power
-Transformations.''} The R Foundation.
-\url{https://doi.org/10.32614/cran.package.power.transform}.
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{refs.bib}
 

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Updated Latex file :)

@@
 - \newtheorem*{definition}{Definition}
-- \newtheorem*{corollary}[Corollary]
+- \newtheorem*{corollary}{Corollary}
 - \DeclareMathOperator*{\argmax}{argmax}
@@
   "Observed type I error rates for central normality tests in the presence of data sampled from ",
-  "central normal distributions with $\\kappa \\in \\left{0.6, 0.8, 1.0\\right}$. ",
+  "central normal distributions with $\\kappa \\in \\left\\{0.6, 0.8, 1.0\\right\\}$. ",
   "Three empirical central normality (ECN) tests were performed, parameterised with same value for the central portion $\\kappa$. ",
-  "Additionally, the Shapiro-Wilk test was assessed.",
+  "Additionally, the Shapiro-Wilk test was assessed. ",
   "ECN: empirical central normality"

### 2025-10-28 manuscript_rmarkdown/manuscript.tex
Updated Latex file :)

@@
 \newtheorem*{definition}{Definition}
-\newtheorem*{corollary}[Corollary]
+\newtheorem*{corollary}{Corollary}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 
-\textbackslash begin\{figure\}
-
-\{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
-
-\}
-
-\textbackslash caption\{Observed type I error rates for central
-normality tests in the presence of data sampled from central normal
-distributions with \$\kappa \in \left{0.6, 0.8, 1.0\right}\}. Three
-empirical central normality (ECN) tests were performed, parameterised
-with same value for the central portion \(\kappa\).Additional, the
-Shapiro-Wilk test was assessed.ECN: empirical central
-normality\}\label{fig:empirical-central-normality-error_rate}
-\textbackslash end\{figure\}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
+
+}
+
+\caption{Observed type I error rates for central normality tests in the presence of data sampled from central normal distributions with $\kappa \in \left\{0.6, 0.8, 1.0\right\}$.
+Three empirical central normality (ECN) tests were performed, parameterised with same value for the central portion $\kappa$.
+Additionally, the Shapiro-Wilk test was assessed.
+ECN: empirical central normality}\label{fig:empirical-central-normality-error_rate}
+\end{figure}
 

### 2025-10-28 manuscript_rmarkdown/manuscript_appendix.Rmd
Update manuscript_appendix.Rmd

@@
 \caption{
-p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
@@
 
-none                                  & & 0.981 & 0.816 & 0.576 & $<0.001$ & 0.099 & 0.011 & $<0.001$ & 0.993 & $<0.001$ \\
-conventional                          & & 0.979 & 0.614 & 0.617 &   0.722  & 0.221 & 0.782 &   0.946  & 0.997 &   0.637 \\
-conventional (z-score norm.)          & & 0.970 & 0.594 & 0.645 &   0.251  & 0.224 & 0.895 &   0.529  & 0.998 &   0.828 \\
-conventional (robust scaling)         & & 0.974 & 0.593 & 0.640 &   0.538  & 0.207 & 0.837 &   0.933  & 0.998 &   0.810 \\
-Raymaekers-Rousseeuw                  & & 0.983 & 0.614 & 0.617 &   0.722  & 0.622 & 0.782 &   0.927  & 0.985 &   0.831 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.985 & 0.594 & 0.645 &   0.251  & 0.623 & 0.895 &   0.564  & 0.984 &   0.807 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.987 & 0.593 & 0.640 &   0.538  & 0.599 & 0.837 &   0.935  & 0.984 &   0.875 \\
-invariant                             & & 0.987 & 0.487 & 0.573 &   0.778  & 0.241 & 0.683 &   0.963  & 0.997 &   0.865 \\
-robust invariant                      & & 0.992 & 0.753 & 0.667 &   0.909  & 0.739 & 0.628 &   0.968  & 0.994 &   0.680 \\
+none                                  & & 0.968 & 0.742 & 0.450 & $<0.001$ & 0.058 & 0.004 & $<0.001$ & 0.963 & $<0.001$ \\
+conventional                          & & 0.966 & 0.494 & 0.498 &   0.632  & 0.145 & 0.707 &   0.828  & 0.986 &   0.358 \\
+conventional (z-score norm.)          & & 0.955 & 0.471 & 0.529 &   0.169  & 0.147 & 0.853 &   0.260  & 0.986 &   0.603 \\
+conventional (robust scaling)         & & 0.957 & 0.470 & 0.523 &   0.674  & 0.146 & 0.775 &   0.828  & 0.986 &   0.347 \\
+Raymaekers-Rousseeuw                  & & 0.972 & 0.494 & 0.498 &   0.632  & 0.518 & 0.707 &   0.785  & 0.939 &   0.609 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.974 & 0.471 & 0.529 &   0.169  & 0.519 & 0.853 &   0.289  & 0.935 &   0.568 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.968 & 0.470 & 0.523 &   0.674  & 0.536 & 0.775 &   0.762  & 0.935 &   0.609 \\
+invariant                             & & 0.978 & 0.353 & 0.448 &   0.701  & 0.161 & 0.588 &   0.872  & 0.984 &   0.667 \\
+robust invariant                      & & 0.985 & 0.660 & 0.556 &   0.871  & 0.652 & 0.525 &   0.888  & 0.968 &   0.411 \\
 
@@
 \caption{
-p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Box-Cox transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
@@
 \midrule
-
-none                                  & & 0.940 & 0.816 & 0.498 & $<0.001$ & 0.098 & 0.714 & $<0.001$ & 0.993 & 0.928 \\
-conventional                          & & 0.948 & 0.695 & 0.603 &   0.798  & 0.163 & 0.791 &   0.931  & 0.997 & 0.658 \\
-conventional (z-score norm.)          & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
-conventional (robust scaling)         & & 0.947 & 0.777 & 0.624 &   0.611  & 0.182 & 0.788 &   0.969  & 0.991 & 0.822 \\
-Raymaekers-Rousseeuw                  & & 0.948 & 0.695 & 0.603 &   0.798  & 0.768 & 0.806 &   0.816  & 0.962 & 0.782 \\
-Raymaekers-Rousseeuw (z-score norm.)  & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.775  & 0.866 & 0.921 \\
-Raymaekers-Rousseeuw (robust scaling) & & 0.947 & 0.777 & 0.753 &   0.913  & 0.255 & 0.784 &   0.807  & 0.866 & 0.921 \\
-invariant                             & & 0.947 & 0.661 & 0.624 &   0.611  & 0.182 & 0.789 &   0.956  & 0.997 & 0.664 \\
-robust invariant                      & & 0.969 & 0.824 & 0.750 &   0.972  & 0.779 & 0.775 &   0.956  & 0.934 & 0.840 \\
+none                                  & & 0.911 & 0.742 & 0.365 & $<0.001$ & 0.057 & 0.623 & $<0.001$ & 0.962 & 0.789 \\
+conventional                          & & 0.922 & 0.590 & 0.482 &   0.726  & 0.100 & 0.718 &   0.794  & 0.984 & 0.380 \\
+conventional (z-score norm.)          & & 0.920 & 0.690 & 0.506 &   0.505  & 0.114 & 0.714 &   0.888  & 0.957 & 0.594 \\
+conventional (robust scaling)         & & 0.920 & 0.690 & 0.506 &   0.504  & 0.114 & 0.714 &   0.887  & 0.957 & 0.594 \\
+Raymaekers-Rousseeuw                  & & 0.922 & 0.590 & 0.482 &   0.726  & 0.689 & 0.736 &   0.582  & 0.869 & 0.531 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.920 & 0.690 & 0.660 &   0.876  & 0.172 & 0.710 &   0.521  & 0.670 & 0.774 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.920 & 0.690 & 0.660 &   0.876  & 0.172 & 0.710 &   0.568  & 0.670 & 0.774 \\
+invariant                             & & 0.921 & 0.549 & 0.506 &   0.505  & 0.114 & 0.715 &   0.853  & 0.986 & 0.387 \\
+robust invariant                      & & 0.954 & 0.752 & 0.656 &   0.956  & 0.699 & 0.698 &   0.852  & 0.801 & 0.623 \\
 
@@
 \caption{
-p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from real-world datasets after Yeo-Johnson transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
@@
 
-none                                  & 0.694 & 0.832     & 0.727     & $< 0.001$ & $< 0.001$ \\
-conventional                          & 0.961 & 0.003     & $< 0.001$ & $< 0.001$ & 0.101 \\
-conventional (z-score norm.)          & 0.919 & 0.021     & $< 0.001$ & $< 0.001$ & 0.099 \\
-conventional (robust scaling)         & 0.933 & $< 0.001$ & $< 0.001$ & $< 0.001$ & 0.129 \\
-Raymaekers-Rousseeuw                  & 0.961 & 0.881     & 0.373     & $< 0.001$ & 0.101 \\
-Raymaekers-Rousseeuw (z-score norm.)  & 0.694 & 0.899     & 0.809     & $< 0.001$ & 0.099 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.900 & 0.880     & 0.811     & $< 0.001$ & 0.144 \\
-invariant                             & 0.976 & 0.147     & $< 0.001$ & $< 0.001$ & 0.276 \\
-robust invariant                      & 0.933 & 0.178     & 0.759     & $< 0.001$ & 0.688 \\
+none                                  & 0.541 &    0.731  &    0.546  & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.919 & $< 0.001$ & $< 0.001$ & $< 0.001$ &    0.032 \\
+conventional (z-score norm.)          & 0.845 &    0.007  & $< 0.001$ & $< 0.001$ &    0.031 \\
+conventional (robust scaling)         & 0.847 & $< 0.001$ & $< 0.001$ & $< 0.001$ &    0.063 \\
+Raymaekers-Rousseeuw                  & 0.919 &    0.802  &    0.205  & $< 0.001$ &    0.032 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.540 &    0.830  &    0.655  & $< 0.001$ &    0.031 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.438 &    0.878  &    0.669  & $< 0.001$ &    0.064 \\
+invariant                             & 0.947 &    0.071  & $< 0.001$ & $< 0.001$ &    0.123 \\
+robust invariant                      & 0.867 &    0.091  &    0.587  & $< 0.001$ &    0.477 \\
 
@@
 \caption{
-p-values of empirical central normality tests for features from real-world datasets after Box-Cox transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from real-world datasets after Box-Cox transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
@@
 
-none                                  & 0.694 & 0.832 & 0.727     & $< 0.001$ & $< 0.001$ \\
-conventional                          & 0.961 & 0.002 & $< 0.001$ & $< 0.001$ & 0.101 \\
-conventional (z-score norm.)          & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
-conventional (robust scaling)         & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
-Raymaekers-Rousseeuw                  & 0.961 & 0.871 & 0.365     & $< 0.001$ & 0.101 \\
-Raymaekers-Rousseeuw (z-score norm.)  & 0.891 & 0.268 & $< 0.001$ & $< 0.001$ &	0.495 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.562 & 0.789 & $< 0.001$ & $< 0.001$ &	0.492 \\
-invariant                             & 0.962 & 0.004 & $< 0.001$ & $< 0.001$ &	0.154 \\
-robust invariant                      & 0.899 & 0.890 & 0.866     & $< 0.001$ & 0.622 \\
+none                                  & 0.541 &    0.731  &    0.546  & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.919 & $< 0.001$ & $< 0.001$ & $< 0.001$ &    0.032 \\
+conventional (z-score norm.)          & 0.815 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+conventional (robust scaling)         & 0.815 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+Raymaekers-Rousseeuw                  & 0.919 &    0.787  &    0.200  & $< 0.001$ &    0.032 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.799 &    0.152  & $< 0.001$ & $< 0.001$ &    0.285 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.799 &    0.674  &    0.060  & $< 0.001$ &    0.285 \\
+invariant                             & 0.919 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+robust invariant                      & 0.813 &    0.817  &    0.736  & $< 0.001$ &    0.403 \\
 

### 2025-10-29 manuscript_rmarkdown/manuscript.Rmd
Updates to text.

@@
 for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
-The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq] (version 1.0.1).
+The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq] (version 1.0.2).
 Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.
@@
 Then, the elements outside the central portion of the sequence (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
-Elements with $pi < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
+Elements with $p_i < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
 and elements with $p_i > \frac{1 + \kappa}{2}$ are replaced with values sampled from $\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)$.
@@
   "Additionally, the Shapiro-Wilk test was assessed. ",
+  "The dashed line indicates the expected type I error rate for rejecting the null hypothesis ",
+  "at significance level $\\alpha = 0.05$. ",
   "ECN: empirical central normality"
@@
   "drawn from normal distributions that are offset by a distance $d$. ",
-  "The empirical central normality ($\\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from ",
+  "The empirical central normality test ($\\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from ",
   "a central normal distribution. As may be observed, with increasing offset $d$ ",
@@
 Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:experimental-results-invariance}:
-no transformation ($\sum r_i = 16.5$, $p = 0.69$);
-conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.96$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.96$);
-location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 8.8$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 1.1$, $p = 0.98$);
-and robust location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 9.3$, $\mu_{YJ} = -1.0$, $\sigma_{YJ} = 1.1$, $p = 0.93$).
+no transformation ($\sum r_i = 16.5$, $p = 0.54$);
+conventional transformation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.92$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 2.0$, $\sum r_i = 11.5$, $\mu_{YJ} = 1.8 \cdot 10^3$, $\sigma_{YJ} = 0.5 \cdot 10^3$, $p = 0.92$);
+location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 8.8$, $\mu_{YJ} = -1.2$, $\sigma_{YJ} = 1.1$, $p = 0.95$);
+and robust location- and scale-invariant transformation ($\lambda = 1.3$, $\sum r_i = 9.3$, $\mu_{YJ} = -1.0$, $\sigma_{YJ} = 1.1$, $p = 0.87$).
 
@@
 Conventional transformations inflated the mean $\mu_{YJ}$ and standard deviation $\sigma_{YJ}$ of the age feature after transformation.
-The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.93$).
+The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.87$).
 
@@
 no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
-conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.10$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.10$);
-location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$, $p = 0.28$);
-and robust location- and scale-invariant transformation ($\lambda = 0.3$, $\sum r_i = 22.0$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.9$, $p = 0.69$).
-
-Location- and scale-invariant transformation produced a lower overall residual errors, indicating a better transformation.
-Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.
-The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.10$).
+conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.03$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.1$, $\sigma_{YJ} = 4 \cdot 10^{-3}$, $p = 0.03$);
+location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$, $p = 0.12$);
+and robust location- and scale-invariant transformation ($\lambda = 0.3$, $\sum r_i = 22.0$, $\mu_{YJ} = 0.7$, $\sigma_{YJ} = 0.9$, $p = 0.48$).
+
+Location- and scale-invariant transformation yielded lower overall residual errors, indicating a better transformation to central normality,
+which is reflected by the empirical central normality test results. 
+Conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation,
+and the null hypothesis of the empirical central normality test was rejected (at significance level $\alpha = 0.05$) in both cases.
 
@@
 
-Every transformation reduced the residual sum.
+Every transformation reduced the residual sum compared to the original data.
 None of the transformations yielded a centrally normal distribution.
@@
 Applying conventional and invariant Yeo-Johnson transformations to the fuel consumption feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 54$, $p=0.72$);
+no transformation (residual sum $\sum r_i = 54$, $p=0.55$);
 conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{YJ} = 3.0$, $\sigma_{YJ} = 0.3$, $p < 0.001$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.37$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{YJ} = 29$, $\sigma_{YJ} = 15$, $p=0.21$);
 location- and scale-invariant transformation ($\lambda = -1.3$, $\sum r_i = 44$, $\mu_{YJ} = 0.5$, $\sigma_{YJ} = 0.1$, $p < 0.001$);
-and robust location- and scale-invariant transformation ($\lambda = 1.0$, $\sum r_i = 56$, $\mu_{YJ} = 1.7$, $\sigma_{YJ} = 1.0$, $p=0.76$).
-
-Outliers cause non-robust transformations to fail to transform the data to a centrally normal distribution
-(empirical central normality test $p < 0.001$ for conventional and invariant transformations).
-Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
+and robust location- and scale-invariant transformation ($\lambda = 1.0$, $\sum r_i = 56$, $\mu_{YJ} = 1.7$, $\sigma_{YJ} = 1.0$, $p=0.59$).
+
+Outliers caused non-robust transformations to fail the empirical central normality test for
+transforming the feature to a centrally normal distribution.
 
@@
 Applying conventional and invariant Yeo-Johnson transformations to this feature yielded the following results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum $\sum r_i = 110$, $p=0.83$);
-conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p=0.003$);
-Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14.3$, $p=0.88$);
-location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.15$);
-and robust location- and scale-invariant transformation ($\lambda = -0.3$, $\sum r_i = 30$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.2$, $p=0.18$).
-
-The conventional non-robust transformation failed to produce a centrally normal distribution (empirical central normality test $p=0.003$).
-Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
+no transformation (residual sum $\sum r_i = 110$, $p=0.73$);
+conventional transformation ($\lambda = -0.7$, $\sum r_i = 30$, $\mu_{YJ} = 1.0$, $\sigma_{YJ} = 0.1$, $p < 0.001$);
+Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 136$, $\mu_{YJ} = 7.2$, $\sigma_{YJ} = 14.3$, $p=0.80$);
+location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 12$, $\mu_{YJ} = -11.8$, $\sigma_{YJ} = 6.9$, $p=0.07$);
+and robust location- and scale-invariant transformation ($\lambda = -0.3$, $\sum r_i = 30$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.2$, $p=0.09$).
+
+For conventional non-robust transformation the null hypothesis was rejected (at $\alpha = 0.05$), indicating
+that they failed to produce a transformed feature that followed a central normal distribution.
 

### 2025-10-29 manuscript_rmarkdown/manuscript_appendix.Rmd
Updates to text.

@@
 metric is computed from the residual error and the residual error of the central
-portion ($\kappa = 0.70$) for sequences without and with outliers, respectively.
+portion ($\kappa = 0.80$) for sequences without and with outliers, respectively.
 In each experiment, the respective residual error was normalised by dividing with
@@
 conventional (z-score norm.)          & 11.5 &  19.3 & 49.0 & 326.2 & 33.3 \\
-conventional (robust scaling)         & 11.3 &  33.3 & 55.9 & 326.4 & 31.5 \\
+conventional (robust scaling)         & 11.3 &  33.3 & 55.9 & 326.5 & 31.5 \\
 Raymaekers-Rousseeuw                  & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
 Raymaekers-Rousseeuw (z-score norm.)  & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
-Raymaekers-Rousseeuw (robust scaling) & 13.4 & 203.0 & 56.9 & 326.4 & 31.5 \\
+Raymaekers-Rousseeuw (robust scaling) & 13.4 & 202.9 & 56.9 & 326.4 & 31.5 \\
 invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\

### 2025-10-29 manuscript_rmarkdown/manuscript_appendix.tex
Updates to text.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-04-11}
+\date{2025-10-29}
 
@@
 
-The parameters of these power transformations can be optimised based by
+The parameters of these power transformations can be optimised by
 maximising the log-likelihood function, under the assumption that the
@@
 \(\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)\)
-its transformed values using conventional and shift and scale invariant
-power transformations, respectively. Since power transformations are
-monotonic, \(\mathbf{Y}\) will likewise be ordered.
+its transformed values using conventional and location- and
+scale-invariant power transformations, respectively. Since power
+transformations are monotonic, \(\mathbf{Y}\) will likewise be ordered.
 
 Below we will focus on criteria based on the empirical density function
-and those based on skewness and kurtosis of the transformed featured.
+and those based on skewness and kurtosis of the transformed feature.
 Other potential criteria, such as the Shapiro-Wilk test statistic
-(Shapiro and Wilk 1965) are not investigated here. In the case of the
+\citep{Shapiro1965-zd} are not investigated here. In the case of the
 Shapiro-Wilk test statistic this is because of lack of scalability to
@@
 EDF and the normal CDF, with differences at the upper and lower end of
-the normal CDF receiving more weight than those at the the centre of the
+the normal CDF receiving more weight than those at the centre of the
 CDF.
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cramér-von Mises test statistic Mises
-(1928), aside from a additive scalar value and the introduction of
-weights. This criterion, like the Anderson-Darling criterion, seeks to
-minimise the squared Euclidean distance between the EDF and the normal
-CDF. Unlike the Anderson-Darling criterion, this criterion weights all
-instances equally.
+The criterion is similar to the Cramér-von Mises test statistic
+\citep[\citet{Von_Mises1928-ef}]{Cramer1928-rc}, aside from an additive
+scalar value and the introduction of weights. This criterion, like the
+Anderson-Darling criterion, seeks to minimise the squared Euclidean
+distance between the EDF and the normal CDF. Unlike the Anderson-Darling
+criterion, this criterion weights all instances equally.
 
@@
 kurtosis of the transformed feature \(\mathbf{Y}\). We will first define
-the location \(\mu\) and scale \(\sigma\) of the the transformed as
+the location \(\mu\) and scale \(\sigma\) of the transformed feature as
 these are required for computing skewness and kurtosis. Here, \(\mu\) is
@@
 
-The location, or mean, is weighted using weights \(w_i\). For non-robust
-transformations, \(w_i = 1\). Then, \(\sigma^2\) is defined as:
+The location, or mean, is weighted by \(w_i\), which equals \(1\) for
+non-robust transformations.
+
+Then, \(\sigma^2\) is defined as:
 
@@
 The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
-test statistic (D'Agostino and Belanger 1990). This test statistic is
-composed of two separate test statistics, one of which is related to
-skewness, and the other to kurtosis. Both test statistics are computed
-in several steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for
-non-robust power transformations, \(\nu = n\).
-
-For the skewness test statistic we first compute (D'Agostino and
-Belanger 1990):
+test statistic \citep{DAgostino1990-kp}. This test statistic is composed
+of two separate test statistics, one of which is related to skewness,
+and the other to kurtosis. Both test statistics are computed in several
+steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for non-robust
+power transformations, \(\nu = n\).
+
+For the skewness test statistic we first compute
+\citep{DAgostino1990-kp}:
 
@@
 
-For the kurtosis test statistic we first compute (D'Agostino and
-Belanger 1990; Anscombe and Glynn 1983):
+For the kurtosis test statistic we first compute
+\citep{DAgostino1990-kp, Anscombe1983-nz}:
 
@@
 
-\section{Appendix D: Central normality test and empirical central
+\section{Appendix C: Central normality test and empirical central
 normality test critical
-values}\label{appendix-d-central-normality-test-and-empirical-central-normality-test-critical-values}
-
-Critical values for the central normality test and its empirical variant
-are found in Table \ref{tab:central-normality-critical-statistic} and
-Table \ref{tab:empirical-central-normality-critical-statistic},
-respectively.
+values}\label{appendix-c-central-normality-test-and-empirical-central-normality-test-critical-values}
+
+Critical values for the empirical central normality test at
+\(\kappa = 0.80\) are found in Table
+\ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 \caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
-\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
-
-\midrule
-    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
-   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
-   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
-   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
-  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
-  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
-  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
- 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
- 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
- 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
-10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
+central normality at $\kappa = 0.80$, as a function of significance 
+level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
 \label{tab:empirical-central-normality-critical-statistic}
@@
 \midrule
-    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
-   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
-   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
-   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
-  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
-  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
-  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
- 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
- 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
- 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
-10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/empirical-central-normality-critical-statistics-1}
-
-}
-
-\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.80}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
-The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
-with fully normal data where 10\% of elements are replaced by outliers.
-emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
-\end{figure}
+    5 & 200.88 & 90.25 & 64.46 & 49.04 & 36.07 & 27.70 & 19.28 & 13.41 & 10.90 \\
+   10 &  45.90 & 31.78 & 27.12 & 24.14 & 21.23 & 18.30 & 13.68 & 10.20 &  8.71 \\
+   20 &  28.88 & 22.53 & 19.96 & 18.01 & 15.99 & 13.93 & 10.67 &  8.26 &  7.27 \\
+   50 &  16.86 & 14.10 & 12.66 & 11.55 & 10.28 &  8.95 &  7.00 &  5.53 &  4.91 \\
+  100 &  12.41 & 10.00 &  9.02 &  8.21 &  7.36 &  6.45 &  5.02 &  4.00 &  3.58 \\
+  200 &   8.65 &  7.05 &  6.37 &  5.80 &  5.22 &  4.57 &  3.57 &  2.84 &  2.54 \\
+  500 &   5.54 &  4.47 &  4.03 &  3.69 &  3.30 &  2.90 &  2.27 &  1.81 &  1.63 \\
+ 1000 &   3.88 &  3.17 &  2.86 &  2.60 &  2.34 &  2.05 &  1.61 &  1.28 &  1.15 \\
+ 2000 &   2.79 &  2.24 &  2.03 &  1.84 &  1.66 &  1.45 &  1.14 &  0.91 &  0.82 \\ 
+ 5000 &   1.75 &  1.41 &  1.27 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.21 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.36 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\section{Appendix E: Assessing transformations using simulated
-data}\label{appendix-e-assessing-transformations-using-simulated-data}
+\section{Appendix D: Assessing transformations using simulated
+data}\label{appendix-d-assessing-transformations-using-simulated-data}
 
@@
 
-none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
-conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
-conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
-conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
-Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
-Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
-Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
-invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
-robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.47  &         6.78 \\
+conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.10 \\
+conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.34  &         5.84 \\
+conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.35  &         5.84 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.30  &         5.80 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.48  &         3.50 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.53  &         3.51 \\
+invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.51} &         5.58 \\
+robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} & \textbf{3.51} & \textbf{2.00} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Ranking Yeo-Johnson transformations: residual
+errors}\label{ranking-yeo-johnson-transformations-residual-errors}
+
+In the main manuscript, transformation methods are compared by ranking
+within each experiment, defined by the combination of sequence and
+absence/presence of outliers. This allows for direct comparison of
+methods between different experiments.
+
+Table \ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual}
+shows the average normalised residual error for these transformation
+methods. This metric is computed from the residual error and the
+residual error of the central portion (\(\kappa = 0.70\)) for sequences
+without and with outliers, respectively. In each experiment, the
+respective residual error was normalised by dividing with the sequence
+length, and can be interpreted as the average residual error of a single
+sample. This reduces the effect of sequence length on the residual
+error.
+
+Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff}
+shows the average normalised difference in residual error between the
+best method for each experiment and the individual methods. As above,
+normalisation was done by division by the length of the sequence
+corresponding to each experiment.
+
+Both tables show that the differences between best-performing methods on
+each dataset are relatively small.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.202 & 0.112 & 0.128 & 0.080 & 0.129 & 0.080 \\
+conventional                          & & 0.058 & 0.094 & 0.085 & 0.078 & 0.129 & 0.080 \\
+conventional (z-score norm.)          & & 0.061 & 0.094 & 0.091 & 0.073 & 0.091 & 0.073 \\
+conventional (robust scaling)         & & 0.058 & 0.094 & 0.090 & 0.073 & 0.090 & 0.073 \\
+Raymaekers-Rousseeuw                  & & 0.071 & 0.055 & 0.092 & 0.279 & 0.129 & 0.080 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.087 & 0.057 & 0.097 & 0.060 & 0.097 & 0.060 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.071 & 0.056 & 0.095 & 0.059 & 0.095 & 0.059 \\
+invariant                             & & 0.057 & 0.099 & 0.083 & 0.075 & 0.083 & 0.075 \\
+robust invariant                      & & 0.069 & 0.066 & 0.084 & 0.058 & 0.084 & 0.058 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average difference in residual error between Yeo-Johnson
+transformation methods and the best transformation method within each
+experiment, normalised by the corresponding sequence length. This is based on
+either the residual error (without outliers) or the residual error of the
+central portion (with outliers; $\kappa = 0.80$) over 3 datasets with 100000
+sequences each. Residual errors were normalised by sequence length to reduce the
+effect of sequence length on the residual error. The clean dataset consists of
+sequences derived through inverse Yeo-Johnson transformation of data sampled
+from a standard normal distribution. The dirty dataset contains sequences
+sampled from asymmetric generalised normal distributions, centred at 0. The
+shifted dataset also contains sequences sampled from asymmetric generalised
+normal distributions, but centred at 100, and scaled by 0.001. Several
+transformation methods include normalisation before transformation, indicated by
+z-score normalisation (norm.) or robust scaling. A lower difference is better,
+with a minimum value of 0.0. Discrepancies with the rank-based method may occur,
+as normalised differences in residual error are not fully comparable between
+experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.150 & 0.068 & 0.050 & 0.033 & 0.049 & 0.033 \\
+conventional                          & & 0.005 & 0.049 & 0.006 & 0.032 & 0.049 & 0.033 \\
+conventional (z-score norm.)          & & 0.008 & 0.049 & 0.013 & 0.026 & 0.011 & 0.025 \\
+conventional (robust scaling)         & & 0.005 & 0.050 & 0.012 & 0.026 & 0.010 & 0.026 \\
+Raymaekers-Rousseeuw                  & & 0.018 & 0.011 & 0.014 & 0.233 & 0.049 & 0.033 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.034 & 0.012 & 0.018 & 0.014 & 0.017 & 0.013 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.018 & 0.011 & 0.017 & 0.012 & 0.015 & 0.011 \\
+invariant                             & & 0.004 & 0.054 & 0.005 & 0.029 & 0.003 & 0.028 \\
+robust invariant                      & & 0.016 & 0.021 & 0.006 & 0.011 & 0.004 & 0.010 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Ranking Yeo-Johnson transformations: effect of sequence
+length}\label{ranking-yeo-johnson-transformations-effect-of-sequence-length}
+
+The main manuscript shows the overall ranking of different
+Yeo-Johnson-based power transformation methods across three sets of
+problems. To assess the effect of limited sequence lengths, we perform
+ranking on subsets of the experiment. The results for ranking the subset
+of sequences with 30 samples or fewer (\(n = 16001\)) is shown in Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_10_30}. In the clean
+dataset without outliers, where all sequences were sampled from a normal
+distribution and then underwent an inverse Yeo-Johnson transformation,
+conventional and location- and scale-invariant are ranked similarly,
+with no clear advantage to any method. For the remaining datasets, the
+robust location- and scale-invariant method ranked best. It slightly
+outperformed (robustly scaled) conventional Yeo-Johnson transformations
+and (z-score normalised) robust transformations
+\citep{Raymaekers2024-zf}, in datasets without and with outliers,
+respectively.
+
+In the subset for larger sequence lengths, with at least 1000 samples
+per sequence (\(n = 33037\)), the cost of making the location- and
+scale-invariant transformation robust against outliers becomes more
+apparent (Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000}). In the
+clean dataset without outliers, down-weighting of the tails of the
+distribution by the robust location- and scale-invariant method leads to
+larger residuals compared to other methods (Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Each sequence contained 30 samples or fewer. 
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.75  &         6.81  &         7.27  &         6.40  &         7.45  &         6.60 \\
+conventional                          & &         4.19  &         5.47  &         4.48  &         5.78  &         6.55  &         5.94 \\
+conventional (z-score norm.)          & &         4.36  &         5.12  &         5.01  &         4.99  &         4.35  &         4.70 \\
+conventional (robust scaling)         & & \textbf{4.05} &         5.63  &         4.38  &         5.49  &         3.76  &         5.20 \\
+Raymaekers-Rousseeuw                  & &         5.36  &         4.19  &         5.14  &         4.62  &         6.32  &         5.83 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.56  &         4.08  &         5.67  &         4.05  &         5.03  &         3.81 \\
+Raymaekers-Rousseeuw (robust scaling) & &         5.33  &         4.14  &         5.15  &         4.15  &         4.55  &         3.94 \\
+invariant                             & &         4.11  &         5.92  &         4.12  &         5.67  &         3.62  &         5.37 \\
+robust invariant                      & &         4.28  & \textbf{3.65} & \textbf{3.79} & \textbf{3.85} & \textbf{3.37} & \textbf{3.61} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained 30 samples or fewer. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.246 & 0.173 & 0.186 & 0.156 & 0.187 & 0.157 \\
+conventional                          & & 0.139 & 0.142 & 0.142 & 0.132 & 0.187 & 0.157 \\
+conventional (z-score norm.)          & & 0.139 & 0.142 & 0.142 & 0.130 & 0.142 & 0.130 \\
+conventional (robust scaling)         & & 0.139 & 0.142 & 0.141 & 0.130 & 0.141 & 0.130 \\
+Raymaekers-Rousseeuw                  & & 0.211 & 0.151 & 0.183 & 1.515 & 0.187 & 0.157 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.161 & 0.175 & 0.156 & 0.175 & 0.156 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.209 & 0.155 & 0.172 & 0.144 & 0.172 & 0.144 \\
+invariant                             & & 0.141 & 0.144 & 0.140 & 0.132 & 0.140 & 0.132 \\
+robust invariant                      & & 0.141 & 0.125 & 0.140 & 0.119 & 0.140 & 0.119 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 33037 sequences each. Each sequence contained between 1000 and 10000 samples.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         8.85  &         8.14  &         7.35  &         6.29  &         7.41  &         6.87 \\
+conventional                          & &         3.50  &         5.65  &         3.52  &         7.78  &         6.42  &         6.05 \\
+conventional (z-score norm.)          & &         5.24  &         5.94  &         7.34  &         4.98  &         6.26  &         4.79 \\
+conventional (robust scaling)         & &         3.53  &         6.98  &         5.97  &         5.79  &         4.87  &         5.65 \\
+Raymaekers-Rousseeuw                  & &         4.15  &         2.81  &         3.37  &         3.72  &         6.19  &         5.79 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.34  & \textbf{1.86} &         6.57  &         3.39  &         5.49  &         3.30 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.20  &         2.33  &         5.21  &         3.36  &         4.12  &         3.23 \\
+invariant                             & & \textbf{2.36} &         7.59  & \textbf{2.66} &         6.82  & \textbf{1.98} &         6.81 \\
+robust invariant                      & &         7.83  &         3.70  &         3.01  & \textbf{2.88} &         2.26  & \textbf{2.51} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained between 1000 and 10000 samples. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors 
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.186 & 0.091 & 0.106 & 0.053 & 0.106 & 0.053 \\
+conventional                          & & 0.014 & 0.076 & 0.060 & 0.057 & 0.106 & 0.053 \\
+conventional (z-score norm.)          & & 0.021 & 0.076 & 0.070 & 0.050 & 0.070 & 0.050 \\
+conventional (robust scaling)         & & 0.014 & 0.076 & 0.069 & 0.050 & 0.069 & 0.050 \\
+Raymaekers-Rousseeuw                  & & 0.015 & 0.021 & 0.060 & 0.030 & 0.106 & 0.053 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.020 & 0.070 & 0.030 & 0.070 & 0.030 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.015 & 0.020 & 0.068 & 0.030 & 0.068 & 0.030 \\
+invariant                             & & 0.013 & 0.084 & 0.059 & 0.055 & 0.059 & 0.055 \\
+robust invariant                      & & 0.034 & 0.042 & 0.061 & 0.032 & 0.061 & 0.032 \\
 
@@
 Box-Cox transformations, each sequence was shifted prior to inverse
-transformation so that every element after inverse transformation would
-be strictly positive.
+transformation ensuring all elements were strictly positive after
+inverse transformation.
 
@@
 
-\section{Appendix F: Experimental
-results}\label{appendix-f-experimental-results}
+\section{Appendix E: Experimental
+results}\label{appendix-e-experimental-results}
 
@@
 conventional (z-score norm.)          & 11.5 &  19.3 & 49.0 & 326.2 & 33.3 \\
-conventional (robust scaling)         & 11.3 &  31.8 & 53.3 & 324.5 & 32.2 \\
+conventional (robust scaling)         & 11.3 &  33.3 & 55.9 & 326.5 & 31.5 \\
 Raymaekers-Rousseeuw                  & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
 Raymaekers-Rousseeuw (z-score norm.)  & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
-Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
+Raymaekers-Rousseeuw (robust scaling) & 13.4 & 202.9 & 56.9 & 326.4 & 31.5 \\
 invariant                             &  8.8 &  12.2 & 44.0 & 326.4 & 26.8 \\
@@
 
-none                                  & 0.694 & 0.832 & 0.727     & $< 0.001$ & $< 0.001$ \\
-conventional                          & 0.961 & 0.002 & $< 0.001$ & $< 0.001$ & 0.101 \\
-conventional (z-score norm.)          & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
-conventional (robust scaling)         & 0.901 & 0.005 & $< 0.001$ & $< 0.001$ & 0.154 \\
-Raymaekers-Rousseeuw                  & 0.961 & 0.871 & 0.365     & $< 0.001$ & 0.101 \\
-Raymaekers-Rousseeuw (z-score norm.)  & 0.891 & 0.268 & $< 0.001$ & $< 0.001$ & 0.495 \\
-Raymaekers-Rousseeuw (robust scaling) & 0.562 & 0.789 & $< 0.001$ & $< 0.001$ & 0.492 \\
-invariant                             & 0.962 & 0.004 & $< 0.001$ & $< 0.001$ & 0.154 \\
-robust invariant                      & 0.899 & 0.890 & 0.866     & $< 0.001$ & 0.622 \\
+none                                  & 0.541 &    0.731  &    0.546  & $< 0.001$ & $< 0.001$ \\
+conventional                          & 0.919 & $< 0.001$ & $< 0.001$ & $< 0.001$ &    0.032 \\
+conventional (z-score norm.)          & 0.815 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+conventional (robust scaling)         & 0.815 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+Raymaekers-Rousseeuw                  & 0.919 &    0.787  &    0.200  & $< 0.001$ &    0.032 \\
+Raymaekers-Rousseeuw (z-score norm.)  & 0.799 &    0.152  & $< 0.001$ & $< 0.001$ &    0.285 \\
+Raymaekers-Rousseeuw (robust scaling) & 0.799 &    0.674  &    0.060  & $< 0.001$ &    0.285 \\
+invariant                             & 0.919 &    0.001  & $< 0.001$ & $< 0.001$ &    0.057 \\
+robust invariant                      & 0.813 &    0.817  &    0.736  & $< 0.001$ &    0.403 \\
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-Anderson1952-gz}
-Anderson, T W, and D A Darling. 1952. {``Asymptotic Theory of Certain
-{`Goodness of Fit'} Criteria Based on Stochastic Processes.''}
-\emph{Annals of Mathematical Statistics} 23 (2): 193--212.
-\url{https://doi.org/10.1214/aoms/1177729437}.
-
-\bibitem[\citeproctext]{ref-Anscombe1983-nz}
-Anscombe, F J, and William J Glynn. 1983. {``Distribution of the
-Kurtosis Statistic B2 for Normal Samples.''} \emph{Biometrika} 70 (1):
-227--34. \url{https://doi.org/10.1093/biomet/70.1.227}.
-
-\bibitem[\citeproctext]{ref-Cramer1928-rc}
-Cramér, Harald. 1928. {``On the Composition of Elementary Errors.''}
-\emph{Scand. Actuar. J.} 1928 (1): 13--74.
-\url{https://doi.org/10.1080/03461238.1928.10416862}.
-
-\bibitem[\citeproctext]{ref-DAgostino1990-kp}
-D'Agostino, Ralph B, and Albert Belanger. 1990. {``A Suggestion for
-Using Powerful and Informative Tests of Normality.''} \emph{Am. Stat.}
-44 (4): 316--21. \url{https://doi.org/10.2307/2684359}.
-
-\bibitem[\citeproctext]{ref-Huber1981-su}
-Huber, Peter J. 1981. \emph{Robust Statistics}. John Wiley \& Sons.
-\url{https://doi.org/10.1002/0471725250}.
-
-\bibitem[\citeproctext]{ref-Jarque1980-hw}
-Jarque, Carlos M, and Anil K Bera. 1980. {``Efficient Tests for
-Normality, Homoscedasticity and Serial Independence of Regression
-Residuals.''} \emph{Econ. Lett.} 6 (3): 255--59.
-\url{https://doi.org/10.1016/0165-1765(80)90024-5}.
-
-\bibitem[\citeproctext]{ref-Von_Mises1928-ef}
-Mises, Richard von. 1928. \emph{Wahrscheinlichkeit Statistik Und
-Wahrheit}. Schriften Zur Wissenschaftlichen Weltauffassung.
-Springer-Verlag Berlin, Heidelberg.
-\url{https://doi.org/10.1007/978-3-662-36230-3}.
-
-\bibitem[\citeproctext]{ref-Raymaekers2024-zf}
-Raymaekers, Jakob, and Peter J Rousseeuw. 2024. {``Transforming
-Variables to Central Normality.''} \emph{Mach. Learn.} 113 (8):
-4953--75. \url{https://doi.org/10.1007/s10994-021-05960-5}.
-
-\bibitem[\citeproctext]{ref-Shapiro1965-zd}
-Shapiro, S S, and M B Wilk. 1965. {``An Analysis of Variance Test for
-Normality (Complete Samples).''} \emph{Biometrika} 52 (3/4): 591--611.
-\url{https://doi.org/10.2307/2333709}.
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{refs.bib}
 

### 2025-10-29 manuscript_rmarkdown/manuscript_appendix.tex
Fixed and updated text.

@@
 methods. This metric is computed from the residual error and the
-residual error of the central portion (\(\kappa = 0.70\)) for sequences
+residual error of the central portion (\(\kappa = 0.80\)) for sequences
 without and with outliers, respectively. In each experiment, the

### 2025-10-29 manuscript_latex/manuscript.tex
Update manuscript.tex

@@
 By default the style follows the most common style.
-To switch between the options you can add or remove “Numbered” in the optional parenthesis.
+To switch between the options you can add or remove ?Numbered?
+in the optional parenthesis.
 %%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst%
@@
 \newtheorem*{definition}{Definition}
+\newtheorem*{corollary}{Corollary}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 
-Statistical tests for normality, such as the Shapiro-Wilk test \citep{Shapiro1965-zd},
-could be automatically applied to transformed features.
-However, given sufficiently large sample sizes such tests can detect
-trivial deviations from normality, and may lead to rejection of
+Statistical tests for normality, such as the Shapiro-Wilk test
+\citep{Shapiro1965-zd}, could be automatically applied to transformed
+features. However, given sufficiently large sample sizes such tests can
+detect trivial deviations from normality, and may lead to rejection of
 sufficiently good power transformations.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_1}
+{\centering \includegraphics{manuscript_files/figure-latex/decreased-normality-1}
 
@@
 i.e.~using maximum likelihood estimation (MLE). A full derivation of the
-log-likelihood function for both transformations is shown in \hyperref[appendix-a]{Appendix A}.
+log-likelihood function for both transformations is shown in Appendix A.
 The location- and scale-invariant Box-Cox log-likelihood function is:
@@
 transformations to normality, as shown in Figure
-\ref{fig:decreased-normality}. As indicated by \citet{Raymaekers2024-zf},
- the general aim of power transformations should be to transform
-non-outlier data to normality, i.e.~achieve \emph{central normality}. To
-achieve this, they devised an iterative procedure to find a robust
-estimate of the transformation parameter \(\lambda\). Briefly, this
-process requires identifying outliers in the data and weighting such
-instances during the optimisation process. \citet{Raymaekers2024-zf}
-achieve this through weighted maximum likelihood estimation.
-However, because this procedure iteratively estimates and updates
-\(\lambda\), it can not be used here to simultaneously estimate
-\(\lambda\), \(x_0\) and \(s\) for location- and scale-invariant power
-transformations. Nonetheless, as a procedure, weighted MLE can be used
-for estimating the transformation, shift and scale parameters.
+\ref{fig:decreased-normality}. As indicated by
+\citet{Raymaekers2024-zf}, the general aim of power transformations
+should be to transform non-outlier data to normality, i.e.~achieve
+\emph{central normality}. To achieve this, they devised an iterative
+procedure to find a robust estimate of the transformation parameter
+\(\lambda\). Briefly, this process requires identifying outliers in the
+data and weighting such instances during the optimisation process.
+\citet{Raymaekers2024-zf} achieve this through weighted maximum
+likelihood estimation. However, because this procedure iteratively
+estimates and updates \(\lambda\), it can not be used here to
+simultaneously estimate \(\lambda\), \(x_0\) and \(s\) for location- and
+scale-invariant power transformations. Nonetheless, as a procedure,
+weighted MLE can be used for estimating the transformation, shift and
+scale parameters.
 
@@
 \ref{eqn:box-cox-invariant-log-likelihood} and
-\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to \citet{Raymaekers2024-zf},
-these log-likelihood functions include additional
-terms to accommodate estimation of \(x_0\) and \(s\). The weighted
-location- and scale-invariant Box-Cox log-likelihood function is:
+\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to
+\citet{Raymaekers2024-zf}, these log-likelihood functions include
+additional terms to accommodate estimation of \(x_0\) and \(s\). The
+weighted location- and scale-invariant Box-Cox log-likelihood function
+is:
 
@@
 \item
-  A tapered cosine function \citep{Tukey1967-eb}, with \(\delta_1 \geq 0\)
-  and \(\delta_2 \geq \delta_1\) as threshold parameters:
-  \begin{equation*}
+  A tapered cosine function \citep{Tukey1967-eb}, with
+  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
+  parameters: \begin{equation*}
   w_i =
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_2}
+{\centering \includegraphics{manuscript_files/figure-latex/weighting-functions-1}
 
@@
   The weighting function uses the z-score of the transformed feature
-  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After \citep{Raymaekers2024-zf},
+  \(\phi^{\lambda, x_0, s} (\mathbf{X})\). After
+  \citet{Raymaekers2024-zf},
   \(z_i = \frac{\phi^{\lambda, x_0, s}(x_i) - \mu_M}{\sigma_M}\). Here,
@@
 for use with unimodal distributions, the generalised normal distribution
-\citep{Subbotin1923-qk, Nadarajah2005-xe} is a suitable option for simulating
-realistic feature distributions. This distribution has the following
-probability density function \(f_{\beta}\) for a value
+\citep{Subbotin1923-qk, Nadarajah2005-xe} is a suitable option for
+simulating realistic feature distributions. This distribution has the
+following probability density function \(f_{\beta}\) for a value
 \(x \in \mathbb{R}\):
@@
 recipe for introducing skewness into the otherwise symmetric generalised
-normal distribution  \citep{Gijbels2019-te}, leading to
-the following probability density function:
+normal distribution \citep{Gijbels2019-te}, leading to the following
+probability density function:
 
@@
 \(F_{\alpha}^{-1}(p)\), which incorporates the quantile function of the
-symmetric generalised normal distribution derived by Griffin \citep{Gijbels2019-te}:
+symmetric generalised normal distribution derived by Griffin
+\citep{Griffin2018-bf}:
 
@@
 Deviations from normality can be detected by normality tests, such as
-the Shapiro-Wilk test  \citep{Gijbels2019-te}. In practice, normality
+the Shapiro-Wilk test \citep{Shapiro1965-zd}. In practice, normality
 tests may be too stringent with large sample sizes, outliers, or both.
-Here we develop a test for central normality.
+Here we develop an empirical test for central normality.
 
@@
 Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$,
-with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ \citep{Huber1981-su}, 
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su],
 and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
@@
 
-\begin{definition}[Central normality of a sequence]
-
-The central portion of a sequence is normally distributed if the sum of residual
-errors is equal to zero:
-$\sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$
+\begin{definition}[Central normal distribution]
+
+A central normal distribution is any distribution whose quantile function is
+defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
+with $\mathcal{N}$ any normal distribution.
 
@@
 
-In practice, a finite sequence sampled from a normal distribution
-\(\mathcal{N}(\mu, \sigma)\) will have a non-zero sum of residual
-errors.
+\begin{corollary}
+
+For $\kappa = 1.0$, the central normal distribution is a normal distribution.
+
+\end{corollary}
+
+\begin{corollary}
+
+By the Glivenko-Cantelli theorem, if a sequence $\mathbf{X}$ was sampled from a
+central normal distribution, then almost surely
+$\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,
+
+\end{corollary}
 
 \begin{definition}[Central normality test]
-The null-hypothesis is defined as: $\mathcal{H}_0: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j = 0$
-
-The alternative hypothesis is defined as: $\mathcal{H}_1: \sum_{r_j \in \mathbf{R}_{\text{c}}}r_j > 0$
+The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+from a population that is distributed according to a central normal distribution.
+
+The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+population that is not distributed according to a central normal distribution.
 \end{definition}
 
-The null-hypothesis \(\mathcal{H}_0\) is that the central portion of
-sequence \(\mathbf{X}\) is normally distributed, with the alternative
-hypothesis \(\mathcal{H}_1\) that it is not normally distributed. To
-test this hypothesis, a test statistic is computed and compared against
-a critical value.
+To test this hypothesis, a test statistic is computed and compared
+against a critical value.
 
 \begin{definition}[Central normality test statistic]
-Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$
+Let $m$ be the number of elements of $\mathbf{R}_{\text{c}}$.
 
@@
 and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA
-algorithm for derivative-free bound constraint optimisation \citep{Powell2009-zb} 
-through maximum likelihood estimation. The required algorithms
-were implemented in the \texttt{power.transform} R software package
-\citep{Zwanenburg2024-kq} (version 1.0.1). Of note, the
+algorithm for derivative-free bound constraint optimisation
+\citep{Powell2009-zb} through maximum likelihood estimation. The
+required algorithms were implemented in the \texttt{power.transform} R
+software package \citep{Zwanenburg2024-kq} (version 1.0.2). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_3}
+{\centering \includegraphics{manuscript_files/figure-latex/shifted-distributions-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_4}
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-combined-1}
 
@@
 
-\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the (empirical) central normality test as function of $n$ for several values of central portion $\kappa$.
-The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
-with fully normal data where 10\% of elements are replaced by outliers.
-emp: empirical}\label{fig:empirical-central-normality-kappa}
+\caption{Critical test statistic $\tau_{\alpha = 0.05, n, \kappa}$ of the empirical central normality test as function of $n$ for several values of central portion $\kappa$ (left) and several values of significance level $\alpha$ (right).}\label{fig:empirical-central-normality-combined}
 \end{figure}
 
-To assess type I error rates for the (empirical) central normality test
+To assess type I error rates for the empirical central normality test
 and compare these to the Shapiro-Wilk test for normality, we randomly
-drew another 1000 sequences from a standard normal distribution
-(\(\mathcal{N}(0,1)\) for each
+drew another 10000 sequences from a central normal distribution for each
 \(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
-A second dataset was created by replacing elements with randomly drawn
-outliers, as described above. For each sequence, the p-value for the
-respective test was used to reject the null hypothesis that the sequence
-is (centrally) normal. The null hypothesis was rejected if
-\(p \leq 0.05\). The type I error rate was determined by computing the
-fraction of centrally normal sequences rejected this way. Figure
-\ref{fig:empirical-central-normality-error_rate} shows that the
-empirical central normality test had the expected type I error rate for
-centrally normal data (with tails of the distribution not being normally
-distributed), whereas the central normality test and Shapiro-Wilk test
-overestimated type I error rates.
+
+Each sequence was created as follows: first \(n\) values are sampled
+from \(\mathcal{N}(0, 1)\). Then, the elements outside the central
+portion of the sequence
+(i.e.~\(\mathbf{X} \setminus \mathbf{X}_{\text{c}}\)) are replaced.
+Elements with \(p_i < \frac{1-\kappa}{2}\) are replaced with values
+sampled from \(\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))\), and
+elements with \(p_i > \frac{1 + \kappa}{2}\) are replaced with values
+sampled from \(\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)\).
+
+For each sequence, the p-value for the respective test was used to
+reject the null hypothesis that sequences were derived from a population
+with a (central) normal distribution. The null hypothesis was rejected
+if \(p \leq 0.05\). The type I error rate was determined by computing
+the fraction of sequences rejected this way. Figure
+\ref{fig:empirical-central-normality-error_rate} shows that when the
+central portion \(\kappa\) of the central normal distribution is equal
+or greater than the \(\kappa\) parameter for the empirical central
+normality test, the tests are accurate. If \(\kappa\) of the empirical
+central normality test exceeds that the central portion \(\kappa\) for
+the underlying distribution, tests are conservative. The Shapiro-Wilk
+test behaves similar to the empirical central normality test with
+\(\kappa = 1.0\), which is expected. For \(n < 10\), the empirical
+central normality test can be optimistic.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_5}
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
 
@@
 
-\caption{Observed type I error rates for (central) normality tests in the presence of data derived from normal distributions with and without outliers.
-The empirical central normality (ECN) test uses $\kappa = 0.80$.
-CN: central normality}\label{fig:empirical-central-normality-error_rate}
+\caption{Observed type I error rates for central normality tests in the presence of data sampled from central normal distributions with $\kappa \in \left\{0.6, 0.8, 1.0\right\}$.
+Three empirical central normality (ECN) tests were performed, parameterised with same value for the central portion $\kappa$.
+Additionally, the Shapiro-Wilk test was assessed.
+The dashed line indicates the expected type I error rate for rejecting the null hypothesis at significance level $\alpha = 0.05$.
+ECN: empirical central normality}\label{fig:empirical-central-normality-error_rate}
 \end{figure}
@@
 In Figure \ref{fig:empirical-central-normality-examples} we apply the
-empirical central normality test to assess central normality of features
-that are composed of a mixture of samples drawn from two normal
-distributions (\(n = 100\) each). With increased separation of the
-underlying normal distributions, the probability of the feature being
-centrally normal decreases, as expected.
+empirical central normality test (with \(\kappa = 0.80\)) to assess
+central normality of features that are composed of a mixture of samples
+drawn from two normal distributions (\(n = 100\) each). With increased
+separation of the underlying normal distributions, the test's p-value
+decreases, as expected.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_6}
+{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-examples-1}
 
@@
 The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) drawn from normal distributions that are offset by a distance $d$.
-We use the empirical centrally normality test to compute the probability for the hypothesis that the distribution is centrally normal.
-As may be observed, with increasing offset $d$ the probability that the feature is centrally normal decreases.
+The empirical central normality test ($\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from a central normal distribution.
+As may be observed, with increasing offset $d$ the test's p-value decreases.
 Quantile-quantile plots are drawn below each distribution.}\label{fig:empirical-central-normality-examples}
 \end{figure}
+
+We use \(\kappa = 0.80\) for the empirical central normality tests in
+the remainder of the manuscript, unless noted otherwise. Critical
+statistic values of the empirical central normality test for
+\(\kappa = 0.80\) are shown in Table
+\ref{tab:empirical-central-normality-critical-statistic}.
 
@@
 \(\gamma \sim U\left(\log_{10}50, 3\right)\), resulting in a set of
-sequences with between \(50\) and \(1000\) elements each.
-
-Outlier values were then drawn to randomly replace 10 percent of the
-elements of \(\mathbf{X}_i\). Outlier values were set according to \citet{Tukey1977-xm},
-as follows. Let \(x^{*} \sim U\left(-2, 2\right)\).
-Then the corresponding outlier value was:
-
-\begin{equation}
+sequences with between \(50\) and \(1000\) elements each. Outlier values
+were then drawn to randomly replace 10 percent of the elements of
+\(\mathbf{X}_i\). These values were set according to
+\citet{Tukey1977-xm}, as follows. Let
+\(x^{*} \sim U\left(-2, 2\right)\). Then the corresponding outlier value
+was:
+
+\begin{equation}
+\label{eqn:outlier_rate}
 x_{out} =
@@
 \begin{table}
+\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
@@
 \end{tabular}
+\end{center}
 \end{table}
@@
 \begin{table}
+\begin{center}
 \caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
@@
 the central portion (data with outliers). The average rank of each
-transformation method was computed over all 10000 sequences in each
+transformation method was computed over all 100000 sequences in each
 dataset. Average ranks for Yeo-Johnson transformations are shown in
-Table \ref{tab:comparison_methods_simulations_yeo_johnson}. Location and
+Table \ref{tab:comparison_methods_simulation_yeo_johnson}. Location and
 shift-invariant Yeo-Johnson transformation ranked best for all datasets
@@
 \begin{table}
+\begin{center}
 \caption{
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in \hyperref[appendix-e]{Appendix E}.
+for Box-Cox transformations are shown in Appendix D.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_7}
+{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-invariance-1}
 
@@
 Loprinzi et al. \citep{Loprinzi1994-cd}. The age in the cohort was
-\(62.4 \pm 9.1\) (mean $\pm$ standard deviation) years. Applying
+\(62.4 \pm 9.1\) (mean ± standard deviation) years. Applying
 conventional and invariant Yeo-Johnson transformations to patient age
@@
 \ref{fig:experimental-results-invariance}: no transformation
-(\(\sum r_i = 16.5\), \(p = 0.69\)); conventional transformation
+(\(\sum r_i = 16.5\), \(p = 0.54\)); conventional transformation
 (\(\lambda = 2.0\), \(\sum r_i = 11.5\), \(\mu_{YJ} = 1.8 \cdot 10^3\),
-\(\sigma_{YJ} = 0.5 \cdot 10^3\), \(p = 0.96\)); Raymaekers and
+\(\sigma_{YJ} = 0.5 \cdot 10^3\), \(p = 0.92\)); Raymaekers and
 Rousseeuw's robust adaptation (\(\lambda = 2.0\), \(\sum r_i = 11.5\),
 \(\mu_{YJ} = 1.8 \cdot 10^3\), \(\sigma_{YJ} = 0.5 \cdot 10^3\),
-\(p = 0.96\)); location- and scale-invariant transformation
+\(p = 0.92\)); location- and scale-invariant transformation
 (\(\lambda = 1.3\), \(\sum r_i = 8.8\), \(\mu_{YJ} = -1.2\),
-\(\sigma_{YJ} = 1.1\), \(p = 0.98\)); and robust location- and
+\(\sigma_{YJ} = 1.1\), \(p = 0.95\)); and robust location- and
 scale-invariant transformation (\(\lambda = 1.3\), \(\sum r_i = 9.3\),
-\(\mu_{YJ} = -1.0\), \(\sigma_{YJ} = 1.1\), \(p = 0.93\)).
+\(\mu_{YJ} = -1.0\), \(\sigma_{YJ} = 1.1\), \(p = 0.87\)).
 
@@
 significant deviations from central normality for any transformation
-(all \(p \geq 0.93\)).
+(all \(p \geq 0.87\)).
 
@@
 Gorman, Williams and Fraser recorded body mass (in grams) of 342
-penguins of three different species \citep{Gorman2014-eo}.
-The body mass was \((4.2 \pm 0.8) \cdot 10^3\) (mean $\pm$ standard
-deviation) grams, and not centrally normal (\(p < 0.001\)). Applying
-conventional and invariant Yeo-Johnson transformations to body mass
-yielded the following results, see Figure
-\ref{fig:experimental-results-invariance}: no transformation (residual
-sum \(\sum r_i = 48.0\), \(p < 0.001\)); conventional transformation
-(\(\lambda = -0.5\), \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\));
-location- and scale-invariant transformation (\(\lambda = 0.5\),
-\(\sum r_i = 26.8\), \(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\),
-\(p = 0.28\)); and robust location- and scale-invariant transformation
-(\(\lambda = 0.3\), \(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\),
-\(\sigma_{YJ} = 0.9\), \(p = 0.69\)).
-
-Location- and scale-invariant transformation produced a lower overall
-residual errors, indicating a better transformation. Moreover,
-conventional transformations led to low standard deviation
-\(\sigma_{YJ}\) of the body mass feature after transformation. The
-empirical central normality test did not detect any statistically
-significant deviations from central normality for any transformation
-(all \(p \geq 0.10\)).
+penguins of three different species \citep{Gorman2014-eo}. The body mass
+was \((4.2 \pm 0.8) \cdot 10^3\) (mean ± standard deviation) grams, and
+not centrally normal (\(p < 0.001\)). Applying conventional and
+invariant Yeo-Johnson transformations to body mass yielded the following
+results, see Figure \ref{fig:experimental-results-invariance}: no
+transformation (residual sum \(\sum r_i = 48.0\), \(p < 0.001\));
+conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
+\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.03\));
+Raymaekers and Rousseeuw's robust adaptation (\(\lambda = -0.5\),
+\(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.03\)); location- and
+scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
+\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\), \(p = 0.12\)); and robust
+location- and scale-invariant transformation (\(\lambda = 0.3\),
+\(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.9\),
+\(p = 0.48\)).
+
+Location- and scale-invariant transformation yielded lower overall
+residual errors, indicating a better transformation to central
+normality, which is reflected by the empirical central normality test
+results. Conventional transformations led to low standard deviation
+\(\sigma_{YJ}\) of the body mass feature after transformation, and the
+null hypothesis of the empirical central normality test was rejected (at
+significance level \(\alpha = 0.05\)) in both cases.
 
@@
 2010 \citep{De-Cock2011-jf}, including their geospatial coordinates. The
-latitude was \(42.03 \pm 0.02)\) (mean $\pm$ standard deviation). Applying
+latitude was \(42.03 \pm 0.02)\) (mean ± standard deviation). Applying
 conventional and invariant Yeo-Johnson transformations to latitude
@@
 
-Every transformation reduced the residual sum. None of the
-transformations yielded a centrally normal distribution. Conventional
-transformations had high values for the \(\lambda\) parameter, which
-could lead to numerical issues.
+Every transformation reduced the residual sum compared to the original
+data. None of the transformations yielded a centrally normal
+distribution. Conventional transformations had high values for the
+\(\lambda\) parameter, which could lead to numerical issues.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_8}
+{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-outlier-robustness-1}
 
@@
 The Top Gear dataset contains data on 297 vehicles that appeared on the
-BBC television show \emph{Top Gear} \citep{Alfons2021-kc}. Within this dataset,
-the fuel consumption feature contains outliers due to highly
+BBC television show \emph{Top Gear} \citep{Alfons2021-kc}. Within this
+dataset, the fuel consumption feature contains outliers due to highly
 fuel-efficient vehicles. Applying conventional and invariant Yeo-Johnson
@@
 results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum \(\sum r_i = 54\), \(p=0.72\));
+no transformation (residual sum \(\sum r_i = 54\), \(p=0.55\));
 conventional transformation (\(\lambda = -0.1\), \(\sum r_i = 55\),
@@
 and Rousseeuw's robust adaptation (\(\lambda = 0.8\), \(\sum r_i = 48\),
-\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.37\)); location- and
+\(\mu_{YJ} = 29\), \(\sigma_{YJ} = 15\), \(p=0.21\)); location- and
 scale-invariant transformation (\(\lambda = -1.3\), \(\sum r_i = 44\),
@@
 \(\sum r_i = 56\), \(\mu_{YJ} = 1.7\), \(\sigma_{YJ} = 1.0\),
-\(p=0.76\)).
-
-Outliers cause non-robust transformations to fail to transform the data
-to a centrally normal distribution (empirical central normality test
-\(p < 0.001\) for conventional and invariant transformations). Robust
-transformations produce distributions that are centrally normal
-(empirical central normality test \(p > 0.05\)).
+\(p=0.59\)).
+
+Outliers caused non-robust transformations to fail the empirical central
+normality test for transforming the feature to a centrally normal
+distribution.
 
@@
 results, see Figure \ref{fig:experimental-results-outlier-robustness}:
-no transformation (residual sum \(\sum r_i = 110\), \(p=0.83\));
+no transformation (residual sum \(\sum r_i = 110\), \(p=0.73\));
 conventional transformation (\(\lambda = -0.7\), \(\sum r_i = 30\),
-\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p=0.003\)); Raymaekers and
-Rousseeuw's robust adaptation (\(\lambda = 1.1\), \(\sum r_i = 136\),
-\(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14.3\), \(p=0.88\)); location- and
-scale-invariant transformation (\(\lambda = 0.2\), \(\sum r_i = 12\),
-\(\mu_{YJ} = -11.8\), \(\sigma_{YJ} = 6.9\), \(p=0.15\)); and robust
-location- and scale-invariant transformation (\(\lambda = -0.3\),
-\(\sum r_i = 30\), \(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.2\),
-\(p=0.18\)).
-
-The conventional non-robust transformation failed to produce a centrally
-normal distribution (empirical central normality test \(p=0.003\)).
-Robust transformations produce distributions that are centrally normal
-(empirical central normality test \(p > 0.05\)).
+\(\mu_{YJ} = 1.0\), \(\sigma_{YJ} = 0.1\), \(p < 0.001\)); Raymaekers
+and Rousseeuw's robust adaptation (\(\lambda = 1.1\),
+\(\sum r_i = 136\), \(\mu_{YJ} = 7.2\), \(\sigma_{YJ} = 14.3\),
+\(p=0.80\)); location- and scale-invariant transformation
+(\(\lambda = 0.2\), \(\sum r_i = 12\), \(\mu_{YJ} = -11.8\),
+\(\sigma_{YJ} = 6.9\), \(p=0.07\)); and robust location- and
+scale-invariant transformation (\(\lambda = -0.3\), \(\sum r_i = 30\),
+\(\mu_{YJ} = 0.8\), \(\sigma_{YJ} = 0.2\), \(p=0.09\)).
+
+For conventional non-robust transformation the null hypothesis was
+rejected (at \(\alpha = 0.05\)), indicating that they failed to produce
+a transformed feature that followed a central normal distribution.
 
@@
 datasets to classification tasks. Using the familiar auto-machine
-learning library \citep{Zwanenburg2021-so} (version 1.5.0), each
-dataset was used to train a model for each of 32 process configurations.
-Each process configuration specifies the learner (generalised linear
-model, L1-regularised linear models (Lasso), gradient boosted linear
-model, or random forest), transformation method (none, conventional
-Yeo-Johnson, robust invariant Yeo-Johnson, robust invariant Yeo-Johnson
-with empirical central normality test (rejecting transformations with
+learning library \citep{Zwanenburg2021-so} (version 1.5.0), each dataset
+was used to train a model for each of 32 process configurations. Each
+process configuration specifies the learner (generalised linear model,
+L1-regularised linear models (Lasso), gradient boosted linear model, or
+random forest), transformation method (none, conventional Yeo-Johnson,
+robust invariant Yeo-Johnson, robust invariant Yeo-Johnson with
+empirical central normality test (rejecting transformations with
 \(p \leq 0.01\)), and normalisation method (none,
@@
 \begin{itemize}
+\tightlist
 \item
@@
 transformation method, we first fit a regression random forest (ranger
-package \citep{Wright2017-rf} version 0.16.0): 2000 trees, node size
-2, other hyperparameters default) with process parameters and task
+package \citep{Wright2017-rf} version 0.16.0): 2000 trees, node size 2,
+other hyperparameters default) with process parameters and task
 difficulty as predictors and normalised rank as response variable. The
@@
 
-{\centering \includegraphics[width=1\linewidth]{figure_9}
+{\centering \includegraphics{manuscript_files/figure-latex/marginal-effect-plot-1}
 
@@
 
-\FloatBarrier
-
 \section{Discussion}\label{discussion}
@@
 \citep{Box1964-mz}, which subsequently became the convention. Yeo and
-Johnson's power transformation lacks a shift parameter altogether \citep{Yeo2000-vw}.
-We showed that these power transformations are
+Johnson's power transformation lacks a shift parameter altogether
+\citep{Yeo2000-vw}. We showed that these power transformations are
 sensitive to location and scale of data distributions. To mitigate this
@@
 sequences deviate from normality in a way that might require closer
-inspection prior to further processing. The empirical test for central
-normality differs from other tests for normality, such as the
-Shapiro-Wilk test \citep{Shapiro1965-zd}, as it assesses normality of
-the central portion of a feature, instead of the entire feature.
-Compared to the central normality test and the Shapiro-Wilk test, the
-empirical central normality test remains consistent in the presence of
-outliers, although the former tests are more powerful when outliers are
-absent.
+inspection prior to further processing. The empirical central normality
+test differs from other tests for normality, such as the Shapiro-Wilk
+test \citep{Shapiro1965-zd}, as it assesses normality of the central
+portion of a feature, instead of the entire feature. Compared to the
+Shapiro-Wilk test, the empirical central normality test remains
+consistent when a feature was not sampled from a normal distribution,
+such as a central normal distribution with \(\kappa < 1.0\). This
+enables assessing reasonable normality of (transformed) features in the
+presence of outliers.
 
 This work has the following limitation: We observed several numerical
-stability issues for optimisation criteria other than MLE (\hyperref[appendix-b]{Appendix B}).
+stability issues for optimisation criteria other than MLE (Appendix B).
 These appear in regions where transformation parameters would lead to

### 2025-10-29 manuscript_latex/manuscript.tex
Update manuscript.tex

@@
 
-\section{Log-likelihood functions for location- and
-scale-invariant power
+\section{Log-likelihood functions for location- and scale-invariant power
 transformation}\label{appendix-a}

### 2025-10-29 manuscript_latex/manuscript.tex
Update Appendix B.

@@
 
+\FloatBarrier
+
 \section{Discussion}\label{discussion}
@@
 \(\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)\)
-its transformed values using conventional and shift and scale invariant
-power transformations, respectively. Since power transformations are
-monotonic, \(\mathbf{Y}\) will likewise be ordered.
+its transformed values using conventional and location- and
+scale-invariant power transformations, respectively. Since power
+transformations are monotonic, \(\mathbf{Y}\) will likewise be ordered.
 
 Below we will focus on criteria based on the empirical density function
-and those based on skewness and kurtosis of the transformed featured.
+and those based on skewness and kurtosis of the transformed feature.
 Other potential criteria, such as the Shapiro-Wilk test statistic
-\citet{Raymaekers2024-zf} are not investigated here. In the case of the
+\citep{Shapiro1965-zd} are not investigated here. In the case of the
 Shapiro-Wilk test statistic this is because of lack of scalability to
@@
 EDF and the normal CDF, with differences at the upper and lower end of
-the normal CDF receiving more weight than those at the the centre of the
+the normal CDF receiving more weight than those at the centre of the
 CDF.
 
-\subsubsection{Cram\'{e}r-von Mises
-criterion}\label{cramer-von-mises-criterion}
-
-The Cram\'{e}r-von Mises criterion is also based on the empirical
-distribution function of \(\mathbf{X}\). We define the Cram\'{e}r-von Mises
+\subsubsection{Cramér-von Mises
+criterion}\label{cramuxe9r-von-mises-criterion}
+
+The Cramér-von Mises criterion is also based on the empirical
+distribution function of \(\mathbf{X}\). We define the Cramér-von Mises
 criterion as follows:
@@
 scale parameters. For non-robust power transformations, all \(w_i = 1\).
-The criterion is similar to the Cram\'{e}r-von Mises test statistic
-\citep{Cramer1928-rc, Von_Mises1928-ef}, aside from a additive scalar value and the introduction of
-weights. This criterion, like the Anderson-Darling criterion, seeks to
-minimise the squared Euclidean distance between the EDF and the normal
-CDF. Unlike the Anderson-Darling criterion, this criterion weights all
-instances equally.
+The criterion is similar to the Cramér-von Mises test statistic
+\citep[\citet{Von_Mises1928-ef}]{Cramer1928-rc}, aside from an additive
+scalar value and the introduction of weights. This criterion, like the
+Anderson-Darling criterion, seeks to minimise the squared Euclidean
+distance between the EDF and the normal CDF. Unlike the Anderson-Darling
+criterion, this criterion weights all instances equally.
 
@@
 substituted by \(\phi^{\lambda} (\mathbf{X})\) in the definition of the
-Cram\'{e}r-von Mises criterion.
+Cramér-von Mises criterion.
 
@@
 The D'Agostino criterion defined here follows the D'Agostino \(K^2\)
-test statistic \citep{DAgostino1990-kp}. This test statistic is
-composed of two separate test statistics, one of which is related to
-skewness, and the other to kurtosis. Both test statistics are computed
-in several steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for
-non-robust power transformations, \(\nu = n\).
-
-For the skewness test statistic we first compute \citep{DAgostino1990-kp}:
+test statistic \citep{DAgostino1990-kp}. This test statistic is composed
+of two separate test statistics, one of which is related to skewness,
+and the other to kurtosis. Both test statistics are computed in several
+steps. Let us first define \(\nu=\sum_{i=1}^n w_i\). Thus for non-robust
+power transformations, \(\nu = n\).
+
+For the skewness test statistic we first compute
+\citep{DAgostino1990-kp}:
 
@@
 
-For the kurtosis test statistic we first compute \citep{Anscombe1983-nz, DAgostino1990-kp}:
+For the kurtosis test statistic we first compute
+\citep{DAgostino1990-kp, Anscombe1983-nz}:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{appendix_figure_1}
+{\centering \includegraphics{manuscript_appendix_files/figure-latex/shifted-distributions-appendix-1}
 

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix C

@@
 
-\section{Central normality test and empirical central
-normality test critical
-values}\label{appendix-c}
-
-Critical values for the central normality test and its empirical variant
-are found in Table \ref{tab:central-normality-critical-statistic} and
-Table \ref{tab:empirical-central-normality-critical-statistic},
-respectively.
+\section{Appendix C: Empirical central normality test critical values}
+\label{appendix-c}
+
+Critical values for the empirical central normality test at
+\(\kappa = 0.80\) are found in Table
+\ref{tab:empirical-central-normality-critical-statistic}.
 
 \begin{table}
+\begin{center}
 \caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
-\textbf{central normality} at $\kappa = 0.80$, as a function of significance
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:central-normality-critical-statistic}
+central normality at $\kappa = 0.80$, as a function of significance
+level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
+\label{tab:empirical-central-normality-critical-statistic}
 \begin{tabular}{l | r r r r r r r r r}
@@
 \midrule
-    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
-   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
-   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
-   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
-  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
-  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\
-  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
- 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
- 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
- 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
-10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
+    5 & 200.88 & 90.25 & 64.46 & 49.04 & 36.07 & 27.70 & 19.28 & 13.41 & 10.90 \\
+   10 &  45.90 & 31.78 & 27.12 & 24.14 & 21.23 & 18.30 & 13.68 & 10.20 &  8.71 \\
+   20 &  28.88 & 22.53 & 19.96 & 18.01 & 15.99 & 13.93 & 10.67 &  8.26 &  7.27 \\
+   50 &  16.86 & 14.10 & 12.66 & 11.55 & 10.28 &  8.95 &  7.00 &  5.53 &  4.91 \\
+  100 &  12.41 & 10.00 &  9.02 &  8.21 &  7.36 &  6.45 &  5.02 &  4.00 &  3.58 \\
+  200 &   8.65 &  7.05 &  6.37 &  5.80 &  5.22 &  4.57 &  3.57 &  2.84 &  2.54 \\
+  500 &   5.54 &  4.47 &  4.03 &  3.69 &  3.30 &  2.90 &  2.27 &  1.81 &  1.63 \\
+ 1000 &   3.88 &  3.17 &  2.86 &  2.60 &  2.34 &  2.05 &  1.61 &  1.28 &  1.15 \\
+ 2000 &   2.79 &  2.24 &  2.03 &  1.84 &  1.66 &  1.45 &  1.14 &  0.91 &  0.82 \\
+ 5000 &   1.75 &  1.41 &  1.27 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
+10000 &   1.21 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.36 \\
 \bottomrule
 \end{tabular}
+\end{center}
 \end{table}
-
-\begin{table}
-\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
-\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance
-level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
-\label{tab:empirical-central-normality-critical-statistic}
-\begin{tabular}{l | r r r r r r r r r}
-
-\toprule
-$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\
-
-\midrule
-    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
-   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
-   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
-   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\
-  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
-  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
-  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
- 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
- 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
- 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
-10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{appendix_figure_2}
-
-}
-
-\caption{Critical test statistic $\tau_{\alpha, n, \kappa = 0.80}$ of the (empirical) central normality test as function of $n$ for several values of significance level $\alpha$.
-The critical test statistics for central normality test are determined using fully normal data, whereas the statistics for the empirical variant are determined using centrally normal data, i.e.
-with fully normal data where 10\% of elements are replaced by outliers.
-emp: empirical}\label{fig:empirical-central-normality-critical-statistics}
-\end{figure}
 

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix d.

@@
 
+\subsection{Ranking Box-Cox
+transformations}\label{ranking-box-cox-transformations}
+
 Box-Cox transformations were assessed in the same manner as Yeo-Johnson
@@
 \begin{table}
+\begin{center}
 \caption{
 Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
@@
 
-none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
-conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
-conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
-conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
-Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\
-Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
-invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
-robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\
+none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.47  &         6.78 \\
+conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.10 \\
+conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.34  &         5.84 \\
+conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.35  &         5.84 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.30  &         5.80 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.48  &         3.50 \\
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.53  &         3.51 \\
+invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.51} &         5.58 \\
+robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} & \textbf{3.51} & \textbf{2.00} \\
 
@@
 \end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Ranking Yeo-Johnson transformations: residual
+errors}\label{ranking-yeo-johnson-transformations-residual-errors}
+
+In the main manuscript, transformation methods are compared by ranking
+within each experiment, defined by the combination of sequence and
+absence/presence of outliers. This allows for direct comparison of
+methods between different experiments.
+
+Table \ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual}
+shows the average normalised residual error for these transformation
+methods. This metric is computed from the residual error and the
+residual error of the central portion (\(\kappa = 0.80\)) for sequences
+without and with outliers, respectively. In each experiment, the
+respective residual error was normalised by dividing with the sequence
+length, and can be interpreted as the average residual error of a single
+sample. This reduces the effect of sequence length on the residual
+error.
+
+Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff}
+shows the average normalised difference in residual error between the
+best method for each experiment and the individual methods. As above,
+normalisation was done by division by the length of the sequence
+corresponding to each experiment.
+
+Both tables show that the differences between best-performing methods on
+each dataset are relatively small.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.202 & 0.112 & 0.128 & 0.080 & 0.129 & 0.080 \\
+conventional                          & & 0.058 & 0.094 & 0.085 & 0.078 & 0.129 & 0.080 \\
+conventional (z-score norm.)          & & 0.061 & 0.094 & 0.091 & 0.073 & 0.091 & 0.073 \\
+conventional (robust scaling)         & & 0.058 & 0.094 & 0.090 & 0.073 & 0.090 & 0.073 \\
+Raymaekers-Rousseeuw                  & & 0.071 & 0.055 & 0.092 & 0.279 & 0.129 & 0.080 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.087 & 0.057 & 0.097 & 0.060 & 0.097 & 0.060 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.071 & 0.056 & 0.095 & 0.059 & 0.095 & 0.059 \\
+invariant                             & & 0.057 & 0.099 & 0.083 & 0.075 & 0.083 & 0.075 \\
+robust invariant                      & & 0.069 & 0.066 & 0.084 & 0.058 & 0.084 & 0.058 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average difference in residual error between Yeo-Johnson
+transformation methods and the best transformation method within each
+experiment, normalised by the corresponding sequence length. This is based on
+either the residual error (without outliers) or the residual error of the
+central portion (with outliers; $\kappa = 0.80$) over 3 datasets with 100000
+sequences each. Residual errors were normalised by sequence length to reduce the
+effect of sequence length on the residual error. The clean dataset consists of
+sequences derived through inverse Yeo-Johnson transformation of data sampled
+from a standard normal distribution. The dirty dataset contains sequences
+sampled from asymmetric generalised normal distributions, centred at 0. The
+shifted dataset also contains sequences sampled from asymmetric generalised
+normal distributions, but centred at 100, and scaled by 0.001. Several
+transformation methods include normalisation before transformation, indicated by
+z-score normalisation (norm.) or robust scaling. A lower difference is better,
+with a minimum value of 0.0. Discrepancies with the rank-based method may occur,
+as normalised differences in residual error are not fully comparable between
+experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_avg_residual_diff}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.150 & 0.068 & 0.050 & 0.033 & 0.049 & 0.033 \\
+conventional                          & & 0.005 & 0.049 & 0.006 & 0.032 & 0.049 & 0.033 \\
+conventional (z-score norm.)          & & 0.008 & 0.049 & 0.013 & 0.026 & 0.011 & 0.025 \\
+conventional (robust scaling)         & & 0.005 & 0.050 & 0.012 & 0.026 & 0.010 & 0.026 \\
+Raymaekers-Rousseeuw                  & & 0.018 & 0.011 & 0.014 & 0.233 & 0.049 & 0.033 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.034 & 0.012 & 0.018 & 0.014 & 0.017 & 0.013 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.018 & 0.011 & 0.017 & 0.012 & 0.015 & 0.011 \\
+invariant                             & & 0.004 & 0.054 & 0.005 & 0.029 & 0.003 & 0.028 \\
+robust invariant                      & & 0.016 & 0.021 & 0.006 & 0.011 & 0.004 & 0.010 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\FloatBarrier
+
+\subsection{Ranking Yeo-Johnson transformations: effect of sequence
+length}\label{ranking-yeo-johnson-transformations-effect-of-sequence-length}
+
+The main manuscript shows the overall ranking of different
+Yeo-Johnson-based power transformation methods across three sets of
+problems. To assess the effect of limited sequence lengths, we perform
+ranking on subsets of the experiment. The results for ranking the subset
+of sequences with 30 samples or fewer (\(n = 16001\)) is shown in Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_10_30}. In the clean
+dataset without outliers, where all sequences were sampled from a normal
+distribution and then underwent an inverse Yeo-Johnson transformation,
+conventional and location- and scale-invariant are ranked similarly,
+with no clear advantage to any method. For the remaining datasets, the
+robust location- and scale-invariant method ranked best. It slightly
+outperformed (robustly scaled) conventional Yeo-Johnson transformations
+and (z-score normalised) robust transformations
+\citep{Raymaekers2024-zf}, in datasets without and with outliers,
+respectively.
+
+In the subset for larger sequence lengths, with at least 1000 samples
+per sequence (\(n = 33037\)), the cost of making the location- and
+scale-invariant transformation robust against outliers becomes more
+apparent (Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000}). In the
+clean dataset without outliers, down-weighting of the tails of the
+distribution by the robust location- and scale-invariant method leads to
+larger residuals compared to other methods (Table
+\ref{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}.
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Each sequence contained 30 samples or fewer.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         7.75  &         6.81  &         7.27  &         6.40  &         7.45  &         6.60 \\
+conventional                          & &         4.19  &         5.47  &         4.48  &         5.78  &         6.55  &         5.94 \\
+conventional (z-score norm.)          & &         4.36  &         5.12  &         5.01  &         4.99  &         4.35  &         4.70 \\
+conventional (robust scaling)         & & \textbf{4.05} &         5.63  &         4.38  &         5.49  &         3.76  &         5.20 \\
+Raymaekers-Rousseeuw                  & &         5.36  &         4.19  &         5.14  &         4.62  &         6.32  &         5.83 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.56  &         4.08  &         5.67  &         4.05  &         5.03  &         3.81 \\
+Raymaekers-Rousseeuw (robust scaling) & &         5.33  &         4.14  &         5.15  &         4.15  &         4.55  &         3.94 \\
+invariant                             & &         4.11  &         5.92  &         4.12  &         5.67  &         3.62  &         5.37 \\
+robust invariant                      & &         4.28  & \textbf{3.65} & \textbf{3.79} & \textbf{3.85} & \textbf{3.37} & \textbf{3.61} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained 30 samples or fewer. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_10_30_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.246 & 0.173 & 0.186 & 0.156 & 0.187 & 0.157 \\
+conventional                          & & 0.139 & 0.142 & 0.142 & 0.132 & 0.187 & 0.157 \\
+conventional (z-score norm.)          & & 0.139 & 0.142 & 0.142 & 0.130 & 0.142 & 0.130 \\
+conventional (robust scaling)         & & 0.139 & 0.142 & 0.141 & 0.130 & 0.141 & 0.130 \\
+Raymaekers-Rousseeuw                  & & 0.211 & 0.151 & 0.183 & 1.515 & 0.187 & 0.157 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.161 & 0.175 & 0.156 & 0.175 & 0.156 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.209 & 0.155 & 0.172 & 0.144 & 0.172 & 0.144 \\
+invariant                             & & 0.141 & 0.144 & 0.140 & 0.132 & 0.140 & 0.132 \\
+robust invariant                      & & 0.141 & 0.125 & 0.140 & 0.119 & 0.140 & 0.119 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 33037 sequences each. Each sequence contained between 1000 and 10000 samples.
+The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & &         8.85  &         8.14  &         7.35  &         6.29  &         7.41  &         6.87 \\
+conventional                          & &         3.50  &         5.65  &         3.52  &         7.78  &         6.42  &         6.05 \\
+conventional (z-score norm.)          & &         5.24  &         5.94  &         7.34  &         4.98  &         6.26  &         4.79 \\
+conventional (robust scaling)         & &         3.53  &         6.98  &         5.97  &         5.79  &         4.87  &         5.65 \\
+Raymaekers-Rousseeuw                  & &         4.15  &         2.81  &         3.37  &         3.72  &         6.19  &         5.79 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.34  & \textbf{1.86} &         6.57  &         3.39  &         5.49  &         3.30 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.20  &         2.33  &         5.21  &         3.36  &         4.12  &         3.23 \\
+invariant                             & & \textbf{2.36} &         7.59  & \textbf{2.66} &         6.82  & \textbf{1.98} &         6.81 \\
+robust invariant                      & &         7.83  &         3.70  &         3.01  & \textbf{2.88} &         2.26  & \textbf{2.51} \\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption{
+Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
+Each sequence contained between 1000 and 10000 samples. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
+A lower normalised residual error is better. Discrepancies with the rank-based method may occur, as even with normalisation residual errors
+are not fully comparable between experiments, whereas the ranks achieved within each experiment are.
+}
+\label{tab:comparison_methods_simulations_yeo_johnson_1000_10000_avg_residual}
+\begin{tabular}{l | l r r r r r r}
+
+\toprule
+& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
+transformation & outliers: & no & yes & no & yes & no & yes \\
+
+\midrule
+
+none                                  & & 0.186 & 0.091 & 0.106 & 0.053 & 0.106 & 0.053 \\
+conventional                          & & 0.014 & 0.076 & 0.060 & 0.057 & 0.106 & 0.053 \\
+conventional (z-score norm.)          & & 0.021 & 0.076 & 0.070 & 0.050 & 0.070 & 0.050 \\
+conventional (robust scaling)         & & 0.014 & 0.076 & 0.069 & 0.050 & 0.069 & 0.050 \\
+Raymaekers-Rousseeuw                  & & 0.015 & 0.021 & 0.060 & 0.030 & 0.106 & 0.053 \\
+Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.020 & 0.070 & 0.030 & 0.070 & 0.030 \\
+Raymaekers-Rousseeuw (robust scaling) & & 0.015 & 0.020 & 0.068 & 0.030 & 0.068 & 0.030 \\
+invariant                             & & 0.013 & 0.084 & 0.059 & 0.055 & 0.059 & 0.055 \\
+robust invariant                      & & 0.034 & 0.042 & 0.061 & 0.032 & 0.061 & 0.032 \\
+
+\bottomrule
+\end{tabular}
+\end{center}
 \end{table}

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix d, part 2.

@@
 \begin{table}
+\begin{center}
 \caption{Residual errors for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
@@
 }
+\end{center}
 \end{table}
@@
 \begin{table}
+\begin{center}
 \caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{
-p-values of empirical central normality tests for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Yeo-Johnson transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}
@@
 \begin{table}
+\begin{center}
 \caption{Residual errors for features from simulated clean data without outliers after Box-Cox transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{Transformation parameter $\lambda$ for features from simulated clean data without outliers after Box-Cox transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{
-p-values of empirical central normality tests for features from simulated clean data without outliers after Box-Cox transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from simulated clean data without outliers after Box-Cox transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.}

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix e.

@@
 
-\section{Experimental
-results}\label{appendix-e}
+\section{Experimental results}\label{appendix-e}
 
@@
 \begin{table}
+\begin{center}
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{
-p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from real-world datasets after Yeo-Johnson transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 \begin{table}
+\begin{center}
 \caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality.
@@
 \begin{table}
+\begin{center}
 \caption{
-p-values of empirical central normality tests for features from real-world datasets after Box-Cox transformation to normality.
+p-values of empirical central normality tests ($\kappa = 0.80$) for features from real-world datasets after Box-Cox transformation to normality.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.

### 2025-10-29 manuscript_latex/manuscript.tex
Update appendix references.

@@
 because of its ability to handle features with negative values. Results
-for Box-Cox transformations are shown in Appendix D.
+for Box-Cox transformations are shown in \hyperref[appendix-d]{Appendix D}.
 
@@
 
-\section{Appendix C: Empirical central normality test critical values}
+\section{Empirical central normality test critical values}
 \label{appendix-c}

### 2025-10-29 manuscript_latex/manuscript.tex
Updated figures from main manuscript.

@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/decreased-normality-1}
+{\centering \includegraphics[width=1\linewidth]{figure-1}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/weighting-functions-1}
+{\centering \includegraphics[width=1\linewidth]{figure-2}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/shifted-distributions-1}
+{\centering \includegraphics[width=1\linewidth]{figure-3}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-combined-1}
+{\centering \includegraphics[width=1\linewidth]{figure-4}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-error_rate-1}
+{\centering \includegraphics[width=1\linewidth]{figure-5}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/empirical-central-normality-examples-1}
+{\centering \includegraphics[width=1\linewidth]{figure-6}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-invariance-1}
+{\centering \includegraphics[width=1\linewidth]{figure-7}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/experimental-results-outlier-robustness-1}
+{\centering \includegraphics[width=1\linewidth]{figure-8}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-latex/marginal-effect-plot-1}
+{\centering \includegraphics[width=1\linewidth]{figure-9}
 

### 2025-10-29 manuscript_latex/manuscript.tex
Added appendix figure.
Remove double defined section heading.

@@
 Let the residual error for each element of $\mathbf{X}$ be $r_i =\left| \frac{x_i - \mu_M}{\sigma_M} - F^{-1}_{\mathcal{N}}(p_i) \right|$,
-with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ [@Huber1981-su],
+with $\mu_M$ and $\sigma_M$ robust Huber M-estimates of location and scale of $\mathbf{X}$ \citep{Huber1981-su},
 and $F^{-1}_{\mathcal{N}}$ the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
@@
 
-{\centering \includegraphics{manuscript_appendix_files/figure-latex/shifted-distributions-appendix-1}
+{\centering \includegraphics[width=1\linewidth]{figure-appendix-1}
 
@@
 data}\label{appendix-d}
-
-\subsection{Ranking Box-Cox
-transformations}\label{ranking-box-cox-transformations}
 
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality.
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality.
@@
 \begin{table}
-\begin{center}
 \caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality.
@@
 \begin{table}
-\begin{center}
 \caption{Transformation parameter $\lambda$ for features from real-world datasets after Box-Cox transformation to normality.

### 2025-10-29 manuscript_rmarkdown/manuscript.tex

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-10-28}
+\date{2025-10-29}
 
@@
 required algorithms were implemented in the \texttt{power.transform} R
-software package \citep{Zwanenburg2024-kq} (version 1.0.1). Of note, the
+software package \citep{Zwanenburg2024-kq} (version 1.0.2). Of note, the
 \texttt{power.transform} package shifts feature values into the positive
@@
 (i.e.~\(\mathbf{X} \setminus \mathbf{X}_{\text{c}}\)) are replaced.
-Elements with \(pi < \frac{1-\kappa}{2}\) are replaced with values
+Elements with \(p_i < \frac{1-\kappa}{2}\) are replaced with values
 sampled from \(\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))\), and
@@
 Additionally, the Shapiro-Wilk test was assessed.
+The dashed line indicates the expected type I error rate for rejecting the null hypothesis at significance level $\alpha = 0.05$.
 ECN: empirical central normality}\label{fig:empirical-central-normality-error_rate}
@@
 The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) drawn from normal distributions that are offset by a distance $d$.
-The empirical central normality ($\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from a central normal distribution.
+The empirical central normality test ($\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from a central normal distribution.
 As may be observed, with increasing offset $d$ the test's p-value decreases.
@@
 conventional transformation (\(\lambda = -0.5\), \(\sum r_i = 32.2\),
-\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\));
+\(\mu_{YJ} = 2.1\), \(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.03\));
 Raymaekers and Rousseeuw's robust adaptation (\(\lambda = -0.5\),
 \(\sum r_i = 32.2\), \(\mu_{YJ} = 2.1\),
-\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.10\)); location- and
+\(\sigma_{YJ} = 4 \cdot 10^{-3}\), \(p = 0.03\)); location- and
 scale-invariant transformation (\(\lambda = 0.5\), \(\sum r_i = 26.8\),
-\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\), \(p = 0.28\)); and robust
+\(\mu_{YJ} = 0.9\), \(\sigma_{YJ} = 0.9\), \(p = 0.12\)); and robust
 location- and scale-invariant transformation (\(\lambda = 0.3\),
 \(\sum r_i = 22.0\), \(\mu_{YJ} = 0.7\), \(\sigma_{YJ} = 0.9\),
-\(p = 0.69\)).
-
-Location- and scale-invariant transformation produced a lower overall
-residual errors, indicating a better transformation. Moreover,
-conventional transformations led to low standard deviation
-\(\sigma_{YJ}\) of the body mass feature after transformation. The
-empirical central normality test did not detect any statistically
-significant deviations from central normality for any transformation
-(all \(p \geq 0.10\)).
+\(p = 0.48\)).
+
+Location- and scale-invariant transformation yielded lower overall
+residual errors, indicating a better transformation to central
+normality, which is reflected by the empirical central normality test
+results. Conventional transformations led to low standard deviation
+\(\sigma_{YJ}\) of the body mass feature after transformation, and the
+null hypothesis of the empirical central normality test was rejected (at
+significance level \(\alpha = 0.05\)) in both cases.
 

### 2025-12-15 manuscript_rmarkdown/manuscript.Rmd
Update scale for assessing transformation using simulated data.

@@
   and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-
-- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)$,
+  The sequences are drawn from a range of skewed, unimodal, but not necessarily 
+  normal, distributions, to reflect less \textit{clean} data then that presented in the first dataset.
+
+- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)$,
   with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
   and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
+  The generating process creates sequences similar to that in the \textit{dirty}
+  dataset, but shifted to a location that leads to normalisation issues observed
+  for conventional power transformation methods, as shown in Figure \ref{fig:decreased-normality}.
 
@@
 Average ranks for Yeo-Johnson transformations are shown in Table
-\ref{tab:comparison_methods_simulation_yeo_johnson}.
+\ref{tab:comparison_methods_simulations_yeo_johnson}.
 Location and shift-invariant Yeo-Johnson transformation ranked best for
@@
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 
-none                                  & &         8.52  &         7.62  &         7.35  &         6.17  &         7.46  &         6.62 \\
-conventional                          & &         3.82  &         5.78  &         3.86  &         7.08  &         6.51  &         5.95 \\
-conventional (z-score norm.)          & &         4.70  &         5.75  &         6.44  &         5.07  &         5.48  &         4.93 \\
-conventional (robust scaling)         & &         3.83  &         6.61  &         5.36  &         5.72  &         4.41  &         5.59 \\
-Raymaekers-Rousseeuw                  & &         4.64  &         3.22  &         4.01  &         4.11  &         6.28  &         5.67 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         5.24  & \textbf{2.66} &         6.25  &         3.62  &         5.29  &         3.50 \\
-Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.99  &         5.22  &         3.69  &         4.27  &         3.57 \\
-invariant                             & & \textbf{3.11} &         6.74  & \textbf{3.24} &         6.13  & \textbf{2.63} &         6.03 \\
-robust invariant                      & &         6.44  &         3.64  &         3.27  & \textbf{3.40} &         2.65  & \textbf{3.14} \\
+none                                  & &         8.52  &         7.62  &         7.35  &         6.17  &         7.45  &         6.61 \\
+conventional                          & &         3.82  &         5.78  &         3.86  &         7.08  &         6.34  &         6.20 \\
+conventional (z-score norm.)          & &         4.70  &         5.75  &         6.44  &         5.07  &         5.50  &         4.99 \\
+conventional (robust scaling)         & &         3.83  &         6.61  &         5.36  &         5.72  &         4.43  &         5.67 \\
+Raymaekers-Rousseeuw                  & &         4.64  &         3.22  &         4.01  &         4.11  &         6.36  &         5.22 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.24  & \textbf{2.66} &         6.25  &         3.62  &         5.31  &         3.53 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.99  &         5.22  &         3.69  &         4.29  &         3.59 \\
+invariant                             & & \textbf{3.11} &         6.74  & \textbf{3.24} &         6.13  & \textbf{2.64} &         6.05 \\
+robust invariant                      & &         6.44  &         3.64  &         3.27  & \textbf{3.40} &         2.66  & \textbf{3.14} \\
 

### 2025-12-15 manuscript_rmarkdown/manuscript_appendix.Rmd
Update scale for assessing transformation using simulated data.

@@
 
-# Appendix C: Central normality test and empirical central normality test critical values
+# Appendix C: Empirical central normality test critical values
 
@@
 of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 1000.
 If necessary, each sequence was shifted so that every element had a strictly positive value. 
@@
 
-none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.47  &         6.78 \\
-conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.10 \\
-conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.34  &         5.84 \\
-conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.35  &         5.84 \\
-Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.30  &         5.80 \\ 
-Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.48  &         3.50 \\ 
-Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.53  &         3.51 \\
-invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.51} &         5.58 \\
-robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} & \textbf{3.51} & \textbf{2.00} \\
+none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.46  &         6.77 \\
+conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.26 \\
+conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.39  &         5.96 \\
+conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.40  &         5.96 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.29  &         5.35 \\ 
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.53  &         3.51 \\ 
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.58  &         3.52 \\
+invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.53} &         5.67 \\
+robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} &         3.55  & \textbf{2.00} \\
 
@@
 none                                  & & 0.202 & 0.112 & 0.128 & 0.080 & 0.129 & 0.080 \\
-conventional                          & & 0.058 & 0.094 & 0.085 & 0.078 & 0.129 & 0.080 \\
+conventional                          & & 0.058 & 0.094 & 0.085 & 0.078 & 0.126 & 0.080 \\
 conventional (z-score norm.)          & & 0.061 & 0.094 & 0.091 & 0.073 & 0.091 & 0.073 \\
 conventional (robust scaling)         & & 0.058 & 0.094 & 0.090 & 0.073 & 0.090 & 0.073 \\
-Raymaekers-Rousseeuw                  & & 0.071 & 0.055 & 0.092 & 0.279 & 0.129 & 0.080 \\
+Raymaekers-Rousseeuw                  & & 0.071 & 0.055 & 0.092 & 0.279 & 0.126 & 0.079 \\
 Raymaekers-Rousseeuw (z-score norm.)  & & 0.087 & 0.057 & 0.097 & 0.060 & 0.097 & 0.060 \\
@@
 shifted dataset also contains sequences sampled from asymmetric generalised
-normal distributions, but centred at 100, and scaled by 0.001. Several
+normal distributions, but centred at 1000. Several
 transformation methods include normalisation before transformation, indicated by
@@
 none                                  & & 0.150 & 0.068 & 0.050 & 0.033 & 0.049 & 0.033 \\
-conventional                          & & 0.005 & 0.049 & 0.006 & 0.032 & 0.049 & 0.033 \\
+conventional                          & & 0.005 & 0.049 & 0.006 & 0.032 & 0.046 & 0.032 \\
 conventional (z-score norm.)          & & 0.008 & 0.049 & 0.013 & 0.026 & 0.011 & 0.025 \\
 conventional (robust scaling)         & & 0.005 & 0.050 & 0.012 & 0.026 & 0.010 & 0.026 \\
-Raymaekers-Rousseeuw                  & & 0.018 & 0.011 & 0.014 & 0.233 & 0.049 & 0.033 \\
+Raymaekers-Rousseeuw                  & & 0.018 & 0.011 & 0.014 & 0.233 & 0.046 & 0.031 \\
 Raymaekers-Rousseeuw (z-score norm.)  & & 0.034 & 0.012 & 0.018 & 0.014 & 0.017 & 0.013 \\
@@
 
-none                                  & &         7.75  &         6.81  &         7.27  &         6.40  &         7.45  &         6.60 \\
-conventional                          & &         4.19  &         5.47  &         4.48  &         5.78  &         6.55  &         5.94 \\
-conventional (z-score norm.)          & &         4.36  &         5.12  &         5.01  &         4.99  &         4.35  &         4.70 \\
-conventional (robust scaling)         & & \textbf{4.05} &         5.63  &         4.38  &         5.49  &         3.76  &         5.20 \\
-Raymaekers-Rousseeuw                  & &         5.36  &         4.19  &         5.14  &         4.62  &         6.32  &         5.83 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         5.56  &         4.08  &         5.67  &         4.05  &         5.03  &         3.81 \\
-Raymaekers-Rousseeuw (robust scaling) & &         5.33  &         4.14  &         5.15  &         4.15  &         4.55  &         3.94 \\
-invariant                             & &         4.11  &         5.92  &         4.12  &         5.67  &         3.62  &         5.37 \\
-robust invariant                      & &         4.28  & \textbf{3.65} & \textbf{3.79} & \textbf{3.85} & \textbf{3.37} & \textbf{3.61} \\
+none                                  & &         7.75  &         6.81  &         7.27  &         6.40  &         7.44  &         6.59 \\
+conventional                          & &         4.19  &         5.47  &         4.48  &         5.78  &         6.35  &         6.00 \\
+conventional (z-score norm.)          & &         4.36  &         5.12  &         5.01  &         4.99  &         4.36  &         4.71 \\
+conventional (robust scaling)         & & \textbf{4.05} &         5.63  &         4.38  &         5.49  &         3.77  &         5.21 \\
+Raymaekers-Rousseeuw                  & &         5.36  &         4.19  &         5.14  &         4.62  &         6.47  &         5.72 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.56  &         4.08  &         5.67  &         4.05  &         5.04  &         3.82 \\
+Raymaekers-Rousseeuw (robust scaling) & &         5.33  &         4.14  &         5.15  &         4.15  &         4.56  &         3.95 \\
+invariant                             & &         4.11  &         5.92  &         4.12  &         5.67  &         3.64  &         5.38 \\
+robust invariant                      & &         4.28  & \textbf{3.65} & \textbf{3.79} & \textbf{3.85} & \textbf{3.37} & \textbf{3.62} \\
 
@@
 none                                  & & 0.246 & 0.173 & 0.186 & 0.156 & 0.187 & 0.157 \\
-conventional                          & & 0.139 & 0.142 & 0.142 & 0.132 & 0.187 & 0.157 \\
+conventional                          & & 0.139 & 0.142 & 0.142 & 0.132 & 0.184 & 0.155 \\
 conventional (z-score norm.)          & & 0.139 & 0.142 & 0.142 & 0.130 & 0.142 & 0.130 \\
 conventional (robust scaling)         & & 0.139 & 0.142 & 0.141 & 0.130 & 0.141 & 0.130 \\
-Raymaekers-Rousseeuw                  & & 0.211 & 0.151 & 0.183 & 1.515 & 0.187 & 0.157 \\
+Raymaekers-Rousseeuw                  & & 0.211 & 0.151 & 0.183 & 1.515 & 0.185 & 0.155 \\
 Raymaekers-Rousseeuw (z-score norm.)  & & 0.295 & 0.161 & 0.175 & 0.156 & 0.175 & 0.156 \\
@@
 
-none                                  & &         8.85  &         8.14  &         7.35  &         6.29  &         7.41  &         6.87 \\
-conventional                          & &         3.50  &         5.65  &         3.52  &         7.78  &         6.42  &         6.05 \\
-conventional (z-score norm.)          & &         5.24  &         5.94  &         7.34  &         4.98  &         6.26  &         4.79 \\
-conventional (robust scaling)         & &         3.53  &         6.98  &         5.97  &         5.79  &         4.87  &         5.65 \\
-Raymaekers-Rousseeuw                  & &         4.15  &         2.81  &         3.37  &         3.72  &         6.19  &         5.79 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         5.34  & \textbf{1.86} &         6.57  &         3.39  &         5.49  &         3.30 \\
-Raymaekers-Rousseeuw (robust scaling) & &         4.20  &         2.33  &         5.21  &         3.36  &         4.12  &         3.23 \\
-invariant                             & & \textbf{2.36} &         7.59  & \textbf{2.66} &         6.82  & \textbf{1.98} &         6.81 \\
-robust invariant                      & &         7.83  &         3.70  &         3.01  & \textbf{2.88} &         2.26  & \textbf{2.51} \\
+none                                  & &         8.85  &         8.14  &         7.35  &         6.29  &         7.40  &         6.86 \\
+conventional                          & &         3.50  &         5.65  &         3.52  &         7.78  &         6.24  &         6.45 \\
+conventional (z-score norm.)          & &         5.24  &         5.94  &         7.34  &         4.98  &         6.29  &         4.92 \\
+conventional (robust scaling)         & &         3.53  &         6.98  &         5.97  &         5.79  &         4.90  &         5.79 \\
+Raymaekers-Rousseeuw                  & &         4.15  &         2.81  &         3.37  &         3.72  &         6.22  &         5.04 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         5.34  & \textbf{1.86} &         6.57  &         3.39  &         5.52  &         3.35 \\
+Raymaekers-Rousseeuw (robust scaling) & &         4.20  &         2.33  &         5.21  &         3.36  &         4.16  &         3.27 \\
+invariant                             & & \textbf{2.36} &         7.59  & \textbf{2.66} &         6.82  & \textbf{1.99} &         6.82 \\
+robust invariant                      & &         7.83  &         3.70  &         3.01  & \textbf{2.88} &         2.27  & \textbf{2.52} \\
 
@@
 none                                  & & 0.186 & 0.091 & 0.106 & 0.053 & 0.106 & 0.053 \\
-conventional                          & & 0.014 & 0.076 & 0.060 & 0.057 & 0.106 & 0.053 \\
+conventional                          & & 0.014 & 0.076 & 0.060 & 0.057 & 0.103 & 0.053 \\
 conventional (z-score norm.)          & & 0.021 & 0.076 & 0.070 & 0.050 & 0.070 & 0.050 \\
 conventional (robust scaling)         & & 0.014 & 0.076 & 0.069 & 0.050 & 0.069 & 0.050 \\
-Raymaekers-Rousseeuw                  & & 0.015 & 0.021 & 0.060 & 0.030 & 0.106 & 0.053 \\
+Raymaekers-Rousseeuw                  & & 0.015 & 0.021 & 0.060 & 0.030 & 0.103 & 0.051 \\
 Raymaekers-Rousseeuw (z-score norm.)  & & 0.021 & 0.020 & 0.070 & 0.030 & 0.070 & 0.030 \\

### 2025-12-15 manuscript_rmarkdown/simulations.R

@@ -1194,8 +1194,8 @@
       # Dirty data that is shifted and scaled to be far from 0.0.
       x <- power.transform::ragn(
         n,
-        location = 100.0,
-        scale = 0.001 * 1 / sqrt(2),
+        location = 1000.0,
+        scale = 1 / sqrt(2),
         alpha = alpha,
         beta = beta
       )

### 2025-12-15 manuscript_rmarkdown/manuscript.Rmd
Grammar and spelling.

@@
   The sequences are drawn from a range of skewed, unimodal, but not necessarily 
-  normal, distributions, to reflect less \textit{clean} data then that presented in the first dataset.
+  normal, distributions, to reflect less \textit{clean} data than presented in the first dataset.
 

### 2025-12-19 manuscript_rmarkdown/manuscript.Rmd
Rework text.

@@
 \begin{definition}[Box-Cox power transformation]
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a finite sequence of length $n > 0$.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i > 0 \right\}$ be a vector of length $n > 0$.
 Let $\lambda \in \mathcal{R}$ be a transformation parameter.
@@
 \begin{definition}[Yeo-Johnson power transformation]
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a finite sequence of length $n > 0$.
+Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ be a vector of length $n > 0$.
 Let $\lambda \in \mathcal{R}$ be a transformation parameter.
@@
 \begin{definition}[Numeric feature]
-A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
+A numeric feature is a vector $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
 \end{definition}
@@
 Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i - x_0 > 0 \right\}$ be a finite sequence of length $n > 0$.
+Let $\mathbf{X}$ be a numeric feature of length $n > 0$.
 
@@
 Let $x_0 \in \mathcal{R}$ and $s > 0$ be location and scale parameters.
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathcal{R} \right\}$ be a finite sequence of length $n > 0$.
+Let $\mathbf{X}$ be a numeric feature of length $n > 0$.
 
@@
 Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
-Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 
@@
 Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
-Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
+Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.
 
@@
 Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
-Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
+Let $\mu_w$ be the weighted mean of the Box-Cox transformed feature:
 \begin{equation*}
@@
 
-Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
+Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed feature:
 \begin{equation*}
@@
 Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
-Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
+Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed feature:
 \begin{equation*}
@@
 
-Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
+Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed feature:
 \begin{equation*}
@@
 
-\begin{definition}[Central portion of a sequence]
-Let $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R}\right\}$ be a finite sequence of length $n > 0$,
+\begin{definition}[Central portion of a numeric feature]
+Let $\mathbf{X}$ be a numeric feature of length $n > 0$, as before,
 ordered so that $x_1 \leq x_2 \leq \ldots \leq x_n$.
@@
 
-\begin{definition}[Residual errors of the central portion of a sequence]
+\begin{definition}[Residual errors of the central portion of a feature]
 
@@
 
-A central normal distribution is any distribution whose quantile function is 
-defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
+A central normal distribution is any distribution whose quantile function 
+satisfies $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
 with $\mathcal{N}$ any normal distribution.
@@
 
-For $\kappa = 1.0$, the central normal distribution is a normal distribution.
+A normal distribution is a central normal distribution.
 
@@
 
-By the Glivenko-Cantelli theorem, if a sequence $\mathbf{X}$ was sampled from a
-central normal distribution, then almost surely
-$\lim_{n \rightarrow \infty} \sum_{r_j \in \mathbf{R}_{\text{c}}} r_j = 0$,
+For $\kappa = 1.0$, a central normal distribution is a normal distribution.
 
@@
 
+\begin{corollary}
+
+If $\mathbf{X}$ is randomly sampled from a normal distribution,
+$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, almost surely.
+That is, every residual of the central portion of $\mathbf{X}$ will asymptotically
+converge to 0.
+
+\end{corollary}
+
+The latter corollary follows from the definition of the residuals and the restriction
+to normal distributions. Residuals were defined using 
+$\frac{x_i - \mu_M}{\sigma_M} \approx \hat{F}^{-1}_{\mathcal{N}}(p_i)$, with
+$\hat{F}^{-1}_{\mathcal{N}}(p_i)$ the empirical quantile function resembling
+the quantile function of the normal distribution $\mathcal{N}(0, 1)$.
+Since $\hat{F}^{-1}_{\mathcal{N}}(p) \rightarrow F^{-1}_{\mathcal{N}}(p)$ with
+$\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$ [@van-der-Vaart1998-xu],
+$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, if and only if
+for all $x_i \in \mathbf{X}_c$, $\frac{x_i - \mu_M}{\sigma_M} \rightarrow \hat{F}^{-1}_{\mathcal{N}}(p_i)$.
+Thus, the requirement is that the estimators for location and scale $\mu_M$ and $\sigma_M$
+are bias-free. M-estimators for location and scale satisfy this condition, if $\mathbf{X}$ 
+follows a normal distribution without contamination [Huber1964-xk; @Huber1981-su].
+
 \begin{definition}[Central normality test]
-The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+The null-hypothesis $\mathcal{H}_0$ is that $\mathbf{X}$ was randomly sampled
 from a population that is distributed according to a central normal distribution.
 
-The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+The alternative hypothesis is that $\mathbf{X}$ was randomly sampled from a
 population that is not distributed according to a central normal distribution.
@@
 To assess whether the proposed power transformations lead to values of $\lambda$ 
-that are invariant to location and scale of the distribution, we simulated three different sequences.
+that are invariant to location and scale of the distribution, we simulated three different numeric features.
 We first randomly drew $10000$ values from a normal distribution: 
@@
 The critical test statistic for the empirical central normality test depends on 
-the significance level $\alpha$, sequence length $n$ and central portion $\kappa$.
+the significance level $\alpha$, feature length $n$ and central portion $\kappa$.
 Critical test statistic values were set through simulation, as described below.
@@
 For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
-we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
-For each sequence we then computed $\tau_{\kappa}$ for 
+we randomly sampled $m_d = 30000$ ordered numeric features $\mathbf{X}$ from $\mathcal{N}(0,1)$.
+For each feature we then computed $\tau_{\kappa}$ for 
 $\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.
@@
 To assess type I error rates for the empirical central normality test and compare
-these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
+these to the Shapiro-Wilk test for normality, we randomly drew another 10000 numeric features from a
 central normal distribution for each 
@@
 
-Each sequence was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
-Then, the elements outside the central portion of the sequence (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
+Each feature was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
+Then, the elements outside the central portion of the feature (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
 Elements with $p_i < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
@@
 
-For each sequence, the p-value for the respective test was used to reject the null
-hypothesis that sequences were derived from a population with a (central) normal distribution.
+For each feature, the p-value for the respective test was used to reject the null
+hypothesis that features were derived from a population with a (central) normal distribution.
 The null hypothesis was rejected if $p \leq 0.05$.
-The type I error rate was determined by computing the fraction of sequences rejected this way.
+The type I error rate was determined by computing the fraction of features rejected this way.
 Figure \ref{fig:empirical-central-normality-error_rate} shows that when the
@@
 To determine the weighting function parameters for each of the nine combinations,
-$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
+$m_d=500$ numeric features $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
 drawn from randomly parametrised asymmetric generalised normal distributions.
@@
 Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
-To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
-with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
+To form each feature $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
+with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of features with between $50$ and $1000$ elements each.
 Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
@@
 The composite loss consisted of two components: a loss term $L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}$, i.e. the mean of residual 
-errors of the central 80% of elements of each sequence $\mathbf{X}_i$, which aimed at optimising central normality; 
+errors of the central 80% of elements of each feature $\mathbf{X}_i$, which aimed at optimising central normality; 
 and a loss term $L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)$,
 with $\lambda_{0,i}$ and $\lambda_{i}$ transformation parameters found
-for sequence $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
+for feature $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
 and tolerance parameter $\xi = 0.5$ for Box-Cox and $\xi = 0.3$ for Yeo-Johnson power transformations.
 This second term aimed to prevent solutions that provide small improvements in central normality
-at the cost of a poor fit of the tails of sequences.
+at the cost of a poor fit of the tails of features.
 
@@
 
-Three datasets with 100000 sequences each were created to assess transformation
-to normality. For each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
+Three datasets with 100000 numeric features each were created to assess transformation
+to normality. For each feature $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
 elements were randomly drawn, with $\gamma \sim U\left(1, 4\right)$, 
-resulting in sequences with between $10$ and $10000$ elements.
-
-- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
+resulting in features with between $10$ and $10000$ elements.
+
+- A \textit{clean} dataset with each feature drawn from $\mathcal{N}(0, 1)$ and
   transformed using an inverse power transformation with randomly drawn 
@@
   were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
-  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
+  respectively. Prior to inverse transformation, features for Box-Cox transformations
   were shifted into the positive domain, with minimum value $1$.
 
-- A \textit{dirty} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
+- A \textit{dirty} dataset with each feature drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
   with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
   and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-  The sequences are drawn from a range of skewed, unimodal, but not necessarily 
-  normal, distributions, to reflect less \textit{clean} data than presented in the first dataset.
-
-- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)$,
+  The features are drawn from a range of asymmetric unimodal distributions 
+  to reflect a wider range of generating distributions than for the 
+  data presented in the first dataset.
+
+- A \textit{shifted} dataset with each feature drawn from
+  $\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)$,
   with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
   and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
-  The generating process creates sequences similar to that in the \textit{dirty}
+  The generating process creates features similar to those in the \textit{dirty}
   dataset, but shifted to a location that leads to normalisation issues observed
@@
 A dataset with outliers was created for each of the above datasets by replacing
-10 percent of elements in each sequence, as described earlier.
+10 percent of elements in each feature, as described earlier.
 
@@
 
-1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of sequence $\mathbf{X}$.
+1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the 
+  mean and standard deviation of $\mathbf{X}$.
 
@@
 This results in nine types of power transformation. Transformation parameters
-were optimised for each sequence. Subsequently the sum of residual errors 
+were optimised for each feature. Subsequently the sum of residual errors 
 and sum of residual errors of the central portion ($\kappa = 0.80$) were computed
-for each sequence after transformation. Then, each method was ranked according 
+for each feature after transformation. Then, each method was ranked according 
 to the sum of residual errors (data without outliers) or the sum of residual
 errors of the central portion (data with outliers). The average rank of each
-transformation method was computed over all 100000 sequences in each dataset.
+transformation method was computed over all 100000 features in each dataset.
 Average ranks for Yeo-Johnson transformations are shown in Table
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 1000.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 
-We introduced an empirical test for central normality to assess whether sequences
+We introduced an empirical test for central normality to assess whether features
 deviate from normality in a way that might require closer inspection prior to further processing.
@@
 
-This work has the following limitation: We observed several numerical stability
+This work has the following limitations: We observed several numerical stability
 issues for optimisation criteria other than MLE (Appendix B). These appear in regions
@@
 For MLE stability issues were not observed.
+Moreover, the definition of central normal distribution does not constrain the
+distribution of the non-central tails. Thus, the test may be conservative for 
+central normal distributions other than uncontaminated normal distributions, because the 
+location and scale estimators are not generally bias-free [@Rousseeuw1994-ur].
 

### 2025-12-19 manuscript_rmarkdown/simulations.R

@@ -1783,3 +1783,158 @@
     value.var = "tau"
   ))
 }
+
+
+
+.assess_gc_theorem <- function(manuscript_dir) {
+  n_rep <- 100L  # number of distributions drawn.
+  # n_rep <- 10L  # number of distributions drawn.
+  kappa <- c(0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00)  # central portion of distribution.
+
+  # From 5 to 10000000 in equal steps (log 10 scale)
+  n <- unique(floor(10^(seq(from = 0.75, to = 7.0, by = 0.25))))
+
+  # Wrapper for multi-processing.
+  .compute_wrapper <- function(
+    n,
+    n_rep,
+    kappa
+  ) {
+    return(
+      lapply(
+        X = n,
+        FUN = ..assess_gc_theorem,
+        n_rep = n_rep,
+        kappa = kappa
+      )
+    )
+  }
+
+  file_name <- file.path(manuscript_dir, "gc_theorem_data.RDS")
+
+  if (!file.exists(file_name)) {
+    n_threads <- 18L
+    assignment_id <- rep_len(seq_len(n_threads), length(n))
+
+    cl <- parallel::makeCluster(n_threads)
+    on.exit(parallel::stopCluster(cl))
+
+    parallel::clusterExport(cl = cl, "..assess_gc_theorem")
+
+    # data <- lapply(
+    #   X = split(n, assignment_id),
+    #   FUN = .compute_wrapper,
+    #   n_rep = n_rep,
+    #   kappa = kappa
+    # )
+
+    data <- parallel::parLapply(
+      cl = cl,
+      X = split(n, assignment_id),
+      fun = .compute_wrapper,
+      n_rep = n_rep,
+      kappa = kappa
+    )
+
+    # Aggregate to single table.
+    data <- data.table::rbindlist(unlist(data, recursive = FALSE))
+    data <- data[order(n, kappa, sum_residual_error)]
+
+    # Store
+    saveRDS(data, file_name)
+
+  } else {
+    data <- readRDS(file_name)
+  }
+
+  return(data)
+}
+
+
+..assess_gc_theorem <- function(n, n_rep, kappa) {
+  set.seed(19L + n)
+
+  data <- list()
+  result_index <- 1L
+
+  # Repeat n_rep times.
+  for (jj in seq_len(n_rep)) {
+    # Standard normal distribution -- we will only assess the central portion
+    # kappa
+    x <- power.transform::ragn(
+      floor(n),
[…]

### 2025-12-19 manuscript_latex/manuscript.tex
Update manuscript for revision 3.

@@
 \begin{definition}[Numeric feature]
-A numeric feature is a finite sequence $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in  \right\}$ of length $n$.
+A numeric feature is a vector $\mathbf{X} = \left\{x_1, x_2, \ldots, x_n | x_i \in \mathbb{R} \right\}$ of length $n$.
 \end{definition}
@@
 
-A central normal distribution is any distribution whose quantile function is
-defined by $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
+A central normal distribution is any distribution whose quantile function
+satisfies $F^{-1}_{\mathcal{N}}$ for $\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}$,
 with $\mathcal{N}$ any normal distribution.
@@
 
+\begin{corollary}
+
+If $\mathbf{X}$ is randomly sampled from a normal distribution,
+$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, almost surely.
+That is, every residual of the central portion of $\mathbf{X}$ will asymptotically
+converge to 0.
+
+\end{corollary}
+
+The latter corollary follows from the definition of the residuals and
+the restriction to normal distributions. Residuals were defined using
+\(\frac{x_i - \mu_M}{\sigma_M} \approx \hat{F}^{-1}_{\mathcal{N}}(p_i)\),
+with \(\hat{F}^{-1}_{\mathcal{N}}(p_i)\) the empirical quantile function
+resembling the quantile function of the normal distribution
+\(\mathcal{N}(0, 1)\). Since
+\(\hat{F}^{-1}_{\mathcal{N}}(p) \rightarrow F^{-1}_{\mathcal{N}}(p)\)
+with \(\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}\)
+\citep{van-der-Vaart1998-xu},
+\(\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0\), if and only
+if for all \(x_i \in \mathbf{X}_c\),
+\(\frac{x_i - \mu_M}{\sigma_M} \rightarrow \hat{F}^{-1}_{\mathcal{N}}(p_i)\).
+Thus, the requirement is that the estimators for location and scale
+\(\mu_M\) and \(\sigma_M\) are bias-free. M-estimators for location and
+scale satisfy this condition, if \(\mathbf{X}\) follows a normal
+distribution without contamination \citep{Huber1964-xk, Huber1981-su}.
+
 \begin{definition}[Central normality test]
-The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+The null-hypothesis $\mathcal{H}_0$ is that $\mathbf{X}$ was randomly sampled
 from a population that is distributed according to a central normal distribution.
 
-The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+The alternative hypothesis is that $\mathbf{X}$ was randomly sampled from a
 population that is not distributed according to a central normal distribution.
@@
 \(\lambda\) that are invariant to location and scale of the
-distribution, we simulated three different sequences. We first randomly
-drew \(10000\) values from a normal distribution:
+distribution, we simulated three different numeric features. We first
+randomly drew \(10000\) values from a normal distribution:
 \(\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)\),
@@
 The critical test statistic for the empirical central normality test
-depends on the significance level \(\alpha\), sequence length \(n\) and
+depends on the significance level \(\alpha\), feature length \(n\) and
 central portion \(\kappa\). Critical test statistic values were set
@@
 \(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}\)
-we randomly sampled \(m_d = 30000\) ordered sequences \(\mathbf{X}\)
-from \(\mathcal{N}(0,1)\). For each sequence we then computed
-\(\tau_{\kappa}\) for
+we randomly sampled \(m_d = 30000\) ordered numeric features
+\(\mathbf{X}\) from \(\mathcal{N}(0,1)\). For each feature we then
+computed \(\tau_{\kappa}\) for
 \(\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}\).
@@
 and compare these to the Shapiro-Wilk test for normality, we randomly
-drew another 10000 sequences from a central normal distribution for each
+drew another 10000 numeric features from a central normal distribution
+for each
 \(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
 
-Each sequence was created as follows: first \(n\) values are sampled
-from \(\mathcal{N}(0, 1)\). Then, the elements outside the central
-portion of the sequence
-(i.e.~\(\mathbf{X} \setminus \mathbf{X}_{\text{c}}\)) are replaced.
-Elements with \(p_i < \frac{1-\kappa}{2}\) are replaced with values
-sampled from \(\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))\), and
-elements with \(p_i > \frac{1 + \kappa}{2}\) are replaced with values
-sampled from \(\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)\).
-
-For each sequence, the p-value for the respective test was used to
-reject the null hypothesis that sequences were derived from a population
-with a (central) normal distribution. The null hypothesis was rejected
-if \(p \leq 0.05\). The type I error rate was determined by computing
-the fraction of sequences rejected this way. Figure
+Each feature was created as follows: first \(n\) values are sampled from
+\(\mathcal{N}(0, 1)\). Then, the elements outside the central portion of
+the feature (i.e.~\(\mathbf{X} \setminus \mathbf{X}_{\text{c}}\)) are
+replaced. Elements with \(p_i < \frac{1-\kappa}{2}\) are replaced with
+values sampled from \(\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))\),
+and elements with \(p_i > \frac{1 + \kappa}{2}\) are replaced with
+values sampled from \(\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)\).
+
+For each feature, the p-value for the respective test was used to reject
+the null hypothesis that features were derived from a population with a
+(central) normal distribution. The null hypothesis was rejected if
+\(p \leq 0.05\). The type I error rate was determined by computing the
+fraction of features rejected this way. Figure
 \ref{fig:empirical-central-normality-error_rate} shows that when the
@@
 To determine the weighting function parameters for each of the nine
-combinations, \(m_d=500\) sequences \(\mathbf{X}_i\)
+combinations, \(m_d=500\) numeric features \(\mathbf{X}_i\)
 (\(i \in \{1, 2, \ldots, m_d\}\)) were randomly drawn from randomly
@@
 were set as \(\mu = 0\) and \(\sigma = 1\), respectively. To form each
-sequence \(\mathbf{X}_i\), \(n = \lceil 10^\gamma \rceil\) instances
-were then randomly drawn, with
-\(\gamma \sim U\left(\log_{10}50, 3\right)\), resulting in a set of
-sequences with between \(50\) and \(1000\) elements each. Outlier values
-were then drawn to randomly replace 10 percent of the elements of
-\(\mathbf{X}_i\). These values were set according to
+feature \(\mathbf{X}_i\), \(n = \lceil 10^\gamma \rceil\) instances were
+then randomly drawn, with \(\gamma \sim U\left(\log_{10}50, 3\right)\),
+resulting in a set of features with between \(50\) and \(1000\) elements
+each. Outlier values were then drawn to randomly replace 10 percent of
+the elements of \(\mathbf{X}_i\). These values were set according to
 \citet{Tukey1977-xm}, as follows. Let
@@
 \(L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}\), i.e.~the mean of
-residual errors of the central 80\% of elements of each sequence
+residual errors of the central 80\% of elements of each feature
 \(\mathbf{X}_i\), which aimed at optimising central normality; and a
@@
 with \(\lambda_{0,i}\) and \(\lambda_{i}\) transformation parameters
-found for sequence \(\mathbf{X}_i\) prior to and after adding outliers,
+found for feature \(\mathbf{X}_i\) prior to and after adding outliers,
 respectively, and tolerance parameter \(\xi = 0.5\) for Box-Cox and
@@
 aimed to prevent solutions that provide small improvements in central
-normality at the cost of a poor fit of the tails of sequences.
+normality at the cost of a poor fit of the tails of features.
 
@@
 
-Three datasets with 100000 sequences each were created to assess
-transformation to normality. For each sequence \(\mathbf{X}_i\),
+Three datasets with 100000 numeric features each were created to assess
+transformation to normality. For each feature \(\mathbf{X}_i\),
 \(n = \lceil 10^\gamma \rceil\) elements were randomly drawn, with
-\(\gamma \sim U\left(1, 4\right)\), resulting in sequences with between
+\(\gamma \sim U\left(1, 4\right)\), resulting in features with between
 \(10\) and \(10000\) elements.
@@
 \item
-  A \textit{clean} dataset with each sequence drawn from
+  A \textit{clean} dataset with each feature drawn from
   \(\mathcal{N}(0, 1)\) and transformed using an inverse power
@@
   transformations for assessing Box-Cox and Yeo-Johnson transformations,
-  respectively. Prior to inverse transformation, sequences for Box-Cox
+  respectively. Prior to inverse transformation, features for Box-Cox
   transformations were shifted into the positive domain, with minimum
@@
 \item
-  A \textit{dirty} dataset with each sequence drawn from
+  A \textit{dirty} dataset with each feature drawn from
   \(\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)\),
@@
   \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-  \(\beta \sim U\left(1.00, 5.00 \right)\).
+  \(\beta \sim U\left(1.00, 5.00 \right)\). The features are drawn from
+  a range of asymmetric unimodal distributions to reflect a wider range
+  of generating distributions than for the data presented in the first
+  dataset.
 \item
-  A \textit{shifted} dataset with each sequence drawn from
-  \(\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)\),
+  A \textit{shifted} dataset with each feature drawn from
+  \(\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)\),
   with randomly drawn skewness parameter
   \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-  \(\beta \sim U\left(1.00, 5.00 \right)\).
+  \(\beta \sim U\left(1.00, 5.00 \right)\). The generating process
+  creates features similar to those in the \textit{dirty} dataset, but
+  shifted to a location that leads to normalisation issues observed for
+  conventional power transformation methods, as shown in Figure
+  \ref{fig:decreased-normality}.
 \end{itemize}
@@
 A dataset with outliers was created for each of the above datasets by
-replacing 10 percent of elements in each sequence, as described earlier.
+replacing 10 percent of elements in each feature, as described earlier.
 
@@
   \(x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma\), with \(\mu\)
-  and \(\sigma\) the mean and standard deviation of sequence
-  \(\mathbf{X}\).
+  and \(\sigma\) the mean and standard deviation of \(\mathbf{X}\).
 \item
@@
 This results in nine types of power transformation. Transformation
-parameters were optimised for each sequence. Subsequently the sum of
+parameters were optimised for each feature. Subsequently the sum of
 residual errors and sum of residual errors of the central portion
-(\(\kappa = 0.80\)) were computed for each sequence after
-transformation. Then, each method was ranked according to the sum of
-residual errors (data without outliers) or the sum of residual errors of
-the central portion (data with outliers). The average rank of each
-transformation method was computed over all 100000 sequences in each
-dataset. Average ranks for Yeo-Johnson transformations are shown in
-Table \ref{tab:comparison_methods_simulations_yeo_johnson}. Location and
+(\(\kappa = 0.80\)) were computed for each feature after transformation.
+Then, each method was ranked according to the sum of residual errors
+(data without outliers) or the sum of residual errors of the central
+portion (data with outliers). The average rank of each transformation
+method was computed over all 100000 features in each dataset. Average
+ranks for Yeo-Johnson transformations are shown in Table
+\ref{tab:comparison_methods_simulations_yeo_johnson}. Location and
 shift-invariant Yeo-Johnson transformation ranked best for all datasets
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 We introduced an empirical test for central normality to assess whether
-sequences deviate from normality in a way that might require closer
+features deviate from normality in a way that might require closer
 inspection prior to further processing. The empirical central normality
@@
 
-This work has the following limitation: We observed several numerical
+This work has the following limitations: We observed several numerical
 stability issues for optimisation criteria other than MLE (\hyperref[appendix-b]{Appendix B}).
@@
 very large or small numbers when using conventional power
-transformations. For MLE stability issues were not observed.
+transformations. For MLE stability issues were not observed. Moreover,
+the definition of central normal distribution does not constrain the
+distribution of the non-central tails. Thus, the test may be
+conservative for central normal distributions other than uncontaminated
+normal distributions, because the location and scale estimators are not
+generally bias-free \citep{Rousseeuw1994-ur}.
 
@@
 Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
-If necessary, each sequence was shifted so that every element had a strictly positive value.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. The clean dataset consists of features derived through inverse Box-Cox transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
+If necessary, each feature was shifted so that every element had a strictly positive value.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 
-none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.47  &         6.78 \\
-conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.10 \\
-conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.34  &         5.84 \\
-conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.35  &         5.84 \\
-Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.30  &         5.80 \\
-Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.48  &         3.50 \\
-Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.53  &         3.51 \\
-invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.51} &         5.58 \\
-robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} & \textbf{3.51} & \textbf{2.00} \\
+none                                  & &         7.63  &         6.95  &         7.29  &         6.59  &         7.46  &         6.77 \\
+conventional                          & &         3.11  &         5.98  &         5.41  &         6.18  &         6.51  &         6.26 \\
+conventional (z-score norm.)          & &         4.15  &         6.27  &         4.83  &         6.32  &         4.39  &         5.96 \\
+conventional (robust scaling)         & &         4.20  &         6.29  &         4.84  &         6.31  &         4.40  &         5.96 \\
+Raymaekers-Rousseeuw                  & &         4.50  &         3.56  &         5.39  &         3.69  &         6.29  &         5.35 \\
+Raymaekers-Rousseeuw (z-score norm.)  & &         6.23  &         3.80  &         4.93  &         3.85  &         4.53  &         3.51 \\
+Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.80  &         4.99  &         3.87  &         4.58  &         3.52 \\
+invariant                             & & \textbf{2.96} &         5.59  & \textbf{3.62} &         6.01  & \textbf{3.53} &         5.67 \\
+robust invariant                      & &         6.10  & \textbf{2.76} &         3.68  & \textbf{2.19} &         3.55  & \textbf{2.00} \\
 
@@
 In the main manuscript, transformation methods are compared by ranking
-within each experiment, defined by the combination of sequence and
+within each experiment, defined by the combination of feature and
 absence/presence of outliers. This allows for direct comparison of
@@
 methods. This metric is computed from the residual error and the
-residual error of the central portion (\(\kappa = 0.80\)) for sequences
+residual error of the central portion (\(\kappa = 0.80\)) for features
 without and with outliers, respectively. In each experiment, the
-respective residual error was normalised by dividing with the sequence
+respective residual error was normalised by dividing with the feature
 length, and can be interpreted as the average residual error of a single
-sample. This reduces the effect of sequence length on the residual
-error.
+sample. This reduces the effect of feature length on the residual error.
 
@@
 best method for each experiment and the individual methods. As above,
-normalisation was done by division by the length of the sequence
-corresponding to each experiment.
+normalisation was done by division by the feature length of each experiment.
 
@@
 Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
-The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. Residual errors were normalised by feature length to reduce the effect of feature length on the residual error.
+The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 transformation methods and the best transformation method within each
-experiment, normalised by the corresponding sequence length. This is based on
+experiment, normalised by the corresponding feature length. This is based on
 either the residual error (without outliers) or the residual error of the
 central portion (with outliers; $\kappa = 0.80$) over 3 datasets with 100000
-sequences each. Residual errors were normalised by sequence length to reduce the
-effect of sequence length on the residual error. The clean dataset consists of
-sequences derived through inverse Yeo-Johnson transformation of data sampled
-from a standard normal distribution. The dirty dataset contains sequences
+features each. Residual errors were normalised by feature length to reduce the
+effect of feature length on the residual error. The clean dataset consists of
+features derived through inverse Yeo-Johnson transformation of data sampled
+from a standard normal distribution. The dirty dataset contains features
 sampled from asymmetric generalised normal distributions, centred at 0. The
-shifted dataset also contains sequences sampled from asymmetric generalised
-normal distributions, but centred at 100, and scaled by 0.001. Several
+shifted dataset also contains features sampled from asymmetric generalised
+normal distributions, but centred at 1000. Several
 transformation methods include normalisation before transformation, indicated by
@@
 
-\subsection{Ranking Yeo-Johnson transformations: effect of sequence
-length}\label{ranking-yeo-johnson-transformations-effect-of-sequence-length}
+\subsection{Ranking Yeo-Johnson transformations: effect of feature
+length}\label{ranking-yeo-johnson-transformations-effect-of-feature-length}
 
@@
 Yeo-Johnson-based power transformation methods across three sets of
-problems. To assess the effect of limited sequence lengths, we perform
+problems. To assess the effect of limited feature lengths, we perform
 ranking on subsets of the experiment. The results for ranking the subset
-of sequences with 30 samples or fewer (\(n = 16001\)) is shown in Table
+of features with 30 samples or fewer (\(n = 16001\)) is shown in Table
 \ref{tab:comparison_methods_simulations_yeo_johnson_10_30}. In the clean
-dataset without outliers, where all sequences were sampled from a normal
+dataset without outliers, where all features were sampled from a normal
 distribution and then underwent an inverse Yeo-Johnson transformation,
@@
 
-In the subset for larger sequence lengths, with at least 1000 samples
-per sequence (\(n = 33037\)), the cost of making the location- and
+In the subset for larger feature lengths, with at least 1000 samples per
+feature (\(n = 33037\)), the cost of making the location- and
 scale-invariant transformation robust against outliers becomes more
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Each sequence contained 30 samples or fewer.
-The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 features each. Each feature contained 30 samples or fewer.
+The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
-Each sequence contained 30 samples or fewer. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 features each. Residual errors were normalised by feature length to reduce the effect of feature length on the residual error.
+Each feature contained 30 samples or fewer. The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 33037 sequences each. Each sequence contained between 1000 and 10000 samples.
-The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 33037 features each. Each feature contained between 1000 and 10000 samples.
+The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 Comparison of average normalised residual error between Yeo-Johnson transformation methods based on either the residual error (without outliers) or the residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Residual errors were normalised by sequence length to reduce the effect of sequence length on the residual error.
-Each sequence contained between 1000 and 10000 samples. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 features each. Residual errors were normalised by feature length to reduce the effect of feature length on the residual error.
+Each feature contained between 1000 and 10000 samples. The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 Robust transformations are hypothesised to have a cost in efficiency for
-data without outliers, i.e.~clean data. Here we draw nine sequences with
+data without outliers, i.e.~clean data. Here we draw nine features with
 elements randomly drawn from a standard normal distribution
@@
 
-The sequences drawn resulted from the permutations of the number of
-elements of each sequence (\(n \in \{30, 100, 500\}\)) and
-transformation parameter for the inverse transformation
-\(\lambda \in \{0.1, 1.0, 1.9\}\). Each sequence then underwent power
+The features drawn resulted from the permutations of the number of
+elements of each feature (\(n \in \{30, 100, 500\}\)) and transformation
+parameter for the inverse transformation
+\(\lambda \in \{0.1, 1.0, 1.9\}\). Each feature then underwent power
 transformation using Box-Cox and Yeo-Johnson transformations. For
-Box-Cox transformations, each sequence was shifted prior to inverse
+Box-Cox transformations, each feature was shifted prior to inverse
 transformation ensuring all elements were strictly positive after

### 2025-12-19 manuscript_rmarkdown/manuscript.tex
Update manuscript for revision 3.

@@
 \author{Alex Zwanenburg, Steffen Löck}
-\date{2025-10-29}
+\date{2025-12-19}
 
@@
 
+\begin{corollary}
+
+If $\mathbf{X}$ is randomly sampled from a normal distribution,
+$\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0$, almost surely.
+That is, every residual of the central portion of $\mathbf{X}$ will asymptotically
+converge to 0.
+
+\end{corollary}
+
+The latter corollary follows from the definition of the residuals and
+the restriction to normal distributions. Residuals were defined using
+\(\frac{x_i - \mu_M}{\sigma_M} \approx \hat{F}^{-1}_{\mathcal{N}}(p_i)\),
+with \(\hat{F}^{-1}_{\mathcal{N}}(p_i)\) the empirical quantile function
+resembling the quantile function of the normal distribution
+\(\mathcal{N}(0, 1)\). Since
+\(\hat{F}^{-1}_{\mathcal{N}}(p) \rightarrow F^{-1}_{\mathcal{N}}(p)\)
+with \(\frac{1-\kappa}{2} \leq  p \leq \frac{1 + \kappa}{2}\)
+\citep{van-der-Vaart1998-xu},
+\(\sup_{r_j \in \mathbf{R}_{\text{c}}} r_j \rightarrow 0\), if and only
+if for all \(x_i \in \mathbf{X}_c\),
+\(\frac{x_i - \mu_M}{\sigma_M} \rightarrow \hat{F}^{-1}_{\mathcal{N}}(p_i)\).
+Thus, the requirement is that the estimators for location and scale
+\(\mu_M\) and \(\sigma_M\) are bias-free. M-estimators for location and
+scale satisfy this condition, if \(\mathbf{X}\) follows a normal
+distribution without contamination {[}Huber1964-xk;
+\citet{Huber1981-su}{]}.
+
 \begin{definition}[Central normality test]
-The null-hypothesis $\mathcal{H}_0$ is that the sequence $\mathbf{X}$ was sampled
+The null-hypothesis $\mathcal{H}_0$ is that $\mathbf{X}$ was randomly sampled
 from a population that is distributed according to a central normal distribution.
 
-The alternative hypothesis is that the sequence $\mathbf{X}$ was sampled from a
+The alternative hypothesis is that $\mathbf{X}$ was randomly sampled from a
 population that is not distributed according to a central normal distribution.
@@
   \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-  \(\beta \sim U\left(1.00, 5.00 \right)\).
-\item
-  A \textit{shifted} dataset with each sequence drawn from
-  \(\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)\),
+  \(\beta \sim U\left(1.00, 5.00 \right)\). The features are drawn from
+  a range of asymmetric unimodal distributions to reflect a wider range
+  of generating distributions than for the data presented in the first
+  dataset.
+\item
+  A \textit{shifted} dataset with each feature drawn from
+  \(\mathcal{AGN}\left(\mu = 1000, \sigma = 1/\sqrt{2} , \alpha, \beta \right)\),
   with randomly drawn skewness parameter
   \(\alpha \sim U\left(0.01, 0.99\right)\) and shape parameter
-  \(\beta \sim U\left(1.00, 5.00 \right)\).
+  \(\beta \sim U\left(1.00, 5.00 \right)\). The generating process
+  creates features similar to those in the \textit{dirty} dataset, but
+  shifted to a location that leads to normalisation issues observed for
+  conventional power transformation methods, as shown in Figure
+  \ref{fig:decreased-normality}.
 \end{itemize}
@@
 This results in nine types of power transformation. Transformation
-parameters were optimised for each sequence. Subsequently the sum of
+parameters were optimised for each feature. Subsequently the sum of
 residual errors and sum of residual errors of the central portion
-(\(\kappa = 0.80\)) were computed for each sequence after
-transformation. Then, each method was ranked according to the sum of
-residual errors (data without outliers) or the sum of residual errors of
-the central portion (data with outliers). The average rank of each
-transformation method was computed over all 100000 sequences in each
-dataset. Average ranks for Yeo-Johnson transformations are shown in
-Table \ref{tab:comparison_methods_simulation_yeo_johnson}. Location and
+(\(\kappa = 0.80\)) were computed for each feature after transformation.
+Then, each method was ranked according to the sum of residual errors
+(data without outliers) or the sum of residual errors of the central
+portion (data with outliers). The average rank of each transformation
+method was computed over all 100000 features in each dataset. Average
+ranks for Yeo-Johnson transformations are shown in Table
+\ref{tab:comparison_methods_simulations_yeo_johnson}. Location and
 shift-invariant Yeo-Johnson transformation ranked best for all datasets
@@
 
-This work has the following limitation: We observed several numerical
+This work has the following limitations: We observed several numerical
 stability issues for optimisation criteria other than MLE (Appendix B).

### 2025-12-19 manuscript_rmarkdown/manuscript_appendix.tex
Update appendix C
Update manuscript for revision 3.

@@
 
-\section{Appendix C: Central normality test and empirical central
-normality test critical
-values}\label{appendix-c-central-normality-test-and-empirical-central-normality-test-critical-values}
+\section{Appendix C: Empirical central normality test critical
+values}\label{appendix-c}
 
@@
 Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
-If necessary, each sequence was shifted so that every element had a strictly positive value. 
+(with outliers; $\kappa = 0.80$) over 3 datasets with 100000 features each. The clean dataset consists of features derived through inverse Box-Cox transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
+If necessary, each feature was shifted so that every element had a strictly positive value. 
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
@@
 Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
-(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 sequences each. Each sequence contained 30 samples or fewer. 
-The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
-of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
-The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
+(with outliers; $\kappa = 0.80$) over 3 datasets with 16001 features each. Each feature contained 30 samples or fewer. 
+The clean dataset consists of features derived through inverse Yeo-Johnson transformation
+of data sampled from a standard normal distribution. The dirty dataset contains features sampled from asymmetric generalised normal distributions, centred at 0.
+The shifted dataset also contains features sampled from asymmetric generalised normal distributions, but centred at 1000.
 Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.

### 2025-12-19 manuscript_rmarkdown/manuscript_appendix.tex
Update manuscript_appendix.tex

@@
 \section{Appendix C: Empirical central normality test critical
-values}\label{appendix-c}
+values}\label{appendix-c-empirical-central-normality-test-critical-values}
 
@@
 best method for each experiment and the individual methods. As above,
-normalisation was done by division by the length of the sequence
+normalisation was done by division by the length of the feature
 corresponding to each experiment.
